### 2014-10-21 paper.tex
added paper

\documentclass{article}

\usepackage[round]{natbib}
\usepackage{amsmath,amssymb,amsthm,bm,enumerate,mathrsfs,mathtools}
\usepackage{latexsym,color,verbatim,multirow}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tikz}
\usepackage[margin=1.3in]{geometry}
\usetikzlibrary{shapes,arrows}
\tikzstyle{block} = [rectangle, draw, fill=white!20,
    text width=7em, text centered, rounded corners, minimum height=4em]
\tikzstyle{title} = [text width=7em, text centered, font=\bfseries]
\tikzstyle{line} = [draw, -latex']

\usepackage{mycommands}

\begin{document}

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{observation}[theorem]{Observation}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{fact}[theorem]{Fact}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{model}[theorem]{Model}

\theoremstyle{definition}
\newtheorem{example}{Example}

\newcommand{\cM}{\mathcal{M}}
\newcommand{\cH}{\mathcal{H}}
\newcommand{\cD}{\mathcal{D}}
\newcommand{\FDR}{\textnormal{FDR}}
\newcommand{\FCR}{\textnormal{FCR}}
\newcommand{\crt}{\phi}
\newcommand{\M}{\mathcal{M}}
\newcommand{\cY}{\mathcal{Y}}
\newcommand{\cX}{\mathcal{X}}
\newcommand{\cV}{\mathcal{V}}
\newcommand{\bX}{\mathbf{X}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\Gv}{\;\;\big|\;\;}
%\newcommand{\cP}{\mathcal{P}}
\newcommand{\proj}{\cP}
\newcommand{\pow}{\text{Pow}}
\newcommand{\sF}{\mathscr{F}}
\newcommand{\cF}{\mathcal{F}}
\newcommand{\sC}{\mathscr{C}}
\newcommand{\hJ}{\widehat{J}}
\newcommand{\bH}{\mathbf{H}}
\newcommand{\bM}{\mathbf{M}}
\newcommand{\hK}{\widehat{K}}
\newcommand{\leqAS}{\overset{\textrm{a.s.}}{\leq}}

\newcommand*\mystrut{\vrule width0pt height0pt depth1.5ex\relax}
\newcommand{\underlabel}{\underbracket[1pt][.5pt]{\mystrut \quad\;\; \sub \quad\;\; }}
\newcommand{\JTcomment}[1]{{\color{blue}{(JT: \bf \sc #1) }}}
\newcommand{\WFcomment}[1]{{\color{red}{(WF: \bf \sc #1) }}}

\title{Adaptive Sequential Model Selection}
\maketitle

\begin{abstract}

\end{abstract}

\begin{example}[Linear Regression Variable Selection]\label{ex:linReg}
  Assume that
  \begin{equation}
    Y \sim N(X\beta, \sigma^2I).
  \end{equation}
  With many of the $\beta_j=0$. Our goal is to find a small subset of variables containing the true model.
\end{example}

\section{The Problem of Sequential Model Selection}

\subsection{Notation and Problem Setting}

Suppose that after proposing a random sequence of nested models 
\begin{equation}
  M_0(Y) \sub M_1(Y) \sub \cdots \sub M_d(Y) \sub \cF,
\end{equation}
we wish to sequentially test each $M_k$ against $M_{k-1}$. Let $\bM(Y)$ denote the sequence $(M_0,\ldots,M_d)$, and define its {\em completion index} $K$ to be the index of the first model containing $F$, i.e.
\begin{equation}
  K(\bM) = \min \{k:\, F\in M_k\}.
\end{equation}
Let $N_k$ denote the event $\{Y\in M_{k-1}\}$, i.e. the event that the null hypothesis for the test at stage $k$ is true. $Y\in N_k$ if and only if $k>K$.

For example, suppose that we use the LARS algorithm to build up a linear regression model. Let $J_k(Y)$ be the index of the $k$th variable to have a nonzero coefficient in the LARS path, so that we add variables to the model in the order $X_{J_1}, X_{J_2}, \ldots, X_{J_d}$. It is natural to take model $M_k(Y)$ as
\begin{equation}
  M_k:\, Y \sim 
  N\left(\alpha + X_{J_{1}} \beta_{J_{1}} + \cdots X_{J_k}\beta_{J_k}, 
  \;\sigma^2I\right).
\end{equation}
If the sequence $J_1,\ldots,J_d$ were deterministic instead of adaptive, then we might construct an ANOVA table for testing whether there was sufficient evidence to add each variable to the model in sequence. If the true mean of $Y$ is linear in $X_{J_1}, X_{J_3},$ and $X_{J_5}$, then $K = 5$ and $Y\in N_k$ for $k=6,\ldots,d$.

Once the algorithm for choosing the sequence $\bM(Y)$ has been specified, we will carry out tests for each comparison. The goal we have set for ourselves is meant to be analogous to sequential testing of ANOVA tables for a fixed sequence of models, though there certainly are other sequential model selection tasks that could be of interest.

\subsection{Stopping Rules}
For our purposes, a {\em stopping rule} $\hK(Y)$ is simply an estimate of the completion index $K$, which we could use to ultimately choose the model $M_{\hK}$ from the sequence.

For our purposes, a good stopping rule is one that does not go on refining the model unnecessarily once it is already adequate; e.g. we might want
\begin{equation}
  \P(\hK > K \gv \bM) \leqAS \alpha, \quad \text{ or } \quad \E\left(\frac{(\hK - K)\vee 1}{\hK \vee 1} \gv \bM \right) \leqAS \alpha
\end{equation}
Although it might initally seem more natural to require $\P(\hK < K \gv \bM) \leq \alpha$ (i.e., that we choose an adequate model with high probability) this goal is typically infeasible as stated. For example, suppose that all the coefficients in a linear regression model are $\beta_j=\ep>0$. As $\ep\to 0$, it becomes impossible to distinguish the true model from the global null model $\beta=0$.

In particular, we will focus on constructing a sequence of $p$-values $(p_1(Y),\ldots,p_d(Y))$. $p_k(Y)$ is a {\em conservative selective $p$-value} if
\begin{equation}
  \P_F\left(p_k(Y) \leq \alpha \gv \bM\right) \leqAS \alpha \text{ on } N_k.
\end{equation}
That is, conditional on the question asked, $p_k$ ought to be stochastically larger than uniform.

We say the sequence $(p_1,\ldots,p_d)$ are {\em jointly conservative selective $p$-values} if
\begin{equation}
  \P_F(p_k(Y) < \alpha_k, \forall k>K \gv \bM) \leqAS \prod_{k>K} \alpha_k, \quad \forall \alpha_1,\ldots, \alpha_d\in (0,1), F\in \cF
\end{equation}
Notice that $K$ is random but the conditional probability above still makes sense because $K$ is a function of $\bM$.

Joint conservatism of $p$-values is a minor generalization of requiring the null $p$-values to be mutually independent given the model sequence $\bM$.

\begin{proposition}
  Suppose that, given $\bM$, the $(p_k)_{k>K}$ are conditionally uniform and mutually independent. Then $(p_k)$ are jointly conservative.
\end{proposition}

A  benefit of having jointly conservative $p$-values is that we can apply the stopping rules of \citet{gsell2013sequential} to control error rates such as the model FDR and model FWER. The next section proposes a generic recipe for obtaining such $p$-values

\subsection{A Sequential Selection Variable for Pathwise FDR}

Define the {\em $k$th pathwise selection variable} as 
\begin{equation}
  S_k(Y) = (M_0,\ldots,M_k,p_1,\ldots,p_{k-1}).
\end{equation}
If for each $k$ the $p$-value $p_k$ is conservative conditional on $S_k$, then the $p$-value sequence $(p_k)$ is jointly conservative. 
\begin{proposition}
  Suppose that for each $k$, $p_k(Y)$ is conservative conditional on $S_k(Y)$ under $F\in M_k(Y)$. Then $(p_k)$ are jointly conservative given $\bM$.
\end{proposition}

Although conditioning on $S_k$ may appear daunting, in many cases it can be a simple matter, as we see next.

\section{Simplifications in Exponential Family Models}

\begin{model}[Exponential Family]\label{mod:exFam}
  Assume that we have finitely many sufficient statistics $T_j(Y)$, and $F$ is in an exponential family after adding enough sufficient statistics to the model:
  \begin{equation}
    Y \sim \exp\left\{\sum_j \theta_j ' T_j(y) - \psi(\theta)\right\} h(y)
  \end{equation}
  Assume we add sufficient statistics to the model one at a time in the order $T_{J_0}, T_{J_1}, \ldots$, so that model $k$ is
  \begin{equation}
    Y \sim \exp\left\{\theta_{J_0} ' T_{J_0}(y) + \theta_{J_1}'T_{J_1}(y) + \cdots + \theta_{J_k}'T_{J_k}(y) - \psi(\cdot)\right\} h(y)
  \end{equation}
\end{model}

Our running Example~\ref{ex:linReg} is a simple case of Model~\ref{mod:exFam}. In that case,
\begin{align*}
  \theta_0 = 1/2\sigma^2, &\qquad T_0(Y) = \|Y\|^2,\\
  \theta_j = \beta_j/\sigma^2, &\qquad 
T_j(Y) = X_j'Y, \text{ for } j>0
\end{align*}

Another example is that of the Wishart, in which we have
\begin{equation}
  Y \sim \exp\left\{ \sum_{ij} \theta_{ij}y_{ij} - \psi(\theta)\right\}h(y)
\end{equation}

To obtain a $p$-value for testing the null hypothesis $M_{k-1}$ against the alternative $M_k \setminus M_{k-1}$, we will typically condition on $U_k=(T_1,\ldots, T_{k-1})$ simply as a way to eliminate $\zeta_k=(\theta_1,\ldots, \theta_{k-1})$ from the problem. In many cases, $p_{k-1}$ is a function only of $U_k$. If so, it is not necessary to condition explicitly on the previous $p$-values.

\section{Simulation: Sequential Model Selection Using the Lasso Path}

\end{document}

### 2014-12-31 knockoffs.tex
added knockoffs and updated paper

\documentclass{article}

\usepackage[round]{natbib}
\usepackage{amsmath,amssymb,amsthm,bm,enumerate,mathrsfs,mathtools}
\usepackage{latexsym,color,verbatim,multirow}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tikz}
\usepackage[margin=1.3in]{geometry}
\usetikzlibrary{shapes,arrows}
\tikzstyle{block} = [rectangle, draw, fill=white!20,
    text width=7em, text centered, rounded corners, minimum height=4em]
\tikzstyle{title} = [text width=7em, text centered, font=\bfseries]
\tikzstyle{line} = [draw, -latex']

\usepackage{mycommands}

\begin{document}

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{observation}[theorem]{Observation}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{fact}[theorem]{Fact}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{model}[theorem]{Model}

\theoremstyle{definition}
\newtheorem{example}{Example}

\newcommand{\cM}{\mathcal{M}}
\newcommand{\cN}{\mathcal{N}}
\newcommand{\cH}{\mathcal{H}}
\newcommand{\cD}{\mathcal{D}}
\newcommand{\FDR}{\textnormal{FDR}}
\newcommand{\FCR}{\textnormal{FCR}}
\newcommand{\crt}{\phi}
\newcommand{\M}{\mathcal{M}}
\newcommand{\cY}{\mathcal{Y}}
\newcommand{\cX}{\mathcal{X}}
\newcommand{\cV}{\mathcal{V}}
\newcommand{\bX}{\mathbf{X}}
\newcommand{\x}{\mathbf{x}}
\newcommand{\Gv}{\;\;\big|\;\;}
%\newcommand{\cP}{\mathcal{P}}
\newcommand{\proj}{\cP}
\newcommand{\pow}{\text{Pow}}
\newcommand{\sF}{\mathscr{F}}
\newcommand{\cF}{\mathcal{F}}
\newcommand{\sC}{\mathscr{C}}
\newcommand{\hJ}{\widehat{J}}
\newcommand{\bH}{\mathbf{H}}
\newcommand{\bM}{\mathbf{M}}
\newcommand{\hK}{\widehat{K}}
\newcommand{\leqAS}{\overset{\textrm{a.s.}}{\leq}}

\newcommand*\mystrut{\vrule width0pt height0pt depth1.5ex\relax}
\newcommand{\underlabel}{\underbracket[1pt][.5pt]{\mystrut \quad\;\; \sub \quad\;\; }}
\newcommand{\JTcomment}[1]{{\color{blue}{(JT: \bf \sc #1) }}}
\newcommand{\WFcomment}[1]{{\color{red}{(WF: \bf \sc #1) }}}

\title{A Construction of Knockoffs for Generic Exponential Family Models}
\maketitle

\begin{abstract}
  We construct ``knockoff'' sufficient statistics having the
  pairwise exchangeability property of \citet{barber2014controlling}
\end{abstract}

\section{Goal}

Consider the linear model
\begin{equation}
  Y \sim \cN(X\beta, \sigma^2I_n)
\end{equation}
which is an exponential family with $d$ sufficient statistics $T=X'Y$ where $X\in \R^{n\times d}$ (although it is not really necessary, assume $\sigma^2$ is known).

The knockoff method of~\citet{barber2014controlling} constructs $\widetilde X\in \R^{n\times d}$ with the properties
\begin{equation}
  \widetilde X'\widetilde X = X'X, \quad\text{and}\quad \widetilde X'X = X'X.
\end{equation}
In effect this creates a parallel set of ``control'' sufficient statistics $U=\widetilde X'Y$. The key stochastic property of $(T,U)$ is the ``pairwise exchangeability'' property: namely, that its law is invariant to swapping $T_M$ with $U_M$ for a ``null'' subset $M\sub \{1,\ldots,d\}$ and $\beta_M=0$.

\section{Abstract Construction for Generic Exponential Families}

The ``simple'' algorithm is as follows. We begin with an exponential
family model, given below:
\begin{equation}
  \L_\theta(T) = e^{\theta't - \psi(\theta)}f^0(t).
\end{equation}
Assume without loss of generality that $f^0$ is a valid density. We introduce the following notation for the conditional laws (under the ``global null'' $\theta=0$):
\begin{equation}
  f_j^0(t_j \gv t_{-j}) = \frac{f^0(t)}{\int f^0(t_1,\ldots,t_{j-1},v,t_{j+1},\ldots,t_d) \,dv}
\end{equation}

To construct the first knockoff sufficient statistic $U_1$ (the ``control'' for $T_1$), we can generate a random variable distributed as
\begin{equation}
  \L(U_1 \gv T) = f_1^0(u_1 \gv t_{-1})
\end{equation}
So that (under $\theta=0$) we obtain distribution 
\begin{equation}
  \L_0(T,U_1) = f^1(t, u_1) = f^0(t)f_1^0(u_1\gv t_{-1})
\end{equation}
and more generally,
\begin{equation}
  \L_\theta(T,U_1) = e^{\theta't - \psi(\theta)}f^0(t)f_1^0(u_1\gv t_{-1}).
\end{equation}

Note that if $\theta_1=0$, the law of $(T,U_1)$ is invariant to swapping $T_1$ with its knockoff $U_1$, since
\begin{equation}
  f^1(t,u) = f_1^0(u \gv t_{-1})\; f_1^0(t_1 \gv t_{-1}) \int f^0(v,t_{-1})\,dv
\end{equation}

To construct the second knockoff $U_2$, we generate
\begin{equation}
  \L(U_2 \gv T, U_1) = f_2^1(u_2 \gv t_{-2}, u_1)
\end{equation}
where $f_2^1$ is the conditional density of $T_2$ given $T_{-2}$ and $U_1$. We can continue on in this manner,
\begin{equation}
  \L(U_j \gv T, U_{1,\ldots,j-1}) = f_{j-1}^j(u_j \gv t_{-j}, u_{1,\ldots,j-1}),
\end{equation}
finally obtaining
\begin{equation}\label{eq:finalKnockoff}
    \L_\theta(T,U) = e^{\theta't - \psi(\theta)}f^d(t,u)
\end{equation}

\begin{theorem}
  For $M \sub \{1,\ldots,d\}$ assume $\theta_M=0$. Then $\L_\theta(T,U)$ is invariant to swapping any subset $T_M$ with $U_M$.
\end{theorem}
\begin{proof}
First, note that this claim is true under the global null. In that case, we can first swap $T_1$ for $U_1$ if $1\in M$, then swap $T_2$ for $U_2$ if $2\in M$, and so on. At step $j$, it is clear from our construction that swapping or not swapping does not change the distribution of $(T,U_{1,\ldots,j-1})$. It also does not change the conditional distribution of $U_{j,\ldots,d}$: since $f^{j-1}$ is invariant to swapping arguments $t_k$ and $u_k$ for $k<j$, so must $f_j^{j-1}$ be.

Next consider the weaker condition $\theta_M=0$. Under the global null, we have shown in particular that the conditional distribution $\L_0(T_M,U_M \gv T_{-M},U_{-M})$ is invariant to swapping $T_M$ and $U_M$. But this conditional law only depends on $\theta_M$, so the conditional distribution is still invariant to swapping. Moreover, the marginal distribution of $(T_{-M},U_{-M})$ is obviously invariant to swapping $T_M$ with $U_M$; as a result, the joint distribution of $(T,U)$ is invariant.
\end{proof}
Note that, in the above proof, it was not really necessary that we actually construct conditionally independent knockoffs at each stage. It would be sufficient simply to construct knockoffs that are conditionally exchangeable with the true variable at each stage.

As a result, we can use our knockoffs $U$ just as \citet{barber2014controlling} did. For any choice of statistics $W(T,U)$ where $W_j\mapsto -W_j$ when we swap $T_j$ with $U_j$, we can plug $W$ into any the knockoff procedures they proposed.

\section{Linear Regression}

For $n\geq 2d$, assume we begin with $T = X'Y$ and let $Z\in\R^{n\times d}$ have orthogonal columns with $X'Z=0$. Then, we can carry out the program above by first choosing 
\begin{equation}
  \widetilde X_1 = \proj_{X_{-1}}X_1 + \|\proj_{X_{-1}}^\perp X_1\|Z_1,
\end{equation}
so that $\widetilde X_1'X_j = X_1'X_j$ for $j\neq 1$, and $\|\widetilde X_1\| = \|X_1\|$. Then, $U_1 = \widetilde X_1'Y$ is a knockoff statistic for $T_1=X_1'Y$.

In the second step, we similarly take 
\begin{equation}
  \widetilde X_2 = \proj_{[X_{-1},U_1]}X_2 + \|\proj_{[X_{-1},U_1]}^\perp X_2\|Z_2
\end{equation}
so that $\widetilde X_2'X_j = X_2'X_j$ for $j\neq 2$, and $\widetilde X_1'\widetilde X_2 = \widetilde X_1'X_2 = X_1'X_2$. $U_2 = \widetilde X_2'Y$ is also a knockoff statistic for $T_2=X_2'Y$. We can continue in the obvious way to construct knockoffs $U_3,\ldots,U_d$.

\section{Poisson Process}

Next consider a simple non-Gaussian example, a Poisson process with sufficient statistics corresponding to overlapping windows. Suppose that $N$ is a Poisson process on $[0,300]$ with intensity equal to $e^{\lambda_1 1\{x<200\} + \lambda_2 1\{x>100\}}$. The sufficient statistics are:
\begin{equation}
  T_1=N([0,200]), \quad T_2=N([100,300]).
\end{equation}
Can we construct knockoffs $U_1,U_2$?

First, consider trying to construct conditionally independent knockoffs. We must construct $U_1$, which (conditionally on $T_2$) is independent of $T_1$ and has its null distribution. In particular, we can construct an artifical process $N^1$ which has $T_2$ points uniformly distributed on $[100,300]$, and is a Poisson process with unit intensity on $[0,100]$, then take $U_1=N^1([0,200])$.

Next, consider constructing $U_2$. Marginally under $\theta=0$, $N$ and $N^1$ are two Poisson processes of unit intensity, conditioned on having the same count falling into $[100,300]$. Then we have to condition further on the values of $N([0,200])$ and $N^1([0,200])$. This does not appear to be a trivial problem...

This is one case where it might be easier not to insist on conditional independence. For example, we could artificially generate $N$ on $[300,400]$ with unit intensity. We can then take $U_1=N([200,400])$ ($T_1$ and $U_1$ are conditionally exchangeable if $\theta_1=0$), and $U_2=N([0,100] \cup [300,400])$ (exchangeable with $T_2=N([100,300])$ given $T_1$ and $U_1$, if $\theta_2=0$).

\bibliographystyle{plainnat}
\bibliography{biblio}

\end{document}

### 2014-12-31 paper.tex
added knockoffs and updated paper

@@
 \usepackage{tikz}
-\usepackage[margin=1.3in]{geometry}
+\usepackage{geometry}
 \usetikzlibrary{shapes,arrows}
@@
 
-\begin{example}[Linear Regression Variable Selection]\label{ex:linReg}
-  Assume that
+\section{Introduction}
+
+In many data analysis problems, it is difficult to specify in advance a probabilistic model that is simultaneously expansive enough to adequately capture important characteristics of the sampling distribution, and parsimonious enough to yield precise and interpretable inferences.
+For example, in a linear regression problem, despite having measured thousands of covariates we commonly expect that only a few are actually relevant.
+That is, if the design matrix is $X\in \R^{n\times d}$, we expect that for some small subset $M\sub \{1,\ldots,d\}$ of the predictor variables, the model
 \begin{equation}
-    Y \sim N(X\beta, \sigma^2I).
+  Y \sim N(X_M\beta^M, \sigma^2I_n)
 \end{equation}
-  With many of the $\beta_j=0$. Our goal is to find a small subset of variables containing the true model.
-\end{example}
+approximately holds. In this setting, a common goal is to find a small subset of variables that include $M$.
+
+Analysis of variance (ANOVA) tables provide a classical tool for selecting from among a fixed sequence of nested linear models 
+\begin{equation}
+  \emptyset = M_0 \sub M_1 \sub \cdots \sub \{1,\ldots,d\},
+\end{equation}
+using $t$- or $F$-tests to compare each $M_k$ to the next larger proposed model $M_{k+1}$.
+
+In exploratory settings it is commonly impossible to specify a sequence of nested models in advance.
+Procedures abound for adaptively selecting a sequence of nested models, such as forward stepwise regression, LARS, the lasso, and many more \citep{ESL}.
+However, there are few accepted methods for choosing a final $M_k$ from among the sequence.
+In particular, using unadjusted $t$- or $F$-tests is no longer a valid option after adaptive selection.
+For example, if our algorithm begins by selecting the variable most correlated with $Y$, then the $t$-statistic for testing $M_1$ against $M_0$ is very likely to be large, even if $M_0$, the global null hypothesis, is true.
+
+Cross-validation treats $k$ as a tuning parameter and seeks a value $k$ that will produce a good predictive model.
+However, it does not yield statements of accuracy or inaccuracy for the models selected.
+
+We propose a modular method for selecting a stopping rule based on the selective framework of \citet{fithian2014optimal}.
+Our method generalizes the ANOVA table to nested sequences of linear or other probabilistic models that are adaptively selected using $Y$.
+Given data $Y$, a nested model sequence, and the algorithm used to form the sequence, we return a $p$-values $p_k$ for the test at each step $k$.
+If $M_K$ is the first correct model (note $K$ is random because the sequence is), the resulting $p$-values $p_{K}, \ldots, p_d$ are independent and (stochastically larger than) uniform, conditional on the selected model sequence.
+By combining our $p$-values with the sequential testing methods of \citet{wager2014}, we can estimate $K$, and control the probability of selecting an index much bigger than the true $K$.
+
+\WFcomment{Put in real data example?}
+
+\WFcomment{Put in non-Gaussian example?}
 
@@
 
-Suppose that after proposing a random sequence of nested models 
+Assume that we observe data $Y\in\cY$, with unknown sampling distribution $F$.
+Our goal is to choose a suitable model for $F$ after using some algorithm to propose an adaptive sequence of nested models:
 \begin{equation}
-  M_0(Y) \sub M_1(Y) \sub \cdots \sub M_d(Y) \sub \cF,
+  M_0(Y) \sub M_1(Y) \sub \cdots \sub M_d(Y) \sub \cF.
 \end{equation}
-we wish to sequentially test each $M_k$ against $M_{k-1}$.
-Let $\bM(Y)$ denote the sequence $(M_0,\ldots,M_d)$, and define its {\em completion index} $K$ to be the index of the first model containing $F$, i.e.
+Specifically, for $k=1,\ldots,d$, we wish to test whether $M_k$ represents a meaningful improvement on $M_{k-1}$; e.g., test whether adding the $k$th selected variable improves the model with the first $k-1$ variables.
+Let $\bM(Y)$ denote the sequence $(M_0,\ldots,M_d)$, and define its {\em completion index} $K$ to be the index of the first model containing $F$:
 \begin{equation}
@@
 \end{equation}
-Let $N_k$ denote the event $\{Y\in M_{k-1}\}$, i.e. the event that the null hypothesis for the test at stage $k$ is true. $Y\in N_k$ if and only if $k>K$.
+Let $N_k$ denote the event $\{F\in M_{k-1}(Y)\}\sub \cY$, the event that the null hypothesis is true for the test at stage $k$.
+That is, if $M_0$ and $M_1$ are false but $M_2$ is true, then $Y\notin N_0$ or $N_1$, but $Y\in N_2 \sub N_3 \sub \cdots$.
+Note $Y_k\in N_k$ if and only if $k>K$.
 
@@
   M_k:\, Y \sim 
-  N\left(\alpha + X_{J_{1}} \beta_{J_{1}} + \cdots X_{J_k}\beta_{J_k}, 
+  \cN\left(\alpha + X_{J_{1}} \beta_{J_{1}} + \cdots X_{J_k}\beta_{J_k}, 
   \;\sigma^2I\right).
 \end{equation}
-If the sequence $J_1,\ldots,J_d$ were deterministic instead of adaptive, then we might construct an ANOVA table for testing whether there was sufficient evidence to add each variable to the model in sequence.
+%If the sequence $J_1,\ldots,J_d$ were deterministic instead of adaptive, then %we might construct an ANOVA table for testing whether there was sufficient %evidence to add each variable to the model in sequence.
 If the true mean of $Y$ is linear in $X_{J_1}, X_{J_3},$ and $X_{J_5}$, then $K = 5$ and $Y\in N_k$ for $k=6,\ldots,d$.
@@
 
+\section{Discussion}
+
+We have seen that sequential selection of models --- the
+
+\bibliographystyle{plainnat}
+\bibliography{biblio}
+
 \end{document}

### 2015-04-28 completeNull/maxt.tex
added directories for subprojects

\documentclass{article}
\usepackage[left=1.25in,top=1.25in,right=1.25in,bottom=1.25in,head=1.25in]{geometry}
\usepackage{amsfonts,amsmath,amssymb,amsthm}
\usepackage{verbatim,float,url,enumerate}
\usepackage{graphicx,subfigure,epsfig,psfrag}
\usepackage{bm,dsfont,color,appendix}
\usepackage{natbib}
\usepackage{latexsym,graphicx}
%\usepackage{showkeys}

\newtheorem{algorithm}{Algorithm}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{remark}{Remark}
\newtheorem{corollary}{Corollary} 
\newtheorem{proposition}{Proposition}

\newcommand{\argmin}{\mathop{\mathrm{argmin}}}
\newcommand{\argmax}{\mathop{\mathrm{argmax}}}
\newcommand{\minimize}{\mathop{\mathrm{minimize}}}
\newcommand{\maximize}{\mathop{\mathrm{maximize}}}
\newcommand{\st}{\mathop{\mathrm{subject\,\,to}}}

\newcommand{\mm}{\mathcal{M}}
\newcommand{\pp}{\mathcal{P}}

\newcommand{\inc}{{\rm inc}}
\newcommand{\comp}{{\rm comp}}
\newcommand{\FM}{ {\rm FM}}
% \def\baselinestretch{1.7931}  
\pdfminorversion=4

\def\real{\mathbb{R}}
\def\epi{\mathbf{epi}}
\def\E{\mathbb{E}}
\def\P{\mathbb{P}}
\def\Var{\mathrm{Var}}
\def\Cov{\mathrm{Cov}}
\def\half{\frac{1}{2}}
\def\th{\mathrm{th}}
\def\tr{\mathrm{tr}}
\def\df{\mathrm{df}}
\def\dim{\mathrm{dim}}
\def\col{\mathrm{col}}
\def\row{\mathrm{row}}
\def\nul{\mathrm{null}}
\def\rank{\mathrm{rank}}
\def\nuli{\mathrm{nullity}}
\def\sign{{\mathrm{sign}}}
\def\supp{\mathrm{supp}}
\def\diag{\mathrm{diag}}
\def\aff{\mathrm{aff}}
\def\hy{\hat{y}}
\def\by{{\bf y}}
\def \sI {{\cal I}}
\def\bfeta{{\bf \eta}}
\def\bmu{{\bf \mu}}
\def\ty{{\tilde y}}
\newcommand{\Nu}{{\cal V}}
\def\bI {{\bf I}}
\def\by{{\bf y}}
\def\bSigma {{\bf\Sigma}}
\def\ty{\tilde{y}}
\def\hbeta{\hat{\beta}}
\def\tbeta{\tilde{\beta}}
\def\hu{\hat{u}}
\def\hv{\hat{v}}
\def\lone{1}
\def\ltwo{2}
\def\linf{\infty}
\def\lzero{0}
\def\T{^T}
\def\R{\mathbb{R}}
\def\cA{\mathcal{A}}
\def\cB{\mathcal{B}}
\def\cD{\mathcal{D}}
\def\cE{\mathcal{E}}
\def\cF{\mathcal{F}}
\def\cG{\mathcal{G}}
\def\cM{\mathcal{M}}
\def\cN{\mathcal{N}}
\def\cO{\mathcal{O}}
\def\cP{\mathcal{P}}
\def\cS{\mathcal{S}}
\def\cT{\mathcal{T}}
\def\cZ{\mathcal{Z}}
\def\Exp{\mathrm{Exp}}
\def\LARS{\mathrm{LARS}}
\def\RSS{\mathrm{RSS}}
\def\spa{\mathrm{span}}
\def\cl{\mathrm{cl}}
\def\relint{\mathrm{relint}}
\def\lin{\mathrm{lin}}
\def\codim{\mathrm{codim}}
\def\F{\mathbb{F}}

\usepackage[T1]{fontenc}
\def\qu{\textquotedbl}

\newcommand{\cd}{\overset{d}{\rightarrow}}
\newcommand{\sphere}[1]{O^{#1-1}}
\newcommand{\hittime}[2]{\gamma(#1,#2)}
\def\strong{S}
\def\lips{\mathcal{L}}
\def\dimens{\text{dim}}
\def\cone{\text{Cone}}
\def\critval{\gamma}
\def\tube{\mathrm{Tube}}
\def\V{\mathcal{V}}
\def\K{\mathcal{K}}
\def\numsignals{L}
\def\numevents{E}
\def\weight{\omega}
\def\winv{\theta}
\def\Pp{\mathbb{P}}
\def\Gfirst{G}
\def\Gsecond{{\bar{G}}}
\def\modelinfo{{\cal Z}}
\def\cS {\cal S}
\def\Prob{{\rm Pr}}

\title{	Bootstrap and analytic conditional tests for  forward stepwise regression}
%\author{Ryan Tibshirani$^1$, Robert J. Tibshirani$^2$, Alessandro Rinaldo$^1$ \and Larry Wasserman$^1$  in some order}
\author{Tibs times two}

%\and Robert Tibshirani$^1$}
%\date{\normalsize $^1$Stanford University,
%$^2$Simon Fraser University, 
%$^3$Carnegie Mellon University}

\begin{document}
\maketitle

\begin{abstract}
We study the standard ``max-t'' test for sequential testing in forward stepwise regression.
We provide computational and analytic approaches for carrying out this test,  that yield type- I error
control  conditional on the selection inherent in the procedure.
We compare the power of this procedure to the max-t test of Buja et al. and the recently proposed ``spacing'' test of \citet{TLTT2014}.
We propose two versions of the procedure: one that uses a  Gaussian error model and the other a  nonparametric bootstrap approach.

Keywords: Forward stepwise regression, multiple testing

\end{abstract}

\section{Introduction}
\label{sec:intro}
We consider observations $y \in \real^n$ drawn from a Gaussian model  
\begin{equation}
y = \theta + \epsilon, \;\;\; \epsilon \sim N(0,\sigma^2 I),
\label{eq:mod}
\end{equation}
Given a fixed matrix $X \in \real^{n\times p}$ of predictor variables,
we  derive new inferential tools for forward stepwise regression, that properly adjust
for the selection inherent to the procedure.

Assume that $X$ has standardized columns and $y$ has mean zero.
Suppose that we run forward stepwise regression for $k-1$ steps, yielding a model $M$ having  least squares coefficients
$\hat\beta$ and residuals $r=y-X\hat\beta$. Let $\tilde X=P_M^\perp X$, the residuals of each column of $X$ after projection onto $M$.
The  standard ``Max'-t" test for entering the next predictor defines
$$ T_j=\frac{\tilde x_j^Tr}{||\tilde x_j||};\; j\notin M$$
and chooses $\hat j ={\rm argmax}_{j\notin M}  |T_j|$.
In this paper we establish some  optimal properties of this sequential test, conditional on the selection, and provide analytic formulae for its implementation.
We call this procedure the {\em Conditional max-t} test.

We wish to stress that this test, of  course, is far from new. The max-t test is the standard test for forward stepwise regression.
Typically it is used without any attempt to adjust for the  effects of selection, that is, to test significance a standardized version of $T_j$ is compared a normal quantile
This yields a test with type I error that is far too large.  An exception is the discussion by Buja et. al. of \citet{LTTT2014}, who
describe their ``max-t'  test that uses permutations at each step to get proper control of type I error.

The conditional max-t test uses the same form of statistic of the max-t test of  described by Buja et. al in \cite{LTTT2014}.
Further, at the first step of forward stepwise (test of the global null), the tests are the same.
They both use parametric or nonparametric sampling of the data to estimate p-values.
it is at subsequent steps that they differ. The max-t test computes residuals from the current model and then applies the global test
to these residuals. In the process, it ignores the selection that has taken place up to the current step. The
conditional max-t test conditions on this selection and uses resamples conditional on this selection.

Because the max-t test ignore the effect of selection, it will tend to be too conservative and lose power in the process.
 Figure \ref{fig:buja} demonstrates this effect in a small example with  three  signal
variables. 
\begin{figure}[hbtp]
\centering
\includegraphics[width=\textwidth]{bujaplot.pdf}  
\caption{\em Simulated data: $n=20, p=10, \sigma=1$, $\beta=(2,2,2,0,\ldots 0)$. Shown are the p-values for the first 6 steps of forward stepwise regression over 2000 realizations.
In steps 4-6, we only report p-values for truly null case, that is, realizations for which the first 3 predictors are chosen in the first 3 steps.
The power of the max-t and conditional max-t tests (parametric residual  bootstrap and its conditional version) at $\alpha=.10$. are indicated in the bottom right of each of the first 3 panels.}
\label{fig:buja}
\end{figure}
The details are in the Figure caption. We see that the tests are identical within numerical simulation error in the first step, but the max-t test has lower power
in steps 2 and 3.  After that, it becomes more and more conservative.

\section{Optimal tests after forward selection}
\subsection{Hypothesis testing in regression}
\citet{fdrlasso} present a clear definition of the different kinds of hypothesis tests that are possible in stepwise regression.
We summarize their definitions here, as they provide an important framework for our comparisons.
Suppose that we have a ${n \times p}$ data  matrix $X$  and an outcome vector 
$y$, and seek to fit the linear regression model
$$ \hat y =X \beta ^*$$
using a sparse weight vector $\beta^*$. Forward stepwise regression yields a sequence of nested models
$$ \emptyset = \mm_0 \subset \mm_1 \subset ... \subset \mm_p, \; \text{with} \; \mm_k = \{j_1, \, ..., \, j_k\}. $$
We wish to  pick one of the models $\mm_k$, and
set to zero all coordinates $\beta^*_j$ with $j \notin \mm_k$.
The $k$-th ordered hypothesis $H_k$ then tests whether
or not adding the $k$-th variable $j_k$ was helpful or not.

The null hypothesis $H_k$ that the addition of the  $k$-th variable
along the regression path was unimportant, can be formalized
in several ways:
\begin{itemize}
\item {\bf The Incremental Null:} Here we construct $H_k$ to measure whether model $\mm_k$ improves over $\mm_{k-1}$.
The null hypothesis states that the best regression
fit for model $\mm_{k-1}$ is the same as the best regression fit for $\mm_k$ or,
more formally, that
\begin{align}
\label{eq:incremental}
&H_k^{\inc}: \pp_{\mm_{k - 1}} X \beta^* = 
\pp_{\mm_{k}} X \beta^*, \; \text{where} \\
&\pp_{\mm} = X_{\mm} \left(X_{\mm}^\top X_{\mm}\right)^{\dagger} X_{\mm}
\end{align}
is a projection onto the column-span of $X_\mm$. Here, we
write $X_{\mm}$ for the matrix comprised of the columns of $X$ contained
in $\mm$, and $A^\dagger$ denotes the Moore-Penrose pseudoinverse of 
a matrix A. \citet{TLTT2014} develop tests for $H_k^\inc$ in the context of
both forward stepwise regression and least-angle regression.
\item {\bf The Complete Null:}  We consider the stronger null hypothesis
that the model $\mm_{k - 1}$  captures all the available signal. More specifically,
writing $\mm^*$ for the support set of $\beta^*$, we define
\begin{equation}
\label{eq:complete}
H_k^\comp: \mm^* \subseteq \mm_{k - 1}.
\end{equation} 
Tests of $H_k^\comp$ for various pathwise regression models
have been studied by, among others by \citet{LTTT2013} and \cite{LT2014}.
\item {\bf The Full-Model Null:} Here we test
\begin{equation}
\label{eq:abs}
H_k^\FM: j_k \in \mm^*,
\end{equation}
i.e., that the $k$-th variable added to the regression path belongs to the support set of
$\beta^*$. Despite its simple appearance, however, the hypothesis $H_k^\FM$ is
difficult to work with. The problem is that the truth of $H_k^\FM$ depends critically on variables
that may not be contained in $\mm_k$, and so $H_k^\FM$ will have a ``high-dimensional''
character even when $k$ is small. We do not consider the
Full-model Null in this  paper.
\end{itemize}
As we will make clear, the conditional max-t test is designed for testing the Complete Null hypothesis,  while the spacing test of \citet{TLTT2014}
is designed for the Incremental null setting. However the simplicity of the spacing test makes it a natural choice for the Complete Null hypothesis yes as well,
and we examine its performance  in that problem.
\subsection{The conditional max-t test}

To define our conditional max-t test and contrast it with other tests,  we need to make precise the notion of selection.
Suppose at a given step we have selected predictors (and their signs) $S_M$ before this step $(M$ for model) , and predictor (and its sign) $S_C$ at the current step $(C$ for current).
Let the support of the true model be $\cS$.
We test the complete null hypothesis
\begin{eqnarray}
H_0: \cS \subseteq M.
\end{eqnarray}
That is, the null hypothesis is that all signal predictors are contained in the model $M$, and all predictors not in $M$ have coefficients of zero.
Then the conditional max-t test  provides a cutpoint $c$ so that the selective type I error is equal to a target level $\alpha$:
\begin{eqnarray}
\Prob_{H_0} (|T_j| \geq t | S_M)=\alpha
\label{eqn:null}
\end{eqnarray}
More precisely,  $S_M$ is the set of $y$ that lead to selection of the same model $M$ and its signs, in the steps leading up to the current one.

Taking $Z$ to be an $n$-vector of standard normal variates , there are three probability mechanisms that we consider for the probability in (\ref{eqn:null}):
\begin{eqnarray}
y^*=P_My+ \sigma Z;  \;{\rm ``residual} \;{\rm  bootstrap}\;{\rm (RB)''}
\label{eqn:parboot}
\end{eqnarray}
and
\begin{eqnarray}
y^*&=&P_My+ \sigma (I-P_M) Z;  \;{\rm ``conditional}\;{\rm  residual} \;{\rm bootstrap (CRB)}- \sigma \;{\rm known''.} \cr
y^*&=&P_My+ k\cdot (I-P_M) Z;  \;{\rm ``conditional}\;{\rm  residual} \;{\rm bootstrap (CRB)} - \sigma \;{\rm unknown ''.}
\label{eqn:selective}
\end{eqnarray}

The first is a standard parametric residual bootstrap sample from the fitted model $P_M y$;
in the second and third model,s the residuals are constrained  to be orthogonal to $P_My$. This is equivalent to
conditioning $y^*$ on  the observed value of $P_My$, and is the test for the complete null hypothesis that is  derived from the selective framework of \citet{FST2014}.
As a result, under the reduced model it has exact type-I error control.  In the last model above, $k$ is chosen so that $||y^*||= ||y||$.
This corresponds to conditioning on the value of $||y||$.

For the test of the global null hypothesis, \citet{ACP2012}  show that under sparse alternatives where $p^{1-\gamma}$ coefficients are non-zero, and $3/4 \leq \gamma \leq 1$,
the max-t  test is asymptotically most powerful. This result does not address the selective  setup: that is, further steps of the forward stepwise procedure.

Hence we consider the alternative
\begin{eqnarray}
H_1: p^{1-\gamma}\;{\rm predictors}\; {\rm not}\;{\rm in}\; M\;{\rm  have}\; {\rm  nonzero}\;{\rm coefficients}
\end{eqnarray}
with $3/4 \leq \gamma \leq 1$.
Building on the result of \citet{ACP2012}, heuristically we would expect that  the conditional max-t test is the asymptotically most powerful selective test for this problem under appropriate conditions.
\subsection{Summary of the various tests studied}

Since a number of different tests are examined in this paper, we summarized them here.
\begin{enumerate}
\item {\em Max-t test} with Gaussian errors (\ref{eq:mod}) . This test does not condition on the selection done in previous steps.
\item {\em Max-t-RB test}. The conditional max-t test using bootstrap  sampling from the  model (\ref{eqn:parboot}).  This is done either parametrically (sampling from the Gaussian)
or non-parametrically (sampling from the model residuals).  The parametric sampling is discussed in Section \ref{sec:exact}, while 
the nonparametric version is illustrated in Section \ref{sec:bootstrap}.
\item {\em Max-t-RB approximate test}. 
 This is an analytic approximation to the  parametric bootstrap Max-t p-value, given below in Section (\ref{sec:analytic}).
\item {\em Max-t-CRB test}. The conditional max-t test using bootstrap  sampling from the reduced model as in (\ref{eqn:selective}).  This is done either parametrically (sampling from the Gaussian)
or non-parametrically (sampling from the model residuals).  T
\end{enumerate}

\section{Computation of the conditional max-t test p-values}

\subsection{Exact computation}
\label{sec:exact}
Given a selected model $M$, we need to estimate the probability in (\ref{eqn:null}).
For the first step of forward stepwise (for which  $M$  is empty), this is straightforward:  we can fix $X$ and generate permutations (or bootstrap samples)
of $y$, compute the corresponding test statistics $T_1^*, T_2^*, \ldots T_p^*$  and use this to estimate the tail probability directly.

For subsequent steps of
forward stepwise regression, the computation is much more challenging. Given a model $M$ with fit $\hat\mu_M$,  we  generate a bootstrap sample
\begin{eqnarray}
y_i^*=\hat\mu_i+ \sigma A \cdot Z_i, \; i=1,2,\ldots N
\label{eq:boot}
\end{eqnarray}
Here $A$ equals the identity for the parametric bootstrap test, and $(I-P_M)$ for the selective test.
(Alternatively we can do residual bootstrap sampling: see Section \ref{sec:bootstrap}) .
 Then we can directly estimate the tail probability in  (\ref{eqn:null}). However we need to retain
only realizations $y^*$ that fall in $S_M$, that is, yield the same selected model $M$.

The task is made much simpler by the fact that the selection event can we written as $S_M=\{y: \Gamma y \geq u\}$.
Hence we don't need to run the entire forward stepwise procedure on $y^*$: instead, we pre-compute $\Gamma$ and $u$ and then check
if $\Gamma y^* \geq u$. Further, since $\Gamma$ has on the  order of $2pk$ rows after $k$ steps,  we can save 
substantially in computation by checking the row-wise inequalities starting from the last row, wherethe inequality  it is most likely to fail.

An additional fact that makes the computation feasible with strong signals:  much  of the time the procedure applied to bootstrap samples
make the same selections as in the original sample. It is only when the selected variables are  noise that the great majority of bootstrap samples
are rejected. Hence we run the procedure for a few steps after  the estimated p-values start to rise, and then stop.

 {\em Importance sampling} is also very helpful here.
Suppose that we want to estimate the conditional cdf  of a statistic $T(y)$ under $N(\mu, \Sigma)$, denoted by
$F_\mu(t \mid S)$.
Then if we instead sample $y_1^*, \ldots y_B^*$ from $N(\mu_0,\Sigma)$,  giving estimates $T_1^*, T_2^*, \ldots T_b^*$, 
an unbiased estimate of $F_\mu(t)$ is given by
\begin{eqnarray}
\frac{1}{B} \sum_b w_b \cdot I(T_b^* \leq t)
\end{eqnarray}
where $w_b=\phi_{\mu,\Sigma}(y_b^*)/\phi_{\mu_0,\Sigma}(y_b^*)$.
The idea is that we can choose $\mu_0$ so that the probability of $y^*$  lying  in the selection set $S$ is larger than under
$\mu$ (ie. fewer rejections).

Here are more details.
We want to choose $\mu_0$ to decrease the proportion of rejections, but don't want $\mu_0$ to be too far from $\mu$, or else the weights
will become too extreme.
 Denote the least squares estimate for the selected model after $k$ steps by $\hat\beta^M$.
We construct $\mu_0$ as follows.
The expected maximum of $(p-k)$  $N(0,1)$ variates is $\sqrt{2\log(p-k)}$.
Thus for  each component $j$, if $|\hat\beta_j^M/{\rm \hat se}_j|< \sqrt{2\log(p-k)}$ we increase $\hat\beta_j^M$ away from zero 
so that it equals $\sqrt{2\log(p-k)}$. For signal predictors, the coefficient will usually be unchanged: but for noise predictors, this will inflate their
coefficient so that there will be fewer rejections.

Table \ref{tab:tab1} shows an example. For the most difficult  case ($p=30$,  8 signals, 10 steps) the procedure required  1.8M samples to obtain 300 good ones.

\begin{table}[ht]
\centering
\begin{tabular}{r|rrr}
  \hline
  p&\multicolumn{3}{c}{Number of signals}\\
 & 2 & 5 & 8 \\ 
  \hline
  10 & 0.08 & 1.62 & 1.23 \\ 
  20 & 0.10 & 1.11 & 14.90\\ 
  30& 0.09 & 0.35 & 24.81 \\ 
   \hline
\end{tabular}
\label{tab:tab1}
\caption{\em Time in seconds to estimate p-values, for nsteps= \# signals+2; min 300 bootstraps; n=50}
\end{table}

\subsection{An analytic approximation}
\label{sec:analytic}
In this section we provide an analytic approximation to the parametric bootstrap (\ref{eqn:parboot}).
First consider the global null case, and let $T_j=v_j^Ty$ be the least squares estimates for predictor $j$  and $T_{max}={\rm max}_j(|T_j|)$ with observed value $M$.
Let $G(t)$ be the CDF of $T_{max}$. The desired p-value is $1-G(M)$.
Recall  Sidak's  lemma (see e.g. \citet{tong1990}) which states that if $y \sim N(0,\Sigma)$ then
\begin{eqnarray}
\Pr(\cap\{ |y_i| \leq t_i\} \geq \prod \Pr(|y_i| \leq t_i)
\end{eqnarray}
with equality if the variables are independent. Using this fact, we can approximate the p-value  $1-G(M)$ by 
$1-\prod \Pr(|T_j| \leq M)$ and this will be conservative by Sidak's lemma.

Now after  $k$ steps of FS with $p$ predictors, denote the selection by $S$. Let $T_j=v_j^Ty$,  $T_{max}={\rm max}_j(|T_j|)$
with observed value $M$. Now at the first step, the value of $T_{\rm max}$ can be any value in $(-\infty, \infty)$.  But after $k$ steps, the selection  events allow us to narrow this range.
The simple max-t test  described by Buja et. al in \cite{LTTT2014} ignores this fact and uses the range $(-\infty, \infty)$ for all steps, which will make quite conservative (as seen in Figure \ref{eqn:bujaplot}).
The narrowing of the range is most easily seen when the predictors are uncorrelated. 
Then $|T_{max}|$ can be no larger than its value at the previous step.  Hence $G(m)$ is the  CDF of largest of the last $p - k + 1$ 
absolute normal order statistics, conditioned on being smaller than the $k$th  order statistic.

Consider then the general non-independent  case where we have made a selection $S$ after $k$ steps.
Let predictor $0$ be the one selected at the current step.
Let $G(t)$ be the CDF of $T_{max} \mid S$. 
Let $V_0^\pm$ be the limits derived from polyhedral lemma applied to $S$ and $v_0$.
This lemma is reviewed in the appendix, and allows us to write the selection set $y \in S$  as
$V_0^- \leq v_0^Ty\leq V_0^+$.

Let $V={\rm max}(|V_0^-|, |V_0^+|)$, $M'=min(V,M)$.
Define
\begin{eqnarray}
F(M')=\Bigl[\frac{\Phi(M'/\sigma)-\Phi(-M'/\sigma)}{\Phi(V/\sigma)-\Phi(-V/\sigma)}\Bigr]^{p-k}.
\label{eq:approx}
\end{eqnarray}
The quantity $1-F(M')$ is our proposed approximate p-value.
We conjecture that $G(M') \geq F(M)$  so that $1-F(M')  \geq 1-G(M) \sim U(0,1)$ under the null. Hence the test based on $1-F(M')$ is conservative under the null.
At the first step (global null)  we have $V=\infty$ and hence this result is already established above. For subsequent steps with independent predictors, it is easy show that 
$V$ equals the $k$th  largest value among the $|T_j|$ and again the result follows.

Now we sketch a proof of the general result.
First we give  a refinement of  Sidak's lemma, which we need later.

\medskip

{\bf Lemma 1.} 
Let $y \sim N(0,\Sigma)$ and define
$$S(t) = \frac{\Pr({\rm max} |y_i| \leq t )}{\prod_i \Pr( |y_i| \leq  t )},$$
Then $S(t)$ is non-increasing in $t$.  [We note that Sidak's lemma implies that $S(t) \geq 1= S(\infty)$.]
A proof, due to Roman Latala is given in the Appendix. We were unable to find the result in the literature.

Now consider $G(M \mid S)=\Pr({\rm max}_{j \notin S} (|T_j|\leq M \mid S)$.

PROOF IS NOT YET CORRECT!

We can write
\begin{eqnarray}
G(M \mid S, V_0^- ,V_0^+)&=&\Pr({\rm max}_{j \notin S} (|T_j|\leq M\mid V_0^- \leq v_0^Ty\leq V_0^+,V_0^- ,V_0^+)\\
&\geq &\Pr({\rm max}_{j \notin S} (|T_j|\leq M\mid -V \leq v_0^Ty\leq V, V_0^- ,V_0^+) )\\
&=&\Pr({\rm max}_{j \notin S} (|T_j|\leq M\mid  -V \leq v_j^Ty\leq V, j\notin S, V_0^- ,V_0^+)) \\
&\geq &\Pr({\rm max}_{j \notin S} (|T_j|\leq M'\mid -V \leq v_j^Ty\leq V, j\notin S, V_0^- ,V_0^+)\\
&=&  \frac{ {\Pr({\rm max}_{j \notin S} |T_j|\leq M'},V_0^- ,V_0^+) } { \Pr({\rm max}_{j \notin S} |T_j|\leq V, V_0^- ,V_0^+)}\\
&\geq & \frac{\Bigl[\prod_{j\notin S} \Pr(|T_j|\leq M', V_0^- ,V_0^+)\Bigr] } { \prod_{j\notin S} \Pr(|T_j|\leq V, V_0^- ,V_0^+]}\\
&=&\Bigl[\frac{\Phi(M'/\sigma)-\Phi(-M'/\sigma)}{\Phi(V/\sigma)-\Phi(-V/\sigma)}\Bigr]^{p-k} \\
\end{eqnarray}
Now $V_0^- ,V_0^+$ are independent of $v_j^Ty$ for $j\notin S$, so (13) holds if we integrate out wrt $V_0^- ,V_0^+$.

Explanations:

(14):  $|V| \geq |V_0^-|, |V_0^+|$.

(15)  predictor 0 was selected among $j \notin S$

(16)  $M' \leq M$.

(18)  follows from Lemma 1

(19) The variance of $v_j^Ty$ is $\leq \sigma$ and under the null each $T_j$ has mean zero.

\subsection{Example}

Figure \ref{fig:fig2} compares the spacing, exact and approximate conditional max-t tests in  two examples.
We see that conditional max-t test has substantially more power. The exact conditional max-tmax-t test test becomes
a little conservative in the top  right panel, as we move farther into the null. The reason is as follows.
Let $\hat \mu_0$ by the least squares fit for the model containing just the signal variables, and suppose that we have fit a larger model
$\hat\mu$ containing the signal variables plus some noise ones.  Then the correct procedure generates
$y^*=\hat\mu_0+\sigma Z$, but instead we generate
$$y'=\hat\mu+\sigma Z=y^*+(\hat\mu-\hat\mu_0)$$
Consider a contrast vector $v$ that is orthogonal to the column spaces yielding both $\hat\mu_0$ and $\hat\mu$.
Then 
$$v^T y'=v^T y^*+v^T(\hat\mu-\hat\mu_0)$$
The second term above in normal  mean zero and some variance $\tau^2$.
Thus the variance of $y'$ is larger than that of $y^*$ and the test becomes conservative.

\begin{figure}[hbtp]
\centering
\includegraphics[width=\textwidth]{fig2.pdf}  
\caption{\em $n=20, p=8$ ,  $\beta=(3,0,0\ldots)$ in top panels and $(3,3,0,0,\ldots)$ in bottom panels.
Estimated power at $\alpha=.10$ for spacing, approximate and exact conditional max-t tests: top left- .60, .80. 80;  bottom left- .59, .98,.98,; bottom middle-
.39, .56, .62. }
\label{fig:fig2}
\end{figure}

\subsection{Relation to the spacing test} 
\label{sec:spacing}
Suppose that we have carried out a forward selection  yielding a set of predictors and their signs, denoted by $S$.
Let $\theta={\rm E}(y)$ and suppose we wish to make inferences about some linear contrast $v^T \theta$ based on
our sample estimate $v^T y$. The common choice for $v$ is the vector that produces the partial regression coefficient for one of the predictors
in the chosen model. Note that $v$ is data-dependent.
\citet{TLTT2014} and \citet{LT2014} propose a closed form method that yields exact tests for $H_0: v^T\theta=0$ and exact confidence intervals
for$v^T\theta$, condition on the selection $S$.
From this description, we see that the spacing test is a test of the incremental null hypothesis.

\cite{FST2014} provide a general framework for inference after selection, and derive uniformly most tests  when the distribution of the data are in  the exponential family.
There are two scenarios in their paper: the {\em saturated model} for which only model ({\ref{eq:mod}) is assumed, and the {\em reduced model} in which we assume
that some submodel of ({\ref{eq:mod}) holds. \citet{FST2014} 
show that when $\sigma$ is known, the spacing test coincides with the uniformly most powerful test in the saturated model. Furthermore. if the predictors are orthogonal
the spacing test is optimal under the reduced model as well.

There is another conceptual difference between the two tests.
If we include the current selection, we can view the overall selection event as $S=\{S_P,S_c\}$. The spacing test conditions on all of $S$.
The conditional max-t test conditions only on $S_M$. Less conditioning  generally yields more power.

The following simple example contrast the two tests explicitly,
Suppose we have $y=(y_1,y_2) \sim N((\theta_1,\theta_2), I)$.
Assume that we select the variable $y_j={\rm max}(y_1,y_2)$.
If we observe $y_1>y_2$, the spacing test  uses  the null hypothesis $\theta_1=0$, and its p-value is easily shown to be
\begin{equation}
{\rm pv}_{\rm spacing}=\frac{1-\Phi(y_1)}{1-\Phi(y_2)}.
\end{equation}

On the other hand, the conditional max-t test uses  the null hypothesis $\theta=(0,0)$ and has p-value
\begin{equation}
{\rm pv}_{\rm max}=1-\Phi(y_1)^2
\end{equation}
Direct inspection shows that ${\rm pv}_{\rm max} < {\rm pv}_{\rm spacing}$  as long as $y_1>y_2 > 0$.

As a second example, \citet{TLTT2014} derive a sequential test  (``the spacing test'') for the Gaussian regression model, with an exact analytic formula for the p-values. This is one of the first
regression tests that properly account for the selection effects and has exact type I error control at each step.

The left panel of Figure \ref{fig:fig1}  shows a simple example of the conditional max-t test compared to the spacing test.
 There are just two orthogonal predictors, whose statistics $T_1, T_2 \sim N(\Delta,1)$.
 We select the largest of the two statistics in absolute value. The power of the conditional max-t test , for target type I error $\alpha=0.10$, is shown by the red curve, as a function of $\mu$.
  Note that at $\Delta=0$ the power is $0.10$, confirming that the test is properly calibrated,
In Figure \ref{fig:fig1} we can see that the spacing test has substantially less power in this example.
The case of two large equal effects turns out to be especially problematic for the spacing test. In the right panel we show the alternative $(\Delta, \Delta/2)$:
the gap between the power curves has narrowed, but the conditional max-t test is still superior.
We study the relationship of the conditional max-t test with the spacing test in Section \ref{sec:spacing}.

\begin{figure}[hbtp]
\centering
\includegraphics[width=.6\textwidth]{diab.pdf}  
\caption{\em Diabetes data: p-values from spacing test, and exact and approximate conditional max-t test}
\label{fig:diab}
\end{figure}

\subsection{Robustness to model misspecification}
Figure \ref{fig:bujaplot-unksig} shows the sequential p-values from a setting in which the reduced is misspecified.
Details are in the Figure caption.  We wee that the type I error of the conditional CRB test  started to drug above the nominal level in steps 
4,5, and 6.
\begin{figure}[hbtp]
\centering
\includegraphics[width=\textwidth]{bujaplot-fixnormy-unksigma.pdf}  
\caption{\em $n=20, p=8, \beta=c(4,3,2,rep(0,5),rep(1,8))$, pairwise correlation $0.5$; only $x_1, \ldots x_8$ are
observed; $\sigma$ unknown}
\label{fig:bujaplot-unknsig}
\end{figure}

\section{Diabetes data}
Here we apply the various selection procedure to some well-known known on diabetes.
These data consists of  observations on $442$ patients, with the response of interest being a quantitative measure of disease progression one year after baseline.
There are ten baseline variables---age, sex, body-mass index, average blood pressure,
and six blood serum measurements---plus quadratic terms, giving a total of 64 features.
\begin{figure}[hbtp]
\centering
\includegraphics[width=.6\textwidth]{diab.pdf}  
\caption{\em Diabetes data: p-values from spacing test, and exact and approximate conditional max-t test}
\label{fig:diab}
\end{figure}
The results are shown in Figure \ref{fig:diab}. We see that there are only 2 or 3 small spacing p-values, while both forms of the conditional max exhibit 5 small p-values.

\section{A nonparametric conditional max-t test }
\label{sec:bootstrap}
The non-parametric version of the conditional max-t test simply replaces the Gaussian generative model (\ref{eq:boot}) by bootstrap residual sampling:
\begin{eqnarray}
y_i^*=\hat\mu_i+ r_i^*,\; i=1,2,\ldots N
\label{eq:bootnp}
\end{eqnarray}
where $r_i*$ is sampled with replacement from the residuals $r_i=y_i-\hat\mu_i$.
We then apply the same exact computation scheme as outlined above.
By avoiding normality assumption for the errors,  the hope is that this procedure will be more robust that the parametric version.
Figure \ref{fig:slash} examines this in a small example with no signal.
In the top row, with Gaussian errors, we see that both methods provide type I error control.
In the bottom two, we have used long-tailed ``slash'' errors (a Gaussian divided by a uniform), and we use that the parametric  test becomes
badly anti-conservative.  However the nonparametric version holds up quite well.
\begin{figure}[hbtp]
\centering
\includegraphics[width=\textwidth]{slash.pdf}  
\caption{\em Parametric and nonparametric bootstrapping;. $n= 15, p=5$. Null model.
Gaussian or slash errors}
\label{fig:slash}
\end{figure}

\subsection{Analytic approximation to the nonparametric bootstrap test}
Given a selection $S$, let $G(M \mid S)=\Pr({\rm max}_{j \notin S} (|T^*_j|\leq M \mid S)$.
Let $V$ be the bounds  in (\ref{eq:approx}) obtained from polyhedral lemma.
Then the same style of  argument that  was used in the Gaussian case suggests the approximation
\begin{eqnarray}
G(M \mid S, V_0^- ,V_0^+) &\approx& \Pr({\rm max}_{j \notin S} (|T^*_j|\leq M\mid V_0(y)^- \leq v_0^Ty^*\leq V_0(y)^+,V_0(y)^- ,V_0(y)^+)\cr
&=& \frac{ {\hat F}(M)-{\hat F}(-M)} {  {\hat F}(V)-{\hat F}(V)}
\label{eq:bootapprox}
\end{eqnarray}
Here $\hat F$ is the bootstrap distribution of the maximum attained  at step $k$. No rejection sampling is needed for account for the selection.
Figure \ref{fig2boot} suggests that this approximation is potentially useful.

\begin{figure}[hbtp]
\centering
\includegraphics[width=\textwidth]{fig2boot.pdf}  
\caption{\em Approximation to non-parametric residual bootstrap p-values using (\ref{eq:bootapprox}). Here  $n=20, p=8$ and  $\beta=(3,0,0\ldots)$ in top panels and $(3,3,0,0,\ldots)$ in bottom panels.}
\label{fig:fig2boot}
\end{figure}

{\bf Proof of correctness of bootstrap procedure}

\section{Other topics and issues}
Here we briefly discuss some further topics.
\begin{itemize}
\item {\it Estimation of the noise variance}
\item{\it Backward stepwise selection}.
The method described here can be also be adapted to backward stepwise regression.
Here are some details.
Suppose that we have dropped some  $k$ predictors: call the drop event $D$.
Consider another  predictor  $j$ as a candidate to drop. Let $r=P_{D-j}^\perp y$, $\tilde X=P_{D-j}^\perp X$,
$W_j=\langle r, \tilde x_j\rangle/||\tilde x_j||$. Then we drop the predictor
\begin{eqnarray}
\hat j={\rm argmin }(|W_j|)
\end{eqnarray}
Denote the attained minimum by $M=W_{\hat j}$. Define
$$G(t)=P(M\geq t | D) $$
then our p-value is $1-G(W_{\rm min})$.
We can use the same form of construction as before to get an analytic approximation for $G$:
\begin{eqnarray}
F(M')=\Bigl[\frac{ (1-\Phi(M'/\sigma)-\Phi(-M'/\sigma))}{(1-\Phi(V/\sigma)-\Phi(-V/\sigma))}\Bigr]^{p-k}\
\label{eq:approx}
\end{eqnarray}
The quantity $1-F(M')$ is then our proposed approximate p-value.
\end{itemize}

\section{Results of various simulations}

\subsection*{Acknowledgements}

\appendix
\section*{Review of Conditional Gaussian inference after polyhedral selection}

This section describes  a powerful result on Gaussian contrasts after polyhedral selection, which
appears in \citet{TLTT2014} and \citet{exactlasso}. 
\begin{equation*}
y \sim N(\theta,\Sigma), 
\end{equation*}
where $\theta \in \R^n$ is unknown but $\Sigma \in \R^{n\times n}$ 
is known. This generalizes our original setup in \eqref{eq:mod} (in
that we allow for a general error covariance structure).  Consider a 
polyhedron 
\begin{equation*}
\cP = \{y : \Gamma y \geq u\},
\end{equation*}
where $\Gamma \in \R^{m\times n}$, $u \in \R^m$ are fixed, and the
above inequality is meant to be interpreted componentwise.  Recall
that we call $y\in\cP$ a polyhedral selection event. Our goal is to
make inferences about $v^T \theta$ conditional on $y \in \cP$.
Indeed $v=v(\cP)$ is allowed to be defined in terms of $\cP$.  The 
next result provides a key alternate representation for
$\cP$.

It is shown in \citet{TLTT2014} and \citet{exactlasso}.  that
for any $\Sigma,v$ such that $v^T \Sigma v \not= 0$, there exists explicitly computable
functions $\V^\mathrm{lo}(y), \V^\mathrm{lo}(y), \V^0(y)$ such that
\begin{equation}
\label{eq:truncate}
\Gamma y \geq u \iff
\V^\mathrm{lo}(y) \leq v^T y \leq \V^\mathrm{up}(y), 
\, \V^0(y) \leq 0,
\end{equation}
Moreover,  it turns out that the triplet $(\V^\mathrm{lo},\V^\mathrm{up},\V^0)(y)$ is
independent of $v^T y$.    The explicit  expressions for these quantities are given  
 in the Appendix.

Hence the distribution of 
any linear function $v^Ty$, conditional on the selection 
$\Gamma y \geq u$,  
can be written as a truncated Gaussian distribution, and this can be used for hypothesis testing and confidence interval
construction. 
\citet{TLTT2014} use this result to derive a test for the entry of predictors in least angle and forward stepwise regression.

\section*{ Proof of Lemma 1}  {\em Due to Rafal Latala, personal communication}

The standard proof of the Khatri-Sidak inequality  (e.g. \citet{J70}) shows that for any
symmetric convex set $K$ and symmetric Gaussian vector $X$ ,the function

$$t \rightarrow  \frac {\Pr(X\in K,|X_1|< t)}{\Pr(|X_1|<t)}$$
is nonincreasing on $(0,\infty)$.
This shows that the function
$$G(t_1,\ldots,t_n)=\frac{\Pr(|X_1|< t_1,\ldots,|X_n|<t)}{\prod_i \Pr(|X_i|<t_i)}$$
is coordinate-wise nonincreasing on$ (0,\infty)^n$, in particular
$S(t)=G(t,\ldots,t)$
is nonincreasing on $(0,\infty)$.

\bibliographystyle{agsm}
\bibliography{ryantibs}

\end{document}

### 2015-05-01 code/bivariateSelVSat.R
done with "single-step" part of talk

@@ -1,0 +1,22 @@
+B <- 10000
+mu <- c(4,4)
+pvals <- NULL
+for(b in 1:B) {
+  y <- mu + rnorm(2)
+  if(abs(y[1]) > abs(y[2])) {
+    pvals <- rbind(pvals, c(
+      integrate(function(x) abs(x)*dnorm(x)/integrate(function(u) u*dnorm(u), 0, 10)$value, 
+                abs(y[1]), 10)$value,
+      pnorm(-abs(y[1]))/pnorm(-abs(y[2]))
+    ))
+  }
+}
+mean(pvals[,1]<.05) # power of selected-model test
+mean(pvals[,2]<.05) # power of saturated-model test
+
+pdf("../talks/figs/bivariateSelVSat.pdf",width=4,height=3)
+par(xaxs="i",mar=c(4.1,4.1,0.2,0.7))
+plot(ecdf(pvals[,2]),xlim=c(0,1),main="",xlab=expression(p[1]),ylab="CDF")
+plot(ecdf(pvals[,1]),col="red",add=T)
+legend("bottomright",legend=c("Saturated","Selected"),col=1:2,lty=1)
+dev.off()

### 2015-05-01 code/sparseSim.R
state at the time of group meeting talk

@@ -1,0 +1,161 @@
+library(xtable)
+
+simulation.data <- read.csv("snr_5_alpha_05.csv")
+names(simulation.data)
+
+
+methods <- c(paste(rep(c("simple", "forward", "strong"),2),
+                   rep(c("selected","saturated"),each=3),
+                  sep="_"),
+             "knockoffs")
+error.types <- c("pscreen","FWER.mod","FDR.mod","FDR.var")
+error.rates <- matrix(NA,7,4,
+                      dimnames=list(methods,error.types))
+for(i in 1:6) {
+  R <- simulation.data[[paste0(methods[i],"_R")]]
+  k0 <- simulation.data[["completion_index"]]
+  V.mod <- pmax(0,R - k0)
+  V.var <- simulation.data[[paste0(methods[i],"_V_var")]]
+  error.rates[i,"pscreen"] <- mean(R >= k0)
+  error.rates[i,"FWER.mod"] <- mean(V.mod > 1)
+  error.rates[i,"FDR.mod"] <- mean(V.mod / pmax(R,1))
+  error.rates[i,"FDR.var"] <- mean(V.var / pmax(R,1))
+}
+
+# What's wrong here?
+with(simulation.data, table(knockoff_R - knockoff_V))
+with(simulation.data, summary(simple_saturated_screen))
+with(simulation.data, summary(simple_saturated_V_model))
+with(simulation.data, which.max(simple_saturated_V_model))
+simulation.data[46, c("completion_index", "simple_saturated_R", "simple_saturated_V_model")]
+simulation.data$completion_index[46]
+simulation.data$simple_saturated_R[46]
+simulation.data$simple_saturated_V_model[46]
+
+
+error.rates["knockoffs", "pscreen"] <- with(simulation.data, mean(knockoff_R - knockoff_V == 7))
+error.rates["knockoffs", "FDR.var"] <- with(simulation.data, mean(knockoff_V / pmax(knockoff_R,1)))
+                                           
+## Fix this later...
+colnames(error.rates) <- c("$p_{\\text{screen}}", 
+                           "$\\text{FWER}_{\\text{model}}$",
+                           "$\\text{FDR}_{\\text{model}}$",
+                           "$\\text{FDR}_{\\text{variable}}$")
[…]

### 2015-05-03 paper.tex
more work on bivariate example, start on paper

--- 2014-10-21 paper.tex
@@
 \newcommand{\bM}{\mathbf{M}}
+\newcommand{\tM}{\widetilde{M}}
+\newcommand{\tE}{\widetilde{E}}
 \newcommand{\hK}{\widehat{K}}
+\newcommand{\hk}{\hat{k}}
 \newcommand{\leqAS}{\overset{\textrm{a.s.}}{\leq}}
@@
 \title{Adaptive Sequential Model Selection}
+\author{William Fithian, Jonathan Taylor, Rob Tibshirani, \and Ryan Tibshirani}
 \maketitle
@@
 
-\begin{example}[Linear Regression Variable Selection]\label{ex:linReg}
-  Assume that
-  \begin{equation}
-    Y \sim N(X\beta, \sigma^2I).
-  \end{equation}
-  With many of the $\beta_j=0$. Our goal is to find a small subset of variables containing the true model.
+\section{Introduction}
+
+... and find the index of the smallest adequate model --- that is, the smallest model that cannot be falsified given available evidence.
+
+\begin{example}[Forward-Stepwise Linear Regression with the LASSO]
+  \citet{taylor2014exact}
 \end{example}
 
-\section{The Problem of Sequential Model Selection}
+\begin{example}[The LARS Algorithm in Regression]
+  \citet{taylor2014exact}
+\end{example}
 
-\subsection{Notation and Problem Setting}
+\begin{example}[Ever-Active Path in $\ell_1$-Regularized Methods]
+  \citet{taylor2014exact}
+\end{example}
 
-Suppose that after proposing a random sequence of nested models 
+\begin{example}[Principal Components Analysis]
+  As a second motivating example, consider model selection for   principal components analysis.
+In that case we are given a data matrix $X \in \R^{n\times d}$, with which we form a sample covariance matrix
+\[
+S = \frac{1}{n-1} \sum_{i=1}^n(x_i - \bar x)^2
+\]
+The first $d$ principal component loadings are the first $d$ eigenvectors of $S$, which call $u_1,\ldots, u_d$. These induce a sequence of nested Wishart models:
+\[
+M_0 \sub M_1 \sub \cdots \sub M_d
+\]
+in which
 \begin{equation}
-  M_0(Y) \sub M_1(Y) \sub \cdots \sub M_d(Y) \sub \cF,
+  M_k:\; (n-1) S \sim W_d\left(\lambda_0 I_d + \sum_{\ell=1}^k     \lambda_\ell u_i u_i', \;\;\; n-1\right).
 \end{equation}
-we wish to sequentially test each $M_k$ against $M_{k-1}$.
-Let $\bM(Y)$ denote the sequence $(M_0,\ldots,M_d)$, and define its {\em completion index} $K$ to be the index of the first model containing $F$, i.e.
-\begin{equation}
-  K(\bM) = \min \{k:\, F\in M_k\}.
-\end{equation}
-Let $N_k$ denote the event $\{Y\in M_{k-1}\}$, i.e. the event that the null hypothesis for the test at stage $k$ is true. $Y\in N_k$ if and only if $k>K$.
+This problem was studied in \citet{choi2014selecting}.
+\end{example}
 
-For example, suppose that we use the LARS algorithm to build up a linear regression model.
-Let $J_k(Y)$ be the index of the $k$th variable to have a nonzero coefficient in the LARS path, so that we add variables to the model in the order $X_{J_1}, X_{J_2}, \ldots, X_{J_d}$.
-It is natural to take model $M_k(Y)$ as
-\begin{equation}
-  M_k:\, Y \sim 
-  N\left(\alpha + X_{J_{1}} \beta_{J_{1}} + \cdots X_{J_k}\beta_{J_k}, 
-  \;\sigma^2I\right).
-\end{equation}
-If the sequence $J_1,\ldots,J_d$ were deterministic instead of adaptive, then we might construct an ANOVA table for testing whether there was sufficient evidence to add each variable to the model in sequence.
-If the true mean of $Y$ is linear in $X_{J_1}, X_{J_3},$ and $X_{J_5}$, then $K = 5$ and $Y\in N_k$ for $k=6,\ldots,d$.
+\subsection{Generic Setting}
 
-Once the algorithm for choosing the sequence $\bM(Y)$ has been specified, we will carry out tests for each comparison.
-The goal we have set for ourselves is meant to be analogous to sequential testing of ANOVA tables for a fixed sequence of models, though there certainly are other sequential model selection tasks that could be of interest.
+More generically, we observe data $Y \in \cY$, with unknown sampling distribution $F$. We then use $Y$ to generate an adaptive sequence of $d$ nested models
+\[
+M_0(Y) \sub M_1(Y) \sub \cdots \sub M_d(Y).
+\]
+Define the {\em completion index} $k_0(Y) = \min\{k:\; F \in M_k\}$, the index of the first correct model.
+
+We will consider two related problems. First, we will consider the problem of obtaining selective single-step $p$-values, where $p_k$ is a $p$-value for testing
+\[
+H_{0,k}:\; F\in M_{k-1}\quad \text{ vs. } \quad 
+H_{1,k}:\; F\in M_k\setminus M_{k-1},
+\]
+while adjusting for the fact that the models are chosen adaptively.
+As we will see, {\em selected-model} tests can be dramatically more powerful than {\em saturated-model} tests at early steps in the model path when most of the signal variables have not yet entered the model.
+
+The second half of the paper concerns {\em stopping rules} $\hk$, estimators of $k_0$.
+We will consider several stopping rules that operate on the full sequence $p_{[d]}$, including two powerful stopping rules proposed in \citet{gsell2013sequential}.
+These latter require guarantees on the joint law of the $p$-values.
+We will prove sufficient conditions for when the $p$-values are Gaussian.
+
+Although we will focus most of our attention and examples on the case of selecting a set of predictors in linear regression, essentially all of our results apply in generic exponential family models.
+
+\subsection{Which Null Hypothesis Should We Test?}
+
+There is some ambiguity involved in deciding how to generalize $z$- and $t$-tests to the selective case.
+For example, \citet{gsell2013sequential} describe three different null hypotheses that we could consider testing at step $k$.
+As discussed at some length in \citet{fithian2014optimal}, there are several different models that
+
+In some cases the design matrix of the full model may represent a  scrupulously curated set of features, and the analysts may know in advance that inferences with respect to the full model are of major scientific interest.
+For example, the scientist may believe, due to theoretical considerations, that a nonzero coefficient of $X_1$ after controlling for $X_2,\ldots,X_p$ would be evidence for a causal effect of $X_1$ on the response.
+
+If the full model has no special scientific status, however, then we see little advantage in insisting that all inferences should adjust for every other predictor in the full design matrix $X$.
+For example, suppose that $X$ contains gene expression measurements for all genes that happened to be measured by a microarry chip.
+Then $\beta_{j,\text{Full}}$ is already fairly arbitrary, since we would be controlling for a different set of genes if we had purchased the chip from a different manufacturer.
+
+\WFcomment{Also say: Controlling for more variables could take us farther from a causal effect.}
+
+\WFcomment{Also say: }
+
+\section{Inference for One Step}
+
+\subsection{}
+
+\subsection{}
+
+\section{Sequential Inference}
 
 \subsection{Stopping Rules}
-For our purposes, a {\em stopping rule} $\hK(Y)$ is simply an estimate of the completion index $K$, which we could use to ultimately choose the model $M_{\hK}$ from the sequence.
 
-For our purposes, a good stopping rule is one that does not go on refining the model unnecessarily once it is already adequate; e.g. we might want
-\begin{equation}
-  \P(\hK > K \gv \bM) \leqAS \alpha, \quad \text{ or } \quad \E\left(\frac{(\hK - K)\vee 1}{\hK \vee 1} \gv \bM \right) \leqAS \alpha
-\end{equation}
-Although it might initally seem more natural to require $\P(\hK < K \gv \bM) \leq \alpha$ (i.e., that we choose an adequate model with high probability) this goal is typically infeasible as stated.
-For example, suppose that all the coefficients in a linear regression model are $\beta_j=\ep>0$.
-As $\ep\to 0$, it becomes impossible to distinguish the true model from the global null model $\beta=0$.
+\subsection{The Excluded-Variable Irrelevance Property}
 
-In particular, we will focus on constructing a sequence of $p$-values $(p_1(Y),\ldots,p_d(Y))$. $p_k(Y)$ is a {\em conservative selective $p$-value} if
-\begin{equation}
-  \P_F\left(p_k(Y) \leq \alpha \gv \bM\right) \leqAS \alpha \text{ on } N_k.
-\end{equation}
-That is, conditional on the question asked, $p_k$ ought to be stochastically larger than uniform.
+We say that a selection algorithm $M_{[d]}(\cdot)$ has the {\em excluded-parameter irrelevance condition} (henceforth EPIC) if, given $M_k(Y)=m_k$, the sequence of models $M_{[k]}$ depends only on sufficient statistics of $m_k$.
+That is
 
-We say the sequence $(p_1,\ldots,p_d)$ are {\em jointly conservative selective $p$-values} if
-\begin{equation}
-  \P_F(p_k(Y) < \alpha_k, \forall k>K \gv \bM) \leqAS \prod_{k>K} \alpha_k, \quad \forall \alpha_1,\ldots, \alpha_d\in (0,1), F\in \cF
-\end{equation}
-Notice that $K$ is random but the conditional probability above still makes sense because $K$ is a function of $\bM$.
+Ever-active coefficients in a regularization path are 
+\begin{proposition}
+  Let $M_\infty$ be a model with parameters $\theta\in\Theta\sub \R^p$, and likelihood $L(\theta; Y)$. For $k\in [d]$, let $P_k(\theta)$ denote a regularizing penalty to use at step $k$. Define
+%  \begin{align*}
+%    \hat\theta^{k}(Y) &\triangleq 
+%    \argmin_{\theta\in\Theta} -\log L(\theta; Y) + P_k(\theta) \\
+%    E_k(Y) &\triangleq \left\{j:\; \hat\theta_j^\ell \neq 0 
+%      \text{ for any \ell \leq k} \right\}\\ 
+%    M_k(Y) &\triangleq \{F\in M_\infty:\; 
+%    \theta_j = 0,\; \forall j \notin E_k\}
+%  \end{align*} 
+  Then $M_d(\cdot)$ satisfies EPIC.
+\end{proposition}
+\WFcomment{Need to be able to prune away the times when $M_k=M_{k-1}$; otherwise how do you construct $p_k$ for that time?}
 
-Joint conservatism of $p$-values is a minor generalization of requiring the null $p$-values to be mutually independent given the model sequence $\bM$.
+\begin{proof}
+  Let 
+\end{proof}
 
 \begin{proposition}
-  Suppose that, given $\bM$, the $(p_k)_{k>K}$ are conditionally uniform and mutually independent. Then $(p_k)$ are jointly conservative.
+
 \end{proposition}
 
-A  benefit of having jointly conservative $p$-values is that we can apply the stopping rules of \citet{gsell2013sequential} to control error rates such as the model FDR and model FWER.
-The next section proposes a generic recipe for obtaining such $p$-values
-
-\subsection{A Sequential Selection Variable for Pathwise FDR}
-
-Define the {\em $k$th pathwise selection variable} as 
-\begin{equation}
-  S_k(Y) = (M_0,\ldots,M_k,p_1,\ldots,p_{k-1}).
-\end{equation}
-If for each $k$ the $p$-value $p_k$ is conservative conditional on $S_k$, then the $p$-value sequence $(p_k)$ is jointly conservative. 
-\begin{proposition}
-  Suppose that for each $k$, $p_k(Y)$ is conservative conditional on $S_k(Y)$ under $F\in M_k(Y)$. Then $(p_k)$ are jointly conservative given $\bM$.
-\end{proposition}
-
-Although conditioning on $S_k$ may appear daunting, in many cases it can be a simple matter, as we see next.
-
-\section{Simplifications in Exponential Family Models}
-
-\begin{model}[Exponential Family]\label{mod:exFam}
-  Assume that we have finitely many sufficient statistics $T_j(Y)$, and $F$ is in an exponential family after adding enough sufficient statistics to the model:
-  \begin{equation}
-    Y \sim \exp\left\{\sum_j \theta_j ' T_j(y) - \psi(\theta)\right\} h(y)
-  \end{equation}
-  Assume we add sufficient statistics to the model one at a time in the order $T_{J_0}, T_{J_1}, \ldots$, so that model $k$ is
-  \begin{equation}
-    Y \sim \exp\left\{\theta_{J_0} ' T_{J_0}(y) + \theta_{J_1}'T_{J_1}(y) + \cdots + \theta_{J_k}'T_{J_k}(y) - \psi(\cdot)\right\} h(y)
-  \end{equation}
-\end{model}
-
-Our running Example~\ref{ex:linReg} is a simple case of Model~\ref{mod:exFam}. In that case,
+Let $\sF_{k,\ell}$ denote the $\sigma$-algebra generated by $M_{[k]}$ and $p_{[\ell]}$.
 \begin{align*}
-  \theta_0 = 1/2\sigma^2, &\qquad T_0(Y) = \|Y\|^2,\\
-  \theta_j = \beta_j/\sigma^2, &\qquad 
-T_j(Y) = X_j'Y, \text{ for } j>0
+  \sF_{k,\ell} &= \sF(M_{[k]},p_{[\ell]})\\
+  \sF_0 &\underlabel_{\text{selection } 1} \sF_{1,0} \underlabel_{\text{inference } 1}
+  \sF_{1,1} \;\;\sub \cdots \sub\;\;
+  \sF_{d-1,d-1} \underlabel_{\text{selection } d} \sF_{d,d-1}
+  \underlabel_{\text{inference } d} \sF_{d,d}
 \end{align*}
 
-Another example is that of the Wishart, in which we have
-\begin{equation}
-  Y \sim \exp\left\{ \sum_{ij} \theta_{ij}y_{ij} - \psi(\theta)\right\}h(y)
-\end{equation}
+\section{Simulation: Sparse Linear Regression}
 
-To obtain a $p$-value for testing the null hypothesis $M_{k-1}$ against the alternative $M_k \setminus M_{k-1}$, we will typically condition on $U_k=(T_1,\ldots, T_{k-1})$ simply as a way to eliminate $\zeta_k=(\theta_1,\ldots, \theta_{k-1})$ from the problem.
-In many cases, $p_{k-1}$ is a function only of $U_k$.
-If so, it is not necessary to condition explicitly on the previous $p$-values.
-
-\section{Simulation: Sequential Model Selection Using the Lasso Path}
+\bibliographystyle{plainnat}
+\bibliography{biblio}
 

### 2015-05-03 code/bivariateSelVSat.R

@@ -1,22 +1,111 @@
+selected.p <- function(y) {
+  total.mass <- integrate(function(x) (1-2*pnorm(-x))*dnorm(x), 
+                          0, abs(y[1])+10)$value
+  tail.mass  <- integrate(function(x) (1-2*pnorm(-x))*dnorm(x), 
+                          abs(y[1]), abs(y[1])+10)$value
+  tail.mass / total.mass
+}
+
+saturated.p <- function(y) {
+  pnorm(-abs(y[1]))/pnorm(-abs(y[2]))  
+}
+
+right.wedge.mass <- function(mu) {
+  # probability that y1 > y2
+  p1 <- pnorm((mu[1]-mu[2])/sqrt(2))    
+  # probability that y1 > -y2
+  p2 <- pnorm((mu[1]+mu[2])/sqrt(2))
+  p1 * p2
+}
+
+rtruncnorm.right <- function(n, mu) {
+  # probability that y1 > y2
+  p1 <- pnorm((mu[1]-mu[2])/sqrt(2))
+  # sample y1 - y2
+  ydiff <- -qnorm(p1*runif(n))*sqrt(2) + mu[1] - mu[2]
+  # probability that y1 > -y2
+  p2 <- pnorm((mu[1]+mu[2])/sqrt(2))
+  # sample y1 + y2
+  ysum <- -qnorm(p2*runif(n))*sqrt(2) + mu[1] + mu[2]
+  y <- cbind((ysum+ydiff) / 2, (ysum - ydiff) / 2)
+}
+
+rtruncnorm <- function(n, mu) {
+  p.tot <- right.wedge.mass(mu) + right.wedge.mass(-mu) 
+  p.right <- right.wedge.mass(mu) / p.tot
+  is.right.lobe <- runif(n) < p.right 
+  y <- matrix(NA,n,2)
+  y[is.right.lobe,] <- rtruncnorm.right(sum(is.right.lobe),mu)
+  y[!is.right.lobe,] <- -rtruncnorm.right(sum(!is.right.lobe),-mu)
+  y
+}
+
+plot(rtruncnorm(1000,c(.5,3)),xlim=c(-5,5),ylim=c(-5,5),asp=1)
+
+
+set.seed(1)
 B <- 10000
-mu <- c(4,4)
-pvals <- NULL
+mu <- c(0,0)
+pvals <- matrix(NA,B,2)
+y <- rtruncnorm(B,mu)
 for(b in 1:B) {
-  y <- mu + rnorm(2)
[…]

### 2015-05-04 paper.tex
more edits of paper, cleaned up bivariate R script

@@
 \begin{abstract}
-
+  Many model selection algorithms produce a ``path'' of fits that can be veiwed as specifying a sequence of models of increasing complexity.
+Given such a sequence of models and the data set used to produce them, we consider the problem of choosing the least complex model that is not falsified by the data, while accounting for the fact that the model path is determined adaptively using the data.
+Extending the tools of Fithian, Sun and Taylor (2014), we construct a $p$-value for each step in the sequence.
+In the case of linear regression, our $p$-values improve on the power of the spacings test of \citet{taylor2014exact}, often dramatically.
+By combining our $p$-values with stopping rules proposed by \citet{gsell2013sequential}, we achieve adaptive control of the pathwise familywise error rate and false discovery rate.
 \end{abstract}
@@
 \section{Introduction}
+
+... generates a sequence of increasingly complex models. Our goal is to choose the simplest model that is not falsified by the available data.
 
@@
 
-\WFcomment{Also say: Controlling for more variables could take us farther from a causal effect.}
-
-\WFcomment{Also say: }
+\WFcomment{Also say: Controlling for more variables could take us farther from a causal effect; furthermore, coefficient in larger model is much harder to interpret.}
 
@@
 
-\subsection{}
+To begin, we consider the problem of constructing a valid selective $p$-value for a particular step. At step $k$, we construct $p_k(Y)$ to test
+\[
+  H_{0,k}:\; F \in M_{k-1}(Y)
+  \quad \text{ vs. } \quad
+  H_{1,k}:\; F \in M_k(Y) \setminus M_{k-1}(Y).
+\]
+The main complication arises from the fact that the null and alternative hypotheses are random.
+To simplify matters, we first consider deriving $p$-values for a fixed candidate pair $m_{k-1} \sub m_k$.
 
-\subsection{}
+The random variable  $p_{k,m_{k-1},m_k}(Y)$ is a valid {\em selective $p$-value} for the fixed candidate pair $(m_{k-1},m_k)$ if it is stochastically larger than uniform under the (fixed) null $m_{k-1}$, given that the pair $(m_{k-1},m_k)$ is selected.
+That is,
+\[
+\P_F\left(p_{k,m_{k-1},m_k}(Y) \leq \alpha \mid M_{k-1}(Y) = m_{k-1}\right) 
+\leq \alpha, \quad \forall F\in m_{k-1}, \alpha \in [0,1].
+\]
+Given selective $p$-values for each fixed candidate pair, we can construct a combined $p$-value for the random null $(M_{k-1},M_k)$: 
+\[
+p_k(y) = p_{k, M_{k-1}(y), M_k(y)}(y),
+\]
+which is a valid $p$-value on the event $\{F \in M_{k-1}(Y)\}$:
+\[
+\P_F\left(p_k \leq \alpha \mid M_{k-1}, \;M_k, 
+  \;F\in M_{k-1}\right) \leq \alpha, \quad \forall \alpha \in [0,1].
+\]
+Recall that, as usual, $F$ is not random, but $M_{k-1}$ is.
+
+Methods for one-step selective $p$-values are by now well-studied. \WFcomment{big list of references.} See \citet{fithian2014optimal} for a general treatment.
+
+\subsection{Selective $p$-Values in Regression}
+
+\subsection{Selected- and Saturated-Model inference}
+
+We illustrate the superior power of the selected-model test in early steps with an extended example.
+
+\begin{example}[Bivariate Regression]
+  Consider forward stepwise selection in a regression model with $n=p=2$, with identity design $X = I_2=\begin{pmatrix} 1 & 0 \\ 0 & 1\end{pmatrix}$ and known $\sigma^2=1$.
+We perform one step of forward stepwise and then test the null model with no variables against the model with one variable.
+
+We will choose variable 1 first on the selection event $A=\{|Y_1| > |Y_2|\}$, which is shown in yellow in Figure~\ref{fig:condSets}. In that case,
+\[
+M_0:\; Y\sim \cN(0,I_2), \quad\text{ and } 
+M_1:\; Y \sim \cN\left(\frac{\mu_1}{0}, \; I_2\right).
+\]
+The selected-model test compares $Y_1$ to its distribution under $M_0$ conditional on $A$, a test of $H_0:\;\mu_1=0$ in model $M_1$.
+
+By contrast, the saturated-model test is a test of $H_0:\; \mu_1=0$ in the model $M_{\text{sat}}:\; Y \sim \cN(\mu, I_2)$.
+Now, $\mu_2$ enters the problem as a nuisance parameter because the saturated-model test refuses to assume that $\mu_2=0$.
+To eliminate the nuisance parameter $\mu_2$, the saturated model test must condition on $Y_2$, and compare $Y_1$ to its null distribution given $A$ {\em and} the observed value of $Y_2$.
+
+Figure~\ref{fig:condSets} shows the conditioning sets for each model when $Y=(2.9,2.5)$.
+Next to it, Figure~\ref{fig:nullDists} shows the null distribution for the test statistic $Y_1$ in each case.
+The $p$-values for the selected and saturated models are 0.007 and 0.3, respectively.
+\end{example}
+
+The case illustrated in Figure~\ref{fig:condSets} exhibits a phenomenon that has been remarked upon elsewhere in the literature of saturated-model tests: when there are near-ties between strong variables that are competing to enter the model, the winning variable may have a very weak $p$-value.
+\WFcomment{add references}.
+Figure~\ref{fig:}
+
+While the selected-model $p$-value is not always smaller than the saturated-model $p$-value, the selected-model test is strictly more powerful when the selected model is correct; i.e., when $\mu_2=0$.
+Figure~\ref{fig:powCurves_0} shows the power curve for each test when $\mu_2=0$.
+While the selected-model test is more powerful, the difference between the two is relatively mild.
+The difference is much more pronounced when $\mu_2=4$, as shown in Figure~\ref{fig:powCurves_4}.
+
+\subsection{Sub-Path Sufficiency Principle}
 

### 2015-05-04 code/bivariateSelVSat.R

@@ -1,15 +1,88 @@
-selected.p <- function(y) {
+#########################################
+# Make plots of the conditioning sets and null distributions
+#########################################
+
+# Conditioning Sets
+pdf("../figs/bivariateSelVSat_condSets.pdf",height=4.3,width=4)
+par(mar=c(4.1,4.1,3.1,0.1))
+y <- t(c(2.9,2.5))
+plot(y,xlim=c(-5,5),ylim=c(-5,5),xlab=expression(Y[1]),ylab=expression(Y[2]),
+     main="Conditioning Sets")#"Full vs. Reduced Model: First Step",asp=1)
+polygon(c(0,10,10),c(0,10,-10),lty=2,col="#F4E918")
+polygon(c(0,-10,-10),c(0,10,-10),lty=2,col="#F4E918")
+abline(h=0)
+abline(v=0)
+text(2,.5,"A")
+text(y+c(.3,-.4),labels="Y")
+lines(c(y[2],10),c(y[2],y[2]),lwd=2,col="brown")
+lines(c(-y[2],-10),c(y[2],y[2]),lwd=2,col="brown")
+points(y,pch=16)
+dev.off()
+
+# Null Distributions
+pdf("../figs/bivariateSelVSat_nullDists.pdf",height=4.3,width=4)
+par(mar=c(4.1,4.1,3.1,0.1),yaxs="i")
+x <- seq(-6,6,.01)
+plot(x,(abs(x)>2.5)*dnorm(x)/2/pnorm(-2.5),ylim=c(0,1.4),lty=1,
+     col="brown",type="l",
+     main="Conditional Null Distributions",
+     ylab="Density",xlab=expression(Y[1]))
+selected.model.density <- (1-2*pnorm(-abs(x)))*dnorm(x) * 2
+polygon(c(x,0),c(selected.model.density,0),lty=2,col="#F4E918")
+lines(x,(abs(x)>2.5)*dnorm(x)/2/pnorm(-2.5),col="brown")
+legend("topleft",legend=c("Saturated Model","Selected Model","Observed Value"),lty=1:3,bg="white", col=c("brown","black","black"))
+abline(v=2.9,lty=3)
+dev.off()
+
+
+####################################
+##  Code to produce p-values
+####################################
+
+# y1 is a (vector of) values of y1
+selected.p <- function(y1) {
[…]

### 2015-05-04 paper.tex
fixed some bugs in paper

@@
 \newcommand{\hk}{\hat{k}}
+\newcommand{\cN}{\mathcal{N}}
 \newcommand{\leqAS}{\overset{\textrm{a.s.}}{\leq}}
@@
 \title{Adaptive Sequential Model Selection}
-\author{William Fithian, Jonathan Taylor, Rob Tibshirani, \and Ryan Tibshirani}
+\author{William Fithian, Jonathan Taylor, Rob Tibshirani, and Ryan Tibshirani}
 \maketitle
@@
 
-\begin{example}[Bivariate Regression]
+\begin{example}[Bivariate Regression]\label{ex:bivariate}
   Consider forward stepwise selection in a regression model with $n=p=2$, with identity design $X = I_2=\begin{pmatrix} 1 & 0 \\ 0 & 1\end{pmatrix}$ and known $\sigma^2=1$.
@@
 
-We will choose variable 1 first on the selection event $A=\{|Y_1| > |Y_2|\}$, which is shown in yellow in Figure~\ref{fig:condSets}. In that case,
+We will choose variable 1 first on the selection event $A=\{|Y_1| > |Y_2|\}$, which is shown in yellow in Figure~\ref{fig:bv_condSets}. In that case,
 \[
@@
 
-Figure~\ref{fig:condSets} shows the conditioning sets for each model when $Y=(2.9,2.5)$.
-Next to it, Figure~\ref{fig:nullDists} shows the null distribution for the test statistic $Y_1$ in each case.
+Figure~\ref{fig:bv_condSets} shows the conditioning sets for each model when $Y=(2.9,2.5)$.
+Next to it, Figure~\ref{fig:bv_nullDists} shows the null distribution for the test statistic $Y_1$ in each case.
 The $p$-values for the selected and saturated models are 0.007 and 0.3, respectively.
+These two plots are reproduced from \citet{fithian2014optimal}, in which the same example was presented in less detail.
 \end{example}
 
-The case illustrated in Figure~\ref{fig:condSets} exhibits a phenomenon that has been remarked upon elsewhere in the literature of saturated-model tests: when there are near-ties between strong variables that are competing to enter the model, the winning variable may have a very weak $p$-value.
+\begin{figure}
+  \centering
+  \begin{subfigure}[t]{.4\textwidth}
+    % source code: bivariateSelVSat.R
+    \includegraphics[width=\textwidth]{figs/bivariateSelVSat_condSets.pdf}
+    \caption{\WFcomment{Copied directly from FST14; reword} 
+      For $Y=(2.9,2.5)$, the selected-model conditioning set is
+      $A=\{y:\;|y_1|>|y_2|\}$, a union of quadrants,
+      plotted in yellow. The saturated-model conditioning set
+      is ${\{y:\; y_2=2.5\}\cap A} = {\{y:\;y_2=2.5, |y_1|>2.5\}}$,
+      a union of rays, plotted in brown.}
+    \label{fig:bv_condSets}
+  \end{subfigure}
+  \hspace{.1\textwidth}
+  \begin{subfigure}[t]{.4\textwidth}
+    % source code: bivariateSelVSat.R
+    \includegraphics[width=\textwidth]{figs/bivariateSelVSat_nullDists.pdf}
+    \caption{\WFcomment{Copied directly from FST14; reword} 
+      Conditional distributions of $Y_1$ under
+      $H_0:\mu_1 = 0$. Under the hypothesis
+      $\mu=0$, the realized  $|Y_1|$ is  quite large given $A$,
+      giving  $p$-value 0.015. By contrast, $|Y_1|$ is not too large
+      given $A \cap \{y:\; y_2=Y_2\}$, giving
+      $p$-value 0.3.}
+  \end{subfigure}
+  \caption{\WFcomment{Copied directly from FST14; reword} 
+    Contrast between the saturated-model and selected-model tests
+    in Example~\ref{ex:bivariate}, in which we fit a one-sparse model with
+    design matrix $\bX=I_2$. The selected-model test is based
+    on  $\L_0(Y_1 \gv A)$, whereas the saturated-model test is based
+    on $\L_0(Y_1  \gv  Y_2, A)$.}
+  \label{fig:bv_nullDists}
+\end{figure}
+
+The case illustrated in Figure~\ref{fig:bv_condSets} exhibits a phenomenon that has been remarked upon elsewhere in the literature of saturated-model tests: when there are near-ties between strong variables that are competing to enter the model, the winning variable may have a very weak $p$-value.
 \WFcomment{add references}.
-Figure~\ref{fig:}
+Figure~\ref{fig:bv_rocCurve} displays the cumulative distribution function for the first $p$-value when $\mu=\binom{4}{4}$, a very strong signal.
+While the selected model test has near perfect power, it is not uncommon for the saturated model test to produce large $p$-values, even in the range of 0.5-0.9.
+These large $p$-values arise exactly when there is a near tie between the variables.
 
-While the selected-model $p$-value is not always smaller than the saturated-model $p$-value, the selected-model test is strictly more powerful when the selected model is correct; i.e., when $\mu_2=0$.
-Figure~\ref{fig:powCurves_0} shows the power curve for each test when $\mu_2=0$.
+\begin{figure}
+  \centering
+  % source code: bivariateSelVSat.R
+  \includegraphics[width=.5\textwidth]{figs/bivariateSelVSat_rocCurve.pdf}
+  \caption{\WFcomment{Write caption here.}}
+  \label{fig:bv_rocCurve}
+\end{figure}
+
+Results in \citet{fithian2014optimal} show that the selected-model test is strictly more powerful when the selected model is correct; i.e., when $\mu_2=0$.
+Figure~\ref{fig:bv_powCurves_0} shows the power curve for each test when $\mu_2=0$.
 While the selected-model test is more powerful, the difference between the two is relatively mild.
-The difference is much more pronounced when $\mu_2=4$, as shown in Figure~\ref{fig:powCurves_4}.
+Interestingly, the difference is much more pronounced when $\mu_2=4$, as shown in Figure~\ref{fig:bv_powCurves_4}.
 
-\subsection{Sub-Path Sufficiency Principle}
+\begin{figure}
+  \centering
+  \begin{subfigure}[t]{.4\textwidth}
+    % source code: bivariateSelVSat.R
+    \includegraphics[width=\textwidth]{figs/bivariateSelVSat_powCurves_0.pdf}
+    \caption{\WFcomment{Write caption here.}}
+    \label{fig:bv_powCurves_0}
+  \end{subfigure}
+  \hspace{.1\textwidth}
+  \begin{subfigure}[t]{.4\textwidth}
+    % source code: bivariateSelVSat.R
+    \includegraphics[width=\textwidth]{figs/bivariateSelVSat_powCurves_4.pdf}
+    \caption{\WFcomment{Write caption here.}}
+  \end{subfigure}
+  \caption{\WFcomment{Write caption here.}}
+   \label{fig:bv_powCurves_4}
+\end{figure}
 
@@
 
-\subsection{The Excluded-Variable Irrelevance Property}
+\subsection{Sub-Path Sufficiency Principle}
 
-We say that a selection algorithm $M_{[d]}(\cdot)$ has the {\em excluded-parameter irrelevance condition} (henceforth EPIC) if, given $M_k(Y)=m_k$, the sequence of models $M_{[k]}$ depends only on sufficient statistics of $m_k$.
+We say that a selection algorithm $M_{[d]}(\cdot)$ satisfies the {\em sub-path sufficiency principle} (henceforth SPSP) if, given $M_k(Y)=m_k$, the sequence of models $M_{[k]}$ depends only on sufficient statistics of $m_k$.
 That is
 
-Ever-active coefficients in a regularization path are 
-\begin{proposition}
-  Let $M_\infty$ be a model with parameters $\theta\in\Theta\sub \R^p$, and likelihood $L(\theta; Y)$. For $k\in [d]$, let $P_k(\theta)$ denote a regularizing penalty to use at step $k$. Define
-%  \begin{align*}
-%    \hat\theta^{k}(Y) &\triangleq 
-%    \argmin_{\theta\in\Theta} -\log L(\theta; Y) + P_k(\theta) \\
-%    E_k(Y) &\triangleq \left\{j:\; \hat\theta_j^\ell \neq 0 
-%      \text{ for any \ell \leq k} \right\}\\ 
-%    M_k(Y) &\triangleq \{F\in M_\infty:\; 
-%    \theta_j = 0,\; \forall j \notin E_k\}
-%  \end{align*} 
-  Then $M_d(\cdot)$ satisfies EPIC.
+A model selection procedure derived from the ``ever-active'' coefficients of a regularized likelihood path always satisfies the SPSP.
+Let $M_\infty$ be a model with parameters $\theta\in\Theta\sub \R^p$, and likelihood $L(\theta; Y)$.
+For $r=0,\ldots,R$, let $P_r(\theta)$ denote a sequence of regularizing penalties, and define
+\begin{align}\label{eq:regPathDef_start}
+  \hat\theta^{r}(Y) &\triangleq 
+  \argmin_{\theta\in\Theta} -\log L(\theta; Y) + P_r(\theta) \\
+  \tE_r(Y) &\triangleq \left\{j:\; \hat\theta_j^s \neq 0 
+    \text{ for any } s \leq r \right\}
+\end{align}
+Note that while the sets $\tE_r$ are nested by definition, we could have $\tE_r = \tE_{r+1}$ for most values of $r$. Let $r(0)=0$, and for $k>0$ define
+\begin{equation}
+r(k) = \min\{s:\; \tE_s \neq \tE_{r(k-1)}\}.
+\end{equation}
+That is, the $r(k)$ are the indices where the active set changes. Let $E_k = \tE_{r(k)}$, and take
+\begin{equation}\label{eq:regPathDef_end}
+M_k \triangleq \{F\in M_\infty:\; \theta_j = 0,\; \forall j \notin E_k\}
+\end{equation}
+Model paths described in this way satisfy the SPSP, as we see next.
+
+\begin{proposition}[Regularized Likelihood Paths Satisfy SPSP]
+  Model paths specified as in (\ref{eq:regPathDef_start}--\ref{eq:regPathDef_end}) satisfy the sub-path sufficiency principle.
 \end{proposition}
-\WFcomment{Need to be able to prune away the times when $M_k=M_{k-1}$; otherwise how do you construct $p_k$ for that time?}
+\WFcomment{Actually, you just keep the times when the model changes... Need to recast notation a bit; indices of $\lambda$ (or $P$, here) need not match up with indices of model.}
 
@@
 
-\begin{proposition}
-
+\begin{proposition}[Forward Stepwise Paths Satisfy SPSP]
+  \WFcomment{Generalized forward stepwise}
 \end{proposition}
 
-Let $\sF_{k,\ell}$ denote the $\sigma$-algebra generated by $M_{[k]}$ and $p_{[\ell]}$.
+\begin{proof}
+  Let
+\end{proof}
+
+\WFcomment{This is a filtration interpretation when you have the appropriate sufficiency properties.} Let $\sF_{k,\ell}$ denote the $\sigma$-algebra generated by $M_{[k]}$ and $p_{[\ell]}$.
 \begin{align*}
@@
 
+\subsection{}
+
 \section{Simulation: Sparse Linear Regression}
+
+\section{Simulation: Principal Components Analysis}
+
+\WFcomment{Can we do this one?}
 

### 2015-05-04 code/bivariateSelVSat.R

@@ -136,11 +136,12 @@ par(xaxs="i",mar=c(4.1,4.1,0.2,0.7))
 plot(ecdf(pvals[,2]),xlim=c(0,1),main="",xlab=expression(p[1]),ylab="CDF")
 plot(ecdf(pvals[,1]),col="red",add=T)
 lines(c(max(pvals[,1]),1),c(1,1),col="red")
-legend("bottomright",legend=c("Saturated","Selected"),col=1:2,lty=1)
+abline(h=1,col="gray",lty=2)
+legend("bottomright",legend=c("Saturated","Selected"),col=1:2,lty=1,bg="white")
 dev.off()
 
 # Now, try for a grid of mu values
-mu1.vals <- c(seq(0,.1,.02),seq(.01,8,.2))
+mu1.vals <- c(seq(0,.1,.02),seq(.2,8,.2))
 mu2.vals <- c(0,4)#seq(0,5,1)
 set.seed(1)
 B <- 1E5
@@ -156,18 +157,13 @@ for(mu2 in mu2.vals) {
     nrej.sat[as.character(mu1), as.character(mu2)] <-
       sum(saturated.test(y[,1], y[,2]))
   }
-  #cat("finished for mu2 = ", mu2, "\n")
+  cat("finished for mu2 = ", mu2, "\n")
 }
 
 
 library(splines)
 
 #mu2 = 4
-plot(mu1.vals,nrej.sat[,"0"]/B)
-points(mu1.vals,nrej.sel[,"0"]/B,col="red")
-y <- rtruncnorm(B, c(2,0))
-points(2,mean(saturated.test(y[,1], y[,2])),col="blue",pch=10)
-
 
 ## Smooth power function using logistic regression with splines
 resp.sat <- cbind(nrej.sat[,"4"],B-nrej.sat[,"4"])
@@ -179,7 +175,7 @@ pred <- c(-rev(mu1.vals),mu1.vals)
 sm.pow.sat <- glm(resp.sat ~ ns(pred,40), family=binomial)
 sm.pow.sel <- glm(resp.sel ~ ns(pred,40), family=binomial)
 
-pdf("../figs/bivariateSelVSat_powCurves_4.pdf",width=6,height=5)
+pdf("../figs/bivariateSelVSat_powCurves_4.pdf",width=4.5,height=4.3)
 par(xaxs="i",mar=c(4.1,4.1,3.1,0.7))
[…]

### 2015-05-05 paper.tex
Added more on stopping rules

@@
 \newcommand{\tE}{\widetilde{E}}
+\newcommand{\tV}{\widetilde{V}}
+\newcommand{\tR}{\widetilde{R}}
 \newcommand{\hK}{\widehat{K}}
@@
 
+Having discussed methods for constructing single-step $p$-values, we turn now to the problem of constructing a {\em stopping rule}; that is, an estimator $\hk$ of the completion index $k_0(Y)\in \{0,\ldots,d,\infty\}$.
+$M_k$ is ``rejected'' if and only if $\hk>k$.
+Thus $\hk$ is the number of models we rejected, while $k_0$ is the number that we should have rejected.
+The number of type I errors, then, is $(\hk-k_0)_+$, while the number of type II errors is $(k_0-\hk)_+$.
+Define $V=(\hk-k_0)_+$, the number of excess rejections.
+
+Note that the type I error $V$ is defined in a ``model-centric,'' as opposed to a ``variable-centric,'' fashion: in regression, our model-wise $V$ counts the number of true null {\em models} incorrectly rejected, instead of counting the number of noise variables included in the final model.
+We can define the (model-wise) familywise error rate (FWER) and false discovery rate (FDR) respectively as $\P_F(V >0)$ and $\E_F[V/(\hk \vee 1)]$.
+
+We could have instead defined the ``variable-centric'' type I error $\tV$ as the number of noise variables incorrectly included in the final model, with $\tR$ denoting the total number of variables included.
+Here, by ``noise variable'' we mean one that has a non-zero coefficient in the full model.
+Using selective inference to control the variable-wise FWER $\P_\mu(\tV>0)$ and FDR $\E[\tV / (\tR \vee 1)]$ is an interesting topic for further study.
+
 \subsection{Stopping Rules}
 
-\subsection{Sub-Path Sufficiency Principle}
+We will consider stopping rules that depend on the sequence $p_1,\ldots,p_d$ of $p$-values, which have been studied in the literature by \WFcomment{literature review here.} We will focus on three such stopping rules: simple stop, proposed by \WFcomment{whom?}, and the more refined strong stop and forward stop, proposed by \citet{gsell2013sequential}.
+
+While strong stop and forward stop tend to give more powerful stopping rules, they require independence among the $p$-values.
+Sections~\ref{sec:modelSPSP}--\ref{sec:pvalSP} discuss conditions on the model sequence $M_{[d]}$ and the $p$-value sequence $p_{[d]}$ under which selective $p$-values are mutually independent.
+We will see that, typically, selected-model $p$-values are sequentially independent while saturated-model $p$-values are not.
+
+\subsubsection{Basic Stop:}
+
+\WFcomment{Is there a better name for this? Can we find it in the literature?}
+
+The most obvious sequential procedure is to reject at each step until the first time $p_k > \alpha$, which we can formalize as
+\[
+\hk_B(Y) = \min\left\{k \in \{1,\ldots,m\} :\;
+  p_k > \alpha\right\} - 1
+\]
+We will call this procedure ``basic stop.''
+
+If $k_0$ is fixed, then it is clear that $\hk_b$ controls the FWER at level $\alpha$ provided that $p_{k}$ is a valid $p$-value for each $k>k_0$. Then, 
+\[
+\P(V>0) = \P(p_{k_0+1} \leq \alpha) \leq \alpha
+\]
+\WFcomment{Trivial, but no doubt someone else proved this first...}
+
+Selectively valid single-step $p$-values do not necessarily guarantee FWER control when $k_0$ is random.
+In that case, we need a bit more, but it is sufficient that we have type I error control for $p_k$ conditional on $k_0=k-1$; i.e., not just conditional on $M_{k-1}$ being a correct model, but given that $M_{k-1}$ is the {\em first} correct model.
+
+\subsubsection{Strong Stop:}
+
+\citet{gsell2013sequential} propose another method for controlling the model-wise FWER, which they called ``strong stop.'' Define
+\[
+  \hk_{S}(Y) = \max\left\{k \in \{1,\ldots,d\} :\;
+    \exp\left(\sum_{i=k}^d \frac{\log p_i}{i}\right) 
+    \leq \frac{\alpha k}{d}\right\}
+\]
+Even if $p_k$ is a little larger than $\alpha$, say $\alpha=0.05$ and $p_k=0.06$, strong stop can still reject $M_{k-1}$ if the next three or four $p$-values are also relatively small.
+
+\citet{gsell2013sequential} show that if $k_0$ is fixed, then $\hk_S$ controls the FWER at level $\alpha$, as long as the null $p$-values are independent given the non-null ones.
+Specifically, they require that
+\[
+\P(p_{k_0+1} \leq \alpha_{k_0+1}, \ldots, p_d \leq \alpha_d
+\mid p_1, \ldots, p_{k_0}) \leq \prod_{\ell=k_0+1}^d \alpha_\ell
+\]
+
+\subsubsection{Forward Stop:}
+
+\subsection{Independent $p$-Values and Sufficiency Principles}
+
+\subsection{}
+
+\subsubsection{A Sub-Path Sufficiency Principle for SelectionAlgorithms}\label{sec:modelSPSP}
+
+\subsubsection{A Sub-Path Sufficiency Principle for SelectionAlgorithms}\label{sec:modelSPSP}
 
@@
 
-\WFcomment{This is a filtration interpretation when you have the appropriate sufficiency properties.} Let $\sF_{k,\ell}$ denote the $\sigma$-algebra generated by $M_{[k]}$ and $p_{[\ell]}$.
+\WFcomment{There is a filtration interpretation when you have the appropriate sufficiency properties.} Let $\sF_{k,\ell}$ denote the $\sigma$-algebra generated by $M_{[k]}$ and $p_{[\ell]}$.
 \begin{align*}

### 2015-05-05 code/bivariateSelVSat.R
added code to show saturated-model p-values not independent

@@ -210,3 +210,45 @@ abline(h=0:1,lty=2,col="gray")
 legend("bottomright",lty=1,col=1:2,legend=c("Saturated","Selected"),
        bg="white")
 dev.off()
+
+
+
+###########################################
+### Dependence of Saturated-Model p-Values
+###########################################
+
+# Second-step p-value for bivariate normal. Note that the saturated and selected models are the same in the second step.
+p2 <- function(y1, y2) {
+  1 - (1-2*pnorm(-abs(y2))) / (1-2*pnorm(-abs(y1)))
+}
+
+set.seed(1)
+B <- 1E5
+y <- matrix(abs(rnorm(2*B)),B,2) # Signs don't matter for the p-values
+p.vals <- matrix(NA,B,2)
+ymax <- apply(y,1,max)
+ymin <- apply(y,1,min)
+saturated.p1 <- saturated.p(ymax,ymin)
+#selected.p1 <- selected.p(ymax)
+both.p2 <- p2(ymax,ymin)
+cor(saturated.p1, both.p2)
+#cor(selected.p1, both.p2)
+
+bin.2 <- function(p.vals) cut(p.vals,breaks=seq(0,1,.2))
+
+rounded <- data.frame(p1 = bin.2(saturated.p1), p2 = bin.2(both.p2))
+tbl <- xtabs(~p1 + p2, data = rounded)
+tbl <- cbind(tbl, Total=rowSums(tbl))
+tbl <- rbind(tbl, Total=colSums(tbl))
+xtable(100*tbl/B, digits=1, align="l|ccccc|c")
+
+
+
+#rounded <- data.frame(p1 = bin.2(selected.p1), p2 = bin.2(both.p2))
+#tbl <- xtabs(~p1 + p2, data = rounded)
+#tbl <- cbind(tbl, Total=rowSums(tbl))
+#tbl <- rbind(tbl, Total=colSums(tbl))
+#xtable(100*tbl/B, digits=1, align="l|ccccc|c")
+
+
+image(xtabs(~rounded[,1] + rounded[,2]))

### 2015-05-05 paper.tex
filled in much of sequential section and simulation section

@@
 In the case of linear regression, our $p$-values improve on the power of the spacings test of \citet{taylor2014exact}, often dramatically.
-By combining our $p$-values with stopping rules proposed by \citet{gsell2013sequential}, we achieve adaptive control of the pathwise familywise error rate and false discovery rate.
+By combining our $p$-values with stopping rules proposed by \citet{gsell2013sequential}, we achieve adaptive control of the familywise error rate and false discovery rate.
 \end{abstract}
@@
 M_0:\; Y\sim \cN(0,I_2), \quad\text{ and } 
-M_1:\; Y \sim \cN\left(\frac{\mu_1}{0}, \; I_2\right).
+M_1:\; Y \sim \cN\left(\binom{\mu_1}{0}, \; I_2\right).
 \]
@@
 \]
-We will call this procedure ``basic stop.''
+We will call this procedure {\em basic stop}.
 
@@
 \[
-\P(V>0) = \P(p_{k_0+1} \leq \alpha) \leq \alpha
+\P(V>0) = \P(p_{k_0+1} \leq \alpha) \leq \alpha.
 \]
@@
 
-\citet{gsell2013sequential} propose another method for controlling the model-wise FWER, which they called ``strong stop.'' Define
+\citet{gsell2013sequential} propose another stopping rule that controls the model-wise FWER, which they called {\em strong stop}. Define
 \[
   \hk_{S}(Y) = \max\left\{k \in \{1,\ldots,d\} :\;
-    \exp\left(\sum_{i=k}^d \frac{\log p_i}{i}\right) 
+    \exp\left(\sum_{\ell=k}^d \frac{\log p_\ell}{\ell}\right) 
     \leq \frac{\alpha k}{d}\right\}
@@
 
-\citet{gsell2013sequential} show that if $k_0$ is fixed, then $\hk_S$ controls the FWER at level $\alpha$, as long as the null $p$-values are independent given the non-null ones.
+\citet{gsell2013sequential} show that if the completion index $k_0$ is fixed, then $\hk_S$ controls the FWER at level $\alpha$, as long as the null $p$-values are independent given the non-null ones.
 Specifically, they require that
-\[
+\begin{equation}\label{eq:indepCond_fixed_k0}
 \P(p_{k_0+1} \leq \alpha_{k_0+1}, \ldots, p_d \leq \alpha_d
 \mid p_1, \ldots, p_{k_0}) \leq \prod_{\ell=k_0+1}^d \alpha_\ell
-\]
+\end{equation}
+
+For random $k_0$, it is easy to see that the $p$-values simply need to satisfy~\eqref{eq:indepCond_fixed_k0} conditional on the value of $k_0$:
+\begin{equation}\label{eq:indepCond_random_k0}
+  \P(p_{k+1} \leq \alpha_{k+1}, \ldots, p_d \leq \alpha_d
+  \mid p_1, \ldots, p_{k_0}, \; k_0 = k) \leq \prod_{\ell=k+1}^d \alpha_\ell
+\end{equation}
+If~\eqref{eq:indepCond_random_k0} holds, then strong stop controls theprobability of $V>0$ conditionally on $k_0$, and thus also marginally.
 
@@
 
-\subsection{Independent $p$-Values and Sufficiency Principles}
+\citet{gsell2013sequential} propose another stopping rule, {\em forward stop}, that controls the model-wise FDR:
+\[
+  \hk_{F}(Y) = \max\left\{k \in \{1,\ldots,d\} :\;
+    -\frac{1}{k}\sum_{\ell=1}^k \log(1-p_\ell) \leq \alpha\right\}
+\]
+If $p_k$ is uniform then  $-\log(1-p_k)$ is an exponential random variable. If all of the null $p$-values are uniform and the others are zero, then only the null ones contribute to the sum, and
+\[
+\widehat{\text{FDR}}_k = -\frac{1}{k}\sum_{\ell=1}^k \log(1-p_\ell)
+\]
+is a Gamma random variable with mean $V/k$ and variance $V/k^2$.
+We stop the last time the estimate of FDR is less than $\alpha$.
+Forward stop requires the same condition independence condition~\eqref{eq:indepCond_random_k0} as strong stop does.
 
-\subsection{}
+\subsection{Saturated- and Selected-Model $p$-Values in Regression}
 
-\subsubsection{A Sub-Path Sufficiency Principle for SelectionAlgorithms}\label{sec:modelSPSP}
+If we wish to use strong stop and forward stop, we need the $p$-values to satisfy~\eqref{eq:indepCond_random_k0}, but there is no guarantee that single-step selective $p$-values are independent of each other.
 
-\subsubsection{A Sub-Path Sufficiency Principle for SelectionAlgorithms}\label{sec:modelSPSP}
+In fact, saturated-model $p$-values typically are not independent.
+Continuing with Example~\ref{ex:bivariate}, Table~\ref{tab:bv_twoWayTable} shows a two-way contingency table for the saturated-model $p$-values $(p_1(Y), p_2(Y))$, binned into cells of height and width 0.2.
+Because the global null is true, both $p$-values are uniform by construction, but the $p$-values are strongly dependent, with correlation $-48\%$.
+
+% latex table generated in R 3.0.2 by xtable 1.7-1 package
+% Mon May  4 21:11:15 2015
+\begin{table}[ht]
+  \centering
+  \begin{tabular}{l|ccccc|c}
+%    & & & $p_2$ & & & \\
+    \multicolumn{7}{c}{$p_2(Y)$}\\[5pt]
+    ${\large p_1(Y)}$ & (0,0.2] & (0.2,0.4] & (0.4,0.6] & (0.6,0.8] & (0.8,1] & \textbf{Total} \\ 
+    \hline
+    (0,0.2] & 1.0 & 2.7 & 4.2 & 5.6 & 6.7 & 20.1 \\ 
+    (0.2,0.4] & 1.4 & 3.4 & 4.5 & 5.2 & 5.5 & 20.0 \\ 
+    (0.4,0.6] & 2.3 & 4.3 & 4.7 & 4.5 & 4.2 & 20.0 \\ 
+    (0.6,0.8] & 4.2 & 5.4 & 4.3 & 3.4 & 2.7 & 20.0 \\ 
+    (0.8,1] & 11.1 & 4.3 & 2.3 & 1.4 & 1.0 & 20.0 \\ 
+    \hline
+    \textbf{Total} & 19.9 & 20.0 & 20.0 & 20.1 & 20.0 & 100.0 \\ 
+  \end{tabular}
+  \caption{Two-way contingency table of saturated-model $p$-values $(p_1(Y), p_2(Y))$ for Example~\ref{ex:bivariate}, after binning into cells of height and width 0.2.
+We report the percentage of $p$-value pairs falling into each cell out of one million simulations from the global null hypothesis, $\mu=0$.
+Both $p$-values are marginally uniform but strongly dependent, with a correlation of $-48\%$.}
+\label{tab:bv_twoWayTable}
+\end{table}
+
+By contrast, the selected-model $p$-values $(p_1,p_2)$ are independent.
+To see why, recall that, on the selection event $A$, $p_1$ is only a function of $Y_1$.
+Moreover, $p_2$ is a selective $p$-value for comparing the null
+\[
+M_1:\; Y \sim \cN\left(\binom{\mu_1}{0}, I_2\right)
+\]
+against the alternative
+\[
+M_2:\; Y \sim \cN\left(\mu, I_2\right), \text{ with } \mu_2 \neq 0.
+\]
+Because $\mu_1$ is a nuisance parameter in $M_1$, an exact selected-model test must condition on the corresponding sufficient statistic $Y_1$.
+But that means $p_2$ is uniform given $Y_1$, and therefore also uniform given $p_1(Y_1)$.
+
+A generalization of this argument shows that the $p$-values for the saturated-model tests are always independent:
+
+\begin{proposition}
+  The selected-model equal-tailed or UMPU $p$-values after forward-stepwise linear regression. satisfy~\eqref{eq:indepCond_random_k0}.
+\end{proposition}
+This is a simple consequence of the much more general results proven in Section~\ref{sec:sufficiency}, but it is instructive to see the proof in the relatively concrete case of linear regression.
+\begin{proof}
+  At step $k$, we are testing the null hypothesis that $\beta_j=0$ in     the model $Y\sim \cN(X_E\beta, I_2)$, for some active set $E\sub [p]$ of size $k$, and some variable $j\in E$.
+Let $A_{j,E}$ denote the selection event for a particular $j$ and $E$, and let $m_{k-1}(j,E)$ denote the null model with active set $X_{E\setminus j}$.
+
+On $A_{j, E}$, the selected-model $p$-value $p_k$ is based on the null distribution of the test statistic $X_j'Y$ conditional on the null sufficient statistics $X_{E\setminus j}'Y$ and on $Y \in A_{j,E}$.
+
+Moreover, on $A_{j,E}$, the entire sequence of previous $p$-values $p_{[k-1]}$ depend only on the sufficient statistics $X_{E\setminus j}'Y$ of $m_{k-1}(j,E)$.
+To see why, \WFcomment{complete this proof.}
+
+We have shown that, if $m_{k-1}(j,E)$ is true, then $p_k$ is uniform and independent of $p_{[k-1]}$ on $A_{j,E}$ for each $j$ and $E$. That is,
+\[
+\P_F\left(p_{k} \leq \alpha \mid p_{[k-1]}, Y \in A_{j,E} \right) = \alpha \quad \forall F \in m_{k-1}.
+\]
+It follows that 
+\[
+\P_F\left(p_{k} \leq \alpha \mid p_{[k-1]}, F \in M_{k-1} \right) = \alpha \quad \forall F.
+\]
+\end{proof}
+The proof turns on two properties of selected-model $p$-values, and one property of forward stepwise regression.
+
+As for the $p$-values, we used the fact that $p_{k,m_{k-1},m_k}$ is uniform conditional on the sufficient statistics of the null model $m_{k-1}$.
+This is true for selected-model and saturated-model $p$-values alike.
+We also used the fact that $p_{k,m_{k-1},m_k}$ follows the {\em sufficiency principle} --- that is, it depends only on the sufficient statistics of $m_k$.
+This is {\em not} true of the saturated-model $p$-values: to compute them, we must know $\proj_{\eta}^\perp Y$, which depends on statistics that are ancillary for $m_{k-1}$.
+
+We also used the fact that, in the forward stepwise algorithm, if $m_k$ is the $k$th model then, on the event $M_k=m_k$, all the models up to step $k$ depend only on sufficient statistics of $m_k$.
+This is a sort of {\em sub-path sufficiency principle} which many other algorithms satisfy, as we will see in Section~\ref{sec:modelSPSP}.
+
+\subsection{Sequential Sufficiency Principles}\label{sec:sufficiency}
+
+\subsubsection{$p$-Values and the Sufficiency Principle}\label{sec:pvalSP}
+
+A $p$-value for testing model $M_{k-1}$ against $M_k$ satisfies the sufficiency principle if it is a function only of the sufficient statistics of $M_k$
+
+\subsubsection{A Sub-Path Sufficiency Principle for Model Selection Procedures}\label{sec:modelSPSP}
 
@@
 \end{equation}
+
 Model paths described in this way satisfy the SPSP, as we see next.
@@
 
-\subsection{}
+\subsection{Independent $p$-Values Via Selection Variables}\label{sec:selectionVariables}
 
-\section{Simulation: Sparse Linear Regression}
+As discussed in \citet{fithian2014optimal}, there are occasions when it is desirable for some reason to condition on more than just the selection event.
+On these occasions we can introduce a {\em selection variable} --- any random variable that is finer than the indicator of the selection event --- and condition on its value.
+Conditioning on a finer selection variable preserves validity conditional on the selection event, but tends to reduce the power of a test.
 
-\section{Simulation: Principal Components Analysis}
+For example, in linear regression after selecting variables via the lasso, \citet{lee2013exact} propose conditioning not only on the set of active variables but also on the signs of the active coefficients.
+The selection event corresponding to a given active set $E\sub [p]$ typically contains up to $2^{|E|}$ convex polytopes, each of which corresponds to a particular sign pattern $\hat z_E$.
+Thus, conditioning on the selection variable $\hat z_E$ can simplify computation considerably.
+
+We say that $p_{k,m_{k-1},m_k}$ is a selective $p$-value with respect to selection variable $S(Y)$ if it is uniform under $F\in m_k$ given $S$.
+
+To achieve joint error control properties among different tests, we may need to condition on a finer variable.
+
+\section{Simulation: Sparse Linear Regression}\label{sec:sparseReg}
+
+Next we compare several model selection procedures in simulation.
+We simlate from a linear regression model with $n=100$ observations and $p=40$ variables.
+The design matrix $X\in\R^{n\times p}$ is a random Gaussian design with pairwise correlations of 0.3 between predictor variables.
+
+The columns of $X$ are normalized to have length 1, and we simulate from $Y \sim \cN(X\beta,I_n)$, using a seven-sparse model with signal-to-noise ratio 5:
+\[
+\beta_j = \left\{\begin{matrix}5 & j = 1,\ldots,7\\ 0 &
+    j>7\end{matrix}\right.
+\]
+We use known $\sigma^2=1$, so that we can compare the saturated-model test with the selected-model test. For our selection algorithm, we use the entire forward-stepwise path, for all 40 steps.
+
+\subsection{Single-Step $p$-Values}
+
+For each step we compute one-step selected-model and saturated-model $p$-values, as well as nominal (unadjusted) $p$-values, conditioning on the signs of the active variables to make the problem more computationally tractable.
+Figure~\ref{fig:simulation_null_false} shows the power of all three tests for each of the first ten steps, conditional on the event that the null hypothesis is false.
+It is clear from Figure~\ref{fig:simulation_null_false} that the selected-model $p$-values are far more powerful than the saturated-model $p$-values.
+The nominal $p$-values are also quite powerful, but they do not have the correct level.
+
+\begin{figure}
+  \centering
+  % source code: ??? for simulation, sparseSim.R for plotting
+  \includegraphics[width=.8\textwidth]{figs/simulation_snr_5_alpha_05_null_false.pdf}
+  \caption{CDFs of saturated-model (black), selected-model (red), and nominal (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on testing a false null hypothesis at step $k$.
+The selected-model test is much more powerful than the saturated-model test.
+The nominal test appears powerful, but is not a correct $p$-value.}
+  \label{fig:simulation_null_false}
+\end{figure}
+
+\begin{figure}
+  \centering
+  % source code: ??? for simulation, sparseSim.R for plotting
+  \includegraphics[width=.8\textwidth]{figs/simulation_snr_5_alpha_05_null_true.pdf}
+  \caption{CDFs of saturated-model (black), selected-model (red), and nominal (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on testing a true null hypothesis at step $k$.
+The selected-model test is much more powerful than the saturated-model test.
+The nominal test is badly anti-conservative.}
+  \label{fig:simulation_null_true}
+\end{figure}
+
+\begin{figure}
+  \centering
+  % source code: ??? for simulation, sparseSim.R for plotting
+  \includegraphics[width=.8\textwidth]{figs/simulation_snr_5_alpha_05_noise_var.pdf}
+  \caption{CDFs of saturated-model (black), selected-model (red), and nominal (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on the event that the variable added at step $k$ is a noise variable in the full model.
+Here, none of the methods produce uniform $p$-values.
+The null hypothesis is false in most cases and so --- in our model-centric point of view --- rejection is the desired outcome.}
+  \label{fig:simulation_noise_var}
+\end{figure}
+
+Figure~\ref{fig:simulation_null_true} shows the distribution of $p_k$ for $k = 8, \ldots, 17$, given that the null hypothesis tested at step $k$ is correct (i.e., that $k_0< k$).
+Because the correct model is seven-sparse, $k=8$ is the first index for which the null can possibly be true.
+Both selective $p$-values are uniform by construction, but the nominal $p$-values are highly anti-conservative, as expected.
+
+Finally, as a warning against misinterpretation of our method, we include Figure~\ref{fig:simulation_noise_var} showing the first ten $p$-values for each method, conditional on event that the variable added at step $k$ is a noise variable in the full model.
+Now, none of the $p$-values are uniform.
+
+This is {\em not} a mistake in our implementation of the method, but rather a consequence of our ``model-centric'' point of view.
+If we try to add a noise variable to the model before we have included all the signal variables, then we are testing a false null hypothesis.
+The test rejects because there is much more signal to find, and as such, it is entirely appropriate for us to reject the null and continue growing the model.
+
+\subsection{Model-Selection Performance}
+
+If we combine the saturated-model or selected-model $p$-values with one of our three stopping rules, we can evaluate the model-selection performance of each method in terms of:
+\begin{itemize}
+\item its probability of selecting a correct model or $p_{\text{screen}}$,
+\item its model-wise FWER,
+\item its model-wise FDR, and
+\item its variable-wise FDR, where we use $\tV= \#\{\text{ noise variables included in } M_{\hk}\}$ instead of $V=(\hk-k_0)_+$.
+\end{itemize} 
+The last measure of performance is not explicitly controlled by any of the selective-inference methods, but we might nevertheless hope to perform reasonably.
+\WFcomment{Comment on results once they are correct.
+These numbers are wrong right now.}
+
+% latex table generated in R 3.0.2 by xtable 1.7-1 package
+% Fri May  1 09:29:28 2015
+\begin{table}[ht]
+  \centering
+  \begin{tabular}{llcccc}
+    \hline
+    & & $p_{\text{screen}}$ & $\text{FWER}_{\text{mod}}$ 
+    & $\text{FDR}_{\text{mod}}$ 
+    & $\text{FDR}_{\text{var}}$ \\ 
+    \hline
+    Selected & Simple & .290 & .000 & .002 & .039 \\ 
+    Selected & Forward & .559 & .027 & .020 & .066 \\ 
+    Selected & Strong & .041 & .014 & .008 & .042 \\ 
+    \hline
+    Saturated & Simple & .000 & .000 & .000 & .028 \\ 
+    Saturated & Forward & .014 & .000 & .000 & .030 \\ 
+    Saturated & Strong & .000 & .000 & .000 & .032 \\ 
+    \hline
+    Knockoffs & & .000 &  &  & .231 \\ 
+    \hline
+  \end{tabular}
+  \caption{\WFcomment{Comment on results once they are correct. These numbers are wrong right now.}}
+\end{table}
+
+\section{Simulation: Principal Components Analysis}\label{sec:pca}
 

### 2015-05-05 paper.tex
forward stepwise definition

@@
 
+\begin{model}[Generalized Sparse Model]\label{mod:genSparse}
+  Let $M_\infty$ be a model parameterized by $\theta\in \Theta \sub \R^p$:
+  \[
+  M_{\infty} = \{F_\theta:\; \theta \in \Theta\}.
+  \]
+  For subsets $E\sub [p]$ define the model $M_E$ with active coefficients $E$ as follows:
+  \[
+  \Theta_E = \{\theta:\; \theta_j = 0, \;\;\forall j \notin E\}, 
+  \quad M(E) = \{F_\theta:\; \theta\in \Theta_E\}.
+  \]
+\end{model}
+
+In Model~\ref{mod:genSparse}, the {\em forward stepwise algorithm} proceeds as follows: we first set $E_0$ arbitrarily, then at step $k=1,\ldots,d$, we define
+\begin{align}\label{eq:forwardDef_start}
+j_k &= \argmax_j \max_{\theta\in\Theta_{E_{k-1}}} \ell(\theta; Y)\\
+E_k &= E_{k-1} \cup \{j_k\}\\\label{eq:forwardDef_end}
+M_k &= M(E_k)
+\end{align}
+
+\begin{proposition}[Forward Stepwise Satisfies SPSP]\label{prop:forwardSPSP}
+  The forward stepwise algorithm defined in (\ref{eq:forwardDef_start}--\ref{eq:forwardDef_end}) satisfies the SPSP.
+\end{proposition}
+\begin{proof}
+  \WFcomment{Write proof.}
+\end{proof}
+
 A model selection procedure derived from the ``ever-active'' coefficients of a regularized likelihood path always satisfies the SPSP.
@@
 
-\begin{proposition}[Regularized Likelihood Paths Satisfy SPSP]
+\begin{proposition}[Regularized Likelihood Paths Satisfy SPSP]\label{prop:regPathSPSP}
   Model paths specified as in (\ref{eq:regPathDef_start}--\ref{eq:regPathDef_end}) satisfy the sub-path sufficiency principle.
@@
 
+\section{Discussion}
+
 \bibliographystyle{plainnat}

### 2015-05-05 code/python/compare_methods.py
initial version of python code

@@ -1,0 +1,257 @@
+import numpy as np, os
+import rpy2.robjects as rpy
+from rpy2.robjects import numpy2ri
+rpy.conversion.py2ri = numpy2ri.numpy2ri
+from matplotlib import mlab
+
+from selection.algorithms.tests.test_forward_step import test_full_pvals
+
+numpy2ri.activate()
+rpy.r("""
+strongstop <- function(p.values,alpha) {
+   d <- length(p.values)
+   lhs <- exp(rev(cumsum(rev(log(p.values)/(1:d))))) # LHS from G'Sell et al.
+   rhs <- alpha * (1:d) / d # RHS from G'Sell et al.
+   return(max(c(0,which(lhs <= rhs))))
+}
+
+forwardstop <- function(p, alpha) {
+   m <- length(p)
+   sums <- -(1/(1:m))*cumsum(log(1-p))
+   return(max(c(0, which(sums < alpha))))
+}
+"""
+)
+numpy2ri.deactivate()
+
+def simple_stop(pvalues, alpha):
+    if not np.all(pvalues <= alpha):
+        return np.min(np.nonzero(pvalues > alpha)[0])
+    else:
+        return pvalues.shape[0]
+
+def strong_stop(pvalues, alpha):
+    numpy2ri.activate()
+    rpy.r.assign('P', pvalues)
+    rpy.r.assign('alpha', alpha)
+    rpy.r('IDX = strongstop(P, alpha)')
+    value = int(np.array(rpy.r('IDX'))[0])
+    numpy2ri.deactivate()
+    return value
+
+def forward_stop(pvalues, alpha):
+    numpy2ri.activate()
+    rpy.r.assign('P', pvalues)
+    rpy.r.assign('alpha', alpha)
+    rpy.r('IDX = forwardstop(P, alpha)')
+    value = int(np.array(rpy.r('IDX'))[0])
+    numpy2ri.deactivate()
+    return value
+
+def run_one(n=100, p=40, rho=0.3, snr=4, alpha=0.05):
+    (X, 
+     y, 
+     beta, 
+     active, 
+     sigma, 
+     results, 
+     completion_index) = test_full_pvals(n=n,
+                                         p=p,
+                                         snr=snr,
+                                         rho=rho)
+    numpy2ri.activate()
+    rpy.r.assign('X', X)
+    rpy.r.assign('y', y)
+
+    # knockoff
+
+    rpy.r.assign('alpha', alpha)
[…]

### 2015-05-05 code/python/compute_pvalues.py

@@ -1,0 +1,100 @@
+import numpy as np
+import pandas as pd
+from selection.algorithms.forward_step import forward_stepwise
+from scipy.stats import norm as ndist
+
+def compute_pvalues(y, X, sigma=1.):
+    """
+    Parameters
+    ----------
+
+    y : np.float(n)
+        The target, in the model $y = X\beta$
+
+    X : np.float((n, p))
+        The data, in the model $y = X\beta$
+
+    sigma : np.float
+        Standard deviation of the gaussian distribution :
+        The covariance matrix is
+        `sigma**2 * np.identity(X.shape[0])`.
+        Defauts to 1.
+
+    Returns
+    -------
+
+    results : pd.DataFrame
+        DataFrame with variables (variable_id, 
+                            selective_pvalue,
+                            saturated_pvalue,
+                            nominal_pvalue)   
+
+    """
+
+    n, p = X.shape
+
+    FS = forward_stepwise(X, y, covariance=sigma**2 * np.identity(n))
+    results = []
+
+    for i in range(min(n, p)):
+        FS.next()
+        var_select, pval_select = FS.model_pivots(i+1, alternative='twosided',
+                                                  which_var=[FS.variables[-1]],
+                                                  saturated=False,
+                                                  burnin=2000,
+                                                  ndraw=8000)[0]
+        pval_saturated = FS.model_pivots(i+1, alternative='twosided',
+                                         which_var=[FS.variables[-1]],
+                                         saturated=True)[0][1]
+
+        # now, nominal ones
+
+        LSfunc = np.linalg.pinv(FS.X[:,FS.variables])
+        Z = np.dot(LSfunc[-1], FS.Y) / (np.linalg.norm(LSfunc[-1]) * sigma)
+        pval_nominal = 2 * ndist.sf(np.fabs(Z))
+        results.append((var_select, pval_select, pval_saturated, pval_nominal))
+            
+    results = np.array(results).T
+    return pd.DataFrame({'variable_selected': results[0].astype(np.int),
+                         'selected_pvalue': results[1],
+                         'saturated_pvalue': results[2],
+                         'nominal_pvalue': results[3]})
+
+def completion_index(selected, active_set):
+    """
+    Compute completion index from a sequence of 
+    selected variables.
+
+    The completion index is the first index
+    of selected that contains the full active_set.
+
+    Parameters
+    ----------
+
+    selected : []
[…]

### 2015-05-05 code/python/rules.py

@@ -1,0 +1,127 @@
+"""
+Stopping rules used
+"""
+
+import numpy as np
+
+def simple_stop(pvalues, alpha):
+    """
+    Compute the number of rejections using 
+    simple stop, the first time a p-value is above
+    alpha.
+
+    Parameters
+    ----------
+
+    pvalues : np.float
+
+    alpha : float
+
+    Returns
+    -------
+
+    num_rejections : int
+
+    """
+    if not np.all(pvalues <= alpha):
+        return np.min(np.nonzero(pvalues > alpha)[0])
+    else:
+        return pvalues.shape[0]
+
+def strong_stop(pvalues, alpha):
+    """
+
+    Compute the number of rejections using 
+    strong stop of `http://arxiv.org/abs/1309.5352`_
+
+    >>> strong_stop(np.array([0.5,0.6,0.7,0.8,0.9]), 0.05)
+    0
+    >>> strong_stop(np.array([0.001, 0.002, 0.0015, 0.0013, 0.05, 0.6]), 0.05)
+    3
+
+    In R:
+
+    > strongstop(c(0.001, 0.002, 0.0015, 0.0013, 0.05, 0.6), 0.05)
+    [1] 3
+    > strongstop(c(0.5,0.6,0.7,0.8,0.9), 0.05)
+    [1] 0
+
+    Parameters
+    ----------
+
+    pvalues : np.float
+
+    alpha : float
+
+    Returns
+    -------
+
+    num_rejections : int
+
+    Based on R code:
+    ----------------
+
+    strongstop <- function(p.values,alpha) {
+       d <- length(p.values)
+       lhs <- exp(rev(cumsum(rev(log(p.values)/(1:d))))) # LHS from G'Sell et al.
+       rhs <- alpha * (1:d) / d # RHS from G'Sell et al.
+       return(max(c(0,which(lhs <= rhs))))
+    }
+
+    """
+    n = pvalues.shape[0]
+    LHS = np.exp(np.cumsum((np.log(pvalues) / np.linspace(1., n, n))[::-1])[::-1])
+    RHS = alpha * np.linspace(1., n, n) / n
+    if np.any(LHS <= RHS):
+        return max(np.nonzero(LHS <= RHS)[0])+1
+    return 0
+
+
+def forward_stop(pvalues, alpha):
+    """
+
+    Compute the number of rejections using 
[…]

### 2015-05-05 paper.tex
Filled in most of "sub-path sufficiency principle" section

--- 2015-05-05 paper.tex
@@
 \newcommand{\tR}{\widetilde{R}}
-\newcommand{\hK}{\widehat{K}}
+\newcommand{\tL}{\widetilde{L}}
 \newcommand{\hk}{\hat{k}}
+\newcommand{\hr}{\hat{r}}
 \newcommand{\cN}{\mathcal{N}}
@@
   \begin{tabular}{l|ccccc|c}
-%    & & & $p_2$ & & & \\
+    \multicolumn{7}{c}{Saturated-Model $p$-Values 
+      (\% of $10^6$ Simulations)}\\
+    \hline
     \multicolumn{7}{c}{$p_2(Y)$}\\[5pt]
@@
     \textbf{Total} & 19.9 & 20.0 & 20.0 & 20.1 & 20.0 & 100.0 \\ 
+    \hline
   \end{tabular}
@@
 
-A $p$-value for testing model $M_{k-1}$ against $M_k$ satisfies the sufficiency principle if it is a function only of the sufficient statistics of $M_k$
+A $p$-value for testing model $M_{k-1}$ against $M_k$ satisfies the (conditional) {\em sufficiency principle} if, on the event $\{M_{k-1}=m_{k-1}, M_k=m_k\}$, it depends only on sufficient statistics of $M_k$.
+
+Based on the law
+\[
+\L_{m_{k-1}}\left(T_k \;\;\mid\;\; T_{k-1},\; M_{k-1}=m_{k-1},\; M_k=m_k\right)
+\]
+
+If $T(Y)$ is a complete sufficient statistic, then any exact $p$-value must have the latter property.
 
@@
 
-We say that a selection algorithm $M_{[d]}(\cdot)$ satisfies the {\em sub-path sufficiency principle} (henceforth SPSP) if, given $M_k(Y)=m_k$, the sequence of models $M_{[k]}$ depends only on sufficient statistics of $m_k$.
-That is
+We say that a selection algorithm $M_{[d]}(\cdot)$ satisfies the {\em sub-path sufficiency principle} (henceforth SPSP) if, on the event $M_k(Y)=m_k$, the sequence of models $M_{[k]}$ depends only on sufficient statistics of $m_k$.
+Again, if
 
-A model selection procedure derived from the ``ever-active'' coefficients of a regularized likelihood path always satisfies the SPSP.
-Let $M_\infty$ be a model with parameters $\theta\in\Theta\sub \R^p$, and likelihood $L(\theta; Y)$.
-For $r=0,\ldots,R$, let $P_r(\theta)$ denote a sequence of regularizing penalties, and define
+It will be useful to introduce the following notation:
+
+\begin{model}[Generalized Sparse Model]\label{mod:genSparse}
+  Let $M$ be a model parameterized by $\theta\in \Theta \sub \R^p$:
+  \[
+  M = \{F_\theta:\; \theta \in \Theta\}.
+  \]
+  For subsets $E\sub [p]$ define the submodel with active coefficients $E$ as follows:
+  \[
+  \Theta(E) = \{\theta:\; \theta_j = 0, \;\;\forall j \notin E\}, 
+  \quad M(E) = \{F_\theta:\; \theta\in \Theta_E\}.
+  \]
+  Let $L(\theta; Y)$ denote the likelihood for model $M$.
+\end{model}
+Note that, although we will make no assumptions about the form of $M$, the results of this section are only useful in cases where we can actually construct single-step $p$-values.
+At present, that caveat largely restricts us to using exponential family models.
+
+In Model~\ref{mod:genSparse}, the {\em forward stepwise} algorithm proceeds as follows: we first set $E_0(Y)$ arbitrarily, then at step $k=1,\ldots,d$, we define
+\begin{align}\label{eq:forwardDef_start}
+j_k &= \argmax_j \;\;\sup \left\{L(\theta; Y):\; \theta\in\Theta(E_{k-1} \cup \{j\})\right\} \\
+E_k &= E_{k-1} \cup \{j_k\}\\\label{eq:forwardDef_end}
+M_k &= M(E_k)
+\end{align}
+
+To make this definition more concrete, we can specialize it to several familiar examples:
+
+\begin{example}[Forward Stepwise Regression]
+In the case of linear regression with known $\sigma^2$, $p$ is the number of variables and $M$ is $\cN(X\beta, \;\sigma^2I)$.
+For active set $E \sub [p]$, $M(E)$ is the model $\cN(X_E\beta_E, \;\sigma^2I)$, and $E_0 = \emptyset$.
+Because the maximized log-likelihood for a linear regression is monotone in the residual sum of squares, the forward stepwise algorithm adds whichever variable reduces the residual sum of squares the most.
+
+For linear regression with unknown $\sigma^2$, there are $p+1$ parameters $(\beta, \sigma^2)$ and $E_0 = \{p+1\}$ because $\sigma^2$ is a parameter for every model.
+\end{example}
+
+\begin{example}[Principle Components Analysis]
+\WFcomment{Do this one.}  
+\end{example}
+
+All forward stepwise procedures satisfy the SPSP.
+
+\begin{proposition}[Forward Stepwise Satisfies SPSP]\label{prop:forwardSPSP}
+  The forward stepwise algorithm in Model~\ref{mod:genSparse} as defined in (\ref{eq:forwardDef_start}--\ref{eq:forwardDef_end}) satisfies the SPSP.
+\end{proposition}
+\begin{proof}
+  Let $A$ denote the event $\{M_k = M(E)\}$ for some fixed active set $E$.
+Letting $T(Y)$ be a sufficient statistic for $m_k$, the likelihood $L$ restricted to $\Theta(E) \times A$ is proportional to a function depending only on $T$.
+If $L_T$ is the likelihood of $T(Y)$, then
+\[
+L(\theta; y) = L_T(\theta; T(y)) \;L_{Y\mid T}(y), \;\; \text{ for } (\theta, y) \in  \Theta(E) \times A.
+\]
+Note that $L_{Y \mid T}(y)$ does not depend on $\theta$ since $T$ is sufficient for $M(E)$.
+
+For $y\in A$ and $\ell \leq k$, we have $E_\ell(y) \sub E$. This means that the maximizer in \eqref{eq:forwardDef_start} is attained for some $j\in E$. So, for $\ell \leq k$, on $A$,
+\begin{align}
+  j_\ell &= \argmax_{j\in E} \;\;\sup \left\{L(\theta; Y):\;
+    \theta\in\Theta(E_{\ell-1} \cup \{j\})\right\} \\
+  &= \argmax_{j\in E} \;\;\sup \left\{L_T(\theta; T(Y)):\;
+    \theta\in\Theta(E_{\ell-1} \cup \{j\})\right\}. \label{eq:onlyT}
+\end{align}
+Replacing $L$ with $L_T$ in~\eqref{eq:onlyT} is justified because $\Theta(E_{\ell - 1} \cup \{j\}) \sub \Theta(E)$ when $E_{\ell-1}\cup E$ and $j\in E$.
+
+Equation~\eqref{eq:onlyT} shows that $j_1,\ldots, j_{k-1}$ all depend on $Y$ only through $T(Y)$. As a result, it also follows that the sequence $M_{[k-1]}(Y)$ depends only of $T(Y)$.
+\end{proof}
+
+Another class of model selection procedures satisfying the SPSP is the sequence ``ever-active'' sets in regularized likelihood methods.
+For $r=0,1,\ldots$, let $P_r(\theta)$ denote a regularizing penalty, and define
 \begin{align}\label{eq:regPathDef_start}
-  \hat\theta^{r}(Y) &\triangleq 
+  \hat\theta^{r}(Y) &= 
   \argmin_{\theta\in\Theta} -\log L(\theta; Y) + P_r(\theta) \\
-  \tE_r(Y) &\triangleq \left\{j:\; \hat\theta_j^s \neq 0 
+  \tE_r(Y) &= \left\{j:\; \hat\theta_j^s \neq 0 
     \text{ for any } s \leq r \right\}
 \end{align}
-Note that while the sets $\tE_r$ are nested by definition, we could have $\tE_r = \tE_{r+1}$ for most values of $r$. Let $r(0)=0$, and for $k>0$ define
-\begin{equation}
-r(k) = \min\{s:\; \tE_s \neq \tE_{r(k-1)}\}.
-\end{equation}
-That is, the $r(k)$ are the indices where the active set changes. Let $E_k = \tE_{r(k)}$, and take
-\begin{equation}\label{eq:regPathDef_end}
-M_k \triangleq \{F\in M_\infty:\; \theta_j = 0,\; \forall j \notin E_k\}
-\end{equation}
 
-Model paths described in this way satisfy the SPSP, as we see next.
+While the sets $\tE_r$ are nested by definition, we could have $\tE_r = \tE_{r+1}$ for most values of $r$.
+We will take the sequence of {\em distinct} ever-active sets.
+Let $R_0=0$, and for $k>0$ let $R_k$ denote the (random) index where the active set actually changes:
+\begin{align}
+  R_k &= \min\{s:\; \tE_s \neq \tE_{R_{k-1}}\}\\
+  E_k &= \tE_{R_k}\\
+  \label{eq:regPathDef_end}
+  M_k &= M(E_k)
+\end{align}
+\WFcomment{Slightly sloppy definitions: note that $d$, the number of models, is random.
+$d$ could be 0 or could be infinite, depending on the data.
+Also, it isn't really necessary that the $r$ values be integers...}
 
-\begin{proposition}[Regularized Likelihood Paths Satisfy SPSP]
-  Model paths specified as in (\ref{eq:regPathDef_start}--\ref{eq:regPathDef_end}) satisfy the sub-path sufficiency principle.
-\end{proposition}
-\WFcomment{Actually, you just keep the times when the model changes... Need to recast notation a bit; indices of $\lambda$ (or $P$, here) need not match up with indices of model.}
-
-\begin{proof}
-  Let 
-\end{proof}
-
-\begin{proposition}[Forward Stepwise Paths Satisfy SPSP]
-  \WFcomment{Generalized forward stepwise}
+\begin{proposition}[Regularized Likelihood Paths Satisfy SPSP]\label{prop:regPathSPSP}
+Ever-active model paths derived from regularized likelihood methods as in (\ref{eq:regPathDef_start}--\ref{eq:regPathDef_end}) satisfy the sub-path sufficiency principle.
 \end{proposition}
@@
 \begin{proof}
-  Let
+  Again, let $A$ denote the event $\{M_k = M(E)\}$ for some fixed $E$, and define $T$ and $L_T$ as in the proof of Proposition~\ref{prop:forwardSPSP}.
+We can partition $A$ into subsets $A_r = A \cap \{R(k)=r\}$.
+
+On the event $A_r$, every $\hat\theta^s$ for $s\leq r$ is supported on $E$. That is, for $s \leq r$, we can write 
+\begin{equation}\label{eq:onlyT_Ar}
+  \hat\theta^{s} = 
+  \argmin_{\theta\in\Theta(E)} -\log L_T(\theta; T(Y)) + P_s(\theta) \quad \text{ on } A_r
+\end{equation}
+Equation~\label{eq:onlyT_Ar} guarantees that $M_{[k-1]}$ is a function of $T(Y)$ on $A_r$. That is, we can write
+\[
+M_{[k-1]}(Y) = M_{[k-1]}^r(T(Y)) \quad \text{ on } A_r
+\]
+It remains only to show that, on $A$, $R_k(Y)$ depends only on $T(Y)$; then we have
+\[
+M_{[k-1]}(Y) = M_{[k-1]}^{R(T)}(T) \quad \text{ on } A
+\]
+But $R_k$ is a function of $T$ on $A$ for the same reason: that $\hat\theta^s$ depends only on $T$ until other coefficients enter the model.
+\WFcomment{Would the proof be cleaner with $A_r$ defined as $A\cap \{R_k \leq r\}$?}
 \end{proof}
@@
 
+\WFcomment{Should we compute the FWER and FDR only conditional on screening? Strong / Forward / Simple should all be correct given screening since they condition on $p_{[k_0-1]}$.}
+
 % latex table generated in R 3.0.2 by xtable 1.7-1 package
@@
     \hline
-    & & $p_{\text{screen}}$ & $\text{FWER}_{\text{mod}}$ 
+    Method & Stopping Rule & $p_{\text{screen}}$ & $\text{FWER}_{\text{mod}}$ 
     & $\text{FDR}_{\text{mod}}$ 
@@
     \hline
-    Selected & Simple & .290 & .000 & .002 & .039 \\ 
+    Selected & Basic & .290 & .000 & .002 & .039 \\ 
     Selected & Forward & .559 & .027 & .020 & .066 \\ 
@@
     \hline
-    Saturated & Simple & .000 & .000 & .000 & .028 \\ 
+    Saturated & Basic & .000 & .000 & .000 & .028 \\ 
     Saturated & Forward & .014 & .000 & .000 & .030 \\ 
@@
     \hline
-    Knockoffs & & .000 &  &  & .231 \\ 
+    Knockoffs & & .000 & --- & --- & .231 \\ 
     \hline

### 2015-05-05 code/python/compare_methods.py
rewrite simulation script

@@ -1,237 +1,112 @@
-import numpy as np, os
-import rpy2.robjects as rpy
-from rpy2.robjects import numpy2ri
-rpy.conversion.py2ri = numpy2ri.numpy2ri
+import os
+import numpy as np, pandas as pd
+from itertools import product
+from selection.algorithms.lasso import instance
+
+from rules import simple_stop, forward_stop, strong_stop
+from compute_pvalues import compute_pvalues, completion_index
+
 from matplotlib import mlab
 
-from selection.algorithms.tests.test_forward_step import test_full_pvals
+def summary(variables, pvalues, active, rule, alpha):
+    """
+    Compute R, V_var, V_model, screening
+    for a sequence of pvalues.
+    """
 
-numpy2ri.activate()
-rpy.r("""
-strongstop <- function(p.values,alpha) {
-   d <- length(p.values)
-   lhs <- exp(rev(cumsum(rev(log(p.values)/(1:d))))) # LHS from G'Sell et al.
-   rhs <- alpha * (1:d) / d # RHS from G'Sell et al.
-   return(max(c(0,which(lhs <= rhs))))
-}
+    completion_idx = completion_index(variables, active)
+    R = rule(pvalues, alpha)
+    V_var = R - len(set(active).intersection(variables[:R]))
+    V_model = max(R - completion_idx, 0)
+    screen = R > completion_idx
+    return R, V_var, V_model, screen
 
-forwardstop <- function(p, alpha) {
-   m <- length(p)
-   sums <- -(1/(1:m))*cumsum(log(1-p))
-   return(max(c(0, which(sums < alpha))))
-}
-"""
-)
-numpy2ri.deactivate()
+def simulate(n=100, p=40, rho=0.3, snr=5,
+             do_knockoff=False,
+             full_results={},
+             alpha=0.05):
 
-def simple_stop(pvalues, alpha):
-    if not np.all(pvalues <= alpha):
-        return np.min(np.nonzero(pvalues > alpha)[0])
-    else:
-        return pvalues.shape[0]
-
-def strong_stop(pvalues, alpha):
-    numpy2ri.activate()
-    rpy.r.assign('P', pvalues)
-    rpy.r.assign('alpha', alpha)
-    rpy.r('IDX = strongstop(P, alpha)')
-    value = int(np.array(rpy.r('IDX'))[0])
-    numpy2ri.deactivate()
-    return value
[…]

### 2015-05-05 code/python/rules.py

@@ -125,3 +125,4 @@ def forward_stop(pvalues, alpha):
         return max(np.nonzero(sums < alpha)[0])+1
     return 0
 
+

### 2015-05-05 code/python/data_carving_batch.py
data carving scripts initial commit

@@ -1,0 +1,84 @@
+import numpy as np, csv, hashlib, os.path
+import uuid
+from selection.algorithms.tests.test_forward_step import test_data_carving_IC
+from data_carving import split_vals, vals, df, dname
+
+# how many points do we want for each fraction
+
+min_sample_size = 100000
+
+vals = vals + [('split_frac', 0.9),
+               ('ndraw', 6000),
+               ('burnin', 5000)]
+
+dtype = np.dtype([('n', np.int), 
+                  ('p', np.int), 
+                  ('s', np.int), 
+                  ('rho', np.float), 
+                  ('snr', np.float), 
+                  ('split_frac', np.float), 
+                  ('ndraw', np.int), 
+                  ('burnin', np.int), 
+                  ('method', 'S5'), 
+                  ('null', np.bool), 
+                  ('pval', np.float),
+                  ('uuid', 'S40')])
+
+for i in range(5000):
+    for split_frac in split_vals[::-1]:
+        opts = dict(vals)
+        opts['split_frac'] = split_frac
+        identifier = str(uuid.uuid1())
+        fname = '%s/results_split_%0.2f.npy' % (dname, split_frac)
+        opts['df'] = np.inf # degrees of freedom for noise
+
+        test = not os.path.exists(fname)
+        if not test:
+            prev_results = np.load(fname)
+            if split_frac not in [0.3, 0.4, 0.5]:
+                test = prev_results.shape[0] < min_sample_size
+            elif split_frac in [0.3, 0.4]:
+                test = prev_results.shape[0] < 5000
+            else:
+                test = prev_results.shape[0] < 10000
+
+        if test:
+            (null_carve, 
+             null_split, 
+             alt_carve,
+             alt_split,
+             counter) = test_data_carving_IC(**opts)[:-2]
+
+            params = [v for _, v in vals]
+            results = []
+            if os.path.exists('%s/screening.npy' % dname):
+                prev_results = np.load('%s/screening.npy' % dname)
+                screening = np.empty(prev_results.shape[0]+1, 
+                                      prev_results.dtype)
+                screening[:-1] = prev_results
+                screening[-1] = (split_frac, counter, identifier)
+                np.save('%s/screening.npy' % dname, screening)
+            else:
+                dtype_screen = np.dtype([('split', np.float),
+                                         ('counter', np.float),
+                                         ('uuid', 'S40')])
+                screening = np.array([(split_frac, counter, identifier)],
+                                     dtype_screen)
[…]

### 2015-05-05 code/python/fwdstep_rob.py
script for Rob to compute pvalues

@@ -1,0 +1,59 @@
+# These lines are specific to miller so that the python used is my local one
+#!/home/stow/swtree/python-2.7.3/bin/python2.7
+
+# Usage example:
+# ./fwdstep_rob.py --X=X.csv --Y=Y.csv --active=active.csv --outfile=results.csv --sigma=5.
+
+import sys, os
+sys.path.append('/home/jtaylo/.local/lib/python2.7/site-packages')
+
+# add the current directory to the path so we can find the 
+# sequential-selection functions
+
+sys.path.append(os.path.abspath(os.path.dirname("__file__")))
+
+# The rest of the script should run as long as selection is installed
+# and the selective-selection python directory is on the path
+
+from argparse import ArgumentParser
+
+import numpy as np, pandas as pd
+from compare_methods import run
+
+def main():
+    parser = ArgumentParser(
+        description= '''
+        Run forward step a full min(n, p) steps, computing
+        selected, saturated and nominal pvalues.
+
+        ''')
+    parser.add_argument('--X',
+                        help='Design matrix (CSV filename).')
+    parser.add_argument('--Y',
+                        help='Response (CSV filename).')
+    parser.add_argument('--sigma',
+                        help='Noise standard deviation.', type=float)
+    parser.add_argument('--active',
+                        help='Active set (CSV filename). Necessary ' + 
+                        'to compute completion index. ASSUMES 1-BASED INDEXING!')    
+    parser.add_argument('--outfile',
+                        help='Where to store output.')
+
+    args = parser.parse_args()
+    X = np.loadtxt(args.X, delimiter=',')
+    Y = np.loadtxt(args.Y, delimiter=',')
+    # ASSUMING 1-BASED INDEXING!!!!
+    active = np.loadtxt(args.active, delimiter=',').astype(np.int) - 1
+    
+    if args.sigma is None:
+        raise ValueError('sigma is needed to compute saturated p-values!')
+
+    full_results = {}
+    run(Y, X, args.sigma, active,
+        full_results=full_results)
+    pd.DataFrame(full_results).to_csv(args.outfile, index=False)
+        
+if __name__ == "__main__":
+    main()
+
+                

### 2015-05-05 code/python/compare_methods.py
Return length of selected as completion_index

@@ -26,21 +26,21 @@ def simulate(n=100, p=40, rho=0.3, snr=5,
              full_results={},
              alpha=0.05):
 
-    X, y, beta, active, sigma = instance(n=n,
-                                         p=p,
-                                         rho=rho,
-                                         snr=snr)
+    X, y, _, active, sigma = instance(n=n,
+                                      p=p,
+                                      rho=rho,
+                                      snr=snr)
     full_results.setdefault('n', []).append(n)
     full_results.setdefault('p', []).append(p)
     full_results.setdefault('rho', []).append(rho)
     full_results.setdefault('s', []).append(len(active))
     full_results.setdefault('snr', []).append(snr)
 
-    return run(y, X, beta, sigma, active, 
+    return run(y, X, sigma, active, 
                do_knockoff=do_knockoff,
                full_results=full_results)
 
-def run(y, X, beta, sigma, active,
+def run(y, X, sigma, active,
         full_results={},
         do_knockoff=False,
         alpha=0.05):
@@ -137,8 +137,8 @@ def batch(outbase, nsim, **simulate_args):
         df = pd.DataFrame(full_results)
         df.to_csv(outbase + '.csv', index=False)
 
-if __name__ == "__main__":
-    batch('test', 20, p=10)
+#if __name__ == "__main__":
+    # batch('test', 20, p=10)
     #D = {};
      # batch('test', nsim=1000, p=40, snr=5)
     #simulate(p=10, full_results=D, do_knockoff=True) # batch('test', nsim=1000, p=40, snr=5)

### 2015-05-05 code/python/compute_pvalues.py

@@ -92,6 +92,7 @@ def completion_index(selected, active_set):
     for i in range(len(selected)):
         if active_set.issubset(selected[:i]):
             return i-1
+    return len(selected)-1
 
 if __name__ == "__main__":
     from selection.algorithms.lasso import instance

### 2015-05-05 code/python/fwdstep_rob.py
edits to pythonpath, usage example

@@ -1,8 +1,9 @@
+#!/software/free/statdept/bin/python
+
 # These lines are specific to miller so that the python used is my local one
-#!/home/stow/swtree/python-2.7.3/bin/python2.7
 
 # Usage example:
-# ./fwdstep_rob.py --X=X.csv --Y=Y.csv --active=active.csv --outfile=results.csv --sigma=5.
+# ./fwdstep_rob.py --X=X.csv --Y=Y.csv --active=active.csv --outfile=results.csv --sigma=5. --outfile=results.csv
 
 import sys, os
 sys.path.append('/home/jtaylo/.local/lib/python2.7/site-packages')
@@ -49,6 +50,7 @@ def main():
         raise ValueError('sigma is needed to compute saturated p-values!')
 
     full_results = {}
+    print 'here'
     run(Y, X, args.sigma, active,
         full_results=full_results)
     pd.DataFrame(full_results).to_csv(args.outfile, index=False)

### 2015-05-05 code/python/fwdstep_rob.py
printing help if problems parsing args

@@ -29,23 +29,27 @@ def main():
 
         ''')
     parser.add_argument('--X',
-                        help='Design matrix (CSV filename).')
+                        help='Design matrix (CSV filename). REQUIRED')
     parser.add_argument('--Y',
-                        help='Response (CSV filename).')
+                        help='Response (CSV filename). REQUIRED')
     parser.add_argument('--sigma',
-                        help='Noise standard deviation.', type=float)
+                        help='Noise standard deviation. REQUIRED', type=float)
     parser.add_argument('--active',
                         help='Active set (CSV filename). Necessary ' + 
-                        'to compute completion index. ASSUMES 1-BASED INDEXING!')    
+                        'to compute completion index. REQUIRED. ASSUMES 1-BASED INDEXING!')    
     parser.add_argument('--outfile',
                         help='Where to store output.')
 
-    args = parser.parse_args()
-    X = np.loadtxt(args.X, delimiter=',')
-    Y = np.loadtxt(args.Y, delimiter=',')
-    # ASSUMING 1-BASED INDEXING!!!!
-    active = np.loadtxt(args.active, delimiter=',').astype(np.int) - 1
-    
+    try: 			
+        args = parser.parse_args()
+        X = np.loadtxt(args.X, delimiter=',')
+        Y = np.loadtxt(args.Y, delimiter=',')
+        # ASSUMING 1-BASED INDEXING!!!!
+        active = np.loadtxt(args.active, delimiter=',').astype(np.int) - 1
+    except:
+        parser.print_help()
+        return 
+
     if args.sigma is None:
         raise ValueError('sigma is needed to compute saturated p-values!')
 

### 2015-05-05 code/python/sqrtlasso_rob.py
script for sqrtlasso (doesn't necessarily belong in this repo i suppse)

@@ -1,0 +1,63 @@
+#!/software/free/statdept/bin/python
+
+import sys
+sys.path.append('/home/jtaylo/.local/lib/python2.7/site-packages')
+
+from argparse import ArgumentParser
+
+import numpy as np, pandas as pd
+from selection.algorithms.sqrt_lasso import sqrt_lasso, choose_lambda
+
+def main():
+    parser = ArgumentParser(
+        description= '''
+        Run LASSO at fixed lambda and save p-values and intervals.
+
+        Objective is
+
+        \|Y-X\beta\|_2 + \lambda \|\beta\|_1
+        
+        ''')
+    parser.add_argument('--X',
+                        help='Design matrix (CSV).')
+    parser.add_argument('--Y',
+                        help='Response.')
+    parser.add_argument('--lam',
+                        help='Multiple of theoretical lambda for sqrt-LASSO.', default=0.9, type=float)
+    parser.add_argument('--outfile',
+                        help='Where to results.')
+
+    args = parser.parse_args()
+    X = np.loadtxt(args.X, delimiter=',')
+    Y = np.loadtxt(args.Y, delimiter=',')
+    
+    L = sqrt_lasso(Y, X, args.lam * choose_lambda(X))
+    L.fit(min_its=150, tol=1.e-10)
+    
+    intervals = []
+    pvalues = []
+    if L.active is not None and args.outfile:
+        intervals = np.array(L.active_gaussian_intervals,
+                             np.dtype([('variable', np.int),
+                                       ('lower', np.float),
+                                       ('upper', np.float)]))
+        pvals = np.array(L.active_gaussian_pval,
+                             np.dtype([('variable', np.int),
+                                       ('pvalue', np.float)]))
+        if np.any(intervals['variable'] != pvals['variable']):
+            raise ValueError('variable lists are different!')
+
+        df = pd.DataFrame({'variable':pvals['variable'],
+                           'pvalue':pvals['pvalue'],
+                           'lower':intervals['lower'],
+                           'upper':intervals['upper']})
+        df.reindex(columns=['variable', 'pvalue', 'lower', 'upper'])
+        df.to_csv(args.outfile, index=False)
+    else:
+        if not args.outfile:
+            sys.stderr.write('Active set was empty!\n')
+        
+if __name__ == "__main__":
+    main()
+
+                

### 2015-05-05 code/python/sqrtlasso_rob.py
sqrt lasso now has array for intervals, pvalues

@@ -37,13 +37,9 @@ def main():
     intervals = []
     pvalues = []
     if L.active is not None and args.outfile:
-        intervals = np.array(L.active_gaussian_intervals,
-                             np.dtype([('variable', np.int),
-                                       ('lower', np.float),
-                                       ('upper', np.float)]))
-        pvals = np.array(L.active_gaussian_pval,
-                             np.dtype([('variable', np.int),
-                                       ('pvalue', np.float)]))
+        intervals = L.active_gaussian_intervals
+        pvals = L.active_gaussian_pval
+
         if np.any(intervals['variable'] != pvals['variable']):
             raise ValueError('variable lists are different!')
 
@@ -51,7 +47,7 @@ def main():
                            'pvalue':pvals['pvalue'],
                            'lower':intervals['lower'],
                            'upper':intervals['upper']})
-        df.reindex(columns=['variable', 'pvalue', 'lower', 'upper'])
+        df = df.reindex(columns=['variable', 'pvalue', 'lower', 'upper'])
         df.to_csv(args.outfile, index=False)
     else:
         if not args.outfile:

### 2015-05-05 code/python/sqrtlasso_rob.py
usage message

@@ -3,6 +3,9 @@
 import sys
 sys.path.append('/home/jtaylo/.local/lib/python2.7/site-packages')
 
+# Usage example:
+# ./sqrtlasso_rob.py --X=X.csv --Y=Y.csv --lam=0.9 --outfile=results.csv
+
 from argparse import ArgumentParser
 
 import numpy as np, pandas as pd

### 2015-05-06 paper.tex
defined "generalized selected-model p-values"

@@
     \multicolumn{7}{c}{Saturated-Model $p$-Values 
-      (\% of $10^6$ Simulations)}\\
+      (\% of $10^6$ Simulations)}\\[7pt]
     \hline
+    \multicolumn{7}{c}{}\\[-1.5ex]
     \multicolumn{7}{c}{$p_2(Y)$}\\[5pt]
@@
 
-\subsection{Sequential Sufficiency Principles}\label{sec:sufficiency}
+\subsection{Conditions for Independent $p$-Values}\label{sec:sufficiency}
 
-\subsubsection{$p$-Values and the Sufficiency Principle}\label{sec:pvalSP}
+To obtain a sequence of $p$-values, we must specify what model selection algorithm $M_{[d]}$ to use, as well as how to construct $p$-values at each step.
+In this section we consider what conditions are sufficient to guarantee that the resulting $p$-value sequence is independent.
+Sections~\ref{sec:pvalSP}--\ref{sec:modelSPSP} discuss sufficient conditions on the $p$-values and the selection algorithm under which single-step $p$-values are automatically independent.
+Section~\ref{sec:selectionVariables} discusses how we can create independent $p$-values by conditioning on finer selection variables at each step.
 
-A $p$-value for testing model $M_{k-1}$ against $M_k$ satisfies the (conditional) {\em sufficiency principle} if, on the event $\{M_{k-1}=m_{k-1}, M_k=m_k\}$, it depends only on sufficient statistics of $M_k$.
+Essentially, we will want to partition the information in $Y$ according to the filtration:
+\begin{align}\nonumber
+  \sF(M_0,T_0) &\underlabel_{\text{selection } 1} 
+  \sF(M_{[1]},T_0) \underlabel_{\text{inference } 1}
+  \sF(M_{[1]},T_1) \quad \sub \;\;\cdots\\[8pt]
+  \label{eq:infoPartition}
+  \cdots\;\; \sub \quad&
+  \sF(M_{[d-1]},T_{d-1}) \underlabel_{\text{selection } d}
+  \sF(M_{[d]},T_{d-1}) 
+  \underlabel_{\text{inference } d}
+  \sF(M_{[d]}, T_d)
+\end{align}
 
-Based on the law
+\subsubsection{Generalized Selected-Model $p$-Values}\label{sec:pvalSP}
+
+For a candidate pair $(m_{k-1},m_k)$ of models to test at step $k$, define the selection event $A = \{M_{k-1},M_k\}$.
+We say $p(Y)$ is a {\em generalized selected-model $p$-value} for testing $m_{k-1}$ against $m_k$ at step $k$ if it is ``based on'' the law
 \[
-\L_{m_{k-1}}\left(T_k \;\;\mid\;\; T_{k-1},\; M_{k-1}=m_{k-1},\; M_k=m_k\right)
+\L_{m_{k-1}}\left(T_k \;\;\mid\;\; T_{k-1},\; M_{k-1}=m_{k-1},\;M_k=m_k\right)
 \]
-
-If $T(Y)$ is a complete sufficient statistic, then any exact $p$-value must have the latter property.
+for some minimal sufficient statistics $T_{k-1}(Y)$ and $T_k(Y)$ of $m_{k-1}$ and $m_k$, respectively.
+More precisely, $p(Y)$ must be uniform or conservative conditional on $T_{k-1}(Y)$, and it must be a function of $T_k(Y)$ on $A$.
 
@@
 
+\begin{comment}
 \WFcomment{There is a filtration interpretation when you have the appropriate sufficiency properties.} Let $\sF_{k,\ell}$ denote the $\sigma$-algebra generated by $M_{[k]}$ and $p_{[\ell]}$.
@@
 \end{align*}
+\end{comment}
 

### 2015-05-07 paper.tex
just commented out the figures
added figures for paper, uncommented figures

@@
 In this section we consider what conditions are sufficient to guarantee that the resulting $p$-value sequence is independent.
+
+\begin{comment}
 Sections~\ref{sec:pvalSP}--\ref{sec:modelSPSP} discuss sufficient conditions on the $p$-values and the selection algorithm under which single-step $p$-values are automatically independent.
@@
 \end{align}
+\end{comment}
 
-\subsubsection{Generalized Selected-Model $p$-Values}\label{sec:pvalSP}
+\subsubsection{Independent $p$-Values and the Sufficiency Principle}\label{sec:pvalSP}
 
-For a candidate pair $(m_{k-1},m_k)$ of models to test at step $k$, define the selection event $A = \{M_{k-1},M_k\}$.
-We say $p(Y)$ is a {\em generalized selected-model $p$-value} for testing $m_{k-1}$ against $m_k$ at step $k$ if it is ``based on'' the law
+Assume that every candidate model $m$ has a set of minimal sufficient statistics $T(y; \;m)$.
+For linear regression, the sufficient statistics for $M(E)$ are $X_E'Y$ if $\sigma^2$ is known and  $(X_E'Y,\|Y\|)$ otherwise.
+At step $k$, let $T_k(Y) = T(Y; \;M_k(Y))$ denote the sufficient statistics for the model actually chosen --- for example, $X_{E_k(Y)}'Y$.
+
+For a candidate pair $(m_{k-1},m_k)$ of models to test at step $k$, define the selection event $A = \{M_{k-1} = m_{k-1}, M_k=m_k\}$.
+The $p$-value for this pair, $p_{k,m_{k-1},m_k}$ satisfies the sufficiency principle if
+\[
+\P_F\left(p_{k,m_{k-1},m_k}(Y) \mid M_{k-1}=m_{k-1}, M_k=m_k\right)
+\]
+This is what we mean when we say $p_{k,m_{k-1},m}$ is ``based on'' the law
 \[

### 2015-05-07 code/python/fwdstep_rob.py
output for Rob now has 1-based indices

@@ -54,9 +54,16 @@ def main():
         raise ValueError('sigma is needed to compute saturated p-values!')
 
     full_results = {}
-    print 'here'
+
     run(Y, X, args.sigma, active,
         full_results=full_results)
+
+    # change output variables to 1-based indices
+
+    for k in full_results.keys():
+        if 'variable_selected' in k:
+            full_results[k][0] = full_results[k][0] + 1
+
     pd.DataFrame(full_results).to_csv(args.outfile, index=False)
         
 if __name__ == "__main__":

### 2015-05-07 code/python/compare_methods.py
how long does a parametric bootstrap take

@@ -46,7 +46,7 @@ def run(y, X, sigma, active,
         alpha=0.05):
 
     n, p = X.shape
-    results = compute_pvalues(y, X, sigma)
+    results, FS = compute_pvalues(y, X, sigma)
     completion_idx = completion_index(results['variable_selected'], active)
     full_results.setdefault('completion_idx', []).append(completion_idx)
 
@@ -106,7 +106,7 @@ def run(y, X, sigma, active,
         full_results.setdefault('%s_%s_V_model' % (pval_name, rule_name), []).append(V_model)
         full_results.setdefault('%s_%s_screen' % (pval_name, rule_name), []).append(screen)
 
-    return full_results
+    return full_results, FS
 
 
 def batch(fbase, nsim=100, n=100, p=40, rho=0.3, snr=4):

### 2015-05-07 code/python/compute_pvalues.py

@@ -3,7 +3,7 @@ import pandas as pd
 from selection.algorithms.forward_step import forward_stepwise
 from scipy.stats import norm as ndist
 
-def compute_pvalues(y, X, sigma=1.):
+def compute_pvalues(y, X, sigma=1., maxstep=np.inf):
     """
     Parameters
     ----------
@@ -36,7 +36,7 @@ def compute_pvalues(y, X, sigma=1.):
     FS = forward_stepwise(X, y, covariance=sigma**2 * np.identity(n))
     results = []
 
-    for i in range(min(n, p)):
+    for i in range(min([n, p, maxstep])):
         FS.next()
         var_select, pval_select = FS.model_pivots(i+1, alternative='twosided',
                                                   which_var=[FS.variables[-1]],
@@ -58,7 +58,7 @@ def compute_pvalues(y, X, sigma=1.):
     return pd.DataFrame({'variable_selected': results[0].astype(np.int),
                          'selected_pvalue': results[1],
                          'saturated_pvalue': results[2],
-                         'nominal_pvalue': results[3]})
+                         'nominal_pvalue': results[3]}), FS
 
 def completion_index(selected, active_set):
     """
@@ -96,6 +96,6 @@ def completion_index(selected, active_set):
 
 if __name__ == "__main__":
     from selection.algorithms.lasso import instance
-    X, y, beta, active, sigma = instance(p=10, snr=4, rho=0.3)
-    R = compute_pvalues(y, X, sigma=sigma)
+    X, y, beta, active, sigma = instance(n=100, p=40, snr=5, rho=0.3)
+    R, FS = compute_pvalues(y, X, sigma=sigma, maxstep=20)
     print completion_index(R['variable_selected'], active)

### 2015-05-08 code/python/compare_methods.py
added maxT as a test -- as expected a little more powerful than conditioning on identity of maximizer

@@ -87,7 +87,8 @@ def run(y, X, sigma, active,
 
     for pval, rule_ in product(['selected_pvalue',
                                 'saturated_pvalue',
-                                'nominal_pvalue'],
+                                'nominal_pvalue',
+                                'maxT_pvalue'],
                                zip([simple_stop, 
                                     strong_stop,
                                     forward_stop],
@@ -108,39 +109,28 @@ def run(y, X, sigma, active,
 
     return full_results, FS
 
-
-def batch(fbase, nsim=100, n=100, p=40, rho=0.3, snr=4):
-    value = []
-    for _ in range(nsim):
-        try:
-            value.append(run_one(n=n, p=p, rho=rho, snr=snr, alpha=0.05))
-        except:
-            pass
-        if not os.path.exists(fbase + '.npy'):
-            V = np.hstack(value)
-            np.save(fbase + '.npy', V)
-            mlab.rec2csv(V, fbase + '.csv')
-        else:
-            V = np.load(fbase + '.npy')
-            V = np.hstack([V] + value[-1:])
-            np.save(fbase + '.npy', V)
-            mlab.rec2csv(V, fbase + '.csv')
-            
-        print np.mean(V['forward_selected_screen']), np.mean(V['forward_saturated_screen']), V.shape
-
 def batch(outbase, nsim, **simulate_args):
 
     full_results = {}
     simulate_args['full_results'] = full_results
-    for _ in range(nsim):
-        simulate(**simulate_args)
-        df = pd.DataFrame(full_results)
-        df.to_csv(outbase + '.csv', index=False)
+    for i in range(nsim):
+        try:
+            print 'run %d' % (i+1)
+            simulate(**simulate_args)
+            df = pd.DataFrame(full_results)
+            df.to_csv(outbase + '.csv', index=False)
+        except:
+            pass
 
-#if __name__ == "__main__":
-    # batch('test', 20, p=10)
-    #D = {};
-     # batch('test', nsim=1000, p=40, snr=5)
-    #simulate(p=10, full_results=D, do_knockoff=True) # batch('test', nsim=1000, p=40, snr=5)
-    #simulate(p=10, full_results=D, do_knockoff=True) # batch('test', nsim=1000, p=40, snr=5)
-    #D
+if __name__ == "__main__":
+    from argparse import ArgumentParser
+    parser = ArgumentParser(
+        description= '''
+Run a batch of simulations.
+        ''')
+    parser.add_argument('--nsim',
[…]

### 2015-05-08 code/python/parametric_bootstrap.R
R script for accept reject

@@ -1,0 +1,63 @@
+set.seed(0)
+X = read.table("X_100_200.csv", sep=',')
+X = as.matrix(X) 
+Y = as.matrix(read.table("y_100.csv", sep=','))
+variables = as.matrix(read.table("variables.csv", sep=','))
+
+print("step 7")
+print(variables[1:7])
+X7 = X[,variables[1:6]]
+
+print("step 8")
+print(variables[1:8])
+X8 = X[,variables[1:7]]
+
+P7 = X7 %*% solve(t(X7) %*% X7) %*% t(X7)
+P8 = X8 %*% solve(t(X8) %*% X8) %*% t(X8)
+
+A7 = as.matrix(read.table("A_step7.csv", sep=','))
+A8 = as.matrix(read.table("A_step8.csv", sep=','))
+
+sigma = as.numeric(read.table("sigma.csv"))
+
+n = nrow(X)
+p = ncol(X)
+R7 = diag(rep(1,n)) - P7
+R8 = diag(rep(1,n)) - P8
+
+P7Y = P7 %*% Y
+P8Y = P8 %*% Y
+
+# here is the null distribution
+
+null_sample = function(R, PY, sigma) {
+     n = nrow(R)
+     Z = PY + R %*% rnorm(n) * sigma
+     return(Z)
+}
+
+# how many tries until we accept
+
+wait_until_accept = function(A, R, PY, sigma) {
+    count = 0
+    while(TRUE) {
+        count = count + 1
+        Z = null_sample(R, PY, sigma)
+	print(max(A %*% Z))
+        if(max(A %*% Z) < 0) {
+            return(count)
+        }
+   }
+}
+
+count7_sample = c()
+for (i in 1:100) {
+    count7_sample = c(count7_sample, wait_until_accept(A7, R7, P7Y, sigma))
+    print(1. / mean(count7_sample))
+}
+
+count8_sample = c()
+for (i in 1:100) {
+    count8_sample = c(count8_sample, wait_until_accept(A8, R8, P8Y, sigma))
+    print(1. / mean(count8_sample))
+}

### 2015-05-08 code/python/parametric_bootstrap.R
R example to reproduce sampling difficulty

@@ -1,63 +1,0 @@
-set.seed(0)
-X = read.table("X_100_200.csv", sep=',')
-X = as.matrix(X) 
-Y = as.matrix(read.table("y_100.csv", sep=','))
-variables = as.matrix(read.table("variables.csv", sep=','))
-
-print("step 7")
-print(variables[1:7])
-X7 = X[,variables[1:6]]
-
-print("step 8")
-print(variables[1:8])
-X8 = X[,variables[1:7]]
-
-P7 = X7 %*% solve(t(X7) %*% X7) %*% t(X7)
-P8 = X8 %*% solve(t(X8) %*% X8) %*% t(X8)
-
-A7 = as.matrix(read.table("A_step7.csv", sep=','))
-A8 = as.matrix(read.table("A_step8.csv", sep=','))
-
-sigma = as.numeric(read.table("sigma.csv"))
-
-n = nrow(X)
-p = ncol(X)
-R7 = diag(rep(1,n)) - P7
-R8 = diag(rep(1,n)) - P8
-
-P7Y = P7 %*% Y
-P8Y = P8 %*% Y
-
-# here is the null distribution
-
-null_sample = function(R, PY, sigma) {
-     n = nrow(R)
-     Z = PY + R %*% rnorm(n) * sigma
-     return(Z)
-}
-
-# how many tries until we accept
-
-wait_until_accept = function(A, R, PY, sigma) {
-    count = 0
-    while(TRUE) {
-        count = count + 1
-        Z = null_sample(R, PY, sigma)
-	print(max(A %*% Z))
-        if(max(A %*% Z) < 0) {
-            return(count)
-        }
-   }
-}
-
-count7_sample = c()
-for (i in 1:100) {
-    count7_sample = c(count7_sample, wait_until_accept(A7, R7, P7Y, sigma))
-    print(1. / mean(count7_sample))
-}
-
-count8_sample = c()
-for (i in 1:100) {
-    count8_sample = c(count8_sample, wait_until_accept(A8, R8, P8Y, sigma))
-    print(1. / mean(count8_sample))
-}

### 2015-05-08 paper.tex
commented out old sequential section
introduced separation and the sufficient filtration

--- 2015-05-06 paper.tex
@@
 \begin{equation}
-  M_k:\; (n-1) S \sim W_d\left(\lambda_0 I_d + \sum_{\ell=1}^k     \lambda_\ell u_i u_i', \;\;\; n-1\right).
+  M_k:\; (n-1) S \sim W_d\left(\lambda_0 I_d + \sum_{i=1}^k     \lambda_i u_i u_i', \;\;\; n-1\right).
 \end{equation}
@@
 The main complication arises from the fact that the null and alternative hypotheses are random.
-To simplify matters, we first consider deriving $p$-values for a fixed candidate pair $m_{k-1} \sub m_k$.
+To simplify matters, we first consider deriving $p$-values for a fixed candidate pair $m_{k-1} \sub m_{k-1}$.
 
-The random variable  $p_{k,m_{k-1},m_k}(Y)$ is a valid {\em selective $p$-value} for the fixed candidate pair $(m_{k-1},m_k)$ if it is stochastically larger than uniform under the (fixed) null $m_{k-1}$, given that the pair $(m_{k-1},m_k)$ is selected.
+The random variable  $p_{k,M}(Y)$ is a valid {\em selective $p$-value} for the fixed candidate pair $(m_{k-1},m_k)$ if it is stochastically larger than uniform under the (fixed) null $m_{k-1}$, given that the pair $(m_{k-1},m_k)$ is selected.
 That is,
 \[
-\P_F\left(p_{k,m_{k-1},m_k}(Y) \leq \alpha \mid M_{k-1}(Y) = m_{k-1}\right) 
+\P_F\left(p_{k,M}(Y) \leq \alpha \mid M_{k-1}(Y) = m_{k-1}\right) 
 \leq \alpha, \quad \forall F\in m_{k-1}, \alpha \in [0,1].
@@
   \hk_{S}(Y) = \max\left\{k \in \{1,\ldots,d\} :\;
-    \exp\left(\sum_{\ell=k}^d \frac{\log p_\ell}{\ell}\right) 
+    \exp\left(\sum_{i=k}^d \frac{\log p_i}{i}\right) 
     \leq \frac{\alpha k}{d}\right\}
@@
 \P(p_{k_0+1} \leq \alpha_{k_0+1}, \ldots, p_d \leq \alpha_d
-\mid p_1, \ldots, p_{k_0}) \leq \prod_{\ell=k_0+1}^d \alpha_\ell
+\mid p_1, \ldots, p_{k_0}) \leq \prod_{i=k_0+1}^d \alpha_i
 \end{equation}
@@
   \P(p_{k+1} \leq \alpha_{k+1}, \ldots, p_d \leq \alpha_d
-  \mid p_1, \ldots, p_{k_0}, \; k_0 = k) \leq \prod_{\ell=k+1}^d \alpha_\ell
+  \mid p_1, \ldots, p_{k_0}, \; k_0 = k) \leq \prod_{i=k+1}^d \alpha_i
 \end{equation}
@@
   \hk_{F}(Y) = \max\left\{k \in \{1,\ldots,d\} :\;
-    -\frac{1}{k}\sum_{\ell=1}^k \log(1-p_\ell) \leq \alpha\right\}
+    -\frac{1}{k}\sum_{i=1}^k \log(1-p_i) \leq \alpha\right\}
 \]
@@
 \[
-\widehat{\text{FDR}}_k = -\frac{1}{k}\sum_{\ell=1}^k \log(1-p_\ell)
+\widehat{\text{FDR}}_k = -\frac{1}{k}\sum_{i=1}^k \log(1-p_i)
 \]
@@
 
-\subsection{Saturated- and Selected-Model $p$-Values in Regression}
+\subsection{$p$-Values and Independence}
 
-If we wish to use strong stop and forward stop, we need the $p$-values to satisfy~\eqref{eq:indepCond_random_k0}, but there is no guarantee that single-step selective $p$-values are independent of each other.
+Recall that $p_k$ is a valid selective $p$-value for $M_{k-1}\sub M_\infty$ if, for all $F\in M_\infty$,
+\[
+\P_F(p_k \leq \alpha \mid M_{k-1},\; F\in M_{k-1}) \leqAS \alpha.
+\]
+Valid single-step selective $p$-values do not generally result in $p$-values that are independent from one step to the next.
 
-In fact, saturated-model $p$-values typically are not independent.
-Continuing with Example~\ref{ex:bivariate}, Table~\ref{tab:bv_twoWayTable} shows a two-way contingency table for the saturated-model $p$-values $(p_1(Y), p_2(Y))$, binned into cells of height and width 0.2.
-Because the global null is true, both $p$-values are uniform by construction, but the $p$-values are strongly dependent, with correlation $-48\%$.
+For example, saturated-model $p$-values are generically non-independent.
+Continuing with Example~\ref{ex:bivariate}, Table~\ref{tab:bv_twoWayTable} shows a two-way contingency table for the saturated-model $p$-values $(p_1(Y), p_2(Y))$, binned into cells of height and width 0.2, simulated under the global null $\mu=0$.
+Because $k_0=0$, both $p$-values are uniform by construction, but the $p$-values are strongly dependent, with correlation $-48\%$.
+By contrast, we will see that the selected-model $p$-values $(p_1,p_2)$ are independent under the global null, a consequence of Proposition~\ref{prop:selectiveModel}.
 
@@
 
-By contrast, the selected-model $p$-values $(p_1,p_2)$ are independent.
-To see why, recall that, on the selection event $A$, $p_1$ is only a function of $Y_1$.
-Moreover, $p_2$ is a selective $p$-value for comparing the null
+First, note that $p_2$ is a selective $p$-value for comparing the null with only variable 1,
 \[
-M_1:\; Y \sim \cN\left(\binom{\mu_1}{0}, I_2\right)
+M_1:\; Y \sim \cN\left(\binom{\mu_1}{0}, I_2\right),
 \]
-against the alternative
+against the alternative with variables 1 and 2:
 \[
-M_2:\; Y \sim \cN\left(\mu, I_2\right), \text{ with } \mu_2 \neq 0.
+M_2:\; Y \sim \cN\left(\mu, I_2\right), \text{ with } 
+\mu_2 \neq 0.
 \]
-Because $\mu_1$ is a nuisance parameter in $M_1$, an exact selected-model test must condition on the corresponding sufficient statistic $Y_1$.
-But that means $p_2$ is uniform given $Y_1$, and therefore also uniform given $p_1(Y_1)$.
+To eliminate the nuisance parameter $\mu_1$, the test conditions on $Y_1$.
+Thus, $p_2$ is uniform given $Y_1$ on $A$.
+Second, note that on $A$, $p_1$ is a function only of $Y_1$.
+Thus, $(p_1,p_2)$ are independent uniform variables given that variable 1 is chosen first.
+By a similar argument, they would also be independent uniforms if $Y_2$ were chosen first; thus, $(p_1, p_2)$ are marginally uniform under the global null.
 
-A generalization of this argument shows that the $p$-values for the saturated-model tests are always independent:
+We now introduce a simple sufficient condition to verify that a given procedure yields independent $p$-values.
+We say that a filtration $\sF_{[d]}$ with $M_{[k]} \in \sF_k$ {\em separates} the $p$-values $p_{[d]}$ if
+\begin{enumerate}
+\item $p_k(Y)$ is conservative given $\sF_{k-1}$ 
+  when $F\in M_{k-1}$, and
+\item $p_k(Y)$ is measurable with respect to $\sF_k$.
+\end{enumerate}
+If we think of $\sF_k$ representing information available at step $k$, then the first condition means that $p_k$ excludes whatever evidence may have accrued against the null by step $k-1$, and the second means that information revealed after step $k$ plays no role in determining $p_k$.
+In that sense, $\sF_k$ forms a ``wall of separation'' between the $p$-values up to $k$ and the ones after $k$.
+As a result, separated null $p$-values are independent.
 
-\begin{proposition}
-  The selected-model equal-tailed or UMPU $p$-values after forward-stepwise linear regression. satisfy~\eqref{eq:indepCond_random_k0}.
+\begin{proposition}[Independence of 
+  Separated $p$-Values]\label{prop:jointConserv}
+
+  Let $p_{[d]}$ be selective $p$-values for $M_{[d]}$, 
+  separated by $\sF_{[d]}$.
+
+  If the $p$-values are exact then $p_{k+1}, \ldots, p_d$ are
+  uniform given $\sF_{k_0}$ on the event $\{k_0=k\}$.
+
+  If they $p$-values are conservative, then for
+  $\alpha_{k+1},\ldots,\alpha_d \in [0,1]$,
+  \begin{equation}\label{eq:jointConserv}
+  \P_F\left(p_{k+1}\leq \alpha_{k+1}, \ldots, p_d \leq \alpha_d\;
+    \mid\; k_0 \leq k, \; \sF_k\right) \leqAS \prod_{i=k+1}^d
+  \alpha_i.
+  \end{equation}
 \end{proposition}
-This is a simple consequence of the much more general results proven in Section~\ref{sec:sufficiency}, but it is instructive to see the proof in the relatively concrete case of linear regression.
+Equation~\ref{eq:jointConserv} is slightly weaker than independence, but it guarantees that both Strong Stop and Forward Stop control their target error rates.
+
 \begin{proof}
-  At step $k$, we are testing the null hypothesis that $\beta_j=0$ in     the model $Y\sim \cN(X_E\beta, I_2)$, for some active set $E\sub [p]$ of size $k$, and some variable $j\in E$.
-Let $A_{j,E}$ denote the selection event for a particular $j$ and $E$, and let $m_{k-1}(j,E)$ denote the null model with active set $X_{E\setminus j}$.
+  We first prove~\eqref{eq:jointConserv}
+  by induction. Define $B_i = \{p_i \leq \alpha_i\}$. 
+  The base case is
+  \[
+  \P_F\left(B_d \mid \sF_{d-1}\right)1_{\{k_0 \leq d-1\}} \leqAS \alpha_d,
+  \]
+  which is true by construction of $p_d$. 
+  For the inductive case, note that the
+  conditional probability on the left-hand side
+  of~\eqref{eq:jointConserv} is
+  \begin{align*}
+    \P_F\left(B_{k+1}, \ldots, B_d
+      \mid \sF_k\right)1_{\{k_0 \leq k\}} 
+    &\eqAS \E_F\bigg[ 1_{B_{k+1}} 
+    \P_F\left(B_{k+2}, \ldots, B_d
+      \mid \sF_{k+1}\right)1_{\{k_0 \leq k+1\}}
+    \mid \sF_k\bigg]1_{\{k_0 \leq k\}}\\
+    &\leqAS \P_F\left[ B_{k+1}
+      \mid \sF_k\right]1_{\{k_0 \leq k+1\}}\prod_{i=k+2}^d \alpha_i
+    \quad\;\;\leqAS\;\; \prod_{i=k+1}^d \alpha_i
+  \end{align*}
 
-On $A_{j, E}$, the selected-model $p$-value $p_k$ is based on the null distribution of the test statistic $X_j'Y$ conditional on the null sufficient statistics $X_{E\setminus j}'Y$ and on $Y \in A_{j,E}$.
+  If the $p_k$ are exact then the above 
+  inequalities become equalities, implying uniformity.
+\end{proof}
 
-Moreover, on $A_{j,E}$, the entire sequence of previous $p$-values $p_{[k-1]}$ depend only on the sufficient statistics $X_{E\setminus j}'Y$ of $m_{k-1}(j,E)$.
-To see why, \WFcomment{complete this proof.}
+\subsection{Sufficient Filtration}
 
-We have shown that, if $m_{k-1}(j,E)$ is true, then $p_k$ is uniform and independent of $p_{[k-1]}$ on $A_{j,E}$ for each $j$ and $E$. That is,
+Assume that each candidate model has a minimal sufficient statistic $T(y; M)$, and write
 \[
-\P_F\left(p_{k} \leq \alpha \mid p_{[k-1]}, Y \in A_{j,E} \right) = \alpha \quad \forall F \in m_{k-1}.
+T_k(Y) = T(Y; M_k(Y)).
 \]
-It follows that 
+For example, in linear regression with known $\sigma^2$, the complete sufficient statistic for model $M(E)$ is $X_E'Y$, and the sufficient statistic at step $k$ is $T_k = X_{E_k}'Y$.
+
+We will concern ourselves mainly with the {\em sufficient filtration} for path $M_{[k]}$, defined as
 \[
-\P_F\left(p_{k} \leq \alpha \mid p_{[k-1]}, F \in M_{k-1} \right) = \alpha \quad \forall F.
+\sF_k = \sF(M_{[k]}, T_k),
 \]
-\end{proof}
-The proof turns on two properties of selected-model $p$-values, and one property of forward stepwise regression.
+The sufficient filtration separates $p_{[d]}$ if $p_k$ is conservative given $\sF_{k-1}$ and $\sF_k$-measurable.\footnote{For readers unfamiliar with $\sigma$-algebra notation, this statement can be read as ``$p_k$ is conservative given $(M_{[k-1]}, T_{k-1})$ and is a function of $(M_{[k]}, T_k)$''}
 
-As for the $p$-values, we used the fact that $p_{k,m_{k-1},m_k}$ is uniform conditional on the sufficient statistics of the null model $m_{k-1}$.
-This is true for selected-model and saturated-model $p$-values alike.
-We also used the fact that $p_{k,m_{k-1},m_k}$ follows the {\em sufficiency principle} --- that is, it depends only on the sufficient statistics of $m_k$.
-This is {\em not} true of the saturated-model $p$-values: to compute them, we must know $\proj_{\eta}^\perp Y$, which depends on statistics that are ancillary for $m_{k-1}$.
+To be a valid selective $p$-value for a single step, $p_k$ only had to condition on $M_{k-1}$.
+Thus, it is a little more stringent to additionally require $p_k$ to condition on the entire subpath $M_{[k-1]}$ as well as $T_{k-1}$.
 
-We also used the fact that, in the forward stepwise algorithm, if $m_k$ is the $k$th model then, on the event $M_k=m_k$, all the models up to step $k$ depend only on sufficient statistics of $m_k$.
-This is a sort of {\em sub-path sufficiency principle} which many other algorithms satisfy, as we will see in Section~\ref{sec:modelSPSP}.
+In most cases of interest, however, the two requirements are equivalent.
+First, we typically construct $p_k$ by conditioning on $T_{k-1}$.
+In fact, $T_{k-1}$ is complete sufficient, then any exact test must condition on $T_{k-1}$ \WFcomment{should we bother proving this?}.
+Second, most path algorithms of interest (including forward stepwise regression, the ever-active LASSO path, and their generalizations) satisfy a {\em subpath sufficiency principle} (henceforth, SSP): that once we know $M_k$ and its sufficient statistics, we can reconstruct the model path up to step $k$.
+Thus, most single-step $p$-values are already conservative conditional on $\sF_{k-1}$.
 
-\subsection{Conditions for Independent $p$-Values}\label{sec:sufficiency}
+The requirement that $p_k$ be $\sF_k$-measurable has more bite.
+For example, it excludes saturated-model tests in linear regression.
+Computing the cutoff for the saturated-model test requires us to know $\proj_{\eta}^\perp Y$, which is not $\sF_k$-measurable.
 
-To obtain a sequence of $p$-values, we must specify what model selection algorithm $M_{[d]}$ to use, as well as how to construct $p$-values at each step.
-In this section we consider what conditions are sufficient to guarantee that the resulting $p$-value sequence is independent.
-Sections~\ref{sec:pvalSP}--\ref{sec:modelSPSP} discuss sufficient conditions on the $p$-values and the selection algorithm under which single-step $p$-values are automatically independent.
-Section~\ref{sec:selectionVariables} discusses how we can create independent $p$-values by conditioning on finer selection variables at each step.
+By contrast, selected-model tests of $M_{k-1}$ against $M_k$ are always $\sF_k$-measurable, because they are based on the law
+\begin{equation}\label{eq:selModel}
+\L\left(X_{j_k}'Y \mid X_{E_k-1}'Y, E_{k-1}, j_k\right),
+\end{equation}
+and all of the random variables appearing in~\eqref{eq:selModel} are $\sF_k$-measurable.
 
-Essentially, we will want to partition the information in $Y$ according to the filtration:
-\begin{align}\nonumber
-  \sF(M_0,T_0) &\underlabel_{\text{selection } 1} 
-  \sF(M_{[1]},T_0) \underlabel_{\text{inference } 1}
-  \sF(M_{[1]},T_1) \quad \sub \;\;\cdots\\[8pt]
-  \label{eq:infoPartition}
-  \cdots\;\; \sub \quad&
-  \sF(M_{[d-1]},T_{d-1}) \underlabel_{\text{selection } d}
-  \sF(M_{[d]},T_{d-1}) 
-  \underlabel_{\text{inference } d}
-  \sF(M_{[d]}, T_d)
-\end{align}
+\begin{comment}
 
-\subsubsection{Generalized Selected-Model $p$-Values}\label{sec:pvalSP}
+\begin{theorem}
+  Assume that $T(Y; M)$ is a complete sufficient statistic, 
+  that $p_k$ are exact selective $p$-values, and that 
+  $M_{[d]}$ satisfies EPIC.
 
-For a candidate pair $(m_{k-1},m_k)$ of models to test at step $k$, define the selection event $A = \{M_{k-1},M_k\}$.
-We say $p(Y)$ is a {\em generalized selected-model $p$-value} for testing $m_{k-1}$ against $m_k$ at step $k$ if it is ``based on'' the law
-\[
-\L_{m_{k-1}}\left(T_k \;\;\mid\;\; T_{k-1},\; M_{k-1}=m_{k-1},\;M_k=m_k\right)
-\]
-for some minimal sufficient statistics $T_{k-1}(Y)$ and $T_k(Y)$ of $m_{k-1}$ and $m_k$, respectively.
-More precisely, $p(Y)$ must be uniform or conservative conditional on $T_{k-1}(Y)$, and it must be a function of $T_k(Y)$ on $A$.
-
-\subsubsection{A Sub-Path Sufficiency Principle for Model Selection Procedures}\label{sec:modelSPSP}
-
-We say that a selection algorithm $M_{[d]}(\cdot)$ satisfies the {\em sub-path sufficiency principle} (henceforth SPSP) if, on the event $M_k(Y)=m_k$, the sequence of models $M_{[k]}$ depends only on sufficient statistics of $m_k$.
-Again, if
+  Then, the sufficient filtration separates $p_{[d]}$ if and only if 
+  $p_k \in \sF_k$.
+\end{theorem}
 
@@
 
-For $y\in A$ and $\ell \leq k$, we have $E_\ell(y) \sub E$. This means that the maximizer in \eqref{eq:forwardDef_start} is attained for some $j\in E$. So, for $\ell \leq k$, on $A$,
+For $y\in A$ and $i \leq k$, we have $E_i(y) \sub E$. This means that the maximizer in \eqref{eq:forwardDef_start} is attained for some $j\in E$. So, for $i \leq k$, on $A$,
 \begin{align}
-  j_\ell &= \argmax_{j\in E} \;\;\sup \left\{L(\theta; Y):\;
-    \theta\in\Theta(E_{\ell-1} \cup \{j\})\right\} \\
+  j_i &= \argmax_{j\in E} \;\;\sup \left\{L(\theta; Y):\;
+    \theta\in\Theta(E_{i-1} \cup \{j\})\right\} \\
   &= \argmax_{j\in E} \;\;\sup \left\{L_T(\theta; T(Y)):\;
-    \theta\in\Theta(E_{\ell-1} \cup \{j\})\right\}. \label{eq:onlyT}
+    \theta\in\Theta(E_{i-1} \cup \{j\})\right\}. \label{eq:onlyT}
 \end{align}
-Replacing $L$ with $L_T$ in~\eqref{eq:onlyT} is justified because $\Theta(E_{\ell - 1} \cup \{j\}) \sub \Theta(E)$ when $E_{\ell-1}\cup E$ and $j\in E$.
+Replacing $L$ with $L_T$ in~\eqref{eq:onlyT} is justified because $\Theta(E_{i - 1} \cup \{j\}) \sub \Theta(E)$ when $E_{i-1}\cup E$ and $j\in E$.
 
@@
 
-\begin{comment}
-\WFcomment{There is a filtration interpretation when you have the appropriate sufficiency properties.} Let $\sF_{k,\ell}$ denote the $\sigma$-algebra generated by $M_{[k]}$ and $p_{[\ell]}$.
-\begin{align*}
-  \sF_{k,\ell} &= \sF(M_{[k]},p_{[\ell]})\\
-  \sF_0 &\underlabel_{\text{selection } 1} \sF_{1,0} \underlabel_{\text{inference } 1}
-  \sF_{1,1} \;\;\sub \cdots \sub\;\;
-  \sF_{d-1,d-1} \underlabel_{\text{selection } d} \sF_{d,d-1}
-  \underlabel_{\text{inference } d} \sF_{d,d}
-\end{align*}
-\end{comment}
-
 \subsection{Independent $p$-Values Via Selection Variables}\label{sec:selectionVariables}
@@
 
-We say that $p_{k,m_{k-1},m_k}$ is a selective $p$-value with respect to selection variable $S(Y)$ if it is uniform under $F\in m_k$ given $S$.
+We say that $p_{k,M}$ is a selective $p$-value with respect to selection variable $S(Y)$ if it is uniform under $F\in m_k$ given $S$.
 
 To achieve joint error control properties among different tests, we may need to condition on a finer variable.
+\end{comment}
 
@@
 \end{document}
+
+\begin{comment}
+Sections~\ref{sec:pvalSP}--\ref{sec:modelSPSP} discuss sufficient conditions on the $p$-values and the selection algorithm under which single-step $p$-values are automatically independent.
+Section~\ref{sec:selectionVariables} discusses how we can create independent $p$-values by conditioning on finer selection variables at each step.
+
+Essentially, we will want to partition the information in $Y$ according to the filtration:
+\begin{align}\nonumber
+  \sF(M_0,T_0) &\underlabel_{\text{selection } 1} 
+  \sF(M_{[1]},T_0) \underlabel_{\text{inference } 1}
+  \sF(M_{[1]},T_1) \quad \sub \;\;\cdots\\[8pt]
+  \label{eq:infoPartition}
+  \cdots\;\; \sub \quad&
+  \sF(M_{[d-1]},T_{d-1}) \underlabel_{\text{selection } d}
+  \sF(M_{[d]},T_{d-1}) 
+  \underlabel_{\text{inference } d}
+  \sF(M_{[d]}, T_d)
+\end{align}
+
+\end{comment}
+
+\begin{comment}
+\WFcomment{There is a filtration interpretation when you have the appropriate sufficiency properties.} Let $\sF_{k,i}$ denote the $\sigma$-algebra generated by $M_{[k]}$ and $p_{[i]}$.
+\begin{align*}
+  \sF_{k,i} &= \sF(M_{[k]},p_{[i]})\\
+  \sF_0 &\underlabel_{\text{selection } 1} \sF_{1,0} \underlabel_{\text{inference } 1}
+  \sF_{1,1} \;\;\sub \cdots \sub\;\;
+  \sF_{d-1,d-1} \underlabel_{\text{selection } d} \sF_{d,d-1}
+  \underlabel_{\text{inference } d} \sF_{d,d}
+\end{align*}
+\end{comment}

### 2015-05-08 paper.tex
more work on 'Sequential Hypothesis Testing' section, reworking some notation.

@@
 H_{0,k}:\; F\in M_{k-1}\quad \text{ vs. } \quad 
-H_{1,k}:\; F\in M_k\setminus M_{k-1},
+H_{1,k}:\; F\in M_\infty\setminus M_{k-1},
 \]
@@
 \subsection{Which Null Hypothesis Should We Test?}
+
+Note that the type I error $V=(\hk-k_0)_+$ is defined in a ``model-centric,'' as opposed to a ``variable-centric,'' fashion: in regression, our model-wise $V$ counts the number of true null {\em models} incorrectly rejected, instead of counting the number of noise variables included in the final model.
+We can define the (model-wise) familywise error rate (FWER) and false discovery rate (FDR) respectively as $\P_F(V >0)$ and $\E_F[V/(\hk \vee 1)]$.
+
+We could have instead defined the ``variable-centric'' type I error $\tV$ as the number of noise variables incorrectly included in the final model, with $\tR$ denoting the total number of variables included.
+Here, by ``noise variable'' we mean one that has a non-zero coefficient in the full model.
+Using selective inference to control the variable-wise FWER $\P_\mu(\tV>0)$ and FDR $\E[\tV / (\tR \vee 1)]$ is an interesting topic for further study.
 
@@
   \quad \text{ vs. } \quad
-  H_{1,k}:\; F \in M_k(Y) \setminus M_{k-1}(Y).
+  H_{1,k}:\; F \in M_\infty \setminus M_{k-1}(Y).
 \]
 The main complication arises from the fact that the null and alternative hypotheses are random.
-To simplify matters, we first consider deriving $p$-values for a fixed candidate pair $m_{k-1} \sub m_{k-1}$.
+To simplify matters, we first consider deriving $p$-values for a fixed candidate hypothesis $M\sub M_{\infty}$.
 
-The random variable  $p_{k,M}(Y)$ is a valid {\em selective $p$-value} for the fixed candidate pair $(m_{k-1},m_k)$ if it is stochastically larger than uniform under the (fixed) null $m_{k-1}$, given that the pair $(m_{k-1},m_k)$ is selected.
-That is,
+The random variable  $p_{k,M}(Y)$ is a valid {\em selective $p$-value} for $M$ if it is stochastically larger than uniform under sampling distributions $F\in M$, given that $M$ is selected. That is,
 \[
-\P_F\left(p_{k,M}(Y) \leq \alpha \mid M_{k-1}(Y) = m_{k-1}\right) 
-\leq \alpha, \quad \forall F\in m_{k-1}, \alpha \in [0,1].
+\P_F\left(p_{k,M}(Y) \leq \alpha \mid M_{k-1}(Y) = M\right) 
+\leq \alpha, \quad \forall F\in M, \alpha \in [0,1].
 \]
-Given selective $p$-values for each fixed candidate pair, we can construct a combined $p$-value for the random null $(M_{k-1},M_k)$: 
+Given selective $p$-values for each fixed candidate pair, we can construct a combined $p$-value for the random null $M_{k-1}(Y)$: 
 \[
-p_k(y) = p_{k, M_{k-1}(y), M_k(y)}(y),
+p_k(y) = p_{k, M_{k-1}(y)}(y),
 \]
@@
 \[
-\P_F\left(p_k \leq \alpha \mid M_{k-1}, \;M_k, 
+\P_F\left(p_k \leq \alpha \mid M_{k-1}, 
   \;F\in M_{k-1}\right) \leq \alpha, \quad \forall \alpha \in [0,1].
 \]
-Recall that, as usual, $F$ is not random, but $M_{k-1}$ is.
+Recall that, as usual $M_{k-1}$ is random and not $F$.
 
@@
 
-\subsection{Selected- and Saturated-Model inference}
+\subsection{Comparison of Selected- and Saturated-Model inference}
 
@@
 The number of type I errors, then, is $(\hk-k_0)_+$, while the number of type II errors is $(k_0-\hk)_+$.
-Define $V=(\hk-k_0)_+$, the number of excess rejections.
-
-Note that the type I error $V$ is defined in a ``model-centric,'' as opposed to a ``variable-centric,'' fashion: in regression, our model-wise $V$ counts the number of true null {\em models} incorrectly rejected, instead of counting the number of noise variables included in the final model.
-We can define the (model-wise) familywise error rate (FWER) and false discovery rate (FDR) respectively as $\P_F(V >0)$ and $\E_F[V/(\hk \vee 1)]$.
-
-We could have instead defined the ``variable-centric'' type I error $\tV$ as the number of noise variables incorrectly included in the final model, with $\tR$ denoting the total number of variables included.
-Here, by ``noise variable'' we mean one that has a non-zero coefficient in the full model.
-Using selective inference to control the variable-wise FWER $\P_\mu(\tV>0)$ and FDR $\E[\tV / (\tR \vee 1)]$ is an interesting topic for further study.
 
@@
 While strong stop and forward stop tend to give more powerful stopping rules, they require independence among the $p$-values.
-Sections~\ref{sec:modelSPSP}--\ref{sec:pvalSP} discuss conditions on the model sequence $M_{[d]}$ and the $p$-value sequence $p_{[d]}$ under which selective $p$-values are mutually independent.
+Sections~\ref{sec:pValsIndep}--\ref{sec:modelSSP} discuss conditions on the model sequence $M_{[d]}$ and the $p$-value sequence $p_{[d]}$ under which selective $p$-values are mutually independent.
 We will see that, typically, selected-model $p$-values are sequentially independent while saturated-model $p$-values are not.
@@
 
-\subsection{$p$-Values and Independence}
+\subsection{$p$-Values and Independence}\label{sec:pValsIndep}
 
@@
   If the $p$-values are exact then $p_{k+1}, \ldots, p_d$ are
-  uniform given $\sF_{k_0}$ on the event $\{k_0=k\}$.
-
-  If they $p$-values are conservative, then for
+  independent and uniform given $\sF_{k_0}$ on the event $\{k_0=k\}$.
+  If they are conservative, then for
   $\alpha_{k+1},\ldots,\alpha_d \in [0,1]$,
@@
 \end{proposition}
-Equation~\ref{eq:jointConserv} is slightly weaker than independence, but it guarantees that both Strong Stop and Forward Stop control their target error rates.
+Equation~\ref{eq:jointConserv} does not quite imply independence, but it guarantees that both Strong Stop and Forward Stop control their target error rates.
 
@@
   If the $p_k$ are exact then the above 
-  inequalities become equalities, implying uniformity.
+  inequalities become equalities, implying uniformity and joint independence.
 \end{proof}
 
-\subsection{Sufficient Filtration}
+\subsection{The Sufficient Filtration}\label{sec:suffFilt}
 
@@
 First, we typically construct $p_k$ by conditioning on $T_{k-1}$.
-In fact, $T_{k-1}$ is complete sufficient, then any exact test must condition on $T_{k-1}$ \WFcomment{should we bother proving this?}.
+In fact, if $T_{k-1}$ is complete sufficient, then any exact test must condition on $T_{k-1}$ (apply the definition of completeness to the function $h(Y) = \P(p_k(Y) \leq \alpha \mid T_{k-1}(Y))$).
 Second, most path algorithms of interest (including forward stepwise regression, the ever-active LASSO path, and their generalizations) satisfy a {\em subpath sufficiency principle} (henceforth, SSP): that once we know $M_k$ and its sufficient statistics, we can reconstruct the model path up to step $k$.
@@
 
-\begin{comment}
+\subsection{Model Paths and the Subpath Sufficiency Principle}\label{sec:modelSSP}
 
-\begin{theorem}
-  Assume that $T(Y; M)$ is a complete sufficient statistic, 
-  that $p_k$ are exact selective $p$-values, and that 
-  $M_{[d]}$ satisfies EPIC.
-
-  Then, the sufficient filtration separates $p_{[d]}$ if and only if 
-  $p_k \in \sF_k$.
-\end{theorem}
+In Section~\ref{sec:suffFilt} we introduced the sufficient filtration as a way to get separated $p$-values.
+Because conditioning on $(M_{[k]}, T_k)$ is nominally different from conditioning only on $(M_k, T_k)$, we might suspect that we will sacrifice power by requiring $p_k$ to condition on the former.
 
@@
 \begin{model}[Generalized Sparse Model]\label{mod:genSparse}
-  Let $M$ be a model parameterized by $\theta\in \Theta \sub \R^p$:
+  Let $M_\infty$ be a model parameterized by $\theta\in \Theta \sub \R^p$:
   \[
-  M = \{F_\theta:\; \theta \in \Theta\}.
+  M_\infty = \{F_\theta:\; \theta \in \Theta\}.
   \]
@@
   \]
-  Let $L(\theta; Y)$ denote the likelihood for model $M$.
+  Let $\ell(\theta; Y)$ denote the log-likelihood for model
+  $M_\infty$, and define the restricted log-likelihood
+  \[
+  \ell_E(\theta; Y) = \left\{\begin{matrix} \ell(\theta; \;Y) 
+      & \theta \in \Theta(E)\\ -\infty  & 
+      \mathrm{ otherwise.}\end{matrix}\right. 
+  \]
 \end{model}
-Note that, although we will make no assumptions about the form of $M$, the results of this section are only useful in cases where we can actually construct single-step $p$-values.
+Although we will make no formal assumptions about the form of $M_\infty$, the results of this section are only useful in cases where we can actually construct single-step $p$-values.
 At present, that caveat  largely restricts us to using exponential family models.
@@
 \begin{example}[Principle Components Analysis]
-\WFcomment{Do this one.}  
+\WFcomment{Do this one. Actually need infinitely many $\theta$s for this.}  
 \end{example}
 
-All forward stepwise procedures satisfy the SPSP.
+All forward stepwise procedures satisfy the SSP.
 
-\begin{proposition}[Forward Stepwise Satisfies SPSP]\label{prop:forwardSPSP}
-  The forward stepwise algorithm in Model~\ref{mod:genSparse} as defined in (\ref{eq:forwardDef_start}--\ref{eq:forwardDef_end}) satisfies the SPSP.
+\begin{proposition}[Forward Stepwise Satisfies SSP]\label{prop:forwardSSP}
+  The forward stepwise algorithm in Model~\ref{mod:genSparse} as defined in (\ref{eq:forwardDef_start}--\ref{eq:forwardDef_end}) satisfies the SSP.
 \end{proposition}
 \begin{proof}
-  Let $A$ denote the event $\{M_k = M(E)\}$ for some fixed active set $E$.
-Letting $T(Y)$ be a sufficient statistic for $m_k$, the likelihood $L$ restricted to $\Theta(E) \times A$ is proportional to a function depending only on $T$.
-If $L_T$ is the likelihood of $T(Y)$, then
+  Let $A$ denote the event $\{M_k = M(E)\}$ for some fixed active set $E$. Writing $T(Y) = T(Y; \;M(E))$, the likelihood restricted to $\Theta(E)$ is proportional to a function depending only on $T$:
 \[
-L(\theta; y) = L_T(\theta; T(y)) \;L_{Y\mid T}(y), \;\; \text{ for } (\theta, y) \in  \Theta(E) \times A.
+\ell(\theta; Y) = \ell_E^T(\theta; T(Y)) + \ell_E^{Y \mid T}(Y, T(Y)),
 \]
-Note that $L_{Y \mid T}(y)$ does not depend on $\theta$ since $T$ is sufficient for $M(E)$.
+where the second term does not depend on $\theta$ because $T$ is sufficient.
 
-For $y\in A$ and $i \leq k$, we have $E_i(y) \sub E$. This means that the maximizer in \eqref{eq:forwardDef_start} is attained for some $j\in E$. So, for $i \leq k$, on $A$,
+On $A$, we have $E_i(Y) \sub E$ for all $i=1,\ldots, k$, meaning that the maximum in \eqref{eq:forwardDef_start} was attained by some $j\in E$ at every step. So, for $i \leq k$, we have
 \begin{align}
-  j_i &= \argmax_{j\in E} \;\;\sup \left\{L(\theta; Y):\;
+  j_i &= \argmax_{j\in E} \;\;\sup \left\{\ell_E(\theta; Y):\;
     \theta\in\Theta(E_{i-1} \cup \{j\})\right\} \\
-  &= \argmax_{j\in E} \;\;\sup \left\{L_T(\theta; T(Y)):\;
+  &= \argmax_{j\in E} \;\;\sup \left\{\ell_E^T(\theta; T(Y)):\;
     \theta\in\Theta(E_{i-1} \cup \{j\})\right\}. \label{eq:onlyT}
 \end{align}
-Replacing $L$ with $L_T$ in~\eqref{eq:onlyT} is justified because $\Theta(E_{i - 1} \cup \{j\}) \sub \Theta(E)$ when $E_{i-1}\cup E$ and $j\in E$.
 
-Equation~\eqref{eq:onlyT} shows that $j_1,\ldots, j_{k-1}$ all depend on $Y$ only through $T(Y)$. As a result, it also follows that the sequence $M_{[k-1]}(Y)$ depends only of $T(Y)$.
+Equation~\eqref{eq:onlyT} shows that $j_1,\ldots, j_{k-1}$ all depend on $Y$ only through $T(Y)$. As a result, it also follows that the enitre sequence $M_{[k]}(Y)$ depends only on $T(Y)$.
 \end{proof}
 
-Another class of model selection procedures satisfying the SPSP is the sequence ``ever-active'' sets in regularized likelihood methods.
-For $r=0,1,\ldots$, let $P_r(\theta)$ denote a regularizing penalty, and define
+Another class of model selection procedures satisfying the SSP is the sequence ``ever-active'' sets in any regularized likelihood method.
+For $r=0,1,\ldots,m$, let $P_r(\theta)$ denote a regularizing penalty, and define
 \begin{align}\label{eq:regPathDef_start}
   \hat\theta^{r}(Y) &= 
-  \argmin_{\theta\in\Theta} -\log L(\theta; Y) + P_r(\theta) \\
+  \argmin_{\theta\in\Theta} -\ell(\theta; Y) + P_r(\theta) \\
   \tE_r(Y) &= \left\{j:\; \hat\theta_j^s \neq 0 
@@
 
-\begin{proposition}[Regularized Likelihood Paths Satisfy SPSP]\label{prop:regPathSPSP}
+\begin{proposition}[Regularized Likelihood Paths Satisfy SSP]\label{prop:regPathSSP}
 Ever-active model paths derived from regularized likelihood methods as in (\ref{eq:regPathDef_start}--\ref{eq:regPathDef_end}) satisfy the sub-path sufficiency principle.
@@
 \begin{proof}
-  Again, let $A$ denote the event $\{M_k = M(E)\}$ for some fixed $E$, and define $T$ and $L_T$ as in the proof of Proposition~\ref{prop:forwardSPSP}.
-We can partition $A$ into subsets $A_r = A \cap \{R(k)=r\}$.
-
-On the event $A_r$, every $\hat\theta^s$ for $s\leq r$ is supported on $E$. That is, for $s \leq r$, we can write 
-\begin{equation}\label{eq:onlyT_Ar}
-  \hat\theta^{s} = 
-  \argmin_{\theta\in\Theta(E)} -\log L_T(\theta; T(Y)) + P_s(\theta) \quad \text{ on } A_r
-\end{equation}
-Equation~\label{eq:onlyT_Ar} guarantees that $M_{[k-1]}$ is a function of $T(Y)$ on $A_r$. That is, we can write
+  Again, let $A$ denote the event $\{M_k = M(E)\}$ for some fixed $E$, and define $T$ and $\ell_E^T$ as in the proof of Proposition~\ref{prop:forwardSSP}. If we define
 \[
-M_{[k-1]}(Y) = M_{[k-1]}^r(T(Y)) \quad \text{ on } A_r
+\hat\theta^{(E,r)} = \argmin_{\theta\in\Theta(E)} -\ell_E^T(\theta; T(Y)) + P_r(\theta),
 \]
-It remains only to show that, on $A$, $R_k(Y)$ depends only on $T(Y)$; then we have
+then 
 \[
-M_{[k-1]}(Y) = M_{[k-1]}^{R(T)}(T) \quad \text{ on } A
+\hat\theta^{(E,r)} \eqAS \hat\theta^{r} \text{ on } 
+A \cap 1_{\{r \leq R_k\}}.
 \]
-But $R_k$ is a function of $T$ on $A$ for the same reason: that $\hat\theta^s$ depends only on $T$ until other coefficients enter the model.
-\WFcomment{Would the proof be cleaner with $A_r$ defined as $A\cap \{R_k \leq r\}$?}
+But $\hat\theta^{(E,r)}$ depends only on $T(Y)$. Therefore, on $A$, we can reconstruct the entire path of solutions once we know $T(Y)$.
 \end{proof}
 
-\subsection{Independent $p$-Values Via Selection Variables}\label{sec:selectionVariables}
+\begin{comment}
 
-As discussed in \citet{fithian2014optimal}, there are occasions when it is desirable for some reason to condition on more than just the selection event.
-On these occasions we can introduce a {\em selection variable} --- any random variable that is finer than the indicator of the selection event --- and condition on its value.
-Conditioning on a finer selection variable preserves validity conditional on the selection event, but tends to reduce the power of a test.
+\begin{theorem}
+  Assume that $T(Y; M)$ is a complete sufficient statistic, 
+  that $p_k$ are exact selective $p$-values, and that 
+  $M_{[d]}$ satisfies EPIC.
 
-For example, in linear regression after selecting variables via the lasso, \citet{lee2013exact} propose conditioning not only on the set of active variables but also on the signs of the active coefficients.
-The selection event corresponding to a given active set $E\sub [p]$ typically contains up to $2^{|E|}$ convex polytopes, each of which corresponds to a particular sign pattern $\hat z_E$.
-Thus, conditioning on the selection variable $\hat z_E$ can simplify computation considerably.
+  Then, the sufficient filtration separates $p_{[d]}$ if and only if 
+  $p_k \in \sF_k$.
+\end{theorem}
 
-We say that $p_{k,M}$ is a selective $p$-value with respect to selection variable $S(Y)$ if it is uniform under $F\in m_k$ given $S$.
-
-To achieve joint error control properties among different tests, we may need to condition on a finer variable.
 \end{comment}
@@
 \begin{comment}
-Sections~\ref{sec:pvalSP}--\ref{sec:modelSPSP} discuss sufficient conditions on the $p$-values and the selection algorithm under which single-step $p$-values are automatically independent.
+Sections~\ref{sec:pvalSP}--\ref{sec:modelSSP} discuss sufficient conditions on the $p$-values and the selection algorithm under which single-step $p$-values are automatically independent.
 Section~\ref{sec:selectionVariables} discusses how we can create independent $p$-values by conditioning on finer selection variables at each step.

### 2015-05-09 code/parametric_bootstrap/parametric_bootstrap.R
added maxT step8 under global null, AIC, BIC

@@ -1,22 +1,66 @@
 set.seed(0)
 X = read.table("X_100_200.csv", sep=',')
+D = X
 X = as.matrix(X) 
 Y = as.matrix(read.table("y_100.csv", sep=','))
 variables = as.matrix(read.table("variables.csv", sep=','))
+D$Y = Y
+
+BICmodel = step(lm(Y ~ 1, data=D), scope=list(upper= ~ V1 + V2 + V3 + V4 + V5 + V6 + V7 + V8 + V9 + V10 + V11 + V12 + V13 + V14 + V15 + V16 + V17 + V18 + V19 + V20 + 
+                                                   V21 + V22 + V23 + V24 + V25 + V26 + V27 + V28 + V29 + V30 + V31 + V32 + V33 + V34 + V35 + V36 + V37 + V38 + V39 + 
+                                                   V40 + V41 + V42 + V43 + V44 + V45 + V46 + V47 + V48 + V49 + V50 + V51 + V52 + V53 + V54 + V55 + V56 + V57 + V58 + V59 + 
+                                                   V60 + V61 + V62 + V63 + V64 + V65 + V66 + V67 + V68 + V69 + V70 + V71 + V72 + V73 + V74 + V75 + V76 + V77 + V78 + V79 + 
+                                                   V80 + V81 + V82 + V83 + V84 + V85 + V86 + V87 + V88 + V89 + V90 + V91 + V92 + V93 + V94 + V95 + V96 + V97 + V98 + V99 + 
+						   V100 + V101 + V102 + V103 + V104 + V105 + V106 + V107 + V108 + V109 + V110 + V111 + V112 + V113 + V114 + V115 + V116 + 
+						   V117 + V118 + V119 + V120 + V121 + V122 + V123 + V124 + V125 + V126 + V127 + V128 + V129 + V130 + V131 + V132 + V133 + 
+						   V134 + V135 + V136 + V137 + V138 + V139 + V140 + V141 + V142 + V143 + V144 + V145 + V146 + V147 + V148 + V149 + V150 + 
+						   V151 + V152 + V153 + V154 + V155 + V156 + V157 + V158 + V159 + V160 + V161 + V162 + V163 + V164 + V165 + V166 + V167 + 
+						   V168 + V169 + V170 + V171 + V172 + V173 + V174 + V175 + V176 + V177 + V178 + V179 + V180 + V181 + V182 + V183 + V184 + 
[…]

### 2015-05-09 code/parametric_bootstrap/parametric_bootstrap.R
message change

@@ -80,7 +80,7 @@ wait_until_accept = function(A, R, PY, sigma) {
         if(max(A %*% Z) < 0) {
             return(count)
         }
-	if(count > 100000) {
+	if(count >= 100000) {
             return(count) # give up...
         }
    }
@@ -107,5 +107,5 @@ for (i in 1:100) {
 print('maxT step 12')
 print(1. / mean(maxT12))
 
-print('maxT step8 under global null with same model, to get one point...')
+print('maxT step8 under global null with same model, to get one point... 100000 means it gave up')
 print(wait_until_accept(A7, R8, P8Y0, sigma))

### 2015-05-09 code/parametric_bootstrap/parametric_bootstrap.R
added null test to R script, looking at talk scenario p=40 snr=5

@@ -3,8 +3,98 @@ X = read.table("X_100_200.csv", sep=',')
 D = X
 X = as.matrix(X) 
 Y = as.matrix(read.table("y_100.csv", sep=','))
+Y0 = as.matrix(read.table("y0_100.csv", sep=','))
 variables = as.matrix(read.table("variables.csv", sep=','))
+variables0 = as.matrix(read.table("variables0.csv", sep=','))
 D$Y = Y
+
+print("step 8")
+print(variables[1:8])
+X8 = X[,variables[1:7]]
+print("step 8 null")
+print(variables[1:8])
+X08 = X[,variables0[1:7]]
+P8 = X8 %*% solve(t(X8) %*% X8) %*% t(X8)
+P08 = X08 %*% solve(t(X08) %*% X08) %*% t(X08)
+A7 = as.matrix(read.table("A_step7.csv", sep=','))
+A07 = as.matrix(read.table("A0_step7.csv", sep=','))
+A8 = as.matrix(read.table("A_step8.csv", sep=','))
+
+print("step 12")
+print(variables[1:12])
+X12 = X[,variables[1:11]]
+P12 = X12 %*% solve(t(X12) %*% X12) %*% t(X12)
+A11 = as.matrix(read.table("A_step11.csv", sep=','))
+
+sigma = as.numeric(read.table("sigma.csv"))
+
+n = nrow(X)
+p = ncol(X)
+R8 = diag(rep(1,n)) - P8
+R08 = diag(rep(1,n)) - P08
+R12 = diag(rep(1,n)) - P12
+
+P8Y = P8 %*% Y
+P8Y0 = P08 %*% Y0
+P12Y = P12 %*% Y
+
+# here is the null distribution
+
+null_sample = function(R, PY, sigma) {
+     n = nrow(R)
+     Z = PY + R %*% rnorm(n) * sigma
+     return(Z)
+}
+
+# how many tries until we accept
+
+wait_until_accept = function(A, R, PY, sigma) {
+    count = 0
+    while(TRUE) {
+        count = count + 1
+        Z = null_sample(R, PY, sigma)
+        if(max(A %*% Z) < 0) {
+            return(count)
+        }
+	if(count >= 100000) {
+            return(count) # give up...
+        }
+   }
+}
+
+
+maxT_sample = c()
+for (i in 1:100) {
+    maxT_sample = c(maxT_sample, wait_until_accept(A7, R8, P8Y, sigma))
+}
+
+print('maxT step8')
[…]

### 2015-05-10 paper.tex
tinkering on paper

@@
 while adjusting for the fact that the models are chosen adaptively.
+In most of our development we will make few formal assumptions about the form of the candidate models.
+Still, our results are only useful in cases where we can actually construct single-step $p$-values.
+At present, that means our applications are largely restricted to exponential family models.
+
+Most of our examples will focus on the important and familiar problem of adaptively selecting a set of predictors in linear regression models.
 As we will see, {\em selected-model} tests can be dramatically more powerful than {\em saturated-model} tests at early steps in the model path when most of the signal variables have not yet entered the model.
 
-The second half of the paper concerns {\em stopping rules} $\hk$, estimators of $k_0$.
+Section~\ref{sec:sequential} concerns {\em stopping rules} $\hk$, estimators of $k_0$.
 We will consider several stopping rules that operate on the full sequence $p_{[d]}$, including two powerful stopping rules proposed in \citet{gsell2013sequential}.
@@
 We will prove sufficient conditions for when the $p$-values are Gaussian.
-
-Although we will focus most of our attention and examples on the case of selecting a set of predictors in linear regression, essentially all of our results apply in generic exponential family models.
 
@@
 There is some ambiguity involved in deciding how to generalize $z$- and $t$-tests to the selective case.
-For example, \citet{gsell2013sequential} describe three different null hypotheses that we could consider testing at step $k$.
-As discussed at some length in \citet{fithian2014optimal}, there are several different models that
+For example, \citet{gsell2013sequential} describe three different null hypotheses that we could consider testing at step $k$ in the case of linear regression under the upper model $M_\infty:\; Y\sim \cN(\mu, \sigma^2 I)$:
+\begin{description}
+\item[Complete Null:] The previous model, $M_{k-1}$, is correct; i.e., that $\mu = X_{E_{k-1}}\beta$ for some $\beta$.
+\item[Incremental Null:] $M_{k-1}$ may not be correct, but $M_k$ is no improvement: i.e., that $X_EX_E\mu = \proj_{X_{E_{k-1}}}\mu$
+\end{description}
+. \citet{fithian2014optimal} discusses several different tests
 
@@
 
-\section{Sequential Inference}
+\section{Sequential Inference}\label{sec:sequential}
 
@@
 
-We will consider stopping rules that depend on the sequence $p_1,\ldots,p_d$ of $p$-values, which have been studied in the literature by \WFcomment{literature review here.} We will focus on three such stopping rules: simple stop, proposed by \WFcomment{whom?}, and the more refined strong stop and forward stop, proposed by \citet{gsell2013sequential}.
+We will consider stopping rules that depend on the sequence $p_1,\ldots,p_d$ of $p$-values, which have been studied in the literature by \WFcomment{literature review here.} We will focus on three such stopping rules: Basic Stop, proposed by \WFcomment{whom?}, and the more refined Strong Stop and Forward Stop, proposed by \citet{gsell2013sequential}.
 
-While strong stop and forward stop tend to give more powerful stopping rules, they require independence among the $p$-values.
+While Strong Stop and Forward Stop tend to give more powerful stopping rules, they require independence among the $p$-values.
 Sections~\ref{sec:pValsIndep}--\ref{sec:modelSSP} discuss conditions on the model sequence $M_{[d]}$ and the $p$-value sequence $p_{[d]}$ under which selective $p$-values are mutually independent.
@@
 \]
-We will call this procedure {\em basic stop}.
+We will call this procedure {\em Basic Stop}.
 
@@
 \end{model}
-Although we will make no formal assumptions about the form of $M_\infty$, the results of this section are only useful in cases where we can actually construct single-step $p$-values.
-At present, that caveat  largely restricts us to using exponential family models.
 

### 2015-05-12 paper.tex
added section on computation, along with theorem about forward stepwise constraints

@@
 
+\section{Computation}
+
+As usual, we have reduced computation of the $p$-value at step $k$ to a sampling problem. Whatever our test statistic at step $k$, we must compute its distribution under the null model $M_{k-1}$.
+
+Happily, the polytope to condition on has only $2p$ constraints in forward stepwise regression, as we see next.
+
+\begin{theorem}
+  Assume the model path $M_{[d]}$ is obtained by forward stepwise 
+  linear regression. Then, for a candidate active set $E$ of size $k$, 
+  the set $A = \{E_k = E, \;X_E'Y = u\}$ is characterized 
+  exactly by the constraints $X_E'Y=u$ and
+  \[
+  v_j^-(E,u) < X_j'Y < v_j^+(E,u), \quad\forall j \notin E,
+  \]
+  for $v_j^-$ and $v_j^+$ given explicitly in
+  Equations~(\ref{eq:vMinus_FS}--\ref{eq:vPlus_FS}).
+  Thus, $A$ corresponds exactly to 
+  a set with $2(p-k)$ linear inequality constraints and $k$
+  linear equality constraints on $Y$.
+\end{theorem}
+
+\begin{proof}
+  At stage $i < k$, the next variable to enter is the one with maximal correlation with the residual vector; i.e.
+  \[
+  j_{i+1} = \argmax_{j \notin E_i} 
+  \frac{\left| X_j ' (Y - X_{E_i}\hat\beta^i) \right|}
+  {\|\proj_{E_i}^\perp X_j\|}
+  \]
+  The crucial observation is that, 
+  once we know the $k$th active set 
+  $E_k$ and its sufficient statistics
+  $X_{E_k}'Y$, we know the entire 
+  path of fitted models up to step $k$.
+  On the set $\{E_k=E\}$, all of the quantities 
+  $j_{i+1}$, $E_i$, $\hat\beta^i$, and $\proj_{E_i}^\perp$
+  depend only on $X_E'Y$. For brevity, write
+  \[
+  C_i^* = \max_{j \notin E_i} 
+  \frac{\left| X_j ' (Y - X_{E_i}\hat\beta^i) \right|}
+  {\|\proj_{E_i}^\perp X_j\|}.
+  \]
+  On $A$, $C_i^*$ is attained at $j=j_{i+1}$, the $i$th 
+  variable added.
+
+  If $X_E'Y$ is fixed at $u$, and $j \notin E$, then the condition for 
+  $X_j$ {\em not} to enter at step $i+1 < k$ is
+  \[
+  \frac{\left| X_j' (Y - X_{E_i}\hat\beta^i) \right|}
+  {\|\proj_{E_i}^\perp X_j\|} 
+  \leq C_i^*,
+  \]
+  or equivalently,
+  \begin{equation}\label{eq:noEnterBounds_FS}
+    X_j' X_{E_i}\hat\beta^i -
+    C_i^*\|\proj_{E_{i}}^\perp X_{j}\|
+    \;\;\;\leq\;\;\;
+    X_j'Y
+    \;\;\;\leq\;\;\;
+    X_j' X_{E_i}\hat\beta^i +
+    C_i^*\|\proj_{E_{i}}^\perp X_{j}\|, 
+  \end{equation}
+  so the set $A$ is equivalent to 
+  Equation~\eqref{eq:noEnterBounds_FS} holding
+  for every $i < k$ and $j \notin E$. Nominally, this gives $2k(p-k)$
+  linear inequality constraints to satisfy, but most of them are
+  non-binding. On $A$, the upper and lower bounds
+  in~\eqref{eq:noEnterBounds_FS}
+  are all known functions of $E$ and $u$, so we can set
+  \begin{align}\label{eq:vMinus_FS}
+    v_j^-(E,u) &= \max_{0 \leq i < k} \;\;X_j' X_{E_i}\hat\beta^i -
+    C_i^*\|\proj_{E_{i}}^\perp X_{j}\| \\
+    \label{eq:vPlus_FS}
+    v_j^+(E,u) &= \min_{0 \leq i < k} \;\;X_j' X_{E_i}\hat\beta^i +
+    C_i^*\|\proj_{E_{i}}^\perp X_{j}\|.
+  \end{align}  
+\end{proof}
+
+As we see below, a similar result holds for the LASSO ever-active path.
+In fact, it holds for a much larger class of $\ell_1$ regularized exponential family models including the graphical LASSO, group LASSO, and...
+\WFcomment{Fill in this proof.}
+
 \section{Simulation: Sparse Linear Regression}\label{sec:sparseReg}
@@
 
-\WFcomment{Should we compute the FWER and FDR only conditional on screening? Strong / Forward / Simple should all be correct given screening since they condition on $p_{[k_0-1]}$.}
+\WFcomment{Should we compute the FWER and FDR only conditional on screening? Strong / Forward / Basic should all be correct given screening since they condition on $p_{[k_0-1]}$.}
 

### 2015-05-12 paper.tex
added theorem on L1 penalized exponential families

@@
 
-\section{Computation}
+\section{Computation}\label{sec:computation}
 
@@
 In fact, it holds for a much larger class of $\ell_1$ regularized exponential family models including the graphical LASSO, group LASSO, and...
-\WFcomment{Fill in this proof.}
+\WFcomment{more?
+group LASSO not really $\ell_1$ penalty, but the result is true for group LASSO}
+
+\begin{theorem}
+  Assume that $M_\infty$ is an exponential family model
+  of the form
+  \[
+  Y \sim \exp\{ \theta'U(y) - \psi(\theta) \}\,d\nu(y),
+  \]
+  with $\Theta \sub \R^p$ convex, and assume
+  that the model path $M_{[d]}$ is given 
+  by the ever-active set for the $\ell_1$-penalized problem
+  \begin{equation}\label{eq:regProblem}
+  \hat\theta^r = \argmin_{\theta\in \Theta} 
+  -\ell(\theta) + \lambda_r\|\theta\|_1,
+  \end{equation}
+  for some sequence $\lambda_r > 0$.
+
+  Then, for a candidate active set $E$ of size $k$, 
+  the set $A = \{E_k = E, \;U_E = u\}$ is characterized 
+  exactly by the constraints $U_E = u$ and
+  \[
+  v_j^-(E,t) \leq U_j(Y) \leq v_j^+(E,t), \quad\forall j \notin E,
+  \]
+  for $v_j^-$ and $v_j^+$ given explicitly in
+  Equations~(\ref{eq:vMinus_L1}--\ref{eq:vPlus_L1}).
+  Thus, $A$ corresponds exactly to 
+  a set with $2(p-k)$ linear inequality constraints and $k$
+  linear equality constraints on $U(Y)$.
+\end{theorem}
+
+\begin{proof}
+  Note that, because exponential family likelihoods are concave in their natural parameters, the problem in~\eqref{eq:regProblem} is convex.
+Again, once we know the sufficient statistics $U_E(Y)$ for model $k$, and that $E_k=E$, we know that
+  \[
+  \hat\theta^r = \hat\theta^{(E,r)}
+  \]
+  for every $r \leq R_k$, so we know the entire path of fits up to and including $R_k$. But then, excluding variable $j \notin E$ at Lagrange parameter $\lambda_r$ is equivalent to
+  \[
+  \lambda_r \geq 
+  \left| \pardd{\ell(\hat\theta^r)}{\theta_j} \right|
+  = \left|U_j - \E_{\hat\theta^r}[U_j]\right|
+  \]
+  On $A$, for $r \leq R_k$, 
+  $\E_{\hat\theta^r}[U_j]$ is a known function of $E$ and $u$,
+  so we can set
+  \begin{align}\label{eq:vMinus_L1}
+    v_j^-(E,u) &= \sup_{r \leq R_k} \;\; 
+    \E_{\hat\theta^r}[U_j] - \lambda_r \\
+    \label{eq:vPlus_L1}
+    v_j^+(E,u) &= \inf_{r \leq R_k} \;\;
+    \E_{\hat\theta^r}[U_j] + \lambda_r .
+  \end{align}
+\end{proof}
 

### 2015-05-13 paper.tex
added more discussion of single-step selective p-values

@@
 \item[Complete Null:] The previous model, $M_{k-1}$, is correct; i.e., that $\mu = X_{E_{k-1}}\beta$ for some $\beta$.
-\item[Incremental Null:] $M_{k-1}$ may not be correct, but $M_k$ is no improvement: i.e., that $X_EX_E\mu = \proj_{X_{E_{k-1}}}\mu$
+\item[Incremental Null:] $M_{k-1}$ may not be correct, but $M_k$ is no improvement: i.e., that $X_EX_E^\dagger\mu = \proj_{E_{k-1}}\mu$
 \end{description}
@@
 
-The random variable  $p_{k,M}(Y)$ is a valid {\em selective $p$-value} for $M$ if it is stochastically larger than uniform under sampling distributions $F\in M$, given that $M$ is selected. That is,
+The random variable  $p_{k,M}(Y)$ is a valid {\em selective $p$-value} for $M$ at step $k$ if it is stochastically larger than uniform under sampling distributions $F\in M$, given that $M$ is selected.
+That is,
 \[
@@
 \]
-Given selective $p$-values for each fixed candidate pair, we can construct a combined $p$-value for the random null $M_{k-1}(Y)$: 
+Given selective $p$-values for each fixed candidate pair, we can construct a combined $p$-value for the random null $M_{k-1}(Y)$. Define
 \[
@@
 \]
-Recall that, as usual $M_{k-1}$ is random and not $F$.
+Recall that $M_{k-1}$ is random and not $F$.
 
@@
 
+\subsection{Selective $p$-Values in Exponential Families}
+
 \subsection{Selective $p$-Values in Regression}
 
-\subsection{Comparison of Selected- and Saturated-Model inference}
+In linear regression, there are several tests that we might perform at step $k$ if we are , which generally use
+
+\subsubsection{Saturated-Model Tests}
+Most of the work in selective inference for linear regression has focused on inference for least-squares parameters of the model $M_\infty:\; Y \sim \cN(\mu, \sigma^2I)$, with $\sigma^2$ assumed known.
+The saturated-model test for significance of variable $j_k$ after adjusting for the variables in $E_{k-1}$ tests whether the coefficient of $X_{j_k}$ is nonzero in the best linear model $X_{E_k}^\dagger Y$.
+
+Let $\proj_{E_{k-1}}\mu$ denote the projection of $\mu$ onto the column space of $X_{E_{k-1}}$, and let $\eta = \proj_{E_{k-1}}^\perp X_{j_k}$.
+Saturated-model $p$-values test the null hypothesis $H_0:\;\eta'\mu = 0$.
+Because $M_\infty$ has $n$ parameters, we must  condition on $\proj_\eta^\perp Y$ to eliminate the effect of the nuisance parameters $\proj_\eta^\perp \mu$ from the problem.
+Thus, the test is based on
+\begin{equation}\label{eq:satModel}
+\L\left( \eta'Y \mid E_{k-1}, j_k, \proj_\eta^\perp Y \right).
+\end{equation}
+To carry out exact tests of this hypothesis, we must assume that $\sigma^2$ is known, or that an independent estimate is available.
+
+\WFcomment{References.} \citet{fithian2014optimal} call $M_\infty$ the {\em saturated model} because there is a parameter for every data point.
+They contrast tests based on $M_\infty$ with {\em selected-model} tests, which do not treat $\proj_\eta^\perp \mu$ as a nuisance parameter.
+
+\subsubsection{Selected-Model Tests}
+Selected-model inference~\citep{fithian2014optimal} presents a more powerful alternative to testing for inclusion of variable $j_k$.
+Selected-model tests are so called because they are based on the statistical model chosen by our selection procedure.
+In this case, the null model selected for testing at step $k$ is simply the model $M_{k-1}:\; Y \sim \cN(X_{E_{k-1}}\beta, \sigma^2 I)$.
+
+There is a UMPU test of the null $M_{k-1}$ against the alternative $M_{k}\setminus M_{k-1}$, which has one more degree of freedom than $M_{k-1}$.
+One option for selected-model inference is to condition on the identity of $j_k$, the next variable added, and perform the UMPU selective test of $M_{k-1}$ against $M_{k}\setminus M_{k-1}$.
+In that case, the test is based on
+\begin{equation}\label{eq:selModel_cond}
+\L\left(X_{j_k}'Y \mid E_{k-1}, \; j_k, \; X_{E_{k-1}}'Y\right).
+\end{equation}
+Note the contrast between~\eqref{eq:selModel_cond} and~\eqref{eq:satModel}.
+In~\eqref{eq:selModel_cond}, we only need to condition on an $|E_{k-1}|$-dimensional projection of $Y$, whereas in~\eqref{eq:satModel} we condition on $n-1$-dimensional projection of $Y$.
+Conditioning on more information tends to sap the power of a test, and we will see in Sections~\ref{sec:bivariate} and~\ref{sec:sparseReg}, the saturated-model tests often pay a heavy price for this extra conditioning.
+
+Although the tests in~\eqref{eq:selModel_cond} and~\eqref{eq:satModel} may appear to be based on different test statistics, they are actually not: because $\eta'Y = X_{j_k}'(Y - \proj_{E_{k-1}}Y)$, one test statistic is a deterministic affine function of the other after we condition on $X_{E_{k-1}}'Y$.
+
+It is not really necessary to condition on $j_k$: instead we could use
+\begin{equation}\label{eq:selModel_marg}
+\L\left(X_{j_k}'Y \mid E_{k-1}, \; X_{E_{k-1}}'Y\right),
+\end{equation}
+or even replace $X_{j_k}'Y$ with any other test statistic.
+Under the null hypothesis, the distribution of $Y$ is completely known once we condition on the value of $X_{E_{k-1}}'Y$, the complete sufficient statistic of $M_{k-1}$.
+
+We will see in Section~\ref{sec:pValsIndep} that conditioning on $j_k$ as in~\eqref{eq:selModel_cond} always results in independent sequential $p$-values, whereas tests of the form~\eqref{eq:selModel_marg} yield independent $p$-values only under certain conditions.
+Section~\ref{sec:pValsIndep} explores the conditions required to obtain independent sequential $p$-values.
+Carrying out a finer comparison of the two selected-model tests is an interesting direction for future work.
+
+Selected-model inference is more computationally involved than saturated-model inference.
+Because the saturated-model test conditions on $n-1$ dimensions, the resulting distribution in~\eqref{eq:satModel} is nothing more than a truncated univariate Gaussian random variable.
+
+However, there are several major benefits to selected-model inference that often justify its use.
+First, selected-model inference allows us to drop the assumption that $\sigma^2$ is known.
+Second, $p$-values for selected-model tests are independent under general conditions, allowing us to plug the $p$-values into powerful sequential stopping rules such as those developed by~\citet{gsell2013sequential}.
+Most importantly, as we will see, selected-model tests can be dramatically more powerful than saturated-model tests.
+See \citet[][Section 5]{fithian2014optimal} for a detailed discussion of the difference between saturated-model and selected-model tests.
+
+\subsection{Comparison of Selected- and Saturated-Model inference}\label{sec:bivariate}
 
@@
 
-As we see below, a similar result holds for the LASSO ever-active path.
-In fact, it holds for a much larger class of $\ell_1$ regularized exponential family models including the graphical LASSO, group LASSO, and...
-\WFcomment{more?
-group LASSO not really $\ell_1$ penalty, but the result is true for group LASSO}
+As we see below, a similar result holds for the LASSO ever-active path. In fact, it holds for a much larger class of $\ell_1$-regularized exponential family models including the graphical LASSO.
 
@@
 \end{proof}
+
+\WFcomment{Give more general characterization of when we can do it, including group lasso? QP? Generalized Lasso?}
 

### 2015-05-15 code/python/compare_methods.py
adding extra arguments for rob running script

@@ -49,11 +49,15 @@ def run(y, X, sigma, active,
         do_knockoff=False,
         alpha=0.05,
         maxstep=np.inf,
-        compute_maxT_identify=True):
+        compute_maxT_identify=True,
+        burnin=2000,
+        ndraw=8000):
 
     n, p = X.shape
     results, FS = compute_pvalues(y, X, sigma, maxstep=maxstep,
-                                  compute_maxT_identify=compute_maxT_identify)
+                                  compute_maxT_identify=compute_maxT_identify,
+                                  burnin=burnin,
+                                  ndraw=ndraw)
     completion_idx = completion_index(results['variable_selected'], active)
     full_results.setdefault('completion_idx', []).append(completion_idx)
 
@@ -108,7 +112,7 @@ def run(y, X, sigma, active,
                                             active,
                                             rule, 
                                             alpha)
-        pval_name = pval.split('_')[0]
+        pval_name = pval.split('_')[:-1]
         full_results.setdefault('%s_%s_R' % (pval_name, rule_name), []).append(R)
         full_results.setdefault('%s_%s_V_var' % (pval_name, rule_name), []).append(V_var)
         full_results.setdefault('%s_%s_V_model' % (pval_name, rule_name), []).append(V_model)

### 2015-05-15 code/python/compute_pvalues.py

@@ -63,7 +63,9 @@ def maxT(FS, sigma=1, burnin=2000, ndraw=8000):
     return pvalue
 
 def compute_pvalues(y, X, sigma=1., maxstep=np.inf,
-                    compute_maxT_identify=True):
+                    compute_maxT_identify=True,
+                    burnin=2000,
+                    ndraw=8000):
     """
     Parameters
     ----------
@@ -116,8 +118,8 @@ def compute_pvalues(y, X, sigma=1., maxstep=np.inf,
             pval_maxT_identify = FS.model_pivots(i+1, alternative='twosided',
                                           which_var=[FS.variables[-1]],
                                           saturated=False,
-                                          burnin=2000,
-                                          ndraw=8000)[0][1]
+                                          burnin=burnin,
+                                          ndraw=ndraw)[0][1]
         else:
             pval_maxT_identify = np.random.sample()
         var_select, pval_saturated = FS.model_pivots(i+1, alternative='twosided',

### 2015-05-15 code/python/fwdstep_rob.py

@@ -34,11 +34,17 @@ def main():
                         help='Response (CSV filename). REQUIRED')
     parser.add_argument('--sigma',
                         help='Noise standard deviation. REQUIRED', type=float)
+    parser.add_argument('--maxstep',
+                        help='How many steps. Defaults to min(n,p)', type=int)
     parser.add_argument('--active',
                         help='Active set (CSV filename). Necessary ' + 
                         'to compute completion index. REQUIRED. ASSUMES 1-BASED INDEXING!')    
     parser.add_argument('--outfile',
                         help='Where to store output.')
+    parser.add_argument('--burnin',
+                        help='How many burnin steps for sampling.', default=2000, type=int)
+    parser.add_argument('--ndraw',
+                        help='How many sampling steps.', default=8000, type=int)
 
     try: 			
         args = parser.parse_args()
@@ -55,8 +61,13 @@ def main():
 
     full_results = {}
 
+    maxstep = args.maxstep or np.inf
+
     run(Y, X, args.sigma, active,
-        full_results=full_results)
+        full_results=full_results,
+        maxstep=args.maxstep,
+        burnin=args.burnin,
+        ndraw=args.ndraw)
 
     # change output variables to 1-based indices
 
@@ -64,7 +75,8 @@ def main():
         if 'variable_selected' in k:
             full_results[k][0] = full_results[k][0] + 1
 
-    pd.DataFrame(full_results).to_csv(args.outfile, index=False)
+    if args.outfile:
+        pd.DataFrame(full_results).to_csv(args.outfile, index=False)
         
 if __name__ == "__main__":
     main()

### 2015-05-22 notes/nonparam.tex
notes on nonparametric bootstrap

\documentclass{article}

\usepackage[round]{natbib}
\usepackage{amsmath,amssymb,amsthm,bm,enumerate,mathrsfs,mathtools}
\usepackage{latexsym,color,verbatim,multirow}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}

\newcommand{\real}{\mathbb{R}}

\newtheorem{theorem}{Theorem}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}

\begin{document}

\title{Nonparametric selective bootstrap}

\maketitle

\section{Non-bootstrap world}

Suppose we have a model selection procedure $\hat{M}$ acting on sufficient
statistics $Z$ choosing among a (finite) set of models. So, our selected model
is $M=\hat{M}(Z)$. Our selected model suggests a hypothesis of interest that we want to test.

We will apply $\hat{M}$ on data of the form
$$
\hat{Z}_n = \frac{1}{n} \sum_{i=1}^n y_i X_i = \frac{1}{n}X^Ty \in \real^p
$$
where $(y_i, X_i) \overset{\text{IID}}{\sim} F$.

Note that $Z$ is a statistical functional that we might write as
$$
Z(F) = E_{F}(yX)
$$
and what we really observe $\hat{Z}_n = Z(\hat{F}_n)$.

We want to carry out a selective test of
$$
H_0: \eta_{M}^T \gamma_M(F) = 0
$$
where $\eta_M \in \real^{p \times k}$ and $\gamma_M$ is some statistical
function possibly different from $Z$. In principle it can depend
on the model we have selected, but we will drop the $M$ from now on.
{\bf We assume that $Z$ is measurable with respect to $\gamma$, i.e.
$\gamma$ includes $Z$ as part of its coordinates or something. This means
that it makes sense to talk about $\hat{M}$ evaluated
on a sample of $\gamma$'s.}

Above, we have chosen $\eta_M$ based on the outcome $\hat{M}(\hat{Z}_n)$. Following the general setup of the
selected model framework, we just have to design
a test for a fixed $M$.

For large enough $n$, under our model $M$ suppose that
$$
\gamma(\hat{F}_n) \approx N(\gamma(F), n^{-1} \Sigma(F))
$$
where, for some distribution $G$ on $\real \times \real^p$
$$
\Sigma(G) = E_{G}\left((\gamma(\tilde{y},\tilde{X}) - \gamma(G))(\gamma(\tilde{y},\tilde{X}) - \gamma(G))^T \right)
$$
where $(\tilde{y}, \tilde{X})$ is a sample of size 1 from $G$.

{\bf Note: as the functional
we choose may depend on $M$, the normal approximation generally depends 
on the selected model -- I am not going to emphasize that here.}

If this is our reference distribution, we should base our distribution on the 
conditional law of $\Sigma(F)^{-1}W | (I - \eta\eta^{\dagger})\Sigma(F)^{-1}W$ where
$W \sim N(\gamma(F), n^{-1} \Sigma(F))$. This is because
we know that, under $H_0:\eta^T\gamma(F)=0$
this distribution does not depend on $(I-\eta\eta^{\dagger})\gamma(F)$ which
is everything except what we wanted to test.

Set $P_{\eta}^{\perp}=I - \eta\eta^{\dagger}$.
This 
distribution has covariance
$$
n^{-1} \left(\Sigma(F)^{-1} - \Sigma(F)^{-1} P_{\eta}^{\perp} \left(P_{\eta}^{\perp} \Sigma(F)^{-1} P_{\eta}^{\perp} \right)^{\dagger}  P_{\eta}^{\perp} \Sigma(F)^{-1}  \right).
$$
Denote this covariance by $\Theta(\eta, \Sigma(F))$.

If $W \sim N(\gamma(F), n^{-1}\Sigma(F))$,
this distribution can be realized by 
$$
\Sigma(F)^{-1} W - \Sigma(F)^{-1} P_{\eta}^{\perp} \left(P_{\eta}^{\perp} 
\Sigma(F)^{-1} P_{\eta}^{\perp} \right)^{\dagger}  P_{\eta}^{\perp} W.
$$
This is an affine function of $W$ determined by $\eta, \Sigma(F)$ 
call this affine function $A(\eta, \Sigma(F), W)$.

In summary, if $W \sim N(\gamma(F), n^{-1}\Sigma(F))$ 
then under $H_0: \eta^T\gamma(F)=0$, asymptotically we have
$$
A(\eta, \Sigma(F), W) \sim N(0, n^{-1} \Theta(\eta, \Sigma(F))).
$$

{\bf Note that, if we can
sample from this distribution, we know how to
do a parametric selective test of $H_0:\eta_M^T\gamma(F)=0$.
Also, I reiterate that the normal approximation could depend on $M$.
}

\section{Bootstrap test}

We apply non-parametric bootstrap to the pairs 
yielding samples
$$
y^{*,b} \in \real^n,  X^{*,b} \in \real^{n \times p}, \qquad 1 \leq  b \leq B
$$
and  bootstrapped sufficient statistics
$$
\gamma^{*,b} = \frac{1}{n} \gamma(y_i^{*,b}, X_i^{*,b}),  \qquad 1 \leq  b \leq B
$$

These are of course just evaluating $\gamma$ on 
IID draws of size $n$ from $\hat{F}_n$. So,
we can write
$\gamma^{*,b} = \gamma(\hat{F}_n^{*,b})$
where $\hat{F}_n^{*,b}$ is the empirical
distribution function of sampling with replacement from the original data.

For large enough $n$, the law of $\gamma(\hat{F}_n^{*,b})$ is hopefully going to be close
to the Gaussian 
$$
\gamma(\hat{F}_n^{*,b}) | \hat{F}_n \sim N(\gamma(\hat{F}_n), n^{-1} \Sigma(\hat{F}_n))
$$
where
$$
\Sigma(\hat{F}_n) = \frac{1}{n} \sum_{i=1}^n \left(\gamma(y_i,X_i) - \bar{\gamma} \right)\left(\gamma(y_i,X_i) - \bar{\gamma} \right)^T
$$
and
$$
\bar{\gamma} = \frac{1}{n} \sum_{i=1}^n \gamma(y_i, X_i).
$$

(Again, the normal approximation may depend on the selected model.)

Therefore, our unselective bootstrap should be giving samples from
roughly $ N(\gamma(\hat{F}_n), n^{-1} \Sigma(\hat{F}_n))$. That is,
taking $B$ very large:
$$
\frac{1}{B}\sum_{b=1}^B \eta^T \gamma(\hat{F}^{*,b}_n) \approx \eta^T\gamma(\hat{F}_n).
$$
This suggests we correct the unselective bootstrap by drawing
$$
\gamma(\hat{F}^{*,b}_n) - \eta \eta^{\dagger}\gamma(\hat{F}_n) | \hat{F}_n \approx N((I - \eta\eta^{\dagger})\gamma(\hat{F}_n) ,n^{-1} \Sigma(\hat{F}_n))
$$
and we see that
$$
A(\eta, \Sigma(\hat{F}_n), \gamma(\hat{F}^{*,b}_n) - \eta \eta^{\dagger} \gamma(\hat{F}_n)) | \hat{F}_n
 \approx N(0 ,n^{-1} \Theta(\eta, \Sigma(\hat{F}_n))).
$$
As $n$ gets larger, $\Sigma(\hat{F}_n)$ will approach $\Sigma(F)$. In general,
plugging in consistent estimates of variance is OK in selective inference (not sure
if Xiaoying put it in the first paper, but definitely in the
randomized setting we have a concrete statement about this). In Rob's example of $n=30$, the sample size might be too small still, but asymptotically
I think it will be fine. Maybe a degrees of freedom correction will help.

So,
we draw nonparametric bootstrap samples
$$
\gamma(\hat{F}^{*,b}_n) - \eta \eta^{\dagger}\gamma(\hat{F}_n)
$$
keeping only those for which $\hat{M}(\gamma(\hat{F}^{*,b}_n) - \eta \eta^{\dagger}\gamma(\hat{F}_n)) = \hat{M}(\gamma(\hat{F}_n))$.
For those that survive, we compute
$$
A(\eta, \Sigma(\hat{F}_n), \gamma(\hat{F}^{*,b}_n) - \eta \eta^{\dagger}\gamma(\hat{F}_n))
$$
and take a test statistic of our choosing.

\section{No nuisance parameters}

When there are no nuisance parameters (i.e. $\eta=I$ or we are make other assumptions about $Z(F)$), we do not need the
function $A$. This is the example I tried this morning. In this
case, $\gamma=Z$ and we just draw nonparametric bootstrap samples
$$
Z(\hat{F}^{*,b}_n) - \eta \eta^{\dagger}Z(\hat{F}_n)
$$
keeping the same ones and compute any test statistic of our 
choosing on these bootstrap samples.

\end{document}

### 2015-05-22 notes/nonparam.tex
added a little on nonparametric bootstrap

@@
 $$
-\hat{Z}_n = \frac{1}{n} \sum_{i=1}^n y_i X_i = \frac{1}{n}X^Ty \in \real^p
+\hat{Z}_n = \frac{1}{n} \sum_{i=1}^n Z(t_i) 
 $$
-where $(y_i, X_i) \overset{\text{IID}}{\sim} F$.
+where $t_i \overset{\text{IID}}{\sim} F$ where $F$
+is some distribution on some sample space ${\cal X}$. In the pairs bootstrap for 
+regression context, we should think of $t_i=(y_i, X_i)$ and $F$ 
+is a distribution on $\real \times \real^p$ and, if we use
+forward stepwise our
+selection is a function of $\frac{1}{n}(X^Ty, X^TX)$
+and our $Z$ should be taken to be
+$$
+Z(t_i) = (y_iX_i, X_iX_i^T).
+$$
 
@@
 $$
-Z(F) = E_{F}(yX)
+Z(F) = E_{F}(Z(\tilde{y}, \tilde{X}))
 $$
@@
 on the model we have selected, but we will drop the $M$ from now on.
+That is, we are now looking at
+$$
+\gamma_M(\hat{F}_n) = \sum_{i=1}^n \gamma(t_i)
+$$
+where $t_i \overset{IID}{\sim} F$.
+
 {\bf We assume that $Z$ is measurable with respect to $\gamma$, i.e.
@@
 
-Above, we have chosen $\eta_M$ based on the outcome $\hat{M}(\hat{Z}_n)$. Following the general setup of the
+\subsection{One sample problem, mean or median?}
+
+Suppose we have a one-sample problem and if the empirical skewness
+is larger than some threshold, we test whether the median of the 
+distribution, otherwise we report the mean. In one case,
+$\gamma_M$ would compute sample skewness and sample median, in the other
+it would compute the sample skewness and sample mean. When choosing to test the sample mean, I suppose the model is all distributions on $\real$ for
+which the sample mean and sample skewness and sample mean 
+are asymptotically Gaussian. When
+reporting the sample median, the model is all distributions on $\real$
+for which the sample skewness and sample median are asymptotically jointly Gaussian.
+
+\subsection{Hypothesis tests under unselective distribution}
+
+We have chosen $\eta_M$ based on the outcome $\hat{M}(\hat{Z}_n)$. Following the general setup of the
 selected model framework, we just have to design
@@
 $$
-\Sigma(G) = E_{G}\left((\gamma(\tilde{y},\tilde{X}) - \gamma(G))(\gamma(\tilde{y},\tilde{X}) - \gamma(G))^T \right)
+\Sigma(F) = E_{F}\left((\gamma(\tilde{t}) - \gamma(F))(\gamma(\tilde{t}) - \gamma(F))^T \right)
 $$
-where $(\tilde{y}, \tilde{X})$ is a sample of size 1 from $G$.
+where $\tilde{t}$ is a sample of size 1 from $F$.
 
@@
 we choose may depend on $M$, the normal approximation generally depends 
-on the selected model -- I am not going to emphasize that here.}
+on the selected model -- I am not always going to emphasize that here.}
 
 If this is our reference distribution, we should base our distribution on the 
-conditional law of $\Sigma(F)^{-1}W | (I - \eta\eta^{\dagger})\Sigma(F)^{-1}W$ where
-$W \sim N(\gamma(F), n^{-1} \Sigma(F))$. This is because
+conditional law 
+$$\Sigma(F)^{-1}\gamma(\hat{F}_n) \big \vert P_{\eta}^{\perp}\Sigma(F)^{-1}\gamma(\hat{F}_n).$$ 
+where $P_{\eta}=\eta\eta^{\dagger}$ and $P_{\eta}^{\perp} = I - P_{\eta}$.
+This is because
 we know that, under $H_0:\eta^T\gamma(F)=0$
-this distribution does not depend on $(I-\eta\eta^{\dagger})\gamma(F)$ which
+this distribution does not depend on $P_{\eta}^{\perp}\gamma(F)$ which
 is everything except what we wanted to test.
 
-Set $P_{\eta}^{\perp}=I - \eta\eta^{\dagger}$.
 This 
-distribution has covariance
+distribution has covariance $n^{-1} \Theta(\eta, \Sigma(F))$ where
 $$
-n^{-1} \left(\Sigma(F)^{-1} - \Sigma(F)^{-1} P_{\eta}^{\perp} \left(P_{\eta}^{\perp} \Sigma(F)^{-1} P_{\eta}^{\perp} \right)^{\dagger}  P_{\eta}^{\perp} \Sigma(F)^{-1}  \right).
+\Theta(\eta, \Sigma(F)) = \left(\Sigma(F)^{-1} - \Sigma(F)^{-1} P_{\eta}^{\perp} \left(P_{\eta}^{\perp} \Sigma(F)^{-1} P_{\eta}^{\perp} \right)^{\dagger}  P_{\eta}^{\perp} \Sigma(F)^{-1}  \right).
 $$
-Denote this covariance by $\Theta(\eta, \Sigma(F))$.
+Under $H_0$, this distribution has mean 0.
 
-If $W \sim N(\gamma(F), n^{-1}\Sigma(F))$,
+If $W \sim N(\gamma(F), n^{-1}\Sigma(F))$, then, under $H_0$
 this distribution can be realized by 
 $$
-\Sigma(F)^{-1} W - \Sigma(F)^{-1} P_{\eta}^{\perp} \left(P_{\eta}^{\perp} 
+L(\eta, \Sigma(F), W) = \Sigma(F)^{-1} W - \Sigma(F)^{-1} P_{\eta}^{\perp} \left(P_{\eta}^{\perp} 
 \Sigma(F)^{-1} P_{\eta}^{\perp} \right)^{\dagger}  P_{\eta}^{\perp} W.
 $$
-This is an affine function of $W$ determined by $\eta, \Sigma(F)$ 
-call this affine function $A(\eta, \Sigma(F), W)$.
 
 In summary, if $W \sim N(\gamma(F), n^{-1}\Sigma(F))$ 
-then under $H_0: \eta^T\gamma(F)=0$, asymptotically we have
+then under $H_0: \eta^T\gamma(F)=0$ we have
 $$
-A(\eta, \Sigma(F), W) \sim N(0, n^{-1} \Theta(\eta, \Sigma(F))).
+L(\eta, \Sigma(F), W) \sim N(0, n^{-1} \Theta(\eta, \Sigma(F))).
 $$
 
-{\bf Note that, if we can
+Therefore, if we can
 sample from this distribution, we know how to
-do a parametric selective test of $H_0:\eta_M^T\gamma(F)=0$.
-Also, I reiterate that the normal approximation could depend on $M$.
+do a parametric unselective test of $H_0:\eta_M^T\gamma(F)=0$.
+
+\subsection{Selective test}
+\label{sec:selective}
+
+The parametric selective test is based on the distribution of 
+$$
+L(\eta, \Sigma(F), \gamma(\hat{F}_n)) \big \vert \hat{M}(\gamma(\hat{F}_n)) = M.
+$$
+To emphasize that this depends on the selected model, we might write
+$$
+L_M(\eta_M, \Sigma_M(F), \gamma_M(\hat{F}_n)) \big \vert \hat{M}(\gamma_M(\hat{F}_n)) = M.
+$$
+
+If $W \sim N(\mu, \Sigma_M(F))$ is any normal distribution
+such that $\eta^T\mu=0$ then by familiar exponential family principles, the above law
+is equal to the law of
+$$
+L_M(\eta_M, \Sigma_M(F), W) \big \vert \hat{M}(W) = M.
+$$
+
+{\bf This means it is enough to find any normal distribution
+$W \sim N(\mu, \Sigma_M(F))$ such that $\eta^T\mu=0$ to carry out the
+selective test.
 }
 
-\section{Bootstrap test}
+\section{Bootstrap world}
 
@@
 $$
-y^{*,b} \in \real^n,  X^{*,b} \in \real^{n \times p}, \qquad 1 \leq  b \leq B
+t^{*,b} \in {\cal X}^n, \qquad 1 \leq  b \leq B
 $$
@@
 $$
-\gamma^{*,b} = \frac{1}{n} \gamma(y_i^{*,b}, X_i^{*,b}),  \qquad 1 \leq  b \leq B
+\gamma^{*,b} = \frac{1}{n} \sum_{i=1}^n \gamma(t_i^{*,b}),  \qquad 1 \leq  b \leq B
 $$
@@
 
-For large enough $n$, the law of $\gamma(\hat{F}_n^{*,b})$ is hopefully going to be close
+For large enough $n$, the law of $\gamma(\hat{F}_n^{*,b})$ is going to be converging
 to the Gaussian 
@@
 $$
-\Sigma(\hat{F}_n) = \frac{1}{n} \sum_{i=1}^n \left(\gamma(y_i,X_i) - \bar{\gamma} \right)\left(\gamma(y_i,X_i) - \bar{\gamma} \right)^T
+\Sigma(\hat{F}_n) = \frac{1}{n} \sum_{i=1}^n \left(\gamma(t_i) - \gamma(\hat{F}_n) \right)\left(\gamma(t_i) - \gamma(\hat{F}_n) \right)^T.
 $$
-and
-$$
-\bar{\gamma} = \frac{1}{n} \sum_{i=1}^n \gamma(y_i, X_i).
-$$
-
-(Again, the normal approximation may depend on the selected model.)
 
@@
 $$
-\gamma(\hat{F}^{*,b}_n) - \eta \eta^{\dagger}\gamma(\hat{F}_n) | \hat{F}_n \approx N((I - \eta\eta^{\dagger})\gamma(\hat{F}_n) ,n^{-1} \Sigma(\hat{F}_n))
+\gamma(\hat{F}^{*,b}_n) - P_{\eta}\gamma(\hat{F}_n) | \hat{F}_n \approx N(P_{\eta}^{\perp}\gamma(\hat{F}_n) ,n^{-1} \Sigma(\hat{F}_n))
 $$
@@
 $$
-A(\eta, \Sigma(\hat{F}_n), \gamma(\hat{F}^{*,b}_n) - \eta \eta^{\dagger} \gamma(\hat{F}_n)) | \hat{F}_n
+L(\eta, \Sigma(\hat{F}_n), \gamma(\hat{F}^{*,b}_n) - P_{\eta} \gamma(\hat{F}_n)) \big \vert \hat{F}_n
  \approx N(0 ,n^{-1} \Theta(\eta, \Sigma(\hat{F}_n))).
 $$
-As $n$ gets larger, $\Sigma(\hat{F}_n)$ will approach $\Sigma(F)$. In general,
+As $n$ gets larger, $\Sigma(\hat{F}_n)$ will approach $\Sigma(F)$. 
+If the RHS above were actually $$N(0 ,n^{-1} \Theta(\eta, \Sigma(\hat{F}_n))),$$ then
+by our earlier observation at the end of Section \ref{sec:selective} this is sufficient to carry out the selective test.
+However, the variance is not exactly correct. In general,
 plugging in consistent estimates of variance is OK in selective inference (not sure
-if Xiaoying put it in the first paper, but definitely in the
-randomized setting we have a concrete statement about this). In Rob's example of $n=30$, the sample size might be too small still, but asymptotically
+if Xiaoying and I put it in the first paper on asymptotics, but definitely in the
+randomized setting we will have a concrete statement about this). In Rob's example of $n=30$, the sample size might be too small still, but asymptotically
 I think it will be fine. Maybe a degrees of freedom correction will help.
@@
 $$
-\gamma(\hat{F}^{*,b}_n) - \eta \eta^{\dagger}\gamma(\hat{F}_n)
+\gamma(\hat{F}^{*,b}_n) - P_{\eta}\gamma(\hat{F}_n)
 $$
-keeping only those for which $\hat{M}(\gamma(\hat{F}^{*,b}_n) - \eta \eta^{\dagger}\gamma(\hat{F}_n)) = \hat{M}(\gamma(\hat{F}_n))$.
+keeping only those for which $\hat{M}(\gamma(\hat{F}^{*,b}_n) - P_{\eta}\gamma(\hat{F}_n)) = \hat{M}(\gamma(\hat{F}_n))$.
+
 For those that survive, we compute
 $$
-A(\eta, \Sigma(\hat{F}_n), \gamma(\hat{F}^{*,b}_n) - \eta \eta^{\dagger}\gamma(\hat{F}_n))
+L(\eta, \Sigma(\hat{F}_n), \gamma(\hat{F}^{*,b}_n) - P_{\eta}\gamma(\hat{F}_n))
 $$
@@
 When there are no nuisance parameters (i.e. $\eta=I$ or we are make other assumptions about $Z(F)$), we do not need the
-function $A$. This is the example I tried this morning. In this
+function $L$. This is the example I tried this morning. In this
 case, $\gamma=Z$ and we just draw nonparametric bootstrap samples
 $$
-Z(\hat{F}^{*,b}_n) - \eta \eta^{\dagger}Z(\hat{F}_n)
+Z(\hat{F}^{*,b}_n) - P_{\eta}Z(\hat{F}_n)
 $$
-keeping the same ones and compute any test statistic of our 
+keeping the ones that satisfy selection and compute any test statistic of our 
 choosing on these bootstrap samples.

### 2015-05-22 notes/nonparam.tex
a little more cleanup on nonparametric notes

@@
 forward stepwise our
-selection is a function of $\frac{1}{n}(X^Ty, X^TX)$
+selection is a function of 
+$$\frac{1}{n}(X^Ty, X^TX)$$
 and our $Z$ should be taken to be
@@
 $$
-Z(F) = E_{F}(Z(\tilde{y}, \tilde{X}))
+Z(F) = E_{F}(Z(t))
 $$
@@
 
-{\bf We assume that $Z$ is measurable with respect to $\gamma$, i.e.
+We assume that $Z$ is measurable with respect to $\gamma$, i.e.
 $\gamma$ includes $Z$ as part of its coordinates or something. This means
 that it makes sense to talk about $\hat{M}$ evaluated
-on a sample of $\gamma$'s.}
+on a sample of $\gamma$'s.
 
@@
 
+Here is an ``example.''
 Suppose we have a one-sample problem and if the empirical skewness
@@
 
-We have chosen $\eta_M$ based on the outcome $\hat{M}(\hat{Z}_n)$. Following the general setup of the
-selected model framework, we just have to design
-a test for a fixed $M$.
-
 For large enough $n$, under our model $M$ suppose that
@@
 $$
-where, for some distribution $G$ on $\real \times \real^p$
+where
 $$
-\Sigma(F) = E_{F}\left((\gamma(\tilde{t}) - \gamma(F))(\gamma(\tilde{t}) - \gamma(F))^T \right)
+\Sigma(F) = E_{F}\left((\gamma(t) - \gamma(F))(\gamma(t) - \gamma(F))^T \right).
 $$
-where $\tilde{t}$ is a sample of size 1 from $F$.
 
@@
 
-If this is our reference distribution, we should base our distribution on the 
+If this is our reference distribution, we should base our test on the 
 conditional law 
-$$\Sigma(F)^{-1}\gamma(\hat{F}_n) \big \vert P_{\eta}^{\perp}\Sigma(F)^{-1}\gamma(\hat{F}_n).$$ 
+$$n^{1/2}\Sigma(F)^{-1}\gamma(\hat{F}_n) \big \vert n^{1/2} P_{\eta}^{\perp}\Sigma(F)^{-1}\gamma(\hat{F}_n).$$ 
 where $P_{\eta}=\eta\eta^{\dagger}$ and $P_{\eta}^{\perp} = I - P_{\eta}$.
@@
 This 
-distribution has covariance $n^{-1} \Theta(\eta, \Sigma(F))$ where
+law has covariance $n^{-1} \Theta(\eta, \Sigma(F))$ where
 $$
-\Theta(\eta, \Sigma(F)) = \left(\Sigma(F)^{-1} - \Sigma(F)^{-1} P_{\eta}^{\perp} \left(P_{\eta}^{\perp} \Sigma(F)^{-1} P_{\eta}^{\perp} \right)^{\dagger}  P_{\eta}^{\perp} \Sigma(F)^{-1}  \right).
+\Theta(\eta, \Sigma(F)) = \left(\Sigma(F)^{-1} - \Sigma(F)^{-1} P_{\eta}^{\perp} \left(P_{\eta}^{\perp} \Sigma(F)^{-1} P_{\eta}^{\perp} \right)^{\dagger}  P_{\eta}^{\perp} \Sigma(F)^{-1}  \right)
 $$
-Under $H_0$, this distribution has mean 0.
+which is an $O(1)$ covariance matrix.
+Under $H_0$, this law has mean 0. Under any alternative, its mean grows like $n^{1/2}$.
 
-If $W \sim N(\gamma(F), n^{-1}\Sigma(F))$, then, under $H_0$
+Under $H_0$
 this distribution can be realized by 
 $$
-L(\eta, \Sigma(F), W) = \Sigma(F)^{-1} W - \Sigma(F)^{-1} P_{\eta}^{\perp} \left(P_{\eta}^{\perp} 
-\Sigma(F)^{-1} P_{\eta}^{\perp} \right)^{\dagger}  P_{\eta}^{\perp} W.
+L(\eta, \Sigma(F), W) = n^{1/2} \left(\Sigma(F)^{-1} W - \Sigma(F)^{-1} P_{\eta}^{\perp} \left(P_{\eta}^{\perp} 
+\Sigma(F)^{-1} P_{\eta}^{\perp} \right)^{\dagger}  P_{\eta}^{\perp} W \right)
 $$
+where $W \sim N(\mu, n^{-1}\Sigma(F))$, is any normal distribution with
+$\eta^T\mu=0$.
 
-In summary, if $W \sim N(\gamma(F), n^{-1}\Sigma(F))$ 
-then under $H_0: \eta^T\gamma(F)=0$ we have
+In other words, if $W \sim N(\mu, n^{-1}\Sigma(F))$ 
+then under $H_0: \eta^T\gamma(F)=0$ we have, asymptotically
 $$
-L(\eta, \Sigma(F), W) \sim N(0, n^{-1} \Theta(\eta, \Sigma(F))).
+ L(\eta, \Sigma(F), \gamma(\hat{F}_n))  \overset{n \rightarrow \infty}{\Rightarrow} L(\eta, \Sigma(F), W) = N(0,  \Theta(\eta, \Sigma(F))).
 $$
@@
 sample from this distribution, we know how to
-do a parametric unselective test of $H_0:\eta_M^T\gamma(F)=0$.
+do an asymptotic unselective test of $H_0:\eta_M^T\gamma(F)=0$ using the 
+normal approximation.
 
@@
 \label{sec:selective}
+
+Recall that we have chosen $\eta_M$ based on the outcome $\hat{M}(\hat{Z}_n)$. Following the general setup of the
+selected model framework, we just have to design
+a test for a fixed $M$.
 
@@
 
-If $W \sim N(\mu, \Sigma_M(F))$ is any normal distribution
+Now, if $W \sim N(\mu, n^{-1}\Sigma(F))$ is any normal distribution
 such that $\eta^T\mu=0$ then by familiar exponential family principles, the above law
@@
 $$
-L_M(\eta_M, \Sigma_M(F), W) \big \vert \hat{M}(W) = M.
+L_M(\eta_M, n^{-1}\Sigma(F), W) \big \vert \hat{M}(W) = M.
 $$
@@
 {\bf This means it is enough to find any normal distribution
-$W \sim N(\mu, \Sigma_M(F))$ such that $\eta^T\mu=0$ to carry out the
+$W \sim N(\mu, n^{-1}\Sigma(F))$ such that $\eta^T\mu=0$ to carry out the
 selective test.
@@
 $$
-\gamma(\hat{F}^{*,b}_n) - P_{\eta}\gamma(\hat{F}_n) | \hat{F}_n \approx N(P_{\eta}^{\perp}\gamma(\hat{F}_n) ,n^{-1} \Sigma(\hat{F}_n))
+\gamma(\hat{F}^{*,b}_n) - P_{\eta}\gamma(\hat{F}_n) | \hat{F}_n \approx N(P_{\eta}^{\perp}\gamma(\hat{F}_n) ,n^{-1} \Sigma(\hat{F}_n)).
 $$
-and we see that
+With this correction, we see that
 $$
@@
 As $n$ gets larger, $\Sigma(\hat{F}_n)$ will approach $\Sigma(F)$. 
-If the RHS above were actually $$N(0 ,n^{-1} \Theta(\eta, \Sigma(\hat{F}_n))),$$ then
-by our earlier observation at the end of Section \ref{sec:selective} this is sufficient to carry out the selective test.
+If the RHS above were actually $$N(0 ,n^{-1} \Theta(\eta, \Sigma(F))),$$ then
+by our earlier observation at the end of Section \ref{sec:selective} this is sufficient to carry out the selective test exactly.
 However, the variance is not exactly correct. In general,
 plugging in consistent estimates of variance is OK in selective inference (not sure
-if Xiaoying and I put it in the first paper on asymptotics, but definitely in the
-randomized setting we will have a concrete statement about this). In Rob's example of $n=30$, the sample size might be too small still, but asymptotically
+if Xiaoying and I put it in the first paper on asymptotics, but she has given me a proof and definitely in the
+randomized paper we will have a concrete statement about this). In Rob's example of $n=30$, the sample size might be too small still, but asymptotically
 I think it will be fine. Maybe a degrees of freedom correction will help.
 
-So,
+The final algorithm is:
 we draw nonparametric bootstrap samples
@@
 $$
-and take a test statistic of our choosing.
+and take any test statistic of our choosing.
 

### 2015-05-27 notes/nonparam.tex
pairs version of maxT

@@
 \newcommand{\real}{\mathbb{R}}
+\usepackage{mycommands}
 
@@
 
-For large enough $n$, under our model $M$ suppose that
+We will follow the convention of \cite{penn_pairs} in stating
+a bootstrap CLT as an $m$-of-$n$ bootstrap with $m \gg n$.
+For large enough $m$, under our model $M$ suppose that
 $$
-\gamma(\hat{F}_n) \approx N(\gamma(F), n^{-1} \Sigma(F))
+\gamma(\hat{F}_m) \approx N(\gamma(F), m^{-1} \Sigma(F))
 $$
@@
 $$
-\Sigma(F) = E_{F}\left((\gamma(t) - \gamma(F))(\gamma(t) - \gamma(F))^T \right).
+\Sigma(F) = \Sigma(F,\gamma) = E_{F}\left((\gamma(t) - \gamma(F))(\gamma(t) - \gamma(F))^T \right).
+$$
+
+The bootstrap version of this result will be denoted by
+$$
+\gamma(\hat{F}^{*,m}_n) \overset{m \to \infty}{\Rightarrow} N(\gamma(\hat{F}_n), m^{-1} \Sigma(\hat{F_n})).
+$$
+That is, if we take $m \gg n$ samples with replacement from $\hat{F}_n$, 
+a CLT holds in $m$. The statement ``the bootstrap works'' can then
+be interpreted as 
+$$
+\gamma(\hat{F}^{*,n}_n) \overset{n \to \infty}{\Rightarrow} N(\gamma(F), n^{-1} \Sigma(F)).
 $$
@@
 conditional law 
-$$n^{1/2}\Sigma(F)^{-1}\gamma(\hat{F}_n) \big \vert n^{1/2} P_{\eta}^{\perp}\Sigma(F)^{-1}\gamma(\hat{F}_n).$$ 
+$$m^{1/2}\Sigma(F)^{-1}\gamma(\hat{F}_m) \big \vert m^{1/2} P_{\eta}^{\perp}\Sigma(F)^{-1}\gamma(\hat{F}_m).$$ 
 where $P_{\eta}=\eta\eta^{\dagger}$ and $P_{\eta}^{\perp} = I - P_{\eta}$.
@@
 This 
-law has covariance $n^{-1} \Theta(\eta, \Sigma(F))$ where
+law has covariance $m^{-1} \Theta(\eta, \Sigma(F))$ where
 $$
@@
 which is an $O(1)$ covariance matrix.
-Under $H_0$, this law has mean 0. Under any alternative, its mean grows like $n^{1/2}$.
+Under $H_0$, this law has mean 0. Under any alternative, its mean grows like $m^{1/2}$.
 
@@
 $$
-L(\eta, \Sigma(F), W) = n^{1/2} \left(\Sigma(F)^{-1} W - \Sigma(F)^{-1} P_{\eta}^{\perp} \left(P_{\eta}^{\perp} 
+L(\eta, \Sigma(F), W) = m^{1/2} \left(\Sigma(F)^{-1} W - \Sigma(F)^{-1} P_{\eta}^{\perp} \left(P_{\eta}^{\perp} 
 \Sigma(F)^{-1} P_{\eta}^{\perp} \right)^{\dagger}  P_{\eta}^{\perp} W \right)
 $$
-where $W \sim N(\mu, n^{-1}\Sigma(F))$, is any normal distribution with
+where $W \sim N(\mu, m^{-1}\Sigma(F))$, is any normal distribution with
 $\eta^T\mu=0$.
 
-In other words, if $W \sim N(\mu, n^{-1}\Sigma(F))$ 
+In other words, if $W_m \sim N(\mu, m^{-1}\Sigma(F))$ 
 then under $H_0: \eta^T\gamma(F)=0$ we have, asymptotically
 $$
- L(\eta, \Sigma(F), \gamma(\hat{F}_n))  \overset{n \rightarrow \infty}{\Rightarrow} L(\eta, \Sigma(F), W) = N(0,  \Theta(\eta, \Sigma(F))).
+ L(\eta, \Sigma(F), \gamma(\hat{F}_m))  \overset{m \rightarrow \infty}{\Rightarrow} L(\eta, \Sigma(F), W_m) = N(0,  \Theta(\eta, \Sigma(F))).
 $$
@@
 
-We apply non-parametric bootstrap to the pairs 
-yielding samples
+We apply non-parametric bootstrap drawing $m$ samples
+yielding 
 $$
-t^{*,b} \in {\cal X}^n, \qquad 1 \leq  b \leq B
+t^{*,b} \in {\cal X}^m, \qquad 1 \leq  b \leq B
 $$
@@
 $$
-\gamma^{*,b} = \frac{1}{n} \sum_{i=1}^n \gamma(t_i^{*,b}),  \qquad 1 \leq  b \leq B
+\gamma^{*,b} = \frac{1}{m} \sum_{i=1}^m \gamma(t_i^{*,b}),  \qquad 1 \leq  b \leq B
 $$
@@
 These are of course just evaluating $\gamma$ on 
-IID draws of size $n$ from $\hat{F}_n$. So,
+IID draws of size $m$ from $\hat{F}_n$. So,
 we can write
-$\gamma^{*,b} = \gamma(\hat{F}_n^{*,b})$
-where $\hat{F}_n^{*,b}$ is the empirical
+$\gamma^{*,b} = \gamma(\hat{F}_n^{*,m})$
+where $\hat{F}_n^{*,m}$ is the empirical
 distribution function of sampling with replacement from the original data.
 
-For large enough $n$, the law of $\gamma(\hat{F}_n^{*,b})$ is going to be converging
+For large enough $m$, the law of $\gamma(\hat{F}_n^{*,m})$ is going to be converging
 to the Gaussian 
 $$
-\gamma(\hat{F}_n^{*,b}) | \hat{F}_n \sim N(\gamma(\hat{F}_n), n^{-1} \Sigma(\hat{F}_n))
-$$
-where
-$$
-\Sigma(\hat{F}_n) = \frac{1}{n} \sum_{i=1}^n \left(\gamma(t_i) - \gamma(\hat{F}_n) \right)\left(\gamma(t_i) - \gamma(\hat{F}_n) \right)^T.
+\gamma(\hat{F}_n^{*,m}) | \hat{F}_n \sim N(\gamma(\hat{F}_n), m^{-1} \Sigma(\hat{F}_n)).
 $$
@@
 Therefore, our unselective bootstrap should be giving samples from
-roughly $ N(\gamma(\hat{F}_n), n^{-1} \Sigma(\hat{F}_n))$. That is,
+roughly $ N(\gamma(\hat{F}_n), m^{-1} \Sigma(\hat{F}_n))$. That is,
 taking $B$ very large:
 $$
-\frac{1}{B}\sum_{b=1}^B \eta^T \gamma(\hat{F}^{*,b}_n) \approx \eta^T\gamma(\hat{F}_n).
+\frac{1}{B}\sum_{b=1}^B \eta^T \gamma(\hat{F}^{*,m}_n) \approx \eta^T\gamma(\hat{F}_n).
 $$
@@
 $$
-\gamma(\hat{F}^{*,b}_n) - P_{\eta}\gamma(\hat{F}_n) | \hat{F}_n \approx N(P_{\eta}^{\perp}\gamma(\hat{F}_n) ,n^{-1} \Sigma(\hat{F}_n)).
+\gamma(\hat{F}^{*,m}_n) - P_{\eta}\gamma(\hat{F}_n) | \hat{F}_n \approx N(P_{\eta}^{\perp}\gamma(\hat{F}_n) ,m^{-1} \Sigma(\hat{F}_n)).
 $$
@@
 $$
-L(\eta, \Sigma(\hat{F}_n), \gamma(\hat{F}^{*,b}_n) - P_{\eta} \gamma(\hat{F}_n)) \big \vert \hat{F}_n
- \approx N(0 ,n^{-1} \Theta(\eta, \Sigma(\hat{F}_n))).
+L(\eta, \Sigma(\hat{F}_n), \gamma(\hat{F}^{*,m}_n) - P_{\eta} \gamma(\hat{F}_n)) \big \vert \hat{F}_n
+ \approx N(0 ,m^{-1} \Theta(\eta, \Sigma(\hat{F}_n))).
 $$
-As $n$ gets larger, $\Sigma(\hat{F}_n)$ will approach $\Sigma(F)$. 
-If the RHS above were actually $$N(0 ,n^{-1} \Theta(\eta, \Sigma(F))),$$ then
+
+Now, as $n$ gets larger, $\Sigma(\hat{F}_n)$ will approach $\Sigma(F)$. 
+If the RHS above were actually $$N(0 ,m^{-1} \Theta(\eta, \Sigma(F))),$$ then
 by our earlier observation at the end of Section \ref{sec:selective} this is sufficient to carry out the selective test exactly.
@@
 $$
-\gamma(\hat{F}^{*,b}_n) - P_{\eta}\gamma(\hat{F}_n)
+\gamma(\hat{F}^{*,m}_n) - P_{\eta}\gamma(\hat{F}_n)
 $$
-keeping only those for which $\hat{M}(\gamma(\hat{F}^{*,b}_n) - P_{\eta}\gamma(\hat{F}_n)) = \hat{M}(\gamma(\hat{F}_n))$.
+keeping only those for which $\hat{M}(\gamma(\hat{F}^{*,m}_n) - P_{\eta}\gamma(\hat{F}_n)) = \hat{M}(\gamma(\hat{F}_n))$.
 
@@
 $$
-L(\eta, \Sigma(\hat{F}_n), \gamma(\hat{F}^{*,b}_n) - P_{\eta}\gamma(\hat{F}_n))
+L(\eta, \Sigma(\hat{F}_n), \gamma(\hat{F}^{*,m}_n) - P_{\eta}\gamma(\hat{F}_n))
 $$
@@
 $$
-Z(\hat{F}^{*,b}_n) - P_{\eta}Z(\hat{F}_n)
+Z(\hat{F}^{*,m}_n) - P_{\eta}Z(\hat{F}_n)
 $$
@@
 
+\section{Pairs bootstrap regression}
+
+A lot of this is based on the excellent survey \cite{penn_pairs}, though 
+I want to turn it into selected model language. In this section,
+$X$ is a random vector on $\real^p$, not a matrix, and $y$ is a 
+random variable, not a vector. 
+Given a distribution $F$ on $\real \times \real^p$, there are various
+quantities that appear in the ``pairs bootstrap'' for regression.
+(We could assume that we have removed a constant mean
+from both $X$ and $y$, so that 
+$\E_F((y,X)) = 0$ though nothing strictly assumes this unless
+we want to talk about ``covariances'' and ``correlations'' instead
+ of ``inner productes'' and ``angles''.)
+
+In any case, least squares regression is a statistical functional
+$$
+\beta(F) = \argmin_{\beta \in \real^P} \E_F((y-X^T\beta)^2)
+$$
+where $(y,X) \sim F$.
+This leads us to the definition of the total error, a random vector 
+defined in \cite{penn_pairs}
+$$
+\delta = \delta(F) = y - X^T\beta,  \qquad (y, X) \sim F.
+$$
+They further decompose this error into two components
+$$
+\delta(F) = \epsilon(F) + \zeta(F) = (y - \E_F(y|X)) + (\E_F(y|X) - X^T\beta).
+$$
+(I don't want to overload $\eta$, so I used $\zeta$ in place of their $\eta$).
+
+Note that we can rewrite the error equation as
+\begin{equation}
+\label{eq:datagen:full}
+y = X^T\beta(F) +\delta(F), \qquad (y, X) \sim F.
+\end{equation}
+This says that, in order to draw from $F$ it is sufficient to know 
+$\beta(F)$ and draw from $(X, \delta(F))$. This formula
+has a more familiar ``data-generating'' mechanism to it.
+
+One of the main takeaways of\cite{penn_pairs} is that the $m$-of-$n$ pairs bootstrap distribution of $\beta(\hat{F}_n)$ for any fixed $n$ (i.e. drawing $m \gg n$ samples with replacement
+from $\hat{F}_n$ and running OLS on this data set) 
+has the following CLT
+$$
+m^{1/2} \left(\beta(\hat{F}^{*,m}_n) - \beta(\hat{F}_n) \right)
+\overset{M\to \infty}{\Rightarrow} N(0, S(\hat{F}_n))
+$$
+where 
+$$
+S(F) = (\E_F(XX^T))^{-1} \E_F(\delta(F)XX^T)(\E_F(XX^T))^{-1}
+$$
+is the sandwich estimator for the law $F$ and 
+$\hat{F}^{*,m}_n$ is a sample of size $M$ with replacement from $\hat{F}_n$.
+
+The fact that the pairs bootstrap works as $n \to \infty$ can therefore be summarized as saying that
+$$
+S(\hat{F}_n) \overset{n \rightarrow \infty}{\to} S(F).
+$$
+
+\subsection{Selected model tests}
+
+We are of course interested in selective versions of these statements.
+Let $A \subset [p]=\{1, \dots, p\}$ and $I=A^c$ partition
+the variables into active and inactive sets.
+Then, we can write
+\begin{equation}
+\label{eq:datagen}
+y = X_A\beta_A(F) +\delta_A(F), \qquad (y, X) \sim F
+\end{equation}
+which makes sense whether or not $\beta(F)[I]=0$ or not. Note that
+we should rewrite \eqref{eq:datagen:full} as
+\begin{equation}
+\label{eq:datagen}
+y = X\beta_{[p]}(F) +\delta_{[p]}(F), \qquad (y, X) \sim F
+\end{equation}
+
+What happens if $\beta(F)_I=0$? In this case,
+$$
+X^T\beta_{[p]}(F) = X_A^T\beta(F)_A.
+$$
+However, it is not immediately clear 
+$$
+\beta_{[p]}(F)_A=\beta_A(F).
+$$
+For this to be the case, it would seem to have to be that
+$$
+\delta_A(F)=\delta_{[p]}(F)
+$$
+which we have no reason to believe is true in general. In fact, I think
+we have
+$$
+\beta_{[p]}(F)_A = \beta_A(F), \beta_{[p]}(F)_I=0 \iff \delta_A(F)=\delta_{[p]}(F).
+$$
+
+\subsection{Analog of max-$T$}
+
+In the fixed design case, our max-$T$ test looks at the law of
+$yX_I$ adjusted for those variables in $A$. In the
+fixed design case, these adjustments are done
+in matricial form which we are not using in this section.
+Morally, though, they use the residuals after having adjusted for $A$
+and see if there are any other important correlations remaining.
+
+A natural analog
+in the pairs bootstrap model is to look at the law of
+$$
+\begin{aligned}
+\delta_A(F) X_I &= (y-X_A^T\beta_A(F)) X_I \\
+&= yX_I - X_I \cdot X_A^T\beta_A(F)
+\end{aligned}
+$$
+and perhaps test $H_0:\E_F(\delta_A(F)X_I)=0$.
+From our earlier discussion we see that
+$$
+\begin{aligned}
+\delta_A(F)X_I &= (X\beta_{[p]}(F) + \delta_{[p]}(F) - X_A\beta_A(F))X_I \\
+&= X_I(X_I^T\beta_{[p]}(F)_I + X_A^T(\beta_{[p]}(F)_A - \beta_A(F))  + \delta_{[p]}(F))
+\end{aligned}
+$$
+and
+$$
+\E_F(\delta_A(F)X_I) = \E_F(X_IX_I^T)\beta_{[p]}(F)_I + \E_F(X_IX_A^T)(\beta_{[p]}(F)_A - \beta_A(F)).
+$$
+Therefore, testing $H_0:\E_F(\delta_A(F)X_I)=0$ is close to
+to testing $\tilde{H}_0:\beta_{[p]}(F)_I=0, \beta_{[p]}(F)_A-\beta_A(F)=0$ but not
+quite the same. However, if we reject $H_0$ then we know one of the
+two conditions in $\tilde{H}_0$ holds.
+The complication here is that $\beta_A(F)$ appears in the law
+we are interested in, i.e. the law of $\delta_A(F)X_I$.
+
+Even in the fixed design case, we use the empirically adjusted residuals
+instead of the actual residuals. We might as well do the same here.
+Define an empirical version as the statistical functional
+$$
+\begin{aligned}
+\tilde{\gamma}_{A}(F) &= \E_F(yX_I - \E_F(X_I X_A^T) \E_F(X_AX_A^T)^{-1} yX_A) \\
+&= \E_F(X_I \delta_A(F) + X_IX_A^T\beta_A(F) - \E_F(X_I X_A^T) \E_F(X_AX_A^T)^{-1} X_A(X_A^T\beta_A(F) + \delta_A(F))) \\
+&= \E_F(X_I \delta_A(F)).
+\end{aligned}
+$$
+This has the same expected value as 
+the law we were interested in and it also has an $m$-of-$n$ bootstrap CLT as well.
+
+If $\mathbf{y}, \mathbf{X}$ is an IID sample of size $n$ from $F$ then
+$$
+\begin{aligned}
+\tilde{\gamma}_A(\hat{F}_n) &= \frac{1}{n} \left(\mathbf{X}_I^T\mathbf{y} - \left(\mathbf{X}_I^T\mathbf{X}_A \right) \left(\mathbf{X}_A^T\mathbf{X}_A \right)^{-1} \mathbf{X}_A^T\mathbf{y} \right) \\
+ &= \frac{1}{n}\mathbf{X}_I^T\mathbf{\delta_A(y)} \\
+\end{aligned}
+$$
+where
+$$
+\mathbf{\delta_A(y)} = \mathbf{y} - \mathbf{X}_A \left(\mathbf{X}_A^T\mathbf{X}_A \right)^{-1} \mathbf{X}_A^T\mathbf{y}
+$$
+are the empirical OLS residuals.
+
+It would seem that we might have
+run OLS for each bootstrap sample which might be expensive. 
+However, following \cite{penn_pairs} it is generally OK to replace
+matrices of cross-products in expressions. It is OK in the sense
+that the bootstrap CLT is generally preserved. For instance, the
+bootstrap distribution of $\beta_{[p]}(\hat{F}_n)$ is essentially the same as
+that of the functional
+$$
+\E_F(XX^T)^{-1} yX
+$$
+which is not a statistical functional per se as it involves $\E_F(XX^T)$.
+
+With this in mind, we might consider the functional 
+$$
+\begin{aligned}
+\gamma_A(\hat{F}_n) &= \frac{1}{n} \left(\mathbf{X}_I^T\mathbf{y} - \E_F(X_IX_A^T) \E_F(X_AX_A^T)^{-1} \mathbf{X}_A^T\mathbf{y} \right)
+\end{aligned}
+$$
+which satisfies
+$$
+\E_F(\gamma_A(\hat{F}_n)) = \E_F(\tilde{\gamma}_A(\hat{F}_n))
+$$
+and
+$$
+n\text{Var}_F(\gamma_A(\hat{F}_n)) \approx n \text{Var}_F(\tilde{\gamma}_A(\hat{F}_n)).
+$$
+This last assertion about the variance depends on a similar argument
+as in the pairs bootstrap which replaces the empirical cross-products
+$n^{-1}\mathbf{X}_I^T\mathbf{X}_A$ and $n^{-1}\mathbf{X}_A^T\mathbf{X}_A$
+with their expectation at the cost of an error of size $n^{-1/2}$ that does not affect the limiting variance.
+
+One advantage of $\gamma_A$ over $\tilde{\gamma}_A$ is that one does 
+not have to solve a least squares problem for every bootstrap sample as
+the bootstrap version of $\gamma_A$ would be
+$$
+\gamma_A(\hat{F}^{*,m}_n) &= \E_{\hat{F}^{*,m}_n}(yX_I) - 
+\E_{\hat{F}_n}(X_IX_A^T) (\E_{\hat{F}_n}(X_AX_A^T))^{-1} \E_{\hat{F}^{*,m}_n}(yX_A).
+$$
+
+This is a statistic and it has an $m$-of-$n$  bootstrap CLT centered at
+$\tilde{\gamma}_A(\hat{F}_n)$ with approximately correct covariance.
+
+\subsection{Sampling correction}
+
+When drawing bootstrap samples we want them to be such that the expected
+value of our statistic is 0 under $H_0$. Our bootstrap samples
+draw samples of $(yX, XX^T)$ and our selection is a function of these.
+To correct, we actually only need to modify the $yX$ part of the sample.
+Define 
+$$
+L_A(F) = \begin{pmatrix} I & \E_{F}(X_IX_A^T) (\E_{F}(X_AX_A^T))^{-1} \end{pmatrix} \in \real^{I \times [p]}
+$$
+so that
+$$
+\gamma_A(\hat{F}_n) =  L_A(F)\E_{\hat{F}_n}(yX).
+$$
+
+Let $\hat{G}^{*,m}_n$ to be the law of 
+$\hat{F}^{*,m}_n$ shifted by
+$$
+\Delta_A(\hat{F}_n) = L_A(\hat{F}_n)^{\dagger}L_A(\hat{F}_n) \E_{\hat{F}_n}(yX)
+$$
+so that
+$$
+\E_{\hat{G}^{*,m}_n}(\gamma_A(\hat{G}^{*,m}_n))= 0.
+$$
+
+Then, $\gamma_A(\hat{G}^{*,m}_n)$ satisfies an $m$-of-$n$ bootstrap CLT
+with the proper mean under $H_0$ and asymptotically (in $n$) correct variance.
+
+\bibliographystyle{alpha}
+\bibliography{nonparam}
+
 \end{document}

### 2015-06-04 code/R/SidakNo.R
counterexample to Sidak lemma for truncated Gaussian

@@ -1,0 +1,21 @@
+B <- 1E5
+rho <- .99
+set.seed(1)
+Z1 <- qnorm(runif(B, pnorm(4), 1)) #Gaussian conditional on being bigger than 4
+Z2.indep <- rnorm(B) # second Gaussian, independent of Z1
+Z2.corr <- Z1*rho + sqrt(1-rho^2)*Z2.indep # second Gaussian, correlated with Z1
+
+pdf("../../figs/SidakNo.pdf",height=6,width=10)
+par(mfrow=c(1,2))
+plot(rep(Z1[1:1E3],2), c(Z2.indep[1:1E3], Z2.corr[1:1E3]), col=rep(1:2,each=1E3), cex=.5, asp=1,
+     main="Truncated Gaussian", xlab="Z1", ylab="Z2")
+abline(v=4,lty=2)
+legend("bottomright", col=1:2, legend=c("Independent", "Correlated"), pch=1)
+
+plot(ecdf(pmax(Z1,Z2.indep)),main="ECDF of max(|Z1|, |Z2|)",xlim=c(4,5.3))
+plot(ecdf(pmax(Z1,Z2.corr)),add=T,col="red")
+legend("bottomright", col=1:2, legend=c("Independent", "Correlated"), lty=1, bg="white")
+dev.off()
+
+hist(,breaks=seq(4,7,.02),border="black",xlim=c(4,5.3))
+hist(pmax(Z1,Z2.corr),breaks=seq(4,7,.02),border="red",xlim=c(4,5.3),add=T)

### 2015-06-09 paper.tex
rewrote abstract

@@
 \begin{abstract}
-  Many model selection algorithms produce a ``path'' of fits that can be veiwed as specifying a sequence of models of increasing complexity.
-Given such a sequence of models and the data set used to produce them, we consider the problem of choosing the least complex model that is not falsified by the data, while accounting for the fact that the model path is determined adaptively using the data.
-Extending the tools of Fithian, Sun and Taylor (2014), we construct a $p$-value for each step in the sequence.
-In the case of linear regression, our $p$-values improve on the power of the spacings test of \citet{taylor2014exact}, often dramatically.
-By combining our $p$-values with stopping rules proposed by \citet{gsell2013sequential}, we achieve adaptive control of the familywise error rate and false discovery rate.
+  Many model selection algorithms produce a ``path'' of fits specifying a sequence of models of increasing complexity.
+Given such a sequence and the data used to produce them, we consider the problem of choosing the least complex model that is not falsified by the data, while accounting for the fact that the model path is determined adaptively using the data.
+Extending the selected-model tests of \citet{fithian2014optimal}, we construct $p$-values for each step in the path.
+In the case of linear regression, our $p$-values improve on the power of the spacings test of \citet{taylor2014exact}.
+The improvement in power can be dramatic for ``early steps'' in the path, when most of the signal has not yet been added to the model.
+To choose a model we apply stopping rules to the resulting $p$-values, achieving adaptive control of the familywise error rate and false discovery rate.
+
+Most of the stopping rules proposed in the literature assume that the single-step null $p$-values are independent of each other and of the non-null $p$-values.
+We derive intuitive and general conditions for independence and show that our proposed construction yields independent $p$-values.
 \end{abstract}
@@
 \section{Introduction}
+
+Many model selection procedures produce a path of fits
 

### 2015-06-10 paper.tex
drafted part of discussion

--- 2015-05-13 paper.tex
@@
 \begin{abstract}
-  Many model selection algorithms produce a ``path'' of fits that can be veiwed as specifying a sequence of models of increasing complexity.
-Given such a sequence of models and the data set used to produce them, we consider the problem of choosing the least complex model that is not falsified by the data, while accounting for the fact that the model path is determined adaptively using the data.
-Extending the tools of Fithian, Sun and Taylor (2014), we construct a $p$-value for each step in the sequence.
+  Many model selection algorithms produce a ``path'' of fits specifying a sequence of increasingly complex models.
+Given such a sequence and the data used to produce them, we consider the problem of choosing the least complex model that is not falsified by the data.
+Extending the selected-model tests of \citet{fithian2014optimal}, we construct $p$-values for each step in the path, accounting for the fact that the model path is determined adaptively using the data.
 In the case of linear regression, our $p$-values improve on the power of the spacings test of \citet{taylor2014exact}, often dramatically.
-By combining our $p$-values with stopping rules proposed by \citet{gsell2013sequential}, we achieve adaptive control of the familywise error rate and false discovery rate.
+
+To choose a model we feed the resulting $p$-values as inputs into sequential stopping rules proposed by \citet{gsell2013sequential} and \citet{li2015accumulation}, achieving adaptive control of the familywise error rate or false discovery rate.
+These stopping rules assume that the single-step null $p$-values are independent of each other and of the non-null $p$-values, a condition that is not satisfied by the $p$-values of \citet{taylor2014exact}.
+We derive intuitive and general conditions for independence and show that our proposed constructions yield independent $p$-values.
 \end{abstract}
@@
 
-Section~\ref{sec:sequential} concerns {\em stopping rules} $\hk$, estimators of $k_0$.
-We will consider several stopping rules that operate on the full sequence $p_{[d]}$, including two powerful stopping rules proposed in \citet{gsell2013sequential}.
-These latter require guarantees on the joint law of the $p$-values.
-We will prove sufficient conditions for when the $p$-values are Gaussian.
+Section~\ref{sec:sequential} concerns {\em stopping rules}, i.e.
+estimators $\hk(Y)$ of $k_0$.
+We will consider several stopping rules that operate on the full sequence $p_{[d]}$, including powerful rules proposed in \citet{gsell2013sequential} and \citet{li2015accumulation}.
+These stopping rules provably control the familywise error rate or the false discovery rate given that the null $p$-values are independent of each other and of the non-null $p$-values.
+Section~\ref{sec:sequential} gives general conditions under which the single-step $p$-values are independent.
+Under weak conditions, selected-model $p$-values are independent.
+By contrast, saturated-model $p$-values generically produce non-independent null $p$-values.
 
@@
 
-In some cases the design matrix of the full model may represent a  scrupulously curated set of features, and the analysts may know in advance that inferences with respect to the full model are of major scientific interest.
-For example, the scientist may believe, due to theoretical considerations, that a nonzero coefficient of $X_1$ after controlling for $X_2,\ldots,X_p$ would be evidence for a causal effect of $X_1$ on the response.
+In some cases the design matrix of the full model may represent a  scrupulously curated set of features, and the analysts may know in advance that inferences with respect to the full model are of primary scientific interest.
+For example, the scientist may believe, due to theoretical considerations, that a nonzero coefficient of $X_1$ after controlling for $X_2,\ldots,X_p$ would constitute strong evidence for a causal effect of $X_1$ on the response.
 
 If the full model has no special scientific status, however, then we see little advantage in insisting that all inferences should adjust for every other predictor in the full design matrix $X$.
-For example, suppose that $X$ contains gene expression measurements for all genes that happened to be measured by a microarry chip.
-Then $\beta_{j,\text{Full}}$ is already fairly arbitrary, since we would be controlling for a different set of genes if we had purchased the chip from a different manufacturer.
+For example, suppose that $X$ contains gene expression measurements for all genes that happened to be measured by a microarray chip, but if we had purchased the chip from a different manufacturer, we would have measured a different set of genes.
+Then, because the full-model coefficients are already artifacts of the somewhat arbitrary choice of which chip we bought, it is less clear that they are more scientifically interesting than the coefficients of a well-chosen submodel.
 
@@
 The main complication arises from the fact that the null and alternative hypotheses are random.
-To simplify matters, we first consider deriving $p$-values for a fixed candidate hypothesis $M\sub M_{\infty}$.
+To simplify matters, we first consider deriving $p$-values for a particular fixed candidate model $M\sub M_{\infty}$.
 
@@
 \]
-Given selective $p$-values for each fixed candidate pair, we can construct a combined $p$-value for the random null $M_{k-1}(Y)$. Define
+
+Given selective $p$-values for each fixed candidate $M$, we can construct a combined $p$-value for the random null $M_{k-1}(Y)$. Define
 \[
@@
 which is a valid $p$-value on the event $\{F \in M_{k-1}(Y)\}$:
-\[
+\begin{equation}\label{eq:selectiveGuaranteePk}
 \P_F\left(p_k \leq \alpha \mid M_{k-1}, 
   \;F\in M_{k-1}\right) \leq \alpha, \quad \forall \alpha \in [0,1].
-\]
-Recall that $M_{k-1}$ is random and not $F$.
+\end{equation}
+Recall that in~\eqref{eq:selectiveGuaranteePk} $M_{k-1}$ is random and not $F$.
 
@@
 \subsubsection{Saturated-Model Tests}
-Most of the work in selective inference for linear regression has focused on inference for least-squares parameters of the model $M_\infty:\; Y \sim \cN(\mu, \sigma^2I)$, with $\sigma^2$ assumed known.
-The saturated-model test for significance of variable $j_k$ after adjusting for the variables in $E_{k-1}$ tests whether the coefficient of $X_{j_k}$ is nonzero in the best linear model $X_{E_k}^\dagger Y$.
+Most of the work in selective inference for linear regression has focused on inference for least-squares parameters of the model $M_{\text{sat}}:\; Y \sim \cN(\mu, \sigma^2I)$, with $\sigma^2$ assumed known.
+The saturated-model test for significance of variable $j_k$ after adjusting for the variables in $E_{k-1}$ tests whether the coefficient of $X_{j_k}$ is nonzero in the best linear approximation to $\mu$, $X_{E_k}^\dagger \mu$.
 
@@
 Saturated-model $p$-values test the null hypothesis $H_0:\;\eta'\mu = 0$.
-Because $M_\infty$ has $n$ parameters, we must  condition on $\proj_\eta^\perp Y$ to eliminate the effect of the nuisance parameters $\proj_\eta^\perp \mu$ from the problem.
+Because $M_{\text{sat}}$ has $n$ parameters, we must  condition on $\proj_\eta^\perp Y$ to eliminate the effect of the nuisance parameters $\proj_\eta^\perp \mu$ from the problem.
 Thus, the test is based on
@@
 
-\WFcomment{References.} \citet{fithian2014optimal} call $M_\infty$ the {\em saturated model} because there is a parameter for every data point.
-They contrast tests based on $M_\infty$ with {\em selected-model} tests, which do not treat $\proj_\eta^\perp \mu$ as a nuisance parameter.
+\WFcomment{References.} \citet{fithian2014optimal} call $M_{\text{sat}}$ the {\em saturated model} because there is a parameter for every data point.
+They contrast tests based on $M_{\text{sat}}$ with {\em selected-model} tests, which assume $\proj_\eta^\perp \mu=0$ rather than treating it as a nuisance parameter.
 
@@
 
+It is a commonplace that ``essentially all models are wrong, but some are useful'' \citep{box1987empirical}.
+In essence, a statistical model is useful if it is large enough to capture the most important features of the data, but still small enough that inference procedures can achieve adequate power and precision.
+Apart from theoretical considerations, the only way to know whether a model is large enough is to test whether it is, using available data.
+
+Although model-checking is commonly recommended to practitioners as an important step in data analysis, it formally invalidates any inferences that are performed with respect to the model selected.
+Our work takes a step in the direction of reconciling that contradiction, but there are important questions left to be resolved.
+In particular: which sorts of model misspecification pose the most threat to our inferential conclusions, and how powerful are our tests against these most troublesome sources of misspecification?
+
+Of course, the answer depends on the scientific context: for example, suppose that at step 1, forward stepwise regression selects a variable $X_1$, and then at some later step it selects another variable $X_2$, which is almost perfectly correlated with $X_1$.
+There may be little evidence to support adding $X_2$ to the model once $X_1$ is already included, even if including $X_2$ would greatly change our inference about the coefficient for $X_1$ (for example, by making its confidence interval much wider).
+Depending on the context, we could draw the conclusion that 1) $X_1$ and $X_2$ are near-duplicate variables and it is therefore unnecessary (and possibly counterproductive) to include both in the model, or 2) $X_2$ is a vital confounding variable for $X_1$ and the confidence interval for $\beta_1$ ought to reflect the resultant uncertainty.
+If the second interpretation is the scientifically appropriate one, then we should probably use a different selection algorithm --- for example, a variant of forward stepwise that always adds both $X_1$ and $X_2$ to the model as a group in the same step.
+
+\section*{Acknowledgments}
+
+The authors are grateful for stimulating and informative conversations with Stefan Wager, Lucas Janson, Trevor Hastie, ....
+
 \bibliographystyle{plainnat}

### 2015-06-12 paper.tex
thorough rewrite of intro

@@
 \newcommand{\cN}{\mathcal{N}}
+\newcommand{\cJ}{\mathcal{J}}
 \newcommand{\leqAS}{\overset{\textrm{a.s.}}{\leq}}
@@
 
-\subsection{Generic Setting}
+\subsection{Notation and Problem Setting}
 
@@
 \[
-M_0(Y) \sub M_1(Y) \sub \cdots \sub M_d(Y).
-\]
-Define the {\em completion index} $k_0(Y) = \min\{k:\; F \in M_k\}$, the index of the first correct model.
-
-We will consider two related problems. First, we will consider the problem of obtaining selective single-step $p$-values, where $p_k$ is a $p$-value for testing
+M_0(Y) \sub M_1(Y) \sub \cdots \sub M_d(Y) \sub M_\infty.
+\]
+By ``model,'' we mean a family of probability distributions for $Y$. Without loss of generality, we can take $M_\infty$ as the union of all models under consideration.
+
+Define the {\em completion index} $k_0(Y) = \min\{k:\; F \in M_k(Y)\}$, the index of the first correct model.
+By construction, $F\in M_k \iff k \geq k_0$.
+Our goal is to examine the data and output a {\em stopping rule}, i.e.
+an estimator $\hk(Y)$ of $k_0$.
+We consider $M_k$ to be ``rejected'' if $k < \hat k$, and ``accepted'' otherwise, so that $V=(\hat k - k_0)_+$ is the number of mistaken rejections.
+
+We will break the problem into two constituent subproblems. First, Section~\ref{sec:singleStep} will discuss how to obtain selective single-step $p$-values, where $p_k$ is a $p$-value for testing
 \[
@@
 \]
-while adjusting for the fact that the models are chosen adaptively.
-In most of our development we will make few formal assumptions about the form of the candidate models.
+adjusting for the fact that the models are chosen adaptively.
+In most of our development we will make few assumptions about what form the candidate models take.
 Still, our results are only useful in cases where we can actually construct single-step $p$-values.
-At present, that means our applications are largely restricted to exponential family models.
+At present, this largely restricts our applications to exponential family models.
+
+Second, Section~\ref{sec:sequential} focuses on the sequential nature of the problem.
+\citet{gsell2013sequential} and \citet{li2015accumulation} propose stopping rules that operate on the sequence $p_{[d]}$.
+These stopping rules provably control type I error rates like the familywise error rate $\P(V>0)$ or the false discovery rate $\E[V/(\hk \vee 1)]$, provided that the null $p$-values are mutually independent and independent of the non-null $p$-values.
+We will discuss general conditions under which single-step $p$-values are independent.
+
+\subsubsection{Linear Regression}
 
 Most of our examples will focus on the important and familiar problem of adaptively selecting a set of predictors in linear regression models.
+In this problem, we observe a random response $Y\in \R^n$ as well as a fixed design matrix $X\in \R^{n \times p}$, whose columns correspond to candidate predictors.
+For each {\em active set} $E \sub [p]$, there is a corresponding candidate model
+\[
+M(E):\; Y \sim \cN( X_E\beta, \sigma^2 I_n),
+\]
+which is a subset of the {\em full model}
+\[
+M_\infty:\; Y \sim \cN(X\beta, \sigma^2I_n).
+\]
+The error variance $\sigma^2$ may be known or unknown depending on the context.
+
+\subsubsection{Sparse Parametric Models}
+
+All of the specific examples considered in this paper will have a common form generalizing the linear regression problem above.
+Let $M_\infty$ be a model parameterized by $\theta\in \Theta \sub \R^{\cJ}$:
+\[
+M_\infty = \{F_\theta:\; \theta \in \Theta\}.
+\]
+For any subset $E\sub \cJ$ define the sparse submodel with active coefficients $E$ as follows:
+\[
+\Theta(E) = \{\theta:\; \theta_j = 0, \;\;\forall j \notin E\}, 
+\quad M(E) = \{F_\theta:\; \theta\in \Theta_E\}.
+\]
+
+This setting includes linear regression problems with known or unknown $\sigma^2$ as special cases. A typical path algorithm returns a sequence of nested active sets 
+\[
+E_0(Y) \sub E_1(Y) \sub \cdots \sub E_d(Y),
+\]
+inducing the model path given by $M_k = M(E_k)$.
+
+As an example, in forward stepwise regression with known $\sigma^2$, we might include an unknown intercept term $\beta_0$ in the global null model $M_0$, then at each step $k\geq 1$, add the predictor $j_k$ that reduces the residual sum of squares the most.
+In that case $E_0 = \{0\}$, and $E_k = E_{k-1} \cup \{j_k\}$ for $k\geq 1$.
+
+\subsubsection{Path Algorithms}
+
+We will be especially interested in two path algorithms in the sparse parametric setting: forward-stepwise paths and ever-active regularization paths.
+
+Let $\ell(\theta; Y)$ denote the log-likelihood for model
+$M_\infty$. The {\em forward stepwise} algorithm proceeds as follows: we begin with some fixed $E_0$, then at step $k=1,\ldots,d$, we define
+\begin{align}\label{eq:forwardDef_start}
+j_k &= \argmax_j \;\;\sup \left\{\ell(\theta; Y):\; \theta\in\Theta(E_{k-1} \cup \{j\})\right\} \\
+E_k &= E_{k-1} \cup \{j_k\}\\\label{eq:forwardDef_end}
+M_k &= M(E_k).
+\end{align}
+That is, at each step we add to the model the one variable that would most increase the likelihood.
+In the case of linear regression, this amounts to adding the variable that maximally reduces the residual sum of squares.
+
+Another class of model selection procedures is the sequence of {\em ever-active} sets for a regularized likelihood path.
+For $r=0,1,\ldots,m$, let $P_r(\theta)$ denote some regularization penalty, and define
+\begin{align}\label{eq:regPathDef_start}
+  \hat\theta^{r}(Y) &= 
+  \argmin_{\theta\in\Theta} -\ell(\theta; Y) + P_r(\theta) \\
+  \tE_r(Y) &= \left\{j:\; \hat\theta_j^s \neq 0 
+    \text{ for any } s \leq r \right\}
+\end{align}
+
+While the sets $\tE_r$ are nested by definition, we could have $\tE_r = \tE_{r+1}$ for most values of $r$.
+Thus, we will take the sequence of distinct ever-active sets.
+Formally, let $R_0=0$, and for $k\geq 1$ let $R_k$ denote the (random) index where the active set actually changes for the $k$th time.
+That is,
+\begin{align}
+  R_k &= \min\{s:\; \tE_s \neq \tE_{R_{k-1}}\}\\
+  E_k &= \tE_{R_k}\\
+  \label{eq:regPathDef_end}
+  M_k &= M(E_k)
+\end{align}
+
+The ever-active lasso path in linear regression is the most familiar example, in which we take $P_r(\beta) = r^{-1}\|\beta\|_1$, and the penalized log-likelihood criterion is
+\begin{equation}
+  \argmin \frac{n}{2}\log(2\pi\sigma^2) 
+  + \frac{1}{2\sigma^2}\|Y - X\beta\|_2^2 + \frac{1}{r}\|\beta\|_1
+\end{equation}
+If $\sigma^2$ is known, then $\hat\beta^{r}$ is the lasso solution with Lagrange parameter $\lambda^r = \sigma^{2}/r$.
+If $\sigma^2$ is unknown, it is the lasso solution for $\lambda^r = (\hat\sigma^r)^2/r$.
+Because $\hat\sigma^r = n^{-1/2}\|Y-X\hat\beta^r\|_2$ is decreasing in $r$, $\lambda^r$ is strictly decreasing in $r$.
+
+Because it is actually possible to solve the lasso at every value of the Lagrange parameter, it is not really necessary that $r$ only take on integer values.
+We can instead notionally take $r\in (0,\infty)$, and use a path algorithm like LARS \citep{larspaper} to find every point in the path where the lasso solution changes.
+
+\subsection{Which Null Hypothesis Should We Test?}
+
+In our formulation of the problem, the type I error $V=(\hk-k_0)_+$ is defined in a ``model-centric'' fashion: at step $k$, we are testing  whether a particular linear model $M(E_{k-1})$ adequately describes the data $Y$.
+Even if the next selected variable $X_{j_k}$ is complete noise, adding it is not a mistake as long as there are other signal variables that have not yet been included.
+
+In some circumstances, we might want to define a type I error at step $k$ differently.
+If so, we must choose a different null hypothesis to test.
+Let $\mu = \E Y$ and let $\nu(E)$ denote the residual after projecting $\mu$ onto the column space of $X_E$:
+\[
+\nu(E) = \mu - X_{E}X_{E}^\dagger\mu,
+\]
+where $A^\dagger$ is the Moore-Penrose pseudoinverse of the matrix $A$. $\nu(E)$ is the portion of $\mu$ that is unexplained by $M(E)$.
+
+\citet{gsell2013sequential} describe three different null hypotheses that we could consider testing at step $k$ in the case of linear regression:
+\begin{description}
+\item[Complete Null:] $M_{k-1}$ is already correct; i.e., $H_k:\;\nu(E_{k-1}) = 0$
+\item[Incremental Null:] $M_{k-1}$ may be incorrect, but $M_k$ is no improvement; i.e., $H_k^I:\; \nu(E_{k-1}) \perp X_{j_k}$.
+\item[Full-Model Null:] The coefficient of $X_{j_k}$ is zero in the     full model; i.e., $H_k^F:\;\nu\left([p]\setminus \{j_k\}\right) \perp X_{j_k}$.
+\end{description}
+
+While the complete null is the strongest hypothesis of the three, the incremental null is neither weaker nor stronger than the full-model null. 
+Defining 
+\begin{align}
+V^{\text{I}} &= \#\{k < k_0:\; H_k^I \text{ is true}\} \quad \text{ and } \\
+V^{\text{F}} &= \#\{k < k_0:\; H_k^F \text{ is true}\},
+\end{align}
+we can define an analogous FWER and FDR with respect to each of these alternative choices, and attempt to control these error rates.
+
+In this article we will concern ourselves primarily with testing the complete null, which conforms most naturally to our stated aim of choosing the least complex model that is not rejected by the data.
+Using selective inference to control the type I error for $H_k^F$ or $H_k^I$ is an interesting topic for further study.
+
+In some cases the design matrix of the full model may represent a  scrupulously curated set of features, and the analysts may know in advance that inferences with respect to the full model are of primary scientific interest.
+For example, the scientist may believe, due to theoretical considerations, that a nonzero coefficient of $X_1$ after controlling for $X_2,\ldots,X_p$ would constitute strong evidence for a causal effect of $X_1$ on the response.
+
+If the full model enjoys no special scientific status, however, there is little advantage in insisting that all inferences should adjust for every other predictor in the full design matrix $X$.
+For example, suppose that $X$ contains gene expression measurements for all genes that happened to be measured by a microarray chip, but if we had purchased the chip from a different manufacturer, we would have measured a different set of genes.
+Because the full-model coefficients are already artifacts of the somewhat arbitrary choice of which chip we bought, there is no clear reason to regard them as more scientifically interesting than the coefficients of a well-chosen submodel.
+
+\begin{comment}
 As we will see, {\em selected-model} tests can be dramatically more powerful than {\em saturated-model} tests at early steps in the model path when most of the signal variables have not yet entered the model.
 
-Section~\ref{sec:sequential} concerns {\em stopping rules}, i.e.
-estimators $\hk(Y)$ of $k_0$.
-We will consider several stopping rules that operate on the full sequence $p_{[d]}$, including powerful rules proposed in \citet{gsell2013sequential} and \citet{li2015accumulation}.
-These stopping rules provably control the familywise error rate or the false discovery rate given that the null $p$-values are independent of each other and of the non-null $p$-values.
-Section~\ref{sec:sequential} gives general conditions under which the single-step $p$-values are independent.
-Under weak conditions, selected-model $p$-values are independent.
-By contrast, saturated-model $p$-values generically produce non-independent null $p$-values.
-
-\subsection{Which Null Hypothesis Should We Test?}
-
-Note that the type I error $V=(\hk-k_0)_+$ is defined in a ``model-centric,'' as opposed to a ``variable-centric,'' fashion: in regression, our model-wise $V$ counts the number of true null {\em models} incorrectly rejected, instead of counting the number of noise variables included in the final model.
-We can define the (model-wise) familywise error rate (FWER) and false discovery rate (FDR) respectively as $\P_F(V >0)$ and $\E_F[V/(\hk \vee 1)]$.
+Under weak conditions, selected-model $p$-values are independent. By contrast, saturated-model $p$-values generically produce non-independent null $p$-values.
 
@@
 
-In some cases the design matrix of the full model may represent a  scrupulously curated set of features, and the analysts may know in advance that inferences with respect to the full model are of primary scientific interest.
-For example, the scientist may believe, due to theoretical considerations, that a nonzero coefficient of $X_1$ after controlling for $X_2,\ldots,X_p$ would constitute strong evidence for a causal effect of $X_1$ on the response.
-
-If the full model has no special scientific status, however, then we see little advantage in insisting that all inferences should adjust for every other predictor in the full design matrix $X$.
-For example, suppose that $X$ contains gene expression measurements for all genes that happened to be measured by a microarray chip, but if we had purchased the chip from a different manufacturer, we would have measured a different set of genes.
-Then, because the full-model coefficients are already artifacts of the somewhat arbitrary choice of which chip we bought, it is less clear that they are more scientifically interesting than the coefficients of a well-chosen submodel.
-
 \WFcomment{Also say: Controlling for more variables could take us farther from a causal effect; furthermore, coefficient in larger model is much harder to interpret.}
-
-\section{Inference for One Step}
+\end{comment}
+
+\section{Inference for One Step}\ref{sec:singleStep}
 
@@
 
+define the restricted log-likelihood
+\[
+\ell_E(\theta; Y) = \left\{\begin{matrix} \ell(\theta; \;Y) 
+    & \theta \in \Theta(E)\\ -\infty  & 
+    \mathrm{ otherwise.}\end{matrix}\right. 
+\]
+
 \begin{proposition}[Regularized Likelihood Paths Satisfy SSP]\label{prop:regPathSSP}

### 2015-06-17 paper.tex
rewrote Section 2 (single-step p-values)

@@
 
-\subsection{Notation and Problem Setting}
+\subsection{Notation and Problem Setting}\label{sec:genericSetting}
 
@@
 
+\paragraph{Forward Stepwise}
 Let $\ell(\theta; Y)$ denote the log-likelihood for model
@@
 
+\paragraph{Ever-Active Regularization Paths}
 Another class of model selection procedures is the sequence of {\em ever-active} sets for a regularized likelihood path.
@@
 Because it is actually possible to solve the lasso at every value of the Lagrange parameter, it is not really necessary that $r$ only take on integer values.
-We can instead notionally take $r\in (0,\infty)$, and use a path algorithm like LARS \citep{larspaper} to find every point in the path where the lasso solution changes.
+We can instead notionally take $r\in (0,\infty)$, and use a path algorithm like LARS \citep{taylor2014exact} to find every point in the path where the lasso solution changes.
 
@@
 If so, we must choose a different null hypothesis to test.
-Let $\mu = \E Y$ and let $\nu(E)$ denote the residual after projecting $\mu$ onto the column space of $X_E$:
-\[
-\nu(E) = \mu - X_{E}X_{E}^\dagger\mu,
-\]
-where $A^\dagger$ is the Moore-Penrose pseudoinverse of the matrix $A$. $\nu(E)$ is the portion of $\mu$ that is unexplained by $M(E)$.
+Let $\mu = \E Y$ and let $\theta^E$ denote the {\em least-squares coefficients} of active set $E$ --- the coefficients of the best linear predictor for the design matrix $X_E$:
+\[
+\theta^E = X_E^\dagger \mu = \argmin_{\theta\in \R^{|E|}} \|\mu - X_E\theta\|_2^2,
+\]
+where $A^\dagger$ is the Moore-Penrose pseudoinverse of the matrix $A$. Write $\proj_E\mu = X_E^\dagger \theta^E$ for the projection of $\mu$ into the column space of $X_E$.
 
@@
 \begin{description}
-\item[Complete Null:] $M_{k-1}$ is already correct; i.e., $H_k:\;\nu(E_{k-1}) = 0$
-\item[Incremental Null:] $M_{k-1}$ may be incorrect, but $M_k$ is no improvement; i.e., $H_k^I:\; \nu(E_{k-1}) \perp X_{j_k}$.
-\item[Full-Model Null:] The coefficient of $X_{j_k}$ is zero in the     full model; i.e., $H_k^F:\;\nu\left([p]\setminus \{j_k\}\right) \perp X_{j_k}$.
+\item[Complete Null:] $M_{k-1}$ is already correct; i.e.,
+\[
+H_k:\;\mu = X_{E_{k-1}} \theta^{E_{k-1}}
+\]
+
+\item[Incremental Null:] $M_{k-1}$ may be incorrect, but $M_k$ is no improvement. That is, 
+\[
+H_k^{\text{inc}}:\; \theta_{j_k}^{E_k} = 0
+\]
+\item[Full-Model Null:] The coefficient of $X_{j_k}$ is zero in the     full model; that is, 
+\[
+H_k^{\text{full}}:\; \theta_{j_k}^{[p]} = 0.
+\]
 \end{description}
-
-While the complete null is the strongest hypothesis of the three, the incremental null is neither weaker nor stronger than the full-model null. 
-Defining 
+See \citet{gsell2013false} for an in-depth discussion of the relative conceptual advantages and disadvantages between these approaches.
+
+While the complete null is the strongest hypothesis of the three, the incremental null is neither weaker nor stronger than the full-model null. Defining 
 \begin{align}
-V^{\text{I}} &= \#\{k < k_0:\; H_k^I \text{ is true}\} \quad \text{ and } \\
-V^{\text{F}} &= \#\{k < k_0:\; H_k^F \text{ is true}\},
+V^{\text{inc}} &= \#\{k < k_0:\; H_k^{\text{inc}} \text{ is true}\} \quad \text{ and } \\
+V^{\text{full}} &= \#\{k < k_0:\; H_k^{\text{full}} \text{ is true}\},
 \end{align}
-we can define an analogous FWER and FDR with respect to each of these alternative choices, and attempt to control these error rates.
-
-In this article we will concern ourselves primarily with testing the complete null, which conforms most naturally to our stated aim of choosing the least complex model that is not rejected by the data.
-Using selective inference to control the type I error for $H_k^F$ or $H_k^I$ is an interesting topic for further study.
-
-In some cases the design matrix of the full model may represent a  scrupulously curated set of features, and the analysts may know in advance that inferences with respect to the full model are of primary scientific interest.
-For example, the scientist may believe, due to theoretical considerations, that a nonzero coefficient of $X_1$ after controlling for $X_2,\ldots,X_p$ would constitute strong evidence for a causal effect of $X_1$ on the response.
+we can define an analogous FWER and FDR with respect to each of these alternative choices, and attempt to control these error rates. For example, we can define
+\[
+\text{FDR}^{\text{full}} = \E[V^{\text{full}} / (\hk \vee 1)],
+\]
+as the false discovery rate with respect to the full-model null.
+
+In this article we will concern ourselves primarily with testing the complete null, which conforms most naturally to our stated aim of choosing the least complex model that is not rejected by the data, and is most easily generalized to the generic setting of Section~\ref{sec:genericSetting}.
+
+The incremental null is closely related to {\em saturated-model} tests for linear regression, as we will discuss further in Section~\ref{sec:singleStep}.
+While we could use the saturated-model framework to construct valid selective tests for $H_k^{\text{inc}}$, as proposed for LARS and lasso regression in \citet{taylor2014exact}, this will often result in a large reduction in power as well as introducing dependence between $p$-values at different steps.
+
+The full-model null is the most different conceptually from the other two.
+It takes a ``variable-centric'' instead of ``model-centric'' point of view: any variable with a nonzero coefficient in the full model is a {\em signal variable} and the others are {\em noise variables}, and $V^{\text{full}}$ counts the number of noise variables incorrectly selected.
+
+If the design matrix of the full model represents a scrupulously curated set of features, then the analyst may be primarily interested in inferences with respect to the full model.
+For example, the scientist may believe, due to theoretical considerations, that the linear model in $X_1, \ldots, X_p$ is fairly credible, and that a nonzero coefficient of $X_1$ after controlling for {\em all} of the other variables would constitute strong evidence for a causal effect of $X_1$ on the response.
 
@@
 Because the full-model coefficients are already artifacts of the somewhat arbitrary choice of which chip we bought, there is no clear reason to regard them as more scientifically interesting than the coefficients of a well-chosen submodel.
+
+See \citet{barber2014controlling} for an exact method to control $\text{FDR}^{\text{full}}$.
+\WFcomment{many high-dim inference references}.
+Using selective inference to control type I error rates for $H_k^{\text{full}}$ is an interesting topic for further study.
 
@@
 
-\section{Inference for One Step}\ref{sec:singleStep}
-
-To begin, we consider the problem of constructing a valid selective $p$-value for a particular step. At step $k$, we construct $p_k(Y)$ to test
+\section{Inference for One Step}\label{sec:singleStep}
+
+In this section we will consider the problem of constructing a valid selective $p$-value for a single step in the path. At step $k$, we construct $p_k(Y)$ to test
 \[
@@
 \]
-The main complication arises from the fact that the null and alternative hypotheses are random.
+As usual, the main complication arises from the fact that the null and alternative hypotheses are random.
 To simplify matters, we first consider deriving $p$-values for a particular fixed candidate model $M\sub M_{\infty}$.
@@
 
-Given selective $p$-values for each fixed candidate $M$, we can construct a combined $p$-value for the random null $M_{k-1}(Y)$. Define
+Once we have constructed selective $p$-values for each fixed candidate $M$, we can use them as building blocks to construct a combined $p$-value for the random null $M_{k-1}(Y)$. Define
 \[
@@
 
-Methods for one-step selective $p$-values are by now well-studied. \WFcomment{big list of references.} See \citet{fithian2014optimal} for a general treatment.
-
-\subsection{Selective $p$-Values in Exponential Families}
+In principle, we can construct selective tests for exponential family models conditional on any selection event. See \citet{fithian2014optimal} for a general treatment.
 
@@
 
-In linear regression, there are several tests that we might perform at step $k$ if we are , which generally use
+In linear regression, there are several tests that we might perform at step $k$.
+Let $E_{k-1}$ and $E_k$ denote the active sets for models $M_{k-1}$ and $M_k$ respectively.
+For simplicity we will assume that only one variable is added at step $k$, so that $E_k = E_{k-1} \cup j_k$.
 
 \subsubsection{Saturated-Model Tests}
-Most of the work in selective inference for linear regression has focused on inference for least-squares parameters of the model $M_{\text{sat}}:\; Y \sim \cN(\mu, \sigma^2I)$, with $\sigma^2$ assumed known.
-The saturated-model test for significance of variable $j_k$ after adjusting for the variables in $E_{k-1}$ tests whether the coefficient of $X_{j_k}$ is nonzero in the best linear approximation to $\mu$, $X_{E_k}^\dagger \mu$.
-
-Let $\proj_{E_{k-1}}\mu$ denote the projection of $\mu$ onto the column space of $X_{E_{k-1}}$, and let $\eta = \proj_{E_{k-1}}^\perp X_{j_k}$.
-Saturated-model $p$-values test the null hypothesis $H_0:\;\eta'\mu = 0$.
-Because $M_{\text{sat}}$ has $n$ parameters, we must  condition on $\proj_\eta^\perp Y$ to eliminate the effect of the nuisance parameters $\proj_\eta^\perp \mu$ from the problem.
-Thus, the test is based on
+
+Many of the selective tests for linear regression use saturated-model tests.
+See for example \citet{lockhart2014significance}, \citet{taylor2013tests}, \citet{taylor2014exact}, \citet{lee2013exact}, and \citet{loftus2014significance}.
+
+The distinction between the incremental null and the complete null is closely related to the distinction between {\em saturated-model} tests and {\em selected-model} tests discussed in \citet{fithian2014optimal}.
+Define the saturated model as
+\[
+M_{\text{sat}}:\; Y \sim \cN(\mu, \sigma^2I_n),
+\]
+so named because there is a mean parameter for every observation. Because the parameter $\mu$ has the same dimension as the data $Y$, we must assume that $\sigma^2$ is known.
+
+Saturated-model tests assume only that $M_{\text{sat}}$ is true, and perform inference on linear functionals of $\mu$.
+For example, in linear regression after selecting active set $E$, we might construct confidence intervals for the coefficients $\theta^E$.
+
+In the context of sequential model selection, we can use the saturated-model framework to test the incremental null $H^{\text{inc}}:\; \theta_{j_k}^{E_k} = 0$.
+The saturated model has $n$-dimensional sufficient statistic $Y$ and natural parameter $\mu$.
+Letting $\eta_k\in \R^n$ denote the vector for which $\theta_{j_k}^{E_k} = \eta_k'\mu$, the UMPU saturated-model test for $\theta_{j_k}^{E_k}$ is based on the distribution
 \begin{equation}\label{eq:satModel}
-\L\left( \eta'Y \mid E_{k-1}, j_k, \proj_\eta^\perp Y \right).
-\end{equation}
-To carry out exact tests of this hypothesis, we must assume that $\sigma^2$ is known, or that an independent estimate is available.
-
-\WFcomment{References.} \citet{fithian2014optimal} call $M_{\text{sat}}$ the {\em saturated model} because there is a parameter for every data point.
-They contrast tests based on $M_{\text{sat}}$ with {\em selected-model} tests, which assume $\proj_\eta^\perp \mu=0$ rather than treating it as a nuisance parameter.
+\L\left( \eta_k'Y \mid E_{k-1}, \;j_k, \;\proj_{\eta_k}^\perp Y \right),
+\end{equation}
+or equivalently on
+\begin{equation}\label{eq:satModelEquiv}
+\L\left( X_{j_k}'Y \mid E_{k-1}, \;j_k, \;\proj_{\eta_k}^\perp Y \right).
+\end{equation}
+The pair $(E_{k-1}, j_k)$ defines the selection event and $\proj_{\eta_k}^\perp Y$ eliminates the nuisance parameter $\proj_{\eta_k}^\perp \mu$.
 
 \subsubsection{Selected-Model Tests}
-Selected-model inference~\citep{fithian2014optimal} presents a more powerful alternative to testing for inclusion of variable $j_k$.
-Selected-model tests are so called because they are based on the statistical model chosen by our selection procedure.
-In this case, the null model selected for testing at step $k$ is simply the model $M_{k-1}:\; Y \sim \cN(X_{E_{k-1}}\beta, \sigma^2 I)$.
-
-There is a UMPU test of the null $M_{k-1}$ against the alternative $M_{k}\setminus M_{k-1}$, which has one more degree of freedom than $M_{k-1}$.
+Selected-model inference~\citep{fithian2014optimal} 
+represents a more powerful option when testing for inclusion of variable $j_k$.
+Selected-model tests are based on the statistical model chosen by our selection procedure.
+At step $k$, we test the selected null model $M(E_{k-1})$ against either the next model $M(E_k)$ or the upper model $M_\infty$.
+
 One option for selected-model inference is to condition on the identity of $j_k$, the next variable added, and perform the UMPU selective test of $M_{k-1}$ against $M_{k}\setminus M_{k-1}$.
-In that case, the test is based on
+If $\sigma^2$ is known, the test is based on
 \begin{equation}\label{eq:selModel_cond}
@@
 In~\eqref{eq:selModel_cond}, we only need to condition on an $|E_{k-1}|$-dimensional projection of $Y$, whereas in~\eqref{eq:satModel} we condition on $n-1$-dimensional projection of $Y$.
-Conditioning on more information tends to sap the power of a test, and we will see in Sections~\ref{sec:bivariate} and~\ref{sec:sparseReg}, the saturated-model tests often pay a heavy price for this extra conditioning.
-
-Although the tests in~\eqref{eq:selModel_cond} and~\eqref{eq:satModel} may appear to be based on different test statistics, they are actually not: because $\eta'Y = X_{j_k}'(Y - \proj_{E_{k-1}}Y)$, one test statistic is a deterministic affine function of the other after we condition on $X_{E_{k-1}}'Y$.
-
-It is not really necessary to condition on $j_k$: instead we could use
+Conditioning on more information tends to sap the power of a test, and we will see in Sections~\ref{sec:bivariate} and~\ref{sec:sparseReg}, the saturated-model tests can pay a heavy price for this extra conditioning.
+
+Instead of conditioning on the next model $M_k$, we could instead use
 \begin{equation}\label{eq:selModel_marg}
@@
 \end{equation}
-or even replace $X_{j_k}'Y$ with any other test statistic.
+or even replace $X_{j_k}'Y$ with any other test statistic $T(Y)$.
 Under the null hypothesis, the distribution of $Y$ is completely known once we condition on the value of $X_{E_{k-1}}'Y$, the complete sufficient statistic of $M_{k-1}$.
 
-We will see in Section~\ref{sec:pValsIndep} that conditioning on $j_k$ as in~\eqref{eq:selModel_cond} always results in independent sequential $p$-values, whereas tests of the form~\eqref{eq:selModel_marg} yield independent $p$-values only under certain conditions.
-Section~\ref{sec:pValsIndep} explores the conditions required to obtain independent sequential $p$-values.
-Carrying out a finer comparison of the two selected-model tests is an interesting direction for future work.
-
-Selected-model inference is more computationally involved than saturated-model inference.
+Selected-model inference is typically more computationally involved than saturated-model inference.
 Because the saturated-model test conditions on $n-1$ dimensions, the resulting distribution in~\eqref{eq:satModel} is nothing more than a truncated univariate Gaussian random variable.
-
-However, there are several major benefits to selected-model inference that often justify its use.
-First, selected-model inference allows us to drop the assumption that $\sigma^2$ is known.
-Second, $p$-values for selected-model tests are independent under general conditions, allowing us to plug the $p$-values into powerful sequential stopping rules such as those developed by~\citet{gsell2013sequential}.
-Most importantly, as we will see, selected-model tests can be dramatically more powerful than saturated-model tests.
-See \citet[][Section 5]{fithian2014optimal} for a detailed discussion of the difference between saturated-model and selected-model tests.
-
-\subsection{Comparison of Selected- and Saturated-Model inference}\label{sec:bivariate}
-
-We illustrate the superior power of the selected-model test in early steps with an extended example.
-
-\begin{example}[Bivariate Regression]\label{ex:bivariate}
-  Consider forward stepwise selection in a regression model with $n=p=2$, with identity design $X = I_2=\begin{pmatrix} 1 & 0 \\ 0 & 1\end{pmatrix}$ and known $\sigma^2=1$.
-We perform one step of forward stepwise and then test the null model with no variables against the model with one variable.
-
-We will choose variable 1 first on the selection event $A=\{|Y_1| > |Y_2|\}$, which is shown in yellow in Figure~\ref{fig:bv_condSets}. In that case,
+By contrast, selected-model tests typically require sampling from a truncated multivariate Gaussian distribution of dimension $p-|E_{k-1}|$.
+
+For some applications, a major conceptual advantage of the saturated-model approach is that, when $M(E)$ is misspecified, $\theta^{E}$ is still a well-defined quantity on which we can perform exact inference.
+For example, a 95\% confidence interval for $\theta_{j_k}^{E_k}$ would cover that coefficient with probability 95\% even if $M(E_k)$ is misspecified.
+In the setting of sequential model selection, this advantage is less compelling.
+We do not necessarily need a scientifically interpretable interval for $\theta_{j_k}^{E_k}$; rather, the goal of the test at step $k$ is to reject and move on if $M(E_{k-1})$ is demonstrably incomplete.
+
+There are several major benefits to selected-model inference.
+First, selected-model inference allows us to drop the assumption that $\sigma^2$ is known by conditioning on $\|Y\|^2$, the sufficient statistic corresponding to the natural parameter $1/2\sigma^2$.
+This is not an option for the saturated model because conditioning on both $\|Y\|^2$ and $\proj_{\eta_k}^\perp Y$ results in a degenerate conditional distribution for $\eta_k'Y$.
+See \citet[][Section 5]{fithian2014optimal} for a detailed discussion of the distinctions between saturated-model and selected-model tests.
+
+Second, we will see in Section~\ref{sec:pValsIndep} that tests of the form~\eqref{eq:selModel_cond} or~\eqref{eq:selModel_marg} yield independent $p$-values under general conditions, allowing us to apply  the sequential stopping rules of~\citet{gsell2013sequential} and \citet{li2015accumulation}.
+Finally, and perhaps most importantly, selected-model tests can be dramatically more powerful than saturated-model tests, as we will see next.
+
+\subsection{Comparison of Selected- and Saturated-Model Inference}\label{sec:bivariate}
+
+In this section we illustrate the superior power of the selected-model test in early steps with an extended example.
+
+\begin{example}[Bivariate Regression with Identity Design]\label{ex:bivariate}
+  Consider forward stepwise selection in a regression model with $n=p=2$, with known $\sigma^2=1$ and identity design 
+\[
+X = I_2=\begin{pmatrix} 1 & 0 \\ 0 & 1\end{pmatrix}.
+\]
+
+The forward stepwise path and the lasso path both select $j_1=1$ if and only if $|Y_1|>|Y_2|$.
+The selection event $A_1=\{|Y_1| > |Y_2|\}$ is shown in yellow in Figure~\ref{fig:bv_condSets}.
+If $j_1=1$ then
 \[
@@
 \]
-The selected-model test compares $Y_1$ to its distribution under $M_0$ conditional on $A$, a test of $H_0:\;\mu_1=0$ in model $M_1$.
-
+
+The selected-model test at step 1 compares $Y_1$ to its distribution under $M_0$ conditional on $A_1$, a test of $H_0:\;\mu_1=0$ in model $M_1$.
 By contrast, the saturated-model test is a test of $H_0:\; \mu_1=0$ in the model $M_{\text{sat}}:\; Y \sim \cN(\mu, I_2)$.
-Now, $\mu_2$ enters the problem as a nuisance parameter because the saturated-model test refuses to assume that $\mu_2=0$.
-To eliminate the nuisance parameter $\mu_2$, the saturated model test must condition on $Y_2$, and compare $Y_1$ to its null distribution given $A$ {\em and} the observed value of $Y_2$.
-
-Figure~\ref{fig:bv_condSets} shows the conditioning sets for each model when $Y=(2.9,2.5)$.
+The saturated model test must condition on $Y_2$ to eliminate the nuisance parameter $\mu_2$, and compare $Y_1$ to its null distribution given $A_1$ {\em and} the observed value of $Y_2$.
+
+Figure~\ref{fig:bv_condSets} shows the conditioning sets for each model when $Y=(2.9, 2.5)$.
 Next to it, Figure~\ref{fig:bv_nullDists} shows the null distribution for the test statistic $Y_1$ in each case.
@@
     \includegraphics[width=\textwidth]{figs/bivariateSelVSat_condSets.pdf}
-    \caption{\WFcomment{Copied directly from FST14; reword} 
-      For $Y=(2.9,2.5)$, the selected-model conditioning set is
-      $A=\{y:\;|y_1|>|y_2|\}$, a union of quadrants,
-      plotted in yellow. The saturated-model conditioning set
-      is ${\{y:\; y_2=2.5\}\cap A} = {\{y:\;y_2=2.5, |y_1|>2.5\}}$,
-      a union of rays, plotted in brown.}
+    \caption{The selected-model test conditions on $j_1=1$ (yellow region), while the saturated-model test also conditions on $Y_2=2.5$ to eliminate the nuisance variable $\mu_2$ (brown region).}
     \label{fig:bv_condSets}
@@
     \includegraphics[width=\textwidth]{figs/bivariateSelVSat_nullDists.pdf}
-    \caption{\WFcomment{Copied directly from FST14; reword} 
-      Conditional distributions of $Y_1$ under
-      $H_0:\mu_1 = 0$. Under the hypothesis
-      $\mu=0$, the realized  $|Y_1|$ is  quite large given $A$,
-      giving  $p$-value 0.015. By contrast, $|Y_1|$ is not too large
-      given $A \cap \{y:\; y_2=Y_2\}$, giving
-      $p$-value 0.3.}
+    \caption{Conditional distributions of $Y_1$ under
+      $M_0:\; Y \sim \cN(0,I_2)$. The realized value $|Y_1|=2.9$ is
+      quite large given that $j_1=1$. By
+      contrast, $|Y_1|=2.9$ is not especially large once we 
+      also condition on $Y_2=2.5$.}
   \end{subfigure}
-  \caption{\WFcomment{Copied directly from FST14; reword} 
-    Contrast between the saturated-model and selected-model tests
-    in Example~\ref{ex:bivariate}, in which we fit a one-sparse model with
-    design matrix $\bX=I_2$. The selected-model test is based
-    on  $\L_0(Y_1 \gv A)$, whereas the saturated-model test is based
-    on $\L_0(Y_1  \gv  Y_2, A)$.}
+  \caption{Contrast between the saturated-model and selected-model
+    tests in Example~\ref{ex:bivariate}. The selected-model test is based on  $\L(Y_1 \gv j_1=1)$,  whereas the saturated-model test is based on $\L(Y_1  \gv Y_2, \; j_1=1)$. 
+    When $Y=(2.9, 2.5)$, the selected- and saturated-model $p$-values are 0.015 and 0.3, respectively.}
   \label{fig:bv_nullDists}
@@
 
-The case illustrated in Figure~\ref{fig:bv_condSets} exhibits a phenomenon that has been remarked upon elsewhere in the literature of saturated-model tests: when there are near-ties between strong variables that are competing to enter the model, the winning variable may have a very weak $p$-value.
-\WFcomment{add references}.
-Figure~\ref{fig:bv_rocCurve} displays the cumulative distribution function for the first $p$-value when $\mu=\binom{4}{4}$, a very strong signal.
+Figure~\ref{fig:bv_condSets} illustrates a certain phenomenon occurring in saturated-model tests: when there are near-ties between strong variables that are competing to enter the model, the resulting $p$-value my be very weak~\citep{lockhart2014significance}.
+Figure~\ref{fig:bv_rocCurve} displays the cumulative distribution function for the first $p$-value when $\mu=(4,4)$, a very strong signal.
 While the selected model test has near perfect power, it is not uncommon for the saturated model test to produce large $p$-values, even in the range of 0.5-0.9.
-These large $p$-values arise exactly when there is a near tie between the variables.
+These large $p$-values arise when there is a near tie between the variables.
 
@@
   \includegraphics[width=.5\textwidth]{figs/bivariateSelVSat_rocCurve.pdf}
-  \caption{\WFcomment{Write caption here.}}
+  \caption{Cumulative distribution function of $p_1(Y)$ for the selected- and saturated-model tests when $\mu=(4,4)$.
+Even though the signal is very strong, the saturated-model test results in large $p$-values in realizations like the one in Figure~\ref{fig:bv_nullDists}, where there is a near-tie between $|Y_1|$ and $|Y_2|$.
+By contrast, the selected-model test has nearly perfect power.}
   \label{fig:bv_rocCurve}
@@
 Results in \citet{fithian2014optimal} show that the selected-model test is strictly more powerful when the selected model is correct; i.e., when $\mu_2=0$.
-Figure~\ref{fig:bv_powCurves_0} shows the power curve for each test when $\mu_2=0$.
-While the selected-model test is more powerful, the difference between the two is relatively mild.
-Interestingly, the difference is much more pronounced when $\mu_2=4$, as shown in Figure~\ref{fig:bv_powCurves_4}.
+Figure~\ref{fig:bv_powCurves} shows the power curve for each test when $\mu_2=0$ (left panel) and $\mu_2=4$ (right panel).
+While the selected-model test is more powerful when the selected model is correct, the difference between the two is relatively small.
+The difference is much more pronounced when $\mu_2=4$.
+
+Note that if $\mu=(0,4)$, the incremental null is true, but the complete null is false.
+That is, the model $M_0 = M(\emptyset)$ is missing an important signal variable, but the missing variable is not $X_1 = (1,0)$.
+Because the saturated-model test is really a test of the incremental null, its power is $\alpha=0.05$.
+By contrast, the selected-model test rejects about half the time when $\mu=(0,4)$, successfully detecting the fact that $M_0$ is a poor fit to the data.
 
@@
     \includegraphics[width=\textwidth]{figs/bivariateSelVSat_powCurves_0.pdf}
-    \caption{\WFcomment{Write caption here.}}
-    \label{fig:bv_powCurves_0}
+    %\caption{\WFcomment{Write caption here.}}
+    %\label{fig:bv_powCurves_0}
   \end{subfigure}
@@
     \includegraphics[width=\textwidth]{figs/bivariateSelVSat_powCurves_4.pdf}
-    \caption{\WFcomment{Write caption here.}}
+    %\caption{\WFcomment{Write caption here.}}
   \end{subfigure}
-  \caption{\WFcomment{Write caption here.}}
-   \label{fig:bv_powCurves_4}
+  \caption{Power at level $\alpha=0.05$ for saturated- and selected-model tests at step 1, given that $j_1=1$.
+The power is plotted as a function of $\mu_1$, for two different values of $\mu_2$.
+When $\mu_2=0$, the selected-model test is strictly more powerful than the saturated-model test, but the difference is slight.
+By contrast, when $\mu_2=4$, the selected-model test is much more powerful.
+The dashed line shows $\alpha=0.05$.}
+   \label{fig:bv_powCurves}
 \end{figure}

### 2015-06-17 code/R/bivariateSelVSat.R

@@ -2,6 +2,8 @@
 # Make plots of the conditioning sets and null distributions
 #########################################
 
+setwd("~/Dropbox/git/sequential-selection/code")
+
 # Conditioning Sets
 pdf("../figs/bivariateSelVSat_condSets.pdf",height=4.3,width=4)
 par(mar=c(4.1,4.1,3.1,0.1))
@@ -12,7 +14,7 @@ polygon(c(0,10,10),c(0,10,-10),lty=2,col="#F4E918")
 polygon(c(0,-10,-10),c(0,10,-10),lty=2,col="#F4E918")
 abline(h=0)
 abline(v=0)
-text(2,.5,"A")
+text(2,.5,expression(A[1]))
 text(y+c(.3,-.4),labels="Y")
 lines(c(y[2],10),c(y[2],y[2]),lwd=2,col="brown")
 lines(c(-y[2],-10),c(y[2],y[2]),lwd=2,col="brown")

### 2015-06-17 paper.tex
some more work, about to do major reorg

@@
 an estimator $\hk(Y)$ of $k_0$.
-We consider $M_k$ to be ``rejected'' if $k < \hat k$, and ``accepted'' otherwise, so that $V=(\hat k - k_0)_+$ is the number of mistaken rejections.
+We consider $M_k$ to be ``rejected'' if $k < \hat k$, and ``accepted'' otherwise.
+Thus $\hk$ is the number of models we rejected, while $k_0$ is the number that we should have rejected.
+The number of type I errors is $V=(\hk-k_0)_+$, while the number of type II errors is $(k_0-\hk)_+$.
 
@@
 Using selective inference to control type I error rates for $H_k^{\text{full}}$ is an interesting topic for further study.
-
-\begin{comment}
-As we will see, {\em selected-model} tests can be dramatically more powerful than {\em saturated-model} tests at early steps in the model path when most of the signal variables have not yet entered the model.
-
-Under weak conditions, selected-model $p$-values are independent. By contrast, saturated-model $p$-values generically produce non-independent null $p$-values.
-
-We could have instead defined the ``variable-centric'' type I error $\tV$ as the number of noise variables incorrectly included in the final model, with $\tR$ denoting the total number of variables included.
-Here, by ``noise variable'' we mean one that has a non-zero coefficient in the full model.
-Using selective inference to control the variable-wise FWER $\P_\mu(\tV>0)$ and FDR $\E[\tV / (\tR \vee 1)]$ is an interesting topic for further study.
-
-There is some ambiguity involved in deciding how to generalize $z$- and $t$-tests to the selective case.
-For example, \citet{gsell2013sequential} describe three different null hypotheses that we could consider testing at step $k$ in the case of linear regression under the upper model $M_\infty:\; Y\sim \cN(\mu, \sigma^2 I)$:
-\begin{description}
-\item[Complete Null:] The previous model, $M_{k-1}$, is correct; i.e., that $\mu = X_{E_{k-1}}\beta$ for some $\beta$.
-\item[Incremental Null:] $M_{k-1}$ may not be correct, but $M_k$ is no improvement: i.e., that $X_EX_E^\dagger\mu = \proj_{E_{k-1}}\mu$
-\end{description}
-. \citet{fithian2014optimal} discusses several different tests
-
-\WFcomment{Also say: Controlling for more variables could take us farther from a causal effect; furthermore, coefficient in larger model is much harder to interpret.}
-\end{comment}
 
@@
 
-Having discussed methods for constructing single-step $p$-values, we turn now to the problem of constructing a {\em stopping rule}; that is, an estimator $\hk$ of the completion index $k_0(Y)\in \{0,\ldots,d,\infty\}$.
-$M_k$ is ``rejected'' if and only if $\hk>k$.
-Thus $\hk$ is the number of models we rejected, while $k_0$ is the number that we should have rejected.
-The number of type I errors, then, is $(\hk-k_0)_+$, while the number of type II errors is $(k_0-\hk)_+$.
-
-\subsection{Stopping Rules}
-
-We will consider stopping rules that depend on the sequence $p_1,\ldots,p_d$ of $p$-values, which have been studied in the literature by \WFcomment{literature review here.} We will focus on three such stopping rules: Basic Stop, proposed by \WFcomment{whom?}, and the more refined Strong Stop and Forward Stop, proposed by \citet{gsell2013sequential}.
-
-While Strong Stop and Forward Stop tend to give more powerful stopping rules, they require independence among the $p$-values.
+Having discussed methods for constructing single-step $p$-values, we turn now to the question of using them to construct a stopping rule $\hat k$.
+Specifically, we will consider stopping rules that depend on the sequence $p_1,\ldots,p_d$ of $p$-values and must always reject an initial block of corresponding hypotheses.
+
+Section~\ref{sec:randomHyp} outlines several proposals for ordered testing of hypotheses and discusses what error rates they control and under what conditions.
+We focus on proposals from \citet{gsell2013sequential} and \citet{li2015accumulation}.
+
+These two papers both assume that the hypotheses $H_1, \ldots, H_d$ are fixed, and in particular that $k_0$ is fixed.
+In our problem, the model path is random and thus so is $k_0$.
+As we will see in Section~\ref{sec:randomHyp}, \WFcomment{...
+that this is not a problem}.
+
+While StrongStop, ForwardStop, and the accumulation tests tend to give more powerful stopping rules, they require independence among the $p$-values.
 Sections~\ref{sec:pValsIndep}--\ref{sec:modelSSP} discuss conditions on the model sequence $M_{[d]}$ and the $p$-value sequence $p_{[d]}$ under which selective $p$-values are mutually independent.
@@
 
-\subsubsection{Basic Stop:}
-
-\WFcomment{Is there a better name for this? Can we find it in the literature?}
-
-The most obvious sequential procedure is to reject at each step until the first time $p_k > \alpha$, which we can formalize as
+\subsection{Ordered Testing of Random Hypotheses}\label{sec:randomHyp}
+
+\subsubsection{BasicStop}
+The most obvious procedure is simply to reject at each step until the first time that $p_k > \alpha$, which we can formalize as
 \[
@@
 \]
-We will call this procedure {\em Basic Stop}.
+We will call this procedure {\em BasicStop}. \WFcomment{Is there a better name for this? Can we find it in the literature?}
 
@@
 \]
-\WFcomment{Trivial, but no doubt someone else proved this first...}
 
@@
 
-\subsubsection{Strong Stop:}
+\subsubsection{StrongStop}
 
@@
 \]
-Even if $p_k$ is a little larger than $\alpha$, say $\alpha=0.05$ and $p_k=0.06$, strong stop can still reject $M_{k-1}$ if the next three or four $p$-values are also relatively small.
+Even if $p_k$ is a little larger than $\alpha$, say $\alpha=0.05$ and $p_k=0.06$, StrongStop can still reject $M_{k-1}$ if the next three or four $p$-values are also relatively small.
 
@@
 \end{equation}
-If~\eqref{eq:indepCond_random_k0} holds, then strong stop controls theprobability of $V>0$ conditionally on $k_0$, and thus also marginally.
-
-\subsubsection{Forward Stop:}
+If~\eqref{eq:indepCond_random_k0} holds, then strong stop controls the probability of $V>0$ conditionally on $k_0$, and thus also marginally.
+
+\subsubsection{ForwardStop}
 
@@
 We stop the last time the estimate of FDR is less than $\alpha$.
-Forward stop requires the same condition independence condition~\eqref{eq:indepCond_random_k0} as strong stop does.
+Forward stop requires the same conditional independence condition~\eqref{eq:indepCond_random_k0} as strong stop does.
+
+\subsubsection{Accumulation Tests and HingeExp}
+
+\WFcomment{Discuss Rina's stuff.}
 

### 2015-06-18 paper.tex
done with reorg

@@
 \newcommand{\leqAS}{\overset{\textrm{a.s.}}{\leq}}
+\newcommand{\Err}{\mathcal{E}}
 
@@
 Thus $\hk$ is the number of models we rejected, while $k_0$ is the number that we should have rejected.
+
 The number of type I errors is $V=(\hk-k_0)_+$, while the number of type II errors is $(k_0-\hk)_+$.
+Depending on the context we might want to control the familywise error rate (FWER) $\P(V>0)$, the false discovery rate (FDR) $\E[V/(\hk \vee 1)]$, or another error rate given by the expectation of some loss function $g$:
+\begin{equation}\label{eq:errRate}
+\Err_F(\hk(\cdot), g) = \E_F\left[ g\left(\hk(Y), k_0(Y)\right)\right]
+\end{equation}
+
+We will assume that for each candidate model $M$ there is a minimal sufficient statistic $T(Y; \,M)$, and write
+\[
+T_k(Y) = T(Y; M_k(Y)).
+\]
 
@@
 \[
-H_{0,k}:\; F\in M_{k-1}\quad \text{ vs. } \quad 
-H_{1,k}:\; F\in M_\infty\setminus M_{k-1},
-\]
-adjusting for the fact that the models are chosen adaptively.
+H_{k}:\; F\in M_{k-1}(Y)
+\]
+against the alternative that $F\in M_\infty\setminus M_{k-1}$.
+The $p$-values must account for the fact that the models are chosen adaptively.
 In most of our development we will make few assumptions about what form the candidate models take.
@@
 \citet{gsell2013sequential} and \citet{li2015accumulation} propose stopping rules that operate on the sequence $p_{[d]}$.
-These stopping rules provably control type I error rates like the familywise error rate $\P(V>0)$ or the false discovery rate $\E[V/(\hk \vee 1)]$, provided that the null $p$-values are mutually independent and independent of the non-null $p$-values.
+These stopping rules provably control certain error rates, provided that the null $p$-values are mutually independent and independent of the non-null $p$-values.
 We will discuss general conditions under which single-step $p$-values are independent.
@@
 \]
-The error variance $\sigma^2$ may be known or unknown depending on the context.
+The error variance $\sigma^2$ may be known or unknown depending on the context. The complete sufficient statistic for $M(E)$ is $X_E'Y$ if $\sigma^2$ is known, or else $\left(X_E'Y, \|Y\|^2\right)$.
 
@@
 \Theta(E) = \{\theta:\; \theta_j = 0, \;\;\forall j \notin E\}, 
-\quad M(E) = \{F_\theta:\; \theta\in \Theta_E\}.
+\quad M(E) = \{F_\theta:\; \theta\in \Theta(E)\}.
 \]
@@
 
-\paragraph{Forward Stepwise}
+\paragraph{Forward Stepwise Paths}
 Let $\ell(\theta; Y)$ denote the log-likelihood for model
@@
 
+\WFcomment{Shorten this section, move most of it to Discussion}
+
 In our formulation of the problem, the type I error $V=(\hk-k_0)_+$ is defined in a ``model-centric'' fashion: at step $k$, we are testing  whether a particular linear model $M(E_{k-1})$ adequately describes the data $Y$.
@@
 Using selective inference to control type I error rates for $H_k^{\text{full}}$ is an interesting topic for further study.
+
+\section{Sequential Inference}\label{sec:sequential}
+
+An ordered hypothesis testing procedure takes in a sequence of $p$-values $p_1, \ldots, p_d$ for null hypotheses $H_{1}, \ldots, H_{d}$, and outputs a decision $\hk$ to reject the initial block $H_{1}, \ldots, H_{\hk}$ and accept the remaining hypotheses.
+We will assume throughout that the hypotheses are nested with $H_{k-1} \sub H_k$.
+Section~\ref{sec:orderedProposals} outlines several proposals for ordered testing and discusses what error rates they control and under what conditions.
+
+We focus on proposals from \citet{gsell2013sequential} and \citet{li2015accumulation}, all of which assume that the null hypotheses $H_{1}, \ldots, H_{d}$ are fixed, and that the null $p$-values are independent of each other and of the non-null $p$-values.
+Specifically, they assume that
+\begin{equation}\label{eq:indepCond_fixed_k0}
+\P(p_{k_0+1} \leq \alpha_{k_0+1}, \ldots, p_d \leq \alpha_d
+\mid p_1, \ldots, p_{k_0}) = \prod_{i=k_0+1}^d \alpha_i,
+\end{equation}
+where $k_0$ is fixed.
+
+In our problem the null hypotheses at each step are random and thus so is the completion index $k_0(Y)$. We will thus require {\em conditional} independence of null $p$-values given $k_0$:
+\begin{equation}\label{eq:indepCond_random_k0}
+  \P(p_{k+1} \leq \alpha_{k+1}, \ldots, p_d \leq \alpha_d
+  \mid p_1, \ldots, p_{k_0}, \; k_0 = k) = \prod_{i=k+1}^d \alpha_i
+\end{equation}
+
+The following proposition shows that~\eqref{eq:indepCond_random_k0} is the correct generalization of~\eqref{eq:indepCond_fixed_k0} to the random-hypothesis case.
+\begin{proposition}
+  Let $\hk$ be a stopping rule that operates on sequences
+  $p_{[d]}(Y)$ of $p$-values for ordered hypotheses $H_{[d]}(Y)$. 
+  For some function $g$, let 
+  \[
+  \Err(\hk,g) = \E\left[ g\left(\hk(Y), k_0(Y)\right)\right],
+  \]
+
+  Suppose that the rule $\hk$ controls $\Err$ at $\alpha$ under
+  independence when $H_{[d]}$ are fixed. That is, suppose that
+  whenever $H_{[d]}$ are fixed and $p_{[d]}$
+  satisfy~\eqref{eq:indepCond_fixed_k0}, we have
+  \[
+  \Err(\hk, g) \leq \alpha
+  \]
+
+  Then, $\hk$ controls $\Err$ at $\alpha$
+  whenever $p_{[d]}$ satisfy~\eqref{eq:indepCond_random_k0}.
+\end{proposition}
+\begin{proof}
+  For any fixed $k_0$ and $q_1,\ldots, q_{k_0}$, we can construct a
+  problem with $\tilde p_k \sim \delta_{q_k}$ for $k\leq k_0$ and 
+  $\tilde p_k \sim \text{Unif}[0,1]$ for $k > k_0$. Then, for the
+  fixed hypotheses 
+  $\widetilde H_k:\; \tilde p_k \sim \text{Unif}[0,1]$, we have
+  \[
+  f(k_0, q_{[k_0]}) 
+  = \E\left[ g\left(\hk(\tilde p_1, \ldots, \tilde p_d),
+      k_0\right) \right],
+  \]
+  which is less than or equal to $\alpha$ by assumption.
+
+  But then, by the condition~\eqref{eq:indepCond_random_k0}, we have
+  \begin{align}
+    \Err(\hk,g) &\;=\; \E\left[ \;\E\left[ g\left(\hk(p_1,\ldots, p_d), k_0(Y)\right) \; \mid \; k_0(Y), \; p_{[k_0]}(Y)\right] \; \right]\\
+    &\;\leq\; \E\left[f(k_0, p_{[k_0]})\right] \;\leq\; \alpha.
+  \end{align}
+\end{proof}
+
+In real applications, null $p$-values may be conservative instead of exactly uniform.
+We can relax~\eqref{eq:indepCond_random_k0} to require only that the joint distribution of null $p$-values is stochastically larger than uniform:
+\begin{equation}\label{eq:stochLargeCond_random_k0}
+  \P(p_{k+1} \leq \alpha_{k+1}, \ldots, p_d \leq \alpha_d
+  \mid p_1, \ldots, p_{k_0}, \; k_0 = k) \leq \prod_{i=k+1}^d \alpha_i
+\end{equation}
+If the null $p$-values are marginally uniform, then the conditions~\eqref{eq:indepCond_random_k0} and~\eqref{eq:stochLargeCond_random_k0} are equivalent.
+For most stopping rules $\hk$, the weaker condition~\eqref{eq:stochLargeCond_random_k0} is enough to guarantee control of whatever error rate holds under~\eqref{eq:indepCond_random_k0}.
+For the sake of brevity, we will say that $p$-value sequences satisfying~\eqref{eq:indepCond_random_k0} are {\em independent on nulls}.
+
+Sections~\ref{sec:pValsIndep}--\ref{sec:modelSSP} discuss conditions on the model sequence $M_{[d]}$ and the $p$-value sequence $p_{[d]}$ under which~\eqref{eq:stochLargeCond_random_k0} is satisfied.
+
+\subsection{Stopping Rules for Ordered Testing}\label{sec:orderedProposals}
+
+\begin{comment}
+\subsubsection{BasicStop}
+The most obvious procedure is simply to reject at each step until the first time that $p_k > \alpha$, which we can formalize as
+\[
+\hk_B(Y) = \min\left\{k \in \{1,\ldots,m\} :\;
+  p_k > \alpha\right\} - 1
+\]
+We will call this procedure {\em BasicStop}. \WFcomment{Is there a better name for this? Can we find it in the literature?}
+
+If $k_0$ is fixed, then it is clear that $\hk_b$ controls the FWER at level $\alpha$ provided that $p_{k}$ is a valid $p$-value for each $k>k_0$. Then, 
+\[
+\P(V>0) = \P(p_{k_0+1} \leq \alpha) \leq \alpha.
+\]
+
+Selectively valid single-step $p$-values do not necessarily guarantee FWER control when $k_0$ is random.
+In that case, we need a bit more, but it is sufficient that we have type I error control for $p_k$ conditional on $k_0=k-1$; i.e., not just conditional on $M_{k-1}$ being a correct model, but given that $M_{k-1}$ is the {\em first} correct model.
+\end{comment}
+
+\subsubsection{StrongStop}
+
+\citet{gsell2013sequential} propose another stopping rule that controls the FWER, which they called {\em strong stop}. Define
+\[
+  \hk_{S}(Y) = \max\left\{k \in \{1,\ldots,d\} :\;
+    \exp\left(\sum_{i=k}^d \frac{\log p_i}{i}\right) 
+    \leq \frac{\alpha k}{d}\right\}
+\]
+Even if $p_k$ is a little larger than $\alpha$, say $\alpha=0.05$ and $p_k=0.06$, StrongStop can still reject $M_{k-1}$ if the next three or four $p$-values are also relatively small.
+
+\citet{gsell2013sequential} show that if the completion index $k_0$ is fixed, then $\hk_S$ controls the FWER at level $\alpha$, as long as the null $p$-values are independent given the non-null ones.
+
+\subsubsection{ForwardStop}
+
+\citet{gsell2013sequential} propose another stopping rule, {\em forward stop}, that controls the FDR:
+\[
+  \hk_{F}(Y) = \max\left\{k \in \{1,\ldots,d\} :\;
+    -\frac{1}{k}\sum_{i=1}^k \log(1-p_i) \leq \alpha\right\}
+\]
+If $p_k$ is uniform then  $-\log(1-p_k)$ is an exponential random variable. If all of the null $p$-values are uniform and the others are zero, then only the null ones contribute to the sum, and
+\[
+\widehat{\text{FDR}}_k = -\frac{1}{k}\sum_{i=1}^k \log(1-p_i)
+\]
+is a Gamma random variable with mean $V/k$ and variance $V/k^2$. We stop the last time the estimate of FDR is less than $\alpha$.
+
+\subsubsection{Accumulation Tests and HingeExp}
+
+\WFcomment{Discuss Rina's stuff.}
+
+\subsection{Conditions for Independence}\label{sec:pValsIndep}
+
+The single-step $p$-value $p_k$ is selectively valid for testing $H_k:\; F \in M_{k-1}(Y)$ against $M_\infty \setminus M_{k-1}$ if, for all $F\in M_\infty$,
+\[
+\P_F(p_k \leq \alpha \mid M_{k-1},\; F\in M_{k-1}) \leqAS \alpha.
+\]
+Valid single-step selective $p$-values are not necessarily independent on nulls, but there is a simple sufficient condition under which independence on nulls is guaranteed.
+
+Consider any filtration
+\[
+\sF_0 \sub \sF_1 \sub \cdots \sub \sF_d
+\]
+for which $M_{[k]}$ is measurable with respect to $\sF_k$. We say that $\sF_{[d]}$ {\em separates} the $p$-values $p_{[d]}$ if 
+\begin{enumerate}
+\item $p_k(Y)$ is conservative given $\sF_{k-1}$ 
+  when $F\in M_{k-1}(Y)$, and
+\item $p_k(Y)$ is measurable with respect to $\sF_k$.
+\end{enumerate}
+If $p_k$ is exact, the first requirement is equivalent to requiring that $p_k$ is independent of $\sF_{k-1}$ on the event $F\in M_{k-1}(Y)$.
+
+If we think of $\sF_k$ as representing information available at step $k$, then the first condition means that $p_k$ ``excludes'' whatever evidence may have accrued against the null by step $k-1$, and the second means that any information revealed after step $k$ plays no role in determining $p_k$.
+In that sense, $\sF_k$ forms a ``wall of separation'' between the $p$-values up to $k$ and the ones after $k$.
+As a result, separated $p$-values are independent on nulls.
+
+\begin{proposition}[Independence of 
+  Separated $p$-Values]\label{prop:jointConserv}
+
+  Let $p_{[d]}$ be selective $p$-values for $M_{[d]}$, 
+  separated by $\sF_{[d]}$.
+
+  If the $p$-values are exact then $p_{k+1}, \ldots, p_d$ are
+  independent and uniform given $\sF_{k_0}$ on the event $\{k_0=k\}$.
+  If they are conservative, then for all
+  $\alpha_{k+1},\ldots,\alpha_d \in [0,1]$,
+  \begin{equation}\label{eq:jointConserv}
+  \P_F\left(p_{k+1}\leq \alpha_{k+1}, \ldots, p_d \leq \alpha_d\;
+    \mid\; k_0 \leq k, \; \sF_k\right) \leqAS \prod_{i=k+1}^d
+  \alpha_i.
+  \end{equation}
+\end{proposition}
+
+\begin{proof}
+  We first prove~\eqref{eq:jointConserv}
+  by induction. Define $B_i = \{p_i \leq \alpha_i\}$. 
+  The base case is
+  \[
+  \P_F\left(B_d \mid \sF_{d-1}\right)1_{\{k_0 \leq d-1\}} \leqAS \alpha_d,
+  \]
+  which is true by construction of $p_d$. 
+  For the inductive case, note that the
+  conditional probability on the left-hand side
+  of~\eqref{eq:jointConserv} is
+  \begin{align*}
+    \P_F\left(B_{k+1}, \ldots, B_d
+      \mid \sF_k\right)1_{\{k_0 \leq k\}} 
+    &\eqAS \E_F\bigg[ 1_{B_{k+1}} 
+    \P_F\left(B_{k+2}, \ldots, B_d
+      \mid \sF_{k+1}\right)1_{\{k_0 \leq k+1\}}
+    \mid \sF_k\bigg]1_{\{k_0 \leq k\}}\\
+    &\leqAS \P_F\left[ B_{k+1}
+      \mid \sF_k\right]1_{\{k_0 \leq k+1\}}\prod_{i=k+2}^d \alpha_i
+    \quad\;\;\leqAS\;\; \prod_{i=k+1}^d \alpha_i
+  \end{align*}
+
+  If the $p_k$ are exact then the above 
+  inequalities become equalities, implying uniformity and joint independence.
+\end{proof}
+
+\subsubsection{The Sufficient Filtration}\label{sec:suffFilt}
+
+We will concern ourselves mainly with the {\em sufficient filtration} for the path $M_{[k]}$, defined as
+\[
+\sF_k^{\text{suff}} = \sF(M_{[k]}, T_k),
+\]
+We will suppress the superscript and use $\sF_k$ to mean $\sF_k^{\text{suff}}$ for the remainder of the article.
+
+The sufficient filtration separates $p_{[d]}$ if and only if 
+\begin{enumerate}
+\item $p_k$ is conservative given $M_{[k-1]}$ and $T_{k-1}$, and
+\item $p_k$ is a function of $M_{[k]}$ and $T_k$.
+\end{enumerate}
+
+To be a valid selective $p$-value for a single step, $p_k$ only needs to be conservative conditional on $M_{k-1}$.
+In general, it is more stringent to additionally require that $p_k$ must also condition on $T_{k-1}$ as well as the entire subpath $M_{[k-1]}$.
+In most cases of interest, however, the ``extra'' conditioning is redundant, for two reasons.
+
+First, most selective tests already condition on $T_{k-1}$ in order to eliminate nuisance parameters. Second, as we will show in Section~\ref{sec:SSP}, for most path algorithms of interest we have
+\begin{equation}\label{eq:SSP}
+\sF(M_k, \, T_k) = \sF(M_{[k]}, \, T_k).
+\end{equation}
+Thus, most single-step selective $p$-values we would contemplate using are already exact conditional on $\sF_{k-1}$.
+
+The second requirement, that $p_k$ be $\sF_k$-measurable, has more bite.
+For example, we will see in Section~\ref{sec:singleStep} that it excludes saturated-model tests in linear regression.
+By contrast, any UMPU selective test for $M_{k-1}$ against $M_k\setminus M_{k-1}$ is typically measurable with respect to $\sF_k$.
+
+\subsubsection{The Subpath Sufficiency Principle}\label{sec:SSP}
+
+We say that a path algorithm $M_{[d]}(\cdot)$ satisfies the {\em subpath sufficiency principle} (SSP) if~\eqref{eq:SSP} holds for every $k$.
+Equation~\eqref{eq:SSP} means that if we know $M_k(Y) = M$, then we can compute the entire subpath of models $M_{[k]}$ knowing only $T(Y, \, M)$, the sufficient statistic of model $M$.
+For example, given that four steps of forward stepwise linear regression yield the active set $E_4 = \{1,5,7,11\}$, the order in which these variables were included can be determined from $T_4 = X_{E_4}'Y$.
+
+For the generic sparse parametric model of Section~\ref{sec:genSparse}, the forward stepwise path satisfies the SSP, as do all ever-active regularization paths. 
+Define the restricted log-likelihood
+\begin{equation}
+\ell_E(\theta; Y) = \left\{\begin{matrix} 
+    \ell(\theta; \;Y) & \theta \in \Theta(E)\\ 
+    -\infty  & \mathrm{ otherwise.}\end{matrix}\right. 
+\end{equation}
+
+\begin{proposition}[Forward Stepwise Satisfies SSP]\label{prop:forwardSSP}
+  The forward stepwise path as defined in
+  (\ref{eq:forwardDef_start}--\ref{eq:forwardDef_end}) satisfies the 
+  sub-path sufficiency principle.
+\end{proposition}
+\begin{proof}
+  For some fixed step $k$ and active set $E$, 
+  let $A$ denote the event $\{M_k = M(E)\}$. 
+  On $A$, the restricted likelihood $\ell_E$ is proportional to a
+  function depending only on $T_k = T(Y; \;M(E))$:
+  \[
+  \ell_E(\theta; Y) = \ell_E^{T_k}(\theta; T_k) 
+  + \ell_E^{Y \mid T_k}(Y \mid T_k).
+  \]
+  The second term does not depend on $\theta$ 
+  because $T(Y; \;M(E))$ is sufficient for $M(E)$.
+
+  On $A$, we have $E_i(Y) \sub E$ for all $i=1,\ldots, k$, meaning   that the maximum in \eqref{eq:forwardDef_start} was attained by some $j\in E$ at every step. So, for $i \leq k$, we have
+\begin{align}
+  j_i &= \argmax_{j\in E} \;\;\sup \left\{\ell_E(\theta; Y):\;
+    \theta\in\Theta(E_{i-1} \cup \{j\})\right\} \\
+  &= \argmax_{j\in E} \;\;\sup \left\{\ell_E^T(\theta; T(Y)):\;
+    \theta\in\Theta(E_{i-1} \cup \{j\})\right\}. \label{eq:onlyT}
+\end{align}
+
+Equation~\eqref{eq:onlyT} shows that $j_1,\ldots, j_{k-1}$ all depend on $Y$ only through $T_k(Y)$. As a result, it also follows that the enitre sequence $M_{[k]}(Y)$ depends only on $T_k(Y)$.
+\end{proof}
+
+\begin{proposition}[Regularized Likelihood Paths Satisfy SSP]\label{prop:regPathSSP}
+Ever-active regularized likelihood paths as defined in 
+(\ref{eq:regPathDef_start}--\ref{eq:regPathDef_end}) satisfy the
+sub-path sufficiency principle.
+\end{proposition}
+
+\begin{proof}
+  Again, for fixed $k$ and $E$ let $A$ denote the event $\{M_k = M(E)\}$ for some fixed $E$, and define $T_k$ as in the proof of Proposition~\ref{prop:forwardSSP}. If we define
+\[
+\hat\theta^{(E,r)} = \argmin_{\theta\in\Theta(E)} -\ell_E^{T_k}(\theta; T(Y)) + P_r(\theta),
+\]
+then 
+\[
+\hat\theta^{(E,r)} \eqAS \hat\theta^{r} \text{ on } 
+A \cap 1_{\{r \leq R_k\}}.
+\]
+But $\hat\theta^{(E,r)}$ depends only on $T_k(Y)$. Therefore, on $A$, we can reconstruct the entire path of solutions once we know $T_k(Y)$.
+\end{proof}
+
+Collecting together the threads of this section, we can arrive at our main result: a fairly general sufficient condition for a $p$-value sequence to be independent on nulls.
+\begin{theorem}[Sufficient Condition for Independence on Nulls] \label{thm:suffCond}
+Assume that $T(Y; M)$ is a complete sufficient statistic for each candidate model $M$, that $p_k$ are exact selective $p$-values for $M_{k-1}$, and that $M_{[d]}(\cdot)$ satisfies the SSP.
+
+If $p_k$ is $\sF_k$-measurable for each $k$, then the $p$-value sequence $p_{[d]}(Y)$ is independent on nulls.
+\end{theorem}
+
+\begin{proof}
+We apply the definition of completeness to the function 
+\[
+h(T(Y; \, M)) \;=\; 
+\P(p_k(Y) \leq \alpha \mid T(Y; \, M), \, M_{k-1}(Y) = M) - \alpha
+\]
+If $p_k$ is exact given $M_{k-1} = M$, then 
+\[
+\E_F[h(T_{k-1}) \mid M_{k-1}=M] \;=\; 0
+\] 
+for every $F\in M$.
+Therefore, we must have $h(T_{k-1})\eqAS 0$, so $p_k$ is independent of $\sF(M_k, T_{k-1})$.
+Because $M_{[d]}$ satisfies the SSP, $p_k$ is also independent of $\sF_{k-1} = \sF(M_{[k-1]}, T_{k-1})$.
+
+If $p_k$ is also $\sF_k$-measurable, then the sufficient filtration separates $p_{[d]}$, implying that $p_{[d]}$ is independent on nulls.
+\end{proof}
 
@@
 
-\section{Sequential Inference}\label{sec:sequential}
-
-Having discussed methods for constructing single-step $p$-values, we turn now to the question of using them to construct a stopping rule $\hat k$.
-Specifically, we will consider stopping rules that depend on the sequence $p_1,\ldots,p_d$ of $p$-values and must always reject an initial block of corresponding hypotheses.
-
-Section~\ref{sec:randomHyp} outlines several proposals for ordered testing of hypotheses and discusses what error rates they control and under what conditions.
-We focus on proposals from \citet{gsell2013sequential} and \citet{li2015accumulation}.
-
-These two papers both assume that the hypotheses $H_1, \ldots, H_d$ are fixed, and in particular that $k_0$ is fixed.
-In our problem, the model path is random and thus so is $k_0$.
-As we will see in Section~\ref{sec:randomHyp}, \WFcomment{...
-that this is not a problem}.
-
-While StrongStop, ForwardStop, and the accumulation tests tend to give more powerful stopping rules, they require independence among the $p$-values.
-Sections~\ref{sec:pValsIndep}--\ref{sec:modelSSP} discuss conditions on the model sequence $M_{[d]}$ and the $p$-value sequence $p_{[d]}$ under which selective $p$-values are mutually independent.
-We will see that, typically, selected-model $p$-values are sequentially independent while saturated-model $p$-values are not.
-
-\subsection{Ordered Testing of Random Hypotheses}\label{sec:randomHyp}
-
-\subsubsection{BasicStop}
-The most obvious procedure is simply to reject at each step until the first time that $p_k > \alpha$, which we can formalize as
-\[
-\hk_B(Y) = \min\left\{k \in \{1,\ldots,m\} :\;
-  p_k > \alpha\right\} - 1
-\]
-We will call this procedure {\em BasicStop}. \WFcomment{Is there a better name for this? Can we find it in the literature?}
-
-If $k_0$ is fixed, then it is clear that $\hk_b$ controls the FWER at level $\alpha$ provided that $p_{k}$ is a valid $p$-value for each $k>k_0$. Then, 
-\[
-\P(V>0) = \P(p_{k_0+1} \leq \alpha) \leq \alpha.
-\]
-
-Selectively valid single-step $p$-values do not necessarily guarantee FWER control when $k_0$ is random.
-In that case, we need a bit more, but it is sufficient that we have type I error control for $p_k$ conditional on $k_0=k-1$; i.e., not just conditional on $M_{k-1}$ being a correct model, but given that $M_{k-1}$ is the {\em first} correct model.
-
-\subsubsection{StrongStop}
-
-\citet{gsell2013sequential} propose another stopping rule that controls the model-wise FWER, which they called {\em strong stop}. Define
-\[
-  \hk_{S}(Y) = \max\left\{k \in \{1,\ldots,d\} :\;
-    \exp\left(\sum_{i=k}^d \frac{\log p_i}{i}\right) 
-    \leq \frac{\alpha k}{d}\right\}
-\]
-Even if $p_k$ is a little larger than $\alpha$, say $\alpha=0.05$ and $p_k=0.06$, StrongStop can still reject $M_{k-1}$ if the next three or four $p$-values are also relatively small.
-
-\citet{gsell2013sequential} show that if the completion index $k_0$ is fixed, then $\hk_S$ controls the FWER at level $\alpha$, as long as the null $p$-values are independent given the non-null ones.
-Specifically, they require that
-\begin{equation}\label{eq:indepCond_fixed_k0}
-\P(p_{k_0+1} \leq \alpha_{k_0+1}, \ldots, p_d \leq \alpha_d
-\mid p_1, \ldots, p_{k_0}) \leq \prod_{i=k_0+1}^d \alpha_i
-\end{equation}
-
-For random $k_0$, it is easy to see that the $p$-values simply need to satisfy~\eqref{eq:indepCond_fixed_k0} conditional on the value of $k_0$:
-\begin{equation}\label{eq:indepCond_random_k0}
-  \P(p_{k+1} \leq \alpha_{k+1}, \ldots, p_d \leq \alpha_d
-  \mid p_1, \ldots, p_{k_0}, \; k_0 = k) \leq \prod_{i=k+1}^d \alpha_i
-\end{equation}
-If~\eqref{eq:indepCond_random_k0} holds, then strong stop controls the probability of $V>0$ conditionally on $k_0$, and thus also marginally.
-
-\subsubsection{ForwardStop}
-
-\citet{gsell2013sequential} propose another stopping rule, {\em forward stop}, that controls the model-wise FDR:
-\[
-  \hk_{F}(Y) = \max\left\{k \in \{1,\ldots,d\} :\;
-    -\frac{1}{k}\sum_{i=1}^k \log(1-p_i) \leq \alpha\right\}
-\]
-If $p_k$ is uniform then  $-\log(1-p_k)$ is an exponential random variable. If all of the null $p$-values are uniform and the others are zero, then only the null ones contribute to the sum, and
-\[
-\widehat{\text{FDR}}_k = -\frac{1}{k}\sum_{i=1}^k \log(1-p_i)
-\]
-is a Gamma random variable with mean $V/k$ and variance $V/k^2$.
-We stop the last time the estimate of FDR is less than $\alpha$.
-Forward stop requires the same conditional independence condition~\eqref{eq:indepCond_random_k0} as strong stop does.
-
-\subsubsection{Accumulation Tests and HingeExp}
-
-\WFcomment{Discuss Rina's stuff.}
-
-\subsection{$p$-Values and Independence}\label{sec:pValsIndep}
-
-Recall that $p_k$ is a valid selective $p$-value for $M_{k-1}\sub M_\infty$ if, for all $F\in M_\infty$,
-\[
-\P_F(p_k \leq \alpha \mid M_{k-1},\; F\in M_{k-1}) \leqAS \alpha.
-\]
-Valid single-step selective $p$-values do not generally result in $p$-values that are independent from one step to the next.
-
-For example, saturated-model $p$-values are generically non-independent.
+Saturated-model $p$-values are generically non-independent.
 Continuing with Example~\ref{ex:bivariate}, Table~\ref{tab:bv_twoWayTable} shows a two-way contingency table for the saturated-model $p$-values $(p_1(Y), p_2(Y))$, binned into cells of height and width 0.2, simulated under the global null $\mu=0$.
 Because $k_0=0$, both $p$-values are uniform by construction, but the $p$-values are strongly dependent, with correlation $-48\%$.
-By contrast, we will see that the selected-model $p$-values $(p_1,p_2)$ are independent under the global null, a consequence of Proposition~\ref{prop:selectiveModel}.
+By contrast, the selected-model $p$-values $(p_1,p_2)$ are independent under the global null, a consequence of Theorem~\ref{thm:suffCond}.
+
+\begin{comment}
+Computing the cutoff for the saturated-model test requires us to know $\proj_{\eta}^\perp Y$, which is not $\sF_k$-measurable.
+
+By contrast, selected-model tests of $M_{k-1}$ against $M_k$ are always $\sF_k$-measurable, because they are based on the law
+\begin{equation}\label{eq:selModel}
+\L\left(X_{j_k}'Y \mid X_{E_k-1}'Y, E_{k-1}, j_k\right),
+\end{equation}
+and all of the random variables appearing in~\eqref{eq:selModel} are $\sF_k$-measurable.
+\end{comment}
 
@@
 
+\begin{comment}
 First, note that $p_2$ is a selective $p$-value for comparing the null with only variable 1,
@@
 By a similar argument, they would also be independent uniforms if $Y_2$ were chosen first; thus, $(p_1, p_2)$ are marginally uniform under the global null.
-
-We now introduce a simple sufficient condition to verify that a given procedure yields independent $p$-values.
-We say that a filtration $\sF_{[d]}$ with $M_{[k]} \in \sF_k$ {\em separates} the $p$-values $p_{[d]}$ if
-\begin{enumerate}
-\item $p_k(Y)$ is conservative given $\sF_{k-1}$ 
-  when $F\in M_{k-1}$, and
-\item $p_k(Y)$ is measurable with respect to $\sF_k$.
-\end{enumerate}
-If we think of $\sF_k$ representing information available at step $k$, then the first condition means that $p_k$ excludes whatever evidence may have accrued against the null by step $k-1$, and the second means that information revealed after step $k$ plays no role in determining $p_k$.
-In that sense, $\sF_k$ forms a ``wall of separation'' between the $p$-values up to $k$ and the ones after $k$.
-As a result, separated null $p$-values are independent.
-
-\begin{proposition}[Independence of 
-  Separated $p$-Values]\label{prop:jointConserv}
-
-  Let $p_{[d]}$ be selective $p$-values for $M_{[d]}$, 
-  separated by $\sF_{[d]}$.
-
-  If the $p$-values are exact then $p_{k+1}, \ldots, p_d$ are
-  independent and uniform given $\sF_{k_0}$ on the event $\{k_0=k\}$.
-  If they are conservative, then for
-  $\alpha_{k+1},\ldots,\alpha_d \in [0,1]$,
-  \begin{equation}\label{eq:jointConserv}
-  \P_F\left(p_{k+1}\leq \alpha_{k+1}, \ldots, p_d \leq \alpha_d\;
-    \mid\; k_0 \leq k, \; \sF_k\right) \leqAS \prod_{i=k+1}^d
-  \alpha_i.
-  \end{equation}
-\end{proposition}
-Equation~\ref{eq:jointConserv} does not quite imply independence, but it guarantees that both Strong Stop and Forward Stop control their target error rates.
-
-\begin{proof}
-  We first prove~\eqref{eq:jointConserv}
-  by induction. Define $B_i = \{p_i \leq \alpha_i\}$. 
-  The base case is
-  \[
-  \P_F\left(B_d \mid \sF_{d-1}\right)1_{\{k_0 \leq d-1\}} \leqAS \alpha_d,
-  \]
-  which is true by construction of $p_d$. 
-  For the inductive case, note that the
-  conditional probability on the left-hand side
-  of~\eqref{eq:jointConserv} is
-  \begin{align*}
-    \P_F\left(B_{k+1}, \ldots, B_d
-      \mid \sF_k\right)1_{\{k_0 \leq k\}} 
-    &\eqAS \E_F\bigg[ 1_{B_{k+1}} 
-    \P_F\left(B_{k+2}, \ldots, B_d
-      \mid \sF_{k+1}\right)1_{\{k_0 \leq k+1\}}
-    \mid \sF_k\bigg]1_{\{k_0 \leq k\}}\\
-    &\leqAS \P_F\left[ B_{k+1}
-      \mid \sF_k\right]1_{\{k_0 \leq k+1\}}\prod_{i=k+2}^d \alpha_i
-    \quad\;\;\leqAS\;\; \prod_{i=k+1}^d \alpha_i
-  \end{align*}
-
-  If the $p_k$ are exact then the above 
-  inequalities become equalities, implying uniformity and joint independence.
-\end{proof}
-
-\subsection{The Sufficient Filtration}\label{sec:suffFilt}
-
-Assume that each candidate model has a minimal sufficient statistic $T(y; M)$, and write
-\[
-T_k(Y) = T(Y; M_k(Y)).
-\]
-For example, in linear regression with known $\sigma^2$, the complete sufficient statistic for model $M(E)$ is $X_E'Y$, and the sufficient statistic at step $k$ is $T_k = X_{E_k}'Y$.
-
-We will concern ourselves mainly with the {\em sufficient filtration} for path $M_{[k]}$, defined as
-\[
-\sF_k = \sF(M_{[k]}, T_k),
-\]
-The sufficient filtration separates $p_{[d]}$ if $p_k$ is conservative given $\sF_{k-1}$ and $\sF_k$-measurable.\footnote{For readers unfamiliar with $\sigma$-algebra notation, this statement can be read as ``$p_k$ is conservative given $(M_{[k-1]}, T_{k-1})$ and is a function of $(M_{[k]}, T_k)$''}
-
-To be a valid selective $p$-value for a single step, $p_k$ only had to condition on $M_{k-1}$.
-Thus, it is a little more stringent to additionally require $p_k$ to condition on the entire subpath $M_{[k-1]}$ as well as $T_{k-1}$.
-
-In most cases of interest, however, the two requirements are equivalent.
-First, we typically construct $p_k$ by conditioning on $T_{k-1}$.
-In fact, if $T_{k-1}$ is complete sufficient, then any exact test must condition on $T_{k-1}$ (apply the definition of completeness to the function $h(Y) = \P(p_k(Y) \leq \alpha \mid T_{k-1}(Y))$).
-Second, most path algorithms of interest (including forward stepwise regression, the ever-active LASSO path, and their generalizations) satisfy a {\em subpath sufficiency principle} (henceforth, SSP): that once we know $M_k$ and its sufficient statistics, we can reconstruct the model path up to step $k$.
-Thus, most single-step $p$-values are already conservative conditional on $\sF_{k-1}$.
-
-The requirement that $p_k$ be $\sF_k$-measurable has more bite.
-For example, it excludes saturated-model tests in linear regression.
-Computing the cutoff for the saturated-model test requires us to know $\proj_{\eta}^\perp Y$, which is not $\sF_k$-measurable.
-
-By contrast, selected-model tests of $M_{k-1}$ against $M_k$ are always $\sF_k$-measurable, because they are based on the law
-\begin{equation}\label{eq:selModel}
-\L\left(X_{j_k}'Y \mid X_{E_k-1}'Y, E_{k-1}, j_k\right),
-\end{equation}
-and all of the random variables appearing in~\eqref{eq:selModel} are $\sF_k$-measurable.
-
-\subsection{Model Paths and the Subpath Sufficiency Principle}\label{sec:modelSSP}
-
-In Section~\ref{sec:suffFilt} we introduced the sufficient filtration as a way to get separated $p$-values.
-Because conditioning on $(M_{[k]}, T_k)$ is nominally different from conditioning only on $(M_k, T_k)$, we might suspect that we will sacrifice power by requiring $p_k$ to condition on the former.
-
-It will be useful to introduce the following notation:
-
-\begin{model}[Generalized Sparse Model]\label{mod:genSparse}
-  Let $M_\infty$ be a model parameterized by $\theta\in \Theta \sub \R^p$:
-  \[
-  M_\infty = \{F_\theta:\; \theta \in \Theta\}.
-  \]
-  For subsets $E\sub [p]$ define the submodel with active coefficients $E$ as follows:
-  \[
-  \Theta(E) = \{\theta:\; \theta_j = 0, \;\;\forall j \notin E\}, 
-  \quad M(E) = \{F_\theta:\; \theta\in \Theta_E\}.
-  \]
-  Let $\ell(\theta; Y)$ denote the log-likelihood for model
-  $M_\infty$, and define the restricted log-likelihood
-  \[
-  \ell_E(\theta; Y) = \left\{\begin{matrix} \ell(\theta; \;Y) 
-      & \theta \in \Theta(E)\\ -\infty  & 
-      \mathrm{ otherwise.}\end{matrix}\right. 
-  \]
-\end{model}
-
-In Model~\ref{mod:genSparse}, the {\em forward stepwise} algorithm proceeds as follows: we first set $E_0(Y)$ arbitrarily, then at step $k=1,\ldots,d$, we define
-\begin{align}\label{eq:forwardDef_start}
-j_k &= \argmax_j \;\;\sup \left\{L(\theta; Y):\; \theta\in\Theta(E_{k-1} \cup \{j\})\right\} \\
-E_k &= E_{k-1} \cup \{j_k\}\\\label{eq:forwardDef_end}
-M_k &= M(E_k)
-\end{align}
-
-To make this definition more concrete, we can specialize it to several familiar examples:
-
-\begin{example}[Forward Stepwise Regression]
-In the case of linear regression with known $\sigma^2$, $p$ is the number of variables and $M$ is $\cN(X\beta, \;\sigma^2I)$.
-For active set $E \sub [p]$, $M(E)$ is the model $\cN(X_E\beta_E, \;\sigma^2I)$, and $E_0 = \emptyset$.
-Because the maximized log-likelihood for a linear regression is monotone in the residual sum of squares, the forward stepwise algorithm adds whichever variable reduces the residual sum of squares the most.
-
-For linear regression with unknown $\sigma^2$, there are $p+1$ parameters $(\beta, \sigma^2)$ and $E_0 = \{p+1\}$ because $\sigma^2$ is a parameter for every model.
-\end{example}
-
-\begin{example}[Principle Components Analysis]
-\WFcomment{Do this one. Actually need infinitely many $\theta$s for this.}  
-\end{example}
-
-All forward stepwise procedures satisfy the SSP.
-
-\begin{proposition}[Forward Stepwise Satisfies SSP]\label{prop:forwardSSP}
-  The forward stepwise algorithm in Model~\ref{mod:genSparse} as defined in (\ref{eq:forwardDef_start}--\ref{eq:forwardDef_end}) satisfies the SSP.
-\end{proposition}
-\begin{proof}
-  Let $A$ denote the event $\{M_k = M(E)\}$ for some fixed active set $E$. Writing $T(Y) = T(Y; \;M(E))$, the likelihood restricted to $\Theta(E)$ is proportional to a function depending only on $T$:
-\[
-\ell(\theta; Y) = \ell_E^T(\theta; T(Y)) + \ell_E^{Y \mid T}(Y, T(Y)),
-\]
-where the second term does not depend on $\theta$ because $T$ is sufficient.
-
-On $A$, we have $E_i(Y) \sub E$ for all $i=1,\ldots, k$, meaning that the maximum in \eqref{eq:forwardDef_start} was attained by some $j\in E$ at every step. So, for $i \leq k$, we have
-\begin{align}
-  j_i &= \argmax_{j\in E} \;\;\sup \left\{\ell_E(\theta; Y):\;
-    \theta\in\Theta(E_{i-1} \cup \{j\})\right\} \\
-  &= \argmax_{j\in E} \;\;\sup \left\{\ell_E^T(\theta; T(Y)):\;
-    \theta\in\Theta(E_{i-1} \cup \{j\})\right\}. \label{eq:onlyT}
-\end{align}
-
-Equation~\eqref{eq:onlyT} shows that $j_1,\ldots, j_{k-1}$ all depend on $Y$ only through $T(Y)$. As a result, it also follows that the enitre sequence $M_{[k]}(Y)$ depends only on $T(Y)$.
-\end{proof}
-
-Another class of model selection procedures satisfying the SSP is the sequence ``ever-active'' sets in any regularized likelihood method.
-For $r=0,1,\ldots,m$, let $P_r(\theta)$ denote a regularizing penalty, and define
-\begin{align}\label{eq:regPathDef_start}
-  \hat\theta^{r}(Y) &= 
-  \argmin_{\theta\in\Theta} -\ell(\theta; Y) + P_r(\theta) \\
-  \tE_r(Y) &= \left\{j:\; \hat\theta_j^s \neq 0 
-    \text{ for any } s \leq r \right\}
-\end{align}
-
-While the sets $\tE_r$ are nested by definition, we could have $\tE_r = \tE_{r+1}$ for most values of $r$.
-We will take the sequence of {\em distinct} ever-active sets.
-Let $R_0=0$, and for $k>0$ let $R_k$ denote the (random) index where the active set actually changes:
-\begin{align}
-  R_k &= \min\{s:\; \tE_s \neq \tE_{R_{k-1}}\}\\
-  E_k &= \tE_{R_k}\\
-  \label{eq:regPathDef_end}
-  M_k &= M(E_k)
-\end{align}
-\WFcomment{Slightly sloppy definitions: note that $d$, the number of models, is random.
-$d$ could be 0 or could be infinite, depending on the data.
-Also, it isn't really necessary that the $r$ values be integers...}
-
-define the restricted log-likelihood
-\[
-\ell_E(\theta; Y) = \left\{\begin{matrix} \ell(\theta; \;Y) 
-    & \theta \in \Theta(E)\\ -\infty  & 
-    \mathrm{ otherwise.}\end{matrix}\right. 
-\]
-
-\begin{proposition}[Regularized Likelihood Paths Satisfy SSP]\label{prop:regPathSSP}
-Ever-active model paths derived from regularized likelihood methods as in (\ref{eq:regPathDef_start}--\ref{eq:regPathDef_end}) satisfy the sub-path sufficiency principle.
-\end{proposition}
-
-\begin{proof}
-  Again, let $A$ denote the event $\{M_k = M(E)\}$ for some fixed $E$, and define $T$ and $\ell_E^T$ as in the proof of Proposition~\ref{prop:forwardSSP}. If we define
-\[
-\hat\theta^{(E,r)} = \argmin_{\theta\in\Theta(E)} -\ell_E^T(\theta; T(Y)) + P_r(\theta),
-\]
-then 
-\[
-\hat\theta^{(E,r)} \eqAS \hat\theta^{r} \text{ on } 
-A \cap 1_{\{r \leq R_k\}}.
-\]
-But $\hat\theta^{(E,r)}$ depends only on $T(Y)$. Therefore, on $A$, we can reconstruct the entire path of solutions once we know $T(Y)$.
-\end{proof}
-
-\begin{comment}
-
-\begin{theorem}
-  Assume that $T(Y; M)$ is a complete sufficient statistic, 
-  that $p_k$ are exact selective $p$-values, and that 
-  $M_{[d]}$ satisfies EPIC.
-
-  Then, the sufficient filtration separates $p_{[d]}$ if and only if 
-  $p_k \in \sF_k$.
-\end{theorem}
-
 \end{comment}

### 2015-06-19 notes/nonparam.tex
included final sentence describing sampling scheme in a little more detail

@@
 
+{\bf Not sure this ``natural'' analog is the correct one.
+That is, it didn't use anything like the linear
+function $L$ in my abstract case. Hmm.....}
+
 A natural analog
@@
 
+The sampling scheme now draws samples from $\hat{G}^{*,m}_n$ retaining those 
+for which the selected model is the same. With these samples, one can standardize each entry of $\gamma_A(\hat{G}^{*,m}_n)$ by its bootstrap SD and finally compute the observed maxT.
+
+{\bf Again, not sure if this is the ``right'' thing to do as it didn't
+use the covariance structure of $yX$ under $\hat{F}_n$ to construct the test.}
+
 \bibliographystyle{alpha}

### 2015-06-19 paper.tex

@@
 \]
+\WFcomment{should probably end with $M_{d-1}(Y)$. need to fix this   throughout paper.}
 By ``model,'' we mean a family of probability distributions for $Y$. Without loss of generality, we can take $M_\infty$ as the union of all models under consideration.
@@
 
-\subsubsection{Sparse Parametric Models}
+\subsubsection{Sparse Parametric Models}\label{sec:genSparse}
 
@@
 
-\WFcomment{Shorten this section, move most of it to Discussion}
+\WFcomment{Shorten this section, move some of it to Discussion?}
 
@@
   \P(p_{k+1} \leq \alpha_{k+1}, \ldots, p_d \leq \alpha_d
-  \mid p_1, \ldots, p_{k_0}, \; k_0 = k) = \prod_{i=k+1}^d \alpha_i
+  \mid p_1, \ldots, p_k, \; k_0 = k) = \prod_{i=k+1}^d \alpha_i
 \end{equation}
@@
 
-Sections~\ref{sec:pValsIndep}--\ref{sec:modelSSP} discuss conditions on the model sequence $M_{[d]}$ and the $p$-value sequence $p_{[d]}$ under which~\eqref{eq:stochLargeCond_random_k0} is satisfied.
+Sections~\ref{sec:pValsIndep}--\ref{sec:SSP} discuss conditions on the model sequence $M_{[d]}$ and the $p$-value sequence $p_{[d]}$ under which~\eqref{eq:stochLargeCond_random_k0} is satisfied.
 
@@
 \end{enumerate}
-If $p_k$ is exact, the first requirement is equivalent to requiring that $p_k$ is independent of $\sF_{k-1}$ on the event $F\in M_{k-1}(Y)$.
+If $p_k$ is exact, the first requirement is equivalent to requiring that $p_k$ is independent of $\sF_{k-1}$ on the event $F\in M_{k-1}(Y)$ (i.e., the event when $p_k$ is null).
 
 If we think of $\sF_k$ as representing information available at step $k$, then the first condition means that $p_k$ ``excludes'' whatever evidence may have accrued against the null by step $k-1$, and the second means that any information revealed after step $k$ plays no role in determining $p_k$.
-In that sense, $\sF_k$ forms a ``wall of separation'' between the $p$-values up to $k$ and the ones after $k$.
+In that sense, $\sF_k$ forms a ``wall of separation'' between the $p$-values up to $k$ and the null $p$-values after $k$.
 As a result, separated $p$-values are independent on nulls.
@@
 
-In this section we will consider the problem of constructing a valid selective $p$-value for a single step in the path. At step $k$, we construct $p_k(Y)$ to test
-\[
-  H_{0,k}:\; F \in M_{k-1}(Y)
-  \quad \text{ vs. } \quad
-  H_{1,k}:\; F \in M_\infty \setminus M_{k-1}(Y).
-\]
-As usual, the main complication arises from the fact that the null and alternative hypotheses are random.
+Having derived conditions for independence, we turn now to the question of how to construct valid selective $p$-values for a single step in the path. At step $k$, we construct $p_k(Y)$ to test
+\[
+  H_{k}:\; F \in M_{k-1}(Y)
+\]
+against the alternative $F \in M_\infty \setminus M_{k-1}(Y)$.
+As usual, the main complication arises from the fact that the null hypothesis is random.
 To simplify matters, we first consider deriving $p$-values for a particular fixed candidate model $M\sub M_{\infty}$.
 
-The random variable  $p_{k,M}(Y)$ is a valid {\em selective $p$-value} for $M$ at step $k$ if it is stochastically larger than uniform under sampling distributions $F\in M$, given that $M$ is selected.
-That is,
+The random variable  $p_{k,M}(Y)$ is a valid {\em selective $p$-value} for $M$ at step $k$ if it is conservative under sampling from $F\in M$, given that $M$ is selected. That is,
 \[
@@
 \end{equation}
-Recall that in~\eqref{eq:selectiveGuaranteePk} $M_{k-1}$ is random and not $F$.
-
-In principle, we can construct selective tests for exponential family models conditional on any selection event. See \citet{fithian2014optimal} for a general treatment.
+%Recall that in~\eqref{eq:selectiveGuaranteePk} $M_{k-1}$ is random and not $F$.
+
+One very useful strategy for constructing exact selective tests is to condition on the sufficient statistics of the null model. Because the law
+\[
+\L(Y \;\mid\; T_{k-1}, \; M_{k-1}=M)
+\]
+is known under any $F\in M$, we can construct selective tests by comparing the distribution of any test statistic $U(Y)$ to its known conditional distribution under the null.
+See \citet{fithian2014optimal} for a general treatment of selective tests in exponential family models.
+
+If we follow this strategy, then the reference distribution for $U(Y)$ is measurable with respect to $T_{k-1}$ and $M_{k-1}$; moreover, if $U(Y)$ is also measurable with respect to $(M_k, T_k)$, the next model in the path and its sufficient statistics, then $p_k$ is measurable with respect to $\sF_k$, and so the $p$-value sequence $p_{[d]}$ will be independent on nulls per Theorem~\ref{thm:suffCond}.
+
+For the remainder of this section we will focus on linear regression.
 
@@
 
-In linear regression, there are several tests that we might perform at step $k$.
+In linear regression, there are several types of selective test that we might contemplate performing at step $k$.
 Let $E_{k-1}$ and $E_k$ denote the active sets for models $M_{k-1}$ and $M_k$ respectively.
 For simplicity we will assume that only one variable is added at step $k$, so that $E_k = E_{k-1} \cup j_k$.
+It is relatively straightforward to generalize the developments in this section to the case where multiple variables are added at once.
 
@@
 
-Many of the selective tests for linear regression use saturated-model tests.
+Many of the selective tests for linear regression use the saturated-model framework.
 See for example \citet{lockhart2014significance}, \citet{taylor2013tests}, \citet{taylor2014exact}, \citet{lee2013exact}, and \citet{loftus2014significance}.
 
-The distinction between the incremental null and the complete null is closely related to the distinction between {\em saturated-model} tests and {\em selected-model} tests discussed in \citet{fithian2014optimal}.
+The distinction between the incremental null and the complete null is closely related to the distinction between {\em saturated-model} tests and {\em selected-model} tests.
 Define the saturated model as
@@
 Saturated-model tests assume only that $M_{\text{sat}}$ is true, and perform inference on linear functionals of $\mu$.
-For example, in linear regression after selecting active set $E$, we might construct confidence intervals for the coefficients $\theta^E$.
+For example, in linear regression after selecting active set $E$, we might construct confidence intervals for the least-squares coefficients $\theta^E$ of active set $E$.
 
@@
 The saturated model has $n$-dimensional sufficient statistic $Y$ and natural parameter $\mu$.
-Letting $\eta_k\in \R^n$ denote the vector for which $\theta_{j_k}^{E_k} = \eta_k'\mu$, the UMPU saturated-model test for $\theta_{j_k}^{E_k}$ is based on the distribution
+Letting $\eta_k\in \R^n$ denote the vector for which $\theta_{j_k}^{E_k} = \eta_k'\mu$, the UMPU saturated-model test for $\eta_k'\mu$ is based on the distribution
 \begin{equation}\label{eq:satModel}
@@
 
+While the test statistic $\eta_k'Y$ is measurable with respect to $\sF_k$, the reference distribution we compare it against is not.
+Specifically, $\proj_{\eta_k}^\perp Y$ is not measurable with respect to $\sF_k$.
+As a result, saturated-model $p$-values constructed as in~\eqref{eq:satModel} are in general {\em not} independent on nulls.
+
 \subsubsection{Selected-Model Tests}
@@
 represents a more powerful option when testing for inclusion of variable $j_k$.
-Selected-model tests are based on the statistical model chosen by our selection procedure.
-At step $k$, we test the selected null model $M(E_{k-1})$ against either the next model $M(E_k)$ or the upper model $M_\infty$.
-
-One option for selected-model inference is to condition on the identity of $j_k$, the next variable added, and perform the UMPU selective test of $M_{k-1}$ against $M_{k}\setminus M_{k-1}$.
+While saturated-model inference always uses the same model $M_{\text{sat}}$, selected-model inference is based on the statistical model that is chosen by our selection procedure.
+In the sequential case this means that at step $k$, we test the selected null model $M(E_{k-1})$, against either the next selected model $M(E_k)$ or the upper model $M_\infty$.
+
+\paragraph{Conditioning on the Next Model}
+One option for selected-model inference is to condition on the identity of $j_k$, the next variable added, and perform a UMPU, equal-tailed, or one-sided selective test of $M_{k-1}$ against $M_{k}\setminus M_{k-1}$ conditional on both models.
 If $\sigma^2$ is known, the test is based on
@@
 In~\eqref{eq:selModel_cond}, we only need to condition on an $|E_{k-1}|$-dimensional projection of $Y$, whereas in~\eqref{eq:satModel} we condition on $n-1$-dimensional projection of $Y$.
-Conditioning on more information tends to sap the power of a test, and we will see in Sections~\ref{sec:bivariate} and~\ref{sec:sparseReg}, the saturated-model tests can pay a heavy price for this extra conditioning.
-
-Instead of conditioning on the next model $M_k$, we could instead use
+Conditioning on more information tends to sap the power of a test, and as we will see in Sections~\ref{sec:bivariate} and~\ref{sec:sparseReg}, the saturated-model tests can pay a heavy price for this extra conditioning.
+
+If we use the construction in~\eqref{eq:selModel_cond} after a subpath-sufficient path algorithm, the $p$-values are always guaranteed to have independence on nulls.
+The test statistic $X_{j_k}'Y$ is measurable with respect to $\sF_k$, and so is the reference distribution, which depends on $E_{k-1}, j_k,$ and $T_{k-1} = X_{E_{k-1}}'Y$.
+
+\paragraph{The Max-$t$ and Next-$\lambda$ Tests}
+Instead of conditioning on the next model $M_k$, we could use
 \begin{equation}\label{eq:selModel_marg}
@@
 \end{equation}
-or even replace $X_{j_k}'Y$ with any other test statistic $T(Y)$.
-Under the null hypothesis, the distribution of $Y$ is completely known once we condition on the value of $X_{E_{k-1}}'Y$, the complete sufficient statistic of $M_{k-1}$.
+or even replace $X_{j_k}'Y$ with any other test statistic $U(Y)$.
+As long as $U(Y)$ is measurable with respect to $(M_k, T_k)$, and the path algorithm is subpath-sufficient, the $p$-values are independent on nulls.
+
+Two specific examples are worth discussing in detail.
+First, consider the {\em selective max-$t$ test}, a generalization of the max-$t$ test of ~\citet{bujaxxxxmaxT}.
+Controlling for $E$, the $z$-statistic for coefficient $j\notin E$ is
+\begin{equation}
+  z_{j,E}(Y) = \frac{(\proj_E^\perp X_j)'Y}
+  {\sigma\|\proj_{E}^\perp X_j\|_2},
+\end{equation}
+and the $t$-statistic is
+\begin{equation}
+  t_{j,E}(Y) = \frac{(\proj_E^\perp X_j)'Y}
+  {\hat \sigma_{j,E} \|\proj_{E}^\perp X_j\|_2},
+  \qquad \text{where} \quad 
+  \hat \sigma_{j,E}^2(Y) = 
+  \frac{\|\proj_{E \cup \{j\}}^\perp Y\|_2^2}{n-|E|-1}.
+\end{equation}
+Note that $z_{j,E}$ is a monotonic function of $X_j'Y$ given $X_E'Y$, and $t_{j, E}$ is a monotonic function of $X_j'Y$ given $(X_E'Y, \|Y\|_2)$. Thus $z_{j,E}$ is measurable with respect 
+to $X_{E \cup \{j\}}'Y$ if $\sigma^2$ is known, and otherwise $t_{j,E}$ is measurable with respect to $(X_{E \cup \{j\}}'Y, \|Y\|_2)$.
+
+We can then define the max-$t$-statistic (respectively max-$z$-statistic) for active set $E$ as
+\begin{equation}\label{eq:maxT}
+  U_E(Y) = \max_{j \notin E} \left|t_{j,E}(Y)\right|,
+\end{equation}
+and use $U_{E_{k-1}}(Y)$ as our test statistic at step $k$.
+
+$U_{E_{k-1}}(Y)$ is $\sF_k$-measurable if and only if $j_k$ is the maximizer in~\eqref{eq:maxT}.
+Consequently, max-$t$ $p$-values are independent on nulls if they are used in tandem with forward-stepwise regression, where the next variable added is always the variable with the largest $t$-statistic.
+
+By contrast, the LASSO path does not in general add the variable with the largest $t$-statistic. In that case, we could use instead the {\em selective next-$\lambda$ test}, which uses test statistic
+\begin{equation}
+  U_k = \inf\{ \lambda:\; \hat\beta^\lambda \notin \beta(E_{k-1})\}.
+\end{equation}
+The next-$\lambda$ statistic is $\sF_k$-measurable if we use the LASSO to select variables, but not if we use forward-stepwise.
+\WFcomment{Relate this to spacings test.
+Is this essentially the ``generalized spacings test?''}
+
+\subsection{Comparison of Selected- and Saturated-Model Inference}\label{sec:bivariate}
 
@@
 For some applications, a major conceptual advantage of the saturated-model approach is that, when $M(E)$ is misspecified, $\theta^{E}$ is still a well-defined quantity on which we can perform exact inference.
-For example, a 95\% confidence interval for $\theta_{j_k}^{E_k}$ would cover that coefficient with probability 95\% even if $M(E_k)$ is misspecified.
-In the setting of sequential model selection, this advantage is less compelling.
-We do not necessarily need a scientifically interpretable interval for $\theta_{j_k}^{E_k}$; rather, the goal of the test at step $k$ is to reject and move on if $M(E_{k-1})$ is demonstrably incomplete.
+For example, a 95\% confidence interval for $\theta_{j_k}^{E_k}$ could cover that coefficient with probability 95\% even if $M(E_k)$ is misspecified.
+In the setting of sequential model selection, this advantage is less important.
+We do not necessarily need a scientifically interpretable interval for $\theta_{j_k}^{E_k}$; rather, the goal of the test at step $k$ is to reject $M_{k-1}$ and move on if it is demonstrably incompatible with the data.
 
@@
 
-Second, we will see in Section~\ref{sec:pValsIndep} that tests of the form~\eqref{eq:selModel_cond} or~\eqref{eq:selModel_marg} yield independent $p$-values under general conditions, allowing us to apply  the sequential stopping rules of~\citet{gsell2013sequential} and \citet{li2015accumulation}.
-Finally, and perhaps most importantly, selected-model tests can be dramatically more powerful than saturated-model tests, as we will see next.
-
-\subsection{Comparison of Selected- and Saturated-Model Inference}\label{sec:bivariate}
-
-In this section we illustrate the superior power of the selected-model test in early steps with an extended example.
+Second, we have seen in Section~\ref{sec:pValsIndep} that tests of the form~\eqref{eq:selModel_cond} or~\eqref{eq:selModel_marg} yield independent $p$-values under general conditions, allowing us to apply  the sequential stopping rules of~\citet{gsell2013sequential} and \citet{li2015accumulation}.
+
+Finally, and perhaps most importantly, selected-model tests can be dramatically more powerful than saturated-model tests, as we illustrate now with an extended example.
 

### 2015-06-30 paper.tex
rob changes

@@
 In the case of linear regression, our $p$-values improve on the power of the spacings test of \citet{taylor2014exact}, often dramatically.
+  We also show that the selected models test  bears a close relationship to parametric bootstrap sampling,
 
@@
 
-Many model selection procedures produce a path of fits
-
-... generates a sequence of increasingly complex models. Our goal is to choose the simplest model that is not falsified by the available data.
-
-... and find the index of the smallest adequate model --- that is, the smallest model that cannot be falsified given available evidence.
-
-\begin{example}[Forward-Stepwise Linear Regression with the LASSO]
-  \citet{taylor2014exact}
-\end{example}
-
-\begin{example}[The LARS Algorithm in Regression]
-  \citet{taylor2014exact}
-\end{example}
-
-\begin{example}[Ever-Active Path in $\ell_1$-Regularized Methods]
-  \citet{taylor2014exact}
-\end{example}
-
-\begin{example}[Principal Components Analysis]
-  As a second motivating example, consider model selection for   principal components analysis.
-In that case we are given a data matrix $X \in \R^{n\times d}$, with which we form a sample covariance matrix
-\[
-S = \frac{1}{n-1} \sum_{i=1}^n(x_i - \bar x)^2
-\]
-The first $d$ principal component loadings are the first $d$ eigenvectors of $S$, which call $u_1,\ldots, u_d$. These induce a sequence of nested Wishart models:
-\[
-M_0 \sub M_1 \sub \cdots \sub M_d
-\]
-in which
-\begin{equation}
-  M_k:\; (n-1) S \sim W_d\left(\lambda_0 I_d + \sum_{i=1}^k     \lambda_i u_i u_i', \;\;\; n-1\right).
-\end{equation}
-This problem was studied in \citet{choi2014selecting}.
-\end{example}
+Many model selection procedures produce  a sequence of increasingly complex models. Often our goal is to choose the simplest model that is not falsified by the available data.
+In other words, we seek  the index of the smallest adequate model ---  the smallest model that cannot be falsified given available evidence.
+Examples of such model selection procedures include  forward stepwise linear regression, least angle regression (LAR) and  the ever-active path in $\ell_!$-regularized models .
+All of these were studied in \citet{taylor2014exact}. Another, different, example is principal components analysis in which we try to determine the correct rank of the
+covariance matrix. This problem was studied in \citet{choi2014selecting}.
+
+A natural way carry out this model selection is as follows: we compute a set  of sequential p-values   at each step of the path of models and then feed them into a stopping rule
+that provides guaranteed control of the false discovery rate (FDR) or a similar quantity.
+Recently  \citet{gsell2013sequential} and \citet{li2015accumulation} proposed sequential stopping rules of this kind.
+Both sets of rules require that once we have reached the correct model in our path, the p-values in subsequent steps are uniform (or sub-uniform)
+and independent.
+In this paper we develop a theoretical framework for the construction of sequential p-values  that are both (sub)-uniform and independent under the correct model,
+and show how to construct them  in forward stepwise regression, LAR and the ever-active path in $\ell_1$-regularized models.
+In the former case, the test is simply a sequential selective version the max-$t$ test of ~\citet{bujaxxxxmaxT}, based one the theory developed \citet{fithian2014optimal}.
+
+Table \ref{tab:diab} shows an example: using the diabetes data taken from \cite{LARS}.
+These data consist of  observations on $442$ patients, with the response of interest being a quantitative measure of disease progression one year after baseline.
+There are ten baseline variables---age, sex, body-mass index, average blood pressure,
+and six blood serum measurements---plus quadratic terms, giving a total of 64 features.
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{rrrrr}
+  \hline\\
+ Step& Predictor& Naive pvalue& Saturated pvalue& Max-t pvalue\cr
+  \hline\\
+1 & 3 & 0.000 & 0.000 & 0.000 \\ 
+  2 & 9& 0.000 & 0.000 & 0.000 \\ 
+  3 & 4 & 0.000 & \color{red}{0.054} & 0.000 \\ 
+  4 & 20& 0.000 & { 0.333} & 0.003 \\ 
+  5 & 37& 0.002 & 0.756 & 0.043 \\ 
+  6 & 7 & 0.003 & 0.254 & 0.037 \\ 
+  7 & 2& 0.000 & 0.003 & 0.000 \\ 
+  8 & 19 & 0.020 & 0.028 &  \color{red}{0.343} \\ 
+  9 & 11 & 0.112 & 0.548 & 0.937 \\ 
+  10 & 52& 0.214 & 0.911 & 0.907 \\ 
+  11 & 57& 0.049 & 0.000 & 0.954 \\ 
+  12 & 24& 0.222 & 0.000 & 0.996 \\ 
+  13 & 5& 0.183 & 0.054 & 0.789 \\ 
+  14 & 49 & 0.158 & 0.333 & 1.000 \\ 
+  15 & 30 & 0.212 & 0.756 & 1.000 \\ 
+  16 & 46 & \color{red}{0.159} & 0.254 & 0.000 \\ 
+  17 & 59 & 0.270 & 0.003 & 0.004 \\ 
+  18 & 33& 0.116 & 0.028 & 1.000 \\ 
+  19 & 62 & 0.081 & 0.548 & 0.668 \\ 
+  20 & 27 & 0.170 & 0.911 & 1.000 \\
+
+   \hline\\
+\end{tabular}
+\caption[tab:diab]{\em Forward stepwise regression for the diabetes data: naive p-values from forward stepwise regression, 
+p-values from the ``saturated'' model, and the ``max-t'' p-values that are presented in this paper. The red annotation indicates the stopping step
+prescribed by {\tt forwardStop}. For the max-t p-values, this gives exact FDR control at the 10\% level.}
+\label{tab:diab}
+\end{table}
+The table shows the naive p-values from forward stepwise regression, that do not account for the selection, as well as the
+p-values from the ``saturated'' model, and the ``max-t'' p-values that are presented in this paper.
+The latter two sets of p-values test someone different hypotheses:  the p-values from the saturated model test whether the partial regression of the predictor just added is zero,
+what we call the ``incremental null''
+In contrast, the max-t pvalues test whether the current model is adequate, that is, all predictors not in the model  have coefficients equal to zero,
+the ``complete null''. (The issue of what hypothesis to test is discussed in section \ref{sec:whichnull}.)
+Now it turns out that the max-t test if often more powerful that the saturated model test, and importantly, the max-t pvalues are exactly
+independent after we reach the point the model sequence when the complete null is true. This independence means that we can apply
+sequential FDR exact stopping rules, such as ``forwardStop'' (\citet{sell2013sequential}). When applied to the max-t pvalues at FDR 0.10, {\tt forwardStop}
+suggested stopping at step 8. In constant, the saturated lvalues gave stopping a point of 3, with no guarantee of FDR control
+due to the lack of independence of the p-values.
+
+An outline of this paper is as follows:
+
+OUTLINE
+
+%\begin{example}[Forward-Stepwise Linear Regression]
+ % \citet{taylor2014exact}
+%end{example}
+
+%\begin{example}[The LARS Algorithm in Regression]
+%  \citet{taylor2014exact}
+%\end{example}
+
+%\begin{example}[Ever-Active Path in $\ell_1$-Regularized Methods]
+ % \citet{taylor2014exact}
+%\end{example}
+
+%\begin{example}[Principal Components Analysis]
+%  As a second motivating example, consider model selection for   principal components analysis.
+In that case we are given a data matrix $X \in \R^{n\times d}$, with which we form a sample %covariance matrix
+%\[
+%S = \frac{1}{n-1} \sum_{i=1}^n(x_i - \bar x)^2
+%\]
+%The first $d$ principal component loadings are the first $d$ eigenvectors of $S$, which call $u_1,\ldots, u_d$. These induce a sequence of nested Wishart models:
+%\[
+%M_0 \sub M_1 \sub \cdots \sub M_d
+%\]
+%n which
+%\begin{equation}
+ % M_k:\; (n-1) S \sim W_d\left(\lambda_0 I_d + \sum_{i=1}^k     \lambda_i u_i u_i', \;\;\; n-1\right).
+%\end{equation}
+%This problem was studied in \citet{choi2014selecting}.
+%\end{example}
 
@@
 \subsection{Which Null Hypothesis Should We Test?}
+\label{sec:whichnull}
 
@@
 
-As usual, we have reduced computation of the $p$-value at step $k$ to a sampling problem. Whatever our test statistic at step $k$, we must compute its distribution under the null model $M_{k-1}$.
-
-Happily, the polytope to condition on has only $2p$ constraints in forward stepwise regression, as we see next.
+As usual, we have reduced computation of the $p$-value at step $k$ to a sampling problem.
+Whatever our test statistic at step $k$, we must compute its distribution under the null model $M_{k-1}$.
+We give details of our  strategy here for the max-t test in forward stepwise regression.
+
+Given a model $M_{k-1}$   we need to sample from 
+\begin{eqnarray}
+L(X'_jY | E_{k-1},X'_{E_{k-1}})
+\label{eqn:maxtsamp}
+\end{eqnarray}
+with $Y\sim N(\mu, I \cdot \sigma^2)$.
+The current model fit is $\proj_{E_{k-1}} Y$ and is is easy to see that (\ref{eqn:maxtsamp})  is equivalent to sampling
+\begin{eqnarray}
+Y^*=\proj_{E_{k-1}}) Y + (I- \cdot \proj_{E_{k-1}}) Z; \;Z\sim N(0,I\sigma^2)
+\label{eq:boot}
+\end{eqnarray}
+and then conditioning on the event $E_{k-1}(Y^*)= E_k(Y)$ (a polyhedron).
+In other words, we need to condition on data vectors  $Y*$ that yields the same
+sequence of variables and coefficient signs up to step $k-1$. This is the challenging aspect of this computation.
+
+Note that  (\ref{eq:boot}  has the form of  (conditional) parametric bootstrap sampling, in which the bootstrap residuals projected into the subspace
+orthogonal to the current fit. In similar fashion, one might use a instead a nonparametric version of (\re{eq:boot}), sampling
+with replacement from the model residuals and then projecting them orthogonal to the current model.
+
+Figure \ref{fig:comparison} shows and example with $n=100,p=20$ and one strong signal, computing the max-t pvalue at Step 3.
+The ``exact'' answer  was determined from bootstrap sampling with $20,000$ replications.
+The Figures show the bias, variance and MSE of the estimates from accept/reject and hit and run, using between 1000 and 8000 samples, averaged
+over 10 different starting seeds.
+\begin{figure}[htp]
+\centering
+  \includegraphics[width=\textwidth]{figs/comparison.pdf}
+  \caption{\em Simulated example with $n= 100, p=20$ and one strong signal, computing the max-t pvalue at Step 3.
+Shown are bias, variance and MSE of the estimates from accept/reject and hit and run, using between 1000 and 8000 samples, averaged
+over 10 different starting seeds.
+}
+  \label{fig:comparison}
+\end{figure}
+We see that accept/reject is far more efficient, yielding a smaller MSE with 1000 samples then hit and run gives for  8000 samples.
+Hence we use accept/reject for as many steps of the sequence is possible,  until the point where the acceptance rate is too low to make is practical.
+
+In detail, we employ a hybrid computational strategy; we use accept-reject sampling, and do so as long as we obtain at least 300 (say) samples in the first set
+of $B$ realizations ($75,000$ in our code)  This is an effective approach until we have reached a point in the model path where all predictors have little or no signal.
+When the acceptance rate gets too low, we switch to a hit-and run Monte Carlo approach. We exploit the fact that each move along a random direction 
+has a truncated Gaussian distribution whose truncation limits are easily computed from the properties of the polyhedral. Unlike accept-reject, sampling
+which produces independent samples,  the hit and run method produces dependent samples. Hence we must run it for longer with an initial burn in period,
+and hope that it has mixed sufficiently well.
+
+The task is made much simpler by the fact that the selection event can we written as $S_M=\{y: \Gamma y \geq u\}$.
+Hence we don't need to run the entire forward stepwise procedure on $y^*$: instead, we pre-compute $\Gamma$ and $u$ and then check
+if $\Gamma y^* \geq u$. Further, since $\Gamma$ has on the  order of $2pk$ rows after $k$ steps,  we can save 
+substantially in computation by checking the row-wise inequalities starting from the last row, where the inequality  it is most likely to fail.
+
+\subsection{Reducing the number of constraints}
+Happily, it turns out that the conditioning polytope can be expressed with only $2p$ constraints in forward stepwise regression, as we show here.
+This increases the speed of both accept/reject and hit andRun sampling for our problem.
 
@@
 
+\subsection{An analytic approximation for the max-t test}
+ Suppose we have taken $k$ steps of forward stepwise regression with $p$ predictors and let the selected model  
+ be $E$. Suppose we choose predictor $j^*$ at the next step.
+
+Let $T_j=\eta_j^Ty, j \notin M$ be the partial regression coefficients for all predictors not in $E$,
+and $(v_j^-, v_j^+)$ be the limits derived from the polyhedral lemma applied to $(E, \eta_j)$  (See e.g. \citet{spacings}).
+These are the upper and lower bounds for the functional $\eta_j^Ty$ implied by the event that model $E$ is selected from the data $y$.
+We chose the sign of $\eta_j$ so that $\eta_j^Ty$ is  $\geq 0$.
+
+ Let $F_{\mu,\sigma^2}^{a,b}(t)$ be the  cumulative distribution function of the truncated normal.
+ Finally let $T=|T_{j^*}|$.
+Then by assuming independence, a  simple approximation to the max-t value is given by
+\begin{equation}
+1-\prod_{j\notin E} \{F_{0,||\eta||^2}^{v_j^-, v_j^+}(T)-   F_{0,||\eta||^2}^{v_j^-, v_j^+}(-T) \}
+\label{eqn:approxpv}
+\end{equation}
+Figure \ref{fig:approxpv} shows an example with $n=50$ and $p=10$ features having pairwise correlation 0.5.
+There are two moderate signals.
+Shown are the p-values from the ``exact'' numerical computation (with $100,000$ bootstrap samples)
+and the approximation based on (\ref{eqn:approxpv}).
+\begin{figure}[htp]
+\centering
+  \includegraphics[width=3in]{figs/approxpv.pdf}
+  \caption{Exact max-t pvalues and approximation based on (\ref{eqn:approxpv}), for a simulated example
+  with $n=50, p=10$.}
+  \label{fig:approxpv}
+\end{figure}
+The approximation seems quite good. By Sidak's lemma, we might conjecture that the approximation is conservative, but we have no proof of this.
+
 \section{Simulation: Sparse Linear Regression}\label{sec:sparseReg}
@@
 Next we compare several model selection procedures in simulation.
-We simlate from a linear regression model with $n=100$ observations and $p=40$ variables.
+We simulate from a linear regression model with $n=100$ observations and $p=40$ variables.
 The design matrix $X\in\R^{n\times p}$ is a random Gaussian design with pairwise correlations of 0.3 between predictor variables.
@@
   \includegraphics[width=.8\textwidth]{figs/simulation_snr_5_alpha_05_null_false.pdf}
-  \caption{CDFs of saturated-model (black), selected-model (red), and nominal (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on testing a false null hypothesis at step $k$.
+  \caption{\em CDFs of saturated-model (black), selected-model (red), and nominal (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on testing a false null hypothesis at step $k$.
 The selected-model test is much more powerful than the saturated-model test.
@@
   \includegraphics[width=.8\textwidth]{figs/simulation_snr_5_alpha_05_null_true.pdf}
-  \caption{CDFs of saturated-model (black), selected-model (red), and nominal (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on testing a true null hypothesis at step $k$.
+  \caption{\em CDFs of saturated-model (black), selected-model (red), and nominal (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on testing a true null hypothesis at step $k$.
 The selected-model test is much more powerful than the saturated-model test.
@@
   \includegraphics[width=.8\textwidth]{figs/simulation_snr_5_alpha_05_noise_var.pdf}
-  \caption{CDFs of saturated-model (black), selected-model (red), and nominal (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on the event that the variable added at step $k$ is a noise variable in the full model.
+  \caption{\em CDFs of saturated-model (black), selected-model (red), and nominal (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on the event that the variable added at step $k$ is a noise variable in the full model.
 Here, none of the methods produce uniform $p$-values.

### 2015-07-01 paper.tex
rob

@@
 \newcommand{\WFcomment}[1]{{\color{red}{(WF: \bf \sc #1) }}}
+\newcommand{\RTcomment}[1]{{\color{red}{(RT: \bf \sc #1) }}}
 
@@
 \section{Introduction}
-
+\label{sec:introduction}
 Many model selection procedures produce  a sequence of increasingly complex models. Often our goal is to choose the simplest model that is not falsified by the available data.
@@
 independent after we reach the point the model sequence when the complete null is true. This independence means that we can apply
-sequential FDR exact stopping rules, such as ``forwardStop'' (\citet{sell2013sequential}). When applied to the max-t pvalues at FDR 0.10, {\tt forwardStop}
+sequential FDR exact stopping rules, such as ``forwardStop'' (\citet{gsell2013sequential}). When applied to the max-t pvalues at FDR 0.10, {\tt forwardStop}
 suggested stopping at step 8. In constant, the saturated lvalues gave stopping a point of 3, with no guarantee of FDR control
@@
 An outline of this paper is as follows:
-
-OUTLINE
+In Section \ref{sec:genericSetting} we setup the problem and give some important definitions. In Section \ref{sec:sequential} we review recently proposed methods for
+testing ordered hypothesis using from a sequence of p-values. These method require that the p-values be uniform and independent under the null, and we derive general conditions
+to ensure this property. In Section \ref{sec:single step}  we  show how to construct valid selective $p$-values for a single step in the path and derive the max-t test
+for  forward stepwise regression and the Next-$\lambda$ test for the lasso.
+Section \ref{sec:computation}  describes our computational strategy for estimating the p-values using accept-reject and hit and run sampling.
+We also relate the method to parametric bootstrap sampling.
+We provide a simulation study in Section \ref{sec:simulation} and briefly discuss applications of our framework to
+principal components analysis in section\ref{sec:pac}. The paper ends with a Discussion.
 
@@
 
-\subsection{Notation and Problem Setting}\label{sec:genericSetting}
+\section{Notation and Problem Setting}
+\label{sec:genericSetting}
 
@@
 
-\section{Sequential Inference}\label{sec:sequential}
+\section{Sequential Inference}
+\label{sec:sequential}
 
@@
 
-\citet{gsell2013sequential} propose another stopping rule that controls the FWER, which they called {\em strong stop}. Define
+\citet{gsell2013sequential} propose a stopping rule that controls the FWER, which they called {\em strong stop}. Define
 \[
@@
 \]
+StrongStop uses all of the p-values after step $k$ in its decision to reject at step $k$. 
 Even if $p_k$ is a little larger than $\alpha$, say $\alpha=0.05$ and $p_k=0.06$, StrongStop can still reject $M_{k-1}$ if the next three or four $p$-values are also relatively small.
@@
 \]
+ForwardStop uses the p-values{\it before} step $k$ in its decision to reject at step $k$. 
 If $p_k$ is uniform then  $-\log(1-p_k)$ is an exponential random variable. If all of the null $p$-values are uniform and the others are zero, then only the null ones contribute to the sum, and
@@
 
-\WFcomment{Discuss Rina's stuff.}
+\citet{li2015accumulation} generalize the ForwardStop rule, introducing the family of {\em accumulation tests}, defined as follows.
+Fix a function $h: [0,1] \rightarrow [0,\infty]$ satisfying $\int _{t=0}^1 h(t)dt=1$, called an ``accumulation function''. They define an estimate of the
+false discovery proportion at a cutoff $k$ as
+\begin{equation}
+\widehat{\rm FDP}_h(k)= \frac{\sum_{i=1}^k h(p_i)}{k}
+\end{equation}
+and then define the adaptive cutoff,  for FDR level $\alpha$, to be                                                                                                                 as
+\begin{equation}
+\hat k_h={\rm max} \Bigl\{k\in \{1,2,\ldots n\}; \widehat{\rm FDP}_h(k) \leq \alpha \Bigr\}
+\end{equation}
+The ForwardStop rule falls into this class:  another  choice they study is the ``HingeExp'' function
+\begin{equation}
+h_{\rm HingeExp}(t)= 
+\begin{cases} C\cdot\log\Bigl(\frac{1}{C(1-t)} \Bigr)& \text{for $T> 1-1/C$}\\
+                         0                                         & \text{for $T\leq 1-1/C$}
+                         \end{cases}
+ \end{equation}
+
+Now let ${\rm FP}(k)$ be the actual number of false positives up to and including step $k$.
+They prove that for any  accumulation function $h$ bounded by a constant $C$,
+\begin{equation}
+{\rm E} \Bigl[ \frac{{\rm FP}(\hat k_h)}{C/\alpha+\hat k_h} \Bigr] \leq \alpha
+\end{equation}
+They demonstrate numerically that the HingeExp rule offers superior empirical performance compared to ForwardStop  (and other competitors) in many situations.
+
+Importantly, this class of accumulation tests require that  the p-values $p_i$  are  $U(0,1)$ if the null hypothesis $H_i$ is true, and are independent of the non-null p-values.
+This implies that the class of sequential pvalues proposed in this paper can be used as inputs to these tests.
 
@@
 
-\section{Inference for One Step}\label{sec:singleStep}
+\section{Inference for One Step}
+\label{sec:singleStep}
 
@@
 
-\section{Computation}\label{sec:computation}
+\section{Computation}
+\label{sec:computation}
 
@@
 \end{equation}
+This approximation is exact if the predictors are uncorrelated.
+
 Figure \ref{fig:approxpv} shows an example with $n=50$ and $p=10$ features having pairwise correlation 0.5.
@@
 
-\section{Simulation: Sparse Linear Regression}\label{sec:sparseReg}
+\section{Simulation: Sparse Linear Regression}
+\label{sec:sparseReg}
 
@@
     \hline
-    Method & Stopping Rule & $p_{\text{screen}}$ & $\text{FWER}_{\text{mod}}$ 
+    Method & Stopping Rule & $p_{\text{screen}}$ & $\text{FWER}_{\text{mod}}$ an
     & $\text{FDR}_{\text{mod}}$ 
@@
 
-\section{Simulation: Principal Components Analysis}\label{sec:pca}
+\section{Principal Components Analysis}
+\label{sec:pca}
 
 \WFcomment{Can we do this one?}
+Here we consider model selection for   principal components analysis. In that case we are given a data matrix $X \in \R^{n\times d}$, with which we form a sample %covariance matrix
+\[
+S = \frac{1}{n-1} \sum_{i=1}^n(x_i - \bar x)^2
+\]
+The first $d$ principal component loadings are the first $d$ eigenvectors of $S$, which call $u_1,\ldots, u_d$. These induce a sequence of nested Wishart models:
+\[
+M_0 \sub M_1 \sub \cdots \sub M_d
+\]
+in which
+\begin{equation}
+  M_k:\; (n-1) S \sim W_d\left(\lambda_0 I_d + \sum_{i=1}^k     \lambda_i u_i u_i', \;\;\; n-1\right).
+\end{equation}
+ By successively conditioning on $u_j, j<k$ one can obtain an exact sequential test
+for the complete null hypothesis $H_0: \lambda_j=0 \; \forall j \neq k$.  The computation of the resulting p-values requires numerical integration.
+The details are given iin \citet{choi2014selecting}.
+
+\RTcomment{Jon- add more details?}
 
 \section{Discussion}
+\label{sec:pca}
 

### 2015-07-02 code/python/maxT.R
added unknown sigma tests, removing maxT function as that is now in the forward_step class

@@ -1,8 +1,8 @@
 pdf('maxT_view.pdf')
-D1 = read.table('test1_new.csv', sep=',', header=TRUE)
-D2 = read.table('test2_new.csv', sep=',', header=TRUE)
-D3 = read.table('test3_new.csv', sep=',', header=TRUE)
-D4 = read.table('test4_new.csv', sep=',', header=TRUE)
+D1 = read.table('test1.csv', sep=',', header=TRUE)
+D2 = read.table('test2.csv', sep=',', header=TRUE)
+D3 = read.table('test3.csv', sep=',', header=TRUE)
+D4 = read.table('test4.csv', sep=',', header=TRUE)
 D = rbind(D1,D2,D3,D4)
 write.table(D, "test.csv", row.names=FALSE, col.names=TRUE, sep=',')
 
@@ -10,36 +10,78 @@ par(mfrow=c(2,2))
 
 # mostly signal
 
-plot(ecdf(D$maxT_pvalue_4), main='Step 4')
+plot(ecdf(D$maxT_pvalue_4), main='Step 4', xlim=c(0,1))
 plot(ecdf(D$maxT_identify_pvalue_4), add=TRUE, col='blue')
-plot(ecdf(D$maxT_pvalue_5), main='Step 5')
+plot(ecdf(D$maxT_identify_unknown_pvalue_4), add=TRUE, col='yellow')
+plot(ecdf(D$maxT_unknown_pvalue_4), add=TRUE, col='green')
+plot(ecdf(D$saturated_pvalue_4), add=TRUE, col='red')
+
+plot(ecdf(D$maxT_pvalue_5), main='Step 5', xlim=c(0,1))
 plot(ecdf(D$maxT_identify_pvalue_5), add=TRUE, col='blue')
-plot(ecdf(D$maxT_pvalue_6), main='Step 6')
+plot(ecdf(D$saturated_pvalue_5), add=TRUE, col='red')
+plot(ecdf(D$maxT_identify_unknown_pvalue_5), add=TRUE, col='yellow')
+plot(ecdf(D$maxT_unknown_pvalue_5), add=TRUE, col='green')
+
+plot(ecdf(D$maxT_pvalue_6), main='Step 6', xlim=c(0,1))
 plot(ecdf(D$maxT_identify_pvalue_6), add=TRUE, col='blue')
-plot(ecdf(D$maxT_pvalue_7), main='Step 7')
[…]

### 2015-07-04 paper.tex
rob's changes

@@
 and show how to construct them  in forward stepwise regression, LAR and the ever-active path in $\ell_1$-regularized models.
-In the former case, the test is simply a sequential selective version the max-$t$ test of ~\citet{bujaxxxxmaxT}, based one the theory developed \citet{fithian2014optimal}.
+In the former case, the test is simply a sequential selective version the max-$t$ test of ~\citet{buja2014}, based on the theory developed \citet{fithian2014optimal}.
 
@@
 testing ordered hypothesis using from a sequence of p-values. These method require that the p-values be uniform and independent under the null, and we derive general conditions
-to ensure this property. In Section \ref{sec:single step}  we  show how to construct valid selective $p$-values for a single step in the path and derive the max-t test
+to ensure this property. In Section \ref{sec:singleStep}  we  show how to construct valid selective $p$-values for a single step in the path and derive the max-t test
 for  forward stepwise regression and the Next-$\lambda$ test for the lasso.
@@
 
-We will break the problem into two constituent subproblems. First, Section~\ref{sec:singleStep} will discuss how to obtain selective single-step $p$-values, where $p_k$ is a $p$-value for testing
+First suppose that we have single step p-values for testing
 \[
@@
 \]
-against the alternative that $F\in M_\infty\setminus M_{k-1}$.
-The $p$-values must account for the fact that the models are chosen adaptively.
-In most of our development we will make few assumptions about what form the candidate models take.
-Still, our results are only useful in cases where we can actually construct single-step $p$-values.
-At present, this largely restricts our applications to exponential family models.
-
-Second, Section~\ref{sec:sequential} focuses on the sequential nature of the problem.
+against the alternative that $F\in M_\infty\setminus M_{k-1}$. The $p$-values must account for the fact that the models are chosen adaptively. 
 \citet{gsell2013sequential} and \citet{li2015accumulation} propose stopping rules that operate on the sequence $p_{[d]}$.
 These stopping rules provably control certain error rates, provided that the null $p$-values are mutually independent and independent of the non-null $p$-values.
-We will discuss general conditions under which single-step $p$-values are independent.
-
-\subsubsection{Linear Regression}
+We review these methods in Section~\ref{sec:sequential} .
+
+Hence the task is to  construct sequential p-values which satisfy the above requirements, and we do this in Section \ref{sec:singleStep}.
+We also discuss general conditions under which single-step $p$-values are independent.
+
+\subsection{Linear Regression}
 
@@
 
-\subsubsection{Sparse Parametric Models}\label{sec:genSparse}
+\subsection{Sparse Parametric Models}\label{sec:genSparse}
 
@@
 
-\subsubsection{Path Algorithms}
+\subsection{Path Algorithms}
 
@@
 
-\subsection{Which Null Hypothesis Should We Test?}
+\subsection{``Model-centric'' and ``variable-centric'' null hypotheses}
 \label{sec:whichnull}
@@
 \]
-where $A^\dagger$ is the Moore-Penrose pseudoinverse of the matrix $A$. Write $\proj_E\mu = X_E^\dagger \theta^E$ for the projection of $\mu$ into the column space of $X_E$.
+where $A^\dagger$ is the Moore-Penrose pseudoinverse of the matrix $A$. 
+These are the true ``partial regression coefficients''  for design matrix $X_E$.
+ Write $\proj_E\mu = X_E^\dagger \theta^E$ for the projection of $\mu$ into the column space of $X_E$.
 
@@
 We will assume throughout that the hypotheses are nested with $H_{k-1} \sub H_k$.
-Section~\ref{sec:orderedProposals} outlines several proposals for ordered testing and discusses what error rates they control and under what conditions.
-
-We focus on proposals from \citet{gsell2013sequential} and \citet{li2015accumulation}, all of which assume that the null hypotheses $H_{1}, \ldots, H_{d}$ are fixed, and that the null $p$-values are independent of each other and of the non-null $p$-values.
+The next Section outlines several proposals for ordered testing and discusses what error rates they control and under what conditions.
+
+\subsection{Stopping Rules for Ordered Testing}
+\label{sec:orderedProposals}
+
+\subsubsection{BasicStop}
+The most obvious procedure is simply to reject at each step until the first time that $p_k > \alpha$, which we can formalize as
+\[
+\hk_B(Y) = \min\left\{k \in \{1,\ldots,m\} :\;
+  p_k > \alpha\right\} - 1
+\]
+We will call this procedure {\em BasicStop}. 
+It is discussed in \citet{marcus1976}.
+
+If $k_0$ is fixed, then it is clear that $\hk_b$ controls the FWER at level $\alpha$ provided that $p_{k}$ is a valid $p$-value for each $k>k_0$. Then, 
+\[
+\P(V>0) = \P(p_{k_0+1} \leq \alpha) \leq \alpha.
+\]
+
+Selectively valid single-step $p$-values do not necessarily guarantee FWER control when $k_0$ is random.
+In that case, we need a bit more, but it is sufficient that we have type I error control for $p_k$ conditional on $k_0=k-1$; i.e., not just conditional on $M_{k-1}$ being a correct model, but given that $M_{k-1}$ is the {\em first} correct model.
+
+\subsubsection{StrongStop}
+
+\citet{gsell2013sequential} propose a stopping rule that controls the FWER, which they called {\em strong stop}. Define
+\[
+  \hk_{S}(Y) = \max\left\{k \in \{1,\ldots,d\} :\;
+    \exp\left(\sum_{i=k}^d \frac{\log p_i}{i}\right) 
+    \leq \frac{\alpha k}{d}\right\}
+\]
+StrongStop uses all of the p-values after step $k$ in its decision to reject at step $k$. 
+Even if $p_k$ is a little larger than $\alpha$, say $\alpha=0.05$ and $p_k=0.06$, StrongStop can still reject $M_{k-1}$ if the next three or four $p$-values are also relatively small.
+
+\citet{gsell2013sequential} show that if the completion index $k_0$ is fixed, then $\hk_S$ controls the FWER at level $\alpha$, as long as the null $p$-values are independent given the non-null ones.
+
+\subsubsection{ForwardStop}
+
+\citet{gsell2013sequential} propose another stopping rule, {\em forward stop}, that controls the FDR:
+\[
+  \hk_{F}(Y) = \max\left\{k \in \{1,\ldots,d\} :\;
+    -\frac{1}{k}\sum_{i=1}^k \log(1-p_i) \leq \alpha\right\}
+\]
+ForwardStop uses the p-values {\it before} step $k$ in its decision to reject at step $k$. 
+If $p_k$ is uniform then  $-\log(1-p_k)$ is an exponential random variable. If all of the null $p$-values are uniform and the others are zero, then only the null ones contribute to the sum, and
+\[
+\widehat{\text{FDR}}_k = -\frac{1}{k}\sum_{i=1}^k \log(1-p_i)
+\]
+is a Gamma random variable with mean $V/k$ and variance $V/k^2$. We stop the last time the estimate of FDR is less than $\alpha$.
+
+\subsubsection{Accumulation Tests and HingeExp}
+
+\citet{li2015accumulation} generalize the ForwardStop rule, introducing the family of {\em accumulation tests}, defined as follows.
+Fix a function $h: [0,1] \rightarrow [0,\infty]$ satisfying $\int _{t=0}^1 h(t)dt=1$, called an ``accumulation function''. They define an estimate of the
+false discovery proportion at a cutoff $k$ as
+\begin{equation}
+\widehat{\rm FDP}_h(k)= \frac{\sum_{i=1}^k h(p_i)}{k}
+\end{equation}
+and then define the adaptive cutoff,  for FDR level $\alpha$, to be                                                                                                                 as
+\begin{equation}
+\hat k_h={\rm max} \Bigl\{k\in \{1,2,\ldots n\}; \widehat{\rm FDP}_h(k) \leq \alpha \Bigr\}
+\end{equation}
+The ForwardStop rule falls into this class:  another  choice they study is the ``HingeExp'' function
+\begin{equation}
+h_{\rm HingeExp}(t)= 
+\begin{cases} C\cdot\log\Bigl(\frac{1}{C(1-t)} \Bigr)& \text{for $T> 1-1/C$}\\
+                         0                                         & \text{for $T\leq 1-1/C$}
+                         \end{cases}
+ \end{equation}
+
+Now let ${\rm FP}(k)$ be the actual number of false positives up to and including step $k$.
+They prove that for any  accumulation function $h$ bounded by a constant $C$,
+\begin{equation}
+{\rm E} \Bigl[ \frac{{\rm FP}(\hat k_h)}{C/\alpha+\hat k_h} \Bigr] \leq \alpha
+\end{equation}
+They demonstrate numerically that the HingeExp rule offers superior empirical performance compared to ForwardStop  (and other competitors) in many situations.
+
+Importantly, this class of accumulation tests require that  the p-values $p_i$  are  $U(0,1)$ if the null hypothesis $H_i$ is true, and are independent of the non-null p-values.
+This implies that the class of sequential pvalues proposed in this paper can be used as inputs to these tests.
+
+\subsection{Random hypotheses}
+The forementioned proposals of  \citet{gsell2013sequential} and \citet{li2015accumulation}, all assume that the null hypotheses $H_{1}, \ldots, H_{d}$ are fixed, and that the null $p$-values are independent of each other and of the non-null $p$-values.
 Specifically, they assume that
@@
 
-In our problem the null hypotheses at each step are random and thus so is the completion index $k_0(Y)$. We will thus require {\em conditional} independence of null $p$-values given $k_0$:
+in our problem the null hypotheses at each step are random and thus so is the completion index $k_0(Y)$. We will thus require {\em conditional} independence of null $p$-values given $k_0$:
 \begin{equation}\label{eq:indepCond_random_k0}
@@
 
-\subsection{Stopping Rules for Ordered Testing}\label{sec:orderedProposals}
-
-\begin{comment}
-\subsubsection{BasicStop}
-The most obvious procedure is simply to reject at each step until the first time that $p_k > \alpha$, which we can formalize as
-\[
-\hk_B(Y) = \min\left\{k \in \{1,\ldots,m\} :\;
-  p_k > \alpha\right\} - 1
-\]
-We will call this procedure {\em BasicStop}. \WFcomment{Is there a better name for this? Can we find it in the literature?}
-
-If $k_0$ is fixed, then it is clear that $\hk_b$ controls the FWER at level $\alpha$ provided that $p_{k}$ is a valid $p$-value for each $k>k_0$. Then, 
-\[
-\P(V>0) = \P(p_{k_0+1} \leq \alpha) \leq \alpha.
-\]
-
-Selectively valid single-step $p$-values do not necessarily guarantee FWER control when $k_0$ is random.
-In that case, we need a bit more, but it is sufficient that we have type I error control for $p_k$ conditional on $k_0=k-1$; i.e., not just conditional on $M_{k-1}$ being a correct model, but given that $M_{k-1}$ is the {\em first} correct model.
-\end{comment}
-
-\subsubsection{StrongStop}
-
-\citet{gsell2013sequential} propose a stopping rule that controls the FWER, which they called {\em strong stop}. Define
-\[
-  \hk_{S}(Y) = \max\left\{k \in \{1,\ldots,d\} :\;
-    \exp\left(\sum_{i=k}^d \frac{\log p_i}{i}\right) 
-    \leq \frac{\alpha k}{d}\right\}
-\]
-StrongStop uses all of the p-values after step $k$ in its decision to reject at step $k$. 
-Even if $p_k$ is a little larger than $\alpha$, say $\alpha=0.05$ and $p_k=0.06$, StrongStop can still reject $M_{k-1}$ if the next three or four $p$-values are also relatively small.
-
-\citet{gsell2013sequential} show that if the completion index $k_0$ is fixed, then $\hk_S$ controls the FWER at level $\alpha$, as long as the null $p$-values are independent given the non-null ones.
-
-\subsubsection{ForwardStop}
-
-\citet{gsell2013sequential} propose another stopping rule, {\em forward stop}, that controls the FDR:
-\[
-  \hk_{F}(Y) = \max\left\{k \in \{1,\ldots,d\} :\;
-    -\frac{1}{k}\sum_{i=1}^k \log(1-p_i) \leq \alpha\right\}
-\]
-ForwardStop uses the p-values{\it before} step $k$ in its decision to reject at step $k$. 
-If $p_k$ is uniform then  $-\log(1-p_k)$ is an exponential random variable. If all of the null $p$-values are uniform and the others are zero, then only the null ones contribute to the sum, and
-\[
-\widehat{\text{FDR}}_k = -\frac{1}{k}\sum_{i=1}^k \log(1-p_i)
-\]
-is a Gamma random variable with mean $V/k$ and variance $V/k^2$. We stop the last time the estimate of FDR is less than $\alpha$.
-
-\subsubsection{Accumulation Tests and HingeExp}
-
-\citet{li2015accumulation} generalize the ForwardStop rule, introducing the family of {\em accumulation tests}, defined as follows.
-Fix a function $h: [0,1] \rightarrow [0,\infty]$ satisfying $\int _{t=0}^1 h(t)dt=1$, called an ``accumulation function''. They define an estimate of the
-false discovery proportion at a cutoff $k$ as
-\begin{equation}
-\widehat{\rm FDP}_h(k)= \frac{\sum_{i=1}^k h(p_i)}{k}
-\end{equation}
-and then define the adaptive cutoff,  for FDR level $\alpha$, to be                                                                                                                 as
-\begin{equation}
-\hat k_h={\rm max} \Bigl\{k\in \{1,2,\ldots n\}; \widehat{\rm FDP}_h(k) \leq \alpha \Bigr\}
-\end{equation}
-The ForwardStop rule falls into this class:  another  choice they study is the ``HingeExp'' function
-\begin{equation}
-h_{\rm HingeExp}(t)= 
-\begin{cases} C\cdot\log\Bigl(\frac{1}{C(1-t)} \Bigr)& \text{for $T> 1-1/C$}\\
-                         0                                         & \text{for $T\leq 1-1/C$}
-                         \end{cases}
- \end{equation}
-
-Now let ${\rm FP}(k)$ be the actual number of false positives up to and including step $k$.
-They prove that for any  accumulation function $h$ bounded by a constant $C$,
-\begin{equation}
-{\rm E} \Bigl[ \frac{{\rm FP}(\hat k_h)}{C/\alpha+\hat k_h} \Bigr] \leq \alpha
-\end{equation}
-They demonstrate numerically that the HingeExp rule offers superior empirical performance compared to ForwardStop  (and other competitors) in many situations.
-
-Importantly, this class of accumulation tests require that  the p-values $p_i$  are  $U(0,1)$ if the null hypothesis $H_i$ is true, and are independent of the non-null p-values.
-This implies that the class of sequential pvalues proposed in this paper can be used as inputs to these tests.
-
-\subsection{Conditions for Independence}\label{sec:pValsIndep}
-
-The single-step $p$-value $p_k$ is selectively valid for testing $H_k:\; F \in M_{k-1}(Y)$ against $M_\infty \setminus M_{k-1}$ if, for all $F\in M_\infty$,
+\section{Conditions for Independence of p-values}\label{sec:pValsIndep}
+
+We say that the  single-step $p$-value $p_k$ is selectively valid for testing $H_k:\; F \in M_{k-1}(Y)$ against $M_\infty \setminus M_{k-1}$ if, for all $F\in M_\infty$,
 \[
@@
 
-Two specific examples are worth discussing in detail.
-First, consider the {\em selective max-$t$ test}, a generalization of the max-$t$ test of ~\citet{bujaxxxxmaxT}.
-Controlling for $E$, the $z$-statistic for coefficient $j\notin E$ is
+Two specific examples are worth discussing in detail. First, consider the {\em selective max-$t$ test}, a generalization of the max-$t$ test described in 
+~\citet{buja2014}. Controlling for $E$, the $z$-statistic for coefficient $j\notin E$ is
 \begin{equation}
@@
 
-\begin{figure}
+\begin{figure}[h]
   \centering
   % source code: ??? for simulation, sparseSim.R for plotting
-  \includegraphics[width=.8\textwidth]{figs/simulation_snr_5_alpha_05_null_false.pdf}
-  \caption{\em CDFs of saturated-model (black), selected-model (red), and nominal (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on testing a false null hypothesis at step $k$.
-The selected-model test is much more powerful than the saturated-model test.
-The nominal test appears powerful, but is not a correct $p$-value.}
+  \includegraphics[width=1\textwidth]{figs/simulation_snr_5_alpha_05_null_false.pdf}
+  \caption{\em  CDFs of nominal (black), MaxT (red), MaxT with unknown $\sigma$ (green) and  saturated-model (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on testing a false null hypothesis at step $k$.
+The MaxT approaches are much more powerful than the saturated-model test.
+The nominal test appears to be powerful,
+  but is not $U(0,1)$ under the null, as shown below in Figure \ref{fig:simulation_null_true}.}
   \label{fig:simulation_null_false}
@@
 
-\begin{figure}
+\begin{figure}[h]
   \centering
   % source code: ??? for simulation, sparseSim.R for plotting
-  \includegraphics[width=.8\textwidth]{figs/simulation_snr_5_alpha_05_null_true.pdf}
-  \caption{\em CDFs of saturated-model (black), selected-model (red), and nominal (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on testing a true null hypothesis at step $k$.
-The selected-model test is much more powerful than the saturated-model test.
-The nominal test is badly anti-conservative.}
+  \includegraphics[width=1\textwidth]{figs/simulation_snr_5_alpha_05_null_true.pdf}
+  \caption{\em  CDFs of nominal (black), MaxT (red), MaxT with unknown $\sigma$ (green) and  saturated-model (blue) p-values in the simulation of Section~\ref{sec:sparseReg}, conditional on testing a true null hypothesis at step $k$.
+The nominal test is badly anti-conservative, while all of the other methods  show uniform p-values as desired.}
   \label{fig:simulation_null_true}
@@
 
-\begin{figure}
+\begin{figure}[h]
   \centering
   % source code: ??? for simulation, sparseSim.R for plotting
-  \includegraphics[width=.8\textwidth]{figs/simulation_snr_5_alpha_05_noise_var.pdf}
-  \caption{\em CDFs of saturated-model (black), selected-model (red), and nominal (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on the event that the variable added at step $k$ is a noise variable in the full model.
+  \includegraphics[width=1\textwidth]{figs/simulation_snr_5_alpha_05_noise_var.pdf}
+  \caption{\em CDFs of nominal (black), MaxT (red), MaxT with unknown $\sigma$ (green) and  saturated-model (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on the event that the variable added at step $k$ is a noise variable in the full model.
 Here, none of the methods produce uniform $p$-values.
@@
 % Fri May  1 09:29:28 2015
+%\begin{table}[ht]
+ % \centering
+ % \begin{tabular}{llcccc}
+  %  \hline
+  %  Method & Stopping Rule & $p_{\text{screen}}$ & $\text{FWER}_{\text{mod}}$ an
+ %   & $\text{FDR}_{\text{mod}}$ 
+  %  & $\text{FDR}_{\text{var}}$ \\ 
+ %   \hline
+    %Selected & Basic & .290 & .000 & .002 & .039 \\ 
+   % Selected & Forward & .559 & .027 & .020 & .066 \\ 
+   % Selected & Strong & .041 & .014 & .008 & .042 \\ 
+  %  \hline
+   % Saturated & Basic & .000 & .000 & .000 & .028 \\ 
+  %  Saturated & Forward & .014 & .000 & .000 & .030 \\ 
+  %  Saturated & Strong & .000 & .000 & .000 & .032 \\ 
+  %  \hline
+  %  Knockoffs & & .000 & --- & --- & .231 \\ 
+  %  \hline
+ % \end{tabular}
+ % \caption{\WFcomment{Comment on results once they are correct. These numbers are wrong right now.}}
+%\end{table}
+
 \begin{table}[ht]
 \centering
-  \begin{tabular}{llcccc}
+\begin{tabular}{rrrrr}
   \hline
-    Method & Stopping Rule & $p_{\text{screen}}$ & $\text{FWER}_{\text{mod}}$ an
-    & $\text{FDR}_{\text{mod}}$ 
-    & $\text{FDR}_{\text{var}}$ \\ 
+  Method, Stopping rule& $p_{\text screen}$ & FWER$_{\text mod}$ & FDR$_{\text mod}$ &FDR$_{\text var}$ \\
+% & \$p\_\{$\backslash$text\{screen\}\} & \$$\backslash$text\{FWER\}\_\{$\backslash$text\{model\}\}\$ & \$$\backslash$text\{FDR\}\_\{$\backslash$text\{model\}\}\$ & \$$\backslash$text\{FDR\}\_\{$\backslash$text\{variable\}\}\$ \\
   \hline
-    Selected & Basic & .290 & .000 & .002 & .039 \\ 
-    Selected & Forward & .559 & .027 & .020 & .066 \\ 
-    Selected & Strong & .041 & .014 & .008 & .042 \\ 
+Nominal, Simple & 0.832 & 0.252 & 0.097 & 0.171 \\ 
+  Nominal, Forward & 0.931 & 0.896 & 0.345 & 0.426 \\ 
+  Nominal, Strong & 0.003 & 0.000 & 0.000 & 0.035 \\ 
   \hline
-    Saturated & Basic & .000 & .000 & .000 & .028 \\ 
-    Saturated & Forward & .014 & .000 & .000 & .030 \\ 
-    Saturated & Strong & .000 & .000 & .000 & .032 \\ 
+  MaxT, Simple & 0.369 & 0.000 & 0.002 & 0.042 \\ 
+  MaxT, Forward & 0.668 & 0.038 & 0.026 & 0.080 \\ 
+  MaxT, Strong & 0.025 & 0.006 & 0.004 & 0.038 \\ 
   \hline
-    Knockoffs & & .000 & --- & --- & .231 \\ 
+  MaxT-Identify, Simple & 0.404 & 0.000 & 0.002 & 0.042 \\ 
+  MaxT-Identify, Forward & 0.688 & 0.039 & 0.026 & 0.082 \\ 
+  MaxT-Identify, Strong & 0.024 & 0.003 & 0.002 & 0.037 \\ 
+  \hline
+  Saturated, Simple & 0.000 & 0.000 & 0.000 & 0.029 \\ 
+  Saturated, Forward & 0.013 & 0.001 & 0.000 & 0.030 \\ 
+  Saturated, Strong & 0.000 & 0.000 & 0.000 & 0.033 \\ 
+  \hline
+  Knockoffs & 0.252 &  &  & 0.103 \\ 
    \hline
 \end{tabular}
-  \caption{\WFcomment{Comment on results once they are correct. These numbers are wrong right now.}}
 \end{table}
 
+We see that 1) strongStop and forwardStop control FWER and modelwise FDR, respectively, using p-values from MaxT or MaxT-identify, as predicted by theory
+2) the nominal p-values do not lead to control of FWER or FDR, as expected, 3) the pvalues from the saturated model show FWER and modelwise-FDR control,
+although not guaranteed by theory, 4) the MaxT rules show higher probabilities of selecting the true model, especially using forwardStop, 5) the knockoff method
+does not yields variable-wise FDR control, as it should.
+5)
 \section{Principal Components Analysis}

### 2015-07-04 code/python/compare_methods.py
added options for more sparsity levels, random signs. also just returned uniform p-values after 30 steps... should customize this choice though.

@@ -21,17 +21,21 @@ def summary(variables, pvalues, active, rule, alpha):
     screen = R > completion_idx
     return R, V_var, V_model, screen
 
-def simulate(n=100, p=40, rho=0.3, snr=5,
+def simulate(n=100, p=200, rho=0.3, snr=5,
              do_knockoff=False,
              full_results={},
+             s=10,
              alpha=0.05,
              maxstep=np.inf,
-             compute_maxT_identify=True):
+             compute_maxT_identify=True,
+             random_signs=True):
 
     X, y, _, active, sigma = instance(n=n,
                                       p=p,
                                       rho=rho,
-                                      snr=snr)
+                                      snr=snr,
+                                      s=s,
+                                      random_signs=random_signs)
     full_results.setdefault('n', []).append(n)
     full_results.setdefault('p', []).append(p)
     full_results.setdefault('rho', []).append(rho)
@@ -54,7 +58,7 @@ def run(y, X, sigma, active,
         ndraw=8000):
 
     n, p = X.shape
-    results, FS = compute_pvalues(y, X, sigma, maxstep=maxstep,
+    results, FS = compute_pvalues(y, X, active, sigma, maxstep=maxstep,
                                   compute_maxT_identify=compute_maxT_identify,
                                   burnin=burnin,
                                   ndraw=ndraw)
@@ -151,5 +155,5 @@ Run a batch of simulations.
     args = parser.parse_args()
     if args.maxstep < 0:
         args.maxstep = np.inf
-    batch(args.outbase, args.nsim, do_knockoff=True,
+    batch(args.outbase, args.nsim, do_knockoff=False, # True,
           maxstep=args.maxstep)

### 2015-07-04 code/python/compute_pvalues.py

@@ -7,7 +7,7 @@ from selection.distributions.discrete_family import discrete_family
 from selection.constraints.affine import gibbs_test
 from scipy.stats import norm as ndist
 
-def compute_pvalues(y, X, sigma=1., maxstep=np.inf,
+def compute_pvalues(y, X, active_set, sigma=1., maxstep=np.inf,
                     compute_maxT_identify=True,
                     burnin=2000,
                     ndraw=8000,
@@ -21,6 +21,11 @@ def compute_pvalues(y, X, sigma=1., maxstep=np.inf,
 
     X : np.float((n, p))
         The data, in the model $y = X\beta$
+
+    active_set : []
+        The true active set. For steps beyond this
+        the selected model methods can (for the purposes
+        of simulation) just draw np.random.sample()
 
     sigma : np.float
         Standard deviation of the gaussian distribution :
@@ -55,37 +60,52 @@ def compute_pvalues(y, X, sigma=1., maxstep=np.inf,
     iter(FS_identity_U); iter(FS_maxT_U)
     results = []
 
+    completed = False
     for i in range(min([n, p, maxstep])):
 
-        # take a step of FS
+        screened = set(active_set).issubset(FS_maxT.variables) or (i > 30) # can't be any power way out there... i figure
 
-        pval_maxT = FS_maxT.next(compute_pval=True,
-                                 use_identity=False,
-                                 ndraw=ndraw,
-                                 burnin=burnin)
-        
-        pval_maxT_U = FS_maxT_U.next(compute_pval=True,
+        if not screened: 
+
+            # take a step of FS
+
+            pval_maxT = FS_maxT.next(compute_pval=True,
                                      use_identity=False,
                                      ndraw=ndraw,
-                                     burnin=burnin,
-                                     sigma_known=False)
-        
-        if compute_maxT_identify:
+                                     burnin=burnin)
 
-            pval_maxT_identify = FS_identity.next(compute_pval=True,
-                                                  use_identity=True,
-                                                  ndraw=ndraw,
-                                                  burnin=burnin)
+            pval_maxT_U = FS_maxT_U.next(compute_pval=True,
+                                         use_identity=False,
+                                         ndraw=ndraw,
+                                         burnin=burnin,
+                                         sigma_known=False)
 
-            pval_maxT_identify_U = FS_identity_U.next(compute_pval=True,
+            if compute_maxT_identify:
+
+                pval_maxT_identify = FS_identity.next(compute_pval=True,
                                                       use_identity=True,
[…]

### 2015-07-04 code/python/compare_methods.py
returning to default parameters

@@ -21,14 +21,14 @@ def summary(variables, pvalues, active, rule, alpha):
     screen = R > completion_idx
     return R, V_var, V_model, screen
 
-def simulate(n=100, p=200, rho=0.3, snr=5,
+def simulate(n=100, p=40, rho=0.3, snr=5,
              do_knockoff=False,
              full_results={},
-             s=10,
+             s=7,
              alpha=0.05,
              maxstep=np.inf,
              compute_maxT_identify=True,
-             random_signs=True):
+             random_signs=False):
 
     X, y, _, active, sigma = instance(n=n,
                                       p=p,

### 2015-07-04 code/R/sparseSim2.R
rob added sparseSim2.R

@@ -1,0 +1,93 @@
+## plots p=40
+
+simulation.data <- read.csv("../snr_5_alpha_05.csv")
+names(simulation.data)
+
+p=40
+s=7
+
+
+pdf("../../figs/simulation_snr_5_alpha_05_signal_var.pdf",
+    width=8,height=5)
+par(xaxs="i",mar=c(2.1,3.1,3.1,0.7), mfrow=c(2,5))
+for(k in 1:10) {
+  which.signal <- which(simulation.data[[paste0("variable_selected_",k)]] <=s)
+  main <- bquote(p*.(k))
+  plot(ecdf(simulation.data[[paste0("nominal_pvalue_",k)]][which.signal]),xlim=c(0,1),
+       main=main,xlab="",ylab="")
+del=runif(length(which.signal),0,.01)
+  plot(ecdf(simulation.data[[paste0("maxT_pvalue_",k)]][which.signal]+del),col="red",add=T)
+    plot(ecdf(simulation.data[[paste0("maxT_unknown_pvalue_",k)]][which.signal]+del),col="green",add=T)
+  plot(ecdf(simulation.data[[paste0("saturated_pvalue_",k)]][which.signal]),col="blue",add=T)
+  abline(0,1,lty=3,col="gray")
+if(k==1) legend (list(x=c(.3,1),y=c(.25,0)),c("Nominal","MaxT","Maxt/Unkn","Saturated"),col=c("black","red","green","blue"),lty=1,cex=.7)
+}
+dev.off()
+
+#null is true
+
+
+##
+pdf("../../figs/simulation_snr_5_alpha_05_null_true.pdf",
+    width=8,height=5)
+par(xaxs="i",mar=c(2.1,3.1,3.1,0.7), mfrow=c(2,5))
+
+
+for(k in (s+1):(s+10)) {
+
+which.null <- which(k> simulation.data$completion_idx+1)  # correction for zero-based
+   main <- bquote(p*.(k))
+  plot(ecdf(simulation.data[[paste0("nominal_pvalue_",k)]][which.null]),xlim=c(0,1),
+       main=main,xlab="",ylab="")
+plot(ecdf(simulation.data[[paste0("maxT_pvalue_",k)]][which.null]),col="red",add=T)
+  plot(ecdf(simulation.data[[paste0("maxT_unknown_pvalue_",k)]][which.null]),col="green",add=T)
[…]

### 2015-07-04 code/R/sparseSim2.R
rob added sparseSim2.R

@@ -12,7 +12,7 @@ pdf("../../figs/simulation_snr_5_alpha_05_signal_var.pdf",
 par(xaxs="i",mar=c(2.1,3.1,3.1,0.7), mfrow=c(2,5))
 for(k in 1:10) {
   which.signal <- which(simulation.data[[paste0("variable_selected_",k)]] <=s)
-  main <- bquote(p*.(k))
+  main <- paste("Step ",as.character(k),sep="")
   plot(ecdf(simulation.data[[paste0("nominal_pvalue_",k)]][which.signal]),xlim=c(0,1),
        main=main,xlab="",ylab="")
 del=runif(length(which.signal),0,.01)
@@ -36,7 +36,7 @@ par(xaxs="i",mar=c(2.1,3.1,3.1,0.7), mfrow=c(2,5))
 for(k in (s+1):(s+10)) {
 
 which.null <- which(k> simulation.data$completion_idx+1)  # correction for zero-based
-   main <- bquote(p*.(k))
+  main <- paste("Step ",as.character(k),sep="")
   plot(ecdf(simulation.data[[paste0("nominal_pvalue_",k)]][which.null]),xlim=c(0,1),
        main=main,xlab="",ylab="")
 plot(ecdf(simulation.data[[paste0("maxT_pvalue_",k)]][which.null]),col="red",add=T)
@@ -48,13 +48,32 @@ if(k==(s+1)) legend (list(x=c(.3,1),y=c(.25,0)),c("Nominal","MaxT","Maxt/Unkn","
 dev.off()
 
 
+pdf("../../figs/simulation_snr_5_alpha_05_null_false.pdf",
+    width=8,height=5)
+par(xaxs="i",mar=c(2.1,3.1,3.1,0.7),
+    mfrow=c(2,5))
+for(k in 1:10) {
+  which.nonnull <- which(simulation.data$completion_idx >= k)
+   main <- paste("Step ",as.character(k),sep="")
+  del=runif(length(which.nonnull),0,.01)
+  plot(ecdf(simulation.data[[paste0("nominal_pvalue_",k)]][which.nonnull]+del),xlim=c(0,1),
+       main=main,xlab="",ylab="")
+plot(ecdf(simulation.data[[paste0("maxT_pvalue_",k)]][which.nonnull]+del),col="red",add=T)
[…]

### 2015-07-04 code/R/sparseSim.R
rob added sparseSim2.R

@@ -1,11 +1,11 @@
 library(xtable)
 
-simulation.data <- read.csv("snr_5_alpha_05.csv")
+simulation.data <- read.csv("../snr_5_alpha_05.csv")
 names(simulation.data)
 
 
 methods <- c(paste(rep(c("simple", "forward", "strong"),2),
-                   rep(c("selected","saturated"),each=3),
+                   rep(c("MaxT","saturated"),each=3),
                   sep="_"),
              "knockoffs")
 error.types <- c("pscreen","FWER.mod","FDR.mod","FDR.var")
@@ -13,7 +13,7 @@ error.rates <- matrix(NA,7,4,
                       dimnames=list(methods,error.types))
 for(i in 1:6) {
   R <- simulation.data[[paste0(methods[i],"_R")]]
-  k0 <- simulation.data[["completion_index"]]
+  k0 <- simulation.data[["completion_idx"]]
   V.mod <- pmax(0,R - k0)
   V.var <- simulation.data[[paste0(methods[i],"_V_var")]]
   error.rates[i,"pscreen"] <- mean(R >= k0)
@@ -34,8 +34,10 @@ simulation.data$simple_saturated_V_model[46]
 
 
 error.rates["knockoffs", "pscreen"] <- with(simulation.data, mean(knockoff_R - knockoff_V == 7))
+#error.rates["knockoffs", "FDR.var"] <- with(simulation.data, mean(knockoff_V / pmax(knockoff_R,1)))
+o=simulation
 error.rates["knockoffs", "FDR.var"] <- with(simulation.data, mean(knockoff_V / pmax(knockoff_R,1)))
-                                           
+
 ## Fix this later...
 colnames(error.rates) <- c("$p_{\\text{screen}}", 
                            "$\\text{FWER}_{\\text{model}}$",
@@ -106,10 +108,10 @@ par(xaxs="i",mar=c(2.1,3.1,3.1,0.7),
 for(k in 1:10) {
   which.noise <- which(simulation.data[[paste0("var_",k)]] >= 7)
   main <- bquote(p*.(k))
-  plot(ecdf(simulation.data[[paste0("saturated_",k)]][which.noise]),xlim=c(0,1),
+  plot(ecdf(simulation.data[[paste0("saturated_pvalue_",k)]][which.noise]),xlim=c(0,1),
        main=main,xlab="",ylab="")
[…]

### 2015-07-07 paper.tex
Rob edited paper

@@
 \newcommand{\WFcomment}[1]{{\color{red}{(WF: \bf \sc #1) }}}
-\newcommand{\RTcomment}[1]{{\color{red}{(RT: \bf \sc #1) }}}
+\newcommand{\RTcomment}[1]{{\color{green}{(RT: \bf \sc #1) }}}
 
@@
 \WFcomment{should probably end with $M_{d-1}(Y)$. need to fix this   throughout paper.}
+\RTcomment{Will- be more specific here---   A parametric model  with specified nonzero support?}
+
 By ``model,'' we mean a family of probability distributions for $Y$. Without loss of generality, we can take $M_\infty$ as the union of all models under consideration.
@@
 
-\section{Sequential Inference}
+\section{Sequential Inference from ordered p-values}
 \label{sec:sequential}
@@
   problem with $\tilde p_k \sim \delta_{q_k}$ for $k\leq k_0$ and 
-  $\tilde p_k \sim \text{Unif}[0,1]$ for $k > k_0$. Then, for the
+  $\tilde p_k \sim \text{Unif}[0,1]$ for $k > k_0$. 
+  \RTcomment{Will; what are the qk?  you mean we just choose and fix some values qk? Say so.}Then, for the
   fixed hypotheses 
@@
 
-\section{Conditions for Independence of p-values}\label{sec:pValsIndep}
-
+\section{Construction of Independence of p-values}\label{sec:pValsIndep}
+
+From the previous section, we see that our task is to construct p-values with the {\it independent on nulls} property (\ref{eq:indepCond_random_k0}).
+In this section we develop a set of sufficient conditions for such a construction. Before doing so, we give an example to motivate
+the general theory.
+
+\medskip
+\begin{example}[The selective max-Z test]
+Consider the forward stepwise path (\ref{eq:forwardDef_start}---\ref{eq:forwardDef_end}) in the case of linear regression, that is 
+$\ell(\theta,Y)=||Y-X\theta||^2$.  Assume that the error variance $\sigma^2$ is known.
+ Given an active set of variables $E$ in our current model $M$, the partial regression coefficients for predictors not in $E$ are
+\begin{equation}
+  z_{j,E}(Y) = \frac{(\proj_E^\perp X_j)'Y}
+  {\sigma\|\proj_{E}^\perp X_j\|_2}.
+\end{equation}
+The max-$z$-statistic for active set $E$  is
+\begin{equation}\label{eq:maxT}
+  U_E(Y) = \max_{j \notin E} \left|t_{j,E}(Y)\right|,
+\end{equation}
+and we use $U_{E_{k-1}}(Y)$ as our test statistic at step $k$. We compute a p-value for this test
+using the conditional law 
+\begin{equation}\label{eq:selModel_marg0}
+\L\left(X_{j_k}'Y \mid E_{k-1}, \; X_{E_{k-1}}'Y\right),
+\end{equation}
+If we reject the hypothesis, we add to $E$ the predictor that maximized $U_E(Y)$ and continue.
+This is the procedure that used to obtain the p-values in Table \ref{tab:diab}  \RTcomment{ Rob- make sure you fix table 1 to make the correct computation}
+
+Happily,  it turns that the  p-values from this procedure have exactly the desired {\it independent on nulls} property~\eqref{eq:indepCond_random_k0}.
+This happens because of some special characteristics of the construction: 
+\begin{description}
+\item{(a)}  the inner products $X_E'Y$ are complete sufficient statistics for the regression
+parameters in each model $M$,
+\item{(b)} knowing only the  model $M$ at step $k$ and its sufficient statistics $X'_EY$, we can compute the entire subpath
+of models from the steps 1 to $k-1$. 
+\end{description}
+ In other words, the inner products $X'_EY$ contains  all of the cumulative information used in the selections from
+steps 1 to $k-1$. This in turn results from    the choice  of test statistic $U_E(Y)$  and the law (\ref{eq:selModel_marg0}) used to calibrate it.
+We call (b)  the {\it Subpath sufficiency principle} (SSP) in Section \ref{sec:SSP}  below.
+In contrast, the {\it saturated model} p-values use a law different from (\ref{eq:selModel_marg0}) to compute p-values, one that uses
+information not contained in $X'_EY$. It thus violates the SSP property and hence does not yield p-values that are independent on nulls.
+We discuss the regression setting in more detail in Section \ref{sec:selective-reg}.  $\square$
+\end{example}
+
+\medskip
+
+We now present a general framework for these ideas, culminating in our theorem \ref{thm:suffCond} at the end of this section below.
 We say that the  single-step $p$-value $p_k$ is selectively valid for testing $H_k:\; F \in M_{k-1}(Y)$ against $M_\infty \setminus M_{k-1}$ if, for all $F\in M_\infty$,
@@
 
-\subsubsection{The Sufficient Filtration}\label{sec:suffFilt}
+\subsection{The Sufficient Filtration}\label{sec:suffFilt}
 
@@
 \]
+\RTcomment{Will- define $M_{[k-1]}$..  it is  the collection $\{ M_1 M_2 \ldots M_{k-1}\}$?}
 We will suppress the superscript and use $\sF_k$ to mean $\sF_k^{\text{suff}}$ for the remainder of the article.
@@
 
-\subsubsection{The Subpath Sufficiency Principle}\label{sec:SSP}
+\subsection{The Subpath Sufficiency Principle}\label{sec:SSP}
 
@@
 
-\section{Inference for One Step}
+\section{Inference for one step in the path}
 \label{sec:singleStep}
@@
 \subsection{Selective $p$-Values in Regression}
-
+\label{sec:selective-reg}
 In linear regression, there are several types of selective test that we might contemplate performing at step $k$.
@@
 \subsubsection{Saturated-Model Tests}
-
 Many of the selective tests for linear regression use the saturated-model framework.
@@
 As a result, saturated-model $p$-values constructed as in~\eqref{eq:satModel} are in general {\em not} independent on nulls.
-
+?Collect
 \subsubsection{Selected-Model Tests}
@@
 Happily, it turns out that the conditioning polytope can be expressed with only $2p$ constraints in forward stepwise regression, as we show here.
-This increases the speed of both accept/reject and hit andRun sampling for our problem.
+This increases the speed of both accept/reject and hit and Run sampling for our problem.
 
@@
 
+We plan to distribute  open source software, in both the R and Python languages, , for computing the max-t and other tests discussed in this paper.
+
 \section*{Acknowledgments}

### 2015-07-07 paper.tex
Rob edited paper

@@
 with $Y\sim N(\mu, I \cdot \sigma^2)$.
-The current model fit is $\proj_{E_{k-1}} Y$ and is is easy to see that (\ref{eqn:maxtsamp})  is equivalent to sampling
+The current model fit is $\proj_{E_{k-1}} Y$ and it is easy to see that (\ref{eqn:maxtsamp})  is equivalent to sampling
 \begin{eqnarray}
-Y^*=\proj_{E_{k-1}}) Y + (I- \cdot \proj_{E_{k-1}}) Z; \;Z\sim N(0,I\sigma^2)
+Y^*=\proj_{E_{k-1}} Y + (I- \proj_{E_{k-1}}) Z; \;Z\sim N(0,I\cdot\sigma^2)
 \label{eq:boot}
@@
 
-Note that  (\ref{eq:boot}  has the form of  (conditional) parametric bootstrap sampling, in which the bootstrap residuals projected into the subspace
+Note that  (\ref{eq:boot})  has the form of  (conditional) parametric bootstrap sampling, in which the bootstrap residuals projected into the subspace
 orthogonal to the current fit. In similar fashion, one might use a instead a nonparametric version of (\re{eq:boot}), sampling

### 2015-07-08 paper.tex
answered Rob's questions and replaced '[k]' notation with '1:k' (or '0:k' as appropriate)

@@
 \WFcomment{should probably end with $M_{d-1}(Y)$. need to fix this throughout paper.}
-\RTcomment{Will- be more specific here---   A parametric model  with specified nonzero support?}
-
-By ``model,'' we mean a family of probability distributions for $Y$. Without loss of generality, we can take $M_\infty$ as the union of all models under consideration.
+\RTcomment{Will- be more specific here---   A parametric model  with
+  specified nonzero support?}
+\WFcomment{Not necessarily. As far as I can tell, there's no need to restrict ourselves in either case. $Y\sim U[0,\theta]$ is a perfectly good model; so is a Cox proportional hazards model.}
+
+By ``model,'' we mean a family of probability distributions for $Y$.
+Without loss of generality, we can take $M_\infty$ as the union of all models under consideration.
+We will use the notation $M_{0:k}$ to denote the sequence of the first $k$ models.
 
@@
 against the alternative that $F\in M_\infty\setminus M_{k-1}$. The $p$-values must account for the fact that the models are chosen adaptively. 
-\citet{gsell2013sequential} and \citet{li2015accumulation} propose stopping rules that operate on the sequence $p_{[d]}$.
+\citet{gsell2013sequential} and \citet{li2015accumulation} propose stopping rules that operate on the sequence $p_{1:d}$.
 These stopping rules provably control certain error rates, provided that the null $p$-values are mutually independent and independent of the non-null $p$-values.
@@
 In this problem, we observe a random response $Y\in \R^n$ as well as a fixed design matrix $X\in \R^{n \times p}$, whose columns correspond to candidate predictors.
-For each {\em active set} $E \sub [p]$, there is a corresponding candidate model
+For each {\em active set} $E \sub \{1,\ldots,p\}$, there is a corresponding candidate model
 \[
@@
   Let $\hk$ be a stopping rule that operates on sequences
-  $p_{[d]}(Y)$ of $p$-values for ordered hypotheses $H_{[d]}(Y)$. 
+  $p_{1:d}(Y)$ of $p$-values for ordered hypotheses $H_{1:d}(Y)$. 
   For some function $g$, let 
@@
   Suppose that the rule $\hk$ controls $\Err$ at $\alpha$ under
-  independence when $H_{[d]}$ are fixed. That is, suppose that
-  whenever $H_{[d]}$ are fixed and $p_{[d]}$
+  independence when $H_{1:d}$ are fixed. That is, suppose that
+  whenever $H_{1:d}$ are fixed and $p_{1:d}$
   satisfy~\eqref{eq:indepCond_fixed_k0}, we have
@@
   Then, $\hk$ controls $\Err$ at $\alpha$
-  whenever $p_{[d]}$ satisfy~\eqref{eq:indepCond_random_k0}.
+  whenever $p_{1:d}$ satisfy~\eqref{eq:indepCond_random_k0}.
 \end{proposition}
 \begin{proof}
-  For any fixed $k_0$ and $q_1,\ldots, q_{k_0}$, we can construct a
-  problem with $\tilde p_k \sim \delta_{q_k}$ for $k\leq k_0$ and 
-  $\tilde p_k \sim \text{Unif}[0,1]$ for $k > k_0$. 
-  \RTcomment{Will; what are the qk?  you mean we just choose and fix some values qk? Say so.}Then, for the
-  fixed hypotheses 
+  Fix any values $k_0\in \{0,\ldots,d, \infty\}$ and $q_{1:k_0} \in [0,1]^{k_0}$.
+We can construct a problem with $\tilde p_k \sim \delta_{q_k}$ for $k\leq k_0$ and  $\tilde p_k \sim \text{Unif}[0,1]$ for $k > k_0$.
+  \RTcomment{Will; what are the qk?  you mean we just choose and fix some values qk? Say so.} \WFcomment{Is it clearer now?}
+
+  Then, for the fixed hypotheses 
   $\widetilde H_k:\; \tilde p_k \sim \text{Unif}[0,1]$, we have
   \[
-  f(k_0, q_{[k_0]}) 
+  f(k_0, q_{1:k_0}) 
   = \E\left[ g\left(\hk(\tilde p_1, \ldots, \tilde p_d),
@@
   \]
-  which is less than or equal to $\alpha$ by assumption.
+  which is less than or equal to $\alpha$, since $\hk$ is assumed to control $\Err$ in the fixed-hypothesis case.
 
@@
   \begin{align}
-    \Err(\hk,g) &\;=\; \E\left[ \;\E\left[ g\left(\hk(p_1,\ldots, p_d), k_0(Y)\right) \; \mid \; k_0(Y), \; p_{[k_0]}(Y)\right] \; \right]\\
-    &\;\leq\; \E\left[f(k_0, p_{[k_0]})\right] \;\leq\; \alpha.
+    \Err(\hk,g) &\;=\; \E\left[ \;\E\left[ g\left(\hk(p_1,\ldots, p_d), k_0(Y)\right) \; \mid \; k_0(Y), \; p_{1:k_0}(Y)\right] \; \right]\\
+    &\;\leq\; \E\left[f(k_0, p_{1:k_0})\right] \;\leq\; \alpha.
   \end{align}
@@
 
-Sections~\ref{sec:pValsIndep}--\ref{sec:SSP} discuss conditions on the model sequence $M_{[d]}$ and the $p$-value sequence $p_{[d]}$ under which~\eqref{eq:stochLargeCond_random_k0} is satisfied.
+Sections~\ref{sec:pValsIndep}--\ref{sec:SSP} discuss conditions on the model sequence $M_{0:d}$ and the $p$-value sequence $p_{1:d}$ under which~\eqref{eq:stochLargeCond_random_k0} is satisfied.
 
@@
 \]
-for which $M_{[k]}$ is measurable with respect to $\sF_k$. We say that $\sF_{[d]}$ {\em separates} the $p$-values $p_{[d]}$ if 
+for which $M_{0:k}$ is measurable with respect to $\sF_k$. We say that $\sF_{0:d}$ {\em separates} the $p$-values $p_{1:d}$ if 
 \begin{enumerate}
@@
 
-  Let $p_{[d]}$ be selective $p$-values for $M_{[d]}$, 
-  separated by $\sF_{[d]}$.
+  Let $p_{1:d}$ be selective $p$-values for $M_{0:d}$, 
+  separated by $\sF_{0:d}$.
 
@@
 
-We will concern ourselves mainly with the {\em sufficient filtration} for the path $M_{[k]}$, defined as
-\[
-\sF_k^{\text{suff}} = \sF(M_{[k]}, T_k),
-\]
-\RTcomment{Will- define $M_{[k-1]}$..  it is  the collection $\{ M_1 M_2 \ldots M_{k-1}\}$?}
+We will concern ourselves mainly with the {\em sufficient filtration} for the path $M_{0:k}$, defined as
+\[
+\sF_k^{\text{suff}} = \sF(M_{0:k}, T_k),
+\]
+\RTcomment{Will- define $M_{0:k-1}$..  it is  the collection $\{ M_1 M_2 \ldots M_{k-1}\}$?} \WFcomment{I've now done so in Section~\ref{sec:notation}}.
 We will suppress the superscript and use $\sF_k$ to mean $\sF_k^{\text{suff}}$ for the remainder of the article.
 
-The sufficient filtration separates $p_{[d]}$ if and only if 
+The sufficient filtration separates $p_{1:d}$ if and only if 
 \begin{enumerate}
-\item $p_k$ is conservative given $M_{[k-1]}$ and $T_{k-1}$, and
-\item $p_k$ is a function of $M_{[k]}$ and $T_k$.
+\item $p_k$ is conservative given $M_{0:k-1}$ and $T_{k-1}$, and
+\item $p_k$ is a function of $M_{0:k}$ and $T_k$.
 \end{enumerate}
@@
 To be a valid selective $p$-value for a single step, $p_k$ only needs to be conservative conditional on $M_{k-1}$.
-In general, it is more stringent to additionally require that $p_k$ must also condition on $T_{k-1}$ as well as the entire subpath $M_{[k-1]}$.
+In general, it is more stringent to additionally require that $p_k$ must also condition on $T_{k-1}$ as well as the entire subpath $M_{0:k-1}$.
 In most cases of interest, however, the ``extra'' conditioning is redundant, for two reasons.
@@
 \begin{equation}\label{eq:SSP}
-\sF(M_k, \, T_k) = \sF(M_{[k]}, \, T_k).
+\sF(M_k, \, T_k) = \sF(M_{0:k}, \, T_k).
 \end{equation}
@@
 
-We say that a path algorithm $M_{[d]}(\cdot)$ satisfies the {\em subpath sufficiency principle} (SSP) if~\eqref{eq:SSP} holds for every $k$.
-Equation~\eqref{eq:SSP} means that if we know $M_k(Y) = M$, then we can compute the entire subpath of models $M_{[k]}$ knowing only $T(Y, \, M)$, the sufficient statistic of model $M$.
+We say that a path algorithm $M_{0:d}(\cdot)$ satisfies the {\em subpath sufficiency principle} (SSP) if~\eqref{eq:SSP} holds for every $k$.
+Equation~\eqref{eq:SSP} means that if we know $M_k(Y) = M$, then we can compute the entire subpath of models $M_{0:k}$ knowing only $T(Y, \, M)$, the sufficient statistic of model $M$.
 For example, given that four steps of forward stepwise linear regression yield the active set $E_4 = \{1,5,7,11\}$, the order in which these variables were included can be determined from $T_4 = X_{E_4}'Y$.
@@
 
-Equation~\eqref{eq:onlyT} shows that $j_1,\ldots, j_{k-1}$ all depend on $Y$ only through $T_k(Y)$. As a result, it also follows that the enitre sequence $M_{[k]}(Y)$ depends only on $T_k(Y)$.
+Equation~\eqref{eq:onlyT} shows that $j_1,\ldots, j_{k-1}$ all depend on $Y$ only through $T_k(Y)$. As a result, it also follows that the enitre sequence $M_{0:k}(Y)$ depends only on $T_k(Y)$.
 \end{proof}
@@
 \begin{theorem}[Sufficient Condition for Independence on Nulls] \label{thm:suffCond}
-Assume that $T(Y; M)$ is a complete sufficient statistic for each candidate model $M$, that $p_k$ are exact selective $p$-values for $M_{k-1}$, and that $M_{[d]}(\cdot)$ satisfies the SSP.
-
-If $p_k$ is $\sF_k$-measurable for each $k$, then the $p$-value sequence $p_{[d]}(Y)$ is independent on nulls.
+Assume that $T(Y; M)$ is a complete sufficient statistic for each candidate model $M$, that $p_k$ are exact selective $p$-values for $M_{k-1}$, and that $M_{0:d}(\cdot)$ satisfies the SSP.
+
+If $p_k$ is $\sF_k$-measurable for each $k$, then the $p$-value sequence $p_{1:d}(Y)$ is independent on nulls.
 \end{theorem}
@@
 Therefore, we must have $h(T_{k-1})\eqAS 0$, so $p_k$ is independent of $\sF(M_k, T_{k-1})$.
-Because $M_{[d]}$ satisfies the SSP, $p_k$ is also independent of $\sF_{k-1} = \sF(M_{[k-1]}, T_{k-1})$.
-
-If $p_k$ is also $\sF_k$-measurable, then the sufficient filtration separates $p_{[d]}$, implying that $p_{[d]}$ is independent on nulls.
+Because $M_{0:d}$ satisfies the SSP, $p_k$ is also independent of $\sF_{k-1} = \sF(M_{0:k-1}, T_{k-1})$.
+
+If $p_k$ is also $\sF_k$-measurable, then the sufficient filtration separates $p_{1:d}$, implying that $p_{1:d}$ is independent on nulls.
 \end{proof}
@@
 
-If we follow this strategy, then the reference distribution for $U(Y)$ is measurable with respect to $T_{k-1}$ and $M_{k-1}$; moreover, if $U(Y)$ is also measurable with respect to $(M_k, T_k)$, the next model in the path and its sufficient statistics, then $p_k$ is measurable with respect to $\sF_k$, and so the $p$-value sequence $p_{[d]}$ will be independent on nulls per Theorem~\ref{thm:suffCond}.
+If we follow this strategy, then the reference distribution for $U(Y)$ is measurable with respect to $T_{k-1}$ and $M_{k-1}$; moreover, if $U(Y)$ is also measurable with respect to $(M_k, T_k)$, the next model in the path and its sufficient statistics, then $p_k$ is measurable with respect to $\sF_k$, and so the $p$-value sequence $p_{1:d}$ will be independent on nulls per Theorem~\ref{thm:suffCond}.
 
@@
 \begin{theorem}
-  Assume the model path $M_{[d]}$ is obtained by forward stepwise 
+  Assume the model path $M_{0:d}$ is obtained by forward stepwise 
   linear regression. Then, for a candidate active set $E$ of size $k$, 
@@
   with $\Theta \sub \R^p$ convex, and assume
-  that the model path $M_{[d]}$ is given 
+  that the model path $M_{0:d}$ is given 
   by the ever-active set for the $\ell_1$-penalized problem
@@
 
-\WFcomment{Should we compute the FWER and FDR only conditional on screening? Strong / Forward / Basic should all be correct given screening since they condition on $p_{[k_0-1]}$.}
+\WFcomment{Should we compute the FWER and FDR only conditional on screening? Strong / Forward / Basic should all be correct given screening since they condition on $p_{1:k_0-1}$.}
 
@@
   \sF(M_0,T_0) &\underlabel_{\text{selection } 1} 
-  \sF(M_{[1]},T_0) \underlabel_{\text{inference } 1}
-  \sF(M_{[1]},T_1) \quad \sub \;\;\cdots\\[8pt]
+  \sF(M_{0:1},T_0) \underlabel_{\text{inference } 1}
+  \sF(M_{0:1},T_1) \quad \sub \;\;\cdots\\[8pt]
   \label{eq:infoPartition}
   \cdots\;\; \sub \quad&
-  \sF(M_{[d-1]},T_{d-1}) \underlabel_{\text{selection } d}
-  \sF(M_{[d]},T_{d-1}) 
+  \sF(M_{0:d-1},T_{d-1}) \underlabel_{\text{selection } d}
+  \sF(M_{0:d},T_{d-1}) 
   \underlabel_{\text{inference } d}
-  \sF(M_{[d]}, T_d)
+  \sF(M_{0:d}, T_d)
 \end{align}
@@
 \begin{comment}
-\WFcomment{There is a filtration interpretation when you have the appropriate sufficiency properties.} Let $\sF_{k,i}$ denote the $\sigma$-algebra generated by $M_{[k]}$ and $p_{[i]}$.
+\WFcomment{There is a filtration interpretation when you have the appropriate sufficiency properties.} Let $\sF_{k,i}$ denote the $\sigma$-algebra generated by $M_{0:k}$ and $p_{1:i}$.
 \begin{align*}
-  \sF_{k,i} &= \sF(M_{[k]},p_{[i]})\\
+  \sF_{k,i} &= \sF(M_{0:k},p_{1:i})\\
   \sF_0 &\underlabel_{\text{selection } 1} \sF_{1,0} \underlabel_{\text{inference } 1}

### 2015-07-08 paper.tex
small changes

--- 2015-07-07 paper.tex
@@
 \end{equation}
-The next-$\lambda$ statistic is $\sF_k$-measurable if we use the LASSO to select variables, but not if we use forward-stepwise.
-\WFcomment{Relate this to spacings test.
-Is this essentially the ``generalized spacings test?''}
+Here $\beta(E_{k-1})$ is the support of the ever-active set $E_{k-1}$, so that $U_k$ is the next $\lambda$ at which a new predictor is added by the lasso.
+\RTcomment{Will: we never define $\beta(E)$: have i said it right?}
+The next-$\lambda$ statistic is $\sF_k$-measurable if we use the LASSO to select variables. \WFcomment{Relate this to spacings test. Is this essentially the ``generalized spacings test?''}
 

### 2015-07-09 paper.tex
rob

@@
   \hline\\
-1 & 3 & 0.000 & 0.000 & 0.000 \\ 
-  2 & 9& 0.000 & 0.000 & 0.000 \\ 
-  3 & 4 & 0.000 & \color{red}{0.054} & 0.000 \\ 
-  4 & 20& 0.000 & { 0.333} & 0.003 \\ 
-  5 & 37& 0.002 & 0.756 & 0.043 \\ 
-  6 & 7 & 0.003 & 0.254 & 0.037 \\ 
-  7 & 2& 0.000 & 0.003 & 0.000 \\ 
-  8 & 19 & 0.020 & 0.028 &  \color{red}{0.343} \\ 
-  9 & 11 & 0.112 & 0.548 & 0.937 \\ 
-  10 & 52& 0.214 & 0.911 & 0.907 \\ 
-  11 & 57& 0.049 & 0.000 & 0.954 \\ 
-  12 & 24& 0.222 & 0.000 & 0.996 \\ 
-  13 & 5& 0.183 & 0.054 & 0.789 \\ 
-  14 & 49 & 0.158 & 0.333 & 1.000 \\ 
-  15 & 30 & 0.212 & 0.756 & 1.000 \\ 
-  16 & 46 & \color{red}{0.159} & 0.254 & 0.000 \\ 
-  17 & 59 & 0.270 & 0.003 & 0.004 \\ 
-  18 & 33& 0.116 & 0.028 & 1.000 \\ 
-  19 & 62 & 0.081 & 0.548 & 0.668 \\ 
-  20 & 27 & 0.170 & 0.911 & 1.000 \\
-
+1 & 3& 0.00 & 0.00 & 0.00 \\ 
+  2 & 9 & 0.00 & 0.00 & 0.00 \\ 
+  3 & 4 & 0.00 & \color{red}{0.05} & 0.00 \\ 
+  4 & 20 & 0.00 & 0.33 & 0.01 \\ 
+  5 & 37 & 0.00 & 0.76 & 0.05 \\ 
+  6 & 7 & 0.00 & 0.25 & 0.05 \\ 
+  7 & 2 & 0.00 & 0.00 & 0.00 \\ 
+  8 & 19 & 0.02 & 0.03 & \color{red}{0.37} \\ 
+  9 & 11 & 0.11 & 0.55 & 0.95 \\ 
+  10 & 52 & 0.21 & 0.91 & 0.92 \\ 
+  11 & 57 & 0.05 & 0.37 & 0.21 \\ 
+  12 & 24 & 0.22 & 0.15 & 0.00 \\ 
+  13 & 5 & 0.18 & 0.07 & 0.10 \\ 
+  14 & 49 & 0.16 & 0.97 & 0.75 \\ 
+  15 & 30& 0.21 & 0.15 & 0.05 \\ 
+  16 & 46 & \color{red}{0.16} & 0.05 & 0.53 \\ 
+  17 & 59 & 0.27 & 0.45 & 0.68 \\ 
+  18 & 33 & 0.12 & 0.71 & 0.80 \\ 
+  19 & 62 & 0.08 & 0.40 & 0.48 \\ 
+  20 & 27 & 0.17 & 0.60 & 0.50 \\ 
    \hline\\
@@
 We also relate the method to parametric bootstrap sampling.
-We provide a simulation study in Section \ref{sec:simulation} and briefly discuss applications of our framework to
-principal components analysis in section\ref{sec:pac}. The paper ends with a Discussion.
+We provide a simulation study in Section \ref{sec:sparseReg} and briefly discuss applications of our framework to
+principal components analysis in section\ref{sec:pca}. The paper ends with a Discussion.
 
@@
 If we reject the hypothesis, we add to $E$ the predictor that maximized $U_E(Y)$ and continue.
-This is the procedure that used to obtain the p-values in Table \ref{tab:diab}  \RTcomment{ Rob- make sure you fix table 1 to make the correct computation}
+This is the procedure that used to obtain the p-values in Table \ref{tab:diab}, except that we sampled
+conditioning on $||Y||$ as well (since $\sigma^2$ is unknown).
 

### 2015-07-10 paper.tex
rob changes

@@
 The next-$\lambda$ statistic is $\sF_k$-measurable if we use the LASSO to select variables. \WFcomment{Relate this to spacings test. Is this essentially the ``generalized spacings test?''}
+\begin{figure}
+  % source code: bivariateSelVSat.R
+  \includegraphics[width=\textwidth]{figs/larexample.pdf}
+  \label{fig:larexample}
+\end{figure}
 
@@
 
+\begin{comment}
 \subsection{An analytic approximation for the max-t test}
@@
 The approximation seems quite good. By Sidak's lemma, we might conjecture that the approximation is conservative, but we have no proof of this.
+\end{comment}
 
@@
 \end{tabular}
+\caption[tab:stopping]{\em Results of various stopping rules applied to simulated data with 7 strong signals,  as described at the beginning of Section \ref{sec:sparseReg}.
+The {\tt Simple} rule stops at the first time that a p-value exceeds $\alpha$, while {\tt Forward} and {\tt Strong} refers ForwardStop and StrongStrop rules of Section \ref{sec:sequential}.}
+\label{tab:stopping}
 \end{table}

### 2015-07-10 paper.tex
rob

@@
 the ``complete null''. (The issue of what hypothesis to test is discussed in section \ref{sec:whichnull}.)
-Now it turns out that the max-t test if often more powerful that the saturated model test, and importantly, the max-t pvalues are exactly
+Now it turns out that the max-t test is often more powerful that the saturated model test, and importantly, the max-t pvalues are exactly
 independent after we reach the point the model sequence when the complete null is true. This independence means that we can apply
@@
 The next-$\lambda$ statistic is $\sF_k$-measurable if we use the LASSO to select variables. \WFcomment{Relate this to spacings test. Is this essentially the ``generalized spacings test?''}
+
+Table \ref{tab:testsummary} summarizes a number of different tests that have been proposed for least angle regression, lasso and forward stepwise regression.
+Some focus on the complete null hypothesis, others on the incremental null (as defined in Section \ref{sec:whichnull}).
+Some use the saturated model, others the selective model framework, as detailed in earlier in the section.
+The saturated model tests are simpler and do not require sampling, while the selective tests are generally more powerful.
+In the next section we compare the  selective and saturated model settings in more detail.
+
+\begin{table}
+\begin{center}
+\begin{tabular}{ l | r|r| r}
+Procedure& Complete Null & Incremental Null\\
+\hline
+LAR /lasso& Covariance test/Saturated&Spacing test/Saturated \\
+        & Next-$\lambda$ test/Selective& \\
+        \hline
+Forward Stepwise &maxT, maxT-identify/Selective &Gaussian pivot/Saturated\\
+\hline
+\end{tabular}
+\end{center}
+\caption{\em Summary of different tests that have been proposed for least angle regression, lasso and forward stepwise regression.
+Tests are described by the null hypothesis that they test (complete or incremental)  and the underlying model (saturated or selective).
+The maxT, maxT-identify and next-$\lambda$ tests are proposed in this paper, while the covariance test was proposed in \citet{lockhart2014significance}, and the spacing and Gaussian
+pivot tests are given in \citet{spacings}. The latter paper shows that the covariance and spacing test are asymptotically equivalent for the first null step, when the complete null holds,}
+\label{tab:testsummary}
+\end{table}
+Figure  \ref{fig:larexample} shows an example of the LAR/lasso tests with $N=50, p=10$ and two strong signals.
+We plot the p-values only for those realizations in which the correct model was selected in the first two steps (about 95\% of the time).
+We see that the next-$\lambda$ test is more powerful , especially at step 1. The next-$\lambda$ and spacing test have the correct type I error in steps 3 and 4,
+while the covariance test becomes conservative in Step 4, as we move further into the null part of the sequence.
 \begin{figure}
-  % source code: bivariateSelVSat.R
   \includegraphics[width=\textwidth]{figs/larexample.pdf}
+  \caption[fig:larexample]{\em  Cumulative distribution functions of p-values for LAR/lasso tests from simulations with $N=50,  p=10$ and two strong signals.}
   \label{fig:larexample}
@@
 if $\Gamma y^* \geq u$. Further, since $\Gamma$ has on the  order of $2pk$ rows after $k$ steps,  we can save 
-substantially in computation by checking the row-wise inequalities starting from the last row, where the inequality  it is most likely to fail.
+substantially in computation by checking the row-wise inequalities starting from the last row, where the inequality is most likely to fail.
 
@@
 \caption[tab:stopping]{\em Results of various stopping rules applied to simulated data with 7 strong signals,  as described at the beginning of Section \ref{sec:sparseReg}.
-The {\tt Simple} rule stops at the first time that a p-value exceeds $\alpha$, while {\tt Forward} and {\tt Strong} refers ForwardStop and StrongStrop rules of Section \ref{sec:sequential}.}
+The {\tt Simple} rule stops at the first time that a p-value exceeds $\alpha$, while {\tt Forward} and {\tt Strong} refers to the ForwardStop and StrongStrop rules of Section \ref{sec:sequential}.}
 \label{tab:stopping}
@@
 for the complete null hypothesis $H_0: \lambda_j=0 \; \forall j \neq k$.  The computation of the resulting p-values requires numerical integration.
-The details are given iin \citet{choi2014selecting}.
+The details are given in \citet{choi2014selecting}, with their ``integrated CSV'' test  corresponding to the sequential selective test of this paper.
 
@@
 
-We plan to distribute  open source software, in both the R and Python languages, , for computing the max-t and other tests discussed in this paper.
+We plan to distribute  open source software, in both the R and Python languages,  for computing the max-t and other tests discussed in this paper.
 

### 2015-07-11 paper.tex
rob made small changes

@@
 In the case of linear regression, our $p$-values improve on the power of the spacings test of \citet{taylor2014exact}, often dramatically.
-  We also show that the selected models test  bears a close relationship to parametric bootstrap sampling,
+  We also show that the selected models test  bears a close relationship to parametric bootstrap sampling.
 
@@
 We can construct a problem with $\tilde p_k \sim \delta_{q_k}$ for $k\leq k_0$ and  $\tilde p_k \sim \text{Unif}[0,1]$ for $k > k_0$.
-  \RTcomment{Will; what are the qk?  you mean we just choose and fix some values qk? Say so.} \WFcomment{Is it clearer now?}
 
@@
 \begin{center}
-\begin{tabular}{ l | r|r| r}
+\begin{tabular}{ | l | r|r| r}
+        \hline
 Procedure& Complete Null & Incremental Null\\
 \hline
-LAR /lasso& Covariance test/Saturated&Spacing test/Saturated \\
+LAR/lasso& Covariance test/Saturated&Spacing test/Saturated \\
         & Next-$\lambda$ test/Selective& \\
@@
 We plot the p-values only for those realizations in which the correct model was selected in the first two steps (about 95\% of the time).
-We see that the next-$\lambda$ test is more powerful , especially at step 1. The next-$\lambda$ and spacing test have the correct type I error in steps 3 and 4,
-while the covariance test becomes conservative in Step 4, as we move further into the null part of the sequence.
+We see that the next-$\lambda$ test is the most powerful , especially at step 1. The next-$\lambda$ and spacing test have the correct type I error in steps 3 and 4,
+while the covariance test becomes conservative in step 4, as we move further into the null part of the sequence.
 \begin{figure}
@@
 Note that  (\ref{eq:boot})  has the form of  (conditional) parametric bootstrap sampling, in which the bootstrap residuals projected into the subspace
-orthogonal to the current fit. In similar fashion, one might use a instead a nonparametric version of (\re{eq:boot}), sampling
-with replacement from the model residuals and then projecting them orthogonal to the current model.
-
-Figure \ref{fig:comparison} shows and example with $n=100,p=20$ and one strong signal, computing the max-t pvalue at Step 3.
+orthogonal to the current fit.  The typical justification for the bootstrap is asymptotic, relying on the convergence of the parameter estimates to the true parameters.
+But something magical  occurs in the selective sampling scheme (\ref{eq:boot}): this  sampling fixes the value of $ X_{E_{k-1}}^Ty$ and this conditioning
+removes the dependence on nuisance parameters.
+Hence  (\ref{eq:boot}) yields the exact (non-asymptotic) conditional distributions and p-values.
+
+Figure \ref{fig:comparison} shows an example with $n=100, p=20$ and one strong signal, computing the max-t pvalue at Step 3.
 The ``exact'' answer  was determined from bootstrap sampling with $20,000$ replications.
@@
   \hline
-  Knockoffs & 0.252 &  &  & 0.103 \\ 
+  Knockoff & 0.252 &  &  & 0.103\\ 
+  Knockoff+& ??   &  &    &   0.000? \\
    \hline

### 2015-07-11 code/python/error_rates.py
python script to compute table

@@ -1,0 +1,66 @@
+from itertools import product
+import pandas as pd, numpy as np
+
+sim_results = pd.read_csv('../snr_5_alpha_05.csv')
+
+names = []
+results = []
+error_types = ['pscreen', 'FWER.mod', 'FDR.var', 'FDR.mod']
+
+name_map = dict(zip(['nominal', 'maxT', 'maxT_identify', 'maxT_unknown', 'saturated'],
+                    ['Nominal', 'MaxT', 'MaxT identify', 'MaxT unknown', 'Saturated']))
+for test, rule in product(['nominal', 'maxT', 'maxT_identify', 'maxT_unknown', 'saturated'],
+                          ['simple', 'forward', 'strong']):
+    name = '_'.join([test, rule])
+    screen = np.mean(sim_results['_'.join([name, 'screen'])])
+
+    R = sim_results['_'.join([name, 'R'])]
+    V_var = sim_results['_'.join([name, 'V_var'])]
+    FDR_var = np.mean(V_var * 1. / np.maximum(R, 1))
+
+    V_model = sim_results['_'.join([name, 'V_model'])]
+    S_var = np.mean(R - V_var)
+    FDR_model = np.mean(V_model * 1. / np.maximum(R, 1))
+    FWER_mod = np.mean(V_model > 0)
+
+    results.append((screen, FWER_mod, FDR_var, FDR_model, S_var))
+    names.append('%s %s' % (name_map[test], rule))
+
+# now knockoffs
+
+R = sim_results['knockoff_R']
+V = sim_results['knockoff_V']
+results.append((np.mean(sim_results['knockoff_screen']),
+                np.nan,
+                np.nan,
+                np.mean(V * 1. / np.maximum(R, 1)),
+                np.mean(R - V)))
+names.append('Knockoff')
+
+R = sim_results['knockoff_plus_R']
+V = sim_results['knockoff_plus_V']
+results.append((np.mean(sim_results['knockoff_plus_screen']),
+                np.nan,
+                np.nan,
+                np.mean(V * 1. / np.maximum(R, 1)),
+                np.mean(R - V)))
+names.append('Knockoff+')
+
+results = np.array(results)
+error_df = pd.DataFrame({'screen':results[:,0],
+                         'fwer.mod':results[:,1],
+                         'fdr.var':results[:,2],
+                         'fdr.mod':results[:,3],
+                         's.var':results[:,4]},
+                        index=names)
+
[…]

### 2015-07-11 code/python/compare_methods.py
added unknown pvalue and knockoff+

@@ -84,19 +84,35 @@ def run(y, X, sigma, active,
         # knockoff
 
         rpy.r.assign('alpha', alpha)
+
         knockoff = np.array(rpy.r("""
         library(knockoff)
-        knockoff.filter(X = X, y = y, fdr=alpha)$selected
+        knockoff.filter(X = X, y = y, fdr=alpha, threshold="knockoff")$selected
     """)) - 1
         numpy2ri.deactivate()
         knockoff_R = knockoff.shape[0]
         knockoff_V = knockoff_R - len(set(active).intersection(knockoff))
         knockoff_screen = knockoff_R > completion_idx
+
+        knockoff_plus = np.array(rpy.r("""
+        knockoff.filter(X = X, y = y, fdr=alpha, threshold="knockoff+")$selected
+    """)) - 1
+        numpy2ri.deactivate()
+        knockoff_plus_R = knockoff_plus.shape[0]
+        knockoff_plus_V = knockoff_plus_R - len(set(active).intersection(knockoff_plus))
+        knockoff_plus_screen = knockoff_plus_R > completion_idx
+
         full_results.setdefault('knockoff_R', []).append(knockoff_R)
         full_results.setdefault('knockoff_V', []).append(knockoff_V)
         full_results.setdefault('knockoff_screen', []).append(knockoff_screen)
 
+        full_results.setdefault('knockoff_plus_R', []).append(knockoff_plus_R)
+        full_results.setdefault('knockoff_plus_V', []).append(knockoff_plus_V)
+        full_results.setdefault('knockoff_plus_screen', []).append(knockoff_plus_screen)
+
     for pval, rule_ in product(['maxT_identify_pvalue',
+                                'maxT_identify_unknown_pvalue',
+                                'maxT_unknown_pvalue',
                                 'saturated_pvalue',
                                 'nominal_pvalue',
                                 'maxT_pvalue'],

### 2015-07-11 paper.tex
rob page 21

@@
 
-Note that  (\ref{eq:boot})  has the form of  (conditional) parametric bootstrap sampling, in which the bootstrap residuals projected into the subspace
+Note that  (\ref{eq:boot})  has the form of  (conditional) parametric bootstrap sampling, but  with  a novel twist:  the bootstrap residuals are projected into the subspace
 orthogonal to the current fit.  The typical justification for the bootstrap is asymptotic, relying on the convergence of the parameter estimates to the true parameters.

### 2015-07-13 code/python/compare_methods.py
fixed V_var bug, computing everything in python now

@@ -1,3 +1,5 @@
+from __future__ import division
+
 import os
 import numpy as np, pandas as pd
 from itertools import product
@@ -17,9 +19,13 @@ def summary(variables, pvalues, active, rule, alpha):
     completion_idx = completion_index(variables, active)
     R = rule(pvalues, alpha)
     V_var = R - len(set(active).intersection(variables[:R]))
-    V_model = max(R - completion_idx, 0)
+    V_model = max(R - 1 - completion_idx, 0)
     screen = R > completion_idx
-    return R, V_var, V_model, screen
+    FWER_model = (V_model > 0)
+    FDP_model = V_model / max(R, 1)
+    FDP_var = V_var / max(R, 1)
+    S_var = R - V_var
+    return R, V_var, V_model, screen, FWER_model, FDP_model, FDP_var, S_var
 
 def simulate(n=100, p=40, rho=0.3, snr=5,
              do_knockoff=False,
@@ -42,7 +48,8 @@ def simulate(n=100, p=40, rho=0.3, snr=5,
                do_knockoff=do_knockoff,
                full_results=full_results,
                maxstep=maxstep,
-               compute_maxT_identify=compute_maxT_identify)
+               compute_maxT_identify=compute_maxT_identify,
+               alpha=alpha)
 
 def run(y, X, sigma, active,
         full_results={},
@@ -83,6 +90,7 @@ def run(y, X, sigma, active,
 
         # knockoff
 
+        print X.shape, 'shape'
         rpy.r.assign('alpha', alpha)
 
         knockoff = np.array(rpy.r("""
@@ -123,21 +131,27 @@ def run(y, X, sigma, active,
                                     'strong',
                                     'forward'])):
         rule, rule_name = rule_
-        R, V_var, V_model, screen = summary(results['variable_selected'],
-                                            results[pval],
-                                            active,
-                                            rule, 
-                                            alpha)
+        (R, 
+         V_var, 
+         V_model, 
+         screen,
+         FWER_model,
+         FDP_model,
+         FDP_var,
+         S_var) = summary(np.asarray(results['variable_selected']),
+                          results[pval],
+                          active,
+                          rule, 
+                          alpha)
 
         pval_name = '_'.join(pval.split('_')[:-1])
-        full_results.setdefault('%s_%s_R' % (pval_name, rule_name), []).append(R)
-        full_results.setdefault('%s_%s_V_var' % (pval_name, rule_name), []).append(V_var)
[…]

### 2015-07-13 code/python/error_rates.py

@@ -1,11 +1,14 @@
+import os
 from itertools import product
 import pandas as pd, numpy as np
+
+os.system('R CMD BATCH maxT.R')
 
 sim_results = pd.read_csv('../snr_5_alpha_05.csv')
 
 names = []
 results = []
-error_types = ['pscreen', 'FWER.mod', 'FDR.var', 'FDR.mod']
+error_types = ['pscreen', 'FWER.mod', 'FWER.mod.cond', 'FDR.var', 'FDR.mod']
 
 name_map = dict(zip(['nominal', 'maxT', 'maxT_identify', 'maxT_unknown', 'saturated'],
                     ['Nominal', 'MaxT', 'MaxT identify', 'MaxT unknown', 'Saturated']))
@@ -13,17 +16,13 @@ for test, rule in product(['nominal', 'maxT', 'maxT_identify', 'maxT_unknown', '
                           ['simple', 'forward', 'strong']):
     name = '_'.join([test, rule])
     screen = np.mean(sim_results['_'.join([name, 'screen'])])
+    FDR_var = np.mean(sim_results['_'.join([name, 'FDP_var'])])
+    FDR_model = np.mean(sim_results['_'.join([name, 'FDP_model'])])
+    FWER_model = np.mean(sim_results['_'.join([name, 'FWER_model'])])
+    FWER_model_cond = FWER_model / screen
+    S_var = np.mean(sim_results['_'.join([name, 'S_var'])])
 
-    R = sim_results['_'.join([name, 'R'])]
-    V_var = sim_results['_'.join([name, 'V_var'])]
-    FDR_var = np.mean(V_var * 1. / np.maximum(R, 1))
-
-    V_model = sim_results['_'.join([name, 'V_model'])]
-    S_var = np.mean(R - V_var)
-    FDR_model = np.mean(V_model * 1. / np.maximum(R, 1))
-    FWER_mod = np.mean(V_model > 0)
-
-    results.append((screen, FWER_mod, FDR_var, FDR_model, S_var))
+    results.append((screen, FWER_model, FWER_model_cond, FDR_var, FDR_model, S_var))
     names.append('%s %s' % (name_map[test], rule))
 
 # now knockoffs
@@ -31,6 +30,7 @@ for test, rule in product(['nominal', 'maxT', 'maxT_identify', 'maxT_unknown', '
 R = sim_results['knockoff_R']
 V = sim_results['knockoff_V']
 results.append((np.mean(sim_results['knockoff_screen']),
[…]

### 2015-07-17 code/python/error_rates.py
runs for different levels of alpha

@@ -4,68 +4,79 @@ import pandas as pd, numpy as np
 
 os.system('R CMD BATCH maxT.R')
 
-sim_results = pd.read_csv('../snr_5_alpha_05.csv')
 
-names = []
-results = []
-error_types = ['pscreen', 'FWER.mod', 'FWER.mod.cond', 'FDR.var', 'FDR.mod']
 
-name_map = dict(zip(['nominal', 'maxT', 'maxT_identify', 'maxT_unknown', 'saturated'],
-                    ['Nominal', 'MaxT', 'MaxT identify', 'MaxT unknown', 'Saturated']))
-for test, rule in product(['nominal', 'maxT', 'maxT_identify', 'maxT_unknown', 'saturated'],
-                          ['simple', 'forward', 'strong']):
-    name = '_'.join([test, rule])
-    screen = np.mean(sim_results['_'.join([name, 'screen'])])
-    FDR_var = np.mean(sim_results['_'.join([name, 'FDP_var'])])
-    FDR_model = np.mean(sim_results['_'.join([name, 'FDP_model'])])
-    FWER_model = np.mean(sim_results['_'.join([name, 'FWER_model'])])
-    FWER_model_cond = FWER_model / screen
-    S_var = np.mean(sim_results['_'.join([name, 'S_var'])])
+for alpha in ['05', '10', '20']:
 
-    results.append((screen, FWER_model, FWER_model_cond, FDR_var, FDR_model, S_var))
-    names.append('%s %s' % (name_map[test], rule))
+    sim_results = pd.read_csv('../snr_5_alpha_%s.csv' % alpha)
 
-# now knockoffs
+    getvar = lambda var: sim_results['_'.join([name, var])]
 
-R = sim_results['knockoff_R']
-V = sim_results['knockoff_V']
-results.append((np.mean(sim_results['knockoff_screen']),
-                np.nan,
-                np.nan,
-                np.nan,
-                np.mean(V * 1. / np.maximum(R, 1)),
-                np.mean(R - V)))
-names.append('Knockoff')
+    names = []
+    results = []
+    #error_types = ['pscreen', 'FWER.mod', 'FWER.mod.cond', 'FDR.var', 'FDR.mod']
 
-R = sim_results['knockoff_plus_R']
-V = sim_results['knockoff_plus_V']
-results.append((np.mean(sim_results['knockoff_plus_screen']),
-                np.nan,
-                np.nan,
-                np.nan,
[…]

### 2015-07-17 code/python/error_rates.py
more options for running script

@@ -2,10 +2,6 @@ import os
 from itertools import product
 import pandas as pd, numpy as np
 
-os.system('R CMD BATCH maxT.R')
-
-
-
 for alpha in ['05', '10', '20']:
 
     sim_results = pd.read_csv('../snr_5_alpha_%s.csv' % alpha)
@@ -14,12 +10,12 @@ for alpha in ['05', '10', '20']:
 
     names = []
     results = []
-    #error_types = ['pscreen', 'FWER.mod', 'FWER.mod.cond', 'FDR.var', 'FDR.mod']
+    guarantees = []
 
     name_map = dict(zip(['nominal', 'maxT', 'maxT_identify', 'maxT_unknown', 'saturated'],
                         ['Nominal', 'MaxT', 'MaxT identify', 'MaxT unknown', 'Saturated']))
     for test, rule in product(['nominal', 'maxT', 'maxT_identify', 'maxT_unknown', 'saturated'],
-                              ['simple', 'forward', 'strong']):
+                              ['simple', 'forward']):
         name = '_'.join([test, rule])
         screen = np.mean(getvar("screen"))
         FDR_var = np.mean(getvar("FDP_var"))
@@ -29,7 +25,14 @@ for alpha in ['05', '10', '20']:
         FWER_model_cond = FWER_model / screen
         S_var = np.mean(getvar("S_var"))
 
-        results.append((screen, FWER_model, FWER_model_cond, FDR_var, FDR_model, FDR_model_cond, S_var))
+        if test in ['maxT', 'maxT_identify', 'maxT_unknown']:
+            if rule == 'simple':
+                guarantees.append((False, True, True, True, False, False))
+            else:
+                guarantees.append((False, False, False, True, False, False))
+        else:
+            guarantees.append((False, False, False, False, False, False))
+        results.append((screen, FWER_model, FWER_model_cond, FDR_model, FDR_var, S_var))
         names.append('%s %s' % (name_map[test], rule))
 
     # now knockoffs
@@ -40,10 +43,10 @@ for alpha in ['05', '10', '20']:
                     np.nan,
                     np.nan,
                     np.nan,
-                    np.nan,
                     np.mean(V * 1. / np.maximum(R, 1)),
                     np.mean(R - V)))
     names.append('Knockoff')
+    guarantees.append((False, False, False, False, False, False))
 
     R = sim_results['knockoff_plus_R']
     V = sim_results['knockoff_plus_V']
[…]

### 2015-07-17 paper.tex
added guarantees to tables, diabetes example

@@
 \centering
-\begin{tabular}{rrrrr}
-  \hline
-  Method, Stopping rule& $p_{\text screen}$ & FWER$_{\text mod}$ & FDR$_{\text mod}$ &FDR$_{\text var}$ \\
-% & \$p\_\{$\backslash$text\{screen\}\} & \$$\backslash$text\{FWER\}\_\{$\backslash$text\{model\}\}\$ & \$$\backslash$text\{FDR\}\_\{$\backslash$text\{model\}\}\$ & \$$\backslash$text\{FDR\}\_\{$\backslash$text\{variable\}\}\$ \\
-  \hline
-Nominal, Simple & 0.832 & 0.252 & 0.097 & 0.171 \\ 
-  Nominal, Forward & 0.931 & 0.896 & 0.345 & 0.426 \\ 
-  Nominal, Strong & 0.003 & 0.000 & 0.000 & 0.035 \\ 
-  \hline
-  MaxT, Simple & 0.369 & 0.000 & 0.002 & 0.042 \\ 
-  MaxT, Forward & 0.668 & 0.038 & 0.026 & 0.080 \\ 
-  MaxT, Strong & 0.025 & 0.006 & 0.004 & 0.038 \\ 
-  \hline
-  MaxT-Identify, Simple & 0.404 & 0.000 & 0.002 & 0.042 \\ 
-  MaxT-Identify, Forward & 0.688 & 0.039 & 0.026 & 0.082 \\ 
-  MaxT-Identify, Strong & 0.024 & 0.003 & 0.002 & 0.037 \\ 
-  \hline
-  Saturated, Simple & 0.000 & 0.000 & 0.000 & 0.029 \\ 
-  Saturated, Forward & 0.013 & 0.001 & 0.000 & 0.030 \\ 
-  Saturated, Strong & 0.000 & 0.000 & 0.000 & 0.033 \\ 
-  \hline
-  Knockoff & 0.252 &  &  & 0.103\\ 
-  Knockoff+& ??   &  &    &   0.000? \\
-   \hline
-\end{tabular}
+\input{tables/error_rates_05.tex}
 \caption[tab:stopping]{\em Results of various stopping rules applied to simulated data with 7 strong signals,  as described at the beginning of Section \ref{sec:sparseReg}.
-The {\tt Simple} rule stops at the first time that a p-value exceeds $\alpha$, while {\tt Forward} and {\tt Strong} refers to the ForwardStop and StrongStrop rules of Section \ref{sec:sequential}.}
-\label{tab:stopping}
+The {\tt Simple} rule stops at the first time that a p-value exceeds $\alpha=0.05$, while {\tt Forward} and {\tt Strong} refers to the ForwardStop and StrongStrop rules of Section \ref{sec:sequential}.
+Results with theoretical guarantees are \guarantee{highlighted}.}
+\label{tab:stopping05}
+\end{table}
+
+\begin{table}[ht]
+\centering
+\input{tables/error_rates_20.tex}
+\caption[tab:stopping]{\em Results of various stopping rules applied to simulated data with 7 strong signals,  as described at the beginning of Section \ref{sec:sparseReg}.
+The {\tt Simple} rule stops at the first time that a p-value exceeds $\alpha=0.2$, while {\tt Forward} and {\tt Strong} refers to the ForwardStop and StrongStrop rules of Section \ref{sec:sequential}.
+Results with theoretical guarantees are \guarantee{highlighted}.}
+\label{tab:stopping20}
 \end{table}

### 2015-07-17 code/python/compare_methods.py

@@ -27,17 +27,22 @@ def summary(variables, pvalues, active, rule, alpha):
     S_var = R - V_var
     return R, V_var, V_model, screen, FWER_model, FDP_model, FDP_var, S_var
 
-def simulate(n=100, p=40, rho=0.3, snr=5,
+def simulate(n=100, p=40, rho=0.3, 
+             snr=5,
              do_knockoff=False,
              full_results={},
              alpha=0.05,
+             s=7,
+             random_signs=False,
              maxstep=np.inf,
              compute_maxT_identify=True):
 
     X, y, _, active, sigma = instance(n=n,
                                       p=p,
                                       rho=rho,
-                                      snr=snr)
+                                      snr=snr,
+                                      s=s,
+                                      random_signs=random_signs)
     full_results.setdefault('n', []).append(n)
     full_results.setdefault('p', []).append(p)
     full_results.setdefault('rho', []).append(rho)
@@ -61,7 +66,7 @@ def run(y, X, sigma, active,
         ndraw=8000):
 
     n, p = X.shape
-    results, FS = compute_pvalues(y, X, sigma, maxstep=maxstep,
+    results, FS = compute_pvalues(y, X, active, sigma, maxstep=maxstep,
                                   compute_maxT_identify=compute_maxT_identify,
                                   burnin=burnin,
                                   ndraw=ndraw)
@@ -90,7 +95,6 @@ def run(y, X, sigma, active,
 
         # knockoff
 
-        print X.shape, 'shape'
         rpy.r.assign('alpha', alpha)
 
         knockoff = np.array(rpy.r("""
@@ -179,9 +183,15 @@ Run a batch of simulations.
     parser.add_argument('--alpha',
                         help='Level for FDR and FWER control.', type=float,
                         default=0.05)
-    parser.add_argument('--p',
+    parser.add_argument('--nsample',
+                        help='Sample size.', type=int,
+                        default=100)
+    parser.add_argument('--nfeature',
                         help='Number of features.', type=int,
                         default=40)
+    parser.add_argument('--sparsity',
+                        help='Sparsity level.', type=int,
+                        default=7)
     parser.add_argument('--outfile',
                         help='Where to store results.')
     parser.add_argument('--maxstep',
@@ -196,7 +206,11 @@ Run a batch of simulations.
     if args.seed is not None:
         np.random.seed(args.seed)
 
[…]

### 2015-07-17 code/python/compute_pvalues.py

@@ -7,11 +7,12 @@ from selection.distributions.discrete_family import discrete_family
 from selection.constraints.affine import gibbs_test
 from scipy.stats import norm as ndist
 
-def compute_pvalues(y, X, sigma=1., maxstep=np.inf,
+def compute_pvalues(y, X, active_set=None, sigma=1., maxstep=np.inf,
                     compute_maxT_identify=True,
                     burnin=2000,
                     ndraw=8000,
-                    accept_reject_params=(100,15,2000)):
+                    accept_reject_params=(100,15,2000),
+                    shortcut=True):
     """
     Parameters
     ----------
@@ -21,6 +22,11 @@ def compute_pvalues(y, X, sigma=1., maxstep=np.inf,
 
     X : np.float((n, p))
         The data, in the model $y = X\beta$
+
+    active_set : [] (optional)
+        The true active set. For steps beyond this
+        the selected model methods can (for the purposes
+        of simulation) just draw np.random.sample()
 
     sigma : np.float
         Standard deviation of the gaussian distribution :
@@ -55,37 +61,55 @@ def compute_pvalues(y, X, sigma=1., maxstep=np.inf,
     iter(FS_identity_U); iter(FS_maxT_U)
     results = []
 
+    completed = False
     for i in range(min([n, p, maxstep])):
 
-        # take a step of FS
+        print i
+        if active_set is not None:
+            screened = set(active_set).issubset(FS_maxT.variables) or ((i > (3 * len(active_set))) and shortcut) # can't be any power way out there... i figure
+        else:
+            screened = False
 
-        pval_maxT = FS_maxT.next(compute_pval=True,
-                                 use_identity=False,
-                                 ndraw=ndraw,
-                                 burnin=burnin)
-        
-        pval_maxT_U = FS_maxT_U.next(compute_pval=True,
+        if not screened: 
+
+            # take a step of FS
+
+            pval_maxT = FS_maxT.next(compute_pval=True,
                                      use_identity=False,
                                      ndraw=ndraw,
-                                     burnin=burnin,
-                                     sigma_known=False)
-        
-        if compute_maxT_identify:
+                                     burnin=burnin)
 
-            pval_maxT_identify = FS_identity.next(compute_pval=True,
-                                                  use_identity=True,
-                                                  ndraw=ndraw,
-                                                  burnin=burnin)
+            pval_maxT_U = FS_maxT_U.next(compute_pval=True,
+                                         use_identity=False,
+                                         ndraw=ndraw,
[…]

### 2015-07-17 code/python/diabetes.py

@@ -1,0 +1,58 @@
+import numpy as np
+import pandas as pd
+from rules import forward_stop
+from selection.algorithms.forward_step import forward_step
+from compute_pvalues import compute_pvalues
+import rpy2.robjects as rpy
+from rpy2.robjects import numpy2ri
+rpy.conversion.py2ri = numpy2ri.numpy2ri
+
+numpy2ri.activate()
+rpy.r('''library(lars)
+library(knockoff)
+data(diabetes)
+X = diabetes$x2
+Y = diabetes$y
+D.lm = lm(Y ~ X)
+sigma = sqrt(sum(resid(D.lm)^2 / D.lm$df.resid))
+knock = knockoff.filter(X = X, y = Y, fdr=0.1, threshold="knockoff+")$selected
+print(knock)
+''')
+X = np.array(rpy.r('X'))
+Y = np.array(rpy.r('Y'))
+sigma = np.array(rpy.r('sigma'))[0]
+
+FS_maxT_U = forward_step(X, Y, covariance=sigma**2 * np.identity(Y.shape[0])) # not really using covariance here...
+iter(FS_maxT_U)
+
+FS_maxT = forward_step(X, Y, covariance=sigma**2 * np.identity(Y.shape[0])) 
+iter(FS_maxT)
+
+results = []
+
+for i in range(64):
+    pval_maxT = FS_maxT.next(compute_pval=True,
+                             use_identity=False,
+                             ndraw=80000,
+                             burnin=20000,
+                             sigma_known=True,
+                             alternative='greater')
+    pval_maxT_U = FS_maxT_U.next(compute_pval=True,
+                                 use_identity=False,
+                                 ndraw=80000,
+                                 burnin=20000,
+                                 sigma_known=False,
+                                 alternative='twosided')
+    results.append((pval_maxT, pval_maxT_U))
+    print results[-1]
+
+results = np.array(results)
+pval_table = pd.DataFrame({'known':results[:,0],
+                           'unknown':results[:,1]})
+
+print pval_table.to_html(float_format = lambda v : '%0.3f' % v)
+
+forward_stop_U = forward_stop(results[:,1], 0.1)
+forward_stop_U = forward_stop(results[:,0], 0.1)
+
+

### 2015-07-17 code/python/error_rates.py

@@ -15,7 +15,7 @@ for alpha in ['05', '10', '20']:
     name_map = dict(zip(['nominal', 'maxT', 'maxT_identify', 'maxT_unknown', 'saturated'],
                         ['Nominal', 'MaxT', 'MaxT identify', 'MaxT unknown', 'Saturated']))
     for test, rule in product(['nominal', 'maxT', 'maxT_identify', 'maxT_unknown', 'saturated'],
-                              ['simple', 'forward']):
+                              ['simple', 'forward', 'strong']):
         name = '_'.join([test, rule])
         screen = np.mean(getvar("screen"))
         FDR_var = np.mean(getvar("FDP_var"))
@@ -28,6 +28,8 @@ for alpha in ['05', '10', '20']:
         if test in ['maxT', 'maxT_identify', 'maxT_unknown']:
             if rule == 'simple':
                 guarantees.append((False, True, True, True, False, False))
+            elif rule == 'strong':
+                guarantees.append((False, True, False, True, False, False))
             else:
                 guarantees.append((False, False, False, True, False, False))
         else:
@@ -85,12 +87,11 @@ for alpha in ['05', '10', '20']:
             yield ' & '.join(val) + r' \\ '
 
     table = r'''
-\renewcommand{\guarantee}[1]{{\color{blue} #1}}
-\begin{tabular}{lrrrrrr}
-\toprule
-{} &  $p_{\text{screen}}$ &  $\text{FWER}_{\text{mod}}$ &  $\text{FWER}_{\text{mod}} \vert \text{screen}$ &  $\text{FDR}_{\text{model}}$ &  $\text{FDR}_{\text{var}}$ &  $\text{S}_{\text{var}}$ \\ \midrule
-%s 
-\bottomrule
+\newcommand{\guarantee}[1]{{\color{blue} #1}}
+\begin{tabular}{|l|rrrrrr|}
+ \hline
+{} &  $p_{\text{screen}}$ &  $\text{FWER}_{\text{mod}}$ &  $\text{FWER}_{\text{mod}} \vert \text{screen}$ &  $\text{FDR}_{\text{model}}$ &  $\text{FDR}_{\text{var}}$ &  $\text{S}_{\text{var}}$ \\ \hline
+%s  \hline
 \end{tabular}''' % '\n'.join(table_generator())
 
     file('../../tables/error_rates_%s.tex' % alpha, 'w').write(table)

### 2015-07-18 code/python/compare_methods.py
added snr as command line option

@@ -183,6 +183,9 @@ Run a batch of simulations.
     parser.add_argument('--alpha',
                         help='Level for FDR and FWER control.', type=float,
                         default=0.05)
+    parser.add_argument('--snr',
+                        help='Signal to noise.', type=float,
+                        default=7)
     parser.add_argument('--nsample',
                         help='Sample size.', type=int,
                         default=100)

### 2015-07-18 code/python/compute_pvalues.py

@@ -64,7 +64,6 @@ def compute_pvalues(y, X, active_set=None, sigma=1., maxstep=np.inf,
     completed = False
     for i in range(min([n, p, maxstep])):
 
-        print i
         if active_set is not None:
             screened = set(active_set).issubset(FS_maxT.variables) or ((i > (3 * len(active_set))) and shortcut) # can't be any power way out there... i figure
         else:

### 2015-07-18 code/python/error_rates.py
error rate script made a little clearner

@@ -2,9 +2,9 @@ import os
 from itertools import product
 import pandas as pd, numpy as np
 
-for alpha in ['05', '10', '20']:
-
-    sim_results = pd.read_csv('../snr_5_alpha_%s.csv' % alpha)
+for results, outbase in zip(['../snr_5_alpha_%s.csv' % alpha for alpha in ['05', '10', '20']] + ['../snr_7_alpha_20_sparsity7_p200.csv'],
+                            ['../../tables/error_rates_%s' % alpha for alpha in ['05', '10', '20']] + ['../../error_rates_p200']):
+    sim_results = pd.read_csv(results)
 
     getvar = lambda var: sim_results['_'.join([name, var])]
 
@@ -39,27 +39,29 @@ for alpha in ['05', '10', '20']:
 
     # now knockoffs
 
-    R = sim_results['knockoff_R']
-    V = sim_results['knockoff_V']
-    results.append((np.mean(sim_results['knockoff_screen']),
-                    np.nan,
-                    np.nan,
-                    np.nan,
-                    np.mean(V * 1. / np.maximum(R, 1)),
-                    np.mean(R - V)))
-    names.append('Knockoff')
-    guarantees.append((False, False, False, False, False, False))
+    if 'knockoff_R' in sim_results:
 
-    R = sim_results['knockoff_plus_R']
-    V = sim_results['knockoff_plus_V']
-    results.append((np.mean(sim_results['knockoff_plus_screen']),
-                    np.nan,
-                    np.nan,
-                    np.nan,
-                    np.mean(V * 1. / np.maximum(R, 1)),
-                    np.mean(R - V)))
-    names.append('Knockoff+')
-    guarantees.append((False, False, False, False, True, False))
+        R = sim_results['knockoff_R']
+        V = sim_results['knockoff_V']
+        results.append((np.mean(sim_results['knockoff_screen']),
+                        np.nan,
+                        np.nan,
+                        np.nan,
+                        np.mean(V * 1. / np.maximum(R, 1)),
+                        np.mean(R - V)))
+        names.append('Knockoff')
+        guarantees.append((False, False, False, False, False, False))
+
+        R = sim_results['knockoff_plus_R']
+        V = sim_results['knockoff_plus_V']
+        results.append((np.mean(sim_results['knockoff_plus_screen']),
+                        np.nan,
+                        np.nan,
+                        np.nan,
[…]

### 2015-07-18 code/python/diabetes.py
tried estimate of variance for comparison

@@ -12,13 +12,17 @@ rpy.r('''library(lars)
 library(knockoff)
 data(diabetes)
 X = diabetes$x2
+#X = scale(X, T, T)
 Y = diabetes$y
+#Y = Y - mean(Y)
 D.lm = lm(Y ~ X)
 sigma = sqrt(sum(resid(D.lm)^2 / D.lm$df.resid))
 knock = knockoff.filter(X = X, y = Y, fdr=0.1, threshold="knockoff+")$selected
 print(knock)
 ''')
 X = np.array(rpy.r('X'))
+n, p = X.shape
+
 Y = np.array(rpy.r('Y'))
 sigma = np.array(rpy.r('sigma'))[0]
 
@@ -28,31 +32,53 @@ iter(FS_maxT_U)
 FS_maxT = forward_step(X, Y, covariance=sigma**2 * np.identity(Y.shape[0])) 
 iter(FS_maxT)
 
+FS_maxT2 = forward_step(X, Y, covariance=sigma**2 * np.identity(Y.shape[0])) 
+iter(FS_maxT2)
+
 results = []
 
 for i in range(64):
     pval_maxT = FS_maxT.next(compute_pval=True,
                              use_identity=False,
-                             ndraw=80000,
-                             burnin=20000,
+                             ndraw=40000,
+                             burnin=10000,
                              sigma_known=True,
-                             alternative='greater')
+                             accept_reject_params=(100, 15, 10000))
     pval_maxT_U = FS_maxT_U.next(compute_pval=True,
                                  use_identity=False,
-                                 ndraw=80000,
-                                 burnin=20000,
-                                 sigma_known=False,
-                                 alternative='twosided')
-    results.append((pval_maxT, pval_maxT_U))
+                                 ndraw=40000,
+                                 burnin=10000,
+                                 sigma_known=False)
+
+    X_cur = X.copy()[:,FS_maxT.variables]
+    X_cur -= X_cur.mean(0)[None,:]
+    P_cur = np.dot(X_cur, np.linalg.pinv(X_cur))
+    sigma_cur = np.linalg.norm(Y - np.dot(P_cur, Y) - Y.mean()) / np.sqrt(n - np.diag(P_cur).sum())
+
+    FS_maxT2.covariance[:] = sigma_cur**2 * np.identity(n)
+    pval_maxT2 = FS_maxT2.next(compute_pval=True,
+                               use_identity=False,
+                               ndraw=40000,
+                               burnin=10000,
[…]

### 2015-07-18 code/python/compare_methods.py
using the unknown sigma nominal pvalue as well

@@ -127,6 +127,7 @@ def run(y, X, sigma, active,
                                 'maxT_unknown_pvalue',
                                 'saturated_pvalue',
                                 'nominal_pvalue',
+                                'nominalT_pvalue',
                                 'maxT_pvalue'],
                                zip([simple_stop, 
                                     strong_stop,

### 2015-07-18 code/python/compute_pvalues.py

@@ -6,6 +6,7 @@ from selection.algorithms.forward_step import forward_step
 from selection.distributions.discrete_family import discrete_family
 from selection.constraints.affine import gibbs_test
 from scipy.stats import norm as ndist
+import statsmodels.api as sm
 
 def compute_pvalues(y, X, active_set=None, sigma=1., maxstep=np.inf,
                     compute_maxT_identify=True,
@@ -109,7 +110,11 @@ def compute_pvalues(y, X, active_set=None, sigma=1., maxstep=np.inf,
                 completed = True
                 completion_idx = i
 
-        var_select, pval_saturated = FS_maxT.model_pivots(i+1, alternative='twosided',
+        alternative = {1:'greater',
+                       -1:'less'}[FS_maxT.signs[-1]]
+
+        var_select, pval_saturated = FS_maxT.model_pivots(i+1, 
+                                                          alternative=alternative,
                                                           which_var=[FS_maxT.variables[-1]],
                                                           saturated=True)[0]
 
@@ -117,17 +122,26 @@ def compute_pvalues(y, X, active_set=None, sigma=1., maxstep=np.inf,
 
         LSfunc = np.linalg.pinv(FS_maxT.X[:,FS_maxT.variables])
         Z = np.dot(LSfunc[-1], FS_maxT.Y) / (np.linalg.norm(LSfunc[-1]) * sigma)
+        # assuming known variance
         pval_nominal = 2 * ndist.sf(np.fabs(Z))
-        results.append((var_select, pval_maxT_identify, pval_saturated, pval_nominal, pval_maxT, pval_maxT_U, pval_maxT_identify_U))
+
+        # using T
+
+        OLS_model = sm.OLS(y, np.hstack([FS_maxT.fixed_regressors,
+                                         X[:,FS_maxT.variables]]))
+        pval_nominalT = OLS_model.fit().pvalues[-1]
+
+        results.append((var_select, pval_maxT_identify, pval_saturated, pval_nominal, pval_nominalT, pval_maxT, pval_maxT_U, pval_maxT_identify_U))
             
     results = np.array(results).T
     return pd.DataFrame({'variable_selected': results[0].astype(np.int),
                          'maxT_identify_pvalue': results[1],
                          'saturated_pvalue': results[2],
                          'nominal_pvalue': results[3],
-                         'maxT_pvalue': results[4],
-                         'maxT_unknown_pvalue': results[5],
[…]

### 2015-07-18 code/python/diabetes.py

@@ -3,6 +3,7 @@ import pandas as pd
 from rules import forward_stop
 from selection.algorithms.forward_step import forward_step
 from compute_pvalues import compute_pvalues
+
 import rpy2.robjects as rpy
 from rpy2.robjects import numpy2ri
 rpy.conversion.py2ri = numpy2ri.numpy2ri
@@ -12,73 +13,73 @@ rpy.r('''library(lars)
 library(knockoff)
 data(diabetes)
 X = diabetes$x2
-#X = scale(X, T, T)
 Y = diabetes$y
-#Y = Y - mean(Y)
 D.lm = lm(Y ~ X)
 sigma = sqrt(sum(resid(D.lm)^2 / D.lm$df.resid))
-knock = knockoff.filter(X = X, y = Y, fdr=0.1, threshold="knockoff+")$selected
+knock_plus = knockoff.filter(X = X, y = Y, fdr=0.1, threshold="knockoff+")$selected
+print('knockoff+')
+print(knock_plus)
+
+knock = knockoff.filter(X = X, y = Y, fdr=0.1, threshold="knockoff")$selected
+print('knockoff')
 print(knock)
+
 ''')
+
 X = np.array(rpy.r('X'))
 n, p = X.shape
 
 Y = np.array(rpy.r('Y'))
+varnames = list(np.array(rpy.r('colnames(X)')))
 sigma = np.array(rpy.r('sigma'))[0]
 
-FS_maxT_U = forward_step(X, Y, covariance=sigma**2 * np.identity(Y.shape[0])) # not really using covariance here...
-iter(FS_maxT_U)
+pvals = compute_pvalues(Y, X, sigma=sigma, 
+                        maxstep=55,
+                        compute_maxT_identify=False,
+                        burnin=10000,
+                        ndraw=40000,
+                        accept_reject_params=(100,15,10000),
+                        shortcut=True)[0]
 
-FS_maxT = forward_step(X, Y, covariance=sigma**2 * np.identity(Y.shape[0])) 
-iter(FS_maxT)
+pvals['Variable'] = [varnames[i] for i in pvals['variable_selected']]
+pvals['Step'] = np.arange(pvals.shape[0]) + 1
+pvals['variable_selected'] += 1
+pvals = pvals.reindex_axis(['Step', 'Variable', 'variable_selected',
+                            'nominalT_pvalue',
+                            'saturated_pvalue',
+                            'maxT_unknown_pvalue'], axis=1)
[…]

### 2015-07-18 code/python/error_rates.py

@@ -3,7 +3,7 @@ from itertools import product
 import pandas as pd, numpy as np
 
 for results, outbase in zip(['../snr_5_alpha_%s.csv' % alpha for alpha in ['05', '10', '20']] + ['../snr_7_alpha_20_sparsity7_p200.csv'],
-                            ['../../tables/error_rates_%s' % alpha for alpha in ['05', '10', '20']] + ['../../error_rates_p200']):
+                            ['../../tables/error_rates_%s' % alpha for alpha in ['05', '10', '20']] + ['../../error_rates_p200'])[:-1]:
     sim_results = pd.read_csv(results)
 
     getvar = lambda var: sim_results['_'.join([name, var])]
@@ -12,9 +12,9 @@ for results, outbase in zip(['../snr_5_alpha_%s.csv' % alpha for alpha in ['05',
     results = []
     guarantees = []
 
-    name_map = dict(zip(['nominal', 'maxT', 'maxT_identify', 'maxT_unknown', 'saturated'],
+    name_map = dict(zip(['nominalT', 'maxT', 'maxT_identify', 'maxT_unknown', 'saturated'],
                         ['Nominal', 'MaxT', 'MaxT identify', 'MaxT unknown', 'Saturated']))
-    for test, rule in product(['nominal', 'maxT', 'maxT_identify', 'maxT_unknown', 'saturated'],
+    for test, rule in product(['nominalT', 'maxT', 'maxT_identify', 'maxT_unknown', 'saturated'],
                               ['simple', 'forward', 'strong']):
         name = '_'.join([test, rule])
         screen = np.mean(getvar("screen"))

### 2015-07-18 code/python/diabetes.py
diabetes example redone -- left rob's table in paper

@@ -47,8 +47,9 @@ pvals['variable_selected'] += 1
 pvals = pvals.reindex_axis(['Step', 'Variable', 'variable_selected',
                             'nominalT_pvalue',
                             'saturated_pvalue',
-                            'maxT_unknown_pvalue'], axis=1)
-pvals.columns = pd.Index(['Step', 'Variable', 'Column number', 'Nominal pvalue', 'Saturated pvalue', 'MaxT pvalue'])
+                            'maxT_unknown_pvalue',
+                            'maxT_pvalue'], axis=1)
+pvals.columns = pd.Index(['Step', 'Variable', 'Column number', 'Nominal pvalue', 'Saturated pvalue', 'MaxT pvalue', 'MaxT pvalue plugin sigma'])
 
 forward_stop_U = forward_stop(pvals['MaxT pvalue'], 0.1)
 forward_stop_N = forward_stop(pvals['Nominal pvalue'], 0.1)

### 2015-07-19 code/robsCodeTab1.R
rob added files

@@ -1,0 +1,54 @@
+
+library(lars)
+library(selectiveInference,lib.loc="/Users/tibs/dropbox/git/R/mylib")
+
+
+library(truncnorm)
+ library(xtable)
+
+source("lassoinf.R")
+source("exact.R")
+source("hitAndRUn.R")
+
+set.seed(635)
+options(error=dump.frames)
+BIG=10e9
+
+
+x=read.table("data64.txt")
+x=as.matrix(x)
+x=scale(x,T,T)
+nams=scan("data64.names",what="")
+y=scan("diab.y")
+y=y-mean(y)
+set.seed(44)
+n=nrow(x)
+p=ncol(x)
+
+maxp=p
+
+# table for paper
+a=lsfit(x,y)
+s=sqrt(sum(a$res^2)/(n-p))
+res2=lassoinf(x,y,type="step",maxp=maxp)  #naive
+
+a=forwardStep(x,y,sigma=s)
+aa=forwardStepInf(a,x,y,nsteps=p)  #saturated
+
+d=lar(x,y,norm=F,maxsteps=25)
+
+dd=larInf(x,y,d,sigma=s)  
+
+#maxt 
+b=forwardStepMaxtInf(a,x,y,nsteps=25,initialStep=1,trace=T,maxboot=50000,fix.normy=T,AbMeth="will")
+
+save(b,file="diab.RData")  #known sigma
+save(b,file="diab.unk.RData")  #unknown sigma
+
+#make table
+load("diab.unk.RData")
+
+b$pv=c(b$pv,rep(NA,p-length(b$pv)))
+    
+tab=cbind(res2$res[,1],round(res2$res[,3],4),round(aa$pv,4),round(b$pv,4))
+xtable(tab)

### 2015-07-19 paper.tex

@@
 \centering
+\input{tables/diabetes_edit.tex}
+\caption[tab:diab]{\em Forward stepwise regression for the diabetes data: naive p-values from forward stepwise regression, 
+p-values from the ``saturated'' model, and the ``max-T'' p-values (computed without knowing $\sigma^2$) that are presented in this paper. The red annotation indicates the stopping step
+prescribed by {\tt forwardStop}. For the max-T p-values, this stopping rule gives exact $\text{FDR}_{\text{model}}$ control at the 10\% level.}
+\label{tab:diab}
+\end{table}
+
+\begin{table}[ht]
+\centering
 \begin{tabular}{rrrrr}
@@
 \end{tabular}
-\caption[tab:diab]{\em Forward stepwise regression for the diabetes data: naive p-values from forward stepwise regression, 
-p-values from the ``saturated'' model, and the ``max-t'' p-values that are presented in this paper. The red annotation indicates the stopping step
-prescribed by {\tt forwardStop}. For the max-t p-values, this gives exact FDR control at the 10\% level.}
-\label{tab:diab}
+\caption{Rob's data}
 \end{table}
+
 The table shows the naive p-values from forward stepwise regression, that do not account for the selection, as well as the

### 2015-07-19 paper.tex
rob added larexample.pdf

@@
 To choose a model we feed the resulting $p$-values as inputs into sequential stopping rules proposed by \citet{gsell2013sequential} and \citet{li2015accumulation}, achieving adaptive control of the familywise error rate or false discovery rate.
-These stopping rules assume that the single-step null $p$-values are independent of each other and of the non-null $p$-values, a condition that is not satisfied by the $p$-values of \citet{taylor2014exact}.
+These stopping rules assume that the single-step null $p$-values are independent of each other and of the non-null $p$-values, a condition that bmiis not satisfied by the $p$-values of \citet{taylor2014exact}.
 We derive intuitive and general conditions for independence and show that our proposed constructions yield independent $p$-values.
@@
 \label{tab:diab}
-\end{table}
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{rrrrr}
-  \hline\\
- Step& Predictor& Naive pvalue& Saturated pvalue& Max-t pvalue\cr
-  \hline\\
-1 & 3& 0.00 & 0.00 & 0.00 \\ 
-  2 & 9 & 0.00 & 0.00 & 0.00 \\ 
-  3 & 4 & 0.00 & \color{red}{0.05} & 0.00 \\ 
-  4 & 20 & 0.00 & 0.33 & 0.01 \\ 
-  5 & 37 & 0.00 & 0.76 & 0.05 \\ 
-  6 & 7 & 0.00 & 0.25 & 0.05 \\ 
-  7 & 2 & 0.00 & 0.00 & 0.00 \\ 
-  8 & 19 & 0.02 & 0.03 & \color{red}{0.37} \\ 
-  9 & 11 & 0.11 & 0.55 & 0.95 \\ 
-  10 & 52 & 0.21 & 0.91 & 0.92 \\ 
-  11 & 57 & 0.05 & 0.37 & 0.21 \\ 
-  12 & 24 & 0.22 & 0.15 & 0.00 \\ 
-  13 & 5 & 0.18 & 0.07 & 0.10 \\ 
-  14 & 49 & 0.16 & 0.97 & 0.75 \\ 
-  15 & 30& 0.21 & 0.15 & 0.05 \\ 
-  16 & 46 & \color{red}{0.16} & 0.05 & 0.53 \\ 
-  17 & 59 & 0.27 & 0.45 & 0.68 \\ 
-  18 & 33 & 0.12 & 0.71 & 0.80 \\ 
-  19 & 62 & 0.08 & 0.40 & 0.48 \\ 
-  20 & 27 & 0.17 & 0.60 & 0.50 \\ 
-   \hline\\
-\end{tabular}
-\caption{Rob's data}
 \end{table}

### 2015-07-20 code/R/exact.R
rob

@@ -1,0 +1,604 @@
+# functions for exact selective max computation
+
+
+
+
+forwardStepMaxtInf=function(fsfit,x,y,nsteps,AbMeth="will",initialStep=1,sigma=1,maxboot=100000,minboot=300,nonpar=F,projres=T,fix.normy=F,scaleres=F,return.ystar=F,ntry=NULL,try.hitAndRun=TRUE,nburn=2000,ntotal=8000,trace=T){
+
+# we do acc/ref as long we can get minboot samples in maxboot tries. o/w we switch to hit and run
+ #   and use  it for the rest of the steps
+#note- i don;t think idea behind  fix.normy works in general (trying to cond on ||y||)
+  
+    n=nrow(x)
+    p=ncol(x)
+
+    sigma.call=sigma
+
+pv=rep(NA,nsteps)
+    pvless=NULL
+    if(!is.null(ntry)){
+         pvless=matrix(NA,nsteps,length(ntry))
+     }
+sel=fsfit$pred
+    
+    if(AbMeth=="big"){
+  fsinf=forwardStepInf(fsfit,x,y,sigma=sigma,nsteps=nsteps)
+    gam=-fsinf$A
+    stepind=fsinf$stepind
+}
+    
+    xr=x;r=y;fit1=rep(0,n)
+ totb=goodb=rep(0,nsteps)
+
+    go.accrej=TRUE
+    for(k in initialStep:nsteps){
+        cat(k,fill=T)
+        if(is.null(sigma.call)) sigma=sqrt(sum(y^2)/n)
+        if(k>1){ sel1=sel[1:(k-1)]
+        xx=x[,sel1]
+                 proj=xx%*%solve(t(xx)%*%xx)%*%t(xx)
+            bhat1=lsfit(x[,sel1,drop=F],y)$coef[-1]
+                 fit1=x[,sel1,drop=F]%*%bhat1  
+     
+             xr=lsfit(xx,x)$res
+            r=lsfit(xx,y)$res
+                  if(is.null(sigma.call)) sigma=sqrt(sum(r^2)/(n-k))
+             }
+ip=t(xr)%*%r/sqrt(diag(t(xr)%*%xr))
+  if(k>1)  ip[sel1]=0
+    tt=max(abs(ip))
+if(k>1){
+         if(AbMeth=="big"){
+       gam0=gam[stepind<=(k-1),]
[…]

### 2015-07-20 code/R/hitAndRun.R

@@ -1,0 +1,225 @@
+
+
+
+hitAndRun=function(x,y,fsfit,A,b,nsteps,sigma,nburn=1000,ntotal=2000,gap.reset=trunc(ntotal/40),ntry=NULL,return.ystar=F,fix.normy=F,trace=F){
+
+    # not yet working for fix.normy=T!!!!
+ 
+if(nsteps==1) stop("not yet implemented for nsteps=1")
+    n=length(y)
+
+pp=nrow(A)
+
+mod=fsfit$pred[1:nsteps]
+ystar=matrix(NA,ntotal,n)
+xx=x[,mod[1:(nsteps-1)]]
+mu=as.vector(y-lsfit(xx,y)$res)
+newb=b-A%*%mu
+ystar[1,]=y-mu
+
+ii=1
+go=T
+
+while(ii<ntotal){
+    z=rnorm(n)   #random direction
+    if(ii%%gap.reset==0) z=sample(c(1,rep(0,n-1)))
+    z=lsfit(xx,z,int=F)$res
+    z=z/sqrt(sum(z*z))
+    junk=tf.jonvs(ystar[ii,],A,newb, z)   #find truncation limits
+    if(junk$vm>junk$vp){
+        temp=junk$vp;junk$vp= junk$vm;junk$vm=temp  # not sure why this is needed(for fix.normy=T only)
+    }
+    
+    if(ii%%100==0 & trace)  cat(c(ii,junk$vm,junk$vp),fill=T)
+    cc=rtruncnorm(1,a=junk$vm,b=junk$vp,mean=0,sd=sigma*sqrt(sum(z*z)))
+    yy=lsfit(z,ystar[ii,],int=F)$res+cc*z
+ 
+    if(fix.normy) {
+         go=F;
+         syy=yy
+         yy=sqrt(sum(y*y))*yy/sqrt(sum(yy*yy))
+         cat(c(max(A%*%(syy+mu)-b),max(A%*%(yy+mu)-b)),fill=T)
+         if (max(A%*%(yy+mu)-b)<0) {go=T}
+         
+     }
+    if(!fix.normy | go)  {
+    ystar[ii+1,]=yy;
+    val=max(A%*%(ystar[ii+1,]+mu)-b)
+    if(val>0) c("Warning max>0",ii+1,val,fill=T)
+    ii=ii+1
+}
+}
+
+ystar=scale(ystar,-mu,F)
+
+ystar=ystar[-(1:nburn),]
+
+ 
[…]

### 2015-07-21 paper.tex
robs small changes

@@
 \begin{eqnarray}
-L(X'_jY | E_{k-1},X'_{E_{k-1}})
+L(X'_jY | E_{k-1},X'_{E_{k-1}}Y)
 \label{eqn:maxtsamp}
@@
 and then conditioning on the event $E_{k-1}(Y^*)= E_k(Y)$ (a polyhedron).
-In other words, we need to condition on data vectors  $Y*$ that yields the same
+In other words, we need to condition on data vectors  $Y^*$ that yields the same
 sequence of variables and coefficient signs up to step $k-1$. This is the challenging aspect of this computation.
@@
 When the acceptance rate gets too low, we switch to a hit-and run Monte Carlo approach. We exploit the fact that each move along a random direction 
-has a truncated Gaussian distribution whose truncation limits are easily computed from the properties of the polyhedral. Unlike accept-reject, sampling
+has a truncated Gaussian distribution whose truncation limits are easily computed from the properties of the polyhedron. Unlike accept-reject, sampling
 which produces independent samples,  the hit and run method produces dependent samples. Hence we must run it for longer with an initial burn in period,

### 2015-07-23 paper.tex
rob added cart example

@@
 
+\section{Decision trees and interaction models}
+In this section we discuss another application of adaptive sequential inference, to binary decision trees.
+Specifically, the classification and regression tree  (CART)  method builds a decision tree in a top down manner, finding the best available split
+at each node.  When considering splitting two daughter nodes, the generic CART algorithm splits the nodes in an arbitrary order, only stopping when a node
+reaches a minimum size.  We instead consider a  special version of CART  algorithm:  the   "best first"  method, which order splits by their achieved reduction in loss.
+This procedure is used, for example in the R package for gradient boosting, called {\tt gbm}.
+The best first tree-builder is thus a sequential selection  procedure,  to which we can apply the results in this paper. Hence we can obtain independent, sequential p-values
+for each successive split in the decision  tree. We now give some details and an example.
+
+Given data $(y_i, x_{i1}, x_{i2},\ldots x_{ip})$ for $i=1,2,\ldots N$,
+we choose a splitting  variable $j_1$ and split point $t_1$, leading to a model of the form
+\begin{eqnarray}
+\hat y_i=\beta_0 +\beta_{11} I(x_{ij_1} \leq t_1) +\beta_{12} I(x_{ij_1} > t_1)
+\end{eqnarray}
+Then at the second stage, we try splitting one of the two daughter nodes, and choose the split and most reduces the residual sum of squares.
+Suppose that a split of the first node is chosen, using predictor $j_2$ at split point $t_2$.  This leads to the interaction model of the form
+\begin{eqnarray}
+\hat y_i=\beta_0 +\beta_{11} I(x_{ij_1} \leq t_1) \cdot \beta_{31} I(x_{ij_2} \leq t_2)+   \beta_{11} I(x_{ij_1} \leq t_1) \cdot  \beta_{31} I(x_{ij_2} > t_2)  +\beta_{12} I(x_{ij_1} > t_1).\cr
+\label{eqn:intmodel}
+\end{eqnarray}
+Note that the search for possible split points can be expressed through an exhaustive set of basis functions $I(x_{ij} \leq  t_i), (x_{ij}  > t_i)$ for each $t_i \in (x_{1j}, x_{2j}, \ldots x_{nj})$.
+The procedure is somewhat similar to foreward stepwise regression, but with differences.
+For example the set of candidate basis functions available at each stage are a function of what has been  previously entered into the model
+Nonetheless,  the selection procedure  for the first set of $k$ splits can be expressed in polyhedral from $Ay \leq b$, and the saturated and selective inference procedures can be applied.
+As before, the saturated p-values have an explicit truncated Gaussian form, while computation of the selective p-values require accept-reject sampling.
+
+Figure \ref{fig:cart} shows an example with independent standard Gaussian features, and $n=30, p=3$.
+We 
+consider just the first two splits, in a null case (top row) and non-null (bottom row).
+In the non-null case,  the data are generated from a  model of the form (\ref{eqn:intmodel}).
+
+As expect, the nominal p-values are far too small and hence don't control type I error.
+Both the saturated and selective p-values control type I error, with the selective p-values are somewhat more powerful in the non-null case.
+Rules like the forwardStop procedure could be applied  to the selective p-values to choose a ``best'' model size.
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.65\textwidth]{figs/cart.pdf}
+  \caption{\em Cumulative distribution functions of P-values for  the first two splits in a CART decision tree. In the top row, the true model is null,
+  while in the bottom row, the data are generated from a  model of the form (\ref{eqn:intmodel})}
+  \label{fig:cart}
+\end{figure}
+
 \section{Discussion}

### 2015-07-23 paper.tex
rob

@@
 \end{eqnarray}
-Note that the search for possible split points can be expressed through an exhaustive set of basis functions $I(x_{ij} \leq  t_i), (x_{ij}  > t_i)$ for each $t_i \in (x_{1j}, x_{2j}, \ldots x_{nj})$.
+Note that the search for possible split points can be expressed through an exhaustive set of basis functions $I(x_{ij} \leq  t_i), I(x_{ij}  > t_i)$ for each $t_i \in (x_{1j}, x_{2j}, \ldots x_{nj})$.
 The procedure is somewhat similar to foreward stepwise regression, but with differences.

### 2015-10-13 paper.tex
major reorg of paper

@@
 \newcommand{\Err}{\mathcal{E}}
+\newcommand{\RSS}{\text{RSS}}
 
@@
 
-\title{Adaptive Sequential Model Selection}
+\title{Selective Sequential Model Selection}
 \author{William Fithian, Jonathan Taylor, Rob Tibshirani, and Ryan Tibshirani}
@@
 \begin{abstract}
-  Many model selection algorithms produce a ``path'' of fits specifying a sequence of increasingly complex models.
+  Many model selection algorithms produce a path of fits specifying a sequence of increasingly complex models.
 Given such a sequence and the data used to produce them, we consider the problem of choosing the least complex model that is not falsified by the data.
 Extending the selected-model tests of \citet{fithian2014optimal}, we construct $p$-values for each step in the path, accounting for the fact that the model path is determined adaptively using the data.
-In the case of linear regression, our $p$-values improve on the power of the spacings test of \citet{taylor2014exact}, often dramatically.
-  We also show that the selected models test  bears a close relationship to parametric bootstrap sampling.
-
-To choose a model we feed the resulting $p$-values as inputs into sequential stopping rules proposed by \citet{gsell2013sequential} and \citet{li2015accumulation}, achieving adaptive control of the familywise error rate or false discovery rate.
-These stopping rules assume that the single-step null $p$-values are independent of each other and of the non-null $p$-values, a condition that bmiis not satisfied by the $p$-values of \citet{taylor2014exact}.
+In the case of linear regression, our $p$-values improve on the power of the spacings test of \citet{taylor2014exact}, sometimes dramatically.
+
+To select a model, we can feed the resulting $p$-values as inputs into sequential stopping rules such as those proposed by \citet{gsell2013sequential} and \citet{li2015accumulation}, achieving control of the familywise error rate or false discovery rate as desired.
+Some stopping rules require the null $p$-values to be independent of each other and of the non-null $p$-values, a condition not satisfied by the saturated-model $p$-values of \citet{taylor2014exact}.
 We derive intuitive and general conditions for independence and show that our proposed constructions yield independent $p$-values.
+We also discuss connections between the selected-model test and parametric bootstrap sampling.
 \end{abstract}
@@
 \label{sec:introduction}
-Many model selection procedures produce  a sequence of increasingly complex models. Often our goal is to choose the simplest model that is not falsified by the available data.
-In other words, we seek  the index of the smallest adequate model ---  the smallest model that cannot be falsified given available evidence.
-Examples of such model selection procedures include  forward stepwise linear regression, least angle regression (LAR) and  the ever-active path in $\ell_!$-regularized models .
-All of these were studied in \citet{taylor2014exact}. Another, different, example is principal components analysis in which we try to determine the correct rank of the
-covariance matrix. This problem was studied in \citet{choi2014selecting}.
-
-A natural way carry out this model selection is as follows: we compute a set  of sequential p-values   at each step of the path of models and then feed them into a stopping rule
-that provides guaranteed control of the false discovery rate (FDR) or a similar quantity.
+Many model selection procedures produce a sequence of increasingly complex models, leaving the data analyst to choose among them. 
+%While methods like cross-validation aim at minimizing out-of-sample prediction error, such methods tend to select a large fraction of noise variables. 
+Given such a path, we consider the problem of choosing the simplest model in the path that is not falsified by the available data.
+Examples of path algorithms include forward stepwise linear regression, least angle regression (LAR) and  the ever-active path in $\ell_1$-regularized models.
+\citet{taylor2014exact} study methods for generating exact $p$-values at each step of these path algorithms, and their methods provide a starting point for our proposals.
+In addition, \citet{choi2014selecting} discuss a selective approach to determining the correct rank of the covariance matrix in principle components analysis.
+
+To select a model, we compute a set of sequential $p$-values at each step of the model path, and then feed them into a stopping rule that is guaranteed to control the FDR, familywise error rate (FWER), or a similar quantity.
 Recently \citet{gsell2013sequential} and \citet{li2015accumulation} proposed sequential stopping rules of this kind.
-Both sets of rules require that once we have reached the correct model in our path, the p-values in subsequent steps are uniform (or sub-uniform)
+Both sets of rules require that once we have reached the first correct model in our path, the $p$-values in subsequent steps are uniform (or conservative),
 and independent.
-In this paper we develop a theoretical framework for the construction of sequential p-values  that are both (sub)-uniform and independent under the correct model,
-and show how to construct them  in forward stepwise regression, LAR and the ever-active path in $\ell_1$-regularized models.
-In the former case, the test is simply a sequential selective version the max-$t$ test of ~\citet{buja2014}, based on the theory developed \citet{fithian2014optimal}.
-
-Table \ref{tab:diab} shows an example: using the diabetes data taken from \cite{LARS}.
-These data consist of  observations on $442$ patients, with the response of interest being a quantitative measure of disease progression one year after baseline.
-There are ten baseline variables---age, sex, body-mass index, average blood pressure,
-and six blood serum measurements---plus quadratic terms, giving a total of 64 features.
+In this paper we develop a theoretical framework for constructing sequential $p$-values that are both uniform and independent under the correct model, and show how to construct them in forward stepwise regression, LAR and the ever-active path in $\ell_1$-regularized models.
+
+While our approach and analysis are quite general, we begin by introducing a specific example: the selective max-$t$ test for forward stepwise regression, a selective sequential version of the max-$t$ test of \citet{buja2014}, generalized using the theory in \citet{fithian2014optimal}.
+
+\subsection{The max-$t$ Test in Forward Stepwise Regression}
+
+Forward stepwise regression is a greedy algorithm for building a sequence of nested linear regression models.
+At each iteration, it augments the current model by including the variable that will minimize the residual sum of squares of the next fitted model.
+Equivalently, it selects the variable with the largest $t$-statistic, controlling for the variables in the current model.
+
+Let $E \sub \{1, \ldots, d\}$ denote the current {\em active set}, the set of predictor variables already selected, and let $\RSS(E)$ denote the residual sum of squares for the corresponding regression model.
+For $j\notin E$ let $t_{j,E}(Y)$ denote the multivariate $t$-statistic of variable $j$, adjusting for the active variables.
+Using the standard result that
+\begin{equation}
+t_{j,E}^2 = (n-|E|-1) \frac{\RSS(E) - \RSS(E \cup \{j\})}{\RSS(E \cup \{j\})},
+\end{equation}
+we see that the next variable selected is 
+\begin{equation}
+j^* = \argmax_{j\notin E} |t_{j,E}| = \argmin_{j\notin E} \RSS(E \cup \{j\}),
+\end{equation}
+with corresponding $t$-statistic $t_E^* = t_{j^*,E}$.
+The selective max-$t$ test rejects for large values of $t_E^*$, compared to an appropriate conditional distribution that accounts for the adaptive selection of the model path.
+
+Table~\ref{tab:diab} illustrates the max-$t$ test and two others applied to the diabetes data from \cite{LARS}, consisting of observations of $n=442$ patients.
+The response of interest is a quantitative measure of disease progression one year after baseline, and there are ten measured predictors --- age, sex, body-mass index, average blood pressure, and six blood serum measurements --- plus quadratic terms, giving a total of $d=64$ features.
+We apply forward stepwise regression to generate the model path (beginning with an intercept term), and then use each of the three methods to generate $p$-values at each step along the path.
+Finally, we use the ForwardStop rule \citep{gsell2013sequential} at FDR level $\alpha=0.1$ to select a model based on the sequence of $p$-values.
+The red entry in each column indicates the last variable selected by ForwardStop.
 
@@
 \input{tables/diabetes_edit.tex}
-\caption[tab:diab]{\em Forward stepwise regression for the diabetes data: naive p-values from forward stepwise regression, 
-p-values from the ``saturated'' model, and the ``max-T'' p-values (computed without knowing $\sigma^2$) that are presented in this paper. The red annotation indicates the stopping step
-prescribed by {\tt forwardStop}. For the max-T p-values, this stopping rule gives exact $\text{FDR}_{\text{model}}$ control at the 10\% level.}
+\caption[tab:diab]{\em Forward stepwise regression for the diabetes data: naive $p$-values, $p$-values from the saturated model, and our max-$t$ $p$-values.
+The red annotation indicates the model selected by ForwardStop with $\alpha=0.1$.
+For the max-$t$ $p$-values, this stopping rule gives exact $\text{FDR}_{\text{model}}$ control at the 10\% level.}
 \label{tab:diab}
@@
 
-The table shows the naive p-values from forward stepwise regression, that do not account for the selection, as well as the
-p-values from the ``saturated'' model, and the ``max-t'' p-values that are presented in this paper.
-The latter two sets of p-values test someone different hypotheses:  the p-values from the saturated model test whether the partial regression of the predictor just added is zero,
-what we call the ``incremental null''
-In contrast, the max-t pvalues test whether the current model is adequate, that is, all predictors not in the model  have coefficients equal to zero,
-the ``complete null''. (The issue of what hypothesis to test is discussed in section \ref{sec:whichnull}.)
-Now it turns out that the max-t test is often more powerful that the saturated model test, and importantly, the max-t pvalues are exactly
-independent after we reach the point the model sequence when the complete null is true. This independence means that we can apply
-sequential FDR exact stopping rules, such as ``forwardStop'' (\citet{gsell2013sequential}). When applied to the max-t pvalues at FDR 0.10, {\tt forwardStop}
-suggested stopping at step 8. In constant, the saturated lvalues gave stopping a point of 3, with no guarantee of FDR control
-due to the lack of independence of the p-values.
-
-An outline of this paper is as follows:
-In Section \ref{sec:genericSetting} we setup the problem and give some important definitions. In Section \ref{sec:sequential} we review recently proposed methods for
-testing ordered hypothesis using from a sequence of p-values. These method require that the p-values be uniform and independent under the null, and we derive general conditions
-to ensure this property. In Section \ref{sec:singleStep}  we  show how to construct valid selective $p$-values for a single step in the path and derive the max-t test
-for  forward stepwise regression and the Next-$\lambda$ test for the lasso.
-Section \ref{sec:computation}  describes our computational strategy for estimating the p-values using accept-reject and hit and run sampling.
-We also relate the method to parametric bootstrap sampling.
-We provide a simulation study in Section \ref{sec:sparseReg} and briefly discuss applications of our framework to
-principal components analysis in section\ref{sec:pca}. The paper ends with a Discussion.
-
-%\begin{example}[Forward-Stepwise Linear Regression]
- % \citet{taylor2014exact}
-%end{example}
-
-%\begin{example}[The LARS Algorithm in Regression]
-%  \citet{taylor2014exact}
-%\end{example}
-
-%\begin{example}[Ever-Active Path in $\ell_1$-Regularized Methods]
- % \citet{taylor2014exact}
-%\end{example}
-
-%\begin{example}[Principal Components Analysis]
-%  As a second motivating example, consider model selection for   principal components analysis.
-In that case we are given a data matrix $X \in \R^{n\times d}$, with which we form a sample %covariance matrix
-%\[
-%S = \frac{1}{n-1} \sum_{i=1}^n(x_i - \bar x)^2
-%\]
-%The first $d$ principal component loadings are the first $d$ eigenvectors of $S$, which call $u_1,\ldots, u_d$. These induce a sequence of nested Wishart models:
-%\[
-%M_0 \sub M_1 \sub \cdots \sub M_d
-%\]
-%n which
-%\begin{equation}
- % M_k:\; (n-1) S \sim W_d\left(\lambda_0 I_d + \sum_{i=1}^k     \lambda_i u_i u_i', \;\;\; n-1\right).
-%\end{equation}
-%This problem was studied in \citet{choi2014selecting}.
-%\end{example}
-
-\section{Notation and Problem Setting}
+The nominal $p$-values in Table~\ref{tab:diab} are computed by comparing $t_E^*$ to the $t_{n-|E|-1}$-distribution, which would be the correct distribution if the sequence of models were selected ahead of time, before observing the data.
+Because the model sequence is in fact selected adaptively to maximize the value of $t_E^*$, this method is highly anti-conservative.
+
+The max-$t$ test $p$-values are computed using the same test statistic $|t_E^*|$, but compared with a more appropriate null distribution.
+We can simulate the distribution of $|t_E^*|$ under the null model, i.e.
+the linear regression model specified by active set $E$:
+\begin{equation}
+  M(E):\; Y \sim N(X_E\beta, \sigma^2 I), \quad \beta\in \R^{|E|}, \sigma^2>0,
+\end{equation}
+where $X_E$ is the matrix with columns $(X_j)_{j\in E}$.
+Because $U_E=\left(X_E'Y,\, \|Y\|^2\right)$ is the complete sufficient statistic for $M(E)$, we can sample from the conditional null distribution of $t_E^*$ given $U_E$ and the current active set $E$, using the {\em selected-model test} framework of \citet{fithian2014optimal}.
+In step one, because $E=\{1_n\}$ is fixed and $t_E^*$ is independent of $U_E$, the conditional test reduces to the max-$t$ test proposed in \citet{buja2014}.
+In later steps, $t_E^*$ and $U_E$ are not conditionally independent of $E$, but we can determine the null conditional distribution through Monte Carlo sampling.
+
+The saturated-model $p$-values also account for selection, but they rely on somewhat different assumptions, condition on more information, and test a slightly less restrictive null hypothesis.
+We discuss these distinctions in detail in Section~\ref{sec:whichnull}.
+
+For most of the early steps in Table~\ref{tab:diab}, the (invalid) nominal test gives the smallest $p$-values, followed by the max-$t$ test, then the saturated-model test.
+Both the max-$t$ and saturated-model $p$-values are exactly uniform under the null, but the max-$t$ $p$-values appear to be more powerful under the alternative.
+As we will discuss in Section~\ref{sec:selective-reg}, selected-model tests such as the max-$t$ can be much more powerful than saturated-model tests in early steps of the path when multiple strong variables are compete with each other to enter the model first.
+Using the max-$t$ $p$-values, ForwardStop selects a model of size 8, compared to size 3 when using the saturated-model $p$-values.
+
+Because the max-$t$ null $p$-values are independent, while the saturated-model $p$-values are not, ForwardStop is guaranteed to control FDR when using the former but not the latter.
+In Section~\ref{sec:pValsIndep} we discuss intuitive sufficient conditions for independence and show that the max-$t$ and many other selected-model tests satisfy these conditions.
+
+\begin{comment}
+The latter two sets of $p$-values test somewhat different hypotheses:  the $p$-values from the saturated model test whether the partial regression of the predictor just added is zero after regressing out the variables already in the model, what we call the ``incremental null''
+In contrast, the max-$t$ $p$-values test whether the current model is adequate; that is, whether all predictors not in the model  have coefficients equal to zero, the ``complete null''.
+(The issue of what hypothesis to test is discussed in section \ref{sec:whichnull}.)
+\end{comment}
+
+\subsection{Outline}
+
+For the remainder of the article we will discuss the problem of selective sequential model selection in some generality, returning periodically to the selective max-$t$ test in forward stepwise regression and its lasso regression counterpart, the next-$\lambda$ test, as intuitive and practically important running examples.
+
+In comparison to their saturated-model counterparts, we will see that the selected-model tests proposed here have three main advantages: power, independence (in some cases), and generalizability beyond linear regression.
+These advantages do not come entirely for free, as we will see: the selected-model methods test a more restrictive null hypothesis than the saturated-model methods.
+
+Section~\ref{sec:genericSetting} introduces the general problem setting, gives important definitions, and reviews relevant literature on testing ordered hypotheses using a sequence of $p$-values.
+Many of these methods require the $p$-values to be uniform and independent under the null, and we derive general conditions to ensure this property in Section~\ref{sec:pValsIndep}.
+In Section~\ref{sec:selective-reg} we contrast selected-model and saturated-model tests for linear regression setting, and illustrate why selected-model tests are often much more powerful in early steps of the path.
+In Section~\ref{sec:computation} we discuss strategies for sampling from the null distribution for our tests, and prove that in forward stepwise regression and $\ell_1$-regularized likelihood methods, the number of constraints in the sampling problem is never more than twice the number of variables.
+We provide a simulation study in Section~\ref{sec:sparseReg} and briefly discuss applications of our framework to principal components analysis in Section~\ref{sec:pca}.
+The paper ends with a Discussion.
+
+\section{Selective Sequential Model Selection}
 \label{sec:genericSetting}
 
-More generically, we observe data $Y \in \cY$, with unknown sampling distribution $F$. We then use $Y$ to generate an adaptive sequence of $d$ nested models
+In the general setting, we observe data $Y \in \cY$ with unknown sampling distribution $F$, and then apply some algorithm to generate an adaptive sequence of $d$ nested statistical models contained in some {\em upper model} $M_{\infty}$.
 \[
@@
 \]
-\WFcomment{should probably end with $M_{d-1}(Y)$. need to fix this throughout paper.}
-\RTcomment{Will- be more specific here---   A parametric model  with
-  specified nonzero support?}
-\WFcomment{Not necessarily. As far as I can tell, there's no need to restrict ourselves in either case. $Y\sim U[0,\theta]$ is a perfectly good model; so is a Cox proportional hazards model.}
-
-By ``model,'' we mean a family of probability distributions for $Y$.
+By {\em model}, we mean a family of candidate probability distributions for $Y$.
 Without loss of generality, we can take $M_\infty$ as the union of all models under consideration.
-We will use the notation $M_{0:k}$ to denote the sequence of the first $k$ models.
+We will use the notation $M_{0:d}$ to denote the sequence $M_0, \ldots, M_d$, which we call the {\em model path}.
+We assume throughout that for each candidate model $M$ there is a minimal sufficient statistic $U(Y; \,M)$, and write
+\[
+U_k(Y) = U(Y; \,M_k(Y)).
+\]
+
+Given data $Y$ and model path $M_{0:d}$, we set ourselves the formal goal of selecting the simplest correct model: the smallest $k$ for which $F\in M_k(Y)$.
+Of course, in most real data problems all of the $M_k(Y)$, as well as all other models under consideration, are incorrect.
+In more informal terms, then, we seek the simplest model in the path that is not refuted by the available data.
 
@@
 By construction, $F\in M_k \iff k \geq k_0$.
-Our goal is to examine the data and output a {\em stopping rule}, i.e.
-an estimator $\hk(Y)$ of $k_0$.
-We consider $M_k$ to be ``rejected'' if $k < \hat k$, and ``accepted'' otherwise.
-Thus $\hk$ is the number of models we rejected, while $k_0$ is the number that we should have rejected.
-
-The number of type I errors is $V=(\hk-k_0)_+$, while the number of type II errors is $(k_0-\hk)_+$.
-Depending on the context we might want to control the familywise error rate (FWER) $\P(V>0)$, the false discovery rate (FDR) $\E[V/(\hk \vee 1)]$, or another error rate given by the expectation of some loss function $g$:
+A {\em stopping rule} is any estimator $\hk(Y)$ of $k_0$, with $M_k$ ``rejected'' if $k < \hk$, and ``accepted'' otherwise.
+Because $\hk$ is the number of models we do reject, and $k_0$ is the number we should reject, the number of type I errors is $V=(\hk-k_0)_+$, while the number of type II errors is $(k_0-\hk)_+$.
+Depending on the scientific context, we might want to control the familywise error rate (FWER) $\P(V>0)$, the false discovery rate (FDR) $\E[V/\hk; \hk>0]$ \citep{benjamini1995controlling}, or another error rate defined by the expectation of some loss function $g(\hk, k_0)$:
 \begin{equation}\label{eq:errRate}
@@
 
-We will assume that for each candidate model $M$ there is a minimal sufficient statistic $T(Y; \,M)$, and write
-\[
-T_k(Y) = T(Y; M_k(Y)).
-\]
-
-First suppose that we have single step p-values for testing
+\subsection{Single-Step $p$-Values}
+
+We will restrict our attention to stopping rules like those recently proposed by \citet{gsell2013sequential} and \citet{li2015accumulation}, which operate on a sequence $p_{1:d}$ of $p$-values.
+At each step $k$, we will construct a $p$-value for testing
 \[
@@
 \]
-against the alternative that $F\in M_\infty\setminus M_{k-1}$. The $p$-values must account for the fact that the models are chosen adaptively. 
-\citet{gsell2013sequential} and \citet{li2015accumulation} propose stopping rules that operate on the sequence $p_{1:d}$.
-These stopping rules provably control certain error rates, provided that the null $p$-values are mutually independent and independent of the non-null $p$-values.
-We review these methods in Section~\ref{sec:sequential} .
-
-Hence the task is to  construct sequential p-values which satisfy the above requirements, and we do this in Section \ref{sec:singleStep}.
-We also discuss general conditions under which single-step $p$-values are independent.
-
-\subsection{Linear Regression}
-
-Most of our examples will focus on the important and familiar problem of adaptively selecting a set of predictors in linear regression models.
-In this problem, we observe a random response $Y\in \R^n$ as well as a fixed design matrix $X\in \R^{n \times p}$, whose columns correspond to candidate predictors.
-For each {\em active set} $E \sub \{1,\ldots,p\}$, there is a corresponding candidate model
+against the alternative that $F\in M_\infty\setminus M_{k-1}$, accounting for the fact that $M_{k-1}$ is chosen adaptively.
+
+Following \citet{fithian2014optimal}, we say that for a fixed candidate model $M$, the random variable $p_{k,M}(Y)$ is a valid {\em selective $p$-value} for $M$ at step $k$ if it is conservative under sampling from any $F\in M$, given that $M$ is selected.
+That is,
+\begin{equation}
+\P_F\left(p_{k,M}(Y) \leq \alpha \mid M_{k-1}(Y) = M\right) 
+\leq \alpha, \quad \forall F\in M, \alpha \in [0,1].
+\end{equation}
+
+Once we have constructed selective $p$-values for each fixed $M$, we can use them as building blocks to construct a combined $p$-value for the random null $M_{k-1}(Y)$. Define
+\[
+p_k(y) = p_{k, M_{k-1}(y)}(y),
+\]
+which is conservative on the event $\{F \in M_{k-1}(Y)\}$:
+\begin{equation}\label{eq:selectiveGuaranteePk}
+\P_F\left(p_k \leq \alpha \mid F\in M_{k-1}\right) \leq \alpha, \quad \forall \alpha \in [0,1].
+\end{equation}
+
+One useful strategy for constructing exact selective tests is to condition on the sufficient statistics of the null model $M_{k-1}$. By sufficiency, the law
+\[
+\L(Y \;\mid\; U_{k-1}, \; M_{k-1}=M)
+\]
+is the same for every $F\in M$.
+Thus, we can construct selective tests by comparing the realized value of any test statistic $T(Y)$ to its known conditional distribution under the null.
+All that remains is to choose the test statistic and compute its conditional null distribution, which can be challenging.
+See \citet{fithian2014optimal} for a general treatment.
+
+\subsection{Sparse Parametric Models}\label{sec:genSparse}
+
+Many familiar path algorithms including lasso, LARS, and forward stepwise regression, are methods for adaptively selecting a set of predictors in linear regression models where we observe a random response $Y\in \R^n$ as well as a fixed design matrix $X\in \R^{n \times d}$, whose columns correspond to candidate predictors.
+For each possible active set $E \sub \{1,\ldots,d\}$, there is a corresponding candidate model
 \[
@@
 \]
-The error variance $\sigma^2$ may be known or unknown depending on the context. The complete sufficient statistic for $M(E)$ is $X_E'Y$ if $\sigma^2$ is known, or else $\left(X_E'Y, \|Y\|^2\right)$.
-
-\subsection{Sparse Parametric Models}\label{sec:genSparse}
-
-All of the specific examples considered in this paper will have a common form generalizing the linear regression problem above.
+If the error variance $\sigma^2$ is known, the complete sufficient statistic for $M(E)$ is $X_E'Y$; otherwise it is $\left(X_E'Y,\, \|Y\|^2\right)$.
+
+For the most part, we can discuss our theory and proposals in a generic parametric setting generalizing the linear regression problem above.
 Let $M_\infty$ be a model parameterized by $\theta\in \Theta \sub \R^{\cJ}$:
@@
 \]
-For any subset $E\sub \cJ$ define the sparse submodel with active coefficients $E$ as follows:
+For any subset $E\sub \cJ$ define the sparse submodel with active set $E$ as follows:
 \[
@@
 
-This setting includes linear regression problems with known or unknown $\sigma^2$ as special cases. A typical path algorithm returns a sequence of nested active sets 
-\[
-E_0(Y) \sub E_1(Y) \sub \cdots \sub E_d(Y),
-\]
-inducing the model path given by $M_k = M(E_k)$.
-
-As an example, in forward stepwise regression with known $\sigma^2$, we might include an unknown intercept term $\beta_0$ in the global null model $M_0$, then at each step $k\geq 1$, add the predictor $j_k$ that reduces the residual sum of squares the most.
-In that case $E_0 = \{0\}$, and $E_k = E_{k-1} \cup \{j_k\}$ for $k\geq 1$.
-
-\subsection{Path Algorithms}
-
-We will be especially interested in two path algorithms in the sparse parametric setting: forward-stepwise paths and ever-active regularization paths.
-
-\paragraph{Forward Stepwise Paths}
+We can consider path algorithms that return a sequence of nested active sets
+\[
+E_0(Y) \sub E_1(Y) \sub \cdots \sub E_d(Y) \sub \cJ,
+\]
+inducing a model path with $M_k = M(E_k),$ for $k=0,\ldots,d$.
+We will be especially interested in two generic path algorithms for the sparse parametric setting: forward stepwise paths and ever-active regularization paths.
+
+\subsubsection{Forward Stepwise Paths and Greedy Likelihood Ratio Tests}
 Let $\ell(\theta; Y)$ denote the log-likelihood for model
-$M_\infty$. The {\em forward stepwise} algorithm proceeds as follows: we begin with some fixed $E_0$, then at step $k=1,\ldots,d$, we define
-\begin{align}\label{eq:forwardDef_start}
-j_k &= \argmax_j \;\;\sup \left\{\ell(\theta; Y):\; \theta\in\Theta(E_{k-1} \cup \{j\})\right\} \\
-E_k &= E_{k-1} \cup \{j_k\}\\\label{eq:forwardDef_end}
-M_k &= M(E_k).
+$M_\infty$. The {\em forward stepwise} algorithm proceeds as follows: we begin with some fixed $E_0$ (such as an intercept-only model), then at step $k=1,\ldots,d$, we set
+\begin{equation}
+j_k = \argmax_j \;\;\sup \left\{\ell(\theta; Y):\; \theta\in\Theta(E_{k-1} \cup \{j\})\right\}, \quad \text{ and } \quad
+E_k = E_{k-1} \cup \{j_k\}.
+\end{equation}
+That is, at each step we select the next variable to maximize the likelihood of the next fitted model.
+
+A natural choice of test statistic is the {\em greedy likelihood ratio statistic}
+\begin{equation}\label{eq:greedyLRT}
+G_k(Y) = \sup_{\theta\in \Theta(E_k)} {\ell(\theta; Y)} \;\;- \sup_{\theta\in \Theta(E_{k-1})} {\ell(\theta; Y)},
+\end{equation}
+the generalized likelihood ratio statistic for testing $M(E_{k-1})$ against the ``greedy'' alternative with one more active parameter:
+\begin{equation}\label{eq:greedyAlternative}
+M(E_{k-1})^{+1} = \bigcup_{j \notin E_{k-1}} M(E_{k-1} \cup \{j\}).
+\end{equation}
+
+We define the {\em selective greedy likelihood ratio test} as the test that rejects for large $G_k$, based on the law
+\begin{equation}
+\L\left(G_k \;\mid\; M_{k-1}, U_{k-1}\right)
+\end{equation}
+
+Because the likelihood in linear regression is a monotone decreasing function of the residual sum of squares, the max-$t$ test is equivalent to the greedy likelihood test.
+
+For simplicity, we have made two assumptions: that only one variable is added at each step, and that the set of candidate variables we choose from is the same in each step.
+It is relatively straightforward to relax either assumption, as we discuss briefly in Section~\ref{sec:discussion}.
+
+\subsubsection{Ever-Active Regularization Paths and next-$\lambda$ Tests}
+Another important class of model selection procedures is the sequence of {\em ever-active} sets for a regularized likelihood path.
+For some ordered set $\Lambda$, let $\lambda\in\Lambda$ parameterize the regularization penalty $P_\lambda(\theta)$.
+In the case of lasso-penalized regression, we could take $P_\lambda = \lambda\|\beta\|_1$ for $\lambda\in (0,\infty]$.
+
+We then define
+\begin{align}
+  \hat\theta^{\lambda}(Y) &= 
+  \argmin_{\theta\in\Theta} -\ell(\theta; Y) + P_\lambda(\theta) \\
+  \tE_\lambda(Y) &= \left\{j:\; \hat\theta_j^\nu \neq 0 
+    \text{ for any } \nu\in \Lambda, \nu \geq \lambda \right\}
 \end{align}
-That is, at each step we add to the model the one variable that would most increase the likelihood.
-In the case of linear regression, this amounts to adding the variable that maximally reduces the residual sum of squares.
-
-\paragraph{Ever-Active Regularization Paths}
-Another class of model selection procedures is the sequence of {\em ever-active} sets for a regularized likelihood path.
-For $r=0,1,\ldots,m$, let $P_r(\theta)$ denote some regularization penalty, and define
-\begin{align}\label{eq:regPathDef_start}
-  \hat\theta^{r}(Y) &= 
-  \argmin_{\theta\in\Theta} -\ell(\theta; Y) + P_r(\theta) \\
-  \tE_r(Y) &= \left\{j:\; \hat\theta_j^s \neq 0 
-    \text{ for any } s \leq r \right\}
-\end{align}
-
-While the sets $\tE_r$ are nested by definition, we could have $\tE_r = \tE_{r+1}$ for most values of $r$.
-Thus, we will take the sequence of distinct ever-active sets.
-Formally, let $R_0=0$, and for $k\geq 1$ let $R_k$ denote the (random) index where the active set actually changes for the $k$th time.
-That is,
-\begin{align}
-  R_k &= \min\{s:\; \tE_s \neq \tE_{R_{k-1}}\}\\
-  E_k &= \tE_{R_k}\\
-  \label{eq:regPathDef_end}
-  M_k &= M(E_k)
-\end{align}
-
-The ever-active lasso path in linear regression is the most familiar example, in which we take $P_r(\beta) = r^{-1}\|\beta\|_1$, and the penalized log-likelihood criterion is
-\begin{equation}
-  \argmin \frac{n}{2}\log(2\pi\sigma^2) 
-  + \frac{1}{2\sigma^2}\|Y - X\beta\|_2^2 + \frac{1}{r}\|\beta\|_1
-\end{equation}
-If $\sigma^2$ is known, then $\hat\beta^{r}$ is the lasso solution with Lagrange parameter $\lambda^r = \sigma^{2}/r$.
-If $\sigma^2$ is unknown, it is the lasso solution for $\lambda^r = (\hat\sigma^r)^2/r$.
-Because $\hat\sigma^r = n^{-1/2}\|Y-X\hat\beta^r\|_2$ is decreasing in $r$, $\lambda^r$ is strictly decreasing in $r$.
-
-Because it is actually possible to solve the lasso at every value of the Lagrange parameter, it is not really necessary that $r$ only take on integer values.
-We can instead notionally take $r\in (0,\infty)$, and use a path algorithm like LARS \citep{taylor2014exact} to find every point in the path where the lasso solution changes.
-
-\subsection{``Model-centric'' and ``variable-centric'' null hypotheses}
+It may be impossible or inconvenient to compute $\hat\theta^\lambda$ for every $\lambda\in(0,\infty]$. If so, we can instead take $\Lambda$ to be a grid of finitely many positive values.
+
+Note that the active sets $\tE_\lambda$ are nested by construction.
+In addition, we will assume that $|\tE_\lambda|<\infty$ for every $\lambda\in\Lambda$.
+Our model path will correspond to the sequence of distinct ever-active sets.
+Formally, let 
+\[
+\lambda_0=\sup \Lambda, \quad \text{ and } \quad 
+E_0 = \bigcap_{\lambda\in \Lambda} \tE^\lambda,
+\]
+and for $k\geq 1$ let $\lambda_k$ denote the (random) value of $\lambda$ where the active set changes for the $k$th time:
+\begin{equation}
+  \Lambda_k = \{\lambda\in \Lambda:\; \tE_\lambda \supsetneq \tE_{\lambda_{k-1}}\},
+  \quad
+  \lambda_k = \sup\Lambda_k,
+  \quad \text{ and } \quad
+  E_k = \bigcap_{\lambda\in\Lambda_k} \tE_{\lambda}.
+\end{equation}
+
+In this setting, $\lambda_k$ is a natural test statistic for model $M_{k-1}$, with larger values suggesting a poorer fit.
+The {\em selective next-$\lambda$ test} is the test that rejects for large $\lambda_k$, based on the law
+\begin{equation}
+\L\left(\lambda_k \;\mid\; U_{k-1}, M_{k-1}\right).
+\end{equation}
+
+\subsection{Which Null Hypothesis?}
 \label{sec:whichnull}
 
-\WFcomment{Shorten this section, move some of it to Discussion?}
-
-In our formulation of the problem, the type I error $V=(\hk-k_0)_+$ is defined in a ``model-centric'' fashion: at step $k$, we are testing  whether a particular linear model $M(E_{k-1})$ adequately describes the data $Y$.
-Even if the next selected variable $X_{j_k}$ is complete noise, adding it is not a mistake as long as there are other signal variables that have not yet been included.
-
-In some circumstances, we might want to define a type I error at step $k$ differently.
-If so, we must choose a different null hypothesis to test.
+Note that in our formulation of the problem, the type I error $V=(\hk-k_0)_+$ is defined in a ``model-centric'' fashion: at step $k$, we are testing  whether a particular linear model $M(E_{k-1})$ adequately describes the data $Y$.
+Even if the next selected variable $X_{j_k}$ is a noise variable, it is not necessarily a mistake to reject $M(E_{k-1})$ provided there are other signal variables that have not yet been included.
+
+In some circumstances, we might want to define a type I error at step $k$ differently, by choosing a different null hypothesis to test.
 Let $\mu = \E Y$ and let $\theta^E$ denote the {\em least-squares coefficients} of active set $E$ --- the coefficients of the best linear predictor for the design matrix $X_E$:
@@
 \]
-where $A^\dagger$ is the Moore-Penrose pseudoinverse of the matrix $A$. 
+where $X_E^\dagger$ is the Moore-Penrose pseudoinverse of the matrix $X_E$. 
 These are the true ``partial regression coefficients''  for design matrix $X_E$.
@@
 \begin{description}
-\item[Complete Null:] $M_{k-1}$ is already correct; i.e.,
-\[
-H_k:\;\mu = X_{E_{k-1}} \theta^{E_{k-1}}
-\]
-
+\item[Complete Null:] $M_{k-1}$ is already correct. That is, 
+\[
+H_k:\;\mu = X_{E_{k-1}} \theta^{E_{k-1}}.
+\]
 \item[Incremental Null:] $M_{k-1}$ may be incorrect, but $M_k$ is no improvement. That is, 
 \[
-H_k^{\text{inc}}:\; \theta_{j_k}^{E_k} = 0
-\]
-\item[Full-Model Null:] The coefficient of $X_{j_k}$ is zero in the     full model; that is, 
-\[
-H_k^{\text{full}}:\; \theta_{j_k}^{[p]} = 0.
+H_k^{\text{inc}}:\; \theta_{j_k}^{E_k} = 0.
+\]
+\item[Full-Model Null:] The coefficient of $X_{j_k}$ is zero in the ``full'' model with all $p$ predictors. That is,
+\[
+H_k^{\text{full}}:\; \theta_{j_k}^{\{1,\ldots,p\}} = 0.
 \]
 \end{description}
-See \citet{gsell2013false} for an in-depth discussion of the relative conceptual advantages and disadvantages between these approaches.
 
@@
 \end{align}
-we can define an analogous FWER and FDR with respect to each of these alternative choices, and attempt to control these error rates. For example, we can define
+we can define an analogous FWER and FDR with respect to each of these alternative choices, and attempt to control these error rates. For example, we could define
 \[
@@
 \]
-as the false discovery rate with respect to the full-model null.
-
-In this article we will concern ourselves primarily with testing the complete null, which conforms most naturally to our stated aim of choosing the least complex model that is not rejected by the data, and is most easily generalized to the generic setting of Section~\ref{sec:genericSetting}.
-
-The incremental null is closely related to {\em saturated-model} tests for linear regression, as we will discuss further in Section~\ref{sec:singleStep}.
-While we could use the saturated-model framework to construct valid selective tests for $H_k^{\text{inc}}$, as proposed for LARS and lasso regression in \citet{taylor2014exact}, this will often result in a large reduction in power as well as introducing dependence between $p$-values at different steps.
-
-The full-model null is the most different conceptually from the other two.
-It takes a ``variable-centric'' instead of ``model-centric'' point of view: any variable with a nonzero coefficient in the full model is a {\em signal variable} and the others are {\em noise variables}, and $V^{\text{full}}$ counts the number of noise variables incorrectly selected.
-
-If the design matrix of the full model represents a scrupulously curated set of features, then the analyst may be primarily interested in inferences with respect to the full model.
-For example, the scientist may believe, due to theoretical considerations, that the linear model in $X_1, \ldots, X_p$ is fairly credible, and that a nonzero coefficient of $X_1$ after controlling for {\em all} of the other variables would constitute strong evidence for a causal effect of $X_1$ on the response.
-
-If the full model enjoys no special scientific status, however, there is little advantage in insisting that all inferences should adjust for every other predictor in the full design matrix $X$.
-For example, suppose that $X$ contains gene expression measurements for all genes that happened to be measured by a microarray chip, but if we had purchased the chip from a different manufacturer, we would have measured a different set of genes.
-Because the full-model coefficients are already artifacts of the somewhat arbitrary choice of which chip we bought, there is no clear reason to regard them as more scientifically interesting than the coefficients of a well-chosen submodel.
-
-See \citet{barber2014controlling} for an exact method to control $\text{FDR}^{\text{full}}$.
-\WFcomment{many high-dim inference references}.
-Using selective inference to control type I error rates for $H_k^{\text{full}}$ is an interesting topic for further study.
-
-\section{Sequential Inference from ordered p-values}
-\label{sec:sequential}
+as the false discovery rate with respect to the full-model null. \citet{barber2014controlling} present a framework for controlling $\text{FDR}^{\text{full}}$.
+
+The full-model null is the most different conceptually from the other two, taking a ``variable-centric'' instead of ``model-centric'' viewpoint, where it is of primary scientific interest to discover variables that have nonzero coefficients after controlling for all $p-1$ other variables that were measured.
+
+To elucidate this distinction consider two methods we could apply in a classical setting where we have some fixed {\em a priori} ordering for the variables, inducing the fixed sequence of nested linear models
+\[
+M(\emptyset) \sub M(\{1\}) \sub M(\{1,2\}) \sub \cdots \sub M(\{1,\ldots, p\}).
+\]
+
+In the first procedure, we construct an analysis-of-variance (ANOVA) table performing an $F$- or $t$-test of each model against the next.
+We then stop and select model $M_k=M(\{1,\ldots,k\})$ the first time the $F$-test $p$-value is larger than $\alpha$.
+This procedure would control the complete-null and incremental-null FWER, but {\em not} the full-model-null FWER.
+The methods we consider in this article are essentially generalizations of this procedure for the setting where the model sequence is selected adaptively, and where we may wish to control FDR instead of FWER.
+
+In the second procedure, we fit the full model $M(\{1,\ldots,p\})$ and compute a $t$-statistic for each variable with respect to that model, yielding $t$-statistics $t_1, t_2, \ldots, t_p$ (not ordered by size).
+Iterating for $j=1,\ldots,p$, we compare each $|t_j|$ to the level-$\alpha$ cutoff for a $t$-test with $n-p$ degrees of freedom, and stop selecting variables the first time $|t_{k+1}|$ is not larger than this cutoff.
+This procedure would control the full-model-null FWER, but {\em not} the incremental-null FWER.
+Note that the {\em model} $M_k$ plays no role in this procedure: its fit to the data is never evaluated and it is not really selected in any meaningful sense.
+Even if the fit of $M_k$ is far worse than that of $M_{k+1}$, we may not be able to select $X_{k+1}$ if (say) it is highly collinear with $X_{k+100}$.
+Conversely, even if the fit of $M_k$ is no better than that of $M_{k-1}$, we may still select $X_k$ if it is a highly useful predictor when used in concert with $X_{k+100}$.
+
+More generally, methods that control full-model error rates are best viewed not as {\em model-selection} procedures --- since all inferences are made with respect to the full model, which is specified ahead of time --- but {\em variable-selection} procedures that test multiple hypotheses with respect to a single fixed model.
+Note that the truth or falsity of these hypotheses can depend sensitively on the set of $p$ predictors: rejecting $H_j^{\text{full}}$ has no meaning without reference to the list of all variables that we controlled for, which can make a rejection difficult to interpret when $p$ is large.
+Thus, error rates like $\text{FDR}^{\text{full}}$ are best motivated when the full model has some special scientific status.
+For example, the scientist may believe due to theoretical considerations that the linear model in $X_1,\ldots,X_p$ is fairly credible, and that a nonzero coefficient $\beta_j$, controlling for all of the other variables, would constitute evidence for a causal effect of $X_j$ on the response.
+
+In this article we will concern ourselves primarily with testing the complete null, reflecting our stated aim of choosing the least complex model that is not rejected by the data.
+As we discuss further in Section~\ref{sec:selective-reg}, the saturated-model tests of \citet{taylor2014exact} are valid selective tests of $H_k^{\text{inc}}$.
+Unfortunately, this requires us to assume $\sigma^2$ is known, can result in a large reduction in power, creates dependence between $p$-values at different steps, and is difficult to generalize beyond the case of linear regression.
+See \citet{gsell2013false} for an in-depth discussion of the relative conceptual advantages and disadvantages between the three approaches.
+
+\subsection{Related Work}
+
+The problem of model selection is an old one, with quite an extensive literature.
+However, with the exception of the works above, very few methods offer finite-sample guarantees on the model that is selected except in the orthogonal-design case.
+One additional exception is the knockoff filter of \citet{barber2014controlling}, a variant of which provably controls the full-model FDR.
+
+Methods like AIC \citet{akaike1974new} and BIC \citep{schwarz1978estimating} are designed for the non-adaptive case, where the sequence of models is determined in advance of observing the data.
+Cross-validation, another general-purpose algorithm for tuning parameter selection, targets out-of-sample error and tends to select many noise variables when the signal is sparse \WFcomment{Whom to cite?}.
+\citet{benjamini2009simple} extend the AIC by using an adaptive penalty $\sigma^2\lambda_{k,p} k$ to select a model, based on generalizing the Benjamini-Hochberg procedure \citet{benjamini1995controlling}, but do not prove finite-sample control.
+The stability selection approach of \citet{meinshausen2010stability} uses an approach based on splitting the data many times and offers asymptotic control of the FDR, but no finite-sample guarantees are available.
+
+\section{Stopping Rules, Ordered Testing, and Independence}
 
 An ordered hypothesis testing procedure takes in a sequence of $p$-values $p_1, \ldots, p_d$ for null hypotheses $H_{1}, \ldots, H_{d}$, and outputs a decision $\hk$ to reject the initial block $H_{1}, \ldots, H_{\hk}$ and accept the remaining hypotheses.
-We will assume throughout that the hypotheses are nested with $H_{k-1} \sub H_k$.
-The next Section outlines several proposals for ordered testing and discusses what error rates they control and under what conditions.
+Note that in our setup the hypotheses are nested, with $H_{k-1} \sub H_k$; as a result, all of the false hypotheses precede all of the true ones.
+
+We first review several proposals for ordered-testing procedures, several of which require independence of the null $p$-values conditional on the non-null ones.
+While these procedures also assume the sequence of hypotheses is fixed, in our setting, the truth or falsehood of $H_k$ is random, depending on which model is selected at step $k$.
+However, we show that the error guarantees for these stopping rules do transfer to the random-hypothesis setting, provided we have the same independence property conditional on $k_0$.
 
@@
 \subsubsection{BasicStop}
-The most obvious procedure is simply to reject at each step until the first time that $p_k > \alpha$, which we can formalize as
-\[
-\hk_B(Y) = \min\left\{k \in \{1,\ldots,m\} :\;
-  p_k > \alpha\right\} - 1
-\]
-We will call this procedure {\em BasicStop}. 
-It is discussed in \citet{marcus1976}.
-
-If $k_0$ is fixed, then it is clear that $\hk_b$ controls the FWER at level $\alpha$ provided that $p_{k}$ is a valid $p$-value for each $k>k_0$. Then, 
-\[
-\P(V>0) = \P(p_{k_0+1} \leq \alpha) \leq \alpha.
-\]
-
-Selectively valid single-step $p$-values do not necessarily guarantee FWER control when $k_0$ is random.
-In that case, we need a bit more, but it is sufficient that we have type I error control for $p_k$ conditional on $k_0=k-1$; i.e., not just conditional on $M_{k-1}$ being a correct model, but given that $M_{k-1}$ is the {\em first} correct model.
-
-\subsubsection{StrongStop}
-
-\citet{gsell2013sequential} propose a stopping rule that controls the FWER, which they called {\em strong stop}. Define
-\[
-  \hk_{S}(Y) = \max\left\{k \in \{1,\ldots,d\} :\;
-    \exp\left(\sum_{i=k}^d \frac{\log p_i}{i}\right) 
-    \leq \frac{\alpha k}{d}\right\}
-\]
-StrongStop uses all of the p-values after step $k$ in its decision to reject at step $k$. 
-Even if $p_k$ is a little larger than $\alpha$, say $\alpha=0.05$ and $p_k=0.06$, StrongStop can still reject $M_{k-1}$ if the next three or four $p$-values are also relatively small.
-
-\citet{gsell2013sequential} show that if the completion index $k_0$ is fixed, then $\hk_S$ controls the FWER at level $\alpha$, as long as the null $p$-values are independent given the non-null ones.
+The simplest procedure is to keep rejecting until the first time that $p_k > \alpha$: 
+\[
+\hk_B(Y) = \min\left\{k:\; p_k > \alpha\right\} - 1
+\]
+We will call this procedure {\em BasicStop}. It is discussed in \citet{marcus1976}.
+
+Because our hypotheses are nested, we have $\hk > k_0$ if and only if $p_{k_0+1} \leq \alpha$.
+If $k_0$ is fixed, then, it is clear that $\hk_B$ controls the FWER at level $\alpha$ provided that $p_k$ is a valid $p$-value for each $k$.
 
@@
 
-\citet{gsell2013sequential} propose another stopping rule, {\em forward stop}, that controls the FDR:
-\[
-  \hk_{F}(Y) = \max\left\{k \in \{1,\ldots,d\} :\;
+\citet{gsell2013sequential} propose the {\em ForwardStop} rule to controls the FDR:
+\[
+  \hk_{F}(Y) = \max\left\{k:\;
     -\frac{1}{k}\sum_{i=1}^k \log(1-p_i) \leq \alpha\right\}
 \]
-ForwardStop uses the p-values {\it before} step $k$ in its decision to reject at step $k$. 
-If $p_k$ is uniform then  $-\log(1-p_k)$ is an exponential random variable. If all of the null $p$-values are uniform and the others are zero, then only the null ones contribute to the sum, and
-\[
-\widehat{\text{FDR}}_k = -\frac{1}{k}\sum_{i=1}^k \log(1-p_i)
-\]
-is a Gamma random variable with mean $V/k$ and variance $V/k^2$. We stop the last time the estimate of FDR is less than $\alpha$.
+ForwardStop uses the $p$-values {\it before} step $k$ in making its decision whether to reject at step $k$.
+If $p_k$ is uniform, then $-\log(1-p_k)$ is an $\text{Exp}(1)$ random variable with expectation 1; thus, the sum can be seen as an estimate of the false discovery proportion (FDP):
+\[
+\widehat{\text{FDP}}_k = -\frac{1}{k}\sum_{i=1}^k \log(1-p_i).
+\]
+ForwardStop chooses the largest model for which $\widehat{\text{FDP}}_k \leq \alpha$.
+\citet{gsell2013sequential} show that ForwardStop controls the FDR if the null $p$-values are independent of each other and of the non-nulls.
 
@@
 
-\citet{li2015accumulation} generalize the ForwardStop rule, introducing the family of {\em accumulation tests}, defined as follows.
-Fix a function $h: [0,1] \rightarrow [0,\infty]$ satisfying $\int _{t=0}^1 h(t)dt=1$, called an ``accumulation function''. They define an estimate of the
-false discovery proportion at a cutoff $k$ as
-\begin{equation}
-\widehat{\rm FDP}_h(k)= \frac{\sum_{i=1}^k h(p_i)}{k}
-\end{equation}
-and then define the adaptive cutoff,  for FDR level $\alpha$, to be                                                                                                                 as
-\begin{equation}
-\hat k_h={\rm max} \Bigl\{k\in \{1,2,\ldots n\}; \widehat{\rm FDP}_h(k) \leq \alpha \Bigr\}
-\end{equation}
-The ForwardStop rule falls into this class:  another  choice they study is the ``HingeExp'' function
-\begin{equation}
-h_{\rm HingeExp}(t)= 
-\begin{cases} C\cdot\log\Bigl(\frac{1}{C(1-t)} \Bigr)& \text{for $T> 1-1/C$}\\
-                         0                                         & \text{for $T\leq 1-1/C$}
+\citet{li2015accumulation} generalize ForwardStop, introducing the family of {\em accumulation tests}, defined as follows.
+Fix a function $h: [0,1] \rightarrow [0,\infty]$ satisfying $\int_{t=0}^1 h(t)dt=1$, called an {\em accumulation function.} Using $h$ they construct an estimate of FDP at step $k$:
+\begin{equation}
+\widehat{\text{FDP}}_h(k)= \frac{\sum_{i=1}^k h(p_i)}{k},
+\end{equation}
+and select the largest model for which $\widehat{\text{FDP}}_k \leq \alpha$:
+\begin{equation}
+\hat k_h=\max \left\{k:\; \widehat{\text{FDP}}_h(k) \leq \alpha \right\}.
+\end{equation}
+Taking $h(p)=-\log(1-p)$ recovers the ForwardStop rule. Another choice they propose is the {\em HingeExp} function
+\begin{equation}
+  h_{\rm HingeExp}(p)= 
+  \begin{cases} 
+    C\log\left(\frac{1}{C(1-p)} \right)& p> 1-1/C\\
+    0 & p\leq 1-1/C
   \end{cases}
@@
 
+\citet{li2015accumulation} show that accumulation tests control a modified FDR criterion provided that the null $p$-values are $U[0,1]$, and are independent of each other and of the non-nulls.
+
+\begin{comment} 
 Now let ${\rm FP}(k)$ be the actual number of false positives up to and including step $k$.
@@
 
-Importantly, this class of accumulation tests require that  the p-values $p_i$  are  $U(0,1)$ if the null hypothesis $H_i$ is true, and are independent of the non-null p-values.
-This implies that the class of sequential pvalues proposed in this paper can be used as inputs to these tests.
-
-\subsection{Random hypotheses}
-The forementioned proposals of  \citet{gsell2013sequential} and \citet{li2015accumulation}, all assume that the null hypotheses $H_{1}, \ldots, H_{d}$ are fixed, and that the null $p$-values are independent of each other and of the non-null $p$-values.
+Importantly, this class of accumulation tests require that  the $p$-values $p_i$  are  $U(0,1)$ if the null hypothesis $H_i$ is true, and are independent of the non-null $p$-values.
+This implies that the class of sequential $p$-values proposed in this paper can be used as inputs to these tests.
+\end{comment}
+
+\subsection{Random Hypotheses}
+
+In our problem setting, the sequence of selected models is random; thus, the truth or falsehood of each $H_k$ is not fixed, as assumed in the previous section.
+This could in principle lead to violation of the error control guarantees that hold in the fixed-hypothesis case.
+
+To see why we a problem could arise, consider apply BasicStop in the following counterexample: let $d=2$ and let $k_0$ be 0 or 1 with probability 1/2 each. Then, let 
+\[
+p_1 \sim U[0,1], \quad p_2 \sim (1-k_0 + U[0,1])/2.
+\]
+Both $p_1$ and $p_2$ are uniform, but
+\begin{align*}
+\P(V>0) &= \P(p_{k_0+1} \leq \alpha)\\ 
+&= \sum_{k=1}^2\P(p_k\leq \alpha \mid k_0+1=k)\cdot \P(k_0+1=k)\\
+&= \alpha\cdot \frac{1}{2} + 2\alpha \cdot \frac{1}{2} = 3\alpha/2
+\end{align*}
+This problem is resolved if $p_k$ is a valid $p$-value conditional on the value of $k_0$.
+
+Similarly, the proposals of  \citet{gsell2013sequential} and \citet{li2015accumulation} assume that the null hypotheses $H_{1}, \ldots, H_{d}$ are fixed, and that the null $p$-values are independent of each other and of the non-null $p$-values.
 Specifically, they assume that
@@
 
-in our problem the null hypotheses at each step are random and thus so is the completion index $k_0(Y)$.     We will thus require {\em conditional} independence of null $p$-values given $k_0$:
+In our problem the null hypotheses at each step are random, and therefore so is the completion index $k_0(Y)$, so the error control guarantees of \citet{gsell2013sequential} and \citet{li2015accumulation} do not directly apply.
+However, we will recover these guarantees if we have {\em conditional} independence of null $p$-values given $k_0$:
 \begin{equation}\label{eq:indepCond_random_k0}
@@
 \begin{proposition}
-  Let $\hk$ be a stopping rule that operates on sequences
-  $p_{1:d}(Y)$ of $p$-values for ordered hypotheses $H_{1:d}(Y)$. 
-  For some function $g$, let 
-  \[
-  \Err(\hk,g) = \E\left[ g\left(\hk(Y), k_0(Y)\right)\right],
-  \]
-
-  Suppose that the rule $\hk$ controls $\Err$ at $\alpha$ under
-  independence when $H_{1:d}$ are fixed. That is, suppose that
-  whenever $H_{1:d}$ are fixed and $p_{1:d}$
-  satisfy~\eqref{eq:indepCond_fixed_k0}, we have
-  \[
-  \Err(\hk, g) \leq \alpha
-  \]
-
-  Then, $\hk$ controls $\Err$ at $\alpha$
-  whenever $p_{1:d}$ satisfy~\eqref{eq:indepCond_random_k0}.
+  Let $\hk$ be a stopping rule operating on $p$-values $p_{1:d}(Y)$ for nested hypotheses $H_{1:d}(Y)$. For some function $g$, let
+  \[
+  \Err_g = \E\left[ g\left(\hk(p_{1:d}(Y)), k_0(Y)\right)\right],
+  \]
+
+  Suppose that $\hk$ controls $\Err_g$ at level $\alpha$ if the $H_{1:d}$ are fixed and $p_{1:d}$ satisfy~\eqref{eq:indepCond_fixed_k0}.
+Then, $\hk$ controls $\Err_g$ at level $\alpha$ whenever $p_{1:d}$ satisfy~\eqref{eq:indepCond_random_k0}.
 \end{proposition}
 \begin{proof}
-  Fix any values $k_0\in \{0,\ldots,d, \infty\}$ and $q_{1:k_0} \in [0,1]^{k_0}$.
-We can construct a problem with $\tilde p_k \sim \delta_{q_k}$ for $k\leq k_0$ and  $\tilde p_k \sim \text{Unif}[0,1]$ for $k > k_0$.
-
-  Then, for the fixed hypotheses 
-  $\widetilde H_k:\; \tilde p_k \sim \text{Unif}[0,1]$, we have
-  \[
-  f(k_0, q_{1:k_0}) 
-  = \E\left[ g\left(\hk(\tilde p_1, \ldots, \tilde p_d),
-      k_0\right) \right],
-  \]
-  which is less than or equal to $\alpha$, since $\hk$ is assumed to control $\Err$ in the fixed-hypothesis case.
-
-  But then, by the condition~\eqref{eq:indepCond_random_k0}, we have
-  \begin{align}
-    \Err(\hk,g) &\;=\; \E\left[ \;\E\left[ g\left(\hk(p_1,\ldots, p_d), k_0(Y)\right) \; \mid \; k_0(Y), \; p_{1:k_0}(Y)\right] \; \right]\\
-    &\;\leq\; \E\left[f(k_0, p_{1:k_0})\right] \;\leq\; \alpha.
-  \end{align}
+    For nested hypotheses, $k_0$ completely determines the truth or falsehood of $H_k(Y)$ for every $k$. If $\hk$ controls $\Err_g$ in the 
+fixed-hypothesis case, it must in particular control $\Err_g$ conditional on any fixed values of $p_{1:k_0}$. Thus, \eqref{eq:indepCond_random_k0} implies
+  \[
+  \E\left[ g\left(\hk(p_1,\ldots,p_d), k_0(Y)\right) \;\mid\;
+      k_0(Y), \; p_{1:k_0}(Y)\right] \leqAS \alpha.
+  \]
+  Marginalizing over $(p_{1:k_0},k_0)$ gives $\Err_g\leq \alpha$.
 \end{proof}
 
-In real applications, null $p$-values may be conservative instead of exactly uniform.
+\begin{comment}
+In real applications, null $p$-values could be conservative instead of exactly uniform, for instance if the test statistic is discrete or the null is composite.
 We can relax~\eqref{eq:indepCond_random_k0} to require only that the joint distribution of null $p$-values is stochastically larger than uniform:
@@
 \end{equation}
-If the null $p$-values are marginally uniform, then the conditions~\eqref{eq:indepCond_random_k0} and~\eqref{eq:stochLargeCond_random_k0} are equivalent.
+If the null $p$-values are each uniform given $(p_{1:k_0},k_0)$, then the conditions~\eqref{eq:indepCond_random_k0} and~\eqref{eq:stochLargeCond_random_k0} are equivalent.
 For most stopping rules $\hk$, the weaker condition~\eqref{eq:stochLargeCond_random_k0} is enough to guarantee control of whatever error rate holds under~\eqref{eq:indepCond_random_k0}.
+\end{comment}
+
 For the sake of brevity, we will say that $p$-value sequences satisfying~\eqref{eq:indepCond_random_k0} are {\em independent on nulls}.
 
-Sections~\ref{sec:pValsIndep}--\ref{sec:SSP} discuss conditions on the model sequence $M_{0:d}$ and the $p$-value sequence $p_{1:d}$ under which~\eqref{eq:stochLargeCond_random_k0} is satisfied.
-
-\section{Construction of Independence of p-values}\label{sec:pValsIndep}
-
-From the previous section, we see that our task is to construct p-values with the {\it independent on nulls} property (\ref{eq:indepCond_random_k0}).
-In this section we develop a set of sufficient conditions for such a construction. Before doing so, we give an example to motivate
-the general theory.
-
-\medskip
-\begin{example}[The selective max-Z test]
-Consider the forward stepwise path (\ref{eq:forwardDef_start}---\ref{eq:forwardDef_end}) in the case of linear regression, that is 
-$\ell(\theta,Y)=||Y-X\theta||^2$.  Assume that the error variance $\sigma^2$ is known.
- Given an active set of variables $E$ in our current model $M$, the partial regression coefficients for predictors not in $E$ are
-\begin{equation}
-  z_{j,E}(Y) = \frac{(\proj_E^\perp X_j)'Y}
-  {\sigma\|\proj_{E}^\perp X_j\|_2}.
-\end{equation}
-The max-$z$-statistic for active set $E$  is
-\begin{equation}\label{eq:maxT}
-  U_E(Y) = \max_{j \notin E} \left|t_{j,E}(Y)\right|,
-\end{equation}
-and we use $U_{E_{k-1}}(Y)$ as our test statistic at step $k$. We compute a p-value for this test
-using the conditional law 
-\begin{equation}\label{eq:selModel_marg0}
-\L\left(X_{j_k}'Y \mid E_{k-1}, \; X_{E_{k-1}}'Y\right),
-\end{equation}
-If we reject the hypothesis, we add to $E$ the predictor that maximized $U_E(Y)$ and continue.
-This is the procedure that used to obtain the p-values in Table \ref{tab:diab}, except that we sampled
-conditioning on $||Y||$ as well (since $\sigma^2$ is unknown).
-
-Happily,  it turns that the  p-values from this procedure have exactly the desired {\it independent on nulls} property~\eqref{eq:indepCond_random_k0}.
-This happens because of some special characteristics of the construction: 
-\begin{description}
-\item{(a)}  the inner products $X_E'Y$ are complete sufficient statistics for the regression
-parameters in each model $M$,
-\item{(b)} knowing only the  model $M$ at step $k$ and its sufficient statistics $X'_EY$, we can compute the entire subpath
-of models from the steps 1 to $k-1$. 
-\end{description}
- In other words, the inner products $X'_EY$ contains  all of the cumulative information used in the selections from
-steps 1 to $k-1$. This in turn results from    the choice  of test statistic $U_E(Y)$  and the law (\ref{eq:selModel_marg0}) used to calibrate it.
-We call (b)  the {\it Subpath sufficiency principle} (SSP) in Section \ref{sec:SSP}  below.
-In contrast, the {\it saturated model} p-values use a law different from (\ref{eq:selModel_marg0}) to compute p-values, one that uses
-information not contained in $X'_EY$. It thus violates the SSP property and hence does not yield p-values that are independent on nulls.
-We discuss the regression setting in more detail in Section \ref{sec:selective-reg}.  $\square$
-\end{example}
-
-\medskip
-
-We now present a general framework for these ideas, culminating in our theorem \ref{thm:suffCond} at the end of this section below.
-We say that the  single-step $p$-value $p_k$ is selectively valid for testing $H_k:\; F \in M_{k-1}(Y)$ against $M_\infty \setminus M_{k-1}$ if, for all $F\in M_\infty$,
+The next Section discusses conditions on the model sequence $M_{0:d}$ and the $p$-value sequence $p_{1:d}$ under which~\eqref{eq:indepCond_random_k0} is satisfied.
+
+\section{Conditions for Independent $p$-values}\label{sec:pValsIndep}
+
+We now develop sufficient conditions for constructing $p$-values with the {\it independent on nulls} property (\ref{eq:indepCond_random_k0}).
+To begin, we motivate the general theory by discussing a specific case, the max-$t$ test for forward stepwise regression.
+
+\subsection{Independence of max-$t$ $p$-values}
+
+Recall that at step $k$, the max-$t$ test rejects for large $T_k=|t_{E_{k-1}}^*|$, comparing its distribution to the conditional law
+\begin{equation}\label{eq:maxTnull}
+\L\left(T_k \; \mid \; E_{k-1}, X_{E_{k-1}}'Y, \|Y\|^2 \right).
+\end{equation}
+This conditional distribution is the same for all $F\in M(E_{k-1})$, because $U_{k-1}=(X_{E_{k-1}}'Y,\; \|Y\|^2)$ is the complete sufficient statistic for $M(E_{k-1})$.
+
+Now, if $F\in M(E_{k-1})$, then $p_k$ is uniform and independent of the previous $p$-values $p_{1:k-1}$, for two reasons:
+\begin{enumerate}
+\item $p_k$ is uniform conditional on $E_{k-1}$ and $U_{k-1}$ by construction, and
+\item $p_1, \ldots, p_{k-1}$ are functions of $E_{k-1}$ and $U_{k-1}$, as we will see shortly.
+\end{enumerate}
+
+Informally, the pair $(E_{k-1}, U_{k-1})$ forms a ``wall of separation'' between $p_k$ and $p_{1:k-1}$, guaranteeing that $\L(p_k \mid p_{1:k-1}) = U[0,1]$ whenever $H_k$ is true.
+
+Next we will explain why $p_1,\ldots,p_{k-1}$ are functions of $E_{k-1}$ and $U_{k-1}$.
+Knowing $E_{k-1}=E$ tells us that the $k-1$ variables in $E$ are selected first, and knowing $(X_{E_{k-1}}'Y,\|Y\|^2)$ is enough information to compute the $t$-statistics $t_{j,D}$ for all $j\in E$ and $D\sub E$.
+As a result, we can reconstruct the order in which those $k-1$ variables were added, since for $s<k$,
+\[
+j_s = \argmax_{j \notin E_{s-1}} |t_{j,E_{s-1}}| 
+= \argmax_{j \in E\setminus E_{s-1}} |t_{j,E_{s-1}}| \quad \text{ on the event } \{E_{k-1}=E\}.
+\]
+In general, we say that the path $M_{0:d}$ follows the {\em subpath sufficiency principle} (SSP) if for each $k$ the subpath $M_{0:k}$ is a function of $(M_k,U_k)$.
+
+Furthermore, for $s<k$, $p_s$ is computed by comparing $T_s=|t_{j_s, E_{s-1}}|$ to the upper $\alpha$ quantile of the distribution $\L(T_s \mid E_{s-1}, U_{s-1})$, computed under the assumption $F\in M(E_{s-1})$.
+The conditional null distribution is a function of $(E_{s-1}, U_{s-1})$, and as we have already seen, the test statistic $T_s$ is a function of $(E_s, U_s)$, both of which are functions of $(E_{k-1},U_{k-1})$.
+
+Because $(E_s,U_s)$ is a function of $(E_k,U_k)$ for $s<k$, we can characterize the above discussion in terms of the filtration $\sF_{0:d}$, where
+\[
+\sF_k = \sF(E_k, U_k),
+\]
+the $\sigma$-algebra generated by random variables $(E_k, U_k)$. We could equivalently write $\sF_k=\sF(M_k, U_k)$, or $\sF_k=\sF(M_{0:k}, U_k)$.
+
+At each $k$, given that $H_k$ is true, $\L(p_k \mid \sF_{k-1})\sim U[0,1]$, and $p_k$ is $\sF_k$-measurable.
+If $H_k$ is true, then so are $H_{k+1},\ldots, H_d$, and applying this conclusion iteratively we can see that all of the remaining $p$-values are uniform and independent.
+
+By contrast, the saturated model $p$-values use a law different from~\eqref{eq:maxTnull} to compute $p$-values, one that uses
+information not contained in $(E_k,U_k)$.
+As a result, saturated-model $p$-values are generally not independent on nulls.
+We discuss the regression setting in more detail in Section \ref{sec:selective-reg}.
+
+\subsection{General Case}
+
+We now present sufficient conditions for independence on nulls generalizing the above ideas, culminating in Theorem~\ref{thm:suffCond} at the end of this section.
+
+A {\em selective $p$-value} $p_k$ for testing $H_k:\; F \in M_{k-1}(Y)$ satisfies for any $F\in M_{\infty}$,
 \[
@@
 \]
-Valid single-step selective $p$-values are not necessarily independent on nulls, but there is a simple sufficient condition under which independence on nulls is guaranteed.
-
-Consider any filtration
-\[
-\sF_0 \sub \sF_1 \sub \cdots \sub \sF_d
-\]
-for which $M_{0:k}$ is measurable with respect to $\sF_k$. We say that $\sF_{0:d}$ {\em separates} the $p$-values $p_{1:d}$ if 
+
+We say that a filtration $\sF_{0:d}$ {\em separates} the $p$-values $p_{1:d}$ if
 \begin{enumerate}
-\item $p_k(Y)$ is conservative given $\sF_{k-1}$ 
-  when $F\in M_{k-1}(Y)$, and
-\item $p_k(Y)$ is measurable with respect to $\sF_k$.
+\item $M_k(Y)$ and $p_k(Y)$ are measurable with respect to $\sF_k$, and 
+\item $p_k(Y)$ is uniform or conservative given $\sF_{k-1}$ on the event $\{F\in M_{k-1}(Y)\}$.
 \end{enumerate}
-If $p_k$ is exact, the first requirement is equivalent to requiring that $p_k$ is independent of $\sF_{k-1}$ on the event $F\in M_{k-1}(Y)$ (i.e., the event when $p_k$ is null).
-
-If we think of $\sF_k$ as representing information available at step $k$, then the first condition means that $p_k$ ``excludes'' whatever evidence may have accrued against the null by step $k-1$, and the second means that any information revealed after step $k$ plays no role in determining $p_k$.
-In that sense, $\sF_k$ forms a ``wall of separation'' between the $p$-values up to $k$ and the null $p$-values after $k$.
-As a result, separated $p$-values are independent on nulls.
+Separated $p$-values are independent on nulls, as we see next.
 
@@
 
-\subsection{The Sufficient Filtration}\label{sec:suffFilt}
-
-We will concern ourselves mainly with the {\em sufficient filtration} for the path $M_{0:k}$, defined as
-\[
-\sF_k^{\text{suff}} = \sF(M_{0:k}, T_k),
-\]
-\RTcomment{Will- define $M_{0:k-1}$..  it is  the collection $\{ M_1 M_2 \ldots M_{k-1}\}$?} \WFcomment{I've now done so in Section~\ref{sec:notation}}.
-We will suppress the superscript and use $\sF_k$ to mean $\sF_k^{\text{suff}}$ for the remainder of the article.
+If we think of $\sF_k$ as representing information available at step $k$, then the first condition means that $p_k$ ``excludes'' whatever evidence may have accrued against the null by step $k-1$, and the second means that any information revealed after step $k$ plays no role in determining $p_k$.
+
+Define the {\em sufficient filtration} for the path $M_{0:d}$ as the filtration $\sF_{0:d}$ with $\sF_k = \sF(M_{0:k}, U_k)$.
+By our assumption of minimal sufficiency, $U_s$ is always a function of $(M_s, U_k)$ for $s<k$.
+The path algorithm satisfies the SSP if and only if
+\begin{equation}\label{eq:SSP}
+  \sF(M_{0:k}, U_k) = \sF(M_k, U_k), \quad k=0,\ldots, d
+\end{equation}
 
@@
 \begin{enumerate}
-\item $p_k$ is conservative given $M_{0:k-1}$ and $T_{k-1}$, and
-\item $p_k$ is a function of $M_{0:k}$ and $T_k$.
+\item $p_k$ is conservative given $M_{0:k-1}$ and $U_{k-1}$, and
+\item $p_k$ is a function of $M_{0:k}$ and $U_k$.
 \end{enumerate}
@@
 To be a valid selective $p$-value for a single step, $p_k$ only needs to be conservative conditional on $M_{k-1}$.
-In general, it is more stringent to additionally require that $p_k$ must also condition on $T_{k-1}$ as well as the entire subpath $M_{0:k-1}$.
-In most cases of interest, however, the ``extra'' conditioning is redundant, for two reasons.
-
-First, most selective tests already condition on $T_{k-1}$ in order to eliminate nuisance parameters. Second, as we will show in Section~\ref{sec:SSP}, for most path algorithms of interest we have
-\begin{equation}\label{eq:SSP}
-\sF(M_k, \, T_k) = \sF(M_{0:k}, \, T_k).
-\end{equation}
+In general, it is more stringent to additionally require that $p_k$ must condition on the entire subpath $M_{0:k-1}$ as well as $U_{k-1}$.
+In most cases of interest, however, (1) we will condition on $(M_{k-1}, U_{k-1})$ anyway, to eliminate nuisance parameters, and (2) the path satisifies the SSP, so conditioning on $(M_{k-1}, U_{k-1})$ is equivalent to conditioning on $(M_{k-1}, U_{k-1})$.
 Thus, most single-step selective $p$-values we would contemplate using are already exact conditional on $\sF_{k-1}$.
 
-The second requirement, that $p_k$ be $\sF_k$-measurable, has more bite.
-For example, we will see in Section~\ref{sec:singleStep} that it excludes saturated-model tests in linear regression.
-By contrast, any UMPU selective test for $M_{k-1}$ against $M_k\setminus M_{k-1}$ is typically measurable with respect to $\sF_k$.
-
-\subsection{The Subpath Sufficiency Principle}\label{sec:SSP}
-
-We say that a path algorithm $M_{0:d}(\cdot)$ satisfies the {\em subpath sufficiency principle} (SSP) if~\eqref{eq:SSP} holds for every $k$.
-Equation~\eqref{eq:SSP} means that if we know $M_k(Y) = M$, then we can compute the entire subpath of models $M_{0:k}$ knowing only $T(Y, \, M)$, the sufficient statistic of model $M$.
-For example, given that four steps of forward stepwise linear regression yield the active set $E_4 = \{1,5,7,11\}$, the order in which these variables were included can be determined from $T_4 = X_{E_4}'Y$.
-
-For the generic sparse parametric model of Section~\ref{sec:genSparse}, the forward stepwise path satisfies the SSP, as do all ever-active regularization paths. 
+The second requirement, that $p_k$ be $\sF_k$-measurable, has more bite. For example, we will see in Section~\ref{sec:selective-reg} that it excludes saturated-model tests in linear regression.
+
+Collecting together the threads of this section, we arrive at our main result: a general sufficient condition for a $p$-value sequence to be independent on nulls.
+\begin{theorem}[Sufficient Condition for Independence on Nulls] \label{thm:suffCond}
+Assume that $U(Y; M)$ is a complete sufficient statistic for each candidate model $M$, that each $p_k$ is an exact selective $p$-value for $M_{k-1}$, and that the path algorithm $M_{0:d}(\cdot)$ satisfies the SSP.
+
+If $p_k$ is $\sF_k$-measurable for each $k$, then the $p$-value sequence $p_{1:d}(Y)$ is independent on nulls.
+\end{theorem}
+
+\begin{proof}
+We apply the definition of completeness to the function 
+\[
+h(U(Y; \, M)) \;=\; 
+\alpha - \P\bigg(p_k(Y) \leq \alpha \mid U(Y; \, M), \, M_{k-1}(Y) = M\bigg)
+\]
+If $p_k$ is exact given $M_{k-1} = M$, then 
+\[
+\E_F[h(U_{k-1}) \mid M_{k-1}=M] \;=\; 0
+\] 
+for every $F\in M$.
+Therefore, we must have $h(U_{k-1})\eqAS 0$, so $p_k$ is independent of $\sF(M_k, U_{k-1})$.
+Because $M_{0:d}$ satisfies the SSP, $p_k$ is also independent of $\sF_{k-1} = \sF(M_{0:k-1}, U_{k-1})$.
+
+If $p_k$ is also $\sF_k$-measurable, then the sufficient filtration separates $p_{1:d}$, implying that the sequence $p_{1:d}$ is independent on nulls.
+\end{proof}
+
+Note that if our path algorithm did not satisfy the SSP, we could in many cases still repair the situation by constructing $p$-values that are uniform conditional on $(M_{0:k-1},U_{k-1})$, not just conditional on $(M_{k-1},U_{k-1})$.
+
+\subsection{Independence for Greedy Likelihood and next-$\lambda$ $p$-values}\label{sec:SSP}
+
+In this section we establish that in the generic sparse parametric model of Section~\ref{sec:genSparse}, the forward stepwise path and all ever-active regularization paths satisfy the SSP.
+What is more, the greedy likelihood ratio statistic $G_k$ is $\sF_k$-measurable with respect to the forward stepwise path's sufficient filtration, and the next-$\lambda$ statistic $\lambda_k$ is $\sF_k$-measurable with respect to the regularized-likelihood path's sufficient filtration.
+As a result, the selective greedy likelihood $p$-values in the forward stepwise path are independent on nulls, as are the next-$\lambda$ $p$-values in the regularized-likelihood path, per Theorem~\ref{thm:suffCond}.
+
 Define the restricted log-likelihood
@@
 
-\begin{proposition}[Forward Stepwise Satisfies SSP]\label{prop:forwardSSP}
-  The forward stepwise path as defined in
-  (\ref{eq:forwardDef_start}--\ref{eq:forwardDef_end}) satisfies the 
-  sub-path sufficiency principle.
+\begin{proposition}\label{prop:forwardSSP}
+  The forward stepwise path satisfies the subpath sufficiency principle.
 \end{proposition}
 \begin{proof}
-  For some fixed step $k$ and active set $E$, 
-  let $A$ denote the event $\{M_k = M(E)\}$. 
-  On $A$, the restricted likelihood $\ell_E$ is proportional to a
-  function depending only on $T_k = T(Y; \;M(E))$:
-  \[
-  \ell_E(\theta; Y) = \ell_E^{T_k}(\theta; T_k) 
-  + \ell_E^{Y \mid T_k}(Y \mid T_k).
-  \]
-  The second term does not depend on $\theta$ 
-  because $T(Y; \;M(E))$ is sufficient for $M(E)$.
-
-  On $A$, we have $E_i(Y) \sub E$ for all $i=1,\ldots, k$, meaning   that the maximum in \eqref{eq:forwardDef_start} was attained by some $j\in E$ at every step. So, for $i \leq k$, we have
+  Recall the forward stepwise path is defined by
+  \begin{equation}\label{eq:forwardStepDef}
+    j_k = \argmax_j \;\;\sup \left\{\ell(\theta; Y):\; \theta\in\Theta(E_{k-1} \cup \{j\})\right\}, \quad \text{ and } \quad
+    E_k = E_{k-1} \cup \{j_k\}
+  \end{equation}
+
+  For some fixed step $k$ and active set $E$, let $A$ denote the event $\{M_k = M(E)\}$.
+Conditioning on $U$, the restricted likelihood is proportional to a function depending only on $U = U(Y; \;M(E))$.
+The log-likelihood decomposes as
+  \[
+  \ell_E(\theta; Y) = \ell_E^{U}(\theta; U) 
+  + \ell_E^{Y \mid U}(Y \mid U).
+  \]
+  The second term, the log-likelihood of $Y$ given $U$, does not depend on $\theta$ because $U$ is sufficient for $M(E)$.
+
+  On $A$, we have $E_s(Y) \sub E$ for all $s\leq k$, meaning that the maximum in~\eqref{eq:forwardStepDef} is attained by some $j\in E$ at every step. So, for $i \leq k$, we have
 \begin{align}
-  j_i &= \argmax_{j\in E} \;\;\sup \left\{\ell_E(\theta; Y):\;
-    \theta\in\Theta(E_{i-1} \cup \{j\})\right\} \\
-  &= \argmax_{j\in E} \;\;\sup \left\{\ell_E^T(\theta; T(Y)):\;
-    \theta\in\Theta(E_{i-1} \cup \{j\})\right\}. \label{eq:onlyT}
+  j_s &= \argmax_{j\in E} \;\;\sup \left\{\ell_E(\theta; Y):\;
+    \theta\in\Theta(E_{s-1} \cup \{j\})\right\} \\
+  &= \argmax_{j\in E} \;\;\sup \left\{\ell_E^U(\theta; U(Y)):\;
+    \theta\in\Theta(E_{s-1} \cup \{j\})\right\}. \label{eq:onlyU}
 \end{align}
 
-Equation~\eqref{eq:onlyT} shows that $j_1,\ldots, j_{k-1}$ all depend on $Y$ only through $T_k(Y)$. As a result, it also follows that the enitre sequence $M_{0:k}(Y)$ depends only on $T_k(Y)$.
+Equation~\eqref{eq:onlyU} shows that $j_1,\ldots, j_{k-1}$ all depend on $Y$ only through $U(Y)$, which equals $U_k(Y)$ on $A$.
+As a result, it also follows that the entire sequence $M_{0:k}(Y)$ depends only on $(M_k, U_k)$.
 \end{proof}
 
-\begin{proposition}[Regularized Likelihood Paths Satisfy SSP]\label{prop:regPathSSP}
-Ever-active regularized likelihood paths as defined in 
-(\ref{eq:regPathDef_start}--\ref{eq:regPathDef_end}) satisfy the
-sub-path sufficiency principle.
+\begin{corollary}\label{cor:greedyLikIndep}
+  If $U(Y; M)$ is complete sufficient for model $M$, the selective greedy likelihood $p$-values computed for the forward stepwise path are independent on nulls.
+\end{corollary}
+\begin{proof}
+  Per Theorem~\ref{thm:suffCond}, it is enough to show that $p_k$ is $\sF_k$-measurable with respect to the forward stepwise path's sufficient filtration.
+First, the test statistic $G_k$ is $\sF_k$-measurable because
+  \begin{align*}
+    G_k(Y) &= \sup_{\theta\in \Theta(E_k)} {\ell(\theta; Y)} \;\;- \sup_{\theta\in \Theta(E_{k-1})} {\ell(\theta; Y)}\\
+    &= \sup_{\theta\in \Theta(E_k)} {\ell_{E_k}^{Y\mid U}(\theta; U_k)} \;\;- \sup_{\theta\in \Theta(E_{k-1})} {\ell_{E_k}^{Y\mid U}(\theta; U_k)}.
+  \end{align*}
+  Second, the null distribution $\L\left(G_k \,\mid\, M_{k-1}, U_{k-1}\right)$ against which we compare $G_k$ is also $\sF_k$-measurable because it depends only on $(M_{k-1}, U_{k-1})$, both of which are $\sF_{k-1}$-measurable.
+\end{proof}
+
+Analogous results hold for ever-active regularized likelihood paths and the next-$\lambda$ $p$-values.
+
+\begin{proposition}\label{prop:regPathSSP}
+Ever-active regularized likelihood paths satisfy the
+subpath sufficiency principle.
 \end{proposition}
@@
 \begin{proof}
-  Again, for fixed $k$ and $E$ let $A$ denote the event $\{M_k = M(E)\}$ for some fixed $E$, and define $T_k$ as in the proof of Proposition~\ref{prop:forwardSSP}. If we define
-\[
-\hat\theta^{(E,r)} = \argmin_{\theta\in\Theta(E)} -\ell_E^{T_k}(\theta; T(Y)) + P_r(\theta),
+  Again, for fixed $k$ and $E$ let $A$ denote the event $\{M_k = M(E)\}$, and define $U=U(Y;\, M(E))$. If we define
+\[
+\hat\theta^{(E,\lambda)} = \argmin_{\theta\in\Theta(E)} -\ell_E^{U}(\theta; U(Y)) + P_\lambda(\theta),
 \]
@@
 \[
-\hat\theta^{(E,r)} \eqAS \hat\theta^{r} \text{ on } 
-A \cap 1_{\{r \leq R_k\}}.
-\]
-But $\hat\theta^{(E,r)}$ depends only on $T_k(Y)$. Therefore, on $A$, we can reconstruct the entire path of solutions once we know $T_k(Y)$.
+\hat\theta^{(E,\lambda)} \eqAS \hat\theta^{\lambda} \text{ on } 
+A \cap 1_{\{\lambda \geq \lambda_k\}}.
+\]
+But $\hat\theta^{(E,\lambda)}$ depends only on $U(Y)$.
+Therefore, on $A$, we can reconstruct the entire path of solutions for $\{\lambda\in \Lambda:\; \lambda\geq \lambda_k\}$, once we know $(M_k,\,U_k)$.
 \end{proof}
 
-Collecting together the threads of this section, we can arrive at our main result: a fairly general sufficient condition for a $p$-value sequence to be independent on nulls.
-\begin{theorem}[Sufficient Condition for Independence on Nulls] \label{thm:suffCond}
-Assume that $T(Y; M)$ is a complete sufficient statistic for each candidate model $M$, that $p_k$ are exact selective $p$-values for $M_{k-1}$, and that $M_{0:d}(\cdot)$ satisfies the SSP.
-
-If $p_k$ is $\sF_k$-measurable for each $k$, then the $p$-value sequence $p_{1:d}(Y)$ is independent on nulls.
-\end{theorem}
-
+\begin{corollary}\label{cor:nextLambdaIndep}
+  If $U(Y; M)$ is complete sufficient for model $M$, the selective next-$\lambda$ $p$-values computed for any ever-active regularized likelihood path are independent on nulls.
+\end{corollary}
 \begin{proof}
-We apply the definition of completeness to the function 
-\[
-h(T(Y; \, M)) \;=\; 
-\P(p_k(Y) \leq \alpha \mid T(Y; \, M), \, M_{k-1}(Y) = M) - \alpha
-\]
-If $p_k$ is exact given $M_{k-1} = M$, then 
-\[
-\E_F[h(T_{k-1}) \mid M_{k-1}=M] \;=\; 0
-\] 
-for every $F\in M$.
-Therefore, we must have $h(T_{k-1})\eqAS 0$, so $p_k$ is independent of $\sF(M_k, T_{k-1})$.
-Because $M_{0:d}$ satisfies the SSP, $p_k$ is also independent of $\sF_{k-1} = \sF(M_{0:k-1}, T_{k-1})$.
-
-If $p_k$ is also $\sF_k$-measurable, then the sufficient filtration separates $p_{1:d}$, implying that $p_{1:d}$ is independent on nulls.
+  Per Theorem~\ref{thm:suffCond}, it is enough to show that $p_k$ is $\sF_k$-measurable with respect to the regularized-likelihood path's sufficient filtration.
+
+  Recall that 
+  \[
+  \Lambda_k = \{\lambda\in \Lambda:\; \tE_\lambda \supsetneq \tE_{\lambda_{k-1}}\},
+  \quad 
+  \lambda_k = \sup\Lambda_k,
+  \quad \text{ and } \quad
+  E_k = \bigcap_{\lambda\in\Lambda_k} \tE_{\lambda},
+  \]
+  and that the $\tE_{\lambda}$ are nested by construction and finite by assumption. As a result, there is some $\nu_k\in \Lambda_k$ for which $\tE_{\nu_k} = E_k$.
+
+  Furthermore, for all $\lambda\in\Lambda, \lambda\geq \nu_k$, we have $\tE_\lambda \sub E_k$.
+As a result, ${\hat\theta^{\lambda} =\hat\theta^{(E_k,\lambda)}}$ for such $\lambda$, so $\tE_{\lambda}$ depends only on $\ell_{E_k}^{Y\mid U}(Y;\, U_k)$, and therefore $\lambda_k$ can be computed from $(E_k, U_k)$.
+
+  Second, the null distribution $\L\left(\lambda_k \,\mid\, M_{k-1}, U_{k-1}\right)$ against which we compare $\lambda_k$ is also $\sF_k$-measurable because it depends only on $(M_{k-1}, U_{k-1})$, both of which are $\sF_{k-1}$-measurable.
 \end{proof}
 
-\section{Inference for one step in the path}
-\label{sec:singleStep}
-
-Having derived conditions for independence, we turn now to the question of how to construct valid selective $p$-values for a single step in the path. At step $k$, we construct $p_k(Y)$ to test
-\[
-  H_{k}:\; F \in M_{k-1}(Y)
-\]
-against the alternative $F \in M_\infty \setminus M_{k-1}(Y)$.
-As usual, the main complication arises from the fact that the null hypothesis is random.
-To simplify matters, we first consider deriving $p$-values for a particular fixed candidate model $M\sub M_{\infty}$.
-
-The random variable  $p_{k,M}(Y)$ is a valid {\em selective $p$-value} for $M$ at step $k$ if it is conservative under sampling from $F\in M$, given that $M$ is selected. That is,
-\[
-\P_F\left(p_{k,M}(Y) \leq \alpha \mid M_{k-1}(Y) = M\right) 
-\leq \alpha, \quad \forall F\in M, \alpha \in [0,1].
-\]
-
-Once we have constructed selective $p$-values for each fixed candidate $M$, we can use them as building blocks to construct a combined $p$-value for the random null $M_{k-1}(Y)$. Define
-\[
-p_k(y) = p_{k, M_{k-1}(y)}(y),
-\]
-which is a valid $p$-value on the event $\{F \in M_{k-1}(Y)\}$:
-\begin{equation}\label{eq:selectiveGuaranteePk}
-\P_F\left(p_k \leq \alpha \mid M_{k-1}, 
-  \;F\in M_{k-1}\right) \leq \alpha, \quad \forall \alpha \in [0,1].
-\end{equation}
-%Recall that in~\eqref{eq:selectiveGuaranteePk} $M_{k-1}$ is random and not $F$.
-
-One very useful strategy for constructing exact selective tests is to condition on the sufficient statistics of the null model. Because the law
-\[
-\L(Y \;\mid\; T_{k-1}, \; M_{k-1}=M)
-\]
-is known under any $F\in M$, we can construct selective tests by comparing the distribution of any test statistic $U(Y)$ to its known conditional distribution under the null.
-See \citet{fithian2014optimal} for a general treatment of selective tests in exponential family models.
-
-If we follow this strategy, then the reference distribution for $U(Y)$ is measurable with respect to $T_{k-1}$ and $M_{k-1}$; moreover, if $U(Y)$ is also measurable with respect to $(M_k, T_k)$, the next model in the path and its sufficient statistics, then $p_k$ is measurable with respect to $\sF_k$, and so the $p$-value sequence $p_{1:d}$ will be independent on nulls per Theorem~\ref{thm:suffCond}.
-
-For the remainder of this section we will focus on linear regression.
-
-\subsection{Selective $p$-Values in Regression}
+As we have shown, the selective max-$t$ test is equivalent to the selective greedy likelihood test in the case of linear regression.
+In the next section, we will compare and contrast the max-$t$ and next-$\lambda$, and other {\em selected-model tests}, with the {\em saturated-model tests} proposed by \citet{taylor2014exact} and others.
+
+\section{Selective $p$-Values in Regression}
 \label{sec:selective-reg}
-In linear regression, there are several types of selective test that we might contemplate performing at step $k$.
-Let $E_{k-1}$ and $E_k$ denote the active sets for models $M_{k-1}$ and $M_k$ respectively.
-For simplicity we will assume that only one variable is added at step $k$, so that $E_k = E_{k-1} \cup j_k$.
-It is relatively straightforward to generalize the developments in this section to the case where multiple variables are added at once.
-
-\subsubsection{Saturated-Model Tests}
-Many of the selective tests for linear regression use the saturated-model framework.
+In linear regression, \citet{fithian2014optimal} draw a distinction between two main types of selective test that we might perform at step $k$: tests in the selected model, and tests in the saturated model.
+For simplicity, we will assume throughout this section that our path algorithm adds exactly one variable at each step.
+We also assume the path algorithm satisfies the SSP, so that we need not worry about the distinction between conditioning on $(M_{k-1}, U_{k-1})$ and conditioning on $(M_{0:k-1}, U_{k-1})$.
+
+\subsection{Saturated-Model Tests}
+Many conditional selective tests for linear regression use the saturated-model framework.
 See for example \citet{lockhart2014significance}, \citet{taylor2013tests}, \citet{taylor2014exact}, \citet{lee2013exact}, and \citet{loftus2014significance}.
-
-The distinction between the incremental null and the complete null is closely related to the distinction between {\em saturated-model} tests and {\em selected-model} tests.
 Define the saturated model as
@@
 \]
-so named because there is a mean parameter for every observation. Because the parameter $\mu$ has the same dimension as the data $Y$, we must assume that $\sigma^2$ is known.
-
-Saturated-model tests assume only that $M_{\text{sat}}$ is true, and perform inference on linear functionals of $\mu$.
-For example, in linear regression after selecting active set $E$, we might construct confidence intervals for the least-squares coefficients $\theta^E$ of active set $E$.
-
-In the context of sequential model selection, we can use the saturated-model framework to test the incremental null $H^{\text{inc}}:\; \theta_{j_k}^{E_k} = 0$.
-The saturated model has $n$-dimensional sufficient statistic $Y$ and natural parameter $\mu$.
-Letting $\eta_k\in \R^n$ denote the vector for which $\theta_{j_k}^{E_k} = \eta_k'\mu$, the UMPU saturated-model test for $\eta_k'\mu$ is based on the distribution
+so named because there is a mean parameter for every observation.
+Because the parameter $\mu$ has the same dimension as the data $Y$, we must assume that $\sigma^2$ is known.
+If this is not the case, we can estimate $\sigma^2$ and apply the test while plugging in the estimated value; this is what we have done with the diabetes data to compute the $p$-values in Table~\ref{tab:diab}.
+
+Saturated-model tests assume only that $M_{\text{sat}}$ is true, and perform inference on selected linear functionals of $\mu$.
+For example, in linear regression after selecting active set $E$, we could construct confidence intervals for the least-squares coefficients $\theta^E$ of active set $E$, defined as
+\[
+\theta^E \;\;=\;\; \argmin_{\theta\in\R^{|E|}} \|\mu - X_E\theta\| \;\;=\;\; (X_E'X_E)^{-1}X_E'\mu
+\]
+
+In the context of sequential model selection, we can use the saturated-model framework to test the incremental null 
+\[
+H^{\text{inc}}:\; \theta_{j_k}^{E_{k-1}\cup \{j_k\}} = 0.
+\]
+Let $\eta_k\in \R^n$ denote the vector for which $\theta_{j_k}^{E_k} = \eta_k'\mu$, and let $\proj_{\eta_k}^\perp$ denote the projection operator into the space orthogonal to $\eta_k$.
+The saturated model has $n$-dimensional natural parameter $\mu$ and complete sufficient statistic $Y$, and the UMPU saturated-model test for $\eta_k'\mu$ is based on the distribution
 \begin{equation}\label{eq:satModel}
@@
 
-While the test statistic $\eta_k'Y$ is measurable with respect to $\sF_k$, the reference distribution we compare it against is not.
-Specifically, $\proj_{\eta_k}^\perp Y$ is not measurable with respect to $\sF_k$.
+While the test statistic $X_{j_k}'Y$ is measurable with respect to $\sF_k$, the null distribution we compare it against is not: it depends on $\proj_{\eta_k}^\perp Y$, which is not measurable with respect to $\sF_k$.
 As a result, saturated-model $p$-values constructed as in~\eqref{eq:satModel} are in general {\em not} independent on nulls.
-?Collect
-\subsubsection{Selected-Model Tests}
+
+\subsection{Selected-Model Tests}
 Selected-model inference~\citep{fithian2014optimal} 
 represents a more powerful option when testing for inclusion of variable $j_k$.
-While saturated-model inference always uses the same model $M_{\text{sat}}$, selected-model inference is based on the statistical model that is chosen by our selection procedure.
-In the sequential case this means that at step $k$, we test the selected null model $M(E_{k-1})$, against either the next selected model $M(E_k)$ or the upper model $M_\infty$.
-
-\paragraph{Conditioning on the Next Model}
-One option for selected-model inference is to condition on the identity of $j_k$, the next variable added, and perform a UMPU, equal-tailed, or one-sided selective test of $M_{k-1}$ against $M_{k}\setminus M_{k-1}$ conditional on both models.
+While saturated-model inference always uses the same model $M_{\text{sat}}$, selected-model inference uses the lower-dimensional statistical model that is chosen by our selection procedure, similarly to data splitting.
+In the sequential case this means that at step $k$, we test the selected null model $M(E_{k-1})$, either against the next selected model $M(E_k)\setminus M(E_{k-1})$ or against the upper model $M_\infty\setminus M(E_{k-1})$.
+
+\subsubsection{Testing Against the Next Model}\label{sec:identify}
+One option for selected-model inference is to perform a selective test of $M_{k-1}$ against $M_{k}\setminus M_{k-1}$, conditioning on both models.
+Conditioning on $(M_{k-1},M_k)$ is equivalent to conditioning on $(E_{k-1}, j_k)$.
 If $\sigma^2$ is known, the test is based on
@@
 \end{equation}
+If $\sigma^2$ were unknown we would also need to condition on $\|Y\|^2$.
+
+We can construct $p$-values from the UMPU, equal-tailed, or one-sided test based on~\eqref{eq:selModel_cond}; the only difference is in how much of the overall level $\alpha$ is apportioned to the right or left tail.
+
+If we use the construction in~\eqref{eq:selModel_cond} after a subpath-sufficient path algorithm, the resulting $p$-values are always guaranteed to have independence on nulls.
+This is because the test statistic $X_{j_k}'Y$ is measurable with respect to $\sF_k$, and so is the null distribution we compare it against, which depends on $E_{k-1}, j_k,$ and $U_{k-1} = X_{E_{k-1}}'Y$.
+
 Note the contrast between~\eqref{eq:selModel_cond} and~\eqref{eq:satModel}.
@@
 
-If we use the construction in~\eqref{eq:selModel_cond} after a subpath-sufficient path algorithm, the $p$-values are always guaranteed to have independence on nulls.
-The test statistic $X_{j_k}'Y$ is measurable with respect to $\sF_k$, and so is the reference distribution, which depends on $E_{k-1}, j_k,$ and $T_{k-1} = X_{E_{k-1}}'Y$.
-
-\paragraph{The Max-$t$ and Next-$\lambda$ Tests}
-Instead of conditioning on the next model $M_k$, we could use
+\subsubsection{Testing Against the Upper Model}
+To avoid conditioning on the next model $M_k$, we could instead test $M_{k-1}$ against the alternative $M_{\infty}\setminus M_{k-1}$.
+In that case, we could modify~\eqref{eq:selModel_cond} and base the test on
 \begin{equation}\label{eq:selModel_marg}
-\L\left(X_{j_k}'Y \mid E_{k-1}, \; X_{E_{k-1}}'Y\right),
-\end{equation}
-or even replace $X_{j_k}'Y$ with any other test statistic $U(Y)$.
-As long as $U(Y)$ is measurable with respect to $(M_k, T_k)$, and the path algorithm is subpath-sufficient, the $p$-values are independent on nulls.
-
+\L\left(X_{j_k}'Y \mid E_{k-1}, \; X_{E_{k-1}}'Y\right).
+\end{equation}
+We could also replace $X_{j_k}'Y$ with any other test statistic $T(Y)$.
+As long as $T(Y)$ is measurable with respect to $(M_k, U_k)$, and the path algorithm is subpath-sufficient, the $p$-values are guaranteed to be independent on nulls.
+The next-$\lambda$ test for the lasso and the max-$t$ test in forward-stepwise regression with unknown $\sigma^2$ are both examples of this approach.
+Note that under resampling from the law in~\eqref{eq:selModel_marg}, the selected variable $j_k$ is random; by contrast, \eqref{eq:selModel_cond} conditions on $j_k$.
+
+In the case of known $\sigma^2$, the greedy likelihood ratio statistic is the max-$z$ statistic: the maximum of $z_{j,E_{k-1}}$, the largest multivariate $z$-statistic of any held-out variable, controlling for the variables in the current model.
+
+For the most part we do not pursue tests like those in Section~\ref{sec:identify}, which condition on the next model $M_k$.
+It is not necessary to condition on $M_k$ to obtain the FDR and FWER guarantees that we want to establish, and so we follow the rule of thumb that we should condition on as little information as possible.
+In Section~\ref{sec:sparseReg}, we empirically compare the max-$t$ test with the test that uses the same test statistic but also conditions on the identity of the next variable $j_k$.
+We call the resulting test the {\em max-$t$-identify test}.
+Its performance in our simulation is similar to that of the max-$t$ test.
+
+\begin{comment}
 Two specific examples are worth discussing in detail. First, consider the {\em selective max-$t$ test}, a generalization of the max-$t$ test described in 
@@
 The next-$\lambda$ statistic is $\sF_k$-measurable if we use the LASSO to select variables. \WFcomment{Relate this to spacings test. Is this essentially the ``generalized spacings test?''}
-
-Table \ref{tab:testsummary} summarizes a number of different tests that have been proposed for least angle regression, lasso and forward stepwise regression.
+\end{comment}
+
+\subsubsection{Classification of Proposed Tests}
+Table \ref{tab:testsummary} classifies a number of different tests that have been proposed for least angle regression, lasso and forward stepwise regression.
 Some focus on the complete null hypothesis, others on the incremental null (as defined in Section \ref{sec:whichnull}).
-Some use the saturated model, others the selective model framework, as detailed in earlier in the section.
-The saturated model tests are simpler and do not require sampling, while the selective tests are generally more powerful.
-In the next section we compare the  selective and saturated model settings in more detail.
+Some use the saturated-model, and others the selected-model framework, as detailed in earlier in the section.
+The saturated-model tests are simpler and do not require sampling, while the selected-model tests are generally more powerful.
+In the next section we compare the selected- and saturated-model settings in more detail.
 
@@
 LAR/lasso& Covariance test/Saturated&Spacing test/Saturated \\
-        & Next-$\lambda$ test/Selective& \\
+        & next-$\lambda$ test/Selected& \\
         \hline
-Forward Stepwise &maxT, maxT-identify/Selective &Gaussian pivot/Saturated\\
+Forward Stepwise &max-$t$, max-$t$-identify/Selected &Gaussian pivot/Saturated\\
 \hline
@@
 \caption{\em Summary of different tests that have been proposed for least angle regression, lasso and forward stepwise regression.
-Tests are described by the null hypothesis that they test (complete or incremental)  and the underlying model (saturated or selective).
-The maxT, maxT-identify and next-$\lambda$ tests are proposed in this paper, while the covariance test was proposed in \citet{lockhart2014significance}, and the spacing and Gaussian
-pivot tests are given in \citet{spacings}. The latter paper shows that the covariance and spacing test are asymptotically equivalent for the first null step, when the complete null holds,}
+Tests are described by the null hypothesis that they test (complete or incremental)  and the underlying model (saturated or selected).
+The max-$t$, max-$t$-identify and next-$\lambda$ tests are proposed in this paper, while the covariance test was proposed in \citet{lockhart2014significance}, and the spacing and Gaussian
+pivot tests are given in \citet{taylor2014exact}. The latter paper shows that the covariance and spacing test are asymptotically equivalent for the first null step, when the complete null holds.}
 \label{tab:testsummary}
@@
 Figure  \ref{fig:larexample} shows an example of the LAR/lasso tests with $N=50, p=10$ and two strong signals.
-We plot the p-values only for those realizations in which the correct model was selected in the first two steps (about 95\% of the time).
-We see that the next-$\lambda$ test is the most powerful , especially at step 1. The next-$\lambda$ and spacing test have the correct type I error in steps 3 and 4,
+We plot the $p$-values only for those realizations in which the correct model was selected in the first two steps (about 95\% of the time).
+We see that the next-$\lambda$ test is the most powerful, especially at step 1. The next-$\lambda$ and spacing test have the correct type I error in steps 3 and 4,
 while the covariance test becomes conservative in step 4, as we move further into the null part of the sequence.
@@
   \includegraphics[width=\textwidth]{figs/larexample.pdf}
-  \caption[fig:larexample]{\em  Cumulative distribution functions of p-values for LAR/lasso tests from simulations with $N=50,  p=10$ and two strong signals.}
+  \caption[fig:larexample]{\em  Cumulative distribution functions of $p$-values for LAR/lasso tests from simulations with $N=50,  p=10$ and two strong signals.}
   \label{fig:larexample}
@@
 \subsection{Comparison of Selected- and Saturated-Model Inference}\label{sec:bivariate}
+
+We now briefly compare the computational, conceptual, and statistical advantages and disadvantages of selected- and saturated-model tests, focusing on the problem of sequential model selection.
+For a detailed general discussion, see \citet[][Section 5]{fithian2014optimal}.
 
@@
 For example, a 95\% confidence interval for $\theta_{j_k}^{E_k}$ could cover that coefficient with probability 95\% even if $M(E_k)$ is misspecified.
-In the setting of sequential model selection, this advantage is less important.
-We do not necessarily need a scientifically interpretable interval for $\theta_{j_k}^{E_k}$; rather, the goal of the test at step $k$ is to reject $M_{k-1}$ and move on if it is demonstrably incompatible with the data.
-
-There are several major benefits to selected-model inference.
-First, selected-model inference allows us to drop the assumption that $\sigma^2$ is known by conditioning on $\|Y\|^2$, the sufficient statistic corresponding to the natural parameter $1/2\sigma^2$.
+In the setting of sequential model selection, this advantage is less important: we are not looking for a scientifically interpretable interval for $\theta_{j_k}^{E_k}$; rather, the goal of the test at step $k$ is to reject $M_{k-1}$ and move on if it is demonstrably incompatible with the data.
+If there are many remaining strong signal variables to be found, then both $M_k$ and $M_{k-1}$ are badly misspecified; the more true this is, the stronger the grounds are for rejecting $M_{k-1}$.
+
+There are several major statistical benefits to selected-model inference.
+First, selected-model inference allows us to drop the assumption that $\sigma^2$ is known.
 This is not an option for the saturated model because conditioning on both $\|Y\|^2$ and $\proj_{\eta_k}^\perp Y$ results in a degenerate conditional distribution for $\eta_k'Y$.
-See \citet[][Section 5]{fithian2014optimal} for a detailed discussion of the distinctions between saturated-model and selected-model tests.
 
@@
 \[
-M_0:\; Y\sim \cN(0,I_2), \quad\text{ and } 
+M_0:\; Y\sim \cN(0,I_2), \quad \text{ and } \quad
 M_1:\; Y \sim \cN\left(\binom{\mu_1}{0}, \; I_2\right).
@@
 
-The selected-model test at step 1 compares $Y_1$ to its distribution under $M_0$ conditional on $A_1$, a test of $H_0:\;\mu_1=0$ in model $M_1$.
-By contrast, the saturated-model test is a test of $H_0:\; \mu_1=0$ in the model $M_{\text{sat}}:\; Y \sim \cN(\mu, I_2)$.
-The saturated model test must condition on $Y_2$ to eliminate the nuisance parameter $\mu_2$, and compare $Y_1$ to its null distribution given $A_1$ {\em and} the observed value of $Y_2$.
+The selected-model test at step 1 compares $Y_1$ to its distribution under $M_0$ conditional on $A_1$, a test of $H_0:\;\mu_1=0$ in model $M_1$.\footnote{In this very simple example, the max-$z$, max-$z$-identify, and next-$\lambda$ tests all coincide.} By contrast, the saturated-model test is a test of $H_0:\; \mu_1=0$ in the model $M_{\text{sat}}:\; Y \sim \cN(\mu, I_2)$.
+Thus, it must condition on $Y_2$ to eliminate the nuisance parameter $\mu_2$, comparing $Y_1$ to its null distribution given $A_1$ and the observed value of $Y_2$.
 
 Figure~\ref{fig:bv_condSets} shows the conditioning sets for each model when $Y=(2.9, 2.5)$.
-Next to it, Figure~\ref{fig:bv_nullDists} shows the null distribution for the test statistic $Y_1$ in each case.
+Beside it, Figure~\ref{fig:bv_nullDists} shows the null conditional distribution of the test statistic $Y_1$ for each test.
 The $p$-values for the selected and saturated models are 0.007 and 0.3, respectively.
@@
     \includegraphics[width=\textwidth]{figs/bivariateSelVSat_condSets.pdf}
-    \caption{The selected-model test conditions on $j_1=1$ (yellow region), while the saturated-model test also conditions on $Y_2=2.5$ to eliminate the nuisance variable $\mu_2$ (brown region).}
+    \caption{The selected-model test conditions on $j_1=1$ (yellow region), while the saturated-model test also conditions on $Y_2=2.5$ to eliminate the nuisance variable $\mu_2$ (brown line segments).}
     \label{fig:bv_condSets}
@@
       also condition on $Y_2=2.5$.}
+  \label{fig:bv_nullDists}
   \end{subfigure}
@@
     tests in Example~\ref{ex:bivariate}. The selected-model test is based on  $\L(Y_1 \gv j_1=1)$,  whereas the saturated-model test is based on $\L(Y_1  \gv Y_2, \; j_1=1)$. 
-    When $Y=(2.9, 2.5)$, the selected- and saturated-model $p$-values are 0.015 and 0.3, respectively.}
-  \label{fig:bv_nullDists}
+    When $Y=(2.9, 2.5)$, the selected- and saturated-model $p$-values are 0.007 and 0.3, respectively.}
+  \label{fig:bv}
 \end{figure}
 
-Figure~\ref{fig:bv_condSets} illustrates a certain phenomenon occurring in saturated-model tests: when there are near-ties between strong variables that are competing to enter the model, the resulting $p$-value my be very weak~\citep{lockhart2014significance}.
+Figure~\ref{fig:bv} illustrates a general phenomenon in saturated-model tests: when there are near-ties between strong variables that are competing to enter the model, the resulting $p$-value my be very weak~\citep{lockhart2014significance}.
 Figure~\ref{fig:bv_rocCurve} displays the cumulative distribution function for the first $p$-value when $\mu=(4,4)$, a very strong signal.
-While the selected model test has near perfect power, it is not uncommon for the saturated model test to produce large $p$-values, even in the range of 0.5-0.9.
+While the selected-model test has nearly perfect power, it is not uncommon for the saturated-model test to produce large $p$-values, even in the range of 0.5-0.9.
 These large $p$-values arise when there is a near tie between the variables.
@@
 That is, the model $M_0 = M(\emptyset)$ is missing an important signal variable, but the missing variable is not $X_1 = (1,0)$.
-Because the saturated-model test is really a test of the incremental null, its power is $\alpha=0.05$.
-By contrast, the selected-model test rejects about half the time when $\mu=(0,4)$, successfully detecting the fact that $M_0$ is a poor fit to the data.
+Because the saturated-model test is a valid test of the incremental null, its power is $\alpha=0.05$.
+By contrast, the selected-model test rejects about half of the time when $\mu=(0,4)$, detecting the fact that $M_0$ is a poor fit to the data.
 
@@
 Continuing with Example~\ref{ex:bivariate}, Table~\ref{tab:bv_twoWayTable} shows a two-way contingency table for the saturated-model $p$-values $(p_1(Y), p_2(Y))$, binned into cells of height and width 0.2, simulated under the global null $\mu=0$.
-Because $k_0=0$, both $p$-values are uniform by construction, but the $p$-values are strongly dependent, with correlation $-48\%$.
+Because $k_0=0$, both $p$-values are uniform by construction, but the $p$-values are strongly dependent, with correlation $-0.48$.
 By contrast, the selected-model $p$-values $(p_1,p_2)$ are independent under the global null, a consequence of Theorem~\ref{thm:suffCond}.
@@
 We report the percentage of $p$-value pairs falling into each cell out of one million simulations from the global null hypothesis, $\mu=0$.
-Both $p$-values are marginally uniform but strongly dependent, with a correlation of $-48\%$.}
+Both $p$-values are marginally uniform but strongly dependent, with a correlation of $-0.48$.}
 \label{tab:bv_twoWayTable}
@@
 
-As usual, we have reduced computation of the $p$-value at step $k$ to a sampling problem.
-Whatever our test statistic at step $k$, we must compute its distribution under the null model $M_{k-1}$.
-We give details of our  strategy here for the max-t test in forward stepwise regression.
+Until this point, we have deferred discussing the question of how to actually compute the rejection cutoffs for the max-$t$ test, next-$\lambda$ test, or any of the other tests we have discussed.
+In each case, we have a test that rejects when some test statistic $T_k$ is large compared to the law
+\begin{equation}\label{eq:genDist}
+\L(T_k \,\mid\, M_{k-1}, U_{k-1}),
+\end{equation}
+under the null hypothesis that $F\in M_{k-1}$.
+As such, we have reduced computation of the $p$-value at step $k$ to a sampling problem: if we can sample values of $Y$ from the law~\eqref{eq:genDist}, we can numerically approximate the $p$-value to whatever precision we desire.
+
+Due to the generality in which we have posed the problem here, we cannot give a general prescription for how to carry out this sampling.
+Instead, we give details of our computational strategy here for one case, the max-$z$ test in forward stepwise regression with known $\sigma^2$.
 
@@
 \begin{eqnarray}
-L(X'_jY | E_{k-1},X'_{E_{k-1}}Y)
+L(X'_jY \,|\, E_{k-1},X'_{E_{k-1}}Y)
 \label{eqn:maxtsamp}
@@
 removes the dependence on nuisance parameters.
-Hence  (\ref{eq:boot}) yields the exact (non-asymptotic) conditional distributions and p-values.
-
-Figure \ref{fig:comparison} shows an example with $n=100, p=20$ and one strong signal, computing the max-t pvalue at Step 3.
+Hence  (\ref{eq:boot}) yields the exact (non-asymptotic) conditional distributions and $p$-values.
+
+Figure \ref{fig:comparison} shows an example with $n=100, p=20$ and one strong signal, computing the max-$t$ $p$-value at Step 3.
 The ``exact'' answer  was determined from bootstrap sampling with $20,000$ replications.
@@
   \includegraphics[width=\textwidth]{figs/comparison.pdf}
-  \caption{\em Simulated example with $n= 100, p=20$ and one strong signal, computing the max-t pvalue at Step 3.
+  \caption{\em Simulated example with $n= 100, p=20$ and one strong signal, computing the max-$t$ $p$-value at Step 3.
 Shown are bias, variance and MSE of the estimates from accept/reject and hit and run, using between 1000 and 8000 samples, averaged
@@
 \begin{comment}
-\subsection{An analytic approximation for the max-t test}
+\subsection{An analytic approximation for the max-$t$ test}
  Suppose we have taken $k$ steps of forward stepwise regression with $p$ predictors and let the selected model  
@@
  Finally let $T=|T_{j^*}|$.
-Then by assuming independence, a  simple approximation to the max-t value is given by
+Then by assuming independence, a  simple approximation to the max-$t$ value is given by
 \begin{equation}
@@
 There are two moderate signals.
-Shown are the p-values from the ``exact'' numerical computation (with $100,000$ bootstrap samples)
+Shown are the $p$-values from the ``exact'' numerical computation (with $100,000$ bootstrap samples)
 and the approximation based on (\ref{eqn:approxpv}).
@@
   \includegraphics[width=3in]{figs/approxpv.pdf}
-  \caption{Exact max-t pvalues and approximation based on (\ref{eqn:approxpv}), for a simulated example
+  \caption{Exact max-$t$ $p$-values and approximation based on (\ref{eqn:approxpv}), for a simulated example
   with $n=50, p=10$.}
@@
   \includegraphics[width=1\textwidth]{figs/simulation_snr_5_alpha_05_null_true.pdf}
-  \caption{\em  CDFs of nominal (black), MaxT (red), MaxT with unknown $\sigma$ (green) and  saturated-model (blue) p-values in the simulation of Section~\ref{sec:sparseReg}, conditional on testing a true null hypothesis at step $k$.
-The nominal test is badly anti-conservative, while all of the other methods  show uniform p-values as desired.}
+  \caption{\em  CDFs of nominal (black), MaxT (red), MaxT with unknown $\sigma$ (green) and  saturated-model (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on testing a true null hypothesis at step $k$.
+The nominal test is badly anti-conservative, while all of the other methods  show uniform $p$-values as desired.}
   \label{fig:simulation_null_true}
@@
 Because the correct model is seven-sparse, $k=8$ is the first index for which the null can possibly be true.
-Both selective $p$-values are uniform by construction, but the nominal $p$-values are highly anti-conservative, as expected.
+All but the nominal $p$-values are uniform by construction, but the nominal $p$-values are highly anti-conservative, as expected.
 
@@
 
-This is {\em not} a mistake in our implementation of the method, but rather a consequence of our ``model-centric'' point of view.
+This is not an error in our implementation of the method, but rather a consequence of our ``model-centric'' point of view.
 If we try to add a noise variable to the model before we have included all the signal variables, then we are testing a false null hypothesis.
@@
 \item its model-wise FDR, and
-\item its variable-wise FDR, where we use $\tV= \#\{\text{ noise variables included in } M_{\hk}\}$ instead of $V=(\hk-k_0)_+$.
+\item its variable-wise FDR, where we use $V^{\text{full}}= \#\{\text{ noise variables included in } M_{\hk}\}$ instead of $V=(\hk-k_0)_+$.
 \end{itemize} 
 The last measure of performance is not explicitly controlled by any of the selective-inference methods, but we might nevertheless hope to perform reasonably.
-\WFcomment{Comment on results once they are correct.
-These numbers are wrong right now.}
-
-\WFcomment{Should we compute the FWER and FDR only conditional on screening? Strong / Forward / Basic should all be correct given screening since they condition on $p_{1:k_0-1}$.}
 
@@
 \caption[tab:stopping]{\em Results of various stopping rules applied to simulated data with 7 strong signals,  as described at the beginning of Section \ref{sec:sparseReg}.
-The {\tt Simple} rule stops at the first time that a p-value exceeds $\alpha=0.05$, while {\tt Forward} and {\tt Strong} refers to the ForwardStop and StrongStrop rules of Section \ref{sec:sequential}.
+The {\tt Simple} rule stops at the first time that a $p$-value exceeds $\alpha=0.05$, while {\tt Forward} and {\tt Strong} refers to the ForwardStop and StrongStrop rules of Section \ref{sec:orderedProposals}.
 Results with theoretical guarantees are \guarantee{highlighted}.}
@@
 \caption[tab:stopping]{\em Results of various stopping rules applied to simulated data with 7 strong signals,  as described at the beginning of Section \ref{sec:sparseReg}.
-The {\tt Simple} rule stops at the first time that a p-value exceeds $\alpha=0.2$, while {\tt Forward} and {\tt Strong} refers to the ForwardStop and StrongStrop rules of Section \ref{sec:sequential}.
+The {\tt Simple} rule stops at the first time that a $p$-value exceeds $\alpha=0.2$, while {\tt Forward} and {\tt Strong} refers to the ForwardStop and StrongStrop rules of Section \ref{sec:orderedProposals}.
 Results with theoretical guarantees are \guarantee{highlighted}.}
@@
 
-We see that 1) strongStop and forwardStop control FWER and modelwise FDR, respectively, using p-values from MaxT or MaxT-identify, as predicted by theory
-2) the nominal p-values do not lead to control of FWER or FDR, as expected, 3) the pvalues from the saturated model show FWER and modelwise-FDR control,
+We see that 1) strongStop and forwardStop control FWER and modelwise FDR, respectively, using $p$-values from MaxT or MaxT-identify, as predicted by theory
+2) the nominal $p$-values do not lead to control of FWER or FDR, as expected, 3) the $p$-values from the saturated model show FWER and modelwise-FDR control,
 although not guaranteed by theory, 4) the MaxT rules show higher probabilities of selecting the true model, especially using forwardStop, 5) the knockoff method
@@
  By successively conditioning on $u_j, j<k$ one can obtain an exact sequential test
-for the complete null hypothesis $H_0: \lambda_j=0 \; \forall j \neq k$.  The computation of the resulting p-values requires numerical integration.
+for the complete null hypothesis $H_0: \lambda_j=0 \; \forall j \neq k$.  The computation of the resulting $p$-values requires numerical integration.
 The details are given in \citet{choi2014selecting}, with their ``integrated CSV'' test  corresponding to the sequential selective test of this paper.
@@
 This procedure is used, for example in the R package for gradient boosting, called {\tt gbm}.
-The best first tree-builder is thus a sequential selection  procedure,  to which we can apply the results in this paper. Hence we can obtain independent, sequential p-values
+The best first tree-builder is thus a sequential selection  procedure,  to which we can apply the results in this paper. Hence we can obtain independent, sequential $p$-values
 for each successive split in the decision  tree. We now give some details and an example.
@@
 Nonetheless,  the selection procedure  for the first set of $k$ splits can be expressed in polyhedral from $Ay \leq b$, and the saturated and selective inference procedures can be applied.
-As before, the saturated p-values have an explicit truncated Gaussian form, while computation of the selective p-values require accept-reject sampling.
+As before, the saturated $p$-values have an explicit truncated Gaussian form, while computation of the selective $p$-values require accept-reject sampling.
 
@@
 
-As expect, the nominal p-values are far too small and hence don't control type I error.
-Both the saturated and selective p-values control type I error, with the selective p-values are somewhat more powerful in the non-null case.
-Rules like the forwardStop procedure could be applied  to the selective p-values to choose a ``best'' model size.
+As expect, the nominal $p$-values are far too small and hence don't control type I error.
+Both the saturated and selective $p$-values control type I error, with the selective $p$-values are somewhat more powerful in the non-null case.
+Rules like the forwardStop procedure could be applied  to the selective $p$-values to choose a ``best'' model size.
 \begin{figure}[h]
@@
   \includegraphics[width=.65\textwidth]{figs/cart.pdf}
-  \caption{\em Cumulative distribution functions of P-values for  the first two splits in a CART decision tree. In the top row, the true model is null,
+  \caption{\em Cumulative distribution functions of $p$-values for  the first two splits in a CART decision tree. In the top row, the true model is null,
   while in the bottom row, the data are generated from a  model of the form (\ref{eqn:intmodel})}
@@
 
-\section{Discussion}
-\label{sec:pca}
+\section{Discussion}\label{sec:discussion}
+
+\WFcomment{Adding variables in groups? Changing the set of variables to select from at each step?}
 
@@
 
-We plan to distribute  open source software, in both the R and Python languages,  for computing the max-t and other tests discussed in this paper.
+We plan to distribute  open source software, in both the R and Python languages,  for computing the max-$t$ and other tests discussed in this paper.
 

### 2015-10-19 paper.tex
rob edited sec 1-5

@@
 In step one, because $E=\{1_n\}$ is fixed and $t_E^*$ is independent of $U_E$, the conditional test reduces to the max-$t$ test proposed in \citet{buja2014}.
-In later steps, $t_E^*$ and $U_E$ are not conditionally independent of $E$, but we can determine the null conditional distribution through Monte Carlo sampling.
+In later steps, $t_E^*$ and $U_E$ are not conditionally independent of $E$, but we can estimate the null conditional distribution through Monte Carlo sampling.
 
@@
 These advantages do not come entirely for free, as we will see: the selected-model methods test a more restrictive null hypothesis than the saturated-model methods.
+In addition, the selected-model test require accept-reject or MCMC sampling, while the saturated-model tests are available in closed form.
 
@@
 \]
-By {\em model}, we mean a family of candidate probability distributions for $Y$.
+By {\em model}, we mean a family of candidate probability distributions for $Y$. For example in  linear regression, a model specifies  the set of
+predictors with nonzero coefficients (but not the  values of their coefficients).
 Without loss of generality, we can take $M_\infty$ as the union of all models under consideration.
@@
 Another important class of model selection procedures is the sequence of {\em ever-active} sets for a regularized likelihood path. 
+The notion of ever-active sets is needed since these solution paths can drop  (as well as add) predictors along the way.
 	For some ordered set $\Lambda$, let $\lambda\in\Lambda$ parameterize the regularization penalty $P_\lambda(\theta)$.
 In the case of lasso-penalized regression, we could take $P_\lambda = \lambda\|\beta\|_1$ for $\lambda\in (0,\infty]$.
+We assume $P_\lambda(\beta)$  has the property  that the model becomes less constrained as $\lambda$ decreases, that is,
+ if $\lambda_2<\lambda_1$ then $\{\beta:P_{\lambda_2}(\beta)\} \supset \{\beta:P_{\lambda_1}(\beta)\}$.  \RTcomment{Will: is this correct?}
 
@@
   \argmin_{\theta\in\Theta} -\ell(\theta; Y) + P_\lambda(\theta) \\
-  \tE_\lambda(Y) &= \left\{j:\; \hat\theta_j^\nu \neq 0 
+  \tE_\lambda(Y) &= \left\{j:\; \hat\theta_j^\nu(Y) \neq 0 
     \text{ for any } \nu\in \Lambda, \nu \geq \lambda \right\}
 \end{align}
-It may be impossible or inconvenient to compute $\hat\theta^\lambda$ for every $\lambda\in(0,\infty]$. If so, we can instead take $\Lambda$ to be a grid of finitely many positive values.
+It may be impossible or inconvenient to compute $\hat\theta^\lambda(Y)$ for every $\lambda\in(0,\infty]$. If so, we can instead take $\Lambda$ to be a grid of finitely many positive values.
 
@@
 \begin{description}
-\item[Complete Null:] $M_{k-1}$ is already correct. That is, 
+\item[Complete Null:] $M_{k-1}$ is (already) correct. That is, 
 \[
@@
 In the second procedure, we fit the full model $M(\{1,\ldots,p\})$ and compute a $t$-statistic for each variable with respect to that model, yielding $t$-statistics $t_1, t_2, \ldots, t_p$ (not ordered by size).
-Iterating for $j=1,\ldots,p$, we compare each $|t_j|$ to the level-$\alpha$ cutoff for a $t$-test with $n-p$ degrees of freedom, and stop selecting variables the first time $|t_{k+1}|$ is not larger than this cutoff.
+For $k=1,\ldots,p$, we compare each $|t_j|$ to the level-$\alpha$ cutoff for a $t$-test with $n-p$ degrees of freedom, and stop selecting variables the first time $|t_{k+1}|$ is not larger than this cutoff.
+We might use this procedure if the variables had a  fixed ordering.
 This procedure would control the full-model-null FWER, but {\em not} the incremental-null FWER.
@@
 More generally, methods that control full-model error rates are best viewed not as {\em model-selection} procedures --- since all inferences are made with respect to the full model, which is specified ahead of time --- but {\em variable-selection} procedures that test multiple hypotheses with respect to a single fixed model.
-Note that the truth or falsity of these hypotheses can depend sensitively on the set of $p$ predictors: rejecting $H_j^{\text{full}}$ has no meaning without reference to the list of all variables that we controlled for, which can make a rejection difficult to interpret when $p$ is large.
+Note that the truth or falsehood of these hypotheses can depend sensitively on the set of $p$ predictors: rejecting $H_j^{\text{full}}$ has no meaning without reference to the list of all variables that we controlled for, which can make a rejection difficult to interpret when $p$ is large.
 Thus, error rates like $\text{FDR}^{\text{full}}$ are best motivated when the full model has some special scientific status.
@@
 However, with the exception of the works above, very few methods offer finite-sample guarantees on the model that is selected except in the orthogonal-design case.
-One additional exception is the knockoff filter of \citet{barber2014controlling}, a variant of which provably controls the full-model FDR.
+One  exception is the knockoff filter of \citet{barber2014controlling}, a variant of which provably controls the full-model FDR.
+We compare our proposal to the Knockoff method in Section \ref{sec:sparseReg}.
 
 Methods like AIC \citet{akaike1974new} and BIC \citep{schwarz1978estimating} are designed for the non-adaptive case, where the sequence of models is determined in advance of observing the data.
-Cross-validation, another general-purpose algorithm for tuning parameter selection, targets out-of-sample error and tends to select many noise variables when the signal is sparse \WFcomment{Whom to cite?}.
+Cross-validation, another general-purpose algorithm for tuning parameter selection, targets out-of-sample error and tends to select many noise variables when the signal is sparse (e.g.
+\cite{LY2015}).
 \citet{benjamini2009simple} extend the AIC by using an adaptive penalty $\sigma^2\lambda_{k,p} k$ to select a model, based on generalizing the Benjamini-Hochberg procedure \citet{benjamini1995controlling}, but do not prove finite-sample control.
@@
 This could in principle lead to violation of the error control guarantees that hold in the fixed-hypothesis case.
-
+\RTcomment{As Will and I discussed, this example is confusing and needs improvement.}
 To see why we a problem could arise, consider apply BasicStop in the following counterexample: let $d=2$ and let $k_0$ be 0 or 1 with probability 1/2 each. Then, let 
@@
 \begin{enumerate}
-\item $p_k$ is conservative given $M_{0:k-1}$ and $U_{k-1}$, and
+\item $p_k$ is conservative given $M_{0:k-1}$ and $U_{k-1}$, and \RTcomment{clearer to say ``sub-uniform''?}
 \item $p_k$ is a function of $M_{0:k}$ and $U_k$.
@@
 
-\section{Discussion}\label{sec:discussion}
+\section{Discussion}
+\label{sec:discussion}
 

### 2015-10-20 paper.tex
some JTcomments

@@
 
+\JTcomment{ Not sure if stopping rule is a great name.
+It suggests it is a ``stopping time'' with respect to some filtration (which it often is).
+But it doesn't have to be.
+Or maybe we should say we call it stopping because we use the property of it being a stopping time?
+
+Another notation quibble for $M(E)$ I kind of prefer the notation $(\beta_E, \sigma^2_E)$ instead of $(\beta, \sigma^2)$. It emphasizes that these are parameters of different models.}
+
 Define the {\em completion index} $k_0(Y) = \min\{k:\; F \in M_k(Y)\}$, the index of the first correct model.
@@
 We assume $P_\lambda(\beta)$  has the property  that the model becomes less constrained as $\lambda$ decreases, that is,
- if $\lambda_2<\lambda_1$ then $\{\beta:P_{\lambda_2}(\beta)\} \supset \{\beta:P_{\lambda_1}(\beta)\}$.  \RTcomment{Will: is this correct?}
+ if $\lambda_2<\lambda_1$ then $\{\beta:P_{\lambda_2}(\beta)\} \supset \{\beta:P_{\lambda_1}(\beta)\}$.
+\RTcomment{Will: is this correct?} \JTcomment{Yes, this looks a little funny to me too.
+Not all penalties produce
+sparsity (for example, the generalized LASSO). So, I think this section is really thinking 
+of $\ell_1$ norm (maybe a group LASSO norm). As written, 
+we are also assuming that
+there is a unique solution to (11) (which is not a big assumption but we might
+as well make it.) Finally, $\ell$ does not actually have to be a likelihood.}
 
@@
 The methods we consider in this article are essentially generalizations of this procedure for the setting where the model sequence is selected adaptively, and where we may wish to control FDR instead of FWER.
+
+\JTcomment{Maybe a comment about Type I and Type III ANOVA tables here?
+I think these are the correct types of ANOVA tables but am not sure...This issue of full vs incremental at least has some classical literature on it.}
 
@@
 \begin{enumerate}
-\item $p_k$ is conservative given $M_{0:k-1}$ and $U_{k-1}$, and \RTcomment{clearer to say ``sub-uniform''?}
+\item $p_k$ is conservative given $M_{0:k-1}$ and $U_{k-1}$, and \RTcomment{clearer to say ``sub-uniform''?} \JTcomment{I think you mean ``super-uniform'', i.e.
+it is larger than a uniform random variable?}
 \item $p_k$ is a function of $M_{0:k}$ and $U_k$.
@@
 
+\JTcomment{I would say our goal is a good stopping rule AND inference on parameters of selected (or perhaps saturated) model. We do not consider it here
+because we don't know how to do it, not because it is not interesting.}
+
 There are several major statistical benefits to selected-model inference.
@@
 
+\JTcomment{We actually don't really need to condition on the signs
+for the theory to go through. It is only so that sampling is feasible.}
+
 Note that  (\ref{eq:boot})  has the form of  (conditional) parametric bootstrap sampling, but  with  a novel twist:  the bootstrap residuals are projected into the subspace
@@
 
+\JTcomment{Is there a reason you have $\psi$ and $\ell$ above instead of just $\ell$? I think the $(E,t)$ above should be $(E,u)$. It seems you never come back to $\psi$.}
+
 \begin{proof}
@@
 \end{proof}
+
+\JTcomment{The notation $R_k$ isn't really defined. I don't think you need an
+index $r$ and $k$, so we could talk about for every $\lambda \in \Lambda \geq \lambda_k$ instead of every $r \leq R_k$. No?
+Here we do use the fact that the objective is a likelihood.}
 
@@
 \WFcomment{Can we do this one?}
+\JTcomment{In principle, we can. But, paper is a little long. Maybe
+just point out in discussion
+that the $p$-values in Choi et al. are examples of this setup?}
+
 Here we consider model selection for principal components analysis. In that case we are given a data matrix $X \in \R^{n\times d}$, with which we form a sample %covariance matrix
@@
 
+\JTcomment{Max $T$ available in python code on github. I suppose R as well, but not sure.}
+
 \section*{Acknowledgments}

### 2015-10-20 paper.tex
clarifying pca comment
some work on computation section

@@
 \citet{taylor2014exact} study methods for generating exact $p$-values at each step of these path algorithms, and their methods provide a starting point for our proposals.
-In addition, \citet{choi2014selecting} discuss a selective approach to determining the correct rank of the covariance matrix in principle components analysis.
+In addition, \citet{choi2014selecting} discuss a selective approach to estimating the rank of the covariance matrix in principle components analysis.
 
@@
 Until this point, we have deferred discussing the question of how to actually compute the rejection cutoffs for the max-$t$ test, next-$\lambda$ test, or any of the other tests we have discussed.
-In each case, we have a test that rejects when some test statistic $T_k$ is large compared to the law
+In each case, the $p$-value at step $k$ comes from a test that rejects when some test statistic $T_k$ is large compared to the law
 \begin{equation}\label{eq:genDist}
-\L(T_k \,\mid\, M_{k-1}, U_{k-1}),
+\L(T_k(Y) \,\mid\, M_{k-1}(Y), U_{k-1}(Y)),
 \end{equation}
 under the null hypothesis that $F\in M_{k-1}$.
-As such, we have reduced computation of the $p$-value at step $k$ to a sampling problem: if we can sample values of $Y$ from the law~\eqref{eq:genDist}, we can numerically approximate the $p$-value to whatever precision we desire.
+Because $U_{k-1}$ is sufficient for $Y$ under the null, this law is completely known.
+Thus, we have reduced computation of the $p$-value at step $k$ to a sampling problem: if we can sample values of $Y$ from its null distribution given $M_{k-1}(Y), U_{k-1}(Y)$, we can numerically approximate the $p$-value to whatever precision we desire.
 
 Due to the generality in which we have posed the problem here, we cannot give a general prescription for how to carry out this sampling.
+However, we can go into some detail for the important case of linear regression.
+
+\subsection{Details for Linear Regression}
+
+As discussed in Section~\ref{sec:singleStep}, we will set
+\[
+p_k(Y) = p_{k, E_{k-1}(Y)}(Y),
+\]
+where $p_{k,E}(Y)$ for each $E\sub \{1,\ldots, p\}$ is a conditional $p$-value for the {\em fixed} null model $M(E):\, Y \sim N(X_E\beta, \sigma^2)$, valid on the selection event $A=\{E_{k-1}(Y)=E\}$.
+
+We focus here on the simpler case where $\sigma^2$ is known.
+For $\sigma^2$ unknown, see \citet{fithian2014optimal}.
+If $\sigma^2$ is known, then $U_{k-1}(Y) = X_{E}'Y$ on $A$, so our task is to sample under the null from
+\[
+\L(Y \mid X_E'Y, Y\in A),
+\]
+a truncated multivariate Gaussian distribution.
+
+\begin{comment}
+ If $\sigma^2$ is unknown, then $U_{k-1}=(X_E'Y,\, \|Y\|^2)$ on $A$ and we must sample under the null from
+\[
+\L(Y \mid X_E'Y,\, \|Y\|^2,\, Y\in A).
+\]
+\end{comment}
+
+\paragraph{Accept-Reject Sampling:}
+Note that sampling from $\L(Y \mid X_E'Y)$ is easy. Letting $\proj_{X_E}$ denote projection into the column space of $X_E$, we can decompose $Y$ into two independent terms:
+\[
+Y = \proj_{X_E}Y + \proj_{X_E}\perp Y.
+\]
+The first term is fixed conditional on $X_E'Y$, while the second is a degenerate Gaussian distribution with mean zero and covariance matrix $\sigma^2\proj_{X_E}^\perp$.
+
+For a fixed set $E$, we must be able to ample from
+
+we must be able to sample from the event on the event $E_{k-1}(Y)=E$,
+
+ if $\sigma^2$ is known, and $(X_{E_{k-1}}'Y, \|Y\|^2)$ otherwise. Thus for a fixed active
+
+ In either case, sampling from $\L(Y \mid U_{k-1})$ is easy under the null hypothesis that $\E Y$ lies in the column span of $X_{E_{k-1}}$. Write
+\[
+Y = \proj_{E_{k-1}} 
+\]
+
+If $\sigma^2$ is known, then
+\[
+Y = 
+\]
+
+, and we need to sample under the null hypothesis that $Y$ is marginally Gaussian, with $\E Y$ in the column span of $X_{E_{k-1}}$.
+
+: if $\sigma^2$ is known, then $Y$
+
+but sampling from $\L(Y \mid E_{k-1}, \, U_{k-1})$ . We focus first on the casIf $\sigma^2$ is known, we can sample
+
+ therefore not a given $U_{k-1}$ is n
+
+The event that $M_{k-1}(Y)$
+
+\subsection{Linear Regression}
+
+See \citet{fithian2014optimal}
+
  Instead, we give details of our computational strategy here for one case, the max-$z$ test in forward stepwise regression with known $\sigma^2$.
@@
 just point out in discussion
-that the $p$-values in Choi et al. are examples of this setup?}
+that the $p$-values in Choi et al.
+are examples of this setup?
+The notation in Yunjin's paper is nothing like this, but we do form selected model tests with the sufficient statistic growing in dimension as we go down the path.
+In this case, ``maxT'' is the same as ``maxT identify'' where we ``identify'' the sets of singular vectors...}
 

### 2015-10-21 paper.tex
merged

@@
 
-For a fixed set $E$, we must be able to ample from
-
-we must be able to sample from the event on the event $E_{k-1}(Y)=E$,
-
- if $\sigma^2$ is known, and $(X_{E_{k-1}}'Y, \|Y\|^2)$ otherwise. Thus for a fixed active
-
- In either case, sampling from $\L(Y \mid U_{k-1})$ is easy under the null hypothesis that $\E Y$ lies in the column span of $X_{E_{k-1}}$. Write
-\[
-Y = \proj_{E_{k-1}} 
-\]
-
-If $\sigma^2$ is known, then
-\[
-Y = 
-\]
-
-, and we need to sample under the null hypothesis that $Y$ is marginally Gaussian, with $\E Y$ in the column span of $X_{E_{k-1}}$.
-
-: if $\sigma^2$ is known, then $Y$
-
-but sampling from $\L(Y \mid E_{k-1}, \, U_{k-1})$ . We focus first on the casIf $\sigma^2$ is known, we can sample
-
- therefore not a given $U_{k-1}$ is n
-
-The event that $M_{k-1}(Y)$
-
 \subsection{Linear Regression}

### 2015-10-24 paper.tex
edits to computation section, light edits to simulation section

@@
 \newcommand{\pow}{\text{Pow}}
+\newcommand{\supp}{\text{supp}}
 \newcommand{\sF}{\mathscr{F}}
@@
 
-\subsection{Single-Step $p$-Values}
+\subsection{Single-Step $p$-Values}\label{sec:singleStep}
 
@@
 
-Due to the generality in which we have posed the problem here, we cannot give a general prescription for how to carry out this sampling.
-However, we can go into some detail for the important case of linear regression.
-
-\subsection{Details for Linear Regression}
+In many cases it is easy to sample $Y$ given $U_{k-1}$, so computational difficulties only arise insofar as the selection event $\{M_{k-1}(Y)=M\}$ is difficult to condition on.
+Due to the generality in which we have posed the problem, we cannot give a general prescription for how to carry out this sampling.
+However, we give some details for the important case of linear regression in Section~\ref{sec:linRegComputation}.
+
+For a completely general selection algorithm, we might expect severe computational difficulties when we attempt to condition on the null model at stage $k$, $M_{k-1}(Y)$.
+For example, in forward stepwise regression, $M_{k-1}$ is the accumulated result of $k-1$ steps of greedy variable selection.
+At each step $s<k$, the selection of variable $j_s$ is a result of comparing its $t$-statistic with the $t$-statistics of the other $p-k$ inactive variables.
+Based on this logic, we might expect to accumulate at least $O(p)$ constraints at each step, resulting in $O(kp)$ constraints after $k$ steps.
+
+For lasso linear regression, the situation seems even more dire: for any fixed value of $\lambda$, the event $\{\supp(\hat\beta^\lambda) = E\}$ is a union of up to $2^{|E|}$ polytopes in $\R^n$, each satisfying $O(p)$ constraints \citep{lee2013exact}.
+\footnote{To finesse this issue, \citet{lee2013exact} propose conditioning on the signs of $\hat\beta^\lambda$ as well.
+If we do this, the conditioning set becomes a single polytope.}.
+But the event $\{E_{k-1}(Y)=E\}$ depends on $\supp(\hat\beta^\lambda)$ for every $\lambda > \lambda_k$ --- i.e., on the solution to {\em infinitely many} lasso problems.
+
+Remarkably, however, in both forward-stepwise and lasso linear regression, the computational problem actually becomes easier when we move to the sequential setting considered here.
+Indeed, Section~\ref{sec:constraints} shows that for both forward-stepwise regression and lasso linear regression --- and their respective generalizations to exponential family models --- conditioning on the event $\{E_{k-1}=E\}$, for any $k$, only introduces $2p$ linear constraints, taking the form of a rectangular constraint on the sufficient statistics for the inactive variables.
+
+\subsection{Details for Linear Regression}\label{sec:linRegComputation}
 
@@
 
-We focus here on the simpler case where $\sigma^2$ is known.
-For $\sigma^2$ unknown, see \citet{fithian2014optimal}.
 If $\sigma^2$ is known, then $U_{k-1}(Y) = X_{E}'Y$ on $A$, so our task is to sample under the null from
-\[
+\begin{equation}\label{eq:truncNorm}
 \L(Y \mid X_E'Y, Y\in A),
-\]
-a truncated multivariate Gaussian distribution.
-
-\begin{comment}
+\end{equation}
+which is a multivariate Gaussian distribution supported on the hyperplane of dimension $n-|E|$ with $X_E'Y$ fixed at its realized value, and truncated to the event $A$.
+
 If $\sigma^2$ is unknown, then $U_{k-1}=(X_E'Y,\, \|Y\|^2)$ on $A$ and we must sample under the null from
-\[
-\L(Y \mid X_E'Y,\, \|Y\|^2,\, Y\in A).
-\]
-\end{comment}
-
-\paragraph{Accept-Reject Sampling:}
-Note that sampling from $\L(Y \mid X_E'Y)$ is easy. Letting $\proj_{X_E}$ denote projection into the column space of $X_E$, we can decompose $Y$ into two independent terms:
-\[
-Y = \proj_{X_E}Y + \proj_{X_E}\perp Y.
-\]
-The first term is fixed conditional on $X_E'Y$, while the second is a degenerate Gaussian distribution with mean zero and covariance matrix $\sigma^2\proj_{X_E}^\perp$.
-
-\subsection{Linear Regression}
-
-See \citet{fithian2014optimal}
-
- Instead, we give details of our computational strategy here for one case, the max-$z$ test in forward stepwise regression with known $\sigma^2$.
-
-Given a model $M_{k-1}$   we need to sample from 
-\begin{eqnarray}
-L(X'_jY \,|\, E_{k-1},X'_{E_{k-1}}Y)
-\label{eqn:maxtsamp}
-\end{eqnarray}
-with $Y\sim N(\mu, I \cdot \sigma^2)$.
-The current model fit is $\proj_{E_{k-1}} Y$ and it is easy to see that (\ref{eqn:maxtsamp})  is equivalent to sampling
-\begin{eqnarray}
-Y^*=\proj_{E_{k-1}} Y + (I- \proj_{E_{k-1}}) Z; \;Z\sim N(0,I\cdot\sigma^2)
-\label{eq:boot}
-\end{eqnarray}
-and then conditioning on the event $E_{k-1}(Y^*)= E_k(Y)$ (a polyhedron).
-In other words, we need to condition on data vectors  $Y^*$ that yields the same
-sequence of variables and coefficient signs up to step $k-1$. This is the challenging aspect of this computation.
-
-\JTcomment{We actually don't really need to condition on the signs
-for the theory to go through. It is only so that sampling is feasible.}
-
-Note that  (\ref{eq:boot})  has the form of  (conditional) parametric bootstrap sampling, but  with  a novel twist:  the bootstrap residuals are projected into the subspace
-orthogonal to the current fit.  The typical justification for the bootstrap is asymptotic, relying on the convergence of the parameter estimates to the true parameters.
-But something magical  occurs in the selective sampling scheme (\ref{eq:boot}): this  sampling fixes the value of $ X_{E_{k-1}}^Ty$ and this conditioning
-removes the dependence on nuisance parameters.
-Hence  (\ref{eq:boot}) yields the exact (non-asymptotic) conditional distributions and $p$-values.
-
-Figure \ref{fig:comparison} shows an example with $n=100, p=20$ and one strong signal, computing the max-$t$ $p$-value at Step 3.
+\begin{equation}\label{eq:truncUnif}
+\L(Y \mid X_E'Y,\, \|Y\|^2,\, Y\in A),
+\end{equation}
+which is a uniform distribution supported on a sphere of dimension $n-|E|-1$ defined by fixing $X_E'Y$ and $\|Y\|^2$ at their realized values, and truncated to the event $A$.
+For the simulations in this article, we use a hybrid of two strategies:
+
+\paragraph{Accept/Reject Sampler} In either of these two cases, sampling would be very easy if $A$ were the entire sample space $\R^n$; the only reason we might encounter difficulty is that $A$ could be an irregular set that is difficult to condition on.
+However, we can always sample from $\L(Y \mid U)$ and then keep each new sample $Y$ if it lies in the selection event.
+
+\paragraph{Hit-and-Run Sampler} \citet{fithian2014optimal} propose sampling from laws like that in~\eqref{eq:truncNorm} via a hit-and-run sampler for the $n-|E|$-dimensional multivariate Gaussian distribution: at each step, to generate $Y^{t+1}$ from $Y^t$, we choose a uniformly random unit vector $\nu$ in the hyperplane, and resample $\nu'Y$ conditional on the constraints $\proj_{\nu}^\perp Y^{t+1}=\proj_{\nu}^\perp Y^t$ and the constraint $Y^{t+1}\in A$.
+If $A$ is a polytope this amounts to resampling a univariate Gaussian distribution with known mean and variance, truncated to an interval defined by maximizing and minimizing $\nu'Y^{t+1}$ subject to linear constraints.
+For the case with $\sigma^2$ unknown, the sampler is a bit more involved; see \citet{fithian2014optimal}.
+
+Figure \ref{fig:comparison} compares the two samplers in an example with $n=100, p=20$ and one strong signal, computing the max-$t$ $p$-value at Step 3.
 The ``exact'' answer was determined from bootstrap sampling with $20,000$ replications.
-The Figures show the bias, variance and MSE of the estimates from accept/reject and hit and run, using between 1000 and 8000 samples, averaged
+The Figures show the bias, variance and mean squared error (MSE) of the estimates from accept/reject and hit and run, using between 1000 and 8000 samples, averaged
 over 10 different starting seeds.
@@
   \caption{\em Simulated example with $n= 100, p=20$ and one strong signal, computing the max-$t$ $p$-value at Step 3.
-Shown are bias, variance and MSE of the estimates from accept/reject and hit and run, using between 1000 and 8000 samples, averaged
-over 10 different starting seeds.
+Shown are bias, variance and MSE of the estimates from accept/reject and hit and run, using between 1000 and 8000 samples, averaged over 10 different starting seeds.
 }
@@
 We see that accept/reject is far more efficient, yielding a smaller MSE with 1000 samples then hit and run gives for 8000 samples.
-Hence we use accept/reject for as many steps of the sequence is possible,  until the point where the acceptance rate is too low to make is practical.
-
-In detail, we employ a hybrid computational strategy; we use accept-reject sampling, and do so as long as we obtain at least 300 (say) samples in the first set
-of $B$ realizations ($75,000$ in our code)  This is an effective approach until we have reached a point in the model path where all predictors have little or no signal.
-When the acceptance rate gets too low, we switch to a hit-and run Monte Carlo approach. We exploit the fact that each move along a random direction 
-has a truncated Gaussian distribution whose truncation limits are easily computed from the properties of the polyhedron. Unlike accept-reject, sampling
-which produces independent samples,  the hit and run method produces dependent samples. Hence we must run it for longer with an initial burn in period,
-and hope that it has mixed sufficiently well.
-
-The task is made much simpler by the fact that the selection event can we written as $S_M=\{y: \Gamma y \geq u\}$.
-Hence we don't need to run the entire forward stepwise procedure on $y^*$: instead, we pre-compute $\Gamma$ and $u$ and then check
-if $\Gamma y^* \geq u$. Further, since $\Gamma$ has on the  order of $2pk$ rows after $k$ steps,  we can save 
-substantially in computation by checking the row-wise inequalities starting from the last row, where the inequality is most likely to fail.
-
-\subsection{Reducing the number of constraints}
-Happily, it turns out that the conditioning polytope can be expressed with only $2p$ constraints in forward stepwise regression, as we show here.
-This increases the speed of both accept/reject and hit and Run sampling for our problem.
+As we move further along the path, the set $A$ becomes more and more constraining, so that we must discard the vast majority of the samples we generate.
+Thus we use accept/reject for as many steps of the sequence is possible,  until the point where the acceptance rate is too low to make is practical.
+
+In detail, we employ a hybrid computational strategy; we use accept-reject sampling, and do so as long as we obtain at least 300 (say) samples in the first set of $B$ realizations ($75,000$ in our code)  This is an effective approach until we have reached a point in the model path where all predictors have little or no signal.
+When the acceptance rate gets too low, we switch to a hit-and run Monte Carlo approach.
+We exploit the fact that each move along a random direction  has a truncated Gaussian distribution whose truncation limits are easily computed from the properties of the polyhedron.
+Unlike accept/reject sampling, which produces independent samples,  the hit and run method produces dependent samples.
+Hence we must run it for longer with an initial burn in period, and hope that it has mixed sufficiently well.
+
+The task is made much simpler by the fact that the selection event can we written as a polytope $A=\{y: \Gamma y \geq u\}$.
+Hence we don't need to run the entire forward stepwise procedure on $Y^*$: instead, we pre-compute $\Gamma$ and $u$ and then check if $\Gamma Y^* \geq u$.
+In the case of forward stepwise and lasso regression, $\Gamma$ has $2(p-k)$ rows after $k$ steps, as we see next.
+
+\subsection{Constraints}\label{sec:constrainsts}
+Happily, it turns out that the conditioning polytope can be expressed with only $2(p-k)$ constraints in forward stepwise regression and the lasso, as we show here.
+This increases the speed of both accept/reject and hit-and-run sampling for our problem.
+
+We begin by proving the result for forward stepwise linear regression, followed by a more general result for the forward stepwise likelihood path in exponential families.
 
@@
   \]
-  The crucial observation is that, 
-  once we know the $k$th active set 
-  $E_k$ and its sufficient statistics
-  $X_{E_k}'Y$, we know the entire 
-  path of fitted models up to step $k$.
-  On the set $\{E_k=E\}$, all of the quantities 
-  $j_{i+1}$, $E_i$, $\hat\beta^i$, and $\proj_{E_i}^\perp$
-  depend only on $X_E'Y$. For brevity, write
+  Because forward stepwise regression satisfies the SSP, we know the entire path of fitted models up to step $k$ once we condition on the $k$th active set $E_k$ and its sufficient statistics $X_{E_k}'Y$.
+On the set $\{E_k=E\}$, all of the quantities $j_{i+1}$, $E_i$, $\hat\beta^i$, and $\proj_{E_i}^\perp$ depend only on $X_E'Y$.
+For brevity, write
   \[
@@
   \]
-  On $A$, $C_i^*$ is attained at $j=j_{i+1}$, the $i$th 
-  variable added.
+  On $A$, $C_i^*$ is attained at $j=j_{i+1}$, the $i$th variable added.
 
@@
 
-As we see below, a similar result holds for the LASSO ever-active path. In fact, it holds for a much larger class of $\ell_1$-regularized exponential family models including the graphical LASSO.
+As we see below, the same result holds for the much larger class of forward stepwise likelihood paths in exponential family models.
+
+\begin{theorem}
+  Assume that $M_\infty$ is an exponential family model
+  of the form
+  \[
+  Y \sim \exp\{ \theta'U(y) - \psi(\theta) \}\,d\nu(y),
+  \]
+  with $\Theta \sub \R^p$ convex, and assume that the model path $M_{0:d}$ is given by the forward stepwise likelihood path with
+\begin{equation}
+j_k = \argmax_j \;\;\sup \left\{\ell(\theta; Y):\; \theta\in\Theta(E_{k-1} \cup \{j\})\right\}, \quad \text{ and } \quad
+E_k = E_{k-1} \cup \{j_k\}.
+\end{equation}
+
+  Then, for a candidate active set $E$ of size $k$, the set $A = \{E_k = E, \;U_E = u\}$ is characterized exactly by the constraints $U_E = u$ and
+  \[
+  v_j^-(E,u) \leq U_j(Y) \leq v_j^+(E,u), \quad\forall j \notin E,
+  \]
+  for $v_j^-$ and $v_j^+$ given in Equations~(\ref{eq:vMinus_FSL}--\ref{eq:vPlus_FSL}).
+  Thus, $A$ corresponds exactly to a set with $2(p-k)$ linear inequality constraints and $k$ linear equality constraints on $U(Y)$.
+\end{theorem}
+
+\begin{proof}
+  For a set $E\sub \{1,\ldots,p\}$ define
+  \[
+  \ell_E^*(U_E(Y)) 
+  = \sup\left\{\theta'U(Y) - \psi(\theta) :\; \theta\in \Theta(E) \right\}
+  = \sup\left\{\theta_E'U_E(Y) - \psi_E(\theta_E):\; \theta\in\Theta(E) \right\},
+  \]
+  where $\psi_E:\; \R^{|E|} \to \R$ is defined such that $\psi(\theta)=\psi_E(\theta_E)$ if $\theta_{\setminus E}=0$.
+Note that $\ell_E^*$ is the convex dual of the convex function $\psi_E$, so $\ell_E^*$ is also convex.
+At stage $i<k$, the next variable to enter is the one that maximizes $\ell_{E_i\cup \{j\}}^*(U(Y))$.
+
+  Because the forward stepwise likelihood path satisfies the SSP, we know the entire path of fitted models up to step $k$ once we condition on $E_k$ and its sufficient statistics $U_{E_k}(Y)$.
+On the set $\{E_k=E\}$, all of the quantities $j_{i+1}$, $E_i$, $\hat\theta^i$, and $\proj_{E_i}^\perp$ depend only on $U_E(Y)$.
+For brevity, write
+  \[
+  C_i^* = \max_{j \notin E_i} 
+    \sup \left\{ \theta ' U - \psi(\theta):\; \theta \in \Theta(E_i\cup \{j\}) \right\}.
+  \]
+  On $A$, $C_i^*$ is attained at $j=j_{i+1}$, the $i$th variable added.
+
+  If $U_E(Y)$ is fixed at $u$, and $j \notin E$, then the condition for 
+  variable $j$ {\em not} to enter at step $i+1 < k$ is
+  \[
+  \ell_{E_i\cup \{j\}}^*(U(Y)) \leq C_i^*.
+  \]
+  Note that $\ell_{E_i\cup \{j\}}^*$ depends only on the sufficient statistics $U_{E_i}(Y)$ --- which is fixed, on $A$ --- and the remaining sufficient statistic $U_j(Y)$.
+Since $\ell_{E_i\cup\{j\}}^*$ is convex, the set
+  \begin{equation}\label{eq:noEnterBounds_FSL}
+  \{U_E=u, \ell_{E_i\cup \{j\}}^*(U(Y)) \leq C_i^*\} 
+  = \{ U_E=u, v_{j,i}^-(E,u) \leq U_j \leq v_{j,i}^+(E,u)\},
+  \end{equation}
+  where $v_j^{-}$ and $v_j^+$ are the two values where $\ell_{E_i\cup \{j\}}^*(U)=C_i^*$.
+
+  The set $A$ is equivalent to 
+  Equation~\eqref{eq:noEnterBounds_FSL} holding
+  for every $i < k$ and $j \notin E$. Nominally, this gives $2k(p-k)$
+  linear inequality constraints to satisfy, but most of them are
+  non-binding. On $A$, the upper and lower bounds
+  in~\eqref{eq:noEnterBounds_FSL}
+  are all known functions of $E$ and $u$, so we can set
+  \begin{align}\label{eq:vMinus_FSL}
+    v_j^-(E,u) &= \max_{0 \leq i < k} \;\;v_{j,i}^-(E,u)\\
+    \label{eq:vPlus_FSL}
+    v_j^+(E,u) &= \min_{0 \leq i < k} \;\;v_{j,i}^-(E,u)
+  \end{align}
+\end{proof}
+
+A similar result also holds for the class of $\ell_1$-regularized exponential family models including the graphical LASSO.
 
@@
   \begin{equation}\label{eq:regProblem}
-  \hat\theta^r = \argmin_{\theta\in \Theta} 
-  -\ell(\theta) + \lambda_r\|\theta\|_1,
-  \end{equation}
-  for some sequence $\lambda_r > 0$.
+  \hat\theta^\lambda = \argmin_{\theta\in \Theta} 
+  -\ell(\theta) + \lambda\|\theta\|_1,
+  \end{equation}
+  for $\lambda\in \Lambda \sub [0,\infty)$.
 
@@
   \[
-  v_j^-(E,t) \leq U_j(Y) \leq v_j^+(E,t), \quad\forall j \notin E,
+  v_j^-(E,u) \leq U_j(Y) \leq v_j^+(E,u), \quad\forall j \notin E,
   \]
@@
 \JTcomment{Is there a reason you have $\psi$ and $\ell$ above instead of just $\ell$? I think the $(E,t)$ above should be $(E,u)$. It seems you never come back to $\psi$.}
+\WFcomment{Fixed t/u. $\psi$ is the log-proportionality constant, I think we need it. Maybe I'm misunderstanding your question?}
 
@@
   \[
-  \hat\theta^r = \hat\theta^{(E,r)}
-  \]
-  for every $r \leq R_k$, so we know the entire path of fits up to and including $R_k$. But then, excluding variable $j \notin E$ at Lagrange parameter $\lambda_r$ is equivalent to
-  \[
-  \lambda_r \geq 
-  \left| \pardd{\ell(\hat\theta^r)}{\theta_j} \right|
-  = \left|U_j - \E_{\hat\theta^r}[U_j]\right|
-  \]
-  On $A$, for $r \leq R_k$, 
-  $\E_{\hat\theta^r}[U_j]$ is a known function of $E$ and $u$,
+  \hat\theta^\lambda = \hat\theta^{(E,\lambda)}
+  \]
+  for every $\lambda \geq \lambda_k$, so we know the entire path of fits up to and including $\lambda_k$. But then, excluding variable $j \notin E$ at Lagrange parameter $\lambda$ is equivalent to
+  \[
+  \lambda \geq 
+  \left| \pardd{\ell(\hat\theta^\lambda)}{\theta_j} \right|
+  = \left|U_j - \E_{\hat\theta^\lambda}[U_j]\right|
+  \]
+  On $A$, for $\lambda \geq\lambda_k$, 
+  $\E_{\hat\theta^\lambda}[U_j]$ is a known function of $E$ and $u$,
   so we can set
   \begin{align}\label{eq:vMinus_L1}
-    v_j^-(E,u) &= \sup_{r \leq R_k} \;\; 
-    \E_{\hat\theta^r}[U_j] - \lambda_r \\
+    v_j^-(E,u) &= \sup_{\lambda \geq \lambda_k} \;\; 
+    \E_{\hat\theta^\lambda}[U_j] - \lambda \\
     \label{eq:vPlus_L1}
-    v_j^+(E,u) &= \inf_{r \leq R_k} \;\;
-    \E_{\hat\theta^r}[U_j] + \lambda_r .
+    v_j^+(E,u) &= \inf_{\lambda \geq \lambda_k} \;\;
+    \E_{\hat\theta^\lambda}[U_j] + \lambda .
   \end{align}
 \end{proof}
-
-\JTcomment{The notation $R_k$ isn't really defined. I don't think you need an
-index $r$ and $k$, so we could talk about for every $\lambda \in \Lambda \geq \lambda_k$ instead of every $r \leq R_k$. No?
-Here we do use the fact that the objective is a likelihood.}
-
-\WFcomment{Give more general characterization of when we can do it, including group lasso? QP? Generalized Lasso?}
 
@@
 
-We see that 1) strongStop and forwardStop control FWER and modelwise FDR, respectively, using $p$-values from MaxT or MaxT-identify, as predicted by theory
-2) the nominal $p$-values do not lead to control of FWER or FDR, as expected, 3) the $p$-values from the saturated model show FWER and modelwise-FDR control,
-although not guaranteed by theory, 4) the MaxT rules show higher probabilities of selecting the true model, especially using forwardStop, 5) the knockoff method
-does not yields variable-wise FDR control, as it should.
-5)
+We see that basicStop and forwardStop control model-wise FWER and FDR, respectively, using $p$-values from max-$t$ or max-$t$-identify, as predicted by the theory, 2) the nominal $p$-values do not lead to control of FWER or FDR, as expected due to their anti-conservatism.
+Although not guaranteed by the theory, the $p$-values from the saturated model do control FWER and model-wise FDR in this example, 4) the max-$t$ rules show higher probabilities of selecting the true model, especially using forwardStop, 5) the knockoff method does not yield variable-wise FDR control.
+
+\section{Futher Examples and Extensions}
+
+Although our our
+
+\subsection{Principal Components Analysis}
+
+\subsection{Classification and Regression Trees}
+
+\subsection{Nonparametric Changepoint Detection}
+
+Consider a problem in which we observe a time series $(y_1,\ldots,y_T)\in \cY^T$, and wish to detect times $t_1,\ldots,t_k$ where the distribution of $y_t$ changes.
+For example, given a historical document, we may wish to determine
+
+Assuming independence between observation times, we can model
+\[
+y_t \simind F_t, \quad t=1,\ldots, T,
+\]
+with the
+
+ independent $y_t$ for $t=1,\ldots,T$, with
+
 \section{Principal Components Analysis}
@@
 Our work takes a step in the direction of reconciling that contradiction, but there are important questions left to be resolved.
-In particular: which sorts of model misspecification pose the most threat to our inferential conclusions, and how powerful are our tests against these most troublesome sources of misspecification?
+In particular: which sorts of model misspecification pose the greatest threat to our inferential conclusions, and how powerful are our tests against these troublesome sources of misspecification?
 

### 2015-10-30 paper.tex
worked on 'extensions' and 'discussion', took out PCA in favor of nonparametric example

@@
 
-Although our our
-
-\subsection{Principal Components Analysis}
-
-\subsection{Classification and Regression Trees}
+Although we have focused our discussion on the case of linear regression --- a very common and important application of sequential model selection --- our theory is quite general and applies to many other parametric and nonparametric models.
+To give a sense of the broader applicability, we now discuss one parametric and one nonparametric example, each of which fit nicely into our theory.
+
+\subsection{Decision Trees}
+In this section we discuss another application of adaptive sequential inference, to binary decision trees.
+Specifically, the classification and regression tree  (CART)  method builds a decision tree in a top down manner, finding the best available split at each node.
+When considering splitting two daughter nodes, the generic CART algorithm splits the nodes in an arbitrary order, only stopping when a node reaches a minimum size.
+We instead consider a  special version of CART  algorithm:  the "best first"  method, which orders splits by their achieved reduction in loss.
+This procedure is used, for example in the R package for gradient boosting, called {\tt gbm}.
+The best-first algorithm is thus a sequential selection  procedure, to which we can apply the results in this paper.
+Hence we can obtain independent, sequential $p$-values for each successive split in the decision tree.
+
+Given predictor matrix $X \in \R^{n\times p}$ and binary response vector $Y\in \{0,1\}^n$, we begin with the intercept-only model $\logit \P(Y_i \mid X_i = x) \equiv \beta_0$.
+
+At the first step we choose a splitting variable $j_1$ and split point $t_1$ which is a midpoint between two observed values of variable $j$, leading to a model of the form
+\begin{equation}
+\logit \P(Y_i \mid X_i=x) = \beta_0 + \beta_{1} 1\{x_{j_1} > t_1\}.
+\end{equation}
+While there are various ways to choose among potential variables and split points, we assume for the sake of simplicity that $(j_1,t_1)$ are chosen to maximize binomial likelihood (or equivalently, to minimize binomial deviance).
+
+At the second stage, we then split one of the two daughter nodes, choosing the split that gives the greatest increase in likelihood.
+If (say) we split the first node using predictor $j_2$ at split point $t_2$, it leads to the interaction model of the form
+\begin{equation}
+\logit \P(Y_i \mid X_i=x) = \beta_0 + \beta_{1} 1\{x_{j_1} > t_1\} + \beta_2 1\{x_{j_1} \leq t_1, x_{j_2} > t_2\}
+\label{eqn:intmodel}
+\end{equation}
+
+After $k$ steps we obtain a logistic regression model with $k$ features, each of which is an interaction of step functions.
+The model $M_k$ is thus an exponential family whose complete sufficient statistics are the sums of $Y_i$ in each leaf in the tree.
+
+The best-first search for split points can be expressed as a modified form of greedy forward stepwise logistic regression in which the set of candidate features changes at each step, depending on what features have already been included.
+Nevertheless, the proof of Proposition~\ref{prop:forwardSSP} goes through essentially without modification to show that this algorithm satisfies the SSP, and the greedy likelihood ratio test gives $p$-values that are independent on nulls\footnote{Since the test statistic is discrete, we need to randomize if we want truly exact $p$-values}.
+
+This development suggests a possible modification to the CART algorithm in which the tree is ``pruned'' using a rule like ForwardStop applied.
+It could be interesting to investigate whether cross-validation using the target FDR $\alpha$ as a tuning parameter might yield better results in predictive performance, compared to (say) using the tree depth or minimum leaf size as a tuning parameter.
+
+\begin{comment}
+For example the set of candidate basis functions available at each stage are a function of what has been  previously entered into the model
+Nonetheless,  the selection procedure  for the first set of $k$ splits can be expressed in polyhedral from $Ay \leq b$, and the saturated and selective inference procedures can be applied.
+As before, the saturated $p$-values have an explicit truncated Gaussian form, while computation of the selective $p$-values require accept-reject sampling.
+
+Figure \ref{fig:cart} shows an example with independent standard Gaussian features, and $n=30, p=3$.
+We 
+consider just the first two splits, in a null case (top row) and non-null (bottom row).
+In the non-null case,  the data are generated from a  model of the form (\ref{eqn:intmodel}).
+
+As expect, the nominal $p$-values are far too small and hence don't control type I error.
+Both the saturated and selective $p$-values control type I error, with the selective $p$-values are somewhat more powerful in the non-null case.
+Rules like the forwardStop procedure could be applied  to the selective $p$-values to choose a ``best'' model size.
+\begin{figure}[h]
+  \centering
+  \includegraphics[width=.65\textwidth]{figs/cart.pdf}
+  \caption{\em Cumulative distribution functions of $p$-values for  the first two splits in a CART decision tree. In the top row, the true model is null,
+  while in the bottom row, the data are generated from a  model of the form (\ref{eqn:intmodel})}
+  \label{fig:cart}
+\end{figure}
+\end{comment}
 
@@
 
-Consider a problem in which we observe a time series $(y_1,\ldots,y_T)\in \cY^T$, and wish to detect times $t_1,\ldots,t_k$ where the distribution of $y_t$ changes.
-For example, given a historical document, we may wish to determine
-
-Assuming independence between observation times, we can model
-\[
-y_t \simind F_t, \quad t=1,\ldots, T,
-\]
-with the
-
- independent $y_t$ for $t=1,\ldots,T$, with
-
+Consider a problem in which we observe a time series $(Y_1,\ldots,Y_T)\in \cY^T$, and wish to detect times $t_1,\ldots,t_k$ where the distribution of $Y_t$ changes.
+For example, given a historical document, we may wish to investigate whether the entire document was written by the same author or different sections were written by different authors, by quantitatively comparing the writing style across different sections.
+For example, $Y_t$ could could represent the usage distribution in section $t$ over ``context-free'' words such as articles, conjunctions, and prepositions \citep[][see e.g.
+]{chen2015graph}.
+
+For simplicity we assume independence between values of $t$ and model the distribution as piecewise constant between $k$ unknown changepoints.
+For an active set of $k$ changepoints $E\sub \{1,\ldots, T-1\}$, let $t(j,E)$ denote the $j$th smallest element of $E$, and let $t(0,E)=0$ and $t(k+1, E)=T$.
+Then, the nonparametric model with changepoints $E$ is given by
+\begin{equation}\label{eq:nonpar-model}
+  M(E):\; Y_t \simind F_j, \quad \text{ for } \;\;t(j,E) < t \leq t(j+1,E).
+\end{equation}
+
+For $t_2>t_1$ let $V(Y;\, t_1, t_2)\in \cY^{t_2-t_1}$ denote the order statistics of $(Y_{t_1+1},\ldots, Y_{t_2})$. Then the complete sufficient statistic for $M(E)$ is
+\[
+U(Y; \, E) = 
+\bigg(
+V(Y; t(0, E), t(1, E)), \ldots, V(Y;\, t(k, E), t(k+1,E))
+\bigg).
+\]
+Under $M(E)$, resampling from $Y$ given $U$ amounts to randomly permuting $(Y_{t(j,E)+1}, \ldots, Y_{t(j+1,E)})$ independently for each $j=0,\ldots,k$.
+
+For $s \notin E$, testing $M(E)$ against $M(E \cup \{s\})$ amounts to a two-sample test comparing the observations coming immediately before $s$ against the observations coming immediately after $s$.
+Specifically, if $t(j, E) < s < t(j+1, E)$, let $W(Y; \, E, s)$ denote any two-sample test statistic that is measurable with respect to the samples $V(Y; t(j, E), s)$ and $V(Y; s, t(j+1, E))$, and large when there is strong evidence that the two samples come from different distributions.
+
+We can build a model in a greedy fashion beginning with $E_0 = \emptyset$ and then at step $k=1,\ldots,d$ setting
+\begin{equation}
+  t_k = \argmax_t \;\; W(Y; \, E_{k-1}, t), 
+  \quad \text{ and } \quad E_k = E_{k-1} \cup \{t_k\}.
+\end{equation}
+This path algorithm satisfies the SSP because $W(Y; \, E_{i}, t_j)$ is measurable with respect to $U_k=U(Y; \, E_k)$ for every $i<j\leq k$.
+Thus, by inspecting $U_k$ we can determine in what order the first $k$ changepoints were added.
+
+By analogy to the greedy likelihood ratio statistic, a natural choice of test statistic at step $k$ is
+\[
+W_k^*(Y) = \max_t W(Y; \, E_{k-1}, t),
+\]
+which is measurable with respect to $\sF(E_k, U_k)$ by construction. Thus, if we base the step-$k$ $p$-value $p_k$ on the law
+\begin{equation}\label{eq:condPerm}
+\L\left(W_k^* \mid E_{k-1}, U_{k-1}\right),
+\end{equation}
+randomizing to obtain an exactly uniform $p$-value under $M(E_{k-1})$, we will obtain an exact $\sF_k$-measurable $p$-value.
+As a result, the $p$-values will be independent on nulls per Theorem~\ref{thm:suffCond}.
+
+Note that sampling from the law in~\eqref{eq:condPerm} can be carried out by permuting subsequences $(Y_{t(j,E)+1}, \ldots, Y_{t(j+1,E)})$ and accepting the permutation only if it gives the same $E_{k-1}$ as we get in the original data.
+When this simple accept/reject algorithm is impractical we may need to resort to an MCMC strategy.
+We do not pursue these computational matters here.
+
+\begin{comment}
 \section{Principal Components Analysis}
@@
 \end{figure}
+\end{comment}
 
@@
 
-\WFcomment{Adding variables in groups? Changing the set of variables to select from at each step?}
-
 It is a commonplace that ``essentially all models are wrong, but some are useful'' \citep{box1987empirical}.
 In essence, a statistical model is useful if it is large enough to capture the most important features of the data, but still small enough that inference procedures can achieve adequate power and precision.
-Apart from theoretical considerations, the only way to know whether a model is large enough is to test whether it is, using available data.
+Apart from theoretical considerations, the only way to know whether a model is large enough is to test it using available data.
 
@@
 
+In future work we also plan to address the matter of performing selective inference for parameters of the model actually selected.
+In principle, we already know how to do this using the framework of \citet{fithian2014optimal} --- we simply condition on the event that $M_{\hk}$ is selected and perform inference with respect to $M_{\hk}$ --- but there is some work to do in understanding the computational complexity of this task.
+
+\begin{comment}
 Of course, the answer depends on the scientific context: for example, suppose that at step 1, forward stepwise regression selects a variable $X_1$, and then at some later step it selects another variable $X_2$, which is almost perfectly correlated with $X_1$.
@@
 If the second interpretation is the scientifically appropriate one, then we should probably use a different selection algorithm --- for example, a variant of forward stepwise that always adds both $X_1$ and $X_2$ to the model as a group in the same step.
-
-We plan to distribute  open source software, in both the R and Python languages,  for computing the max-$t$ and other tests discussed in this paper.
-
-\JTcomment{Max $T$ available in python code on github. I suppose R as well, but not sure.}
+\end{comment}
+
+Open-source code is available at this article's github repository at \texttt{www.github.com/wfithian/sequential-selection}.
 
@@
 
-The authors are grateful for stimulating and informative conversations with Stefan Wager, Lucas Janson, Trevor Hastie, ....
+The authors are grateful for stimulating and informative conversations with Stefan Wager, Lucas Janson, Trevor Hastie, Yoav Benjamini, and Larry Brown. \WFcomment{Funding}.
 

### 2015-11-01 paper.tex
a few more comments: computation of next-lambda tests, notation for log-likelihood

@@
 there is a unique solution to (11) (which is not a big assumption but we might
-as well make it.) Finally, $\ell$ does not actually have to be a likelihood.}
+as well make it.) Finally, $\ell$ does not actually have to be a likelihood.
+BUT later on we implicitly assume that it is the likelihood for 
+$M_{\infty}$. Why not just say this here, i.e. when talking
+about $\ell_1$ regularization paths the models we have in mind are typically
+going to be ones which correspond to a likelihood for some ``largest model'' for which we are trying to find a sub-model.
+
+An example where we might not take $\ell$ to be a likelihood would be
+if it were a pseudo-likelihood as in the neighborhood selection 
+instead of graphical LASSO.
+}
 
@@
 \end{equation}
+
+\JTcomment{The next-$\lambda$ test is actually a little expensive to compute if we a finite set of $\lambda$  as, say, glmnet does. Should we not mention this? On the plus side, 
+it gives us tests that do not require LAR and can be used with glmnet. 
+Formally, to carry this out test, we need to draw samples of $Y$ conditioned
+on the necessary sufficient statistics satisfying the necessary constraints 
+(for LASSO, there $2(p-k)$ constraints on $X_{-E}^Ty$). Then, for each sample
+we restart glmnet and run until the active set changes, storing this
+value of $\lambda$. Then we compare our observed next-$\lambda$ to this
+sample of $\lambda$'s.
+
+In the LAR-LASSO case, this is easier as there is an explicit formula
+to compute the next event. But, LAR is limited to LASSO whereas next-$\lambda$
+in principle works for group LASSO.
+
+Finally, the simplification of constraints we get really only works for
+almost only for $\ell_1$. For group LASSO, the simplification
+doesn't happen. The constraints after $k$ steps look like
+$$
+\left\{y: P_Ey=z, \|X_g^T(y - \hat{\mu}_l(z))\|_2 \leq \lambda_l \omega_g, 
+g \in -E, 1 \leq l \leq k, \|(I-P_E)y\|=c\right\}
+$$
+where $z$ is the projection of our observed $y$ onto the ever active set.
+So, for a given ``ever inactive group'', this says that $X_g^Ty$ is in
+an intersection of $k$ balls which does not necessarily have a simple structure.
+When all the groups are of size 1, i.e. in the LASSO, the intersection 
+of $k$ balls is the intersection of $k$ intervals and we get a simplification.
+
+All this to say, that we are really only talking about $\ell_1$ here, and 
+next-$\lambda$ is essentially only practical BLAH
+}
 
@@
 \subsubsection{Testing Against the Upper Model}
+
 To avoid conditioning on the next model $M_k$, we could instead test $M_{k-1}$ against the alternative $M_{\infty}\setminus M_{k-1}$.
@@
 \WFcomment{Fixed t/u. $\psi$ is the log-proportionality constant, I think we need it. Maybe I'm misunderstanding your question?}
+\JTcomment{I guess I meant that I am assuming that
+$$
+\ell(\theta) = \theta'U(y) - \psi(\theta)
+$$
+but I don't see how I would conclude that from what is written.
+In principle, as written the $\psi$ above might be some other function but
+(44) and (45) imply that the objective function is of the form
+$$
+\ell(\theta) = \tilde{\psi}(\theta) - \theta'U(y)
+$$
+for some $\tilde{\psi}$.
+In the proof when you say $\mathbb{E}_{\hat{\theta}^{\lambda}}[U_j]$ I am assuming this is $e_j^T\nabla \psi (\hat{\theta}^{\lambda})$ soshould I conclude that $\tilde{\psi}=\psi$?
+}
 

### 2015-11-01 paper.tex
a few more small ones

@@
 
-  On $A$, we have $E_s(Y) \sub E$ for all $s\leq k$, meaning that the maximum in~\eqref{eq:forwardStepDef} is attained by some $j\in E$ at every step. So, for $i \leq k$, we have
+  On $A$, we have $E_s(Y) \sub E$ for all $s\leq k$, meaning that the maximum in~\eqref{eq:forwardStepDef} is attained by some $j\in E$ at every step. So, for $s \leq k$, we have
 \begin{align}
@@
 \end{proof}
+
+\JTcomment{Are we assuming that $U(Y;M)$ is minimal sufficient but not 
+necessarily minimal
+complete sufficient in the above proposition? The statement doesn't say
+what $U$ satisfies. I ask because in the corollary we do say something
+about completeness.  Our earlier assumption was just that $U$ always
+denotes a minimal sufficient statistic.
+}
 
@@
 
+\JTcomment{I don't really see the point of Figure 5. It is abundandtly clear
+that if feasible an IID sample is preferable over an MCMC sample. It is
+just as easy to make plots where accept/reject will take forever to get even one sample point. 
+}
+
 In detail, we employ a hybrid computational strategy; we use accept-reject sampling, and do so as long as we obtain at least 300 (say) samples in the first set of $B$ realizations ($75,000$ in our code)  This is an effective approach until we have reached a point in the model path where all predictors have little or no signal.
@@
 Note that $\ell_E^*$ is the convex dual of the convex function $\psi_E$, so $\ell_E^*$ is also convex.
-At stage $i<k$, the next variable to enter is the one that maximizes $\ell_{E_i\cup \{j\}}^*(U(Y))$.
+\JTcomment{It would be convex even if $\psi_E$ were not convex.
+No?
+I also think $U$ in the definition of $\ell^*_E$ can be just $U(Y)$.} At stage $i<k$, the next variable to enter is the one that maximizes $\ell_{E_i\cup \{j\}}^*(U(Y))$.
 

### 2015-11-06 paper.tex
changed color to bold

@@
 \caption[tab:diab]{\em Forward stepwise regression for the diabetes data: naive $p$-values, $p$-values from the saturated model, and our max-$t$ $p$-values.
-The red annotation indicates the model selected by ForwardStop with $\alpha=0.1$.
+The {\bf bold annotation} indicates the model selected by ForwardStop with $\alpha=0.1$.
 For the max-$t$ $p$-values, this stopping rule gives exact $\text{FDR}_{\text{model}}$ control at the 10\% level.}
@@
 The {\tt Simple} rule stops at the first time that a $p$-value exceeds $\alpha=0.05$, while {\tt Forward} and {\tt Strong} refers to the ForwardStop and StrongStrop rules of Section \ref{sec:orderedProposals}.
-Results with theoretical guarantees are \guarantee{highlighted}.}
+Results with theoretical guarantees are in \guarantee{bold highlighted}.}
 \label{tab:stopping05}
@@
 The {\tt Simple} rule stops at the first time that a $p$-value exceeds $\alpha=0.2$, while {\tt Forward} and {\tt Strong} refers to the ForwardStop and StrongStrop rules of Section \ref{sec:orderedProposals}.
-Results with theoretical guarantees are \guarantee{highlighted}.}
+Results with theoretical guarantees are \guarantee{bold highlighted}.}
 \label{tab:stopping20}

### 2015-11-06 code/python/error_rates.py

@@ -89,7 +89,7 @@ for results, outbase in zip(['../snr_5_alpha_%s.csv' % alpha for alpha in ['05',
             yield ' & '.join(val) + r' \\ '
 
     table = r'''
-\newcommand{\guarantee}[1]{{\color{blue} #1}}
+\newcommand{\guarantee}[1]{{\bf #1}}
 \begin{tabular}{|l|rrrrrr|}
  \hline
 {} &  $p_{\text{screen}}$ &  $\text{FWER}_{\text{mod}}$ &  $\text{FWER}_{\text{mod}} \vert \text{screen}$ &  $\text{FDR}_{\text{model}}$ &  $\text{FDR}_{\text{var}}$ &  $\text{S}_{\text{var}}$ \\ \hline

### 2015-11-06 code/python/error_rates.py
edits to tables and script generating tables

@@ -13,9 +13,9 @@ for results, outbase in zip(['../snr_5_alpha_%s.csv' % alpha for alpha in ['05',
     guarantees = []
 
     name_map = dict(zip(['nominalT', 'maxT', 'maxT_identify', 'maxT_unknown', 'saturated'],
-                        ['Nominal', 'MaxT', 'MaxT identify', 'MaxT unknown', 'Saturated']))
+                        ['Nominal', 'MaxZ', 'MaxZ identify', 'MaxT', 'Saturated']))
     for test, rule in product(['nominalT', 'maxT', 'maxT_identify', 'maxT_unknown', 'saturated'],
-                              ['simple', 'forward', 'strong']):
+                              ['simple', 'forward']):
         name = '_'.join([test, rule])
         screen = np.mean(getvar("screen"))
         FDR_var = np.mean(getvar("FDP_var"))

### 2015-11-06 paper.tex
maxT to maxZ changes, removing unused R script, shortening used R script to only produce figs

@@
   \includegraphics[width=1\textwidth]{figs/simulation_snr_5_alpha_05_null_false.pdf}
-  \caption{\em  CDFs of nominal (black), MaxT (red), MaxT with unknown $\sigma$ (green) and  saturated-model (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on testing a false null hypothesis at step $k$.
+  \caption{\em  CDFs of nominal (black), MaxZ (red), MaxT (green) and  saturated-model (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on testing a false null hypothesis at step $k$.
 The MaxT approaches are much more powerful than the saturated-model test.
@@
   \includegraphics[width=1\textwidth]{figs/simulation_snr_5_alpha_05_null_true.pdf}
-  \caption{\em  CDFs of nominal (black), MaxT (red), MaxT with unknown $\sigma$ (green) and  saturated-model (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on testing a true null hypothesis at step $k$.
+  \caption{\em  CDFs of nominal (black), MaxZ (red), MaxT (green) and  saturated-model (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on testing a true null hypothesis at step $k$.
 The nominal test is badly anti-conservative, while all of the other methods  show uniform $p$-values as desired.}
@@
   \includegraphics[width=1\textwidth]{figs/simulation_snr_5_alpha_05_noise_var.pdf}
-  \caption{\em CDFs of nominal (black), MaxT (red), MaxT with unknown $\sigma$ (green) and  saturated-model (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on the event that the variable added at step $k$ is a noise variable in the full model.
+  \caption{\em CDFs of nominal (black), MaxZ (red), MaxT $\sigma$ (green) and  saturated-model (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on the event that the variable added at step $k$ is a noise variable in the full model.
 Here, none of the methods produce uniform $p$-values.

### 2015-11-06 code/R/sparseSim.R

@@ -1,163 +1,0 @@
-library(xtable)
-
-simulation.data <- read.csv("../snr_5_alpha_05.csv")
-names(simulation.data)
-
-
-methods <- c(paste(rep(c("simple", "forward", "strong"),2),
-                   rep(c("MaxT","saturated"),each=3),
-                  sep="_"),
-             "knockoffs")
-error.types <- c("pscreen","FWER.mod","FDR.mod","FDR.var")
-error.rates <- matrix(NA,7,4,
-                      dimnames=list(methods,error.types))
-for(i in 1:6) {
-  R <- simulation.data[[paste0(methods[i],"_R")]]
-  k0 <- simulation.data[["completion_idx"]]
-  V.mod <- pmax(0,R - k0)
-  V.var <- simulation.data[[paste0(methods[i],"_V_var")]]
-  error.rates[i,"pscreen"] <- mean(R >= k0)
-  error.rates[i,"FWER.mod"] <- mean(V.mod > 1)
-  error.rates[i,"FDR.mod"] <- mean(V.mod / pmax(R,1))
-  error.rates[i,"FDR.var"] <- mean(V.var / pmax(R,1))
-}
-
-# What's wrong here?
-with(simulation.data, table(knockoff_R - knockoff_V))
-with(simulation.data, summary(simple_saturated_screen))
-with(simulation.data, summary(simple_saturated_V_model))
-with(simulation.data, which.max(simple_saturated_V_model))
-simulation.data[46, c("completion_index", "simple_saturated_R", "simple_saturated_V_model")]
-simulation.data$completion_index[46]
-simulation.data$simple_saturated_R[46]
-simulation.data$simple_saturated_V_model[46]
-
-
-error.rates["knockoffs", "pscreen"] <- with(simulation.data, mean(knockoff_R - knockoff_V == 7))
-#error.rates["knockoffs", "FDR.var"] <- with(simulation.data, mean(knockoff_V / pmax(knockoff_R,1)))
-o=simulation
-error.rates["knockoffs", "FDR.var"] <- with(simulation.data, mean(knockoff_V / pmax(knockoff_R,1)))
-
-## Fix this later...
-colnames(error.rates) <- c("$p_{\\text{screen}}", 
-                           "$\\text{FWER}_{\\text{model}}$",
[…]

### 2015-11-06 code/R/sparseSim2.R

@@ -1,206 +1,0 @@
-## plots p=40
-
-simulation.data <- read.csv("../snr_5_alpha_05.csv")
-names(simulation.data)
-
-p=40
-s=7
-
-
-pdf("../../figs/simulation_snr_5_alpha_05_signal_var.pdf",
-    width=8,height=5)
-par(xaxs="i",mar=c(2.1,3.1,3.1,0.7), mfrow=c(2,5))
-for(k in 1:10) {
-  which.signal <- which(simulation.data[[paste0("variable_selected_",k)]] <=s)
-  main <- paste("Step ",as.character(k),sep="")
-  plot(ecdf(simulation.data[[paste0("nominal_pvalue_",k)]][which.signal]),xlim=c(0,1),
-       main=main,xlab="",ylab="")
-del=runif(length(which.signal),0,.01)
-  plot(ecdf(simulation.data[[paste0("maxT_pvalue_",k)]][which.signal]+del),col="red",add=T)
-    plot(ecdf(simulation.data[[paste0("maxT_unknown_pvalue_",k)]][which.signal]+del),col="green",add=T)
-  plot(ecdf(simulation.data[[paste0("saturated_pvalue_",k)]][which.signal]),col="blue",add=T)
-  abline(0,1,lty=3,col="gray")
-if(k==1) legend (list(x=c(.3,1),y=c(.25,0)),c("Nominal","MaxT","Maxt/Unkn","Saturated"),col=c("black","red","green","blue"),lty=1,cex=.7)
-}
-dev.off()
-
-#null is true
-
-
-##
-pdf("../../figs/simulation_snr_5_alpha_05_null_true.pdf",
-    width=8,height=5)
-par(xaxs="i",mar=c(2.1,3.1,3.1,0.7), mfrow=c(2,5))
-
-
-for(k in (s+1):(s+10)) {
-
-which.null <- which(k> simulation.data$completion_idx+1)  # correction for zero-based
-  main <- paste("Step ",as.character(k),sep="")
-  plot(ecdf(simulation.data[[paste0("nominal_pvalue_",k)]][which.null]),xlim=c(0,1),
-       main=main,xlab="",ylab="")
-plot(ecdf(simulation.data[[paste0("maxT_pvalue_",k)]][which.null]),col="red",add=T)
-  plot(ecdf(simulation.data[[paste0("maxT_unknown_pvalue_",k)]][which.null]),col="green",add=T)
[…]

### 2015-11-06 code/R/sparseSim_figs.R

@@ -1,0 +1,87 @@
+## plots p=40
+
+simulation.data <- read.csv("../snr_5_alpha_05.csv")
+names(simulation.data)
+
+p=40
+s=7
+
+
+pdf("../../figs/simulation_snr_5_alpha_05_signal_var.pdf",
+    width=8,height=5)
+par(xaxs="i",mar=c(2.1,3.1,3.1,0.7), mfrow=c(2,5))
+for(k in 1:10) {
+  which.signal <- which(simulation.data[[paste0("variable_selected_",k)]] <=s)
+  main <- paste("Step ",as.character(k),sep="")
+  plot(ecdf(simulation.data[[paste0("nominal_pvalue_",k)]][which.signal]),xlim=c(0,1),
+       main=main,xlab="",ylab="")
+del=runif(length(which.signal),0,.01)
+  plot(ecdf(simulation.data[[paste0("maxT_pvalue_",k)]][which.signal]+del),col="red",add=T)
+    plot(ecdf(simulation.data[[paste0("maxT_unknown_pvalue_",k)]][which.signal]+del),col="green",add=T)
+  plot(ecdf(simulation.data[[paste0("saturated_pvalue_",k)]][which.signal]),col="blue",add=T)
+  abline(0,1,lty=3,col="gray")
+if(k==1) legend (list(x=c(.3,1),y=c(.25,0)),c("Nominal","MaxZ","MaxT","Saturated"),col=c("black","red","green","blue"),lty=1,cex=.7)
+}
+dev.off()
+
+#null is true
+
+
+##
+pdf("../../figs/simulation_snr_5_alpha_05_null_true.pdf",
+    width=8,height=5)
+par(xaxs="i",mar=c(2.1,3.1,3.1,0.7), mfrow=c(2,5))
+
+
+for(k in (s+1):(s+10)) {
+
+which.null <- which(k> simulation.data$completion_idx+1)  # correction for zero-based
+  main <- paste("Step ",as.character(k),sep="")
+  plot(ecdf(simulation.data[[paste0("nominal_pvalue_",k)]][which.null]),xlim=c(0,1),
+       main=main,xlab="",ylab="")
+plot(ecdf(simulation.data[[paste0("maxT_pvalue_",k)]][which.null]),col="red",add=T)
+  plot(ecdf(simulation.data[[paste0("maxT_unknown_pvalue_",k)]][which.null]),col="green",add=T)
[…]

### 2015-11-06 code/R/sparseSim_figs.R
4 panel figures

@@ -8,9 +8,10 @@ s=7
 
 
 pdf("../../figs/simulation_snr_5_alpha_05_signal_var.pdf",
-    width=8,height=5)
-par(xaxs="i",mar=c(2.1,3.1,3.1,0.7), mfrow=c(2,5))
-for(k in 1:10) {
+    width=8,height=2.5)
+par(xaxs="i",mar=c(2.1,3.1,3.1,0.7), mfrow=c(1,4))
+
+for(k in 4:7) {
   which.signal <- which(simulation.data[[paste0("variable_selected_",k)]] <=s)
   main <- paste("Step ",as.character(k),sep="")
   plot(ecdf(simulation.data[[paste0("nominal_pvalue_",k)]][which.signal]),xlim=c(0,1),
@@ -20,7 +21,7 @@ del=runif(length(which.signal),0,.01)
     plot(ecdf(simulation.data[[paste0("maxT_unknown_pvalue_",k)]][which.signal]+del),col="green",add=T)
   plot(ecdf(simulation.data[[paste0("saturated_pvalue_",k)]][which.signal]),col="blue",add=T)
   abline(0,1,lty=3,col="gray")
-if(k==1) legend (list(x=c(.3,1),y=c(.25,0)),c("Nominal","MaxZ","MaxT","Saturated"),col=c("black","red","green","blue"),lty=1,cex=.7)
+if(k==4) legend (list(x=c(.3,1),y=c(.25,0)),c("Nominal","MaxZ","MaxT","Saturated"),col=c("black","red","green","blue"),lty=1,cex=.7)
 }
 dev.off()
 
@@ -29,11 +30,11 @@ dev.off()
 
 ##
 pdf("../../figs/simulation_snr_5_alpha_05_null_true.pdf",
-    width=8,height=5)
-par(xaxs="i",mar=c(2.1,3.1,3.1,0.7), mfrow=c(2,5))
+    width=8,height=2.5)
+par(xaxs="i",mar=c(2.1,3.1,3.1,0.7), mfrow=c(1,4))
 
 
-for(k in (s+1):(s+10)) {
+for(k in (s+1):(s+4)) {
 
 which.null <- which(k> simulation.data$completion_idx+1)  # correction for zero-based
   main <- paste("Step ",as.character(k),sep="")
@@ -49,10 +50,10 @@ dev.off()
 
 
 pdf("../../figs/simulation_snr_5_alpha_05_null_false.pdf",
-    width=8,height=5)
[…]

### 2015-11-06 code/python/maxT.R
unused scripts

@@ -1,110 +1,0 @@
-pdf('maxT_view.pdf')
-D1 = read.table('test1.csv', sep=',', header=TRUE)
-D2 = read.table('test2.csv', sep=',', header=TRUE)
-D3 = read.table('test3.csv', sep=',', header=TRUE)
-D4 = read.table('test4.csv', sep=',', header=TRUE)
-D = rbind(D1,D2,D3,D4)
-write.table(D, "test.csv", row.names=FALSE, col.names=TRUE, sep=',')
-
-par(mfrow=c(2,2))
-
-# mostly signal
-
-plot(ecdf(D$maxT_pvalue_4), main='Step 4', xlim=c(0,1))
-plot(ecdf(D$maxT_identify_pvalue_4), add=TRUE, col='blue')
-plot(ecdf(D$maxT_identify_unknown_pvalue_4), add=TRUE, col='yellow')
-plot(ecdf(D$maxT_unknown_pvalue_4), add=TRUE, col='green')
-plot(ecdf(D$saturated_pvalue_4), add=TRUE, col='red')
-
-plot(ecdf(D$maxT_pvalue_5), main='Step 5', xlim=c(0,1))
-plot(ecdf(D$maxT_identify_pvalue_5), add=TRUE, col='blue')
-plot(ecdf(D$saturated_pvalue_5), add=TRUE, col='red')
-plot(ecdf(D$maxT_identify_unknown_pvalue_5), add=TRUE, col='yellow')
-plot(ecdf(D$maxT_unknown_pvalue_5), add=TRUE, col='green')
-
-plot(ecdf(D$maxT_pvalue_6), main='Step 6', xlim=c(0,1))
-plot(ecdf(D$maxT_identify_pvalue_6), add=TRUE, col='blue')
-plot(ecdf(D$saturated_pvalue_6), add=TRUE, col='red')
-plot(ecdf(D$maxT_identify_unknown_pvalue_6), add=TRUE, col='yellow')
-plot(ecdf(D$maxT_unknown_pvalue_6), add=TRUE, col='green')
-
-plot(ecdf(D$maxT_pvalue_7), main='Step 7', xlim=c(0,1))
-plot(ecdf(D$maxT_identify_pvalue_7), add=TRUE, col='blue')
-plot(ecdf(D$saturated_pvalue_7), add=TRUE, col='red')
[…]

### 2015-11-06 code/python/sqrt_lasso_rob.py

@@ -1,62 +1,0 @@
-#!/software/free/statdept/bin/python
-
-# These lines are specific to miller so that the python used is my local one
-
-# Usage example:
-# ./fwdstep_rob.py --X=X.csv --Y=Y.csv --active=active.csv --outfile=results.csv --sigma=5. --outfile=results.csv
-
-import sys, os
-
-# add the current directory to the path so we can find the 
-# sequential-selection functions
-
-sys.path.append(os.path.abspath(os.path.dirname("__file__")))
-
-# The rest of the script should run as long as selection is installed
-# and the selective-selection python directory is on the path
-
-from argparse import ArgumentParser
-
-import numpy as np, pandas as pd
-from selection.algorithms.sqrt_lasso import sqrt_lasso, choose_lambda
-
-def main():
-    parser = ArgumentParser(
-        description= '''
-        Run forward step a full min(n, p) steps, computing
-        selected, saturated and nominal pvalues.
-
-        ''')
-    parser.add_argument('--X',
-                        help='Design matrix (CSV filename). REQUIRED')
-    parser.add_argument('--Y',
-                        help='Response (CSV filename). REQUIRED')
-    parser.add_argument('--lam',
-                        help='What proportion of population lambda should we use?', type=float, default=0.5)
-    parser.add_argument('--outfile',
-                        help='Where to store output.')
-
-    try: 			
-        args = parser.parse_args()
-        X = np.loadtxt(args.X, delimiter=',')
-        Y = np.loadtxt(args.Y, delimiter=',')
-    except:
-        parser.print_help()
-        return 
-
-    results = {}
-
-    L = args.lam * choose_lambda(X)
-    SQL = sqrt_lasso(Y, X, args.lam)
-    SQL.fit(min_its=100)
-
-    V = SQL.active_gaussian_pval
-    V['variable'] += 1
-    if args.outfile:
-        pd.DataFrame(V).to_csv(args.outfile, index=False)
-
-        
-if __name__ == "__main__":
-    main()
-
-                

### 2015-11-06 code/python/sqrtlasso_rob.py

@@ -1,62 +1,0 @@
-#!/software/free/statdept/bin/python
-
-import sys
-sys.path.append('/home/jtaylo/.local/lib/python2.7/site-packages')
-
-# Usage example:
-# ./sqrtlasso_rob.py --X=X.csv --Y=Y.csv --lam=0.9 --outfile=results.csv
-
-from argparse import ArgumentParser
-
-import numpy as np, pandas as pd
-from selection.algorithms.sqrt_lasso import sqrt_lasso, choose_lambda
-
-def main():
-    parser = ArgumentParser(
-        description= '''
-        Run LASSO at fixed lambda and save p-values and intervals.
-
-        Objective is
-
-        \|Y-X\beta\|_2 + \lambda \|\beta\|_1
-        
-        ''')
-    parser.add_argument('--X',
-                        help='Design matrix (CSV).')
-    parser.add_argument('--Y',
-                        help='Response.')
-    parser.add_argument('--lam',
-                        help='Multiple of theoretical lambda for sqrt-LASSO.', default=0.9, type=float)
-    parser.add_argument('--outfile',
-                        help='Where to results.')
-
-    args = parser.parse_args()
-    X = np.loadtxt(args.X, delimiter=',')
-    Y = np.loadtxt(args.Y, delimiter=',')
-    
-    L = sqrt_lasso(Y, X, args.lam * choose_lambda(X))
-    L.fit(min_its=150, tol=1.e-10)
-    
-    intervals = []
-    pvalues = []
-    if L.active is not None and args.outfile:
-        intervals = L.active_gaussian_intervals
-        pvals = L.active_gaussian_pval
-
-        if np.any(intervals['variable'] != pvals['variable']):
-            raise ValueError('variable lists are different!')
-
-        df = pd.DataFrame({'variable':pvals['variable'],
-                           'pvalue':pvals['pvalue'],
-                           'lower':intervals['lower'],
-                           'upper':intervals['upper']})
-        df = df.reindex(columns=['variable', 'pvalue', 'lower', 'upper'])
-        df.to_csv(args.outfile, index=False)
-    else:
-        if not args.outfile:
-            sys.stderr.write('Active set was empty!\n')
-        
-if __name__ == "__main__":
-    main()
-
-                

### 2015-11-06 code/python/data_carving.py
unused files

@@ -1,77 +1,0 @@
-import numpy as np, os
-import matplotlib.pyplot as plt
-from matplotlib.mlab import rec2csv
-
-split_vals = ([0.5, 0.55, 0.6, 0.65, 0.7, 0.75, 0.8, 0.85, 0.875, 0.9,
-              0.925, 0.95, 0.96, 0.97, 0.98, 0.99, 1.])
-vals = [('n', 100),
-        ('p', 50),
-        ('s', 7),
-        ('rho', 0.3),
-        ('snr', 7.)
-        ]
-opts = dict(vals)
-
-df = np.inf
-if df == np.inf:
-    dname = 'gaussian'
-else:
-    dname = 'df_%d' % df
-
-def summary(save=True):
-
-    results = []
-
-    try:
-        screen = np.load('%s/screening.npy' % dname)
-
-        for split_frac in sorted(np.unique(split_vals)):
-            fname = '%s/results_split_%0.2f.npy' % (dname, split_frac)
-            data = np.load(fname)
-            null_carve = np.array([d['pval'] for d in data if d['method'] == 'carve' 
-                                   and d['null'] == True])
-            null_split = np.array([d['pval'] for d in data if d['method'] == 'split' 
-                                   and d['null'] == True])
-            alt_carve = np.array([d['pval'] for d in data if d['method'] == 'carve' 
-                                   and d['null'] == False])
-            alt_split = np.array([d['pval'] for d in data if d['method'] == 'split' 
-                                   and d['null'] == False])
-
-            power_carve = np.mean(alt_carve < 0.05)
-            power_split = np.mean(alt_split < 0.05)
-            level_carve = np.mean(null_carve < 0.05)
-            level_split = np.mean(null_split < 0.05)
-
-            p_screen = 1. / np.mean(screen[screen['split'] == split_frac]['counter'])
-            result = (split_frac, level_carve, level_split, power_carve, power_split, p_screen)
-            results.append(result)
-
-            print split_frac
-
-        R = np.array(results, np.dtype([ \
-                    ('split', np.float),
-                    ('level_carve', np.float),
-                    ('level_split', np.float),
-                    ('power_carve', np.float),
-                    ('power_split', np.float),
-                    ('p_screen', np.float)]))
-
-        if save:
-            np.save('%s/summary.npy' % dname, R)
-            rec2csv(R, '%s/summary.csv' % dname)
[…]

### 2015-11-06 code/python/data_carving_batch.py

@@ -1,84 +1,0 @@
-import numpy as np, csv, hashlib, os.path
-import uuid
-from selection.algorithms.tests.test_forward_step import test_data_carving_IC
-from data_carving import split_vals, vals, df, dname
-
-# how many points do we want for each fraction
-
-min_sample_size = 100000
-
-vals = vals + [('split_frac', 0.9),
-               ('ndraw', 6000),
-               ('burnin', 5000)]
-
-dtype = np.dtype([('n', np.int), 
-                  ('p', np.int), 
-                  ('s', np.int), 
-                  ('rho', np.float), 
-                  ('snr', np.float), 
-                  ('split_frac', np.float), 
-                  ('ndraw', np.int), 
-                  ('burnin', np.int), 
-                  ('method', 'S5'), 
-                  ('null', np.bool), 
-                  ('pval', np.float),
-                  ('uuid', 'S40')])
-
-for i in range(5000):
-    for split_frac in split_vals[::-1]:
-        opts = dict(vals)
-        opts['split_frac'] = split_frac
-        identifier = str(uuid.uuid1())
-        fname = '%s/results_split_%0.2f.npy' % (dname, split_frac)
-        opts['df'] = np.inf # degrees of freedom for noise
-
-        test = not os.path.exists(fname)
-        if not test:
-            prev_results = np.load(fname)
-            if split_frac not in [0.3, 0.4, 0.5]:
-                test = prev_results.shape[0] < min_sample_size
-            elif split_frac in [0.3, 0.4]:
-                test = prev_results.shape[0] < 5000
-            else:
-                test = prev_results.shape[0] < 10000
-
-        if test:
-            (null_carve, 
-             null_split, 
-             alt_carve,
-             alt_split,
-             counter) = test_data_carving_IC(**opts)[:-2]
-
-            params = [v for _, v in vals]
-            results = []
-            if os.path.exists('%s/screening.npy' % dname):
-                prev_results = np.load('%s/screening.npy' % dname)
-                screening = np.empty(prev_results.shape[0]+1, 
-                                      prev_results.dtype)
-                screening[:-1] = prev_results
-                screening[-1] = (split_frac, counter, identifier)
-                np.save('%s/screening.npy' % dname, screening)
-            else:
-                dtype_screen = np.dtype([('split', np.float),
-                                         ('counter', np.float),
-                                         ('uuid', 'S40')])
-                screening = np.array([(split_frac, counter, identifier)],
-                                     dtype_screen)
[…]

### 2015-11-06 code/python/fwdstep_rob.py

@@ -1,84 +1,0 @@
-#!/software/free/statdept/bin/python
-
-# These lines are specific to miller so that the python used is my local one
-
-# Usage example:
-# ./fwdstep_rob.py --X=X.csv --Y=Y.csv --active=active.csv --outfile=results.csv --sigma=5. --outfile=results.csv
-
-import sys, os
-sys.path.append('/home/jtaylo/.local/lib/python2.7/site-packages')
-
-# add the current directory to the path so we can find the 
-# sequential-selection functions
-
-sys.path.append(os.path.abspath(os.path.dirname("__file__")))
-
-# The rest of the script should run as long as selection is installed
-# and the selective-selection python directory is on the path
-
-from argparse import ArgumentParser
-
-import numpy as np, pandas as pd
-from compare_methods import run
-
-def main():
-    parser = ArgumentParser(
-        description= '''
-        Run forward step a full min(n, p) steps, computing
-        selected, saturated and nominal pvalues.
-
-        ''')
-    parser.add_argument('--X',
-                        help='Design matrix (CSV filename). REQUIRED')
-    parser.add_argument('--Y',
-                        help='Response (CSV filename). REQUIRED')
-    parser.add_argument('--sigma',
-                        help='Noise standard deviation. REQUIRED', type=float)
-    parser.add_argument('--maxstep',
-                        help='How many steps. Defaults to min(n,p)', type=int)
-    parser.add_argument('--active',
-                        help='Active set (CSV filename). Necessary ' + 
-                        'to compute completion index. REQUIRED. ASSUMES 1-BASED INDEXING!')    
-    parser.add_argument('--outfile',
-                        help='Where to store output.')
-    parser.add_argument('--burnin',
-                        help='How many burnin steps for sampling.', default=2000, type=int)
-    parser.add_argument('--ndraw',
-                        help='How many sampling steps.', default=8000, type=int)
-
-    try: 			
-        args = parser.parse_args()
-        X = np.loadtxt(args.X, delimiter=',')
-        Y = np.loadtxt(args.Y, delimiter=',')
-        # ASSUMING 1-BASED INDEXING!!!!
-        active = np.loadtxt(args.active, delimiter=',').astype(np.int) - 1
-    except:
-        parser.print_help()
-        return 
-
-    if args.sigma is None:
-        raise ValueError('sigma is needed to compute saturated p-values!')
-
-    full_results = {}
-
[…]

### 2015-11-09 paper.tex
4 panel figures
paper.tex

@@
 \usepackage{amsmath,amssymb,amsthm,bm,enumerate,mathrsfs,mathtools}
+\usepackage{thmtools, thm-restate}
 \usepackage{latexsym,color,verbatim,multirow}
@@
 Extending the selected-model tests of \citet{fithian2014optimal}, we construct $p$-values for each step in the path, accounting for the fact that the model path is determined adaptively using the data.
-In the case of linear regression, our $p$-values improve on the power of the spacings test of \citet{taylor2014exact}, sometimes dramatically.
-
-To select a model, we can feed the resulting $p$-values as inputs into sequential stopping rules such as those proposed by \citet{gsell2013sequential} and \citet{li2015accumulation}, achieving control of the familywise error rate or false discovery rate as desired.
-Some stopping rules require the null $p$-values to be independent of each other and of the non-null $p$-values, a condition not satisfied by the saturated-model $p$-values of \citet{taylor2014exact}.
+In the case of linear regression, we propose two specific tests, the max-$t$ test for forward stepwise regression (generalizing a proposal of \citet{buja2014}), and the next-$\lambda$ test for the lasso.
+These tests improve on the power of the spacings test of \citet{taylor2014exact}, sometimes dramatically.
+In addition, our framework extends beyond linear regression to a much more general class of parametric and nonparametric model selection problems.
+
+To select a model, we can feed our single-step $p$-values as inputs into sequential stopping rules such as those proposed by \citet{gsell2013sequential} and \citet{li2015accumulation}, achieving control of the familywise error rate (FWER) or false discovery rate (FDR) as desired.
+The FDR-controlling rules require the null $p$-values to be independent of each other and of the non-null $p$-values, a condition not satisfied by the saturated-model $p$-values of \citet{taylor2014exact}.
 We derive intuitive and general conditions for independence and show that our proposed constructions yield independent $p$-values.
@@
 
-To select a model, we compute a set of sequential $p$-values at each step of the model path, and then feed them into a stopping rule that is guaranteed to control the FDR, familywise error rate (FWER), or a similar quantity.
+To select a model, we compute a set of sequential $p$-values at each step of the model path, and then feed them into a stopping rule that is guaranteed to control the false discovery rate (FDR) \citep{benjamini1995controlling}, familywise error rate (FWER), or a similar quantity.
 Recently \citet{gsell2013sequential} and \citet{li2015accumulation} proposed sequential stopping rules of this kind.
-Both sets of rules require that once we have reached the first correct model in our path, the $p$-values in subsequent steps are uniform (or conservative),
-and independent.
-In this paper we develop a theoretical framework for constructing sequential $p$-values that are both uniform and independent under the correct model, and show how to construct them in forward stepwise regression, LAR and the ever-active path in $\ell_1$-regularized models.
+Both sets of rules require that once we have reached the first correct model in our path, the $p$-values in subsequent steps are uniform and independent.
+In this paper we develop a theoretical framework for constructing sequential $p$-values satisfying these properties, and show how to construct them in many settings.
 
@@
 
-Let $E \sub \{1, \ldots, d\}$ denote the current {\em active set}, the set of predictor variables already selected, and let $\RSS(E)$ denote the residual sum of squares for the corresponding regression model.
-For $j\notin E$ let $t_{j,E}(Y)$ denote the multivariate $t$-statistic of variable $j$, adjusting for the active variables.
+For design matrix $X\in \R^{n\times p}$ and response vector $Y\in \R^n$, let $E \sub \{1, \ldots, d\}$ denote the current {\em active set}, the set of predictor variables already selected, and let $\RSS(E)$ denote the residual sum of squares for the corresponding regression model.
+For $j\notin E$ let $t_{j,E}(Y)$ denote the multivariate $t$-statistic of variable $j$, adjusting for the active variables and an intercept term.
 Using the standard result that
 \begin{equation}
-t_{j,E}^2 = (n-|E|-1) \frac{\RSS(E) - \RSS(E \cup \{j\})}{\RSS(E \cup \{j\})},
+t_{j,E}^2 = (n-|E|-2) \frac{\RSS(E) - \RSS(E \cup \{j\})}{\RSS(E \cup \{j\})},
 \end{equation}
@@
 
-The nominal $p$-values in Table~\ref{tab:diab} are computed by comparing $t_E^*$ to the $t_{n-|E|-1}$-distribution, which would be the correct distribution if the sequence of models were selected ahead of time, before observing the data.
+The nominal $p$-values in Table~\ref{tab:diab} are computed by comparing $t_E^*$ to the $t$ distribution with $n-|E|-2$ degrees of freedom, which would be the correct distribution if the sequence of models were selected before observing the data.
 Because the model sequence is in fact selected adaptively to maximize the value of $t_E^*$, this method is highly anti-conservative.
@@
 Because $U_E=\left(X_E'Y,\, \|Y\|^2\right)$ is the complete sufficient statistic for $M(E)$, we can sample from the conditional null distribution of $t_E^*$ given $U_E$ and the current active set $E$, using the {\em selected-model test} framework of \citet{fithian2014optimal}.
-In step one, because $E=\{1_n\}$ is fixed and $t_E^*$ is independent of $U_E$, the conditional test reduces to the max-$t$ test proposed in \citet{buja2014}.
-In later steps, $t_E^*$ and $U_E$ are not conditionally independent of $E$, but we can estimate the null conditional distribution through Monte Carlo sampling.
-
-The saturated-model $p$-values also account for selection, but they rely on somewhat different assumptions, condition on more information, and test a slightly less restrictive null hypothesis.
+In step one, because $E=\emptyset$ is fixed and $t_E^*$ is independent of $U_E$, the conditional test reduces to the max-$t$ test proposed in \citet{buja2014}.
+In later steps, $t_E^*$ and $U_E$ are not conditionally independent given $E$, but we can approximate the null conditional distribution through Monte Carlo sampling.
+
+The saturated-model $p$-values also account for selection, but they rely on somewhat different assumptions, condition on more information, and test a slightly different null hypothesis.
 We discuss these distinctions in detail in Section~\ref{sec:whichnull}.
@@
 Both the max-$t$ and saturated-model $p$-values are exactly uniform under the null, but the max-$t$ $p$-values appear to be more powerful under the alternative.
-As we will discuss in Section~\ref{sec:selective-reg}, selected-model tests such as the max-$t$ can be much more powerful than saturated-model tests in early steps of the path when multiple strong variables are compete with each other to enter the model first.
+As we will discuss in Section~\ref{sec:selective-reg}, selected-model tests such as the max-$t$ can be much more powerful than saturated-model tests in early steps of the path when multiple strong variables are competing with each other to enter the model first.
 Using the max-$t$ $p$-values, ForwardStop selects a model of size 8, compared to size 3 when using the saturated-model $p$-values.
 
-Because the max-$t$ null $p$-values are independent, while the saturated-model $p$-values are not, ForwardStop is guaranteed to control FDR when using the former but not the latter.
+Because the null max-$t$ $p$-values are independent, while the saturated-model $p$-values are not, ForwardStop is guaranteed to control FDR when using the former but not the latter.
 In Section~\ref{sec:pValsIndep} we discuss intuitive sufficient conditions for independence and show that the max-$t$ and many other selected-model tests satisfy these conditions.
 
-\begin{comment}
-The latter two sets of $p$-values test somewhat different hypotheses:  the $p$-values from the saturated model test whether the partial regression of the predictor just added is zero after regressing out the variables already in the model, what we call the ``incremental null''
-In contrast, the max-$t$ $p$-values test whether the current model is adequate; that is, whether all predictors not in the model  have coefficients equal to zero, the ``complete null''.
-(The issue of what hypothesis to test is discussed in section \ref{sec:whichnull}.)
-\end{comment}
-
 \subsection{Outline}
@@
 
-In comparison to their saturated-model counterparts, we will see that the selected-model tests proposed here have three main advantages: power, independence (in some cases), and generalizability beyond linear regression.
+In comparison to their saturated-model counterparts, we will see that the selected-model tests proposed here have three main advantages: power, independence (in many cases), and generalizability beyond linear regression.
 These advantages do not come entirely for free, as we will see: the selected-model methods test a more restrictive null hypothesis than the saturated-model methods.
-In addition, the selected-model test require accept-reject or MCMC sampling, while the saturated-model tests are available in closed form.
+In addition, the selected-model tests require accept-reject or MCMC sampling, while the saturated-model tests can be carried out in closed form.
 
@@
 In Section~\ref{sec:computation} we discuss strategies for sampling from the null distribution for our tests, and prove that in forward stepwise regression and $\ell_1$-regularized likelihood methods, the number of constraints in the sampling problem is never more than twice the number of variables.
-We provide a simulation study in Section~\ref{sec:sparseReg} and briefly discuss applications of our framework to principal components analysis in Section~\ref{sec:pca}.
-The paper ends with a Discussion.
+We provide a simulation study in Section~\ref{sec:sparseReg} and briefly discuss applications of our framework to decision trees, nonparametric changepoint detection, and principal components analysis in Section~\ref{sec:further-examples}.
+The paper ends with a discussion.
 
@@
 By {\em model}, we mean a family of candidate probability distributions for $Y$. For example in  linear regression, a model specifies  the set of
-predictors with nonzero coefficients (but not the  values of their coefficients).
+predictors allowed to have nonzero coefficients (but not the  values of their coefficients).
 Without loss of generality, we can take $M_\infty$ as the union of all models under consideration.
@@
 Given data $Y$ and model path $M_{0:d}$, we set ourselves the formal goal of selecting the simplest correct model: the smallest $k$ for which $F\in M_k(Y)$.
-Of course, in most real data problems all of the $M_k(Y)$, as well as all other models under consideration, are incorrect.
+Of course, in most real data problems all of the $M_k(Y)$ in the path, as well as all other models under consideration, are incorrect.
 In more informal terms, then, we seek the simplest model in the path that is not refuted by the available data.
-
-\JTcomment{ Not sure if stopping rule is a great name.
-It suggests it is a ``stopping time'' with respect to some filtration (which it often is).
-But it doesn't have to be.
-Or maybe we should say we call it stopping because we use the property of it being a stopping time?
-
-Another notation quibble for $M(E)$ I kind of prefer the notation $(\beta_E, \sigma^2_E)$ instead of $(\beta, \sigma^2)$. It emphasizes that these are parameters of different models.}
 
@@
 Because $\hk$ is the number of models we do reject, and $k_0$ is the number we should reject, the number of type I errors is $V=(\hk-k_0)_+$, while the number of type II errors is $(k_0-\hk)_+$.
-Depending on the scientific context, we might want to control the familywise error rate (FWER) $\P(V>0)$, the false discovery rate (FDR) $\E[V/\hk; \hk>0]$ \citep{benjamini1995controlling}, or another error rate defined by the expectation of some loss function $g(\hk, k_0)$:
+Depending on the scientific context, we might want to control the FWER $\P(V>0)$, the FDR $\E[V/\hk; \hk>0]$ , or another error rate defined by the expectation of some loss function $g(\hk, k_0)$:
 \begin{equation}\label{eq:errRate}
@@
 
-Many familiar path algorithms including lasso, LARS, and forward stepwise regression, are methods for adaptively selecting a set of predictors in linear regression models where we observe a random response $Y\in \R^n$ as well as a fixed design matrix $X\in \R^{n \times d}$, whose columns correspond to candidate predictors.
-For each possible active set $E \sub \{1,\ldots,d\}$, there is a corresponding candidate model
-\[
-M(E):\; Y \sim \cN( X_E\beta, \sigma^2 I_n),
+Many familiar path algorithms including lasso, LARS, and forward stepwise regression, are methods for adaptively selecting a set of predictors in linear regression models where we observe a random response $Y\in \R^n$ as well as a fixed design matrix $X\in \R^{n \times p}$, whose columns correspond to candidate predictors.
+For each possible active set $E \sub \{1,\ldots,p\}$, there is a corresponding candidate model
+\[
+M(E):\; Y \sim \cN( X_E\beta_E, \sigma^2 I_n),
 \]
@@
 
-We define the {\em selective greedy likelihood ratio test} as the test that rejects for large $G_k$, based on the law
+The {\em selective greedy likelihood ratio test} rejects for large $G_k$, based on the law
 \begin{equation}
@@
 
-Because the likelihood in linear regression is a monotone decreasing function of the residual sum of squares, the max-$t$ test is equivalent to the greedy likelihood test.
+Because the likelihood in linear regression is a monotone decreasing function of the residual sum of squares, the max-$t$ test is equivalent to the greedy likelihood ratio test.
+The counterpart of the max-$t$ test in linear regression with known $\sigma^2$ is the max-$z$ test, which differs only in replacing the $t$-statistics $t_{j,E}$ with their corresponding $z$-statistics.
+The max-$z$ test is also equivalent to the selective greedy likelihood ratio test.
 
 For simplicity, we have made two assumptions: that only one variable is added at each step, and that the set of candidate variables we choose from is the same in each step.
-It is relatively straightforward to relax either assumption, as we discuss briefly in Section~\ref{sec:discussion}.
+It is relatively straightforward to relax either assumption, but we do not pursue such generalizations here.
 
@@
 In the case of lasso-penalized regression, we could take $P_\lambda = \lambda\|\beta\|_1$ for $\lambda\in (0,\infty]$.
-We assume $P_\lambda(\beta)$  has the property  that the model becomes less constrained as $\lambda$ decreases, that is,
- if $\lambda_2<\lambda_1$ then $\{\beta:P_{\lambda_2}(\beta)\} \supset \{\beta:P_{\lambda_1}(\beta)\}$.
-\RTcomment{Will: is this correct?} \JTcomment{Yes, this looks a little funny to me too.
-Not all penalties produce
+
+\JTcomment{Yes, this looks a little funny to me too. Not all penalties produce
 sparsity (for example, the generalized LASSO). So, I think this section is really thinking 
@@
 
-We then define
-\begin{align}
-  \hat\theta^{\lambda}(Y) &= 
-  \argmin_{\theta\in\Theta} -\ell(\theta; Y) + P_\lambda(\theta) \\
-  \tE_\lambda(Y) &= \left\{j:\; \hat\theta_j^\nu(Y) \neq 0 
+Assuming for simplicity that there is a unique solution to each penalized problem of the form
+\begin{equation}\label{eq:pen-lik}
+  \hat\theta^{\lambda}(Y) =
+  \argmin_{\theta\in\Theta} -\ell(\theta; Y) + P_\lambda(\theta),
+\end{equation}
+we define the ever-active set for $\lambda\in\Lambda$ as
+\begin{equation}
+  \tE_\lambda(Y) = \left\{j:\; \hat\theta_j^\nu(Y) \neq 0 
     \text{ for any } \nu\in \Lambda, \nu \geq \lambda \right\}
-\end{align}
-It may be impossible or inconvenient to compute $\hat\theta^\lambda(Y)$ for every $\lambda\in(0,\infty]$. If so, we can instead take $\Lambda$ to be a grid of finitely many positive values.
-
-Note that the active sets $\tE_\lambda$ are nested by construction.
+\end{equation}
+It may be impossible or inconvenient to compute $\hat\theta^\lambda(Y)$ for every $\lambda\in(0,\infty]$. If so, we can instead take $\Lambda$ to be a grid of finitely many values.
+
+Note that the ever-active sets $\tE_\lambda$ are nested by construction.
 In addition, we will assume that $|\tE_\lambda|<\infty$ for every $\lambda\in\Lambda$.
@@
 \lambda_0=\sup \Lambda, \quad \text{ and } \quad 
-E_0 = \bigcap_{\lambda\in \Lambda} \tE^\lambda,
+E_0 = \bigcap_{\lambda\in \Lambda} \tE_\lambda,
 \]
@@
 
+\paragraph{Remark} In its usual formulation, the lasso coefficients for linear regression minimize a penalized RSS criterion:
+\[
+\hat\beta_{\text{lasso}}^\lambda = \frac{1}{2}\|Y-X\beta\|_2^2 + \lambda\|\beta\|_1.
+\]
+
+This is slightly different from the penalized log-likelihood criterion of \eqref{eq:pen-lik}, which would minimize
+\begin{equation}\label{eq:lasso-beta-sigma}
+  -\ell(\beta, \sigma;\,Y) + P_\lambda(\beta)=
+  \frac{n}{2}\log(2\pi\sigma^2) + \frac{1}{2\sigma^2}\|Y-X\beta\|_2^2 
+  + \lambda\|\beta\|_1.
+\end{equation}
+
+If $\sigma^2$ is known, the two methods are equivalent up to rescaling $\lambda$.
+If $\sigma^2$ is unknown, then partially minimizing~\eqref{eq:lasso-beta-sigma} over $\sigma^2$ gives $\hat\sigma^2(\beta) = n^{-1}\|Y-X\beta\|_2^2$, leading to profile log-likelihood
+\[
+-\ell^*(\beta;\,Y) \;\triangleq\; -\ell(\beta, \hat\sigma^2(\beta); \,Y)
+\;=\;  \frac{n}{2}
+\left(\log\left(\frac{2\pi\|Y-X\beta\|_2^2}{n}\right)+1\right)
+\]
+
+As a result the lasso coefficient vector $\hat\beta_{\text{lasso}}^\lambda$ minimizes
+\[
+\frac{n}{2\pi e}\exp\left\{-2\ell^*(\beta;\,Y)/n\right\} + \lambda\|\beta\|_1,
+\]
+while the regularized log-likelihood estimator $\hat\beta^\lambda$ minimizes $-\ell^*(\beta;\,Y)+\lambda\|\beta\|_1$.
+If $\Lambda=(0,\infty)$ then both methods result in exactly the same ever-active path.
+Otherwise, at the cost of some notational bloat, we could generalize \eqref{eq:pen-lik} by replacing $-\ell(\cdot)$ with $\rho(\ell(\cdot))$, where $\rho$ is any strictly decreasing function, and the theory in this article would go through unchanged.
+
+\begin{comment}
+\begin{align*}
+  \hat\sigma^2(\beta)
+  &\triangleq \argmax_{\sigma^2>0} \ell(\beta,\sigma^2;\,Y) 
+  = \frac{1}{n}\|Y-X\beta\|_2^2, \quad \text{ and }
+  \\
+  \ell^*(\beta; \,Y)
+  &\triangleq \sup_{\sigma^2>0} \ell(\beta, \sigma^2;\,Y) 
+  = -\frac{n}{2} - \frac{n}{2}\log\left(\frac{2\pi\|Y-X\beta\|_2^2}{n}\right).
+\end{align*}
+
+As a result the lasso coefficients $\hat\beta_{\text{lasso}}^\lambda$ minimize
+\[
+\exp\{-2\ell(\beta,\sigma;\,Y)/n \,+\, \text{const.}\} + \lambda\|\beta\|_1,
+\]
+instead of the criterion in \eqref{eq:pen-lik}.
+If $\Lambda=(0,\infty)$ then both methods result in exactly the same ever-active path.
+Otherwise, we could ``fix'' this issue (at the cost of some notational bloat) by replacing $-\ell(\cdot)$ with $\rho(\ell(\cdot))$, where $\rho$ is any strictly decreasing function, and all of the theory in this article would go through unchanged.
+\end{comment}
+
 \subsection{Which Null Hypothesis?}
@@
 Note that in our formulation of the problem, the type I error $V=(\hk-k_0)_+$ is defined in a ``model-centric'' fashion: at step $k$, we are testing  whether a particular linear model $M(E_{k-1})$ adequately describes the data $Y$.
-Even if the next selected variable $X_{j_k}$ is a noise variable, it is not necessarily a mistake to reject $M(E_{k-1})$ provided there are other signal variables that have not yet been included.
+Even if the next selected variable $X_{j_k}$ is a noise variable in the full model, it is not necessarily a mistake to reject $M(E_{k-1})$ provided there are other signal variables that have not yet been included.
 
@@
 \[
-\theta^E = X_E^\dagger \mu = \argmin_{\theta\in \R^{|E|}} \|\mu - X_E\theta\|_2^2,
-\]
-where $X_E^\dagger$ is the Moore-Penrose pseudoinverse of the matrix $X_E$. 
+\theta^E = X_E^+ \mu = \argmin_{\theta\in \R^{|E|}} \|\mu - X_E\theta\|_2^2,
+\]
+where $X_E^+$ is the Moore-Penrose pseudoinverse of the matrix $X_E$. 
 These are the true ``partial regression coefficients''  for design matrix $X_E$.
- Write $\proj_E\mu = X_E^\dagger \theta^E$ for the projection of $\mu$ into the column space of $X_E$.
+ Write $\proj_E\mu = X_E \theta^E$ for the projection of $\mu$ into the column space of $X_E$.
 
@@
 \begin{align}
-V^{\text{inc}} &= \#\{k < k_0:\; H_k^{\text{inc}} \text{ is true}\} \quad \text{ and } \\
-V^{\text{full}} &= \#\{k < k_0:\; H_k^{\text{full}} \text{ is true}\},
+V^{\text{inc}} &= \#\{k < \hk:\; H_k^{\text{inc}} \text{ is true}\} \quad \text{ and } \\
+V^{\text{full}} &= \#\{k < \hk:\; H_k^{\text{full}} \text{ is true}\},
 \end{align}
@@
 
-In the first procedure, we construct an analysis-of-variance (ANOVA) table performing an $F$- or $t$-test of each model against the next.
+In the first procedure, we construct a (type I) analysis-of-variance (ANOVA) table performing an $F$- or $t$-test of each model against the next.
 We then stop and select model $M_k=M(\{1,\ldots,k\})$ the first time the $F$-test $p$-value is larger than $\alpha$.
@@
 The methods we consider in this article are essentially generalizations of this procedure for the setting where the model sequence is selected adaptively, and where we may wish to control FDR instead of FWER.
-
-\JTcomment{Maybe a comment about Type I and Type III ANOVA tables here?
-I think these are the correct types of ANOVA tables but am not sure...This issue of full vs incremental at least has some classical literature on it.}
 
@@
 One  exception is the knockoff filter of \citet{barber2014controlling}, a variant of which provably controls the full-model FDR.
-We compare our proposal to the Knockoff method in Section \ref{sec:sparseReg}.
+We compare our proposal to the knockoff method in Section \ref{sec:sparseReg}.
 
@@
 We first review several proposals for ordered-testing procedures, several of which require independence of the null $p$-values conditional on the non-null ones.
-While these procedures also assume the sequence of hypotheses is fixed, in our setting, the truth or falsehood of $H_k$ is random, depending on which model is selected at step $k$.
+These procedures also assume the sequence of hypotheses is fixed, whereas in our setting the truth or falsehood of $H_k$ is random, depending on which model is selected at step $k$.
 However, we show that the error guarantees for these stopping rules do transfer to the random-hypothesis setting, provided we have the same independence property conditional on $k_0$.
 
-\subsection{Stopping Rules for Ordered Testing}
+\subsection{Proposals for Ordered Testing of Fixed Hypotheses}
 \label{sec:orderedProposals}
+
+We now review several proposals for ordered testing of fixed null hypotheses along with their error guarantees in the case where the sequence of null hypotheses is fixed.
+In the next section we will extend the analysis to random null hypotheses.
 
@@
 
-Because our hypotheses are nested, we have $\hk > k_0$ if and only if $p_{k_0+1} \leq \alpha$.
-If $k_0$ is fixed, then, it is clear that $\hk_B$ controls the FWER at level $\alpha$ provided that $p_k$ is a valid $p$-value for each $k$.
+Since $k_0$ is the index of the first null hypothesis, we have
+\begin{equation}\label{eq:basic-stop-control}
+ \text{FWER} \leq \P(p_{k_0} \leq \alpha) \leq alpha.
+\end{equation}
 
@@
 \citet{li2015accumulation} generalize ForwardStop, introducing the family of {\em accumulation tests}, defined as follows.
-Fix a function $h: [0,1] \rightarrow [0,\infty]$ satisfying $\int_{t=0}^1 h(t)dt=1$, called an {\em accumulation function.} Using $h$ they construct an estimate of FDP at step $k$:
+Fix a function $h: [0,1] \rightarrow [0,\infty]$ satisfying $\int_{t=0}^1 h(t)dt=1$, called an {\em accumulation function.} Using $h$ we can construct an estimate of FDP at step $k$:
 \begin{equation}
@@
 \end{equation}
-and select the largest model for which $\widehat{\text{FDP}}_k \leq \alpha$:
-\begin{equation}
-\hat k_h=\max \left\{k:\; \widehat{\text{FDP}}_h(k) \leq \alpha \right\}.
-\end{equation}
-Taking $h(p)=-\log(1-p)$ recovers the ForwardStop rule. Another choice they propose is the {\em HingeExp} function
+and select the largest model for which $\widehat{\text{FDP}}_k \leq \alpha$. Taking $h(p)=-\log(1-p)$ recovers the ForwardStop rule. Another choice they propose is the {\em HingeExp} function
 \begin{equation}
@@
 
-\begin{comment} 
-Now let ${\rm FP}(k)$ be the actual number of false positives up to and including step $k$.
-They prove that for any  accumulation function $h$ bounded by a constant $C$,
-\begin{equation}
-{\rm E} \Bigl[ \frac{{\rm FP}(\hat k_h)}{C/\alpha+\hat k_h} \Bigr] \leq \alpha
-\end{equation}
-They demonstrate numerically that the HingeExp rule offers superior empirical performance compared to ForwardStop  (and other competitors) in many situations.
-
-Importantly, this class of accumulation tests require that  the $p$-values $p_i$  are  $U(0,1)$ if the null hypothesis $H_i$ is true, and are independent of the non-null $p$-values.
-This implies that the class of sequential $p$-values proposed in this paper can be used as inputs to these tests.
-\end{comment}
-
-\subsection{Random Hypotheses}
+\subsection{Ordered Testing of Random Hypotheses}
 
 In our problem setting, the sequence of selected models is random; thus, the truth or falsehood of each $H_k$ is not fixed, as assumed in the previous section.
-This could in principle lead to violation of the error control guarantees that hold in the fixed-hypothesis case.
-\RTcomment{As Will and I discussed, this example is confusing and needs improvement.}
-To see why we a problem could arise, consider apply BasicStop in the following counterexample: let $d=2$ and let $k_0$ be 0 or 1 with probability 1/2 each. Then, let 
-\[
-p_1 \sim U[0,1], \quad p_2 \sim (1-k_0 + U[0,1])/2.
-\]
-Both $p_1$ and $p_2$ are uniform, but
-\begin{align*}
-\P(V>0) &= \P(p_{k_0+1} \leq \alpha)\\ 
-&= \sum_{k=1}^2\P(p_k\leq \alpha \mid k_0+1=k)\cdot \P(k_0+1=k)\\
-&= \alpha\cdot \frac{1}{2} + 2\alpha \cdot \frac{1}{2} = 3\alpha/2
-\end{align*}
-This problem is resolved if $p_k$ is a valid $p$-value conditional on the value of $k_0$.
-
-Similarly, the proposals of  \citet{gsell2013sequential} and \citet{li2015accumulation} assume that the null hypotheses $H_{1}, \ldots, H_{d}$ are fixed, and that the null $p$-values are independent of each other and of the non-null $p$-values.
-Specifically, they assume that
-\begin{equation}\label{eq:indepCond_fixed_k0}
-\P(p_{k_0+1} \leq \alpha_{k_0+1}, \ldots, p_d \leq \alpha_d
-\mid p_1, \ldots, p_{k_0}) = \prod_{i=k_0+1}^d \alpha_i,
-\end{equation}
-where $k_0$ is fixed.
-
-In our problem the null hypotheses at each step are random, and therefore so is the completion index $k_0(Y)$, so the error control guarantees of \citet{gsell2013sequential} and \citet{li2015accumulation} do not directly apply.
-However, we will recover these guarantees if we have {\em conditional} independence of null $p$-values given $k_0$:
+It is not immediately clear, then, how much we should condition on when we construct the $p$-value at step $k$.
+In particular, is it sufficient to condition only on the current null model $M_{k-1}$, or must we condition on the entire subpath of models $M_0,\ldots,M_{k-1}$?
+
+Note that conditioning on the current null model {\em is} enough to guarantee that $p_k$ is uniform conditional on the event $\{F\in M_{k-1}(Y)\}$ that the $k$th null is true.
+Unfortunately, however, it is {\em not} enough to guarantee that the above stopping rules actually control FDR or FWER.
+
+Let $z_{\alpha/2}$ denote the upper-$\alpha/2$ quantile of the $\cN(0,1)$ distribution and consider the following counterexample:
+
+\begin{restatable}{proposition}{counterex}\label{prop:counterexample}
+  Consider linear regression with $n=p=3$, known $\sigma^2=1$ and identity design $X=I_3$.
+
+  Suppose that we construct the model path as follows: if $|Y_3|>z_{\alpha/2}$, we add $X_1$ to the active set first, then $X_2$, then $X_3$.
+Otherwise, we add $X_2$, then $X_1$, then $X_3$.
+At each step we construct selective max-$z$ $p$-values and finally choose $\hat k$ using BasicStop.
+
+  If $\mu=(0,C,0)$, then the FWER for the entire procedure tends to $2\alpha-\alpha^2$ as $C\to\infty$.
+\end{restatable}
+
+A proof of Proposition~\ref{prop:counterexample} is given in the appendix.
+The crucial issue is that $k_0$ is no longer a fixed index.
+Consequently, even though $\L(p_k \mid H_k \text{ true})=U[0,1]$ for each fixed $k$, the $p$-value $p_{k_0}$ corresponding to the {\em first} true null hypothesis is stochastically smaller than uniform.
+
+If, however, we could guarantee that
+\[
+\P(p_k \leq \alpha \mid H_k \text{ true}, k_0=k) \leq \alpha
+\]
+then BasicStop again would control FWER, by the argument in \ref{eq:basic-stop-control}.
+
+Note that $k_0$ is unknown, so we cannot directly condition on its value when we construct our single-step $p$-values. However, because
+\[
+\{k_0=k\} = \{F\in M_{k-1}, F\notin M_{k-2}\},
+\]
+it is enough to condition on $(M_{k-2}, M_{k-1})$ when we construct our $p$-values.
+As we will see in Section~\ref{sec:pValsIndep}, conditioning on $(M_{k-1}, U_{k-1})$ is equivalent to conditioning on $(M_{0:k-1}, U_{k-1})$ in most cases of interest including forward stepwise likelihood paths and ever-active regularization paths (but not the path in Proposition~\ref{prop:counterexample}).
+
+Similarly, the error control guarantees of \citet{gsell2013sequential} and \citet{li2015accumulation} do not directly apply to case where the null hypotheses are random.
+However, we recover these guarantees if we have {\em conditional} independence and uniformity of null $p$-values given $k_0$:
 \begin{equation}\label{eq:indepCond_random_k0}
@@
 \end{equation}
-
-The following proposition shows that~\eqref{eq:indepCond_random_k0} is the correct generalization of~\eqref{eq:indepCond_fixed_k0} to the random-hypothesis case.
+For the sake of brevity, we will say that $p$-value sequences satisfying~\eqref{eq:indepCond_random_k0} are {\em independent on nulls}.
+
+The following proposition shows that independence on nulls allows us to transfer the error-control guarantees of ForwardStop and accumulation tests to the random-hypothesis case.
 \begin{proposition}
@@
 
-  Suppose that $\hk$ controls $\Err_g$ at level $\alpha$ if the $H_{1:d}$ are fixed and $p_{1:d}$ satisfy~\eqref{eq:indepCond_fixed_k0}.
+  Suppose that $\hk$ controls $\Err_g$ at level $\alpha$ if the $H_{1:d}$ are fixed and the null $p$-values are uniform and independent of each other and the non-nulls.
 Then, $\hk$ controls $\Err_g$ at level $\alpha$ whenever $p_{1:d}$ satisfy~\eqref{eq:indepCond_random_k0}.
@@
 
-\begin{comment}
-In real applications, null $p$-values could be conservative instead of exactly uniform, for instance if the test statistic is discrete or the null is composite.
-We can relax~\eqref{eq:indepCond_random_k0} to require only that the joint distribution of null $p$-values is stochastically larger than uniform:
-\begin{equation}\label{eq:stochLargeCond_random_k0}
-  \P(p_{k+1} \leq \alpha_{k+1}, \ldots, p_d \leq \alpha_d
-  \mid p_1, \ldots, p_{k_0}, \; k_0 = k) \leq \prod_{i=k+1}^d \alpha_i
-\end{equation}
-If the null $p$-values are each uniform given $(p_{1:k_0},k_0)$, then the conditions~\eqref{eq:indepCond_random_k0} and~\eqref{eq:stochLargeCond_random_k0} are equivalent.
-For most stopping rules $\hk$, the weaker condition~\eqref{eq:stochLargeCond_random_k0} is enough to guarantee control of whatever error rate holds under~\eqref{eq:indepCond_random_k0}.
-\end{comment}
-
-For the sake of brevity, we will say that $p$-value sequences satisfying~\eqref{eq:indepCond_random_k0} are {\em independent on nulls}.
-
-The next Section discusses conditions on the model sequence $M_{0:d}$ and the $p$-value sequence $p_{1:d}$ under which~\eqref{eq:indepCond_random_k0} is satisfied.
+Note that~\eqref{eq:indepCond_random_k0} implies in particular that each $p_k$ is uniform given $k_0=k$; thus, independence on nulls is enough to guarantee that BasicStop and ForwardStop control FWER and FDR, respectively.
+The next section discusses conditions on the model sequence $M_{0:d}$ and the $p$-value sequence $p_{1:d}$ under which~\eqref{eq:indepCond_random_k0} is satisfied.
 
@@
 \begin{equation}\label{eq:maxTnull}
-\L\left(T_k \; \mid \; E_{k-1}, X_{E_{k-1}}'Y, \|Y\|^2 \right).
+\L\left(T_k \; \mid \; E_{k-1}, \,X_{E_{k-1}}'Y, \,\|Y\|^2 \right).
 \end{equation}
@@
 Its performance in our simulation is similar to that of the max-$t$ test.
-
-\begin{comment}
-Two specific examples are worth discussing in detail. First, consider the {\em selective max-$t$ test}, a generalization of the max-$t$ test described in 
-~\citet{buja2014}. Controlling for $E$, the $z$-statistic for coefficient $j\notin E$ is
-\begin{equation}
-  z_{j,E}(Y) = \frac{(\proj_E^\perp X_j)'Y}
-  {\sigma\|\proj_{E}^\perp X_j\|_2},
-\end{equation}
-and the $t$-statistic is
-\begin{equation}
-  t_{j,E}(Y) = \frac{(\proj_E^\perp X_j)'Y}
-  {\hat \sigma_{j,E} \|\proj_{E}^\perp X_j\|_2},
-  \qquad \text{where} \quad 
-  \hat \sigma_{j,E}^2(Y) = 
-  \frac{\|\proj_{E \cup \{j\}}^\perp Y\|_2^2}{n-|E|-1}.
-\end{equation}
-Note that $z_{j,E}$ is a monotonic function of $X_j'Y$ given $X_E'Y$, and $t_{j, E}$ is a monotonic function of $X_j'Y$ given $(X_E'Y, \|Y\|_2)$. Thus $z_{j,E}$ is measurable with respect 
-to $X_{E \cup \{j\}}'Y$ if $\sigma^2$ is known, and otherwise $t_{j,E}$ is measurable with respect to $(X_{E \cup \{j\}}'Y, \|Y\|_2)$.
-
-We can then define the max-$t$-statistic (respectively max-$z$-statistic) for active set $E$ as
-\begin{equation}\label{eq:maxT}
-  U_E(Y) = \max_{j \notin E} \left|t_{j,E}(Y)\right|,
-\end{equation}
-and use $U_{E_{k-1}}(Y)$ as our test statistic at step $k$.
-
-$U_{E_{k-1}}(Y)$ is $\sF_k$-measurable if and only if $j_k$ is the maximizer in~\eqref{eq:maxT}.
-Consequently, max-$t$ $p$-values are independent on nulls if they are used in tandem with forward-stepwise regression, where the next variable added is always the variable with the largest $t$-statistic.
-
-By contrast, the LASSO path does not in general add the variable with the largest $t$-statistic. In that case, we could use instead the {\em selective next-$\lambda$ test}, which uses test statistic
-\begin{equation}
-  U_k = \inf\{ \lambda:\; \hat\beta^\lambda \notin \beta(E_{k-1})\}.
-\end{equation}
-Here $\beta(E_{k-1})$ is the support of the ever-active set $E_{k-1}$, so that $U_k$ is the next $\lambda$ at which a new predictor is added by the lasso.
-\RTcomment{Will: we never define $\beta(E)$: have i said it right?}
-The next-$\lambda$ statistic is $\sF_k$-measurable if we use the LASSO to select variables. \WFcomment{Relate this to spacings test. Is this essentially the ``generalized spacings test?''}
-\end{comment}
 
@@
 
-\begin{comment}
-Computing the cutoff for the saturated-model test requires us to know $\proj_{\eta}^\perp Y$, which is not $\sF_k$-measurable.
-
-By contrast, selected-model tests of $M_{k-1}$ against $M_k$ are always $\sF_k$-measurable, because they are based on the law
-\begin{equation}\label{eq:selModel}
-\L\left(X_{j_k}'Y \mid X_{E_k-1}'Y, E_{k-1}, j_k\right),
-\end{equation}
-and all of the random variables appearing in~\eqref{eq:selModel} are $\sF_k$-measurable.
-\end{comment}
-
 % latex table generated in R 3.0.2 by xtable 1.7-1 package
@@
 
-\begin{comment}
-First, note that $p_2$ is a selective $p$-value for comparing the null with only variable 1,
-\[
-M_1:\; Y \sim \cN\left(\binom{\mu_1}{0}, I_2\right),
-\]
-against the alternative with variables 1 and 2:
-\[
-M_2:\; Y \sim \cN\left(\mu, I_2\right), \text{ with } 
-\mu_2 \neq 0.
-\]
-To eliminate the nuisance parameter $\mu_1$, the test conditions on $Y_1$.
-Thus, $p_2$ is uniform given $Y_1$ on $A$.
-Second, note that on $A$, $p_1$ is a function only of $Y_1$.
-Thus, $(p_1,p_2)$ are independent uniform variables given that variable 1 is chosen first.
-By a similar argument, they would also be independent uniforms if $Y_2$ were chosen first; thus, $(p_1, p_2)$ are marginally uniform under the global null.
-\end{comment}
-
 \section{Computation}
@@
 Due to the generality in which we have posed the problem, we cannot give a general prescription for how to carry out this sampling.
-However, we give some details for the important case of linear regression in Section~\ref{sec:linRegComputation}.
+However, we give some details for the important case of linear regression in Appendix~\ref{sec:linRegComputation}.
 
@@
 
-\subsection{Details for Linear Regression}\label{sec:linRegComputation}
-
-As discussed in Section~\ref{sec:singleStep}, we will set
-\[
-p_k(Y) = p_{k, E_{k-1}(Y)}(Y),
-\]
-where $p_{k,E}(Y)$ for each $E\sub \{1,\ldots, p\}$ is a conditional $p$-value for the {\em fixed} null model $M(E):\, Y \sim N(X_E\beta, \sigma^2)$, valid on the selection event $A=\{E_{k-1}(Y)=E\}$.
-
-If $\sigma^2$ is known, then $U_{k-1}(Y) = X_{E}'Y$ on $A$, so our task is to sample under the null from
-\begin{equation}\label{eq:truncNorm}
-\L(Y \mid X_E'Y, Y\in A),
-\end{equation}
-which is a multivariate Gaussian distribution supported on the hyperplane of dimension $n-|E|$ with $X_E'Y$ fixed at its realized value, and truncated to the event $A$.
-
-If $\sigma^2$ is unknown, then $U_{k-1}=(X_E'Y,\, \|Y\|^2)$ on $A$ and we must sample under the null from
-\begin{equation}\label{eq:truncUnif}
-\L(Y \mid X_E'Y,\, \|Y\|^2,\, Y\in A),
-\end{equation}
-which is a uniform distribution supported on a sphere of dimension $n-|E|-1$ defined by fixing $X_E'Y$ and $\|Y\|^2$ at their realized values, and truncated to the event $A$.
-For the simulations in this article, we use a hybrid of two strategies:
-
-\paragraph{Accept/Reject Sampler} In either of these two cases, sampling would be very easy if $A$ were the entire sample space $\R^n$; the only reason we might encounter difficulty is that $A$ could be an irregular set that is difficult to condition on.
-However, we can always sample from $\L(Y \mid U)$ and then keep each new sample $Y$ if it lies in the selection event.
-
-\paragraph{Hit-and-Run Sampler} \citet{fithian2014optimal} propose sampling from laws like that in~\eqref{eq:truncNorm} via a hit-and-run sampler for the $n-|E|$-dimensional multivariate Gaussian distribution: at each step, to generate $Y^{t+1}$ from $Y^t$, we choose a uniformly random unit vector $\nu$ in the hyperplane, and resample $\nu'Y$ conditional on the constraints $\proj_{\nu}^\perp Y^{t+1}=\proj_{\nu}^\perp Y^t$ and the constraint $Y^{t+1}\in A$.
-If $A$ is a polytope this amounts to resampling a univariate Gaussian distribution with known mean and variance, truncated to an interval defined by maximizing and minimizing $\nu'Y^{t+1}$ subject to linear constraints.
-For the case with $\sigma^2$ unknown, the sampler is a bit more involved; see \citet{fithian2014optimal}.
-
-Figure \ref{fig:comparison} compares the two samplers in an example with $n=100, p=20$ and one strong signal, computing the max-$t$ $p$-value at Step 3.
-The ``exact'' answer was determined from bootstrap sampling with $20,000$ replications.
-The Figures show the bias, variance and mean squared error (MSE) of the estimates from accept/reject and hit and run, using between 1000 and 8000 samples, averaged
-over 10 different starting seeds.
-\begin{figure}[htp]
-\centering
-  \includegraphics[width=\textwidth]{figs/comparison.pdf}
-  \caption{\em Simulated example with $n= 100, p=20$ and one strong signal, computing the max-$t$ $p$-value at Step 3.
-Shown are bias, variance and MSE of the estimates from accept/reject and hit and run, using between 1000 and 8000 samples, averaged over 10 different starting seeds.
-}
-\label{fig:comparison}
-\end{figure}
-We see that accept/reject is far more efficient, yielding a smaller MSE with 1000 samples then hit and run gives for 8000 samples.
-As we move further along the path, the set $A$ becomes more and more constraining, so that we must discard the vast majority of the samples we generate.
-Thus we use accept/reject for as many steps of the sequence is possible,  until the point where the acceptance rate is too low to make is practical.
-
-\JTcomment{I don't really see the point of Figure 5. It is abundandtly clear
-that if feasible an IID sample is preferable over an MCMC sample. It is
-just as easy to make plots where accept/reject will take forever to get even one sample point. 
-}
-
-In detail, we employ a hybrid computational strategy; we use accept-reject sampling, and do so as long as we obtain at least 300 (say) samples in the first set of $B$ realizations ($75,000$ in our code)  This is an effective approach until we have reached a point in the model path where all predictors have little or no signal.
-When the acceptance rate gets too low, we switch to a hit-and run Monte Carlo approach.
-We exploit the fact that each move along a random direction  has a truncated Gaussian distribution whose truncation limits are easily computed from the properties of the polyhedron.
-Unlike accept/reject sampling, which produces independent samples,  the hit and run method produces dependent samples.
-Hence we must run it for longer with an initial burn in period, and hope that it has mixed sufficiently well.
-
-The task is made much simpler by the fact that the selection event can we written as a polytope $A=\{y: \Gamma y \geq u\}$.
-Hence we don't need to run the entire forward stepwise procedure on $Y^*$: instead, we pre-compute $\Gamma$ and $u$ and then check if $\Gamma Y^* \geq u$.
-In the case of forward stepwise and lasso regression, $\Gamma$ has $2(p-k)$ rows after $k$ steps, as we see next.
-
-\subsection{Constraints}\label{sec:constrainsts}
+\subsection{Constraints}\label{sec:constraints}
 Happily, it turns out that the conditioning polytope can be expressed with only $2(p-k)$ constraints in forward stepwise regression and the lasso, as we show here.
@@
 \end{proof}
-
-\begin{comment}
-\subsection{An analytic approximation for the max-$t$ test}
- Suppose we have taken $k$ steps of forward stepwise regression with $p$ predictors and let the selected model  
- be $E$. Suppose we choose predictor $j^*$ at the next step.
-
-Let $T_j=\eta_j^Ty, j \notin M$ be the partial regression coefficients for all predictors not in $E$,
-and $(v_j^-, v_j^+)$ be the limits derived from the polyhedral lemma applied to $(E, \eta_j)$  (See e.g. \citet{spacings}).
-These are the upper and lower bounds for the functional $\eta_j^Ty$ implied by the event that model $E$ is selected from the data $y$.
-We chose the sign of $\eta_j$ so that $\eta_j^Ty$ is  $\geq 0$.
-
- Let $F_{\mu,\sigma^2}^{a,b}(t)$ be the  cumulative distribution function of the truncated normal.
- Finally let $T=|T_{j^*}|$.
-Then by assuming independence, a  simple approximation to the max-$t$ value is given by
-\begin{equation}
-1-\prod_{j\notin E} \{F_{0,||\eta||^2}^{v_j^-, v_j^+}(T)-   F_{0,||\eta||^2}^{v_j^-, v_j^+}(-T) \}
-\label{eqn:approxpv}
-\end{equation}
-This approximation is exact if the predictors are uncorrelated.
-
-Figure \ref{fig:approxpv} shows an example with $n=50$ and $p=10$ features having pairwise correlation 0.5.
-There are two moderate signals.
-Shown are the $p$-values from the ``exact'' numerical computation (with $100,000$ bootstrap samples)
-and the approximation based on (\ref{eqn:approxpv}).
-\begin{figure}[htp]
-\centering
-  \includegraphics[width=3in]{figs/approxpv.pdf}
-  \caption{Exact max-$t$ $p$-values and approximation based on (\ref{eqn:approxpv}), for a simulated example
-  with $n=50, p=10$.}
-  \label{fig:approxpv}
-\end{figure}
-The approximation seems quite good. By Sidak's lemma, we might conjecture that the approximation is conservative, but we have no proof of this.
-\end{comment}
 
@@
 The {\tt Simple} rule stops at the first time that a $p$-value exceeds $\alpha=0.05$, while {\tt Forward} and {\tt Strong} refers to the ForwardStop and StrongStrop rules of Section \ref{sec:orderedProposals}.
-Results with theoretical guarantees are in \guarantee{bold highlighted}.}
+Results with theoretical guarantees are in \guarantee{bold} type.}
 \label{tab:stopping05}
@@
 The {\tt Simple} rule stops at the first time that a $p$-value exceeds $\alpha=0.2$, while {\tt Forward} and {\tt Strong} refers to the ForwardStop and StrongStrop rules of Section \ref{sec:orderedProposals}.
-Results with theoretical guarantees are \guarantee{bold highlighted}.}
+Results with theoretical guarantees are in \guarantee{bold} type.}
 \label{tab:stopping20}
@@
 
-\section{Futher Examples and Extensions}
+\section{Futher Examples and Extensions}\label{sec:further-examples}
 
@@
 
-\begin{comment}
-For example the set of candidate basis functions available at each stage are a function of what has been  previously entered into the model
-Nonetheless,  the selection procedure  for the first set of $k$ splits can be expressed in polyhedral from $Ay \leq b$, and the saturated and selective inference procedures can be applied.
-As before, the saturated $p$-values have an explicit truncated Gaussian form, while computation of the selective $p$-values require accept-reject sampling.
-
-Figure \ref{fig:cart} shows an example with independent standard Gaussian features, and $n=30, p=3$.
-We 
-consider just the first two splits, in a null case (top row) and non-null (bottom row).
-In the non-null case,  the data are generated from a  model of the form (\ref{eqn:intmodel}).
-
-As expect, the nominal $p$-values are far too small and hence don't control type I error.
-Both the saturated and selective $p$-values control type I error, with the selective $p$-values are somewhat more powerful in the non-null case.
-Rules like the forwardStop procedure could be applied  to the selective $p$-values to choose a ``best'' model size.
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.65\textwidth]{figs/cart.pdf}
-  \caption{\em Cumulative distribution functions of $p$-values for  the first two splits in a CART decision tree. In the top row, the true model is null,
-  while in the bottom row, the data are generated from a  model of the form (\ref{eqn:intmodel})}
-  \label{fig:cart}
-\end{figure}
-\end{comment}
-
 \subsection{Nonparametric Changepoint Detection}
@@
 
-\begin{comment}
-\section{Principal Components Analysis}
-\label{sec:pca}
-
-\WFcomment{Can we do this one?}
-\JTcomment{In principle, we can. But, paper is a little long. Maybe
-just point out in discussion
-that the $p$-values in Choi et al.
-are examples of this setup?
-The notation in Yunjin's paper is nothing like this, but we do form selected model tests with the sufficient statistic growing in dimension as we go down the path.
-In this case, ``maxT'' is the same as ``maxT identify'' where we ``identify'' the sets of singular vectors...}
-
-Here we consider model selection for principal components analysis. In that case we are given a data matrix $X \in \R^{n\times d}$, with which we form a sample %covariance matrix
-\[
-S = \frac{1}{n-1} \sum_{i=1}^n(x_i - \bar x)^2
-\]
-The first $d$ principal component loadings are the first $d$ eigenvectors of $S$, which call $u_1,\ldots, u_d$. These induce a sequence of nested Wishart models:
-\[
-M_0 \sub M_1 \sub \cdots \sub M_d
-\]
-in which
-\begin{equation}
-  M_k:\; (n-1) S \sim W_d\left(\lambda_0 I_d + \sum_{i=1}^k     \lambda_i u_i u_i', \;\;\; n-1\right).
-\end{equation}
- By successively conditioning on $u_j, j<k$ one can obtain an exact sequential test
-for the complete null hypothesis $H_0: \lambda_j=0 \; \forall j \neq k$.  The computation of the resulting $p$-values requires numerical integration.
-The details are given in \citet{choi2014selecting}, with their ``integrated CSV'' test  corresponding to the sequential selective test of this paper.
-
-\RTcomment{Jon- add more details?}
-
-\section{Decision trees and interaction models}
-In this section we discuss another application of adaptive sequential inference, to binary decision trees.
-Specifically, the classification and regression tree  (CART)  method builds a decision tree in a top down manner, finding the best available split
-at each node.  When considering splitting two daughter nodes, the generic CART algorithm splits the nodes in an arbitrary order, only stopping when a node
-reaches a minimum size.  We instead consider a  special version of CART  algorithm:  the   "best first"  method, which order splits by their achieved reduction in loss.
-This procedure is used, for example in the R package for gradient boosting, called {\tt gbm}.
-The best first tree-builder is thus a sequential selection  procedure,  to which we can apply the results in this paper. Hence we can obtain independent, sequential $p$-values
-for each successive split in the decision  tree. We now give some details and an example.
-
-Given data $(y_i, x_{i1}, x_{i2},\ldots x_{ip})$ for $i=1,2,\ldots N$,
-we choose a splitting  variable $j_1$ and split point $t_1$, leading to a model of the form
-\begin{eqnarray}
-\hat y_i=\beta_0 +\beta_{11} I(x_{ij_1} \leq t_1) +\beta_{12} I(x_{ij_1} > t_1)
-\end{eqnarray}
-Then at the second stage, we try splitting one of the two daughter nodes, and choose the split and most reduces the residual sum of squares.
-Suppose that a split of the first node is chosen, using predictor $j_2$ at split point $t_2$.  This leads to the interaction model of the form
-\begin{eqnarray}
-\hat y_i=\beta_0 +\beta_{11} I(x_{ij_1} \leq t_1) \cdot \beta_{31} I(x_{ij_2} \leq t_2)+   \beta_{11} I(x_{ij_1} \leq t_1) \cdot  \beta_{31} I(x_{ij_2} > t_2)  +\beta_{12} I(x_{ij_1} > t_1).\cr
-\label{eqn:intmodel}
-\end{eqnarray}
-Note that the search for possible split points can be expressed through an exhaustive set of basis functions $I(x_{ij} \leq  t_i), I(x_{ij}  > t_i)$ for each $t_i \in (x_{1j}, x_{2j}, \ldots x_{nj})$.
-The procedure is somewhat similar to foreward stepwise regression, but with differences.
-For example the set of candidate basis functions available at each stage are a function of what has been  previously entered into the model
-Nonetheless,  the selection procedure  for the first set of $k$ splits can be expressed in polyhedral from $Ay \leq b$, and the saturated and selective inference procedures can be applied.
-As before, the saturated $p$-values have an explicit truncated Gaussian form, while computation of the selective $p$-values require accept-reject sampling.
-
-Figure \ref{fig:cart} shows an example with independent standard Gaussian features, and $n=30, p=3$.
-We 
-consider just the first two splits, in a null case (top row) and non-null (bottom row).
-In the non-null case,  the data are generated from a  model of the form (\ref{eqn:intmodel}).
-
-As expect, the nominal $p$-values are far too small and hence don't control type I error.
-Both the saturated and selective $p$-values control type I error, with the selective $p$-values are somewhat more powerful in the non-null case.
-Rules like the forwardStop procedure could be applied  to the selective $p$-values to choose a ``best'' model size.
-\begin{figure}[h]
-  \centering
-  \includegraphics[width=.65\textwidth]{figs/cart.pdf}
-  \caption{\em Cumulative distribution functions of $p$-values for  the first two splits in a CART decision tree. In the top row, the true model is null,
-  while in the bottom row, the data are generated from a  model of the form (\ref{eqn:intmodel})}
-  \label{fig:cart}
-\end{figure}
-\end{comment}
-
 \section{Discussion}
@@
 In future work we also plan to address the matter of performing selective inference for parameters of the model actually selected.
-In principle, we already know how to do this using the framework of \citet{fithian2014optimal} --- we simply condition on the event that $M_{\hk}$ is selected and perform inference with respect to $M_{\hk}$ --- but there is some work to do in understanding the computational complexity of this task.
-
-\begin{comment}
-Of course, the answer depends on the scientific context: for example, suppose that at step 1, forward stepwise regression selects a variable $X_1$, and then at some later step it selects another variable $X_2$, which is almost perfectly correlated with $X_1$.
-There may be little evidence to support adding $X_2$ to the model once $X_1$ is already included, even if including $X_2$ would greatly change our inference about the coefficient for $X_1$ (for example, by making its confidence interval much wider).
-Depending on the context, we could draw the conclusion that 1) $X_1$ and $X_2$ are near-duplicate variables and it is therefore unnecessary (and possibly counterproductive) to include both in the model, or 2) $X_2$ is a vital confounding variable for $X_1$ and the confidence interval for $\beta_1$ ought to reflect the resultant uncertainty.
-If the second interpretation is the scientifically appropriate one, then we should probably use a different selection algorithm --- for example, a variant of forward stepwise that always adds both $X_1$ and $X_2$ to the model as a group in the same step.
-\end{comment}
-
-Open-source code is available at this article's github repository at \texttt{www.github.com/wfithian/sequential-selection}.
+In principle, we already know how to do this using the framework of \citet{fithian2014optimal} --- we simply condition on the event that $M_{\hk}$ is selected and perform inference with respect to $M_{\hk}$ --- but the task may be computationally awkward in general.
+
+Open-source code is available at this article's github repository at:
+\begin{center}
+\texttt{www.github.com/wfithian/sequential-selection}
+\end{center}
 
@@
 
+\newpage
+
+\begin{appendix}
+
+\section{Proof of Proposition~\ref{prop:counterexample}}
+
+In this section we prove Proposition~\ref{prop:counterexample}, restated below for convenience.
+
+\counterex*
+
+\begin{proof}
+Denote the events where we choose the two possible orderings as $A_{123}$ and $A_{213}$, respectively. Because the variables are orthogonal, the max-$z$ statistic for any model $M(E)$ is 
+\[
+|z_E^*|=\max_{j\notin E} |Y_j|.
+\]
+
+At each step we use the selective max-$z$ test conditioning on the current null model, where $p_k$ is small when $|z_{E_{k-1}}^*|$ is large compared to the law
+\[
+\L(|z_{E_{k-1}}^*| \,\mid\, E_{k-1}, Y_{E_{k-1}}).
+\]
+
+Note that whichever of $A_{123}$ and $A_{213}$ occurs, $E_2=\{1,2\}$ and $|z_{E_2}^*|=|Y_3|$.
+Thus, $E_2$ is nonrandom and $p_3=2\Phi(-|Y_3|)$, the usual two-sided normal $p$-value based on $Y_3$.
+Finally, note that $k_0=2$ on $A_{123}$ and $k_0=1$ on $A_{213}$.
+
+We have
+\begin{align*}
+  FWER 
+  &= \P(A_{123})\P(p_1<\alpha, p_2<\alpha, p_3<\alpha \mid A_{123})
+  + \P(A_{213})\P(p_1<\alpha, p_2<\alpha \mid A_{213})\\
+  &= \alpha \P(p_1<\alpha, p_2<\alpha \mid A_{123})
+  + (1-\alpha) \P(p_1<\alpha, p_2<\alpha \mid A_{213}),
+\end{align*}
+using the fact that $A_{123}$ corresponds exactly to the event that $p_3<\alpha$.
+As $K\to\infty$, we have $z_\emptyset^*=z_{\{1\}}^*=Y_2>K/2$ with probability tending to 1.
+Thus, $p_1,p_2\approx 0$ on $A_{123}$, while $p_1\approx 0$ on $A_{213}$.
+Continuing the calculation, then,
+\begin{align*}
+  \lim_{K\to\infty} FWER 
+  &= \alpha\cdot 1 + (1-\alpha)\cdot \P(p_2 < \alpha \mid A_{213})\\
+  &= \alpha + (1-\alpha)\alpha = 2\alpha-\alpha^2.
+\end{align*}
+\end{proof}
+
+\section{Computational Details for Linear Regression}\label{sec:linRegComputation}
+
+As discussed in Section~\ref{sec:singleStep}, we will set
+\[
+p_k(Y) = p_{k, E_{k-1}(Y)}(Y),
+\]
+where $p_{k,E}(Y)$ for each $E\sub \{1,\ldots, p\}$ is a conditional $p$-value for the {\em fixed} null model $M(E):\, Y \sim N(X_E\beta, \sigma^2)$, valid on the selection event $A=\{E_{k-1}(Y)=E\}$.
+
+If $\sigma^2$ is known, then $U_{k-1}(Y) = X_{E}'Y$ on $A$, so our task is to sample under the null from
+\begin{equation}\label{eq:truncNorm}
+\L(Y \mid X_E'Y, Y\in A),
+\end{equation}
+which is a multivariate Gaussian distribution supported on the hyperplane of dimension $n-|E|$ with $X_E'Y$ fixed at its realized value, and truncated to the event $A$.
+
+If $\sigma^2$ is unknown, then $U_{k-1}=(X_E'Y,\, \|Y\|^2)$ on $A$ and we must sample under the null from
+\begin{equation}\label{eq:truncUnif}
+\L(Y \mid X_E'Y,\, \|Y\|^2,\, Y\in A),
+\end{equation}
+which is a uniform distribution supported on a sphere of dimension $n-|E|-1$ defined by fixing $X_E'Y$ and $\|Y\|^2$ at their realized values, and truncated to the event $A$.
+For the simulations in this article, we use a hybrid of two strategies:
+
+\paragraph{Accept/Reject Sampler} In either of these two cases, sampling would be very easy if $A$ were the entire sample space $\R^n$; the only reason we might encounter difficulty is that $A$ could be an irregular set that is difficult to condition on.
+However, we can always sample from $\L(Y \mid U)$ and then keep each new sample $Y$ if it lies in the selection event.
+
+\paragraph{Hit-and-Run Sampler} \citet{fithian2014optimal} propose sampling from laws like that in~\eqref{eq:truncNorm} via a hit-and-run sampler for the $n-|E|$-dimensional multivariate Gaussian distribution: at each step, to generate $Y^{t+1}$ from $Y^t$, we choose a uniformly random unit vector $\nu$ in the hyperplane, and resample $\nu'Y$ conditional on the constraints $\proj_{\nu}^\perp Y^{t+1}=\proj_{\nu}^\perp Y^t$ and the constraint $Y^{t+1}\in A$.
+If $A$ is a polytope this amounts to resampling a univariate Gaussian distribution with known mean and variance, truncated to an interval defined by maximizing and minimizing $\nu'Y^{t+1}$ subject to linear constraints.
+For the case with $\sigma^2$ unknown, the sampler is a bit more involved; see \citet{fithian2014optimal}.
+
+Figure \ref{fig:comparison} compares the two samplers in an example with $n=100, p=20$ and one strong signal, computing the max-$t$ $p$-value at Step 3.
+The ``exact'' answer was determined from bootstrap sampling with $20,000$ replications.
+The Figures show the bias, variance and mean squared error (MSE) of the estimates from accept/reject and hit and run, using between 1000 and 8000 samples, averaged
+over 10 different starting seeds.
+\begin{figure}[htp]
+\centering
+  \includegraphics[width=\textwidth]{figs/comparison.pdf}
+  \caption{\em Simulated example with $n= 100, p=20$ and one strong signal, computing the max-$t$ $p$-value at Step 3.
+Shown are bias, variance and MSE of the estimates from accept/reject and hit and run, using between 1000 and 8000 samples, averaged over 10 different starting seeds.
+}
+\label{fig:comparison}
+\end{figure}
+We see that accept/reject is far more efficient, yielding a smaller MSE with 1000 samples then hit and run gives for 8000 samples.
+As we move further along the path, the set $A$ becomes more and more constraining, so that we must discard the vast majority of the samples we generate.
+Thus we use accept/reject for as many steps of the sequence is possible,  until the point where the acceptance rate is too low for accept/reject to be practical.
+
+\JTcomment{I don't really see the point of Figure 5. It is abundandtly clear
+that if feasible an IID sample is preferable over an MCMC sample. It is
+just as easy to make plots where accept/reject will take forever to get even one sample point.}
+
+In detail, we employ a hybrid computational strategy; we use accept-reject sampling, and do so as long as we obtain at least 300 (say) samples in the first set of $B$ realizations ($75,000$ in our code)  This is an effective approach until we have reached a point in the model path where all predictors have little or no signal.
+When the acceptance rate gets too low, we switch to a hit-and run Monte Carlo approach.
+We exploit the fact that each move along a random direction  has a truncated Gaussian distribution whose truncation limits are easily computed from the properties of the polyhedron.
+Unlike accept/reject sampling, which produces independent samples,  the hit and run method produces dependent samples.
+Hence we must run it for longer with an initial burn in period, and hope that it has mixed sufficiently well.
+
+The task is made much simpler by the fact that the selection event can we written as a polytope $A=\{y: \Gamma y \geq u\}$.
+Hence we don't need to run the entire forward stepwise procedure on $Y^*$: instead, we pre-compute $\Gamma$ and $u$ and then check if $\Gamma Y^* \geq u$.
+In the case of forward stepwise and lasso regression, $\Gamma$ has $2(p-k)$ rows after $k$ steps, as we see next.
+
+\end{appendix}
+
 \end{document}
-
-\begin{comment}
-Sections~\ref{sec:pvalSP}--\ref{sec:modelSSP} discuss sufficient conditions on the $p$-values and the selection algorithm under which single-step $p$-values are automatically independent.
-Section~\ref{sec:selectionVariables} discusses how we can create independent $p$-values by conditioning on finer selection variables at each step.
-
-Essentially, we will want to partition the information in $Y$ according to the filtration:
-\begin{align}\nonumber
-  \sF(M_0,T_0) &\underlabel_{\text{selection } 1} 
-  \sF(M_{0:1},T_0) \underlabel_{\text{inference } 1}
-  \sF(M_{0:1},T_1) \quad \sub \;\;\cdots\\[8pt]
-  \label{eq:infoPartition}
-  \cdots\;\; \sub \quad&
-  \sF(M_{0:d-1},T_{d-1}) \underlabel_{\text{selection } d}
-  \sF(M_{0:d},T_{d-1}) 
-  \underlabel_{\text{inference } d}
-  \sF(M_{0:d}, T_d)
-\end{align}
-
-\end{comment}
-
-\begin{comment}
-\WFcomment{There is a filtration interpretation when you have the appropriate sufficiency properties.} Let $\sF_{k,i}$ denote the $\sigma$-algebra generated by $M_{0:k}$ and $p_{1:i}$.
-\begin{align*}
-  \sF_{k,i} &= \sF(M_{0:k},p_{1:i})\\
-  \sF_0 &\underlabel_{\text{selection } 1} \sF_{1,0} \underlabel_{\text{inference } 1}
-  \sF_{1,1} \;\;\sub \cdots \sub\;\;
-  \sF_{d-1,d-1} \underlabel_{\text{selection } d} \sF_{d,d-1}
-  \underlabel_{\text{inference } d} \sF_{d,d}
-\end{align*}
-\end{comment}

### 2015-11-09 paper.tex
replied to some of Jon's comments, changed next-$\lambda$ to next-entry

@@
 Extending the selected-model tests of \citet{fithian2014optimal}, we construct $p$-values for each step in the path, accounting for the fact that the model path is determined adaptively using the data.
-In the case of linear regression, we propose two specific tests, the max-$t$ test for forward stepwise regression (generalizing a proposal of \citet{buja2014}), and the next-$\lambda$ test for the lasso.
+In the case of linear regression, we propose two specific tests, the max-$t$ test for forward stepwise regression (generalizing a proposal of \citet{buja2014}), and the next-entry test for the lasso.
 These tests improve on the power of the spacings test of \citet{taylor2014exact}, sometimes dramatically.
@@
 
-For the remainder of the article we will discuss the problem of selective sequential model selection in some generality, returning periodically to the selective max-$t$ test in forward stepwise regression and its lasso regression counterpart, the next-$\lambda$ test, as intuitive and practically important running examples.
+For the remainder of the article we will discuss the problem of selective sequential model selection in some generality, returning periodically to the selective max-$t$ test in forward stepwise regression and its lasso regression counterpart, the next-entry test, as intuitive and practically important running examples.
 
@@
 
-\subsubsection{Ever-Active Regularization Paths and next-$\lambda$ Tests}
+\subsubsection{Ever-Active Regularization Paths and next-entry Tests}
 Another important class of model selection procedures is the sequence of {\em ever-active} sets for a regularized likelihood path. 
@@
 }
+\WFcomment{Took out some of the confusing language.
+We are explicitly defining $\ell$ as the likelihood for $M_\infty$ at the beginning of 2.2.1.
+I have a remark later that $\ell$ doesn't really need to be a likelihood, but we can't accommodate completely general loss functions without losing some of the results.
+What we do ``need'' for the SSP is that $\ell(\theta)$ is a function of $U(Y; E)$ for $\theta \in \Theta(E)$.
+This is true for likelihoods but probably not true in general for pseudo-likelihoods.
+I'm not sure why you say this is ``really'' about $\ell_1$: it's also about elastic net, group lasso, nuclear norm, $\ell_0$, and much more.
+True, $\ell_1$ is the prototypical example, but I don't see how we get anything from explicitly restricting ourselves to $\ell_1$.}
 
@@
 In this setting, $\lambda_k$ is a natural test statistic for model $M_{k-1}$, with larger values suggesting a poorer fit.
-The {\em selective next-$\lambda$ test} is the test that rejects for large $\lambda_k$, based on the law
+The {\em selective next-entry test} is the test that rejects for large $\lambda_k$, based on the law
 \begin{equation}
@@
 
-\JTcomment{The next-$\lambda$ test is actually a little expensive to compute if we a finite set of $\lambda$  as, say, glmnet does. Should we not mention this? On the plus side, 
+\JTcomment{The next-entry test is actually a little expensive to compute if we a finite set of $\lambda$  as, say, glmnet does. Should we not mention this? On the plus side, 
 it gives us tests that do not require LAR and can be used with glmnet. 
@@
 we restart glmnet and run until the active set changes, storing this
-value of $\lambda$. Then we compare our observed next-$\lambda$ to this
+value of $\lambda$. Then we compare our observed next-entry to this
 sample of $\lambda$'s.
@@
 In the LAR-LASSO case, this is easier as there is an explicit formula
-to compute the next event. But, LAR is limited to LASSO whereas next-$\lambda$
+to compute the next event. But, LAR is limited to LASSO whereas next-entry
 in principle works for group LASSO.
@@
 All this to say, that we are really only talking about $\ell_1$ here, and 
-next-$\lambda$ is essentially only practical BLAH
+next-entry is essentially only practical BLAH
 }
+\WFcomment{Actually I don't think the computation is bad at all.
+The point is that you don't need to compute the distribution of $\lambda_k$, you only need the probability, under the null hypothesis, that $\lambda_k$ is larger than its observed value.
+So at least for convex penalized-likelihood problems, you should only ever have to compute one path --- the path for the observed data set --- and then store those fits.
+And then when you resample, you just check whether $Y^*$ satisfies the KKT conditions at $\hat\theta^\lambda(Y)$ for each $\lambda > \lambda_k$.
+Then $\lambda_k(Y^*)>\lambda_k(Y)$ iff the KKT conditions are violated for any $\lambda > \lambda_k(Y)$.
+(It is probably worth pointing this out in the computation section somewhere).
+
+For group lasso, it's the same situation -- we should only need to compute one path.
+Though it sounds like you think in that case the sampling itself is a problem.
+But at least for non-overlapping groups, is it really so bad?
+
+I think it is worth distinguishing between what procedures we can define and prove things about (which we discuss here), and what subset of those we currently have efficient algorithms for (which we discuss in the Computation section).
+After all, some problems are small, some scientists are willing to wait a while for their results, and we have various cudgels at our disposal such as accept/reject, Besag \& Clifford MCMC tests that are exact even if the chain doesn't mix, etc.
+etc.}
 
@@
 
-\subsection{Independence for Greedy Likelihood and next-$\lambda$ $p$-values}\label{sec:SSP}
+\subsection{Independence for Greedy Likelihood and next-entry $p$-values}\label{sec:SSP}
 
 In this section we establish that in the generic sparse parametric model of Section~\ref{sec:genSparse}, the forward stepwise path and all ever-active regularization paths satisfy the SSP.
-What is more, the greedy likelihood ratio statistic $G_k$ is $\sF_k$-measurable with respect to the forward stepwise path's sufficient filtration, and the next-$\lambda$ statistic $\lambda_k$ is $\sF_k$-measurable with respect to the regularized-likelihood path's sufficient filtration.
-As a result, the selective greedy likelihood $p$-values in the forward stepwise path are independent on nulls, as are the next-$\lambda$ $p$-values in the regularized-likelihood path, per Theorem~\ref{thm:suffCond}.
+What is more, the greedy likelihood ratio statistic $G_k$ is $\sF_k$-measurable with respect to the forward stepwise path's sufficient filtration, and the next-entry statistic $\lambda_k$ is $\sF_k$-measurable with respect to the regularized-likelihood path's sufficient filtration.
+As a result, the selective greedy likelihood $p$-values in the forward stepwise path are independent on nulls, as are the next-entry $p$-values in the regularized-likelihood path, per Theorem~\ref{thm:suffCond}.
 
@@
 
-Analogous results hold for ever-active regularized likelihood paths and the next-$\lambda$ $p$-values.
+Analogous results hold for ever-active regularized likelihood paths and the next-entry $p$-values.
 
@@
 \begin{corollary}\label{cor:nextLambdaIndep}
-  If $U(Y; M)$ is complete sufficient for model $M$, the selective next-$\lambda$ $p$-values computed for any ever-active regularized likelihood path are independent on nulls.
+  If $U(Y; M)$ is complete sufficient for model $M$, the selective next-entry $p$-values computed for any ever-active regularized likelihood path are independent on nulls.
 \end{corollary}
@@
 As we have shown, the selective max-$t$ test is equivalent to the selective greedy likelihood test in the case of linear regression.
-In the next section, we will compare and contrast the max-$t$ and next-$\lambda$, and other {\em selected-model tests}, with the {\em saturated-model tests} proposed by \citet{taylor2014exact} and others.
+In the next section, we will compare and contrast the max-$t$ and next-entry, and other {\em selected-model tests}, with the {\em saturated-model tests} proposed by \citet{taylor2014exact} and others.
 
@@
 As long as $T(Y)$ is measurable with respect to $(M_k, U_k)$, and the path algorithm is subpath-sufficient, the $p$-values are guaranteed to be independent on nulls.
-The next-$\lambda$ test for the lasso and the max-$t$ test in forward-stepwise regression with unknown $\sigma^2$ are both examples of this approach.
+The next-entry test for the lasso and the max-$t$ test in forward-stepwise regression with unknown $\sigma^2$ are both examples of this approach.
 Note that under resampling from the law in~\eqref{eq:selModel_marg}, the selected variable $j_k$ is random; by contrast, \eqref{eq:selModel_cond} conditions on $j_k$.
@@
 LAR/lasso& Covariance test/Saturated&Spacing test/Saturated \\
-        & next-$\lambda$ test/Selected& \\
+        & next-entry test/Selected& \\
         \hline
@@
 Tests are described by the null hypothesis that they test (complete or incremental)  and the underlying model (saturated or selected).
-The max-$t$, max-$t$-identify and next-$\lambda$ tests are proposed in this paper, while the covariance test was proposed in \citet{lockhart2014significance}, and the spacing and Gaussian
+The max-$t$, max-$t$-identify and next-entry tests are proposed in this paper, while the covariance test was proposed in \citet{lockhart2014significance}, and the spacing and Gaussian
 pivot tests are given in \citet{taylor2014exact}. The latter paper shows that the covariance and spacing test are asymptotically equivalent for the first null step, when the complete null holds.}
@@
 We plot the $p$-values only for those realizations in which the correct model was selected in the first two steps (about 95\% of the time).
-We see that the next-$\lambda$ test is the most powerful, especially at step 1. The next-$\lambda$ and spacing test have the correct type I error in steps 3 and 4,
+We see that the next-entry test is the most powerful, especially at step 1. The next-entry and spacing test have the correct type I error in steps 3 and 4,
 while the covariance test becomes conservative in step 4, as we move further into the null part of the sequence.
@@
 
-The selected-model test at step 1 compares $Y_1$ to its distribution under $M_0$ conditional on $A_1$, a test of $H_0:\;\mu_1=0$ in model $M_1$.\footnote{In this very simple example, the max-$z$, max-$z$-identify, and next-$\lambda$ tests all coincide.} By contrast, the saturated-model test is a test of $H_0:\; \mu_1=0$ in the model $M_{\text{sat}}:\; Y \sim \cN(\mu, I_2)$.
+The selected-model test at step 1 compares $Y_1$ to its distribution under $M_0$ conditional on $A_1$, a test of $H_0:\;\mu_1=0$ in model $M_1$.\footnote{In this very simple example, the max-$z$, max-$z$-identify, and next-entry tests all coincide.} By contrast, the saturated-model test is a test of $H_0:\; \mu_1=0$ in the model $M_{\text{sat}}:\; Y \sim \cN(\mu, I_2)$.
 Thus, it must condition on $Y_2$ to eliminate the nuisance parameter $\mu_2$, comparing $Y_1$ to its null distribution given $A_1$ and the observed value of $Y_2$.
@@
 
-Until this point, we have deferred discussing the question of how to actually compute the rejection cutoffs for the max-$t$ test, next-$\lambda$ test, or any of the other tests we have discussed.
+Until this point, we have deferred discussing the question of how to actually compute the rejection cutoffs for the max-$t$ test, next-entry test, or any of the other tests we have discussed.
 In each case, the $p$-value at step $k$ comes from a test that rejects when some test statistic $T_k$ is large compared to the law

### 2015-11-10 paper.tex
more edits

@@
 
-To select a model, we can feed our single-step $p$-values as inputs into sequential stopping rules such as those proposed by \citet{gsell2013sequential} and \citet{li2015accumulation}, achieving control of the familywise error rate (FWER) or false discovery rate (FDR) as desired.
+To select a model, we can feed our single-step $p$-values as inputs into sequential stopping rules such as those proposed by \citet{gsell2013sequential} and \citet{li2015accumulation}, achieving control of the familywise error rate or false discovery rate (FDR) as desired.
 The FDR-controlling rules require the null $p$-values to be independent of each other and of the non-null $p$-values, a condition not satisfied by the saturated-model $p$-values of \citet{taylor2014exact}.
-We derive intuitive and general conditions for independence and show that our proposed constructions yield independent $p$-values.
-We also discuss connections between the selected-model test and parametric bootstrap sampling.
+We derive intuitive and general sufficient conditions for independence and show that our proposed constructions yield independent $p$-values.
 \end{abstract}
@@
 Both sets of rules require that once we have reached the first correct model in our path, the $p$-values in subsequent steps are uniform and independent.
-In this paper we develop a theoretical framework for constructing sequential $p$-values satisfying these properties, and show how to construct them in many settings.
+In this paper we develop a theoretical framework for constructing sequential $p$-values satisfying these properties, and give explicit constructions.
 
@@
 Forward stepwise regression is a greedy algorithm for building a sequence of nested linear regression models.
-At each iteration, it augments the current model by including the variable that will minimize the residual sum of squares of the next fitted model.
-Equivalently, it selects the variable with the largest $t$-statistic, controlling for the variables in the current model.
+At each iteration, it augments the current model by including the variable that will minimize the residual sum of squares (RSS) of the next fitted model.
+Equivalently, it selects the variable with the largest $t$-statistic, adjusting for the variables in the current model.
 
 For design matrix $X\in \R^{n\times p}$ and response vector $Y\in \R^n$, let $E \sub \{1, \ldots, d\}$ denote the current {\em active set}, the set of predictor variables already selected, and let $\RSS(E)$ denote the residual sum of squares for the corresponding regression model.
-For $j\notin E$ let $t_{j,E}(Y)$ denote the multivariate $t$-statistic of variable $j$, adjusting for the active variables and an intercept term.
+For $j\notin E$ let $t_{j,E}(Y)$ denote the multivariate $t$-statistic of variable $j$, adjusting for the active variables as well as an intercept term.
 Using the standard result that
@@
 with corresponding $t$-statistic $t_E^* = t_{j^*,E}$.
-The selective max-$t$ test rejects for large values of $t_E^*$, compared to an appropriate conditional distribution that accounts for the adaptive selection of the model path.
+The selective max-$t$ test rejects for large values of $|t_E^*|$, compared to an appropriate conditional null distribution that accounts for the adaptive selection of the model path.
 
@@
 Finally, we use the ForwardStop rule \citep{gsell2013sequential} at FDR level $\alpha=0.1$ to select a model based on the sequence of $p$-values.
-The red entry in each column indicates the last variable selected by ForwardStop.
+The bolded entry in each column indicates the last variable selected by ForwardStop.
 
@@
 The nominal $p$-values in Table~\ref{tab:diab} are computed by comparing $t_E^*$ to the $t$ distribution with $n-|E|-2$ degrees of freedom, which would be the correct distribution if the sequence of models were selected before observing the data.
-Because the model sequence is in fact selected adaptively to maximize the value of $t_E^*$, this method is highly anti-conservative.
+Because the model sequence is in fact selected adaptively to maximize the absolute value of $t_E^*$, this method is highly anti-conservative.
 
@@
 \begin{equation}
-  M(E):\; Y \sim N(X_E\beta, \sigma^2 I), \quad \beta\in \R^{|E|}, \sigma^2>0,
+  M(E):\; Y \sim \cN(X_E\beta_E, \sigma^2 I), \quad \beta_E\in \R^{|E|}, \sigma^2>0,
 \end{equation}
@@
 Because $U_E=\left(X_E'Y,\, \|Y\|^2\right)$ is the complete sufficient statistic for $M(E)$, we can sample from the conditional null distribution of $t_E^*$ given $U_E$ and the current active set $E$, using the {\em selected-model test} framework of \citet{fithian2014optimal}.
-In step one, because $E=\emptyset$ is fixed and $t_E^*$ is independent of $U_E$, the conditional test reduces to the max-$t$ test proposed in \citet{buja2014}.
-In later steps, $t_E^*$ and $U_E$ are not conditionally independent given $E$, but we can approximate the null conditional distribution through Monte Carlo sampling.
+In step one, because $E=\emptyset$ is fixed and $t_E^*$ is independent of $U_\emptyset=\|Y\|^2$, the conditional test reduces to the max-$t$ test proposed in \citet{buja2014}.
+In later steps, $t_E^*$ and $U_E$ are not conditionally independent given $E$, but we can numerically approximate the null conditional distribution through Monte Carlo sampling.
 
@@
 
-Section~\ref{sec:genericSetting} introduces the general problem setting, gives important definitions, and reviews relevant literature on testing ordered hypotheses using a sequence of $p$-values.
+Sections~\ref{sec:genericSetting}--\ref{sec:ordered} introduce the general problem setting, give important definitions, and review relevant literature on testing ordered hypotheses using a sequence of $p$-values.
 Many of these methods require the $p$-values to be uniform and independent under the null, and we derive general conditions to ensure this property in Section~\ref{sec:pValsIndep}.
-In Section~\ref{sec:selective-reg} we contrast selected-model and saturated-model tests for linear regression setting, and illustrate why selected-model tests are often much more powerful in early steps of the path.
+In Section~\ref{sec:selective-reg} we contrast selected-model and saturated-model tests in the linear regression setting, and explain why selected-model tests are often much more powerful in early steps of the path.
 In Section~\ref{sec:computation} we discuss strategies for sampling from the null distribution for our tests, and prove that in forward stepwise regression and $\ell_1$-regularized likelihood methods, the number of constraints in the sampling problem is never more than twice the number of variables.
-We provide a simulation study in Section~\ref{sec:sparseReg} and briefly discuss applications of our framework to decision trees, nonparametric changepoint detection, and principal components analysis in Section~\ref{sec:further-examples}.
+We provide a simulation study in Section~\ref{sec:sparseReg}, and Section~\ref{sec:further-examples} discusses applications of our framework beyond linear regression to decision trees nonparametric changepoint detection.
 The paper ends with a discussion.
@@
 
-In the general setting, we observe data $Y \in \cY$ with unknown sampling distribution $F$, and then apply some algorithm to generate an adaptive sequence of $d$ nested statistical models contained in some {\em upper model} $M_{\infty}$.
+In the general setting, we observe data $Y \in \cY$ with unknown sampling distribution $F$, and then apply some algorithm to generate an adaptive sequence of $d$ nested statistical models contained in some {\em upper model} $M_{\infty}$:
 \[
@@
 predictors allowed to have nonzero coefficients (but not the  values of their coefficients).
-Without loss of generality, we can take $M_\infty$ as the union of all models under consideration.
+Note the ``assumption'' of an upper model involves no loss of generality --- we can always take $M_\infty$ as the union of all models under consideration.
 We will use the notation $M_{0:d}$ to denote the sequence $M_0, \ldots, M_d$, which we call the {\em model path}.
@@
 By construction, $F\in M_k \iff k \geq k_0$.
-A {\em stopping rule} is any estimator $\hk(Y)$ of $k_0$, with $M_k$ ``rejected'' if $k < \hk$, and ``accepted'' otherwise.
+A {\em stopping rule} is any estimator $\hk$ of $k_0$, with $M_k$ ``rejected'' if $k < \hk$, and ``accepted'' otherwise.
 Because $\hk$ is the number of models we do reject, and $k_0$ is the number we should reject, the number of type I errors is $V=(\hk-k_0)_+$, while the number of type II errors is $(k_0-\hk)_+$.
-Depending on the scientific context, we might want to control the FWER $\P(V>0)$, the FDR $\E[V/\hk; \hk>0]$ , or another error rate defined by the expectation of some loss function $g(\hk, k_0)$:
+Depending on the scientific context, we might want to control the FWER $\P(V>0)$, the FDR $\E[V/\hk; \hk>0]$ , or another error rate such as a modified FDR, defined by the expectation of some loss function $g(\hk, k_0)$:
 \begin{equation}\label{eq:errRate}
@@
 
-Following \citet{fithian2014optimal}, we say that for a fixed candidate model $M$, the random variable $p_{k,M}(Y)$ is a valid {\em selective $p$-value} for $M$ at step $k$ if it is conservative under sampling from any $F\in M$, given that $M$ is selected.
+Following \citet{fithian2014optimal}, we say that for a fixed candidate null model $M$, the random variable $p_{k,M}(Y)$ is a valid {\em selective $p$-value} for $M$ at step $k$ if it is stochastically larger than uniform (henceforth super-uniform) under sampling from any $F\in M$, given that $M$ is selected.
 That is,
@@
 \]
-which is conservative on the event $\{F \in M_{k-1}(Y)\}$:
+which is super-uniform on the event $\{F \in M_{k-1}(Y)\}$:
 \begin{equation}\label{eq:selectiveGuaranteePk}
@@
 We will be especially interested in two generic path algorithms for the sparse parametric setting: forward stepwise paths and ever-active regularization paths.
+For a nonparametric example see Section~\ref{sec:further-examples}
 
@@
 
-\subsubsection{Ever-Active Regularization Paths and next-entry Tests}
+\subsubsection{Ever-Active Regularization Paths and Next-Entry Tests}
 Another important class of model selection procedures is the sequence of {\em ever-active} sets for a regularized likelihood path. 
@@
 	For some ordered set $\Lambda$, let $\lambda\in\Lambda$ parameterize the regularization penalty $P_\lambda(\theta)$.
-In the case of lasso-penalized regression, we could take $P_\lambda = \lambda\|\beta\|_1$ for $\lambda\in (0,\infty]$.
+In the case of lasso-penalized regression, we could take $P_\lambda = \lambda\|\beta\|_1$ for $\lambda\in (0,\infty)$.
 
@@
 This is true for likelihoods but probably not true in general for pseudo-likelihoods.
-I'm not sure why you say this is ``really'' about $\ell_1$: it's also about elastic net, group lasso, nuclear norm, $\ell_0$, and much more.
+I'm not sure why you say this is ``really'' about $\ell_1$: it's also about elastic net, group lasso, nuclear norm, $\ell_0$ (such as in PCA), and much more.
 True, $\ell_1$ is the prototypical example, but I don't see how we get anything from explicitly restricting ourselves to $\ell_1$.}
 
-Assuming for simplicity that there is a unique solution to each penalized problem of the form
+Assume for simplicity that there is a unique solution to each penalized problem of the form
 \begin{equation}\label{eq:pen-lik}
   \hat\theta^{\lambda}(Y) =
-  \argmin_{\theta\in\Theta} -\ell(\theta; Y) + P_\lambda(\theta),
-\end{equation}
-we define the ever-active set for $\lambda\in\Lambda$ as
+  \argmin_{\theta\in\Theta} -\ell(\theta; Y) + P_\lambda(\theta).
+\end{equation}
+It may be impossible or inconvenient to compute $\hat\theta^\lambda(Y)$ for every $\lambda\in(0,\infty)$.
+If so, we can instead take $\Lambda$ to be a grid of finitely many values.
+We define the {\em ever-active set} for $\lambda\in\Lambda$ as
 \begin{equation}
@@
 \end{equation}
-It may be impossible or inconvenient to compute $\hat\theta^\lambda(Y)$ for every $\lambda\in(0,\infty]$. If so, we can instead take $\Lambda$ to be a grid of finitely many values.
 
@@
 \[
-\hat\beta_{\text{lasso}}^\lambda = \frac{1}{2}\|Y-X\beta\|_2^2 + \lambda\|\beta\|_1.
-\]
-
+\hat\beta_{\text{lasso}}^\lambda = \frac{1}{2}\|Y-X\beta\|_2^2 + \lambda\|\beta\|_1,
+\]
+which is slightly different from the penalized log-likelihood criterion of \eqref{eq:pen-lik}.
+One can show that the lasso coefficients $\hat\beta_{\text{lasso}}^\lambda$, along with $\hat\sigma^2=n^{-1}\|Y-X\hat\beta^\lambda_{\text{lasso}}\|^2$, minimize
+\[
+\frac{n}{2\pi e}\exp\left\{-2\ell(\beta, \sigma^2;\,Y)/n\right\} + \lambda\|\beta\|_1.
+\]
+If $\Lambda=(0,\infty)$ then both methods result in exactly the same ever-active path.
+Otherwise, at the cost of some notational bloat, we could generalize \eqref{eq:pen-lik} by replacing $-\ell(\cdot)$ with $\rho(\ell(\cdot))$ for any strictly decreasing $\rho$, and the theory in this article would go through unchanged.
+
+\begin{comment}
 This is slightly different from the penalized log-likelihood criterion of \eqref{eq:pen-lik}, which would minimize
@@
 Otherwise, at the cost of some notational bloat, we could generalize \eqref{eq:pen-lik} by replacing $-\ell(\cdot)$ with $\rho(\ell(\cdot))$, where $\rho$ is any strictly decreasing function, and the theory in this article would go through unchanged.
+\end{comment}
 
@@
 \end{align*}
-
-As a result the lasso coefficients $\hat\beta_{\text{lasso}}^\lambda$ minimize
-\[
-\exp\{-2\ell(\beta,\sigma;\,Y)/n \,+\, \text{const.}\} + \lambda\|\beta\|_1,
-\]
-instead of the criterion in \eqref{eq:pen-lik}.
-If $\Lambda=(0,\infty)$ then both methods result in exactly the same ever-active path.
-Otherwise, we could ``fix'' this issue (at the cost of some notational bloat) by replacing $-\ell(\cdot)$ with $\rho(\ell(\cdot))$, where $\rho$ is any strictly decreasing function, and all of the theory in this article would go through unchanged.
 \end{comment}
@@
 
-Note that in our formulation of the problem, the type I error $V=(\hk-k_0)_+$ is defined in a ``model-centric'' fashion: at step $k$, we are testing  whether a particular linear model $M(E_{k-1})$ adequately describes the data $Y$.
+Note that in our formulation of the problem, the type I error $V=(\hk-k_0)_+$ is defined in a ``model-centric'' fashion: at step $k$ in linear regression, we are testing  whether a particular linear model $M(E_{k-1})$ adequately describes the data $Y$.
 Even if the next selected variable $X_{j_k}$ is a noise variable in the full model, it is not necessarily a mistake to reject $M(E_{k-1})$ provided there are other signal variables that have not yet been included.
@@
 where $X_E^+$ is the Moore-Penrose pseudoinverse of the matrix $X_E$.
-These are the true ``partial regression coefficients''  for design matrix $X_E$.
- Write $\proj_E\mu = X_E \theta^E$ for the projection of $\mu$ into the column space of $X_E$.
 
@@
 This procedure would control the complete-null and incremental-null FWER, but {\em not} the full-model-null FWER.
-The methods we consider in this article are essentially generalizations of this procedure for the setting where the model sequence is selected adaptively, and where we may wish to control FDR instead of FWER.
+The methods we consider in this article are essentially generalizations of this procedure to the setting where the model sequence is selected adaptively, and where we may wish to control FDR instead of FWER.
 
@@
 
-\section{Stopping Rules, Ordered Testing, and Independence}
+\section{Stopping Rules and Ordered Testing}\label{sec:ordered}
 
@@
 \begin{equation}\label{eq:basic-stop-control}
- \text{FWER} \leq \P(p_{k_0} \leq \alpha) \leq alpha.
+ \text{FWER} \leq \P(p_{k_0} \leq \alpha) \leq \alpha.
 \end{equation}
@@
 
-\citet{gsell2013sequential} propose the {\em ForwardStop} rule to controls the FDR:
+\citet{gsell2013sequential} propose the {\em ForwardStop} rule to control the FDR:
 \[
@@
 \]
-ForwardStop uses the $p$-values {\it before} step $k$ in making its decision whether to reject at step $k$.
 If $p_k$ is uniform, then $-\log(1-p_k)$ is an $\text{Exp}(1)$ random variable with expectation 1; thus, the sum can be seen as an estimate of the false discovery proportion (FDP):
 \[
-\widehat{\text{FDP}}_k = -\frac{1}{k}\sum_{i=1}^k \log(1-p_i).
-\]
-ForwardStop chooses the largest model for which $\widehat{\text{FDP}}_k \leq \alpha$.
+\widehat{\text{FDP}}_k = -\frac{1}{k}\sum_{i=1}^k \log(1-p_i),
+\]
+and we choose the {\em largest} model with $\widehat{\text{FDP}}_k \leq \alpha$.
 \citet{gsell2013sequential} show that ForwardStop controls the FDR if the null $p$-values are independent of each other and of the non-nulls.
@@
 In our problem setting, the sequence of selected models is random; thus, the truth or falsehood of each $H_k$ is not fixed, as assumed in the previous section.
-It is not immediately clear, then, how much we should condition on when we construct the $p$-value at step $k$.
-In particular, is it sufficient to condition only on the current null model $M_{k-1}$, or must we condition on the entire subpath of models $M_0,\ldots,M_{k-1}$?
+It is clear that we will recover all of the same guarantees if we construct our $p$-values conditional on the entire path $M_{0:d}(Y)$; however, this option is rather unappealing for both computational and statistical reasons.
+In constructing the $p$-value for step $k$, is it sufficient to condition only on the subpath of models $M_{0:k-1}(Y)$, or better yet to condition only on the current null model $M_{k-1}$?
 
@@
 Unfortunately, however, it is {\em not} enough to guarantee that the above stopping rules actually control FDR or FWER.
-
-Let $z_{\alpha/2}$ denote the upper-$\alpha/2$ quantile of the $\cN(0,1)$ distribution and consider the following counterexample:
+Let $z_{\alpha/2}$ denote the upper-$\alpha/2$ quantile of the $\cN(0,1)$ distribution, and consider the following counterexample:
 
 \begin{restatable}{proposition}{counterex}\label{prop:counterexample}
-  Consider linear regression with $n=p=3$, known $\sigma^2=1$ and identity design $X=I_3$.
-
-  Suppose that we construct the model path as follows: if $|Y_3|>z_{\alpha/2}$, we add $X_1$ to the active set first, then $X_2$, then $X_3$.
+  Consider linear regression with $n=p=3$, known $\sigma^2=1$ and identity design $X=I_3$, and suppose that we construct the model path as follows: if $|Y_3|>z_{\alpha/2}$, we add $X_1$ to the active set first, then $X_2$, then $X_3$.
 Otherwise, we add $X_2$, then $X_1$, then $X_3$.
@@
 
-  If $\mu=(0,C,0)$, then the FWER for the entire procedure tends to $2\alpha-\alpha^2$ as $C\to\infty$.
+  If $\mu=(0,C,0)$, then the FWER for this procedure tends to $2\alpha-\alpha^2$ as $C\to\infty$.
 \end{restatable}
@@
 \[
-\P(p_k \leq \alpha \mid H_k \text{ true}, k_0=k) \leq \alpha
+\P(p_k \leq \alpha \mid H_k \text{ true},\, k_0=k) \leq \alpha
 \]
@@
 Knowing $E_{k-1}=E$ tells us that the $k-1$ variables in $E$ are selected first, and knowing $(X_{E_{k-1}}'Y,\|Y\|^2)$ is enough information to compute the $t$-statistics $t_{j,D}$ for all $j\in E$ and $D\sub E$.
-As a result, we can reconstruct the order in which those $k-1$ variables were added, since for $s<k$,
-\[
-j_s = \argmax_{j \notin E_{s-1}} |t_{j,E_{s-1}}| 
-= \argmax_{j \in E\setminus E_{s-1}} |t_{j,E_{s-1}}| \quad \text{ on the event } \{E_{k-1}=E\}.
+As a result, we can reconstruct the order in which those $k-1$ variables were added, since for $i<k$,
+\[
+j_i = \argmax_{j \notin E_{i-1}} |t_{j,E_{i-1}}| 
+= \argmax_{j \in E\setminus E_{i-1}} |t_{j,E_{i-1}}| \quad \text{ on the event } \{E_{k-1}=E\}.
 \]
@@
 
-Furthermore, for $s<k$, $p_s$ is computed by comparing $T_s=|t_{j_s, E_{s-1}}|$ to the upper $\alpha$ quantile of the distribution $\L(T_s \mid E_{s-1}, U_{s-1})$, computed under the assumption $F\in M(E_{s-1})$.
-The conditional null distribution is a function of $(E_{s-1}, U_{s-1})$, and as we have already seen, the test statistic $T_s$ is a function of $(E_s, U_s)$, both of which are functions of $(E_{k-1},U_{k-1})$.
-
-Because $(E_s,U_s)$ is a function of $(E_k,U_k)$ for $s<k$, we can characterize the above discussion in terms of the filtration $\sF_{0:d}$, where
-\[
-\sF_k = \sF(E_k, U_k),
-\]
-the $\sigma$-algebra generated by random variables $(E_k, U_k)$. We could equivalently write $\sF_k=\sF(M_k, U_k)$, or $\sF_k=\sF(M_{0:k}, U_k)$.
+Furthermore, for $i<k$, $p_i$ is computed by comparing $T_i=|t_{j_i, E_{i-1}}|$ to the upper $\alpha$ quantile of the distribution $\L(T_i \mid E_{i-1}, U_{i-1})$, computed under the assumption $F\in M(E_{i-1})$.
+The conditional null distribution is a function of $(E_{i-1}, U_{i-1})$, and as we have already seen, the test statistic $T_i$ is a function of $(E_i, U_i)$, both of which are functions of $(E_{k-1},U_{k-1})$.
+
+Because $(E_i,U_i)$ is a function of $(E_k,U_k)$ for $i<k$, we can characterize the above discussion in terms of the filtration $\sF_{0:d}$, where
+\[
+\sF_k = \sF(E_{0:k}, U_k),
+\]
+the $\sigma$-algebra generated by random variables $(E_k, U_k)$. When the SSP is in force, we can equivalently write $\sF_k=\sF(E_k, U_k)$, or $\sF_k=\sF(M_k, U_k)$.
 
@@
 \[
-\P_F(p_k \leq \alpha \mid M_{k-1},\; F\in M_{k-1}) \leqAS \alpha.
+\P_F(p_k \leq \alpha \mid M_{k-1}) \leqAS \alpha \quad \text{ on } \{F\in M_{k-1}\}
 \]
@@
 \item $M_k(Y)$ and $p_k(Y)$ are measurable with respect to $\sF_k$, and 
-\item $p_k(Y)$ is uniform or conservative given $\sF_{k-1}$ on the event $\{F\in M_{k-1}(Y)\}$.
+\item $p_k(Y)$ is uniform or super-uniform given $\sF_{k-1}$ on the event $\{F\in M_{k-1}(Y)\}$.
 \end{enumerate}
@@
   independent and uniform given $\sF_{k_0}$ on the event $\{k_0=k\}$.
-  If they are conservative, then for all
+  If they are super-uniform, then for all
   $\alpha_{k+1},\ldots,\alpha_d \in [0,1]$,
@@
       \mid \sF_k\right]1_{\{k_0 \leq k+1\}}\prod_{i=k+2}^d \alpha_i
-    \quad\;\;\leqAS\;\; \prod_{i=k+1}^d \alpha_i
+    &\leqAS \prod_{i=k+1}^d \alpha_i
   \end{align*}
@@
 Define the {\em sufficient filtration} for the path $M_{0:d}$ as the filtration $\sF_{0:d}$ with $\sF_k = \sF(M_{0:k}, U_k)$.
-By our assumption of minimal sufficiency, $U_s$ is always a function of $(M_s, U_k)$ for $s<k$.
+By our assumption of minimal sufficiency, $U_i$ is always a function of $(M_i, U_i)$ for $i<k$.
 The path algorithm satisfies the SSP if and only if
@@
 \begin{enumerate}
-\item $p_k$ is conservative given $M_{0:k-1}$ and $U_{k-1}$, and \RTcomment{clearer to say ``sub-uniform''?} \JTcomment{I think you mean ``super-uniform'', i.e.
-it is larger than a uniform random variable?}
+\item $p_k$ is super-uniform given $M_{0:k-1}$ and $U_{k-1}$, and 
 \item $p_k$ is a function of $M_{0:k}$ and $U_k$.
@@
 
-To be a valid selective $p$-value for a single step, $p_k$ only needs to be conservative conditional on $M_{k-1}$.
+To be a valid selective $p$-value for a single step, $p_k$ only needs to be super-uniform conditional on $M_{k-1}$.
 In general, it is more stringent to additionally require that $p_k$ must condition on the entire subpath $M_{0:k-1}$ as well as $U_{k-1}$.
-In most cases of interest, however, (1) we will condition on $(M_{k-1}, U_{k-1})$ anyway, to eliminate nuisance parameters, and (2) the path satisifies the SSP, so conditioning on $(M_{k-1}, U_{k-1})$ is equivalent to conditioning on $(M_{k-1}, U_{k-1})$.
+In most cases of interest, however, (1) we will want to condition on $(M_{k-1}, U_{k-1})$ anyway, to eliminate nuisance parameters of $M_{k-1}$, and (2) the path satisifies the SSP, so conditioning on $(M_{k-1}, U_{k-1})$ is equivalent to conditioning on $(M_{k-1}, U_{k-1})$.
 Thus, most single-step selective $p$-values we would contemplate using are already exact conditional on $\sF_{k-1}$.
@@
 
-\subsection{Independence for Greedy Likelihood and next-entry $p$-values}\label{sec:SSP}
-
-In this section we establish that in the generic sparse parametric model of Section~\ref{sec:genSparse}, the forward stepwise path and all ever-active regularization paths satisfy the SSP.
+\subsection{Independence for Greedy Likelihood and Next-Entry $p$-values}\label{sec:SSP}
+
+In this section we apply Theorem~\ref{thm:suffCond} to establish that in the generic sparse parametric model of Section~\ref{sec:genSparse}, the forward stepwise path and all ever-active regularization paths satisfy the SSP.
 What is more, the greedy likelihood ratio statistic $G_k$ is $\sF_k$-measurable with respect to the forward stepwise path's sufficient filtration, and the next-entry statistic $\lambda_k$ is $\sF_k$-measurable with respect to the regularized-likelihood path's sufficient filtration.
@@
 }
+\WFcomment{Yes, it is assumed at the beginning of the paper that $U$ is minimal sufficient.
+Perhaps I should just say complete sufficient at the beginning so I don't need to strengthen the assumption later...}
 
@@
   \includegraphics[width=1\textwidth]{figs/simulation_snr_5_alpha_05_null_false.pdf}
-  \caption{\em  CDFs of nominal (black), MaxZ (red), MaxT (green) and  saturated-model (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on testing a false null hypothesis at step $k$.
-The MaxT approaches are much more powerful than the saturated-model test.
+  \caption{\em  CDFs of nominal (black), max-$z$ (red), max-$t$ (green) and  saturated-model (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on testing a false null hypothesis at step $k$.
+The max-$t$ approaches are much more powerful than the saturated-model test.
 The nominal test appears to be powerful,
@@
   \includegraphics[width=1\textwidth]{figs/simulation_snr_5_alpha_05_null_true.pdf}
-  \caption{\em  CDFs of nominal (black), MaxZ (red), MaxT (green) and  saturated-model (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on testing a true null hypothesis at step $k$.
+  \caption{\em  CDFs of nominal (black), max-$z$ (red), max-$t$ (green) and  saturated-model (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on testing a true null hypothesis at step $k$.
 The nominal test is badly anti-conservative, while all of the other methods  show uniform $p$-values as desired.}
@@
   \includegraphics[width=1\textwidth]{figs/simulation_snr_5_alpha_05_noise_var.pdf}
-  \caption{\em CDFs of nominal (black), MaxZ (red), MaxT $\sigma$ (green) and  saturated-model (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on the event that the variable added at step $k$ is a noise variable in the full model.
+  \caption{\em CDFs of nominal (black), max-$z$ (red), max-$t$ $\sigma$ (green) and  saturated-model (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on the event that the variable added at step $k$ is a noise variable in the full model.
 Here, none of the methods produce uniform $p$-values.
@@
 \]
-where $p_{k,E}(Y)$ for each $E\sub \{1,\ldots, p\}$ is a conditional $p$-value for the {\em fixed} null model $M(E):\, Y \sim N(X_E\beta, \sigma^2)$, valid on the selection event $A=\{E_{k-1}(Y)=E\}$.
+where $p_{k,E}(Y)$ for each $E\sub \{1,\ldots, p\}$ is a conditional $p$-value for the {\em fixed} null model $M(E):\, Y \sim \cN(X_E\beta, \sigma^2)$, valid on the selection event $A=\{E_{k-1}(Y)=E\}$.
 

### 2015-11-10 paper.tex
discussion of which penalties we handle

@@
 
+\JTcomment{I agree that it is worth distinguishing ``provable''
+and ``computable''.
+I agree that the collection of things we can prove things about
+includes group LASSO perhaps other many penalties, but
+I would say
+that we only have a practical algorithm for computing
+when the penalty is $\ell_1$ in this paper. I'm not sure exactly
+how to fit nuclear norm here because I'm not sure what
+we should call the  ever active set when the loss is not Frobenius (i.e. unless
+we are doing PCA or SVT). That is, in matrix
+completion or reduced-rank regression I'm not sure what I would call the 
+ever-active set. For PCA or SVT, the spaces produced by the data's left and
+right singular vectors make up (a reasonable version of an) ever-active set.
+
+I also understand your argument that for $\ell_1$  the next entry test is less bad than I thought. 
+To check the KKT conditions at $\hat{\theta}^{\lambda}(Y)$ for $\ell_1$ is 
+straightforward, as $\hat{\theta}^{\lambda}(Y^*)$ has 
+an explicit candidate form so we have a very good guess
+at what the residuals at slightly later $\lambda$'s should look like and we
+can check the KKT conditions easily.
+For group LASSO, things are more complicated
+as we probably will not want to condition on the ``signs'' which are now
+a collection of unit vectors. If we don't condition on them
+then we have to solve group LASSO for response $Y^*$ until the set of
+active groups changes.
+Even if we do condition on them (which is OK
+as they are measurable with respect the right $\sigma$-algebra) the simplification for sampling that we observe for $\ell_1$ goes away. After $k$ steps on the
+discrete $\Lambda$ path I think we are back to having 
+$O(p*k)$ constraints roughly for any penalty that is not $\ell_1$. (For group
+LASSO, the $O(p*k)$ estimate assumes $k$ is not huge, and each group has roughly bounded number of features in it.)}
+
 \paragraph{Remark} In its usual formulation, the lasso coefficients for linear regression minimize a penalized RSS criterion:

### 2015-11-10 paper.tex
one more comment on sampling group LASSO conditional on directions

@@
 $O(p*k)$ constraints roughly for any penalty that is not $\ell_1$. (For group
-LASSO, the $O(p*k)$ estimate assumes $k$ is not huge, and each group has roughly bounded number of features in it.)}
+LASSO, the $O(p*k)$ estimate assumes $k$ is not huge, and each group has roughly bounded number of features in it.) Finally, conditioning on the
+direction vectors needs a sampling algorithm that is not as simple as
+Gaussian subject to affine constraints. Describing this sampler is something
+I think we just don't want to touch in this paper.
+}
 

### 2015-11-22 paper.tex
clarifying conditional expectations?
fixing one of the conditional expectations
notation of k_0
Made a bunch of edits and comments.

@@
 \newcommand{\RTcomment}[1]{{\color{green}{(RT: \bf \sc #1) }}}
-
+\newcommand{\RJTcomment}[1]{{\color{magenta}{(RJT: \bf \sc #1) }}}
 \title{Selective Sequential Model Selection}
-\author{William Fithian, Jonathan Taylor, Rob Tibshirani, and Ryan Tibshirani}
+\author{William Fithian, Jonathan Taylor, Rob Tibshirani, and Ryan J. Tibshirani} 
 \maketitle
@@
 In the case of linear regression, we propose two specific tests, the max-$t$ test for forward stepwise regression (generalizing a proposal of \citet{buja2014}), and the next-entry test for the lasso.
-These tests improve on the power of the spacings test of \citet{taylor2014exact}, sometimes dramatically.
+These tests improve on the power of the saturated-model test of \citet{taylor2014exact}, sometimes dramatically.
 In addition, our framework extends beyond linear regression to a much more general class of parametric and nonparametric model selection problems.
@@
 The FDR-controlling rules require the null $p$-values to be independent of each other and of the non-null $p$-values, a condition not satisfied by the saturated-model $p$-values of \citet{taylor2014exact}.
-We derive intuitive and general sufficient conditions for independence and show that our proposed constructions yield independent $p$-values.
+We derive intuitive and general sufficient conditions for independence, and show that our proposed constructions yield independent $p$-values.
 \end{abstract}
@@
 Given such a path, we consider the problem of choosing the simplest model in the path that is not falsified by the available data.
-Examples of path algorithms include forward stepwise linear regression, least angle regression (LAR) and  the ever-active path in $\ell_1$-regularized models.
+Examples of path algorithms include forward stepwise linear regression, least angle regression (LAR) and the ever-active path in lasso ($\ell_1$-regularized) regression.
 \citet{taylor2014exact} study methods for generating exact $p$-values at each step of these path algorithms, and their methods provide a starting point for our proposals.
-In addition, \citet{choi2014selecting} discuss a selective approach to estimating the rank of the covariance matrix in principle components analysis.
-
-To select a model, we compute a set of sequential $p$-values at each step of the model path, and then feed them into a stopping rule that is guaranteed to control the false discovery rate (FDR) \citep{benjamini1995controlling}, familywise error rate (FWER), or a similar quantity.
+Other related works include \citet{loftus2014significance}, who describe $p$-values for path algorithms that enter groups of variables (instead of individual variables) at each step, and \citet{choi2014selecting}, who describe $p$-values for steps of principle components analysis (each step marking the estimation of a principle component direction).
+
+We consider the following workflow: to select a model, we compute a set of sequential $p$-values at each step of the model path, and then feed them into a stopping rule that is guaranteed to control the false discovery rate (FDR) \citep{benjamini1995controlling}, familywise error rate (FWER), or a similar quantity.
 Recently \citet{gsell2013sequential} and \citet{li2015accumulation} proposed sequential stopping rules of this kind.
 Both sets of rules require that once we have reached the first correct model in our path, the $p$-values in subsequent steps are uniform and independent.
-In this paper we develop a theoretical framework for constructing sequential $p$-values satisfying these properties, and give explicit constructions.
-
-While our approach and analysis are quite general, we begin by introducing a specific example: the selective max-$t$ test for forward stepwise regression, a selective sequential version of the max-$t$ test of \citet{buja2014}, generalized using the theory in \citet{fithian2014optimal}.
+While this is not true of the $p$-values constructed in \citet{taylor2014exact}, in this paper we develop a theoretical framework for constructing sequential $p$-values satisfying these properties, and give explicit constructions.
+
+Our approach and analysis are quite general, but we begin by introducing a specific example: the selective max-$t$ test for forward stepwise regression.
+This is a selective sequential version of the max-$t$ test of \citet{buja2014}, generalized using the theory in \citet{fithian2014optimal}.
 
@@
 
-For design matrix $X\in \R^{n\times p}$ and response vector $Y\in \R^n$, let $E \sub \{1, \ldots, d\}$ denote the current {\em active set}, the set of predictor variables already selected, and let $\RSS(E)$ denote the residual sum of squares for the corresponding regression model.
+For a design matrix $X\in \R^{n\times p}$, with columns $X_1,\ldots X_p \in \R^n$, and a response vector $Y\in \R^n$, let $E \sub \{1, \ldots, p\}$ denote the current {\em active set}, the set of predictor variables already selected, and let $\RSS(E)$ denote the residual sum of squares for the corresponding regression model.
 For $j\notin E$ let $t_{j,E}(Y)$ denote the multivariate $t$-statistic of variable $j$, adjusting for the active variables as well as an intercept term.
 Using the standard result that
-\begin{equation}
+\begin{equation*}
 t_{j,E}^2 = (n-|E|-2) \frac{\RSS(E) - \RSS(E \cup \{j\})}{\RSS(E \cup \{j\})},
-\end{equation}
+\end{equation*}
 we see that the next variable selected is 
-\begin{equation}
-j^* = \argmax_{j\notin E} |t_{j,E}| = \argmin_{j\notin E} \RSS(E \cup \{j\}),
-\end{equation}
+\begin{equation*}
+j^* = \argmax_{j\notin E} \, |t_{j,E}| = \argmin_{j\notin E} \, \RSS(E \cup \{j\}), 
+\end{equation*}
 with corresponding $t$-statistic $t_E^* = t_{j^*,E}$.
@@
 Table~\ref{tab:diab} illustrates the max-$t$ test and two others applied to the diabetes data from \cite{LARS}, consisting of observations of $n=442$ patients.
-The response of interest is a quantitative measure of disease progression one year after baseline, and there are ten measured predictors --- age, sex, body-mass index, average blood pressure, and six blood serum measurements --- plus quadratic terms, giving a total of $d=64$ features.
+The response of interest is a quantitative measure of disease progression one year after baseline, and there are ten measured predictors --- age, sex, body-mass index, average blood pressure, and six blood serum measurements --- plus quadratic terms, giving a total of $p=64$ features.
 We apply forward stepwise regression to generate the model path (beginning with an intercept term), and then use each of the three methods to generate $p$-values at each step along the path.
@@
 \caption[tab:diab]{\em Forward stepwise regression for the diabetes data: naive $p$-values, $p$-values from the saturated model, and our max-$t$ $p$-values.
-The {\bf bold annotation} indicates the model selected by ForwardStop with $\alpha=0.1$.
-For the max-$t$ $p$-values, this stopping rule gives exact $\text{FDR}_{\text{model}}$ control at the 10\% level.}
+The bold annotation in each column marks the model selected by ForwardStop with $\alpha=0.1$.
+For the max-$t$ $p$-values, this stopping rule gives exact FDR control at the 10\% level.}
 \label{tab:diab}
@@
 
-The nominal $p$-values in Table~\ref{tab:diab} are computed by comparing $t_E^*$ to the $t$ distribution with $n-|E|-2$ degrees of freedom, which would be the correct distribution if the sequence of models were selected before observing the data.
+The nominal $p$-values in column 1 of Table~\ref{tab:diab} are computed by comparing $t_E^*$ to the $t$-distribution with $n-|E|-2$ degrees of freedom, which would be the correct distribution if the sequence of models were selected before observing the data.
 Because the model sequence is in fact selected adaptively to maximize the absolute value of $t_E^*$, this method is highly anti-conservative.
 
-The max-$t$ test $p$-values are computed using the same test statistic $|t_E^*|$, but compared with a more appropriate null distribution.
-We can simulate the distribution of $|t_E^*|$ under the null model, i.e.
-the linear regression model specified by active set $E$:
-\begin{equation}
-  M(E):\; Y \sim \cN(X_E\beta_E, \sigma^2 I), \quad \beta_E\in \R^{|E|}, \sigma^2>0,
-\end{equation}
+The max-$t$ test $p$-values in column 3 are computed using the same test statistic $|t_E^*|$, but compared with a more appropriate null distribution.
+We can simulate the distribution of $|t_E^*|$ under the null model, i.e., the linear regression model specified by active set $E$:
+\begin{equation*}
+M(E):\; Y \sim \cN(X_E\beta_E, \sigma^2 I_n), \quad \beta_E\in \R^{|E|}, \sigma^2>0,
+\end{equation*}
 where $X_E$ is the matrix with columns $(X_j)_{j\in E}$.
-Because $U_E=\left(X_E'Y,\, \|Y\|^2\right)$ is the complete sufficient statistic for $M(E)$, we can sample from the conditional null distribution of $t_E^*$ given $U_E$ and the current active set $E$, using the {\em selected-model test} framework of \citet{fithian2014optimal}.
+As $U_E=\left(X_E'Y,\, \|Y\|^2\right)$ is the complete sufficient statistic for $M(E)$, we can sample from the conditional null distribution of $t_E^*$ given $U_E$ and the current active set $E$, using the {\em selected-model test} framework of \citet{fithian2014optimal}.
 In step one, because $E=\emptyset$ is fixed and $t_E^*$ is independent of $U_\emptyset=\|Y\|^2$, the conditional test reduces to the max-$t$ test proposed in \citet{buja2014}.
@@
 
-The saturated-model $p$-values also account for selection, but they rely on somewhat different assumptions, condition on more information, and test a slightly different null hypothesis.
+The saturated-model $p$-values in column 2 also account for selection, but they rely on somewhat different assumptions, condition on more information, and test a slightly different null hypothesis.
 We discuss these distinctions in detail in Section~\ref{sec:whichnull}.
@@
 As we will discuss in Section~\ref{sec:selective-reg}, selected-model tests such as the max-$t$ can be much more powerful than saturated-model tests in early steps of the path when multiple strong variables are competing with each other to enter the model first.
-Using the max-$t$ $p$-values, ForwardStop selects a model of size 8, compared to size 3 when using the saturated-model $p$-values.
-
-Because the null max-$t$ $p$-values are independent, while the saturated-model $p$-values are not, ForwardStop is guaranteed to control FDR when using the former but not the latter.
+In the diabetes example, using the max-$t$ $p$-values, ForwardStop selects a model of size 8, compared to size 3 when using the saturated-model $p$-values.
+
+The null max-$t$ $p$-values are independent, while the saturated-model $p$-values are not, and hence ForwardStop is guaranteed to control FDR when using the former but not the latter.
 In Section~\ref{sec:pValsIndep} we discuss intuitive sufficient conditions for independence and show that the max-$t$ and many other selected-model tests satisfy these conditions.
@@
 
-Sections~\ref{sec:genericSetting}--\ref{sec:ordered} introduce the general problem setting, give important definitions, and review relevant literature on testing ordered hypotheses using a sequence of $p$-values.
+Sections~\ref{sec:genericSetting} and \ref{sec:ordered} introduce the general problem setting, give important definitions, and review relevant literature on testing ordered hypotheses using a sequence of $p$-values.
 Many of these methods require the $p$-values to be uniform and independent under the null, and we derive general conditions to ensure this property in Section~\ref{sec:pValsIndep}.
 In Section~\ref{sec:selective-reg} we contrast selected-model and saturated-model tests in the linear regression setting, and explain why selected-model tests are often much more powerful in early steps of the path.
-In Section~\ref{sec:computation} we discuss strategies for sampling from the null distribution for our tests, and prove that in forward stepwise regression and $\ell_1$-regularized likelihood methods, the number of constraints in the sampling problem is never more than twice the number of variables.
-We provide a simulation study in Section~\ref{sec:sparseReg}, and Section~\ref{sec:further-examples} discusses applications of our framework beyond linear regression to decision trees nonparametric changepoint detection.
+In Section~\ref{sec:computation} we discuss strategies for sampling from the null distribution for our tests, and prove that in forward stepwise regression and lasso ($\ell_1$-regularized) regression, the number of constraints in the sampling problem is never more than twice the number of variables.
+We provide a simulation study in Section~\ref{sec:sparseReg}, and Section~\ref{sec:further-examples} discusses applications of our framework beyond linear regression to decision trees and nonparametric changepoint detection.
 The paper ends with a discussion.
@@
 \[
-M_0(Y) \sub M_1(Y) \sub \cdots \sub M_d(Y) \sub M_\infty.
-\]
+M_0 \sub M_1(Y) \sub \cdots \sub M_d(Y) \sub M_\infty.
+\]
+\RJTcomment{Why does $M_0(Y)$ depend on $Y$ but $M_\infty$ does not?
+I'm thinking of $M_0(Y)$ as the model with intercept, no?
+For simplicity I've removed the dependence of $M_0$ on $Y$, but if I'm wrong about this, sorry, just change it back.} \\
 By {\em model}, we mean a family of candidate probability distributions for $Y$. For example in  linear regression, a model specifies  the set of
@@
 \]
-
 Given data $Y$ and model path $M_{0:d}$, we set ourselves the formal goal of selecting the simplest correct model: the smallest $k$ for which $F\in M_k(Y)$.
@@
 
-Define the {\em completion index} $k_0(Y) = \min\{k:\; F \in M_k(Y)\}$, the index of the first correct model.
+\JTcomment{Maybe we should write $k_F$ instead of $k_0$? Or $k_0(Y,F)$?}
+
+Define the {\em completion index} by $k_0(Y) = \min\{k:\; F \in M_k(Y)\}$, the index of the first correct model.
 By construction, $F\in M_k \iff k \geq k_0$.
-A {\em stopping rule} is any estimator $\hk$ of $k_0$, with $M_k$ ``rejected'' if $k < \hk$, and ``accepted'' otherwise.
-Because $\hk$ is the number of models we do reject, and $k_0$ is the number we should reject, the number of type I errors is $V=(\hk-k_0)_+$, while the number of type II errors is $(k_0-\hk)_+$.
-Depending on the scientific context, we might want to control the FWER $\P(V>0)$, the FDR $\E[V/\hk; \hk>0]$ , or another error rate such as a modified FDR, defined by the expectation of some loss function $g(\hk, k_0)$:
+A {\em stopping rule} is any estimator \smash{$\hk$} of $k_0$, with $M_k$ ``rejected'' if $k < \hk$, and ``accepted'' otherwise.
+Because \smash{$\hk$} is the number of models we do reject, and $k_0$ is the number we should reject, the number of type I errors is \smash{$V=(\hk-k_0)_+$}, while the number of type II errors is \smash{$(k_0-\hk)_+$}.
+Depending on the scientific context, we might want to control the FWER: $\P(V>0)$, the FDR: \smash{$\E[V/\hk; \hk>0]$}, or another error rate such as a modified FDR, defined by the expectation of some loss function \smash{$g(\hk, k_0)$}:
 \begin{equation}\label{eq:errRate}
-\Err_F(\hk(\cdot), g) = \E_F\left[ g\left(\hk(Y), k_0(Y)\right)\right]
+\Err_F(\hk(\cdot), g) = \E_F\left[ g\left(\hk(Y), k_0(Y)\right)\right].
 \end{equation}
@@
 That is,
-\begin{equation}
+\begin{equation*}
 \P_F\left(p_{k,M}(Y) \leq \alpha \mid M_{k-1}(Y) = M\right) 
-\leq \alpha, \quad \forall F\in M, \alpha \in [0,1].
-\end{equation}
-
+\leq \alpha, \quad \forall F\in M, \; \alpha \in [0,1].
+\end{equation*}
 Once we have constructed selective $p$-values for each fixed $M$, we can use them as building blocks to construct a combined $p$-value for the random null $M_{k-1}(Y)$. Define
@@
 \end{equation}
-
 One useful strategy for constructing exact selective tests is to condition on the sufficient statistics of the null model $M_{k-1}$. By sufficiency, the law
@@
 Thus, we can construct selective tests by comparing the realized value of any test statistic $T(Y)$ to its known conditional distribution under the null.
-All that remains is to choose the test statistic and compute its conditional null distribution, which can be challenging.
+What remains is to choose the test statistic and compute its conditional null distribution, which can be challenging.
 See \citet{fithian2014optimal} for a general treatment.
@@
 
-Many familiar path algorithms including lasso, LARS, and forward stepwise regression, are methods for adaptively selecting a set of predictors in linear regression models where we observe a random response $Y\in \R^n$ as well as a fixed design matrix $X\in \R^{n \times p}$, whose columns correspond to candidate predictors.
+Many familiar path algorithms, including forward stepwise regression, least angle regression (LAR), and the lasso regression (thought of as producing a path of models over its regularization parameter $\lambda$), are methods for adaptively selecting a set of predictors in linear regression models where we observe a random response $Y\in \R^n$ as well as a fixed design matrix $X\in \R^{n \times p}$, whose columns correspond to candidate predictors.
 For each possible active set $E \sub \{1,\ldots,p\}$, there is a corresponding candidate model
@@
 
-For the most part, we can discuss our theory and proposals in a generic parametric setting generalizing the linear regression problem above.
-Let $M_\infty$ be a model parameterized by $\theta\in \Theta \sub \R^{\cJ}$:
+For the most part, we can discuss our theory and proposals in a parametric setting generalizing the linear regression problem above.
+Let $M_\infty$ be a model parameterized by $\theta\in \Theta \sub \R^J$:
 \[
@@
 \]
-For any subset $E\sub \cJ$ define the sparse submodel with active set $E$ as follows:
+For any subset $E\sub \{1,\ldots J\}$ define the sparse submodel with active set $E$ as follows:
 \[
@@
 \]
-
 We can consider path algorithms that return a sequence of nested active sets
 \[
-E_0(Y) \sub E_1(Y) \sub \cdots \sub E_d(Y) \sub \cJ,
-\]
+E_0 \sub E_1(Y) \sub \cdots \sub E_d(Y) \sub \{1,\ldots J\},
+\]
+\RJTcomment{Again here I've removed the dependence of $E_0$ on $Y$.} \\
 inducing a model path with $M_k = M(E_k),$ for $k=0,\ldots,d$.
 We will be especially interested in two generic path algorithms for the sparse parametric setting: forward stepwise paths and ever-active regularization paths.
-For a nonparametric example see Section~\ref{sec:further-examples}
+For a nonparametric example see Section~\ref{sec:further-examples}.
 
@@
 Let $\ell(\theta; Y)$ denote the log-likelihood for model
-$M_\infty$. The {\em forward stepwise} algorithm proceeds as follows: we begin with some fixed $E_0$ (such as an intercept-only model), then at step $k=1,\ldots,d$, we set
+$M_\infty$. A generic {\em forward stepwise} algorithm proceeds as follows: we begin with some fixed set $E_0$ (such as an intercept-only model), then at step $k=1,\ldots,d$, we set
 \begin{equation}
-j_k = \argmax_j \;\;\sup \left\{\ell(\theta; Y):\; \theta\in\Theta(E_{k-1} \cup \{j\})\right\}, \quad \text{ and } \quad
+j_k = \argmax_j \; \sup \left\{\ell(\theta; Y):\; \theta\in\Theta(E_{k-1} \cup \{j\})\right\}, \quad \text{ and } \quad
 E_k = E_{k-1} \cup \{j_k\}.
@@
 \end{equation}
-
 The {\em selective greedy likelihood ratio test} rejects for large $G_k$, based on the law
-\begin{equation}
+\begin{equation*}
 \L\left(G_k \;\mid\; M_{k-1}, U_{k-1}\right)
-\end{equation}
-
+\end{equation*}
 Because the likelihood in linear regression is a monotone decreasing function of the residual sum of squares, the max-$t$ test is equivalent to the greedy likelihood ratio test.
@@
 \subsubsection{Ever-Active Regularization Paths and Next-Entry Tests}
-Another important class of model selection procedures is the sequence of {\em ever-active} sets for a regularized likelihood path. 
+Another important class of model selection procedures is the sequence of {\em ever-active} sets for a regularized likelihood path, under a sparsity inducing regularizer.
 The notion of ever-active sets is needed since these solution paths can drop  (as well as add) predictors along the way.
 	For some ordered set $\Lambda$, let $\lambda\in\Lambda$ parameterize the regularization penalty $P_\lambda(\theta)$.
-In the case of lasso-penalized regression, we could take $P_\lambda = \lambda\|\beta\|_1$ for $\lambda\in (0,\infty)$.
+As an example, for a lasso penalty, this is $P_\lambda = \lambda\|\beta\|_1$ with $\lambda\in (0,\infty)$.
 
@@
 BUT later on we implicitly assume that it is the likelihood for 
-$M_{\infty}$. Why not just say this here, i.e. when talking
+$M_{\infty}$. Why not just say this here, i.e., when talking
 about $\ell_1$ regularization paths the models we have in mind are typically
@@
 
+\RJTcomment{Hopefully without too much drama :), I've added the words ``sparsity-inducing regularizer'' above.
+This is because, at least as we've written it, we are thinking of the regularizer as selecting out a submodel.
+I understand that it may be more general than this, but our notation above and in the previous subsection all have to do with sparsity directly.}
+
 Assume for simplicity that there is a unique solution to each penalized problem of the form
@@
 \end{equation}
-It may be impossible or inconvenient to compute $\hat\theta^\lambda(Y)$ for every $\lambda\in(0,\infty)$.
-If so, we can instead take $\Lambda$ to be a grid of finitely many values.
+It may be impossible or inconvenient to compute \smash{$\hat\theta^\lambda(Y)$} for every $\lambda\in(0,\infty)$. If so, we can instead take $\Lambda$ to be a grid of finitely many values.
+
 We define the {\em ever-active set} for $\lambda\in\Lambda$ as
@@
 \end{equation}
-
-Note that the ever-active sets $\tE_\lambda$ are nested by construction.
-In addition, we will assume that $|\tE_\lambda|<\infty$ for every $\lambda\in\Lambda$.
+Note that the ever-active sets \smash{$\tE_\lambda$} are nested by construction.
+In addition, we will assume \smash{$|\tE_\lambda|<\infty$} for every $\lambda\in\Lambda$.
 Our model path will correspond to the sequence of distinct ever-active sets.
@@
 \]
-and for $k\geq 1$ let $\lambda_k$ denote the (random) value of $\lambda$ where the active set changes for the $k$th time:
-\begin{equation}
+and for $k\geq 1$, let $\lambda_k$ denote the (random) value of $\lambda$ where the active set changes for the $k$th time:
+\begin{equation*}
   \Lambda_k = \{\lambda\in \Lambda:\; \tE_\lambda \supsetneq \tE_{\lambda_{k-1}}\},
@@
   E_k = \bigcap_{\lambda\in\Lambda_k} \tE_{\lambda}.
-\end{equation}
+\end{equation*}
 
@@
 The {\em selective next-entry test} is the test that rejects for large $\lambda_k$, based on the law
-\begin{equation}
+\begin{equation*}
 \L\left(\lambda_k \;\mid\; U_{k-1}, M_{k-1}\right).
-\end{equation}
+\end{equation*}
 
@@
 an intersection of $k$ balls which does not necessarily have a simple structure.
-When all the groups are of size 1, i.e. in the LASSO, the intersection 
+When all the groups are of size 1, i.e., in the LASSO, the intersection 
 of $k$ balls is the intersection of $k$ intervals and we get a simplification.
@@
 how to fit nuclear norm here because I'm not sure what
-we should call the  ever active set when the loss is not Frobenius (i.e. unless
+we should call the  ever active set when the loss is not Frobenius (i.e., unless
 we are doing PCA or SVT). That is, in matrix
@@
 
-\paragraph{Remark} In its usual formulation, the lasso coefficients for linear regression minimize a penalized RSS criterion:
-\[
-\hat\beta_{\text{lasso}}^\lambda = \frac{1}{2}\|Y-X\beta\|_2^2 + \lambda\|\beta\|_1,
+\paragraph{Remark.} In its usual formulation, the lasso coefficients for linear regression minimize a penalized RSS criterion:
+\[
+\hat\beta_{\text{lasso}}^\lambda = \argmin_{\beta\in\R^p} \; \frac{1}{2}\|Y-X\beta\|_2^2 + \lambda\|\beta\|_1, 
 \]
@@
 Otherwise, at the cost of some notational bloat, we could generalize \eqref{eq:pen-lik} by replacing $-\ell(\cdot)$ with $\rho(\ell(\cdot))$ for any strictly decreasing $\rho$, and the theory in this article would go through unchanged.
+\RJTcomment{I actually find this last remark more confusing than helpful, and I think we should take it out.
+I think it's pretty obvious that you can transform the loss by a monotone function ...
+and what is the $\exp$ doing up there?
+As it's written that's not a log-likelihood, that's a likelihood?}
 
@@
 
-Note that in our formulation of the problem, the type I error $V=(\hk-k_0)_+$ is defined in a ``model-centric'' fashion: at step $k$ in linear regression, we are testing  whether a particular linear model $M(E_{k-1})$ adequately describes the data $Y$.
-Even if the next selected variable $X_{j_k}$ is a noise variable in the full model, it is not necessarily a mistake to reject $M(E_{k-1})$ provided there are other signal variables that have not yet been included.
+Note that in our formulation of the problem, the type I error \smash{$V=(\hk-k_0)_+$} is defined in a ``model-centric'' fashion: at step $k$ in linear regression, we are testing  whether a particular linear model $M(E_{k-1})$ adequately describes the data $Y$.
+Even if the next selected variable $X_{j_k}$ is a noise variable (i.e., truly inactive) in the full model, it is not necessarily a mistake to reject $M(E_{k-1})$ provided there are signal variables (i.e., truly active ones) that have not yet been included.
 
@@
 \[
-\theta^E = X_E^+ \mu = \argmin_{\theta\in \R^{|E|}} \|\mu - X_E\theta\|_2^2,
+\theta^E = X_E^+ \mu = \argmin_{\theta\in \R^{|E|}} \; \|\mu - X_E\theta\|_2^2,
 \]
@@
 \[
-H_k:\;\mu = X_{E_{k-1}} \theta^{E_{k-1}}.
+H_k^{\text{comp}}:\;\mu = X_{E_{k-1}} \theta^{E_{k-1}}.
 \]
@@
 While the complete null is the strongest hypothesis of the three, the incremental null is neither weaker nor stronger than the full-model null. Defining 
-\begin{align}
-V^{\text{inc}} &= \#\{k < \hk:\; H_k^{\text{inc}} \text{ is true}\} \quad \text{ and } \\
+\begin{align*}
+V^{\text{inc}} &= \#\{k < \hk:\; H_k^{\text{inc}} \text{ is true}\}, \quad \text{ and } \\
 V^{\text{full}} &= \#\{k < \hk:\; H_k^{\text{full}} \text{ is true}\},
-\end{align}
+\end{align*}
 we can define an analogous FWER and FDR with respect to each of these alternative choices, and attempt to control these error rates. For example, we could define
@@
 \]
-
+The goal of course is to select a model along this sequence.
 In the first procedure, we construct a (type I) analysis-of-variance (ANOVA) table performing an $F$- or $t$-test of each model against the next.
@@
 As we discuss further in Section~\ref{sec:selective-reg}, the saturated-model tests of \citet{taylor2014exact} are valid selective tests of $H_k^{\text{inc}}$.
-Unfortunately, this requires us to assume $\sigma^2$ is known, can result in a large reduction in power, creates dependence between $p$-values at different steps, and is difficult to generalize beyond the case of linear regression.
-See \citet{gsell2013false} for an in-depth discussion of the relative conceptual advantages and disadvantages between the three approaches.
+The advantage of these tests is that they are highly computationally efficient (they do not require sampling).
+But, unfortunately, they also carry a number of drawbacks: they require us to assume $\sigma^2$ is known, can result in a large reduction in power, create dependence between $p$-values at different steps, and are difficult to generalize beyond the case of linear regression.
+%See \citet{gsell2013false} for an in-depth discussion of the relative conceptual advantages and disadvantages between the three approaches.
 
@@
 Methods like AIC \citet{akaike1974new} and BIC \citep{schwarz1978estimating} are designed for the non-adaptive case, where the sequence of models is determined in advance of observing the data.
-Cross-validation, another general-purpose algorithm for tuning parameter selection, targets out-of-sample error and tends to select many noise variables when the signal is sparse (e.g.
-\cite{LY2015}).
+Cross-validation, another general-purpose algorithm for tuning parameter selection, targets out-of-sample error and tends to select many noise variables when the signal is sparse (e.g., \cite{LY2015}).
 \citet{benjamini2009simple} extend the AIC by using an adaptive penalty $\sigma^2\lambda_{k,p} k$ to select a model, based on generalizing the Benjamini-Hochberg procedure \citet{benjamini1995controlling}, but do not prove finite-sample control.
@@
 
-An ordered hypothesis testing procedure takes in a sequence of $p$-values $p_1, \ldots, p_d$ for null hypotheses $H_{1}, \ldots, H_{d}$, and outputs a decision $\hk$ to reject the initial block $H_{1}, \ldots, H_{\hk}$ and accept the remaining hypotheses.
+An ordered hypothesis testing procedure takes in a sequence of $p$-values $p_1, \ldots, p_d$ for null hypotheses $H_{1}, \ldots, H_{d}$, and outputs a decision \smash{$\hk$} to reject the initial block \smash{$H_{1}, \ldots, H_{\hk}$} and accept the remaining hypotheses.
 Note that in our setup the hypotheses are nested, with $H_{k-1} \sub H_k$; as a result, all of the false hypotheses precede all of the true ones.
@@
 These procedures also assume the sequence of hypotheses is fixed, whereas in our setting the truth or falsehood of $H_k$ is random, depending on which model is selected at step $k$.
-However, we show that the error guarantees for these stopping rules do transfer to the random-hypothesis setting, provided we have the same independence property conditional on $k_0$.
+Fortuitously, we show that the error guarantees for these stopping rules do transfer to the random-hypothesis setting, provided we have the same independence property conditional on the completion index $k_0$ (recall
+$k_0(Y) = \min\{k:\; F \in M_k(Y)\}$).
 
@@
 We will call this procedure {\em BasicStop}. It is discussed in \citet{marcus1976}.
-
 Since $k_0$ is the index of the first null hypothesis, we have
@@
 \]
-and we choose the {\em largest} model with $\widehat{\text{FDP}}_k \leq \alpha$.
+and we choose the {\em largest} model with \smash{$\widehat{\text{FDP}}_k \leq \alpha$}.
 \citet{gsell2013sequential} show that ForwardStop controls the FDR if the null $p$-values are independent of each other and of the non-nulls.
@@
 \citet{li2015accumulation} generalize ForwardStop, introducing the family of {\em accumulation tests}, defined as follows.
-Fix a function $h: [0,1] \rightarrow [0,\infty]$ satisfying $\int_{t=0}^1 h(t)dt=1$, called an {\em accumulation function.} Using $h$ we can construct an estimate of FDP at step $k$:
-\begin{equation}
+Fix a function $h: [0,1] \rightarrow [0,\infty]$ satisfying \smash{$\int_{t=0}^1 h(t)dt=1$}, called an {\em accumulation function.} Using $h$ we can construct an estimate of FDP at step $k$:
+\begin{equation*}
 \widehat{\text{FDP}}_h(k)= \frac{\sum_{i=1}^k h(p_i)}{k},
-\end{equation}
-and select the largest model for which $\widehat{\text{FDP}}_k \leq \alpha$. Taking $h(p)=-\log(1-p)$ recovers the ForwardStop rule. Another choice they propose is the {\em HingeExp} function
-\begin{equation}
+\end{equation*}
+and select the largest model for which \smash{$\widehat{\text{FDP}}_k \leq \alpha$}. Taking $h(p)=-\log(1-p)$ recovers the ForwardStop rule. Another choice they propose is the {\em HingeExp} function
+\begin{equation*}
   h_{\rm HingeExp}(p)= 
@@
   \end{cases}
-\end{equation}
-
+\end{equation*}
 \citet{li2015accumulation} show that accumulation tests control a modified FDR criterion provided that the null $p$-values are $U[0,1]$, and are independent of each other and of the non-nulls.
@@
 
+\RJTcomment{I've not changed anything with the notation below that conditions on $k_0=k$.
+I'll leave it up to you Will, but as we discussed, I think it would be most useful to either write $P(p_k \leq \alpha | M_{k-1})$ on the set $\{ F \in M_{k-1}\}$, or to plainly explain what you mean by conditioning on an event involving $k_0$.} \\
 In our problem setting, the sequence of selected models is random; thus, the truth or falsehood of each $H_k$ is not fixed, as assumed in the previous section.
 It is clear that we will recover all of the same guarantees if we construct our $p$-values conditional on the entire path $M_{0:d}(Y)$; however, this option is rather unappealing for both computational and statistical reasons.
-In constructing the $p$-value for step $k$, is it sufficient to condition only on the subpath of models $M_{0:k-1}(Y)$, or better yet to condition only on the current null model $M_{k-1}$?
-
-Note that conditioning on the current null model {\em is} enough to guarantee that $p_k$ is uniform conditional on the event $\{F\in M_{k-1}(Y)\}$ that the $k$th null is true.
+In constructing the $p$-value for step $k$, is it sufficient to condition on the subpath of models $M_{0:(k-1)}(Y)$, or better yet, only on the current null model $M_{k-1}$?
+
+We note that conditioning on the current null model {\em is} enough to guarantee that $p_k$ is uniform conditional on the event $\{F\in M_{k-1}(Y)\}$ that the $k$th null is true.
 Unfortunately, however, it is {\em not} enough to guarantee that the above stopping rules actually control FDR or FWER.
-Let $z_{\alpha/2}$ denote the upper-$\alpha/2$ quantile of the $\cN(0,1)$ distribution, and consider the following counterexample:
+Let $z_{\alpha/2}$ denote the upper-$\alpha/2$ quantile of the $\cN(0,1)$ distribution, and consider the following counterexample.
 
@@
 Otherwise, we add $X_2$, then $X_1$, then $X_3$.
-At each step we construct selective max-$z$ $p$-values and finally choose $\hat k$ using BasicStop.
+At each step we construct selective max-$z$ $p$-values and finally choose \smash{$\hk$} using BasicStop.
 
@@
 The crucial issue is that $k_0$ is no longer a fixed index.
-Consequently, even though $\L(p_k \mid H_k \text{ true})=U[0,1]$ for each fixed $k$, the $p$-value $p_{k_0}$ corresponding to the {\em first} true null hypothesis is stochastically smaller than uniform.
+Consequently, even though $\L(p_k \mid H_k \text{ true})=U[0,1]$ for each fixed $k$, the $p$-value \smash{$p_{k_0}$} corresponding to the {\em first} true null hypothesis is stochastically smaller than uniform.
 
@@
 \]
-then BasicStop again would control FWER, by the argument in \ref{eq:basic-stop-control}.
-
-Note that $k_0$ is unknown, so we cannot directly condition on its value when we construct our single-step $p$-values. However, because
+then BasicStop again would control FWER, by the same argument as given in \eqref{eq:basic-stop-control}.
+Note that $k_0$ is unknown, so we cannot directly condition on its value when we construct our single-step $p$-values.
+However, because
 \[
@@
 it is enough to condition on $(M_{k-2}, M_{k-1})$ when we construct our $p$-values.
-As we will see in Section~\ref{sec:pValsIndep}, conditioning on $(M_{k-1}, U_{k-1})$ is equivalent to conditioning on $(M_{0:k-1}, U_{k-1})$ in most cases of interest including forward stepwise likelihood paths and ever-active regularization paths (but not the path in Proposition~\ref{prop:counterexample}).
+As we will see in Section~\ref{sec:pValsIndep}, conditioning on $(M_{k-1}, U_{k-1})$ is equivalent to conditioning on $(M_{0:(k-1)}, U_{k-1})$ in most cases of interest including forward stepwise likelihood paths and ever-active regularization paths (but not the path in Proposition~\ref{prop:counterexample}).
 
@@
   \P(p_{k+1} \leq \alpha_{k+1}, \ldots, p_d \leq \alpha_d
-  \mid p_1, \ldots, p_k, \; k_0 = k) = \prod_{i=k+1}^d \alpha_i
+  \mid p_1, \ldots, p_k, k_0 = k ) \eqAS \left(\prod_{i=k+1}^d \alpha_i\right) .
 \end{equation}
@@
   \]
-
-  Suppose that $\hk$ controls $\Err_g$ at level $\alpha$ if the $H_{1:d}$ are fixed and the null $p$-values are uniform and independent of each other and the non-nulls.
-Then, $\hk$ controls $\Err_g$ at level $\alpha$ whenever $p_{1:d}$ satisfy~\eqref{eq:indepCond_random_k0}.
+  Suppose that \smash{$\hk$} controls $\Err_g$ at level $\alpha$ if $H_{1:d}$ are fixed and the null $p$-values are uniform and independent of each other and the non-nulls.
+Then, \smash{$\hk$} controls $\Err_g$ at level $\alpha$ whenever $p_{1:d}$ satisfy~\eqref{eq:indepCond_random_k0}.
 \end{proposition}
 \begin{proof}
-    For nested hypotheses, $k_0$ completely determines the truth or falsehood of $H_k(Y)$ for every $k$. If $\hk$ controls $\Err_g$ in the 
+    For nested hypotheses, $k_0$ completely determines the truth or falsehood of $H_k(Y)$ for every $k$. If \smash{$\hk$} controls $\Err_g$ in the 
 fixed-hypothesis case, it must in particular control $\Err_g$ conditional on any fixed values of $p_{1:k_0}$. Thus, \eqref{eq:indepCond_random_k0} implies
@@
 
-Recall that at step $k$, the max-$t$ test rejects for large $T_k=|t_{E_{k-1}}^*|$, comparing its distribution to the conditional law
+Recall that at step $k$, the max-$t$ test rejects for large \smash{$T_k=|t_{E_{k-1}}^*|$}, comparing its distribution to the conditional law
 \begin{equation}\label{eq:maxTnull}
@@
 \end{equation}
-This conditional distribution is the same for all $F\in M(E_{k-1})$, because $U_{k-1}=(X_{E_{k-1}}'Y,\; \|Y\|^2)$ is the complete sufficient statistic for $M(E_{k-1})$.
-
-Now, if $F\in M(E_{k-1})$, then $p_k$ is uniform and independent of the previous $p$-values $p_{1:k-1}$, for two reasons:
+This conditional distribution is the same for all $F\in M(E_{k-1})$, because \smash{$U_{k-1}=(X_{E_{k-1}}'Y,\; \|Y\|^2)$} is the complete sufficient statistic for $M(E_{k-1})$.
+
+Now, if $F\in M(E_{k-1})$, then $p_k$ is uniform and independent of the previous $p$-values $p_{1:(k-1)}$, for two reasons:
 \begin{enumerate}
@@
 \end{enumerate}
-
-Informally, the pair $(E_{k-1}, U_{k-1})$ forms a ``wall of separation'' between $p_k$ and $p_{1:k-1}$, guaranteeing that $\L(p_k \mid p_{1:k-1}) = U[0,1]$ whenever $H_k$ is true.
+Informally, the pair $(E_{k-1}, U_{k-1})$ forms a ``wall of separation'' between $p_k$ and $p_{1:(k-1)}$, guaranteeing that $\L(p_k \mid p_{1:(k-1)}) = U[0,1]$ whenever $H_k$ is true.
 
 Next we will explain why $p_1,\ldots,p_{k-1}$ are functions of $E_{k-1}$ and $U_{k-1}$.
-Knowing $E_{k-1}=E$ tells us that the $k-1$ variables in $E$ are selected first, and knowing $(X_{E_{k-1}}'Y,\|Y\|^2)$ is enough information to compute the $t$-statistics $t_{j,D}$ for all $j\in E$ and $D\sub E$.
+Knowing $E_{k-1}=E$ tells us that the $k-1$ variables in $E$ are selected first, and knowing \smash{$(X_{E_{k-1}}'Y,\|Y\|^2)$} is enough information to compute the $t$-statistics $t_{j,D}$ for all $j\in E$ and $D\sub E$.
 As a result, we can reconstruct the order in which those $k-1$ variables were added, since for $i<k$,
 \[
-j_i = \argmax_{j \notin E_{i-1}} |t_{j,E_{i-1}}| 
-= \argmax_{j \in E\setminus E_{i-1}} |t_{j,E_{i-1}}| \quad \text{ on the event } \{E_{k-1}=E\}.
+j_i = \argmax_{j \notin E_{i-1}} \, |t_{j,E_{i-1}}| 
+= \argmax_{j \in E\setminus E_{i-1}} \, |t_{j,E_{i-1}}| \quad \text{ on the event } \{E_{k-1}=E\}.
 \]
 In general, we say that the path $M_{0:d}$ follows the {\em subpath sufficiency principle} (SSP) if for each $k$ the subpath $M_{0:k}$ is a function of $(M_k,U_k)$.
+\RJTcomment{A thought / suggestion: we define the SSP here, but I think it's not really that helpful, because we're only talking about max-t.
+We refer to SSP in the next section; in defining the sufficient filtration.
+Why not just define SSP there?
+In general, I think there's quite a lot of lingo being thrown around, and though I like all the new terms, I think we could do a little better job consolidating where they're defined.}
 
@@
 
-A {\em selective $p$-value} $p_k$ for testing $H_k:\; F \in M_{k-1}(Y)$ satisfies for any $F\in M_{\infty}$,
-\[
-\P_F(p_k \leq \alpha \mid M_{k-1}) \leqAS \alpha \quad \text{ on } \{F\in M_{k-1}\}
-\]
-
+A {\em selective $p$-value} $p_k$ for testing $H_k:\; F \in M_{k-1}(Y)$ satisfies, for any $F\in M_{\infty}$,
+\[
+\P_F(p_k \leq \alpha \mid M_{k-1}) \leqAS \alpha \quad \text{ on } \{F\in M_{k-1}\}.
+\]
 We say that a filtration $\sF_{0:d}$ {\em separates} the $p$-values $p_{1:d}$ if
@@
 \end{enumerate}
+If we think of $\sF_k$ as representing information available at step $k$, then the first condition means that $p_k$ ``excludes'' whatever evidence may have accrued against the null by step $k-1$, and the second means that any information revealed after step $k$ plays no role in determining $p_k$.
 Separated $p$-values are independent on nulls, as we see next.
@@
   \begin{equation}\label{eq:jointConserv}
-  \P_F\left(p_{k+1}\leq \alpha_{k+1}, \ldots, p_d \leq \alpha_d\;
-    \mid\; k_0 \leq k, \; \sF_k\right) \leqAS \prod_{i=k+1}^d
-  \alpha_i.
+  \P_F\left(p_{k+1}\leq \alpha_{k+1}, \ldots, p_d \leq \alpha_d, \; k_0 \leq k \;
+    \mid\; \sF_k\right) \leqAS \left(\prod_{i=k+1}^d
+  \alpha_i\right) 1_{\{k_0 \leq k\}}.
   \end{equation}
@@
   \[
-  \P_F\left(B_d \mid \sF_{d-1}\right)1_{\{k_0 \leq d-1\}} \leqAS \alpha_d,
+  \P_F\left(B_d \mid \sF_{d-1}\right)1_{\{k_0 \leq d-1\}} \leqAS \alpha_d
+1_{\{k_0 \leq d-1\}},
   \]
@@
     &\leqAS \P_F\left[ B_{k+1}
-      \mid \sF_k\right]1_{\{k_0 \leq k+1\}}\prod_{i=k+2}^d \alpha_i
-    &\leqAS \prod_{i=k+1}^d \alpha_i
+      \mid \sF_k\right]1_{\{k_0 \leq k\}}\prod_{i=k+2}^d \alpha_i \\
+    &\leqAS \left(\prod_{i=k+1}^d \alpha_i \right)1_{\{k_0 \leq k\}} .
   \end{align*}
-
-  If the $p_k$ are exact then the above 
-  inequalities become equalities, implying uniformity and joint independence.
+Lastly, if the $p_k$ are exact then the inequalities above become equalities, implying uniformity and joint independence.
 \end{proof}
-
-If we think of $\sF_k$ as representing information available at step $k$, then the first condition means that $p_k$ ``excludes'' whatever evidence may have accrued against the null by step $k-1$, and the second means that any information revealed after step $k$ plays no role in determining $p_k$.
 
@@
 \end{equation}
-
 The sufficient filtration separates $p_{1:d}$ if and only if 
 \begin{enumerate}
-\item $p_k$ is super-uniform given $M_{0:k-1}$ and $U_{k-1}$, and 
+\item $p_k$ is super-uniform given $M_{0:(k-1)}$ and $U_{k-1}$, and 
 \item $p_k$ is a function of $M_{0:k}$ and $U_k$.
 \end{enumerate}
-
 To be a valid selective $p$-value for a single step, $p_k$ only needs to be super-uniform conditional on $M_{k-1}$.
-In general, it is more stringent to additionally require that $p_k$ must condition on the entire subpath $M_{0:k-1}$ as well as $U_{k-1}$.
-In most cases of interest, however, (1) we will want to condition on $(M_{k-1}, U_{k-1})$ anyway, to eliminate nuisance parameters of $M_{k-1}$, and (2) the path satisifies the SSP, so conditioning on $(M_{k-1}, U_{k-1})$ is equivalent to conditioning on $(M_{k-1}, U_{k-1})$.
+In general, it is more stringent to additionally require that $p_k$ must condition on the entire subpath $M_{0:(k-1)}$ as well as $U_{k-1}$.
+In most cases of interest, however, (i) we will want to condition on $(M_{0:(k-1)}, U_{k-1})$ anyway, to eliminate nuisance parameters of $M_{k-1}$, and (ii) the path satisifies the SSP, so conditioning on $(M_{0:(k-1)}, U_{k-1})$ is equivalent to conditioning on $(M_{k-1}, U_{k-1})$.
 Thus, most single-step selective $p$-values we would contemplate using are already exact conditional on $\sF_{k-1}$.
-
-The second requirement, that $p_k$ be $\sF_k$-measurable, has more bite. For example, we will see in Section~\ref{sec:selective-reg} that it excludes saturated-model tests in linear regression.
+The second requirement, that $p_k$ be $\sF_k$-measurable, has more bite.
+For example, we will see in Section~\ref{sec:selective-reg} that it excludes saturated-model tests in linear regression.
 
@@
 Therefore, we must have $h(U_{k-1})\eqAS 0$, so $p_k$ is independent of $\sF(M_k, U_{k-1})$.
-Because $M_{0:d}$ satisfies the SSP, $p_k$ is also independent of $\sF_{k-1} = \sF(M_{0:k-1}, U_{k-1})$.
+Because $M_{0:d}$ satisfies the SSP, $p_k$ is also independent of $\sF_{k-1} = \sF(M_{0:(k-1)}, U_{k-1})$.
 
@@
 
-Note that if our path algorithm did not satisfy the SSP, we could in many cases still repair the situation by constructing $p$-values that are uniform conditional on $(M_{0:k-1},U_{k-1})$, not just conditional on $(M_{k-1},U_{k-1})$.
+Note that if our path algorithm did not satisfy the SSP, we could in many cases still repair the situation by constructing $p$-values that are uniform conditional on $(M_{0:(k-1)},U_{k-1})$, not just conditional on $(M_{k-1},U_{k-1})$.
 
@@
 
-In this section we apply Theorem~\ref{thm:suffCond} to establish that in the generic sparse parametric model of Section~\ref{sec:genSparse}, the forward stepwise path and all ever-active regularization paths satisfy the SSP.
-What is more, the greedy likelihood ratio statistic $G_k$ is $\sF_k$-measurable with respect to the forward stepwise path's sufficient filtration, and the next-entry statistic $\lambda_k$ is $\sF_k$-measurable with respect to the regularized-likelihood path's sufficient filtration.
+In this section, we apply Theorem~\ref{thm:suffCond} to establish that in the generic sparse parametric model of Section~\ref{sec:genSparse}, the forward stepwise path and all ever-active regularization paths satisfy the SSP.
+Moreover, the greedy likelihood ratio statistic $G_k$ is $\sF_k$-measurable with respect to the sufficient filtration of the forward stepwise path, and the next-entry statistic $\lambda_k$ is $\sF_k$-measurable with respect to the sufficient filtration of the regularized-likelihood path.
 As a result, the selective greedy likelihood $p$-values in the forward stepwise path are independent on nulls, as are the next-entry $p$-values in the regularized-likelihood path, per Theorem~\ref{thm:suffCond}.
@@
 Define the restricted log-likelihood
-\begin{equation}
+\begin{equation*}
 \ell_E(\theta; Y) = \left\{\begin{matrix} 
@@
     -\infty  & \mathrm{ otherwise.}\end{matrix}\right. 
-\end{equation}
-
+\end{equation*}
+
+We begin by stating our result for forward stepwise.
 \begin{proposition}\label{prop:forwardSSP}
-  The forward stepwise path satisfies the subpath sufficiency principle.
+Forward stepwise paths satisfy the subpath sufficiency principle.
 \end{proposition}
@@
   \begin{equation}\label{eq:forwardStepDef}
-    j_k = \argmax_j \;\;\sup \left\{\ell(\theta; Y):\; \theta\in\Theta(E_{k-1} \cup \{j\})\right\}, \quad \text{ and } \quad
-    E_k = E_{k-1} \cup \{j_k\}
-  \end{equation}
-
-  For some fixed step $k$ and active set $E$, let $A$ denote the event $\{M_k = M(E)\}$.
+    j_k = \argmax_j \;\sup \left\{\ell(\theta; Y):\; \theta\in\Theta(E_{k-1} \cup \{j\})\right\}, \quad \text{ and } \quad
+    E_k = E_{k-1} \cup \{j_k\}.
+  \end{equation}
+  For some fixed step $k$ and active set $E$, denote the event $A=\{M_k = M(E)\}$.
 Conditioning on $U$, the restricted likelihood is proportional to a function depending only on $U = U(Y; \;M(E))$.
@@
   On $A$, we have $E_s(Y) \sub E$ for all $s\leq k$, meaning that the maximum in~\eqref{eq:forwardStepDef} is attained by some $j\in E$ at every step. So, for $s \leq k$, we have
-\begin{align}
-  j_s &= \argmax_{j\in E} \;\;\sup \left\{\ell_E(\theta; Y):\;
+\begin{align*}
+  j_s &= \argmax_{j\in E} \;\sup \left\{\ell_E(\theta; Y):\;
     \theta\in\Theta(E_{s-1} \cup \{j\})\right\} \\
-  &= \argmax_{j\in E} \;\;\sup \left\{\ell_E^U(\theta; U(Y)):\;
-    \theta\in\Theta(E_{s-1} \cup \{j\})\right\}. \label{eq:onlyU}
-\end{align}
-
-Equation~\eqref{eq:onlyU} shows that $j_1,\ldots, j_{k-1}$ all depend on $Y$ only through $U(Y)$, which equals $U_k(Y)$ on $A$.
+  &= \argmax_{j\in E} \;\sup \left\{\ell_E^U(\theta; U(Y)):\;
+    \theta\in\Theta(E_{s-1} \cup \{j\})\right\}. 
+\end{align*}
+The above shows that $j_1,\ldots, j_{k-1}$ all depend on $Y$ only through $U(Y)$, which equals $U_k(Y)$ on $A$.
 As a result, it also follows that the entire sequence $M_{0:k}(Y)$ depends only on $(M_k, U_k)$.
@@
 \begin{corollary}\label{cor:greedyLikIndep}
-  If $U(Y; M)$ is complete sufficient for model $M$, the selective greedy likelihood $p$-values computed for the forward stepwise path are independent on nulls.
+  If $U(Y; M)$ is complete and sufficient for model $M$, then the selective greedy likelihood $p$-values computed for the forward stepwise path are independent on nulls.
 \end{corollary}
 \begin{proof}
-  Per Theorem~\ref{thm:suffCond}, it is enough to show that $p_k$ is $\sF_k$-measurable with respect to the forward stepwise path's sufficient filtration.
+  Per Theorem~\ref{thm:suffCond}, it is enough to show that $p_k$ is $\sF_k$-measurable with respect to the sufficient filtration for the forward stepwise path.
 First, the test statistic $G_k$ is $\sF_k$-measurable because
@@
 
-Analogous results hold for ever-active regularized likelihood paths and the next-entry $p$-values.
+The analogous result hold for ever-active regularized likelihood paths.
 
@@
 \begin{proof}
-  Again, for fixed $k$ and $E$ let $A$ denote the event $\{M_k = M(E)\}$, and define $U=U(Y;\, M(E))$. If we define
-\[
-\hat\theta^{(E,\lambda)} = \argmin_{\theta\in\Theta(E)} -\ell_E^{U}(\theta; U(Y)) + P_\lambda(\theta),
+  Again, for fixed $k$ and $E$, let $A=\{M_k = M(E)\}$, and define $U=U(Y;\, M(E))$. If we denote 
+\[
+\hat\theta^{(E,\lambda)} = \argmin_{\theta\in\Theta(E)} \; -\ell_E^{U}(\theta; U(Y)) + P_\lambda(\theta),
 \]
@@
 \]
-But $\hat\theta^{(E,\lambda)}$ depends only on $U(Y)$.
+But \smash{$\hat\theta^{(E,\lambda)}$} depends only on $U(Y)$.
 Therefore, on $A$, we can reconstruct the entire path of solutions for $\{\lambda\in \Lambda:\; \lambda\geq \lambda_k\}$, once we know $(M_k,\,U_k)$.
@@
 \begin{corollary}\label{cor:nextLambdaIndep}
-  If $U(Y; M)$ is complete sufficient for model $M$, the selective next-entry $p$-values computed for any ever-active regularized likelihood path are independent on nulls.
+  If $U(Y; M)$ is complete and sufficient for model $M$, then the selective next-entry $p$-values computed for any ever-active regularized likelihood path are independent on nulls.
 \end{corollary}
 \begin{proof}
-  Per Theorem~\ref{thm:suffCond}, it is enough to show that $p_k$ is $\sF_k$-measurable with respect to the regularized-likelihood path's sufficient filtration.
-
-  Recall that 
+  Per Theorem~\ref{thm:suffCond}, it is enough to show that $p_k$ is $\sF_k$-measurable with respect to the  sufficient filtration for the regularized-likelihood path.  Recall that 
   \[
@@
   \]
-  and that the $\tE_{\lambda}$ are nested by construction and finite by assumption. As a result, there is some $\nu_k\in \Lambda_k$ for which $\tE_{\nu_k} = E_k$.
-
-  Furthermore, for all $\lambda\in\Lambda, \lambda\geq \nu_k$, we have $\tE_\lambda \sub E_k$.
-As a result, ${\hat\theta^{\lambda} =\hat\theta^{(E_k,\lambda)}}$ for such $\lambda$, so $\tE_{\lambda}$ depends only on $\ell_{E_k}^{Y\mid U}(Y;\, U_k)$, and therefore $\lambda_k$ can be computed from $(E_k, U_k)$.
-
+  and that the \smash{$\tE_{\lambda}$} are nested by construction and finite by assumption. As a result, there is some $\nu_k\in \Lambda_k$ for which \smash{$\tE_{\nu_k} = E_k$}.
+
+  Furthermore, for all $\lambda\in\Lambda, \lambda\geq \nu_k$, we have \smash{$\tE_\lambda \sub E_k$}.
+As a result, \smash{${\hat\theta^{\lambda} =\hat\theta^{(E_k,\lambda)}}$} for such $\lambda$, so $\tE_{\lambda}$ depends only on \smash{$\ell_{E_k}^{Y\mid U}(Y;\, U_k)$}, and therefore $\lambda_k$ can be computed from $(E_k, U_k)$.
 Second, the null distribution $\L\left(\lambda_k \,\mid\, M_{k-1}, U_{k-1}\right)$ against which we compare $\lambda_k$ is also $\sF_k$-measurable because it depends only on $(M_{k-1}, U_{k-1})$, both of which are $\sF_{k-1}$-measurable.
@@
 For simplicity, we will assume throughout this section that our path algorithm adds exactly one variable at each step.
-We also assume the path algorithm satisfies the SSP, so that we need not worry about the distinction between conditioning on $(M_{k-1}, U_{k-1})$ and conditioning on $(M_{0:k-1}, U_{k-1})$.
+We also assume the path algorithm satisfies the SSP, so that we need not worry about the distinction between conditioning on $(M_{k-1}, U_{k-1})$ and conditioning on $(M_{0:(k-1)}, U_{k-1})$.
 
@@
 See for example \citet{lockhart2014significance}, \citet{taylor2013tests}, \citet{taylor2014exact}, \citet{lee2013exact}, and \citet{loftus2014significance}.
-Define the saturated model as
+\RJTcomment{So, I would claim that the covariance test is NOT a saturated model test, but instead a select model test.
+Of course it doesn't really fit into this framework with exact p-values, but if you look at what we proved, it was a selected model null result.
+Also, Taylor (2013) tests the global null throughout, so again, I don't think it makes much sense to say that it's a saturated model test ...} Define the saturated model as
 \[
@@
 If this is not the case, we can estimate $\sigma^2$ and apply the test while plugging in the estimated value; this is what we have done with the diabetes data to compute the $p$-values in Table~\ref{tab:diab}.
+\citet{tibshirani2015uniform} discuss this and other strategies for unknown $\sigma^2$.
 
@@
 \[
-\theta^E \;\;=\;\; \argmin_{\theta\in\R^{|E|}} \|\mu - X_E\theta\| \;\;=\;\; (X_E'X_E)^{-1}X_E'\mu
-\]
-
+\theta^E \;\;=\;\; \argmin_{\theta\in\R^{|E|}} \; \|\mu - X_E\theta\|_2 \;\;=\;\; (X_E'X_E)^{-1}X_E'\mu.
+\]
 In the context of sequential model selection, we can use the saturated-model framework to test the incremental null 
@@
 \]
-Let $\eta_k\in \R^n$ denote the vector for which $\theta_{j_k}^{E_k} = \eta_k'\mu$, and let $\proj_{\eta_k}^\perp$ denote the projection operator into the space orthogonal to $\eta_k$.
+Let $\eta_k\in \R^n$ denote the vector for which \smash{$\theta_{j_k}^{E_k} = \eta_k'\mu$}, and let \smash{$\proj_{\eta_k}^\perp$} denote the projection operator into the space orthogonal to $\eta_k$.
 The saturated model has $n$-dimensional natural parameter $\mu$ and complete sufficient statistic $Y$, and the UMPU saturated-model test for $\eta_k'\mu$ is based on the distribution
+\begin{equation*}
+\L\left( \eta_k'Y \mid E_{k-1}, \;j_k, \;\proj_{\eta_k}^\perp Y \right),
+\end{equation*}
+or equivalently on
 \begin{equation}\label{eq:satModel}
-\L\left( \eta_k'Y \mid E_{k-1}, \;j_k, \;\proj_{\eta_k}^\perp Y \right),
-\end{equation}
-or equivalently on
-\begin{equation}\label{eq:satModelEquiv}
 \L\left( X_{j_k}'Y \mid E_{k-1}, \;j_k, \;\proj_{\eta_k}^\perp Y \right).
 \end{equation}
-The pair $(E_{k-1}, j_k)$ defines the selection event and $\proj_{\eta_k}^\perp Y$ eliminates the nuisance parameter $\proj_{\eta_k}^\perp \mu$.
-
-While the test statistic $X_{j_k}'Y$ is measurable with respect to $\sF_k$, the null distribution we compare it against is not: it depends on $\proj_{\eta_k}^\perp Y$, which is not measurable with respect to $\sF_k$.
+The pair $(E_{k-1}, j_k)$ defines the selection event and \smash{$\proj_{\eta_k}^\perp Y$} eliminates the nuisance parameter \smash{$\proj_{\eta_k}^\perp \mu$}.
+While the test statistic $X_{j_k}'Y$ is measurable with respect to $\sF_k$, the null distribution we compare it against is not: it depends on \smash{$\proj_{\eta_k}^\perp Y$}, which is not measurable with respect to $\sF_k$.
 As a result, saturated-model $p$-values constructed as in~\eqref{eq:satModel} are in general {\em not} independent on nulls.
@@
 \subsection{Selected-Model Tests}
-Selected-model inference~\citep{fithian2014optimal} 
+Selected-model inference, introduced in \citet{fithian2014optimal},
 represents a more powerful option when testing for inclusion of variable $j_k$.
@@
 If $\sigma^2$ were unknown we would also need to condition on $\|Y\|^2$.
-
 We can construct $p$-values from the UMPU, equal-tailed, or one-sided test based on~\eqref{eq:selModel_cond}; the only difference is in how much of the overall level $\alpha$ is apportioned to the right or left tail.
-
 If we use the construction in~\eqref{eq:selModel_cond} after a subpath-sufficient path algorithm, the resulting $p$-values are always guaranteed to have independence on nulls.
-This is because the test statistic $X_{j_k}'Y$ is measurable with respect to $\sF_k$, and so is the null distribution we compare it against, which depends on $E_{k-1}, j_k,$ and $U_{k-1} = X_{E_{k-1}}'Y$.
+This is because the test statistic $X_{j_k}'Y$ is measurable with respect to $\sF_k$, and so is the null distribution we compare it against, which depends on $E_{k-1}, j_k,$ and \smash{$U_{k-1} = X_{E_{k-1}}'Y$}.
 
 Note the contrast between~\eqref{eq:selModel_cond} and~\eqref{eq:satModel}.
-In~\eqref{eq:selModel_cond}, we only need to condition on an $|E_{k-1}|$-dimensional projection of $Y$, whereas in~\eqref{eq:satModel} we condition on $n-1$-dimensional projection of $Y$.
-Conditioning on more information tends to sap the power of a test, and as we will see in Sections~\ref{sec:bivariate} and~\ref{sec:sparseReg}, the saturated-model tests can pay a heavy price for this extra conditioning.
+In~\eqref{eq:selModel_cond}, we only need to condition on an $|E_{k-1}|$-dimensional projection of $Y$, whereas in~\eqref{eq:satModel} we condition on an $(n-1)$-dimensional projection of $Y$.
+Conditioning on more information can sap the power of a test; as we will see in Sections~\ref{sec:bivariate} and~\ref{sec:sparseReg}, the saturated-model tests tend to pay a heavy price for this extra conditioning.
 
@@
 \end{equation}
+Again, if $\sigma^2$ were unknown we would also need to condition on $\|Y\|^2$.
 We could also replace $X_{j_k}'Y$ with any other test statistic $T(Y)$.
@@
 
-In the case of known $\sigma^2$, the greedy likelihood ratio statistic is the max-$z$ statistic: the maximum of $z_{j,E_{k-1}}$, the largest multivariate $z$-statistic of any held-out variable, controlling for the variables in the current model.
+%In the case of known $\sigma^2$, the greedy likelihood ratio statistic is the max-$z$ statistic: the maximum of $z_{j,E_{k-1}}$, the largest multivariate $z$-statistic of any held-out variable, controlling for the variables in the current model.
 
@@
 In the next section we compare the selected- and saturated-model settings in more detail.
+
+\RJTcomment{I don't mean to sound overtly critical but I don't find this table helpful; in fact I find it confusing.
+Why is the covariance test called a saturated model test?
+Sure, it tests the complete null, but why does it use a saturated model framework?
+I doesn't use the law at all as described in \eqref{eq:satModel}.
+Also, the name ``spacings'' is reserved for a very specific approximation to the saturated-model test in LAR, that only uses knots in the LAR path, so I wouldn't use this name.
+I think the last 3 subsections have explained the differences nicely, and I think this table doesn't really add much beyond them, and confuses me.
+
+Also, I don't know why we give the example below.
+What does it show?
+More examples follow in the next section, adn more detailed experiments follow later, and this just shows that the covariance test is conservative, but by now, we know this and have showed this many, many times over.
+And it's not at all the point of this paper.}
 
@@
 \end{table}
-Figure  \ref{fig:larexample} shows an example of the LAR/lasso tests with $N=50, p=10$ and two strong signals.
+Figure  \ref{fig:larexample} shows an example of the LAR/lasso tests with $n=50, p=10$ and two strong signals.
 We plot the $p$-values only for those realizations in which the correct model was selected in the first two steps (about 95\% of the time).
-We see that the next-entry test is the most powerful, especially at step 1. The next-entry and spacing test have the correct type I error in steps 3 and 4,
-while the covariance test becomes conservative in step 4, as we move further into the null part of the sequence.
+We see that the next-entry test is the most powerful, especially at step 1.
+The next-entry and spacing test have the correct type I error in steps 3 and 4, while the covariance test becomes conservative in step 4, as we move further into the null part of the sequence.
 \begin{figure}
@@
 We now briefly compare the computational, conceptual, and statistical advantages and disadvantages of selected- and saturated-model tests, focusing on the problem of sequential model selection.
-For a detailed general discussion, see \citet[][Section 5]{fithian2014optimal}.
+For a detailed general discussion, see Section 5 of \citet{fithian2014optimal}.
 
 Selected-model inference is typically more computationally involved than saturated-model inference.
-Because the saturated-model test conditions on $n-1$ dimensions, the resulting distribution in~\eqref{eq:satModel} is nothing more than a truncated univariate Gaussian random variable.
+Because the saturated-model test conditions on $n-1$ dimensions, the resulting distribution in~\eqref{eq:satModel} is nothing more than a truncated univariate Gaussian random variable, and inference can be carried out analytically.
 By contrast, selected-model tests typically require sampling from a truncated multivariate Gaussian distribution of dimension $p-|E_{k-1}|$.
@@
 
-\JTcomment{I would say our goal is a good stopping rule AND inference on parameters of selected (or perhaps saturated) model. We do not consider it here
-because we don't know how to do it, not because it is not interesting.}
+\JTcomment{I would say our goal is a good stopping rule AND inference on parameters of selected (or perhaps saturated) model.
+We do not consider it here because we don't know how to do it, not because it is not interesting.}
+
+\RJTcomment{I don't see why what you mention is a disadvantage of the saturated-model perspective.
+In fact, I would argue that it is an advantage.
+The model is always going to be misspecified.
+What is being covered is a well-defined quantity, regardless of the true model.
+That is precisely why I like the saturated-model perspective for confidence intervals...}
 
@@
 First, selected-model inference allows us to drop the assumption that $\sigma^2$ is known.
-This is not an option for the saturated model because conditioning on both $\|Y\|^2$ and $\proj_{\eta_k}^\perp Y$ results in a degenerate conditional distribution for $\eta_k'Y$.
-
+This is not an option for the saturated model because conditioning on both $\|Y\|^2$ and \smash{$\proj_{\eta_k}^\perp Y$} results in a degenerate conditional distribution for $\eta_k'Y$.
 Second, we have seen in Section~\ref{sec:pValsIndep} that tests of the form~\eqref{eq:selModel_cond} or~\eqref{eq:selModel_marg} yield independent $p$-values under general conditions, allowing us to apply  the sequential stopping rules of~\citet{gsell2013sequential} and \citet{li2015accumulation}.
-
 Finally, and perhaps most importantly, selected-model tests can be dramatically more powerful than saturated-model tests, as we illustrate now with an extended example.
@@
 \] 
-
 The forward stepwise path and the lasso path both select $j_1=1$ if and only if $|Y_1|>|Y_2|$.
@@
 \]
-
 The selected-model test at step 1 compares $Y_1$ to its distribution under $M_0$ conditional on $A_1$, a test of $H_0:\;\mu_1=0$ in model $M_1$.\footnote{In this very simple example, the max-$z$, max-$z$-identify, and next-entry tests all coincide.} By contrast, the saturated-model test is a test of $H_0:\; \mu_1=0$ in the model $M_{\text{sat}}:\; Y \sim \cN(\mu, I_2)$.
@@
 
-Figure~\ref{fig:bv} illustrates a general phenomenon in saturated-model tests: when there are near-ties between strong variables that are competing to enter the model, the resulting $p$-value my be very weak~\citep{lockhart2014significance}.
+Figure~\ref{fig:bv} illustrates a general phenomenon in saturated-model tests: when there are near-ties between strong variables that are competing to enter the model, the resulting $p$-value my be very weak.
 Figure~\ref{fig:bv_rocCurve} displays the cumulative distribution function for the first $p$-value when $\mu=(4,4)$, a very strong signal.
@@
 
-Results in \citet{fithian2014optimal} show that the selected-model test is strictly more powerful when the selected model is correct; i.e., when $\mu_2=0$.
+Results in \citet{fithian2014optimal} show that the selected-model test is strictly more powerful when the selected model is correct, i.e., when $\mu_2=0$.
 Figure~\ref{fig:bv_powCurves} shows the power curve for each test when $\mu_2=0$ (left panel) and $\mu_2=4$ (right panel).
@@
 The difference is much more pronounced when $\mu_2=4$.
-
 Note that if $\mu=(0,4)$, the incremental null is true, but the complete null is false.
@@
 Continuing with Example~\ref{ex:bivariate}, Table~\ref{tab:bv_twoWayTable} shows a two-way contingency table for the saturated-model $p$-values $(p_1(Y), p_2(Y))$, binned into cells of height and width 0.2, simulated under the global null $\mu=0$.
-Because $k_0=0$, both $p$-values are uniform by construction, but the $p$-values are strongly dependent, with correlation $-0.48$.
+Because $k_0=0$, both $p$-values are uniform by construction, but the saturated-model $p$-values are strongly dependent, with correlation $-0.48$.
 By contrast, the selected-model $p$-values $(p_1,p_2)$ are independent under the global null, a consequence of Theorem~\ref{thm:suffCond}.
@@
 
-For lasso linear regression, the situation seems even more dire: for any fixed value of $\lambda$, the event $\{\supp(\hat\beta^\lambda) = E\}$ is a union of up to $2^{|E|}$ polytopes in $\R^n$, each satisfying $O(p)$ constraints \citep{lee2013exact}.
+For lasso linear regression, the situation seems even more dire: for any fixed value of $\lambda$, the event \smash{$\{\supp(\hat\beta^\lambda) = E\}$} is a union of up to $2^{|E|}$ polytopes in $\R^n$, each satisfying $O(p)$ constraints \citep{lee2013exact}.
 \footnote{To finesse this issue, \citet{lee2013exact} propose conditioning on the signs of $\hat\beta^\lambda$ as well.
@@
 But the event $\{E_{k-1}(Y)=E\}$ depends on $\supp(\hat\beta^\lambda)$ for every $\lambda > \lambda_k$ --- i.e., on the solution to {\em infinitely many} lasso problems.
+
+\RJTcomment{I don't get this last paragraph really.
+First of all, just to point out, for each of the forward stepwise, LAR, and lasso paths, we condition on the signs in our 2014 paper.
+So this issue is not specific to the lasso.
+Also, why are we talking about lasso for fixed $\lambda$ here?
+This is confusing.
+We have exact inference for steps along the lasso path (not the same problem as fixed $\lambda$ lasso) in our 2014 paper.
+Our constructions give $\approx 2pk$ constraints for forward stepwise after $k$ steps, and $\approx 3pk$ constraints for LAR r lasso after $k$ steps.
+Now, what is given below is still a big improvement over that!
+But it's worth being very clear about the predecessors to it.}
 
@@
   \]
-  for $v_j^-$ and $v_j^+$ given explicitly in
-  Equations~(\ref{eq:vMinus_FS}--\ref{eq:vPlus_FS}).
-  Thus, $A$ corresponds exactly to 
-  a set with $2(p-k)$ linear inequality constraints and $k$
-  linear equality constraints on $Y$.
+  for $v_j^-$ and $v_j^+$ given explicitly in (\ref{eq:vMinus_FS}--\ref{eq:vPlus_FS}).
+Thus, $A$ corresponds exactly to a set with $2(p-k)$ linear inequality constraints and $k$ linear equality constraints on $Y$.
 \end{theorem}
 
+\RJTcomment{Can't we get rid of the strict inequalities above?
+The proof writes things in terms of nonstrict inequalities, and I think it makes sense to drop them; in the case of ties, $E$ is still a valid selected model.}
+
 \begin{proof}
-  At stage $i < k$, the next variable to enter is the one with maximal correlation with the residual vector; i.e.
+  At stage $i < k$, the next variable to enter is the one with maximal correlation with the residual vector, i.e.,
   \[
@@
   \frac{\left| X_j ' (Y - X_{E_i}\hat\beta^i) \right|}
-  {\|\proj_{E_i}^\perp X_j\|}
-  \]
-  Because forward stepwise regression satisfies the SSP, we know the entire path of fitted models up to step $k$ once we condition on the $k$th active set $E_k$ and its sufficient statistics $X_{E_k}'Y$.
-On the set $\{E_k=E\}$, all of the quantities $j_{i+1}$, $E_i$, $\hat\beta^i$, and $\proj_{E_i}^\perp$ depend only on $X_E'Y$.
+  {\|\proj_{E_i}^\perp X_j\|_2}
+  \]
+  Because forward stepwise regression satisfies the SSP, we know the entire path of fitted models up to step $k$ once we condition on the $k$th active set $E_k$ and its sufficient statistics \smash{$X_{E_k}'Y$}.
+On the set $\{E_k=E\}$, all of the quantities $j_{i+1}$, $E_i$, $\hat\beta^i$, and \smash{$\proj_{E_i}^\perp$} depend only on $X_E'Y$.
 For brevity, write
@@
   \frac{\left| X_j ' (Y - X_{E_i}\hat\beta^i) \right|}
-  {\|\proj_{E_i}^\perp X_j\|}.
-  \]
-  On $A$, $C_i^*$ is attained at $j=j_{i+1}$, the $i$th variable added.
-
-  If $X_E'Y$ is fixed at $u$, and $j \notin E$, then the condition for 
-  $X_j$ {\em not} to enter at step $i+1 < k$ is
+  {\|\proj_{E_i}^\perp X_j\|_2}.
+  \]
+  On $A$, $C_i^*$ is attained at $j=j_{i+1}$, the $(i+1)$st variable added.
+
+  If $X_E'Y$ is fixed at $u$, and $j \notin E$, then the condition for $X_j$ {\em not} to enter at step $i+1 < k$ is
   \[
   \frac{\left| X_j' (Y - X_{E_i}\hat\beta^i) \right|}
-  {\|\proj_{E_i}^\perp X_j\|} 
+  {\|\proj_{E_i}^\perp X_j\|_2} 
   \leq C_i^*,
@@
     X_j' X_{E_i}\hat\beta^i -
-    C_i^*\|\proj_{E_{i}}^\perp X_{j}\|
+    C_i^*\|\proj_{E_{i}}^\perp X_{j}\|_2
     \;\;\;\leq\;\;\;
@@
     X_j' X_{E_i}\hat\beta^i +
-    C_i^*\|\proj_{E_{i}}^\perp X_{j}\|, 
+    C_i^*\|\proj_{E_{i}}^\perp X_{j}\|_2, 
   \end{equation}
   so the set $A$ is equivalent to 
-  Equation~\eqref{eq:noEnterBounds_FS} holding
+  \eqref{eq:noEnterBounds_FS} holding
   for every $i < k$ and $j \notin E$. Nominally, this gives $2k(p-k)$
@@
     v_j^-(E,u) &= \max_{0 \leq i < k} \;\;X_j' X_{E_i}\hat\beta^i -
-    C_i^*\|\proj_{E_{i}}^\perp X_{j}\| \\
+    C_i^*\|\proj_{E_{i}}^\perp X_{j}\|_2, \\
     \label{eq:vPlus_FS}
     v_j^+(E,u) &= \min_{0 \leq i < k} \;\;X_j' X_{E_i}\hat\beta^i +
-    C_i^*\|\proj_{E_{i}}^\perp X_{j}\|.
+    C_i^*\|\proj_{E_{i}}^\perp X_{j}\|_2.
   \end{align}  
@@
 \begin{equation}
-j_k = \argmax_j \;\;\sup \left\{\ell(\theta; Y):\; \theta\in\Theta(E_{k-1} \cup \{j\})\right\}, \quad \text{ and } \quad
+j_k = \argmax_j \;\sup \left\{\ell(\theta; Y):\; \theta\in\Theta(E_{k-1} \cup \{j\})\right\}, \quad \text{ and } \quad
 E_k = E_{k-1} \cup \{j_k\}.
 \end{equation}
-
   Then, for a candidate active set $E$ of size $k$, the set $A = \{E_k = E, \;U_E = u\}$ is characterized exactly by the constraints $U_E = u$ and
@@
   \]
-  for $v_j^-$ and $v_j^+$ given in Equations~(\ref{eq:vMinus_FSL}--\ref{eq:vPlus_FSL}).
+  for $v_j^-$ and $v_j^+$ given in (\ref{eq:vMinus_FSL}--\ref{eq:vPlus_FSL}).
 Thus, $A$ corresponds exactly to a set with $2(p-k)$ linear inequality constraints and $k$ linear equality constraints on $U(Y)$.
@@
   \[
-  \ell_E^*(U_E(Y)) 
+  \ell_E^*(U(Y)) 
   = \sup\left\{\theta'U(Y) - \psi(\theta) :\; \theta\in \Theta(E) \right\}
@@
   \]
-  where $\psi_E:\; \R^{|E|} \to \R$ is defined such that $\psi(\theta)=\psi_E(\theta_E)$ if $\theta_{\setminus E}=0$.
+  where $\psi_E:\; \R^{|E|} \to \R$ is defined such that $\psi(\theta)=\psi_E(\theta_E)$ if $\theta_{\{1,\ldots p\}\setminus E}=0$.
 Note that $\ell_E^*$ is the convex dual of the convex function $\psi_E$, so $\ell_E^*$ is also convex.
@@
 No?
-I also think $U$ in the definition of $\ell^*_E$ can be just $U(Y)$.} At stage $i<k$, the next variable to enter is the one that maximizes $\ell_{E_i\cup \{j\}}^*(U(Y))$.
-
-  Because the forward stepwise likelihood path satisfies the SSP, we know the entire path of fitted models up to step $k$ once we condition on $E_k$ and its sufficient statistics $U_{E_k}(Y)$.
-On the set $\{E_k=E\}$, all of the quantities $j_{i+1}$, $E_i$, $\hat\theta^i$, and $\proj_{E_i}^\perp$ depend only on $U_E(Y)$.
+I also think $U$ in the definition of $\ell^*_E$ can be just $U(Y)$.} \RJTcomment{I would call this the ``convex conjugate'' or just the ``conjugate'' of $\ell$; and yes this will always be convex.
+It doesn't require knowing anything about $\psi_E$ or even $\Theta$.} At stage $i<k$, the next variable to enter is the one that maximizes \smash{$\ell_{E_i\cup \{j\}}^*(U(Y))$}.
+
+  Because the forward stepwise likelihood path satisfies the SSP, we know the entire path of fitted models up to step $k$ once we condition on $E_k$ and its sufficient statistics \smash{$U_{E_k}(Y)$}.
+On the set $\{E_k=E\}$, all of the quantities $j_{i+1}$, $E_i$, \smash{$\hat\theta^i$}, and \smash{$\proj_{E_i}^\perp$} depend only on $U_E(Y)$.
 For brevity, write
   \[
-  C_i^* = \max_{j \notin E_i} 
+  C_i^* = \max_{j \notin E_i} \;
     \sup \left\{ \theta ' U - \psi(\theta):\; \theta \in \Theta(E_i\cup \{j\}) \right\}.
   \]
-  On $A$, $C_i^*$ is attained at $j=j_{i+1}$, the $i$th variable added.
-
-  If $U_E(Y)$ is fixed at $u$, and $j \notin E$, then the condition for 
+  On $A$, $C_i^*$ is attained at $j=j_{i+1}$, the $(i+1)$st variable added.  If $U_E(Y)$ is fixed at $u$, and $j \notin E$, then the condition for 
   variable $j$ {\em not} to enter at step $i+1 < k$ is
@@
   \]
-  Note that $\ell_{E_i\cup \{j\}}^*$ depends only on the sufficient statistics $U_{E_i}(Y)$ --- which is fixed, on $A$ --- and the remaining sufficient statistic $U_j(Y)$.
-Since $\ell_{E_i\cup\{j\}}^*$ is convex, the set
+  Note that \smash{$\ell_{E_i\cup \{j\}}^*$} depends only on the sufficient statistic \smash{$U_{E_i}(Y)$} --- which is fixed, on $A$ --- and the remaining sufficient statistic $U_j(Y)$.
+Since \smash{$\ell_{E_i\cup\{j\}}^*$} is convex in $U$, if we fix \smash{$U_{E_i}$}, and consider it as a function of $U_j$ alone, then it is still convex in this single argument.
+Hence any sublevel set of \smash{$\ell_{E_i\cup\{j\}}^*$}, with \smash{$U_{E_i}$} fixed, is just a 1-dimensional convex set, i.e., an interval, so we may write
   \begin{equation}\label{eq:noEnterBounds_FSL}
-  \{U_E=u, \ell_{E_i\cup \{j\}}^*(U(Y)) \leq C_i^*\} 
-  = \{ U_E=u, v_{j,i}^-(E,u) \leq U_j \leq v_{j,i}^+(E,u)\},
-  \end{equation}
-  where $v_j^{-}$ and $v_j^+$ are the two values where $\ell_{E_i\cup \{j\}}^*(U)=C_i^*$.
+  \{U_E=u, \; \ell_{E_i\cup \{j\}}^*(U(Y)) \leq C_i^*\} 
+  = \{ U_E=u, \; v_{j,i}^-(E,u) \leq U_j \leq v_{j,i}^+(E,u)\},
+  \end{equation}
+  where $v_{j,i}^{-}$ and $v_{j,i}^+$ are the two values of $U_j$ such that \smash{$\ell_{E_i\cup \{j\}}^*(U)=C_i^*$}.
 
   The set $A$ is equivalent to 
-  Equation~\eqref{eq:noEnterBounds_FSL} holding
-  for every $i < k$ and $j \notin E$. Nominally, this gives $2k(p-k)$
+  \eqref{eq:noEnterBounds_FSL} holding
+  for every $i < k$ and $j \notin E$. As before, nominally, this gives $2k(p-k)$ 
   linear inequality constraints to satisfy, but most of them are
@@
   \begin{align}\label{eq:vMinus_FSL}
-    v_j^-(E,u) &= \max_{0 \leq i < k} \;\;v_{j,i}^-(E,u)\\
+    v_j^-(E,u) &= \max_{0 \leq i < k} \;\;v_{j,i}^-(E,u),\\
     \label{eq:vPlus_FSL}
-    v_j^+(E,u) &= \min_{0 \leq i < k} \;\;v_{j,i}^-(E,u)
+    v_j^+(E,u) &= \min_{0 \leq i < k} \;\;v_{j,i}^-(E,u).
   \end{align}
@@
 
-A similar result also holds for the class of $\ell_1$-regularized exponential family models including the graphical LASSO.
+A similar result also holds for the class of $\ell_1$-regularized exponential family models. % including the graphical LASSO.
 
@@
   \begin{equation}\label{eq:regProblem}
-  \hat\theta^\lambda = \argmin_{\theta\in \Theta} 
+  \hat\theta^\lambda = \argmin_{\theta\in \Theta} \;
   -\ell(\theta) + \lambda\|\theta\|_1,
@@
   for $\lambda\in \Lambda \sub [0,\infty)$.
-
   Then, for a candidate active set $E$ of size $k$, 
@@
   for $v_j^-$ and $v_j^+$ given explicitly in
-  Equations~(\ref{eq:vMinus_L1}--\ref{eq:vPlus_L1}).
+  (\ref{eq:vMinus_L1}--\ref{eq:vPlus_L1}).
   Thus, $A$ corresponds exactly to 
@@
   \left| \pardd{\ell(\hat\theta^\lambda)}{\theta_j} \right|
-  = \left|U_j - \E_{\hat\theta^\lambda}[U_j]\right|
+  = \left|U_j - \E_{\hat\theta^\lambda}[U_j]\right|.
   \]
   On $A$, for $\lambda \geq\lambda_k$, 
-  $\E_{\hat\theta^\lambda}[U_j]$ is a known function of $E$ and $u$,
+  \smash{$\E_{\hat\theta^\lambda}[U_j]$} is a known function of $E$ and $u$,
   so we can set
@@
     v_j^-(E,u) &= \sup_{\lambda \geq \lambda_k} \;\; 
-    \E_{\hat\theta^\lambda}[U_j] - \lambda \\
+    \E_{\hat\theta^\lambda}[U_j] - \lambda, \\
     \label{eq:vPlus_L1}
     v_j^+(E,u) &= \inf_{\lambda \geq \lambda_k} \;\;
-    \E_{\hat\theta^\lambda}[U_j] + \lambda .
+    \E_{\hat\theta^\lambda}[U_j] + \lambda.
   \end{align}
@@
 
-Next we compare several model selection procedures in simulation.
-We simulate from a linear regression model with $n=100$ observations and $p=40$ variables.
+Here we compare several model selection procedures in simulation.
+We generate data from a linear regression model with $n=100$ observations and $p=40$ variables.
 The design matrix $X\in\R^{n\times p}$ is a random Gaussian design with pairwise correlations of 0.3 between predictor variables.
 
-The columns of $X$ are normalized to have length 1, and we simulate from $Y \sim \cN(X\beta,I_n)$, using a seven-sparse model with signal-to-noise ratio 5:
+The columns of $X$ are normalized to have length 1, and we simulate $Y \sim \cN(X\beta,I_n)$, using a seven-sparse model with signal-to-noise ratio 5:
 \[
 \beta_j = \left\{\begin{matrix}5 & j = 1,\ldots,7\\ 0 &
-    j>7\end{matrix}\right.
+    j = 7,\ldots 100 \end{matrix}\right. .
 \]
@@
 For each step we compute one-step selected-model and saturated-model $p$-values, as well as nominal (unadjusted) $p$-values, conditioning on the signs of the active variables to make the problem more computationally tractable.
-Figure~\ref{fig:simulation_null_false} shows the power of all three tests for each of the first ten steps, conditional on the event that the null hypothesis is false.
+Figure~\ref{fig:simulation_null_false} shows the power of all three tests for each of the first ten steps \RJTcomment{I only see steps 6 through 9?}, conditional on the event that the null hypothesis is false.
 It is clear from Figure~\ref{fig:simulation_null_false} that the selected-model $p$-values are far more powerful than the saturated-model $p$-values.
@@
 The nominal test appears to be powerful,
-  but is not $U(0,1)$ under the null, as shown below in Figure \ref{fig:simulation_null_true}.}
+  but is not $U[0,1]$ under the null, as shown below in Figure \ref{fig:simulation_null_true}.}
   \label{fig:simulation_null_false}
@@
 \end{figure}
+
+ \RJTcomment{The lines in Fig 6 are super thin, thinner than Fig 5?}
 
@@
 
-Figure~\ref{fig:simulation_null_true} shows the distribution of $p_k$ for $k = 8, \ldots, 17$, given that the null hypothesis tested at step $k$ is correct (i.e., that $k_0< k$).
+Figure~\ref{fig:simulation_null_true} shows the distribution of $p_k$ for $k = 8, \ldots, 17$ \RJTcomment{I only see steps 9 through 11?}, given that the null hypothesis tested at step $k$ is correct (i.e., that $k_0< k$).
 Because the correct model is seven-sparse, $k=8$ is the first index for which the null can possibly be true.
-All but the nominal $p$-values are uniform by construction, but the nominal $p$-values are highly anti-conservative, as expected.
-
-Finally, as a warning against misinterpretation of our method, we include Figure~\ref{fig:simulation_noise_var} showing the first ten $p$-values for each method, conditional on event that the variable added at step $k$ is a noise variable in the full model.
+All but the nominal $p$-values are uniform by construction, and the nominal $p$-values are highly anti-conservative, as expected.
+
+Finally, as a warning against misinterpretation of our method, we include Figure~\ref{fig:simulation_noise_var} showing the first ten $p$-values for each method \RJTcomment{I only see steps 1 through 4?}, conditional on event that the variable added at the given step is a noise variable in the full model.
 Now, none of the $p$-values are uniform.
-
 This is not an error in our implementation of the method, but rather a consequence of our ``model-centric'' point of view.
@@
 \item its model-wise FDR, and
-\item its variable-wise FDR, where we use $V^{\text{full}}= \#\{\text{ noise variables included in } M_{\hk}\}$ instead of $V=(\hk-k_0)_+$.
+\item its variable-wise FDR, where we use $V^{\text{full}}= \#\{\text{ noise variables included in } M_{\hk}\}$ instead of \smash{$V=(\hk-k_0)_+$}.
 \end{itemize} 
@@
 
-We see that basicStop and forwardStop control model-wise FWER and FDR, respectively, using $p$-values from max-$t$ or max-$t$-identify, as predicted by the theory, 2) the nominal $p$-values do not lead to control of FWER or FDR, as expected due to their anti-conservatism.
-Although not guaranteed by the theory, the $p$-values from the saturated model do control FWER and model-wise FDR in this example, 4) the max-$t$ rules show higher probabilities of selecting the true model, especially using forwardStop, 5) the knockoff method does not yield variable-wise FDR control.
+\RJTcomment{I think these captions / labels in the tables should be fixed up, but I didn't want to mess with anything.
+Few pedantic things: ``simple'' should be ``basic'', because now we're calling it BasicStop, and why not just write BasicStop and ForwardStop instead of ``basic'' and ``forward''?
+Also, ``MaxZ'' should be ``Max-$z$'' and same for ``Max-$t$''.
+Lastly, what is with the typewriter type?}
+
+We see that BasicStop and ForwardStop control model-wise FWER and FDR, respectively, using $p$-values from max-$t$ or max-$t$-identify, as predicted by the theory, and the nominal $p$-values do not lead to control of FWER or FDR, as expected due to their anti-conservatism.
+Although not guaranteed by the theory, the $p$-values from the saturated model do control FWER and model-wise FDR in this example.
+The max-$t$ rules show higher probabilities of selecting the true model, especially using ForwardStop.
+Lastly, the knockoff method does not yield variable-wise FDR control.
+\RJTcomment{Ummm isn't this an unexpected failure of knockoffs?
+Should we comment on this?
+Or am I missing something?
+Oh, I see.
+We've also included knockoff+, which does control the FDR (because it adds something to the denominator, right?).
+We should say something about this.}
 
@@
 When considering splitting two daughter nodes, the generic CART algorithm splits the nodes in an arbitrary order, only stopping when a node reaches a minimum size.
-We instead consider a  special version of CART  algorithm:  the "best first"  method, which orders splits by their achieved reduction in loss.
+We instead consider a  special version of CART: the ``best first'' method, which orders splits by their achieved reduction in loss.
 This procedure is used, for example in the R package for gradient boosting, called {\tt gbm}.
@@
 Hence we can obtain independent, sequential $p$-values for each successive split in the decision tree.
+\RJTcomment{Just for what it's worth, I always think of CART as performing best-first splits.
+This is how Jerry teaches it, at least.}
 
 Given predictor matrix $X \in \R^{n\times p}$ and binary response vector $Y\in \{0,1\}^n$, we begin with the intercept-only model $\logit \P(Y_i \mid X_i = x) \equiv \beta_0$.
-
 At the first step we choose a splitting variable $j_1$ and split point $t_1$ which is a midpoint between two observed values of variable $j$, leading to a model of the form
-\begin{equation}
-\logit \P(Y_i \mid X_i=x) = \beta_0 + \beta_{1} 1\{x_{j_1} > t_1\}.
-\end{equation}
+\begin{equation*}
+\logit \,\P(Y_i \mid X_i=x) = \beta_0 + \beta_{1} 1\{x_{j_1} > t_1\}.
+\end{equation*}
 While there are various ways to choose among potential variables and split points, we assume for the sake of simplicity that $(j_1,t_1)$ are chosen to maximize binomial likelihood (or equivalently, to minimize binomial deviance).
-
 At the second stage, we then split one of the two daughter nodes, choosing the split that gives the greatest increase in likelihood.
@@
 \begin{equation}
-\logit \P(Y_i \mid X_i=x) = \beta_0 + \beta_{1} 1\{x_{j_1} > t_1\} + \beta_2 1\{x_{j_1} \leq t_1, x_{j_2} > t_2\}
+\logit \,\P(Y_i \mid X_i=x) = \beta_0 + \beta_{1} 1\{x_{j_1} > t_1\} + \beta_2 1\{x_{j_1} \leq t_1, x_{j_2} > t_2\}.
 \label{eqn:intmodel}
 \end{equation}
-
 After $k$ steps we obtain a logistic regression model with $k$ features, each of which is an interaction of step functions.
@@
 The best-first search for split points can be expressed as a modified form of greedy forward stepwise logistic regression in which the set of candidate features changes at each step, depending on what features have already been included.
-Nevertheless, the proof of Proposition~\ref{prop:forwardSSP} goes through essentially without modification to show that this algorithm satisfies the SSP, and the greedy likelihood ratio test gives $p$-values that are independent on nulls\footnote{Since the test statistic is discrete, we need to randomize if we want truly exact $p$-values}.
+Nevertheless, the proof of Proposition~\ref{prop:forwardSSP} goes through essentially without modification to show that this algorithm satisfies the SSP, and the greedy likelihood ratio test gives $p$-values that are independent on nulls.
+Since the test statistic is discrete, we need to randomize if we want truly exact $p$-values.
 
@@
 For example, given a historical document, we may wish to investigate whether the entire document was written by the same author or different sections were written by different authors, by quantitatively comparing the writing style across different sections.
-For example, $Y_t$ could could represent the usage distribution in section $t$ over ``context-free'' words such as articles, conjunctions, and prepositions \citep[][see e.g.
-]{chen2015graph}.
+For example, $Y_t$ could could represent the usage distribution in section $t$ over ``context-free'' words such as articles, conjunctions, and prepositions \citep[see e.g.,][]{chen2015graph}.
 
@@
 \end{equation}
-
 For $t_2>t_1$ let $V(Y;\, t_1, t_2)\in \cY^{t_2-t_1}$ denote the order statistics of $(Y_{t_1+1},\ldots, Y_{t_2})$. Then the complete sufficient statistic for $M(E)$ is
@@
 We can build a model in a greedy fashion beginning with $E_0 = \emptyset$ and then at step $k=1,\ldots,d$ setting
-\begin{equation}
-  t_k = \argmax_t \;\; W(Y; \, E_{k-1}, t), 
+\begin{equation*}
+  t_k = \argmax_t \; W(Y; \, E_{k-1}, t), 
   \quad \text{ and } \quad E_k = E_{k-1} \cup \{t_k\}.
-\end{equation}
+\end{equation*}
 This path algorithm satisfies the SSP because $W(Y; \, E_{i}, t_j)$ is measurable with respect to $U_k=U(Y; \, E_k)$ for every $i<j\leq k$.
@@
 \[
-W_k^*(Y) = \max_t W(Y; \, E_{k-1}, t),
+W_k^*(Y) = \max_t \; W(Y; \, E_{k-1}, t),
 \]
@@
 As a result, the $p$-values will be independent on nulls per Theorem~\ref{thm:suffCond}.
-
 Note that sampling from the law in~\eqref{eq:condPerm} can be carried out by permuting subsequences $(Y_{t(j,E)+1}, \ldots, Y_{t(j+1,E)})$ and accepting the permutation only if it gives the same $E_{k-1}$ as we get in the original data.
@@
 
-It is a commonplace that ``essentially all models are wrong, but some are useful'' \citep{box1987empirical}.
+A common proverb in statistics states that ``essentially all models are wrong, but some are useful'' \citep{box1987empirical}.
 In essence, a statistical model is useful if it is large enough to capture the most important features of the data, but still small enough that inference procedures can achieve adequate power and precision.
@@
 In future work we also plan to address the matter of performing selective inference for parameters of the model actually selected.
-In principle, we already know how to do this using the framework of \citet{fithian2014optimal} --- we simply condition on the event that $M_{\hk}$ is selected and perform inference with respect to $M_{\hk}$ --- but the task may be computationally awkward in general.
+In principle, we already know how to do this using the framework of \citet{fithian2014optimal} --- we simply condition on the event that $M_{\hk}$ is selected and perform inference with respect to \smash{$M_{\hk}$} --- but the task may be computationally awkward in general.
 
@@
 
-\paragraph{Hit-and-Run Sampler} \citet{fithian2014optimal} propose sampling from laws like that in~\eqref{eq:truncNorm} via a hit-and-run sampler for the $n-|E|$-dimensional multivariate Gaussian distribution: at each step, to generate $Y^{t+1}$ from $Y^t$, we choose a uniformly random unit vector $\nu$ in the hyperplane, and resample $\nu'Y$ conditional on the constraints $\proj_{\nu}^\perp Y^{t+1}=\proj_{\nu}^\perp Y^t$ and the constraint $Y^{t+1}\in A$.
+\paragraph{Hit-and-Run Sampler} \citet{fithian2014optimal} propose sampling from laws like that in~\eqref{eq:truncNorm} via a hit-and-run sampler for the $(n-|E|)$-dimensional multivariate Gaussian distribution: at each step, to generate $Y^{t+1}$ from $Y^t$, we choose a uniformly random unit vector $\nu$ in the hyperplane, and resample $\nu'Y$ conditional on the constraints $\proj_{\nu}^\perp Y^{t+1}=\proj_{\nu}^\perp Y^t$ and the constraint $Y^{t+1}\in A$.
 If $A$ is a polytope this amounts to resampling a univariate Gaussian distribution with known mean and variance, truncated to an interval defined by maximizing and minimizing $\nu'Y^{t+1}$ subject to linear constraints.

### 2015-12-03 paper.tex
more edits

@@
 \title{Selective Sequential Model Selection}
-\author{William Fithian, Jonathan Taylor, Rob Tibshirani, and Ryan J. Tibshirani} 
+\author{William Fithian, Jonathan Taylor, Robert Tibshirani, and Ryan J. Tibshirani} 
 \maketitle
@@
 \citet{taylor2014exact} study methods for generating exact $p$-values at each step of these path algorithms, and their methods provide a starting point for our proposals.
-Other related works include \citet{loftus2014significance}, who describe $p$-values for path algorithms that enter groups of variables (instead of individual variables) at each step, and \citet{choi2014selecting}, who describe $p$-values for steps of principle components analysis (each step marking the estimation of a principle component direction).
+Other related works include \citet{loftus2014significance}, who describe $p$-values for path algorithms that add groups of variables (instead of individual variables) at each step, and \citet{choi2014selecting}, who describe $p$-values for steps of principle components analysis (each step marking the estimation of a principle component direction).
 
@@
 For a design matrix $X\in \R^{n\times p}$, with columns $X_1,\ldots X_p \in \R^n$, and a response vector $Y\in \R^n$, let $E \sub \{1, \ldots, p\}$ denote the current {\em active set}, the set of predictor variables already selected, and let $\RSS(E)$ denote the residual sum of squares for the corresponding regression model.
-For $j\notin E$ let $t_{j,E}(Y)$ denote the multivariate $t$-statistic of variable $j$, adjusting for the active variables as well as an intercept term.
+For $j\notin E$ let $t_{j,E}(Y)$ denote the multivariate $t$-statistic of variable $j$, adjusting for the active variables.
 Using the standard result that
 \begin{equation*}
-t_{j,E}^2 = (n-|E|-2) \frac{\RSS(E) - \RSS(E \cup \{j\})}{\RSS(E \cup \{j\})},
+t_{j,E}^2 = (n-|E|-1) \frac{\RSS(E) - \RSS(E \cup \{j\})}{\RSS(E \cup \{j\})},
 \end{equation*}
@@
 The response of interest is a quantitative measure of disease progression one year after baseline, and there are ten measured predictors --- age, sex, body-mass index, average blood pressure, and six blood serum measurements --- plus quadratic terms, giving a total of $p=64$ features.
-We apply forward stepwise regression to generate the model path (beginning with an intercept term), and then use each of the three methods to generate $p$-values at each step along the path.
+We apply forward stepwise regression to generate the model path (beginning with an intercept term, which we represent as a predictor $X_0=(1,\ldots,1)'$), and then use each of the three methods to generate $p$-values at each step along the path.
 Finally, we use the ForwardStop rule \citep{gsell2013sequential} at FDR level $\alpha=0.1$ to select a model based on the sequence of $p$-values.
@@
 
-The nominal $p$-values in column 1 of Table~\ref{tab:diab} are computed by comparing $t_E^*$ to the $t$-distribution with $n-|E|-2$ degrees of freedom, which would be the correct distribution if the sequence of models were selected before observing the data.
+The nominal (unadjusted) $p$-values in column 1 of Table~\ref{tab:diab} are computed by comparing $t_E^*$ to the $t$-distribution with $n-|E|-2$ degrees of freedom, which would be the correct distribution if the sequence of models were selected before observing the data.
 Because the model sequence is in fact selected adaptively to maximize the absolute value of $t_E^*$, this method is highly anti-conservative.
@@
 As $U_E=\left(X_E'Y,\, \|Y\|^2\right)$ is the complete sufficient statistic for $M(E)$, we can sample from the conditional null distribution of $t_E^*$ given $U_E$ and the current active set $E$, using the {\em selected-model test} framework of \citet{fithian2014optimal}.
-In step one, because $E=\emptyset$ is fixed and $t_E^*$ is independent of $U_\emptyset=\|Y\|^2$, the conditional test reduces to the max-$t$ test proposed in \citet{buja2014}.
+In step one, because $E=\{0\}$ is fixed and $t_E^*$ is independent of $U_{\{0\}}=(\sum_i Y_i, \|Y\|^2)$, the conditional test reduces to the max-$t$ test proposed in \citet{buja2014}.
 In later steps, $t_E^*$ and $U_E$ are not conditionally independent given $E$, but we can numerically approximate the null conditional distribution through Monte Carlo sampling.
@@
 In comparison to their saturated-model counterparts, we will see that the selected-model tests proposed here have three main advantages: power, independence (in many cases), and generalizability beyond linear regression.
-These advantages do not come entirely for free, as we will see: the selected-model methods test a more restrictive null hypothesis than the saturated-model methods.
-In addition, the selected-model tests require accept-reject or MCMC sampling, while the saturated-model tests can be carried out in closed form.
-
-Sections~\ref{sec:genericSetting} and \ref{sec:ordered} introduce the general problem setting, give important definitions, and review relevant literature on testing ordered hypotheses using a sequence of $p$-values.
+These advantages do not come entirely for free, as we will see: when $\sigma^2$ is known, the selected-model methods test a more restrictive null hypothesis than the saturated-model methods.
+In addition, the selected-model tests require accept-reject or Markov Chain Monte Carlo (MCMC) sampling, while the saturated-model tests can be carried out in closed form.
+
+Sections~\ref{sec:genericSetting} and \ref{sec:ordered} introduce the general problem setting and review relevant literature on testing ordered hypotheses using a sequence of $p$-values.
 Many of these methods require the $p$-values to be uniform and independent under the null, and we derive general conditions to ensure this property in Section~\ref{sec:pValsIndep}.
@@
 \[
-M_0 \sub M_1(Y) \sub \cdots \sub M_d(Y) \sub M_\infty.
+M_0(Y) \sub M_1(Y) \sub \cdots \sub M_d(Y) \sub M_\infty.
 \]
@@
 For simplicity I've removed the dependence of $M_0$ on $Y$, but if I'm wrong about this, sorry, just change it back.} \\
+\WFcomment{In many cases $M_0$ is deterministic but (for example) the way we define the ever-active path, we have $E_0(Y) = \bigcup_{\lambda\in\Lambda} \tE_\lambda(Y)$, which could be random if $\Lambda$ is a fixed grid.
+$M_\infty$ doesn't really need to be selected since we can just take it as a union over all possible models.}\\
 By {\em model}, we mean a family of candidate probability distributions for $Y$. For example in  linear regression, a model specifies  the set of
@@
 
-\JTcomment{Maybe we should write $k_F$ instead of $k_0$? Or $k_0(Y,F)$?}
-
-Define the {\em completion index} by $k_0(Y) = \min\{k:\; F \in M_k(Y)\}$, the index of the first correct model.
+\JTcomment{Maybe we should write $k_F$ instead of $k_0$? Or $k_0(Y,F)$?}\\
+\WFcomment{$\hat k_F$ is already the ForwardStop estimator. But $k_0(Y,F)$ is good.}\\
+
+Define the {\em completion index} by $k_0(Y, F) = \min\{k:\; F \in M_k(Y)\}$, the index of the first correct model.
 By construction, $F\in M_k \iff k \geq k_0$.
@@
 \begin{equation}\label{eq:errRate}
-\Err_F(\hk(\cdot), g) = \E_F\left[ g\left(\hk(Y), k_0(Y)\right)\right].
+\Err_F(\hk(\cdot), g) = \E_F\left[ g\left(\hk(Y), k_0(Y, F)\right)\right].
 \end{equation}
@@
 See \citet{fithian2014optimal} for a general treatment.
+Section~\ref{sec:computation} discusses computational strategies for the tests we propose in this article.
 
@@
 \]
-For any subset $E\sub \{1,\ldots J\}$ define the sparse submodel with active set $E$ as follows:
+For any subset $E\sub \{1,\ldots, J\}$ define the sparse submodel with active set $E$ as follows:
 \[
@@
 \[
-E_0 \sub E_1(Y) \sub \cdots \sub E_d(Y) \sub \{1,\ldots J\},
-\]
-\RJTcomment{Again here I've removed the dependence of $E_0$ on $Y$.} \\
+E_0(Y) \sub E_1(Y) \sub \cdots \sub E_d(Y) \sub \{1,\ldots J\},
+\]
 inducing a model path with $M_k = M(E_k),$ for $k=0,\ldots,d$.
 We will be especially interested in two generic path algorithms for the sparse parametric setting: forward stepwise paths and ever-active regularization paths.
-For a nonparametric example see Section~\ref{sec:further-examples}.
+As we will see in Section~\ref{sec:indep-greedy-entry}, both methods generically result in independent null $p$-values.
+For a nonparametric example see Section~\ref{sec:nonpar}.
 
@@
 \end{equation}
-the generalized likelihood ratio statistic for testing $M(E_{k-1})$ against the ``greedy'' alternative with one more active parameter:
-\begin{equation}\label{eq:greedyAlternative}
-M(E_{k-1})^{+1} = \bigcup_{j \notin E_{k-1}} M(E_{k-1} \cup \{j\}).
-\end{equation}
+which is the generalized likelihood ratio statistic for testing $M(E_{k-1})$ against the ``greedy'' alternative with one more active parameter, $\bigcup_{j \notin E_{k-1}} M(E_{k-1} \cup \{j\})$.
 The {\em selective greedy likelihood ratio test} rejects for large $G_k$, based on the law
@@
 \subsubsection{Ever-Active Regularization Paths and Next-Entry Tests}
-Another important class of model selection procedures is the sequence of {\em ever-active} sets for a regularized likelihood path, under a sparsity inducing regularizer.
+Another important class of model selection procedures is the sequence of {\em ever-active} sets for a regularized likelihood path, under a sparsity-inducing regularizer such as a scaled $\ell_1$ norm.
 The notion of ever-active sets is needed since these solution paths can drop  (as well as add) predictors along the way.
@@
 As an example, for a lasso penalty, this is $P_\lambda = \lambda\|\beta\|_1$ with $\lambda\in (0,\infty)$.
-
-\JTcomment{Yes, this looks a little funny to me too. Not all penalties produce
-sparsity (for example, the generalized LASSO). So, I think this section is really thinking 
-of $\ell_1$ norm (maybe a group LASSO norm). As written, 
-we are also assuming that
-there is a unique solution to (11) (which is not a big assumption but we might
-as well make it.) Finally, $\ell$ does not actually have to be a likelihood.
-BUT later on we implicitly assume that it is the likelihood for 
-$M_{\infty}$. Why not just say this here, i.e., when talking
-about $\ell_1$ regularization paths the models we have in mind are typically
-going to be ones which correspond to a likelihood for some ``largest model'' for which we are trying to find a sub-model.
-
-An example where we might not take $\ell$ to be a likelihood would be
-if it were a pseudo-likelihood as in the neighborhood selection 
-instead of graphical LASSO.
-}
-\WFcomment{Took out some of the confusing language.
-We are explicitly defining $\ell$ as the likelihood for $M_\infty$ at the beginning of 2.2.1.
-I have a remark later that $\ell$ doesn't really need to be a likelihood, but we can't accommodate completely general loss functions without losing some of the results.
-What we do ``need'' for the SSP is that $\ell(\theta)$ is a function of $U(Y; E)$ for $\theta \in \Theta(E)$.
-This is true for likelihoods but probably not true in general for pseudo-likelihoods.
-I'm not sure why you say this is ``really'' about $\ell_1$: it's also about elastic net, group lasso, nuclear norm, $\ell_0$ (such as in PCA), and much more.
-True, $\ell_1$ is the prototypical example, but I don't see how we get anything from explicitly restricting ourselves to $\ell_1$.}
-
-\RJTcomment{Hopefully without too much drama :), I've added the words ``sparsity-inducing regularizer'' above.
-This is because, at least as we've written it, we are thinking of the regularizer as selecting out a submodel.
-I understand that it may be more general than this, but our notation above and in the previous subsection all have to do with sparsity directly.}
 
@@
 \begin{equation}
-  \tE_\lambda(Y) = \left\{j:\; \hat\theta_j^\nu(Y) \neq 0 
-    \text{ for any } \nu\in \Lambda, \nu \geq \lambda \right\}
+  \tE_\lambda(Y) = \left\{j:\; \hat\theta_j^\gamma(Y) \neq 0 
+    \text{ for any } \gamma\in \Lambda, \gamma \geq \lambda \right\}
 \end{equation}
@@
 
-\JTcomment{The next-entry test is actually a little expensive to compute if we a finite set of $\lambda$  as, say, glmnet does. Should we not mention this? On the plus side, 
-it gives us tests that do not require LAR and can be used with glmnet. 
-Formally, to carry this out test, we need to draw samples of $Y$ conditioned
-on the necessary sufficient statistics satisfying the necessary constraints 
-(for LASSO, there $2(p-k)$ constraints on $X_{-E}^Ty$). Then, for each sample
-we restart glmnet and run until the active set changes, storing this
-value of $\lambda$. Then we compare our observed next-entry to this
-sample of $\lambda$'s.
-
-In the LAR-LASSO case, this is easier as there is an explicit formula
-to compute the next event. But, LAR is limited to LASSO whereas next-entry
-in principle works for group LASSO.
-
-Finally, the simplification of constraints we get really only works for
-almost only for $\ell_1$. For group LASSO, the simplification
-doesn't happen. The constraints after $k$ steps look like
-$$
-\left\{y: P_Ey=z, \|X_g^T(y - \hat{\mu}_l(z))\|_2 \leq \lambda_l \omega_g, 
-g \in -E, 1 \leq l \leq k, \|(I-P_E)y\|=c\right\}
-$$
-where $z$ is the projection of our observed $y$ onto the ever active set.
-So, for a given ``ever inactive group'', this says that $X_g^Ty$ is in
-an intersection of $k$ balls which does not necessarily have a simple structure.
-When all the groups are of size 1, i.e., in the LASSO, the intersection 
-of $k$ balls is the intersection of $k$ intervals and we get a simplification.
-
-All this to say, that we are really only talking about $\ell_1$ here, and 
-next-entry is essentially only practical BLAH
-}
-\WFcomment{Actually I don't think the computation is bad at all.
-The point is that you don't need to compute the distribution of $\lambda_k$, you only need the probability, under the null hypothesis, that $\lambda_k$ is larger than its observed value.
-So at least for convex penalized-likelihood problems, you should only ever have to compute one path --- the path for the observed data set --- and then store those fits.
-And then when you resample, you just check whether $Y^*$ satisfies the KKT conditions at $\hat\theta^\lambda(Y)$ for each $\lambda > \lambda_k$.
-Then $\lambda_k(Y^*)>\lambda_k(Y)$ iff the KKT conditions are violated for any $\lambda > \lambda_k(Y)$.
-(It is probably worth pointing this out in the computation section somewhere).
-
-For group lasso, it's the same situation -- we should only need to compute one path.
-Though it sounds like you think in that case the sampling itself is a problem.
-But at least for non-overlapping groups, is it really so bad?
-
-I think it is worth distinguishing between what procedures we can define and prove things about (which we discuss here), and what subset of those we currently have efficient algorithms for (which we discuss in the Computation section).
-After all, some problems are small, some scientists are willing to wait a while for their results, and we have various cudgels at our disposal such as accept/reject, Besag \& Clifford MCMC tests that are exact even if the chain doesn't mix, etc.
-etc.}
-
-\JTcomment{I agree that it is worth distinguishing ``provable''
-and ``computable''.
-I agree that the collection of things we can prove things about
-includes group LASSO perhaps other many penalties, but
-I would say
-that we only have a practical algorithm for computing
-when the penalty is $\ell_1$ in this paper. I'm not sure exactly
-how to fit nuclear norm here because I'm not sure what
-we should call the  ever active set when the loss is not Frobenius (i.e., unless
-we are doing PCA or SVT). That is, in matrix
-completion or reduced-rank regression I'm not sure what I would call the 
-ever-active set. For PCA or SVT, the spaces produced by the data's left and
-right singular vectors make up (a reasonable version of an) ever-active set.
-
-I also understand your argument that for $\ell_1$  the next entry test is less bad than I thought. 
-To check the KKT conditions at $\hat{\theta}^{\lambda}(Y)$ for $\ell_1$ is 
-straightforward, as $\hat{\theta}^{\lambda}(Y^*)$ has 
-an explicit candidate form so we have a very good guess
-at what the residuals at slightly later $\lambda$'s should look like and we
-can check the KKT conditions easily.
-For group LASSO, things are more complicated
-as we probably will not want to condition on the ``signs'' which are now
-a collection of unit vectors. If we don't condition on them
-then we have to solve group LASSO for response $Y^*$ until the set of
-active groups changes.
-Even if we do condition on them (which is OK
-as they are measurable with respect the right $\sigma$-algebra) the simplification for sampling that we observe for $\ell_1$ goes away. After $k$ steps on the
-discrete $\Lambda$ path I think we are back to having 
-$O(p*k)$ constraints roughly for any penalty that is not $\ell_1$. (For group
-LASSO, the $O(p*k)$ estimate assumes $k$ is not huge, and each group has roughly bounded number of features in it.) Finally, conditioning on the
-direction vectors needs a sampling algorithm that is not as simple as
-Gaussian subject to affine constraints. Describing this sampler is something
-I think we just don't want to touch in this paper.
-}
-
-\paragraph{Remark.} In its usual formulation, the lasso coefficients for linear regression minimize a penalized RSS criterion:
-\[
-\hat\beta_{\text{lasso}}^\lambda = \argmin_{\beta\in\R^p} \; \frac{1}{2}\|Y-X\beta\|_2^2 + \lambda\|\beta\|_1, 
-\]
-which is slightly different from the penalized log-likelihood criterion of \eqref{eq:pen-lik}.
-One can show that the lasso coefficients $\hat\beta_{\text{lasso}}^\lambda$, along with $\hat\sigma^2=n^{-1}\|Y-X\hat\beta^\lambda_{\text{lasso}}\|^2$, minimize
-\[
-\frac{n}{2\pi e}\exp\left\{-2\ell(\beta, \sigma^2;\,Y)/n\right\} + \lambda\|\beta\|_1.
-\]
-If $\Lambda=(0,\infty)$ then both methods result in exactly the same ever-active path.
-Otherwise, at the cost of some notational bloat, we could generalize \eqref{eq:pen-lik} by replacing $-\ell(\cdot)$ with $\rho(\ell(\cdot))$ for any strictly decreasing $\rho$, and the theory in this article would go through unchanged.
-\RJTcomment{I actually find this last remark more confusing than helpful, and I think we should take it out.
-I think it's pretty obvious that you can transform the loss by a monotone function ...
-and what is the $\exp$ doing up there?
-As it's written that's not a log-likelihood, that's a likelihood?}
-
-\begin{comment}
-This is slightly different from the penalized log-likelihood criterion of \eqref{eq:pen-lik}, which would minimize
-\begin{equation}\label{eq:lasso-beta-sigma}
-  -\ell(\beta, \sigma;\,Y) + P_\lambda(\beta)=
-  \frac{n}{2}\log(2\pi\sigma^2) + \frac{1}{2\sigma^2}\|Y-X\beta\|_2^2 
-  + \lambda\|\beta\|_1.
-\end{equation}
-
-If $\sigma^2$ is known, the two methods are equivalent up to rescaling $\lambda$.
-If $\sigma^2$ is unknown, then partially minimizing~\eqref{eq:lasso-beta-sigma} over $\sigma^2$ gives $\hat\sigma^2(\beta) = n^{-1}\|Y-X\beta\|_2^2$, leading to profile log-likelihood
-\[
--\ell^*(\beta;\,Y) \;\triangleq\; -\ell(\beta, \hat\sigma^2(\beta); \,Y)
-\;=\;  \frac{n}{2}
-\left(\log\left(\frac{2\pi\|Y-X\beta\|_2^2}{n}\right)+1\right)
-\]
-
-As a result the lasso coefficient vector $\hat\beta_{\text{lasso}}^\lambda$ minimizes
-\[
-\frac{n}{2\pi e}\exp\left\{-2\ell^*(\beta;\,Y)/n\right\} + \lambda\|\beta\|_1,
-\]
-while the regularized log-likelihood estimator $\hat\beta^\lambda$ minimizes $-\ell^*(\beta;\,Y)+\lambda\|\beta\|_1$.
-If $\Lambda=(0,\infty)$ then both methods result in exactly the same ever-active path.
-Otherwise, at the cost of some notational bloat, we could generalize \eqref{eq:pen-lik} by replacing $-\ell(\cdot)$ with $\rho(\ell(\cdot))$, where $\rho$ is any strictly decreasing function, and the theory in this article would go through unchanged.
-\end{comment}
-
-\begin{comment}
-\begin{align*}
-  \hat\sigma^2(\beta)
-  &\triangleq \argmax_{\sigma^2>0} \ell(\beta,\sigma^2;\,Y) 
-  = \frac{1}{n}\|Y-X\beta\|_2^2, \quad \text{ and }
-  \\
-  \ell^*(\beta; \,Y)
-  &\triangleq \sup_{\sigma^2>0} \ell(\beta, \sigma^2;\,Y) 
-  = -\frac{n}{2} - \frac{n}{2}\log\left(\frac{2\pi\|Y-X\beta\|_2^2}{n}\right).
-\end{align*}
-\end{comment}
+\paragraph{Remark.} In its usual formulation, the lasso coefficients for linear regression minimize a penalized RSS criterion.
+If we replace $-\ell$ with any strictly decreasing function of the log-likelihood, all of the results in this article hold without modification.
 
@@
 
-Methods like AIC \citet{akaike1974new} and BIC \citep{schwarz1978estimating} are designed for the non-adaptive case, where the sequence of models is determined in advance of observing the data.
-Cross-validation, another general-purpose algorithm for tuning parameter selection, targets out-of-sample error and tends to select many noise variables when the signal is sparse (e.g., \cite{LY2015}).
-\citet{benjamini2009simple} extend the AIC by using an adaptive penalty $\sigma^2\lambda_{k,p} k$ to select a model, based on generalizing the Benjamini-Hochberg procedure \citet{benjamini1995controlling}, but do not prove finite-sample control.
+Methods like AIC \citep{akaike1974new} and BIC \citep{schwarz1978estimating} are designed for the non-adaptive case, where the sequence of models is determined in advance of observing the data.
+Cross-validation, another general-purpose algorithm for tuning parameter selection, targets out-of-sample error and tends to select many noise variables when the signal is sparse \citep[e.g.][]{LY2015}.
+\citet{benjamini2009simple} extend the AIC by using an adaptive penalty $\sigma^2\lambda_{k,p} k$ to select a model, based on generalizing the Benjamini-Hochberg procedure \citep{benjamini1995controlling}, but do not prove finite-sample control.
 The stability selection approach of \citet{meinshausen2010stability} uses an approach based on splitting the data many times and offers asymptotic control of the FDR, but no finite-sample guarantees are available.
@@
 These procedures also assume the sequence of hypotheses is fixed, whereas in our setting the truth or falsehood of $H_k$ is random, depending on which model is selected at step $k$.
-Fortuitously, we show that the error guarantees for these stopping rules do transfer to the random-hypothesis setting, provided we have the same independence property conditional on the completion index $k_0$ (recall
-$k_0(Y) = \min\{k:\; F \in M_k(Y)\}$).
+In Section~\ref{sec:random-hyp} we show that the error guarantees for these stopping rules do transfer to the random-hypothesis setting, provided we have the same independence property conditional on the completion index $k_0$ (recall
+$k_0(Y, F) = \min\{k:\; F \in M_k(Y)\}$).
 
@@
 
-We now review several proposals for ordered testing of fixed null hypotheses along with their error guarantees in the case where the sequence of null hypotheses is fixed.
+We now review several proposals for ordered testing of fixed null hypotheses along with their error guarantees in the traditional setting, where the sequence of null hypotheses is fixed.
 In the next section we will extend the analysis to random null hypotheses.
 
-\subsubsection{BasicStop}
 The simplest procedure is to keep rejecting until the first time that $p_k > \alpha$: 
@@
 
-\subsubsection{ForwardStop}
-
-\citet{gsell2013sequential} propose the {\em ForwardStop} rule to control the FDR:
+To control the FDR, \citet{gsell2013sequential} propose the {\em ForwardStop} rule:
 \[
@@
 
-\subsubsection{Accumulation Tests and HingeExp}
-
-\citet{li2015accumulation} generalize ForwardStop, introducing the family of {\em accumulation tests}, defined as follows.
-Fix a function $h: [0,1] \rightarrow [0,\infty]$ satisfying \smash{$\int_{t=0}^1 h(t)dt=1$}, called an {\em accumulation function.} Using $h$ we can construct an estimate of FDP at step $k$:
-\begin{equation*}
-\widehat{\text{FDP}}_h(k)= \frac{\sum_{i=1}^k h(p_i)}{k},
-\end{equation*}
-and select the largest model for which \smash{$\widehat{\text{FDP}}_k \leq \alpha$}. Taking $h(p)=-\log(1-p)$ recovers the ForwardStop rule. Another choice they propose is the {\em HingeExp} function
-\begin{equation*}
-  h_{\rm HingeExp}(p)= 
-  \begin{cases} 
-    C\log\left(\frac{1}{C(1-p)} \right)& p> 1-1/C\\
-    0 & p\leq 1-1/C
-  \end{cases}
-\end{equation*}
+\citet{li2015accumulation} generalize ForwardStop, introducing the family of {\em accumulation tests}, which replace $-\log(1-p_i)$ with a generic {\em accumulation function} $h: [0,1] \rightarrow [0,\infty]$ satisfying \smash{$\int_{t=0}^1 h(t)dt=1$}.
 \citet{li2015accumulation} show that accumulation tests control a modified FDR criterion provided that the null $p$-values are $U[0,1]$, and are independent of each other and of the non-nulls.
 
-\subsection{Ordered Testing of Random Hypotheses}
+\subsection{Ordered Testing of Random Hypotheses}\label{sec:random-hyp}
 
@@
 In our problem setting, the sequence of selected models is random; thus, the truth or falsehood of each $H_k$ is not fixed, as assumed in the previous section.
-It is clear that we will recover all of the same guarantees if we construct our $p$-values conditional on the entire path $M_{0:d}(Y)$; however, this option is rather unappealing for both computational and statistical reasons.
-In constructing the $p$-value for step $k$, is it sufficient to condition on the subpath of models $M_{0:(k-1)}(Y)$, or better yet, only on the current null model $M_{k-1}$?
-
-We note that conditioning on the current null model {\em is} enough to guarantee that $p_k$ is uniform conditional on the event $\{F\in M_{k-1}(Y)\}$ that the $k$th null is true.
-Unfortunately, however, it is {\em not} enough to guarantee that the above stopping rules actually control FDR or FWER.
+Trivially, we can recover the guarantees from the fixed setting if we construct our $p$-values conditional on the entire path $M_{0:d}(Y)$; however, this option is rather unappealing for both computational and statistical reasons.
+In this section we discuss what sort of conditional control the single-step $p$-values must satisfy to recover each of the guarantees in Section~\ref{sec:orderedProposals}.
+
+We note that conditioning on the current null model does guarantee that $p_k$ is uniform conditional on the event $\{F\in M_{k-1}(Y)\}$, the event that the $k$th null is true.
+Unfortunately, however, it does {\em not} guarantee that the stopping rules of Section~\ref{sec:orderedProposals} actually control FDR or FWER.
 Let $z_{\alpha/2}$ denote the upper-$\alpha/2$ quantile of the $\cN(0,1)$ distribution, and consider the following counterexample.
@@
 A proof of Proposition~\ref{prop:counterexample} is given in the appendix.
-The crucial issue is that $k_0$ is no longer a fixed index.
+The problem is that $k_0$ is no longer a fixed index.
 Consequently, even though $\L(p_k \mid H_k \text{ true})=U[0,1]$ for each fixed $k$, the $p$-value \smash{$p_{k_0}$} corresponding to the {\em first} true null hypothesis is stochastically smaller than uniform.
-
-If, however, we could guarantee that
+In this counterexample, $k_0=2 \iff |Y_3|>z_{\alpha/2}$, giving $\text{FWER}\eqAS 1$ on the event $k_0=2$ and leaving no room for error when $k_0\neq 2$.
+
+If, however, we could guarantee that for each $k=1,\ldots, d$,
 \[
@@
 \]
-then BasicStop again would control FWER, by the same argument as given in \eqref{eq:basic-stop-control}.
-Note that $k_0$ is unknown, so we cannot directly condition on its value when we construct our single-step $p$-values.
-However, because
+then BasicStop again would control FWER, by \eqref{eq:basic-stop-control}. Note that $k_0$ is unknown, so we cannot directly condition on its value when constructing $p$-values. However, because
 \[
@@
 \]
-it is enough to condition on $(M_{k-2}, M_{k-1})$ when we construct our $p$-values.
+it is enough to condition on $(M_{k-2}, M_{k-1})$.
 As we will see in Section~\ref{sec:pValsIndep}, conditioning on $(M_{k-1}, U_{k-1})$ is equivalent to conditioning on $(M_{0:(k-1)}, U_{k-1})$ in most cases of interest including forward stepwise likelihood paths and ever-active regularization paths (but not the path in Proposition~\ref{prop:counterexample}).
@@
   \[
-  \Err_g = \E\left[ g\left(\hk(p_{1:d}(Y)), k_0(Y)\right)\right],
+  \Err_g = \E\left[ g\left(\hk(p_{1:d}(Y)), k_0(Y, F)\right)\right],
   \]
@@
   \[
-  \E\left[ g\left(\hk(p_1,\ldots,p_d), k_0(Y)\right) \;\mid\;
-      k_0(Y), \; p_{1:k_0}(Y)\right] \leqAS \alpha.
+  \E\left[ g\left(\hk(p_1,\ldots,p_d), k_0(Y, F)\right) \;\mid\;
+      k_0(Y, F), \; p_{1:k_0}(Y)\right] \leqAS \alpha.
   \]
@@
 
-Now, if $F\in M(E_{k-1})$, then $p_k$ is uniform and independent of the previous $p$-values $p_{1:(k-1)}$, for two reasons:
+If $F\in M(E_{k-1})$, then $p_k$ is uniform and independent of the previous $p$-values $p_{1:(k-1)}$ since:
 \begin{enumerate}
@@
 
-Next we will explain why $p_1,\ldots,p_{k-1}$ are functions of $E_{k-1}$ and $U_{k-1}$.
-Knowing $E_{k-1}=E$ tells us that the $k-1$ variables in $E$ are selected first, and knowing \smash{$(X_{E_{k-1}}'Y,\|Y\|^2)$} is enough information to compute the $t$-statistics $t_{j,D}$ for all $j\in E$ and $D\sub E$.
-As a result, we can reconstruct the order in which those $k-1$ variables were added, since for $i<k$,
-\[
-j_i = \argmax_{j \notin E_{i-1}} \, |t_{j,E_{i-1}}| 
-= \argmax_{j \in E\setminus E_{i-1}} \, |t_{j,E_{i-1}}| \quad \text{ on the event } \{E_{k-1}=E\}.
-\]
-In general, we say that the path $M_{0:d}$ follows the {\em subpath sufficiency principle} (SSP) if for each $k$ the subpath $M_{0:k}$ is a function of $(M_k,U_k)$.
-\RJTcomment{A thought / suggestion: we define the SSP here, but I think it's not really that helpful, because we're only talking about max-t.
-We refer to SSP in the next section; in defining the sufficient filtration.
-Why not just define SSP there?
-In general, I think there's quite a lot of lingo being thrown around, and though I like all the new terms, I think we could do a little better job consolidating where they're defined.}
-
-Furthermore, for $i<k$, $p_i$ is computed by comparing $T_i=|t_{j_i, E_{i-1}}|$ to the upper $\alpha$ quantile of the distribution $\L(T_i \mid E_{i-1}, U_{i-1})$, computed under the assumption $F\in M(E_{i-1})$.
-The conditional null distribution is a function of $(E_{i-1}, U_{i-1})$, and as we have already seen, the test statistic $T_i$ is a function of $(E_i, U_i)$, both of which are functions of $(E_{k-1},U_{k-1})$.
-
-Because $(E_i,U_i)$ is a function of $(E_k,U_k)$ for $i<k$, we can characterize the above discussion in terms of the filtration $\sF_{0:d}$, where
-\[
-\sF_k = \sF(E_{0:k}, U_k),
-\]
-the $\sigma$-algebra generated by random variables $(E_k, U_k)$. When the SSP is in force, we can equivalently write $\sF_k=\sF(E_k, U_k)$, or $\sF_k=\sF(M_k, U_k)$.
-
-At each $k$, given that $H_k$ is true, $\L(p_k \mid \sF_{k-1})\sim U[0,1]$, and $p_k$ is $\sF_k$-measurable.
-If $H_k$ is true, then so are $H_{k+1},\ldots, H_d$, and applying this conclusion iteratively we can see that all of the remaining $p$-values are uniform and independent.
+Next we will see why $p_1,\ldots,p_{k-1}$ are functions of $(E_{k-1},U_{k-1})$.
+Observe that knowing $E_{k-1}=E$ tells us that the $k-1$ variables in $E$ are selected first, and knowing \smash{$U_{k-1}=(X_{E_{k-1}}'Y,\|Y\|^2)$} is enough information to compute the $t$-statistics $t_{j,D}$ for all $j\in E$ and $D\sub E$.
+As a result, we can reconstruct the order in which those $k-1$ variables were added.
+In other words, $(E_i, U_i)$ is a function of $(E_{k-1}, U_{k-1})$ for all $i<k$.
+
+Furthermore, for $i<k$, $p_i$ is computed by comparing $T_i=|t_{j_i, E_{i-1}}|$, which is a function of $(E_i, U_i)$, to the reference null distribution $\L_{H_i}(T_i \mid E_{i-1}, U_{i-1})$, which is a function of $(E_{i-1}, U_{i-1})$.
+Having verified that under $H_k$, $p_k$ is uniform and independent of $p_{1:{k-1}}$, we can apply this conclusion iteratively to see that all remaining $p$-values are also uniform and independent.
 
@@
 
-A {\em selective $p$-value} $p_k$ for testing $H_k:\; F \in M_{k-1}(Y)$ satisfies, for any $F\in M_{\infty}$,
+Define the {\em sufficient filtration} for the path $M_{0:d}$ as the filtration $\sF_{0:d}$ with 
+\[
+\sF_k = \sF(M_{0:k}, U_k).
+\]
+By our assumption of minimal sufficiency, $\sF_i \sub \sF_k$ for $i\leq k$.
+
+For most path algorithms of interest, including forward stepwise and regularized likelihood paths, observing $(M_{k-1}, U_{k-1})$ is equivalent to observing $(M_{0:k-1},U_{k-1})$, because knowing $(M_{k-1}, U_{k-1})$ is enough to  reconstruct the subpath $M_{0:k-1}$.
+We say that $M_{0:d}$ satisfies the {\em subpath sufficiency principle} (henceforth SSP) if \begin{equation}\label{eq:SSP}
+  \sF(M_{0:k}, U_k) = \sF(M_k, U_k), \quad k=0,\ldots, d.
+\end{equation}
+
+A valid {\em selective $p$-value} $p_k$ for testing $H_k:\; F \in M_{k-1}(Y)$ satisfies, for any $F\in M_{\infty}$,
 \[
@@
 \begin{enumerate}
-\item $M_k(Y)$ and $p_k(Y)$ are measurable with respect to $\sF_k$, and 
-\item $p_k(Y)$ is uniform or super-uniform given $\sF_{k-1}$ on the event $\{F\in M_{k-1}(Y)\}$.
+\item $p_k(Y)$ is super-uniform given $\sF_{k-1}$ on the event $\{F\in M_{k-1}(Y)\}$, and
+\item $M_k(Y)$ and $p_k(Y)$ are measurable with respect to $\sF_k$.
 \end{enumerate}
-If we think of $\sF_k$ as representing information available at step $k$, then the first condition means that $p_k$ ``excludes'' whatever evidence may have accrued against the null by step $k-1$, and the second means that any information revealed after step $k$ plays no role in determining $p_k$.
+
+If we think of $\sF_k$ as representing information available at step $k$, then the first condition means that $p_k$ ``excludes'' whatever evidence may have accrued against the null by step $k-1$, and the second means that any information revealed after step $k$ is likewise irrelevant to determining $p_k$.
 Separated $p$-values are independent on nulls, as we see next.
@@
   If the $p$-values are exact then $p_{k+1}, \ldots, p_d$ are
-  independent and uniform given $\sF_{k_0}$ on the event $\{k_0=k\}$.
+  independent and uniform given $\sF_k$ on the event $\{k_0=k\}$.
   If they are super-uniform, then for all
   $\alpha_{k+1},\ldots,\alpha_d \in [0,1]$,
-  \begin{equation}\label{eq:jointConserv}
-  \P_F\left(p_{k+1}\leq \alpha_{k+1}, \ldots, p_d \leq \alpha_d, \; k_0 \leq k \;
+  \begin{equation}\label{eq:jointSuper}
+  \P_F\left(p_{k+1}\leq \alpha_{k+1}, \ldots, p_d \leq \alpha_d, \;
     \mid\; \sF_k\right) \leqAS \left(\prod_{i=k+1}^d
-  \alpha_i\right) 1_{\{k_0 \leq k\}}.
+  \alpha_i\right) \quad \text{ on } \{k_0 = k\}.
   \end{equation}
@@
 \begin{proof}
-  We first prove~\eqref{eq:jointConserv}
-  by induction. Define $B_i = \{p_i \leq \alpha_i\}$. 
-  The base case is
+  Noting that $\{k_0=k\}$ is $\F_k$-measurable, it is enough to show that 
+  \begin{equation}\label{eq:jointSuper2}
+  \P_F\left(p_{k+1}\leq \alpha_{k+1}, \ldots, p_d \leq \alpha_d, \;
+    \mid\; \sF_k\right)\1_{\{k_0\leq k\}} \leqAS \left(\prod_{i=k+1}^d
+  \alpha_i\right)\1_{\{k_0\leq k\}}
+  \end{equation}
+  We now prove~\eqref{eq:jointSuper2} by induction. Define $B_i = \{p_i \leq \alpha_i\}$. The base case is
   \[
@@
   which is true by construction of $p_d$. 
-  For the inductive case, note that the
-  conditional probability on the left-hand side
-  of~\eqref{eq:jointConserv} is
+  For the inductive case, note that 
   \begin{align*}
@@
 
-Define the {\em sufficient filtration} for the path $M_{0:d}$ as the filtration $\sF_{0:d}$ with $\sF_k = \sF(M_{0:k}, U_k)$.
-By our assumption of minimal sufficiency, $U_i$ is always a function of $(M_i, U_i)$ for $i<k$.
-The path algorithm satisfies the SSP if and only if
-\begin{equation}\label{eq:SSP}
-  \sF(M_{0:k}, U_k) = \sF(M_k, U_k), \quad k=0,\ldots, d
-\end{equation}
-The sufficient filtration separates $p_{1:d}$ if and only if 
-\begin{enumerate}
-\item $p_k$ is super-uniform given $M_{0:(k-1)}$ and $U_{k-1}$, and 
-\item $p_k$ is a function of $M_{0:k}$ and $U_k$.
-\end{enumerate}
+The sufficient filtration separates $p_{1:d}$ if and only if (1) $p_k$ is super-uniform given $M_{0:(k-1)}$ and $U_{k-1}$, and (2) $p_k$ is a function of $M_{0:k}$ and $U_k$.
+
 To be a valid selective $p$-value for a single step, $p_k$ only needs to be super-uniform conditional on $M_{k-1}$.
-In general, it is more stringent to additionally require that $p_k$ must condition on the entire subpath $M_{0:(k-1)}$ as well as $U_{k-1}$.
-In most cases of interest, however, (i) we will want to condition on $(M_{0:(k-1)}, U_{k-1})$ anyway, to eliminate nuisance parameters of $M_{k-1}$, and (ii) the path satisifies the SSP, so conditioning on $(M_{0:(k-1)}, U_{k-1})$ is equivalent to conditioning on $(M_{k-1}, U_{k-1})$.
-Thus, most single-step selective $p$-values we would contemplate using are already exact conditional on $\sF_{k-1}$.
-The second requirement, that $p_k$ be $\sF_k$-measurable, has more bite.
-For example, we will see in Section~\ref{sec:selective-reg} that it excludes saturated-model tests in linear regression.
-
-Collecting together the threads of this section, we arrive at our main result: a general sufficient condition for a $p$-value sequence to be independent on nulls.
+It may appear more stringent to additionally require that $p_k$ must condition on the entire subpath $M_{0:(k-1)}$ as well as $U_{k-1}$, but in practice there is often no difference: if $M_{0:d}$ satisfies the SSP and each $U_{k-1}$ is a complete sufficient statistic, then {\em every} exact selective $p$-value $p_k$ is also uniform conditional on $\sF_{k-1}$.
+
+The requirement that $p_k$ must be $\sF_k$-measurable has more bite. For example, we will see in Section~\ref{sec:selective-reg} that it excludes saturated-model tests in linear regression.
+
+Collecting together the threads of this section, we arrive at our main result: a sufficient condition for $p_{1:d}$ to be independent on nulls.
 \begin{theorem}[Sufficient Condition for Independence on Nulls] \label{thm:suffCond}
@@
 
-Note that if our path algorithm did not satisfy the SSP, we could in many cases still repair the situation by constructing $p$-values that are uniform conditional on $(M_{0:(k-1)},U_{k-1})$, not just conditional on $(M_{k-1},U_{k-1})$.
-
-\subsection{Independence for Greedy Likelihood and Next-Entry $p$-values}\label{sec:SSP}
+\paragraph{Remark} If our path algorithm does not satisfy the SSP, we can repair the situation by constructing $p$-values that are uniform conditional on $(M_{0:k-1},U_{k-1})$.
+
+\subsection{Independence for Greedy Likelihood and Next-Entry $p$-values}\label{sec:indep-greedy-entry}
 
 In this section, we apply Theorem~\ref{thm:suffCond} to establish that in the generic sparse parametric model of Section~\ref{sec:genSparse}, the forward stepwise path and all ever-active regularization paths satisfy the SSP.
-Moreover, the greedy likelihood ratio statistic $G_k$ is $\sF_k$-measurable with respect to the sufficient filtration of the forward stepwise path, and the next-entry statistic $\lambda_k$ is $\sF_k$-measurable with respect to the sufficient filtration of the regularized-likelihood path.
-As a result, the selective greedy likelihood $p$-values in the forward stepwise path are independent on nulls, as are the next-entry $p$-values in the regularized-likelihood path, per Theorem~\ref{thm:suffCond}.
-
-Define the restricted log-likelihood
+Moreover, the greedy likelihood ratio statistic $G_k$ and the next-entry statistic $\lambda_k$ are $\sF_k$-measurable with respect to the sufficient filtrations of the forward stepwise and ever-active regularization paths, respectively.
+As a result, the $p$-value sequences in each setting are independent on nulls, per Theorem~\ref{thm:suffCond}.
+
+We begin by proving both paths satisfy the SSP
+\begin{proposition}\label{prop:forwardSSP}
+Forward stepwise paths and ever-active regularization paths both satisfy the subpath sufficiency principle.
+\end{proposition}
+
+\begin{proof}
+First define the restricted log-likelihood
 \begin{equation*}
@@
 \end{equation*}
-
-We begin by stating our result for forward stepwise.
-\begin{proposition}\label{prop:forwardSSP}
-Forward stepwise paths satisfy the subpath sufficiency principle.
-\end{proposition}
-\begin{proof}
+For some fixed step $k$ and active set $E$, denote the event $A=\{M_k = M(E)\}$.
+Conditioning on $U$, the restricted likelihood is proportional to a function depending only on $U = U(Y; \;M(E))$.
+The log-likelihood decomposes as
+\[
+\ell_E(\theta; Y) = \ell_E^{U}(\theta; U) 
++ \ell_E^{Y \mid U}(Y \mid U).
+\]
+The second term, the log-likelihood of $Y$ given $U$, does not depend on $\theta$ because $U$ is sufficient.
+
 Recall the forward stepwise path is defined by
@@
   \end{equation}
-  For some fixed step $k$ and active set $E$, denote the event $A=\{M_k = M(E)\}$.
-Conditioning on $U$, the restricted likelihood is proportional to a function depending only on $U = U(Y; \;M(E))$.
-The log-likelihood decomposes as
-  \[
-  \ell_E(\theta; Y) = \ell_E^{U}(\theta; U) 
-  + \ell_E^{Y \mid U}(Y \mid U).
-  \]
-  The second term, the log-likelihood of $Y$ given $U$, does not depend on $\theta$ because $U$ is sufficient for $M(E)$.
 
@@
 As a result, it also follows that the entire sequence $M_{0:k}(Y)$ depends only on $(M_k, U_k)$.
+
+As for ever-active regularization paths, if we denote 
+\[
+\hat\theta^{(E,\lambda)} = \argmin_{\theta\in\Theta(E)} \; -\ell_E^{U}(\theta; U(Y)) + P_\lambda(\theta),
+\]
+then 
+\[
+\hat\theta^{(E,\lambda)} \eqAS \hat\theta^{\lambda} \text{ on } 
+A \cap 1_{\{\lambda \geq \lambda_k\}}.
+\]
+But \smash{$\hat\theta^{(E,\lambda)}$} depends only on $U(Y)$.
+Therefore, on $A$, we can reconstruct the entire path of solutions for $\{\lambda\in \Lambda:\; \lambda\geq \lambda_k\}$, once we know $(M_k,\,U_k)$.
 \end{proof}
 
-\JTcomment{Are we assuming that $U(Y;M)$ is minimal sufficient but not 
-necessarily minimal
-complete sufficient in the above proposition? The statement doesn't say
-what $U$ satisfies. I ask because in the corollary we do say something
-about completeness.  Our earlier assumption was just that $U$ always
-denotes a minimal sufficient statistic.
-}
-\WFcomment{Yes, it is assumed at the beginning of the paper that $U$ is minimal sufficient.
-Perhaps I should just say complete sufficient at the beginning so I don't need to strengthen the assumption later...}
+As a direct consequence of Proposition~\ref{prop:forwardSSP}, greedy likelihood $p$-values for the forward-stepwise path, and next-entry $p$-values for ever-active regularization paths, are independent on nulls.
 
@@
 
-The analogous result hold for ever-active regularized likelihood paths.
-
-\begin{proposition}\label{prop:regPathSSP}
-Ever-active regularized likelihood paths satisfy the
-subpath sufficiency principle.
-\end{proposition}
-
-\begin{proof}
-  Again, for fixed $k$ and $E$, let $A=\{M_k = M(E)\}$, and define $U=U(Y;\, M(E))$. If we denote 
-\[
-\hat\theta^{(E,\lambda)} = \argmin_{\theta\in\Theta(E)} \; -\ell_E^{U}(\theta; U(Y)) + P_\lambda(\theta),
-\]
-then 
-\[
-\hat\theta^{(E,\lambda)} \eqAS \hat\theta^{\lambda} \text{ on } 
-A \cap 1_{\{\lambda \geq \lambda_k\}}.
-\]
-But \smash{$\hat\theta^{(E,\lambda)}$} depends only on $U(Y)$.
-Therefore, on $A$, we can reconstruct the entire path of solutions for $\{\lambda\in \Lambda:\; \lambda\geq \lambda_k\}$, once we know $(M_k,\,U_k)$.
-\end{proof}
-
 \begin{corollary}\label{cor:nextLambdaIndep}
@@
   \]
-  and that the \smash{$\tE_{\lambda}$} are nested by construction and finite by assumption. As a result, there is some $\nu_k\in \Lambda_k$ for which \smash{$\tE_{\nu_k} = E_k$}.
-
-  Furthermore, for all $\lambda\in\Lambda, \lambda\geq \nu_k$, we have \smash{$\tE_\lambda \sub E_k$}.
+  and that the \smash{$\tE_{\lambda}$} are nested by construction and finite by assumption. As a result, there is some $\gamma_k\in \Lambda_k$ for which \smash{$\tE_{\gamma_k} = E_k$}.
+
+  Furthermore, for all $\lambda\in\Lambda, \lambda\geq \gamma_k$, we have \smash{$\tE_\lambda \sub E_k$}.
 As a result, \smash{${\hat\theta^{\lambda} =\hat\theta^{(E_k,\lambda)}}$} for such $\lambda$, so $\tE_{\lambda}$ depends only on \smash{$\ell_{E_k}^{Y\mid U}(Y;\, U_k)$}, and therefore $\lambda_k$ can be computed from $(E_k, U_k)$.
@@
 Many conditional selective tests for linear regression use the saturated-model framework.
-See for example \citet{lockhart2014significance}, \citet{taylor2013tests}, \citet{taylor2014exact}, \citet{lee2013exact}, and \citet{loftus2014significance}.
-\RJTcomment{So, I would claim that the covariance test is NOT a saturated model test, but instead a select model test.
-Of course it doesn't really fit into this framework with exact p-values, but if you look at what we proved, it was a selected model null result.
-Also, Taylor (2013) tests the global null throughout, so again, I don't think it makes much sense to say that it's a saturated model test ...} Define the saturated model as
+See for example \citet{taylor2013tests}, \citet{taylor2014exact}, \citet{lee2013exact}, and \citet{loftus2014significance}.
+Define the saturated model as
 \[
@@
 Because the parameter $\mu$ has the same dimension as the data $Y$, we must assume that $\sigma^2$ is known.
-If this is not the case, we can estimate $\sigma^2$ and apply the test while plugging in the estimated value; this is what we have done with the diabetes data to compute the $p$-values in Table~\ref{tab:diab}.
+If $\sigma^2$ is unknown, we can estimate $\sigma^2$ and apply the test while plugging in the estimated value; this is what we have done with the diabetes data to compute the $p$-values in Table~\ref{tab:diab}.
 \citet{tibshirani2015uniform} discuss this and other strategies for unknown $\sigma^2$.
 
-Saturated-model tests assume only that $M_{\text{sat}}$ is true, and perform inference on selected linear functionals of $\mu$.
-For example, in linear regression after selecting active set $E$, we could construct confidence intervals for the least-squares coefficients $\theta^E$ of active set $E$, defined as
-\[
-\theta^E \;\;=\;\; \argmin_{\theta\in\R^{|E|}} \; \|\mu - X_E\theta\|_2 \;\;=\;\; (X_E'X_E)^{-1}X_E'\mu.
-\]
-In the context of sequential model selection, we can use the saturated-model framework to test the incremental null 
-\[
-H^{\text{inc}}:\; \theta_{j_k}^{E_{k-1}\cup \{j_k\}} = 0.
-\]
+Saturated-model tests perform inference on selected linear functionals of $\mu$.
+In the context of sequential linear model selection,  we can use the saturated-model framework to test the incremental null
+\[
+H^{\text{inc}}:\; \theta_{j_k}^{E_k} = 0,
+\]
+where $\theta^E$ minimizes \smash{$\|\mu - X_E\theta\|_2$}.
+
 Let $\eta_k\in \R^n$ denote the vector for which \smash{$\theta_{j_k}^{E_k} = \eta_k'\mu$}, and let \smash{$\proj_{\eta_k}^\perp$} denote the projection operator into the space orthogonal to $\eta_k$.
-The saturated model has $n$-dimensional natural parameter $\mu$ and complete sufficient statistic $Y$, and the UMPU saturated-model test for $\eta_k'\mu$ is based on the distribution
+The UMPU saturated-model test for $\eta_k'\mu$ is based on the distribution
 \begin{equation*}
-\L\left( \eta_k'Y \mid E_{k-1}, \;j_k, \;\proj_{\eta_k}^\perp Y \right),
+\L\left( \eta_k'Y \mid E_{k-1}, \;j_k, \;\proj_{\eta_k}^\perp Y \right).
 \end{equation*}
-or equivalently on
-\begin{equation}\label{eq:satModel}
-\L\left( X_{j_k}'Y \mid E_{k-1}, \;j_k, \;\proj_{\eta_k}^\perp Y \right).
-\end{equation}
-The pair $(E_{k-1}, j_k)$ defines the selection event and \smash{$\proj_{\eta_k}^\perp Y$} eliminates the nuisance parameter \smash{$\proj_{\eta_k}^\perp \mu$}.
+
 While the test statistic $X_{j_k}'Y$ is measurable with respect to $\sF_k$, the null distribution we compare it against is not: it depends on \smash{$\proj_{\eta_k}^\perp Y$}, which is not measurable with respect to $\sF_k$.
@@
 represents a more powerful option when testing for inclusion of variable $j_k$.
-While saturated-model inference always uses the same model $M_{\text{sat}}$, selected-model inference uses the lower-dimensional statistical model that is chosen by our selection procedure, similarly to data splitting.
+While saturated-model inference always uses the same model $M_{\text{sat}}$, selected-model inference uses the lower-dimensional statistical model chosen by our selection procedure, similarly to data splitting.
 In the sequential case this means that at step $k$, we test the selected null model $M(E_{k-1})$, either against the next selected model $M(E_k)\setminus M(E_{k-1})$ or against the upper model $M_\infty\setminus M(E_{k-1})$.
@@
 If $\sigma^2$ were unknown we would also need to condition on $\|Y\|^2$.
-We can construct $p$-values from the UMPU, equal-tailed, or one-sided test based on~\eqref{eq:selModel_cond}; the only difference is in how much of the overall level $\alpha$ is apportioned to the right or left tail.
+We can construct $p$-values from the UMPU, equal-tailed, or one-sided test based on~\eqref{eq:selModel_cond}; the only difference is in how much of the overall level $\alpha$ is apportioned to the right or left tail \citep{fithian2014optimal}.
 If we use the construction in~\eqref{eq:selModel_cond} after a subpath-sufficient path algorithm, the resulting $p$-values are always guaranteed to have independence on nulls.
@@
 
-\subsubsection{Classification of Proposed Tests}
-Table \ref{tab:testsummary} classifies a number of different tests that have been proposed for least angle regression, lasso and forward stepwise regression.
-Some focus on the complete null hypothesis, others on the incremental null (as defined in Section \ref{sec:whichnull}).
-Some use the saturated-model, and others the selected-model framework, as detailed in earlier in the section.
-The saturated-model tests are simpler and do not require sampling, while the selected-model tests are generally more powerful.
-In the next section we compare the selected- and saturated-model settings in more detail.
-
-\RJTcomment{I don't mean to sound overtly critical but I don't find this table helpful; in fact I find it confusing.
-Why is the covariance test called a saturated model test?
-Sure, it tests the complete null, but why does it use a saturated model framework?
-I doesn't use the law at all as described in \eqref{eq:satModel}.
-Also, the name ``spacings'' is reserved for a very specific approximation to the saturated-model test in LAR, that only uses knots in the LAR path, so I wouldn't use this name.
-I think the last 3 subsections have explained the differences nicely, and I think this table doesn't really add much beyond them, and confuses me.
-
-Also, I don't know why we give the example below.
-What does it show?
-More examples follow in the next section, adn more detailed experiments follow later, and this just shows that the covariance test is conservative, but by now, we know this and have showed this many, many times over.
-And it's not at all the point of this paper.}
-
-\begin{table}
-\begin{center}
-\begin{tabular}{ | l | r|r| r}
-        \hline
-Procedure& Complete Null & Incremental Null\\
-\hline
-LAR/lasso& Covariance test/Saturated&Spacing test/Saturated \\
-        & next-entry test/Selected& \\
-        \hline
-Forward Stepwise &max-$t$, max-$t$-identify/Selected &Gaussian pivot/Saturated\\
-\hline
-\end{tabular}
-\end{center}
-\caption{\em Summary of different tests that have been proposed for least angle regression, lasso and forward stepwise regression.
-Tests are described by the null hypothesis that they test (complete or incremental)  and the underlying model (saturated or selected).
-The max-$t$, max-$t$-identify and next-entry tests are proposed in this paper, while the covariance test was proposed in \citet{lockhart2014significance}, and the spacing and Gaussian
-pivot tests are given in \citet{taylor2014exact}. The latter paper shows that the covariance and spacing test are asymptotically equivalent for the first null step, when the complete null holds.}
-\label{tab:testsummary}
-\end{table}
-Figure  \ref{fig:larexample} shows an example of the LAR/lasso tests with $n=50, p=10$ and two strong signals.
-We plot the $p$-values only for those realizations in which the correct model was selected in the first two steps (about 95\% of the time).
-We see that the next-entry test is the most powerful, especially at step 1.
-The next-entry and spacing test have the correct type I error in steps 3 and 4, while the covariance test becomes conservative in step 4, as we move further into the null part of the sequence.
-\begin{figure}
-  \includegraphics[width=\textwidth]{figs/larexample.pdf}
-  \caption[fig:larexample]{\em  Cumulative distribution functions of $p$-values for LAR/lasso tests from simulations with $N=50,  p=10$ and two strong signals.}
-  \label{fig:larexample}
-\end{figure}
-
 \subsection{Comparison of Selected- and Saturated-Model Inference}\label{sec:bivariate}
@@
 For example, a 95\% confidence interval for $\theta_{j_k}^{E_k}$ could cover that coefficient with probability 95\% even if $M(E_k)$ is misspecified.
-In the setting of sequential model selection, this advantage is less important: we are not looking for a scientifically interpretable interval for $\theta_{j_k}^{E_k}$; rather, the goal of the test at step $k$ is to reject $M_{k-1}$ and move on if it is demonstrably incompatible with the data.
-If there are many remaining strong signal variables to be found, then both $M_k$ and $M_{k-1}$ are badly misspecified; the more true this is, the stronger the grounds are for rejecting $M_{k-1}$.
-
-\JTcomment{I would say our goal is a good stopping rule AND inference on parameters of selected (or perhaps saturated) model.
-We do not consider it here because we don't know how to do it, not because it is not interesting.}
-
-\RJTcomment{I don't see why what you mention is a disadvantage of the saturated-model perspective.
-In fact, I would argue that it is an advantage.
-The model is always going to be misspecified.
-What is being covered is a well-defined quantity, regardless of the true model.
-That is precisely why I like the saturated-model perspective for confidence intervals...}
+In the setting of sequential model selection, this advantage is less important: we are not constructing intervals for $\theta_{j_k}^{E_k}$; rather, the goal of the test at step $k$ is to reject $M_{k-1}$ and move on if it is demonstrably incompatible with the data.
 
@@
 
-In many cases it is easy to sample $Y$ given $U_{k-1}$, so computational difficulties only arise insofar as the selection event $\{M_{k-1}(Y)=M\}$ is difficult to condition on.
+In many cases it is easy to sample $Y$ given $U_{k-1}$, in which case computational difficulties only arise insofar as the selection event $\{M_{k-1}(Y)=M\}$ is difficult to condition on.
 Due to the generality in which we have posed the problem, we cannot give a general prescription for how to carry out this sampling.
@@
 For a completely general selection algorithm, we might expect severe computational difficulties when we attempt to condition on the null model at stage $k$, $M_{k-1}(Y)$.
-For example, in forward stepwise regression, $M_{k-1}$ is the accumulated result of $k-1$ steps of greedy variable selection.
+For example, in forward stepwise regression, $M_{k-1}$ is the cumulative result of $k-1$ steps of greedy variable selection.
 At each step $s<k$, the selection of variable $j_s$ is a result of comparing its $t$-statistic with the $t$-statistics of the other $p-k$ inactive variables.
-Based on this logic, we might expect to accumulate at least $O(p)$ constraints at each step, resulting in $O(kp)$ constraints after $k$ steps.
-
-For lasso linear regression, the situation seems even more dire: for any fixed value of $\lambda$, the event \smash{$\{\supp(\hat\beta^\lambda) = E\}$} is a union of up to $2^{|E|}$ polytopes in $\R^n$, each satisfying $O(p)$ constraints \citep{lee2013exact}.
-\footnote{To finesse this issue, \citet{lee2013exact} propose conditioning on the signs of $\hat\beta^\lambda$ as well.
-If we do this, the conditioning set becomes a single polytope.}.
-But the event $\{E_{k-1}(Y)=E\}$ depends on $\supp(\hat\beta^\lambda)$ for every $\lambda > \lambda_k$ --- i.e., on the solution to {\em infinitely many} lasso problems.
-
-\RJTcomment{I don't get this last paragraph really.
-First of all, just to point out, for each of the forward stepwise, LAR, and lasso paths, we condition on the signs in our 2014 paper.
-So this issue is not specific to the lasso.
-Also, why are we talking about lasso for fixed $\lambda$ here?
-This is confusing.
-We have exact inference for steps along the lasso path (not the same problem as fixed $\lambda$ lasso) in our 2014 paper.
-Our constructions give $\approx 2pk$ constraints for forward stepwise after $k$ steps, and $\approx 3pk$ constraints for LAR r lasso after $k$ steps.
-Now, what is given below is still a big improvement over that!
-But it's worth being very clear about the predecessors to it.}
+Based on this logic, we might expect to accumulate at least $O(p)$ constraints at each step, resulting in $O(kp)$ constraints after $k$ steps \citep{taylor2014exact}.
 
@@
 \subsection{Constraints}\label{sec:constraints}
-Happily, it turns out that the conditioning polytope can be expressed with only $2(p-k)$ constraints in forward stepwise regression and the lasso, as we show here.
+We now show that the conditioning polytope can be expressed with only $2(p-k)$ constraints in forward stepwise regression and the lasso.
 This increases the speed of both accept/reject and hit-and-run sampling for our problem.
@@
 The proof writes things in terms of nonstrict inequalities, and I think it makes sense to drop them; in the case of ties, $E$ is still a valid selected model.}
+\WFcomment{Thanks for noticing the inconsistency. }
 
@@
 In this section we discuss another application of adaptive sequential inference, to binary decision trees.
-Specifically, the classification and regression tree  (CART)  method builds a decision tree in a top down manner, finding the best available split at each node.
+Specifically, the classification and regression tree  (CART)  method builds a decision tree in a top down manner, finding the best available split at each node \citep{breiman1984classification}.
 When considering splitting two daughter nodes, the generic CART algorithm splits the nodes in an arbitrary order, only stopping when a node reaches a minimum size.
@@
 
-\subsection{Nonparametric Changepoint Detection}
+\subsection{Nonparametric Changepoint Detection}\label{sec:nonpar}
 
@@
 
-Open-source code is available at this article's github repository at:
-\begin{center}
-\texttt{www.github.com/wfithian/sequential-selection}
-\end{center}
+Open-source code is available at this article's github repository at \texttt{github.com/wfithian/sequential-selection}.
+R and Python code are available to implement proposals in this paper are available at \texttt{github.com/selective-inference}.
 
@@
 
-The authors are grateful for stimulating and informative conversations with Stefan Wager, Lucas Janson, Trevor Hastie, Yoav Benjamini, and Larry Brown. \WFcomment{Funding}.
+The authors are grateful for stimulating and informative conversations with Stefan Wager, Lucas Janson, Trevor Hastie, Yoav Benjamini, and Larry Brown.
+William Fithian was supported in part by the Gerald J.
+Lieberman Fellowship.
+Robert Tibshirani was supported by NSF grant DMS-9971405 and NIH grant N01-HV-28183.
+Ryan Tibshirani was supported by NSF grant DMS-1309174.
+Jonathan Taylor was supported by NSF grant DMS-1208857.
 

### 2015-12-04 paper.tex
minor edits, resyncing for new computer

@@
 The selection event $A_1=\{|Y_1| > |Y_2|\}$ is shown in yellow in Figure~\ref{fig:bv_condSets}.
-If $j_1=1$ then
+If $j_1=1$ the first two models are
 \[
@@
 
-Figure~\ref{fig:bv} illustrates a general phenomenon in saturated-model tests: when there are near-ties between strong variables that are competing to enter the model, the resulting $p$-value my be very weak.
+Figure~\ref{fig:bv} illustrates a general phenomenon in saturated-model tests: when there are near-ties between strong variables that are competing to enter the model, the resulting $p$-value may be very weak.
 Figure~\ref{fig:bv_rocCurve} displays the cumulative distribution function for the first $p$-value when $\mu=(4,4)$, a very strong signal.
@@
 
-Until this point, we have deferred discussing the question of how to actually compute the rejection cutoffs for the max-$t$ test, next-entry test, or any of the other tests we have discussed.
+Until this point, we have deferred discussing the question of how to actually compute the $p$-values for the max-$t$ test, next-entry test, or any of the other tests we have discussed.
 In each case, the $p$-value at step $k$ comes from a test that rejects when some test statistic $T_k$ is large compared to the law

### 2015-12-05 paper.tex
lots of small changes

@@
 
-\JTcomment{Maybe we should write $k_F$ instead of $k_0$? Or $k_0(Y,F)$?}\\
-\WFcomment{$\hat k_F$ is already the ForwardStop estimator. But $k_0(Y,F)$ is good.}\\
-
 Define the {\em completion index} by $k_0(Y, F) = \min\{k:\; F \in M_k(Y)\}$, the index of the first correct model.
@@
 \end{equation}
-That is, at each step we select the next variable to maximize the likelihood of the next fitted model.
+That is, at each step we select the next variable to maximize the likelihood of the next fitted model (in the case of ties, we could either choose randomly or select both variables).
 
@@
 
-For simplicity, we have made two assumptions: that only one variable is added at each step, and that the set of candidate variables we choose from is the same in each step.
+For simplicity, we have implicitly made two assumptions: that only one variable is added at each step, and that the set of candidate variables we choose from is the same in each step.
 It is relatively straightforward to relax either assumption, but we do not pursue such generalizations here.
@@
 \paragraph{Remark.} In its usual formulation, the lasso coefficients for linear regression minimize a penalized RSS criterion.
-If we replace $-\ell$ with any strictly decreasing function of the log-likelihood, all of the results in this article hold without modification.
+If we replace $-\ell$ with any strictly decreasing function of the log-likelihood such as RSS, all of the results in this article hold without modification.
 
@@
 \begin{equation}\label{eq:indepCond_random_k0}
-  \P(p_{k+1} \leq \alpha_{k+1}, \ldots, p_d \leq \alpha_d
-  \mid p_1, \ldots, p_k, k_0 = k ) \eqAS \left(\prod_{i=k+1}^d \alpha_i\right) .
+  \P_F(p_{k+1} \leq \alpha_{k+1}, \ldots, p_d \leq \alpha_d
+  \mid p_1, \ldots, p_k) \eqAS \left(\prod_{i=k+1}^d \alpha_i\right) \quad \text{on } \{k_0(Y, F) = k\}.
 \end{equation}
@@
   \end{align*}
-Lastly, if the $p_k$ are exact then the inequalities above become equalities, implying uniformity and joint independence.
+Lastly, if the $p_k$ are exact then the inequalities above become equalities, implying uniformity and mutual independence.
 \end{proof}
@@
 
-We begin by proving both paths satisfy the SSP
+We begin by proving both paths satisfy the SSP:
 \begin{proposition}\label{prop:forwardSSP}
@@
 Saturated-model tests perform inference on selected linear functionals of $\mu$.
-In the context of sequential linear model selection,  we can use the saturated-model framework to test the incremental null
-\[
-H^{\text{inc}}:\; \theta_{j_k}^{E_k} = 0,
-\]
-where $\theta^E$ minimizes \smash{$\|\mu - X_E\theta\|_2$}.
+In the context of sequential linear model selection,  we can use the saturated-model framework to test the incremental null \smash{$H^{\text{inc}}:\; \theta_{j_k}^{E_k} = 0$}, where $\theta^E$ minimizes \smash{$\|\mu - X_E\theta\|_2$}.
 
 Let $\eta_k\in \R^n$ denote the vector for which \smash{$\theta_{j_k}^{E_k} = \eta_k'\mu$}, and let \smash{$\proj_{\eta_k}^\perp$} denote the projection operator into the space orthogonal to $\eta_k$.
-The UMPU saturated-model test for $\eta_k'\mu$ is based on the distribution
-\begin{equation*}
-\L\left( \eta_k'Y \mid E_{k-1}, \;j_k, \;\proj_{\eta_k}^\perp Y \right).
-\end{equation*}
+The UMPU saturated-model test for $\eta_k'\mu$ is based on the distribution \smash{$\L\left( \eta_k'Y \mid E_{k-1}, \;j_k, \;\proj_{\eta_k}^\perp Y \right)$}, or equivalently
+\begin{equation}\label{eq:satModel}
+\L\left( X_k'Y \mid E_{k-1}, \;j_k, \;\proj_{\eta_k}^\perp Y \right)
+\end{equation}
 
 While the test statistic $X_{j_k}'Y$ is measurable with respect to $\sF_k$, the null distribution we compare it against is not: it depends on \smash{$\proj_{\eta_k}^\perp Y$}, which is not measurable with respect to $\sF_k$.
-As a result, saturated-model $p$-values constructed as in~\eqref{eq:satModel} are in general {\em not} independent on nulls.
+As a result, saturated-model $p$-values are in general {\em not} independent on nulls.
 
@@
 If $\sigma^2$ were unknown we would also need to condition on $\|Y\|^2$.
-We can construct $p$-values from the UMPU, equal-tailed, or one-sided test based on~\eqref{eq:selModel_cond}; the only difference is in how much of the overall level $\alpha$ is apportioned to the right or left tail \citep{fithian2014optimal}.
+We can construct $p$-values from the UMPU, equal-tailed, or one-sided test based on the one-parameter exponential family~\eqref{eq:selModel_cond}; the only difference is in how much of the overall level $\alpha$ is apportioned to the right or left tail \citep{fithian2014optimal}.
 If we use the construction in~\eqref{eq:selModel_cond} after a subpath-sufficient path algorithm, the resulting $p$-values are always guaranteed to have independence on nulls.
-This is because the test statistic $X_{j_k}'Y$ is measurable with respect to $\sF_k$, and so is the null distribution we compare it against, which depends on $E_{k-1}, j_k,$ and \smash{$U_{k-1} = X_{E_{k-1}}'Y$}.
+This is because the test statistic $X_{j_k}'Y$ is measurable with respect to $\sF_k$, as is the reference null distribution.
 
@@
 In~\eqref{eq:selModel_cond}, we only need to condition on an $|E_{k-1}|$-dimensional projection of $Y$, whereas in~\eqref{eq:satModel} we condition on an $(n-1)$-dimensional projection of $Y$.
-Conditioning on more information can sap the power of a test; as we will see in Sections~\ref{sec:bivariate} and~\ref{sec:sparseReg}, the saturated-model tests tend to pay a heavy price for this extra conditioning.
+Conditioning on more information can sap the power of a test; as we will see in Sections~\ref{sec:bivariate} and~\ref{sec:sparseReg}, the saturated-model tests often pay a heavy price for this extra conditioning.
 
@@
 
-%In the case of known $\sigma^2$, the greedy likelihood ratio statistic is the max-$z$ statistic: the maximum of $z_{j,E_{k-1}}$, the largest multivariate $z$-statistic of any held-out variable, controlling for the variables in the current model.
-
-For the most part we do not pursue tests like those in Section~\ref{sec:identify}, which condition on the next model $M_k$.
-It is not necessary to condition on $M_k$ to obtain the FDR and FWER guarantees that we want to establish, and so we follow the rule of thumb that we should condition on as little information as possible.
-In Section~\ref{sec:sparseReg}, we empirically compare the max-$t$ test with the test that uses the same test statistic but also conditions on the identity of the next variable $j_k$.
-We call the resulting test the {\em max-$t$-identify test}.
-Its performance in our simulation is similar to that of the max-$t$ test.
+We will not pursue tests like those in Section~\ref{sec:identify}.
+It is not necessary to condition on $M_k$ to obtain the FDR and FWER guarantees that we want to establish, and so we follow the general principle of conditioning on as little information as possible.
+In Section~\ref{sec:sparseReg}, we empirically compare the max-$z$ test with the test that uses the same test statistic but also conditions on the identity of the next variable $j_k$.
+In our simulation this test, which we call the {\em max-$z$-identify test}, performs similarly to the max-$z$ test.
 
@@
 Because the saturated-model test conditions on $n-1$ dimensions, the resulting distribution in~\eqref{eq:satModel} is nothing more than a truncated univariate Gaussian random variable, and inference can be carried out analytically.
-By contrast, selected-model tests typically require sampling from a truncated multivariate Gaussian distribution of dimension $p-|E_{k-1}|$.
-
-For some applications, a major conceptual advantage of the saturated-model approach is that, when $M(E)$ is misspecified, $\theta^{E}$ is still a well-defined quantity on which we can perform exact inference.
+By contrast, selected-model tests typically require MCMC sampling from a truncated multivariate Gaussian distribution of dimension $p-|E_{k-1}|$.
+
+For some applications, a major conceptual advantage of the saturated-model approach is that, even when $M(E)$ is misspecified, $\theta^{E}$ is still a well-defined quantity on which we can perform exact inference.
 For example, a 95\% confidence interval for $\theta_{j_k}^{E_k}$ could cover that coefficient with probability 95\% even if $M(E_k)$ is misspecified.
-In the setting of sequential model selection, this advantage is less important: we are not constructing intervals for $\theta_{j_k}^{E_k}$; rather, the goal of the test at step $k$ is to reject $M_{k-1}$ and move on if it is demonstrably incompatible with the data.
+In the setting of sequential model selection, this advantage is less important: at an intermediate step $k$, we are not constructing intervals for $\theta_{j_k}^{E_k}$ but rather deciding whether to reject $M_{k-1}$ and move on.
+However, the saturated-model framework could be quite helpful after selection is finished, if we want to construct intervals for the least-squares parameters for the selected active set $E_{\hk}$.
 
@@
 First, selected-model inference allows us to drop the assumption that $\sigma^2$ is known.
-This is not an option for the saturated model because conditioning on both $\|Y\|^2$ and \smash{$\proj_{\eta_k}^\perp Y$} results in a degenerate conditional distribution for $\eta_k'Y$.
+This is not possible for the saturated model because conditioning on both $\|Y\|^2$ and \smash{$\proj_{\eta_k}^\perp Y$} results in a degenerate conditional distribution for $\eta_k'Y$.
 Second, we have seen in Section~\ref{sec:pValsIndep} that tests of the form~\eqref{eq:selModel_cond} or~\eqref{eq:selModel_marg} yield independent $p$-values under general conditions, allowing us to apply  the sequential stopping rules of~\citet{gsell2013sequential} and \citet{li2015accumulation}.
-Finally, and perhaps most importantly, selected-model tests can be dramatically more powerful than saturated-model tests, as we illustrate now with an extended example.
+Finally, and most importantly, selected-model tests can be dramatically more powerful than saturated-model tests, as we illustrate in the next example.
 
@@
 The selection event $A_1=\{|Y_1| > |Y_2|\}$ is shown in yellow in Figure~\ref{fig:bv_condSets}.
-If $j_1=1$ the first two models are
+On $A_1$, the first two models are
 \[
@@
 \]
-The selected-model test at step 1 compares $Y_1$ to its distribution under $M_0$ conditional on $A_1$, a test of $H_0:\;\mu_1=0$ in model $M_1$.\footnote{In this very simple example, the max-$z$, max-$z$-identify, and next-entry tests all coincide.} By contrast, the saturated-model test is a test of $H_0:\; \mu_1=0$ in the model $M_{\text{sat}}:\; Y \sim \cN(\mu, I_2)$.
+The selected-model test at step 1 compares $Y_1$ to its distribution under $M_0$, the global null, conditional on $A_1$.\footnote{In this very simple example, the max-$z$, max-$z$-identify, and next-entry tests are all equivalent.} By contrast, the saturated-model test is a test of $H_0:\; \mu_1=0$ in the model $M_{\text{sat}}:\; Y \sim \cN(\mu, I_2)$.
 Thus, it must condition on $Y_2$ to eliminate the nuisance parameter $\mu_2$, comparing $Y_1$ to its null distribution given $A_1$ and the observed value of $Y_2$.
@@
 The $p$-values for the selected and saturated models are 0.007 and 0.3, respectively.
-These two plots are reproduced from \citet{fithian2014optimal}, in which the same example was presented in less detail.
+Figure~\ref{fig:bv} is reproduced from \citet{fithian2014optimal}, where the same example was presented in less detail.
 \end{example}
@@
 Note that if $\mu=(0,4)$, the incremental null is true, but the complete null is false.
-That is, the model $M_0 = M(\emptyset)$ is missing an important signal variable, but the missing variable is not $X_1 = (1,0)$.
+That is, the global null model $M_0$ is missing an important signal variable, but the missing variable is not $X_1 = (1,0)$.
 Because the saturated-model test is a valid test of the incremental null, its power is $\alpha=0.05$.
-By contrast, the selected-model test rejects about half of the time when $\mu=(0,4)$, detecting the fact that $M_0$ is a poor fit to the data.
+By contrast, the selected-model test rejects about half of the time when $\mu=(0,4)$, successfully detecting the poor fit of $M_0$ to the data.
 
@@
 Because $k_0=0$, both $p$-values are uniform by construction, but the saturated-model $p$-values are strongly dependent, with correlation $-0.48$.
-By contrast, the selected-model $p$-values $(p_1,p_2)$ are independent under the global null, a consequence of Theorem~\ref{thm:suffCond}.
+By contrast, the selected-model $p$-values $(p_1,p_2)$ are independent under the global null.
 
@@
 
-Until this point, we have deferred discussing the question of how to actually compute the $p$-values for the max-$t$ test, next-entry test, or any of the other tests we have discussed.
+Until this point, we have deferred discussing how to compute the reference distribution for the max-$t$ test, next-entry test, or any of the other tests we have discussed.
 In each case, the $p$-value at step $k$ comes from a test that rejects when some test statistic $T_k$ is large compared to the law
@@
 
-In many cases it is easy to sample $Y$ given $U_{k-1}$, in which case computational difficulties only arise insofar as the selection event $\{M_{k-1}(Y)=M\}$ is difficult to condition on.
+In many cases it is easy to sample $Y$ given $U_{k-1}$, in which case computational difficulties only arise insofar as the selection event $\{M_{k-1}(Y)=M\}$ is difficult to condition on, or the test statistic $T_k(Y)$ is difficult to evaluate.
 Due to the generality in which we have posed the problem, we cannot give a general prescription for how to carry out this sampling.
 However, we give some details for the important case of linear regression in Appendix~\ref{sec:linRegComputation}.
+
+\subsection{Constraints}\label{sec:constraints}
 
@@
 Remarkably, however, in both forward-stepwise and lasso linear regression, the computational problem actually becomes easier when we move to the sequential setting considered here.
-Indeed, Section~\ref{sec:constraints} shows that for both forward-stepwise regression and lasso linear regression --- and their respective generalizations to exponential family models --- conditioning on the event $\{E_{k-1}=E\}$, for any $k$, only introduces $2p$ linear constraints, taking the form of a rectangular constraint on the sufficient statistics for the inactive variables.
-
-\subsection{Constraints}\label{sec:constraints}
-We now show that the conditioning polytope can be expressed with only $2(p-k)$ constraints in forward stepwise regression and the lasso.
+We now show that for both forward-stepwise regression and lasso linear regression --- and their respective generalizations to exponential family models --- conditioning on the event $\{E_{k-1}=E, U_k=u\}$, for any $k, E,$ and $u$, only introduces $2(p-k)$ linear inequality constraints, amounting to a rectangular constraint for the sufficient statistics of excluded variables.
 This increases the speed of both accept/reject and hit-and-run sampling for our problem.
-
 We begin by proving the result for forward stepwise linear regression, followed by a more general result for the forward stepwise likelihood path in exponential families.
+To avoid trivialities, we assume throughout that there are almost surely no ties.
+This is true, for example, in linear regression if $X$ is in general position.
 
@@
   \[
-  v_j^-(E,u) < X_j'Y < v_j^+(E,u), \quad\forall j \notin E,
+  v_j^-(E,u) \leq X_j'Y \leq v_j^+(E,u), \quad\forall j \notin E,
   \]
@@
 \end{theorem}
-
-\RJTcomment{Can't we get rid of the strict inequalities above?
-The proof writes things in terms of nonstrict inequalities, and I think it makes sense to drop them; in the case of ties, $E$ is still a valid selected model.}
-\WFcomment{Thanks for noticing the inconsistency. }
 
@@
   \]
-  with $\Theta \sub \R^p$ convex, and assume that the model path $M_{0:d}$ is given by the forward stepwise likelihood path with
+  with $\Theta \sub \R^p$ convex, and assume that $M_{0:d}$ is the forward stepwise likelihood path with
 \begin{equation}
@@
   \]
-  where $\psi_E:\; \R^{|E|} \to \R$ is defined such that $\psi(\theta)=\psi_E(\theta_E)$ if $\theta_{\{1,\ldots p\}\setminus E}=0$.
-Note that $\ell_E^*$ is the convex dual of the convex function $\psi_E$, so $\ell_E^*$ is also convex.
-\JTcomment{It would be convex even if $\psi_E$ were not convex.
-No?
-I also think $U$ in the definition of $\ell^*_E$ can be just $U(Y)$.} \RJTcomment{I would call this the ``convex conjugate'' or just the ``conjugate'' of $\ell$; and yes this will always be convex.
-It doesn't require knowing anything about $\psi_E$ or even $\Theta$.} At stage $i<k$, the next variable to enter is the one that maximizes \smash{$\ell_{E_i\cup \{j\}}^*(U(Y))$}.
-
-  Because the forward stepwise likelihood path satisfies the SSP, we know the entire path of fitted models up to step $k$ once we condition on $E_k$ and its sufficient statistics \smash{$U_{E_k}(Y)$}.
-On the set $\{E_k=E\}$, all of the quantities $j_{i+1}$, $E_i$, \smash{$\hat\theta^i$}, and \smash{$\proj_{E_i}^\perp$} depend only on $U_E(Y)$.
-For brevity, write
-  \[
-  C_i^* = \max_{j \notin E_i} \;
-    \sup \left\{ \theta ' U - \psi(\theta):\; \theta \in \Theta(E_i\cup \{j\}) \right\}.
-  \]
-  On $A$, $C_i^*$ is attained at $j=j_{i+1}$, the $(i+1)$st variable added.  If $U_E(Y)$ is fixed at $u$, and $j \notin E$, then the condition for 
-  variable $j$ {\em not} to enter at step $i+1 < k$ is
-  \[
-  \ell_{E_i\cup \{j\}}^*(U(Y)) \leq C_i^*.
-  \]
-  Note that \smash{$\ell_{E_i\cup \{j\}}^*$} depends only on the sufficient statistic \smash{$U_{E_i}(Y)$} --- which is fixed, on $A$ --- and the remaining sufficient statistic $U_j(Y)$.
-Since \smash{$\ell_{E_i\cup\{j\}}^*$} is convex in $U$, if we fix \smash{$U_{E_i}$}, and consider it as a function of $U_j$ alone, then it is still convex in this single argument.
-Hence any sublevel set of \smash{$\ell_{E_i\cup\{j\}}^*$}, with \smash{$U_{E_i}$} fixed, is just a 1-dimensional convex set, i.e., an interval, so we may write
+  where $\psi_E:\; \R^{|E|} \to \R$ is defined such that $\psi(\theta)=\psi_E(\theta_E)$ for $\theta\in \Theta(E)$.
+Note that $\ell_E^*$ is the convex conjugate of the function $\psi_E$.
+For step $i<k$, write
+  \[
+  C_i^* = \max_{j \notin E_i} \;\ell_{E_i\cup \{j\}}^*(U(Y)).
+  \]
+  On $A$, $C_i^*$ is attained at $j=j_{i+1}\in E$, the next variable to enter.
+For $j\notin E$, the condition for $j$ {\em not} to enter at step $i+1$ is \smash{$\ell_{E_i\cup \{j\}}^*(U(Y)) \leq C_i^*$}.
+Due to the SSP, on the set $\{E_k=E\}$ all of the quantities $j_{i+1}$, $E_i$, and $C_i^*$ are functions of $U_E(Y)$.
+
+  Because $\ell_{E_i\cup \{j\}^*}(U)$ depends only on $U_{E_i}$ and $U_j$, the former of which is fixed at $u_{E_i}$, we can consider it as a one-dimensional convex function of $U_j$; hence any sublevel set is an interval with possibly infinite endpoints.
+Thus
   \begin{equation}\label{eq:noEnterBounds_FSL}
@@
   \end{equation}
-  where $v_{j,i}^{-}$ and $v_{j,i}^+$ are the two values of $U_j$ such that \smash{$\ell_{E_i\cup \{j\}}^*(U)=C_i^*$}.
+  where $v_{j,i}^{-}$ and $v_{j,i}^+$ are the two values of $U_j$ such that \smash{$\ell_{E_i\cup \{j\}}^*(U)=C_i^*$}, or else $\pm\infty$.
 
@@
 
-\JTcomment{Is there a reason you have $\psi$ and $\ell$ above instead of just $\ell$? I think the $(E,t)$ above should be $(E,u)$. It seems you never come back to $\psi$.}
-\WFcomment{Fixed t/u. $\psi$ is the log-proportionality constant, I think we need it. Maybe I'm misunderstanding your question?}
-\JTcomment{I guess I meant that I am assuming that
-$$
-\ell(\theta) = \theta'U(y) - \psi(\theta)
-$$
-but I don't see how I would conclude that from what is written.
-In principle, as written the $\psi$ above might be some other function but
-(44) and (45) imply that the objective function is of the form
-$$
-\ell(\theta) = \tilde{\psi}(\theta) - \theta'U(y)
-$$
-for some $\tilde{\psi}$.
-In the proof when you say $\mathbb{E}_{\hat{\theta}^{\lambda}}[U_j]$ I am assuming this is $e_j^T\nabla \psi (\hat{\theta}^{\lambda})$ soshould I conclude that $\tilde{\psi}=\psi$?
-}
-
 \begin{proof}
-  Note that, because exponential family likelihoods are concave in their natural parameters, the problem in~\eqref{eq:regProblem} is convex.
+  Note that, because exponential family log-likelihoods are concave in their natural parameters, the problem in~\eqref{eq:regProblem} is convex.
 Again, once we know the sufficient statistics $U_E(Y)$ for model $k$, and that $E_k=E$, we know that
@@
 
-The columns of $X$ are normalized to have length 1, and we simulate $Y \sim \cN(X\beta,I_n)$, using a seven-sparse model with signal-to-noise ratio 5:
+The columns of $X$ are normalized to have length $\|X_j\|=1$, and we simulate $Y \sim \cN(X\beta,I_n)$, using a seven-sparse model with signal-to-noise ratio 5:
 \[
@@
 
-For each step we compute one-step selected-model and saturated-model $p$-values, as well as nominal (unadjusted) $p$-values, conditioning on the signs of the active variables to make the problem more computationally tractable.
-Figure~\ref{fig:simulation_null_false} shows the power of all three tests for each of the first ten steps \RJTcomment{I only see steps 6 through 9?}, conditional on the event that the null hypothesis is false.
-It is clear from Figure~\ref{fig:simulation_null_false} that the selected-model $p$-values are far more powerful than the saturated-model $p$-values.
+For each step we compute saturated-model $p$-values, max-$t$ and max-$z$ $p$-values, and nominal (unadjusted) $p$-values, conditioning on the signs of the active variables to make the problem more computationally tractable.
+\WFcomment{Do we really need to condition on the signs?
+There's just two polytopes in each case, right?} Figure~\ref{fig:simulation_null_false} shows the power of all three tests for each of the first eight steps, conditional on the event that the null hypothesis is false.
+It is clear from Figure~\ref{fig:simulation_null_false} that the selected-model $p$-values are far more powerful than the saturated-model $p$-values, especially for the first five steps, where they have near-perfect power.
 The nominal $p$-values are also quite powerful, but they do not have the correct level.
@@
 
-Figure~\ref{fig:simulation_null_true} shows the distribution of $p_k$ for $k = 8, \ldots, 17$ \RJTcomment{I only see steps 9 through 11?}, given that the null hypothesis tested at step $k$ is correct (i.e., that $k_0< k$).
-Because the correct model is seven-sparse, $k=8$ is the first index for which the null can possibly be true.
+Figure~\ref{fig:simulation_null_true} shows the distribution of $p_k$ for steps eight to eleven, given that the null hypothesis tested at step $k$ is correct (i.e., that $k_0\leq k$).
+Because the correct model is seven-sparse, $k=8$ is the first index for which the null model can possibly be true.
 All but the nominal $p$-values are uniform by construction, and the nominal $p$-values are highly anti-conservative, as expected.
 
-Finally, as a warning against misinterpretation of our method, we include Figure~\ref{fig:simulation_noise_var} showing the first ten $p$-values for each method \RJTcomment{I only see steps 1 through 4?}, conditional on event that the variable added at the given step is a noise variable in the full model.
+Finally, as a warning against misinterpretation of our method, we include Figure~\ref{fig:simulation_noise_var} showing the first four $p$-values for each method, conditional on event that the variable added at the given step is a noise variable in the full model.
 Now, none of the $p$-values are uniform.
@@
 If we try to add a noise variable to the model before we have included all the signal variables, then we are testing a false null hypothesis.
-The test rejects because there is much more signal to find, and as such, it is entirely appropriate for us to reject the null and continue growing the model.
+The test rejects because there is much more signal to find, and as such, the null model is false.
 
@@
 
-If we combine the saturated-model or selected-model $p$-values with one of our three stopping rules, we can evaluate the model-selection performance of each method in terms of:
-\begin{itemize}
-\item its probability of selecting a correct model or $p_{\text{screen}}$,
-\item its model-wise FWER,
-\item its model-wise FDR, and
-\item its variable-wise FDR, where we use $V^{\text{full}}= \#\{\text{ noise variables included in } M_{\hk}\}$ instead of \smash{$V=(\hk-k_0)_+$}.
-\end{itemize} 
-The last measure of performance is not explicitly controlled by any of the selective-inference methods, but we might nevertheless hope to perform reasonably.
+If we combine the saturated-model or selected-model $p$-values with one of our three stopping rules, we can evaluate the model-selection performance of each method.
+Table~\ref{tab:stopping05} reports several model-selection metrics for each combination of test and stopping rule carried out at level $\alpha=0.05$, while Table~\ref{tab:stopping20} reports the same results for the more liberal $\alpha=0.2$.
+The metrics are $p_{\text{screen}}$, the probability of selecting a correct model; the (model-wise) FWER, both marginally and conditional on screening, the model-wise FDR and variable-wise $\text{FDR}^{\text{full}}$, and the average number of signal variables selected.
+Note that $\text{FDR}^{\text{full}}$ is not explicitly controlled by any of the methods other than knockoff+, but we might nevertheless hope to perform reasonably well.
+Also, note that the model-wise FWER and FDR and not defined for knockoffs, since it does not select variables sequentially, and never formally tests any model.
 
@@
 \input{tables/error_rates_05.tex}
-\caption[tab:stopping]{\em Results of various stopping rules applied to simulated data with 7 strong signals,  as described at the beginning of Section \ref{sec:sparseReg}.
-The {\tt Simple} rule stops at the first time that a $p$-value exceeds $\alpha=0.05$, while {\tt Forward} and {\tt Strong} refers to the ForwardStop and StrongStrop rules of Section \ref{sec:orderedProposals}.
-Results with theoretical guarantees are in \guarantee{bold} type.}
+\caption[tab:stopping]{Model-selection performance of various stopping rules applied to simulated data with 7 strong signals,  as described at the beginning of Section~\ref{sec:sparseReg}.
+Results that are theoretically guaranteed to be less than $\alpha$ are in \guarantee{bold} type.}
 \label{tab:stopping05}
@@
 \input{tables/error_rates_20.tex}
-\caption[tab:stopping]{\em Results of various stopping rules applied to simulated data with 7 strong signals,  as described at the beginning of Section \ref{sec:sparseReg}.
-The {\tt Simple} rule stops at the first time that a $p$-value exceeds $\alpha=0.2$, while {\tt Forward} and {\tt Strong} refers to the ForwardStop and StrongStrop rules of Section \ref{sec:orderedProposals}.
-Results with theoretical guarantees are in \guarantee{bold} type.}
+\caption[tab:stopping]{Model-selection performance of various stopping rules applied to simulated data with 7 strong signals,  as described at the beginning of Section~\ref{sec:sparseReg}.
+Results that are theoretically guaranteed to be less than $\alpha$ are in \guarantee{bold} type.}
 \label{tab:stopping20}
@@
 
-\RJTcomment{I think these captions / labels in the tables should be fixed up, but I didn't want to mess with anything.
-Few pedantic things: ``simple'' should be ``basic'', because now we're calling it BasicStop, and why not just write BasicStop and ForwardStop instead of ``basic'' and ``forward''?
-Also, ``MaxZ'' should be ``Max-$z$'' and same for ``Max-$t$''.
-Lastly, what is with the typewriter type?}
-
 We see that BasicStop and ForwardStop control model-wise FWER and FDR, respectively, using $p$-values from max-$t$ or max-$t$-identify, as predicted by the theory, and the nominal $p$-values do not lead to control of FWER or FDR, as expected due to their anti-conservatism.
-Although not guaranteed by the theory, the $p$-values from the saturated model do control FWER and model-wise FDR in this example.
+Although not guaranteed by the theory, the $p$-values from the saturated model do control model-wise FDR in this example.
 The max-$t$ rules show higher probabilities of selecting the true model, especially using ForwardStop.
 Lastly, the knockoff method does not yield variable-wise FDR control.
-\RJTcomment{Ummm isn't this an unexpected failure of knockoffs?
-Should we comment on this?
-Or am I missing something?
-Oh, I see.
-We've also included knockoff+, which does control the FDR (because it adds something to the denominator, right?).
-We should say something about this.}
+While the more conservative knockoff+ method does achieve variable-wise FDR control, this comes at a cost of very low power.
+Because of the discrete nature of knockoff and knockoff+, these two methods might perform better in an example with a larger number of signal variables to find.
 

### 2015-12-05 code/python/error_rates.py

@@ -13,7 +13,9 @@ for results, outbase in zip(['../snr_5_alpha_%s.csv' % alpha for alpha in ['05',
     guarantees = []
 
     name_map = dict(zip(['nominalT', 'maxT', 'maxT_identify', 'maxT_unknown', 'saturated'],
-                        ['Nominal', 'MaxZ', 'MaxZ identify', 'MaxT', 'Saturated']))
+                        ['Nominal', 'Max-z', 'Max-z-identify', 'Max-t', 'Saturated']))
+    rule_map = dict(zip(['simple', 'forward'],
+                        ['BasicStop', 'ForwardStop']))
     for test, rule in product(['nominalT', 'maxT', 'maxT_identify', 'maxT_unknown', 'saturated'],
                               ['simple', 'forward']):
         name = '_'.join([test, rule])
@@ -35,7 +37,7 @@ for results, outbase in zip(['../snr_5_alpha_%s.csv' % alpha for alpha in ['05',
         else:
             guarantees.append((False, False, False, False, False, False))
         results.append((screen, FWER_model, FWER_model_cond, FDR_model, FDR_var, S_var))
-        names.append('%s %s' % (name_map[test], rule))
+        names.append('%s %s' % (name_map[test], rule_map[rule]))
 
     # now knockoffs
 

### 2015-12-06 paper.tex
more editing, especially of simulation section; tweaked error rate tables code

@@
 \input{tables/diabetes_edit.tex}
-\caption[tab:diab]{\em Forward stepwise regression for the diabetes data: naive $p$-values, $p$-values from the saturated model, and our max-$t$ $p$-values.
+\caption[tab:diab]{Forward stepwise regression for the diabetes data: naive $p$-values, $p$-values from the saturated model, and our max-$t$ $p$-values.
 The bold annotation in each column marks the model selected by ForwardStop with $\alpha=0.1$.
@@
 
-\subsection{Constraints}\label{sec:constraints}
+\subsection{Constraints for the Greedy Likelihood and Next-Entry Tests}\label{sec:constraints}
 
@@
 
-A similar result also holds for the class of $\ell_1$-regularized exponential family models. % including the graphical LASSO.
+A similar result also holds for the class of $\ell_1$-regularized exponential family models.
 
@@
 
-For each step we compute saturated-model $p$-values, max-$t$ and max-$z$ $p$-values, and nominal (unadjusted) $p$-values, conditioning on the signs of the active variables to make the problem more computationally tractable.
-\WFcomment{Do we really need to condition on the signs?
-There's just two polytopes in each case, right?} Figure~\ref{fig:simulation_null_false} shows the power of all three tests for each of the first eight steps, conditional on the event that the null hypothesis is false.
+For each step we compute saturated-model $p$-values, max-$z$ $p$-values, and nominal (unadjusted) $p$-values, conditioning on the signs of the active variables as proposed by \citet{lee2013exact}.
+Figure~\ref{fig:simulation_null_false} shows the power of all three tests for each of the first eight steps, conditional on the event that the null hypothesis is false.
 It is clear from Figure~\ref{fig:simulation_null_false} that the selected-model $p$-values are far more powerful than the saturated-model $p$-values, especially for the first five steps, where they have near-perfect power.
@@
   \includegraphics[width=1\textwidth]{figs/simulation_snr_5_alpha_05_null_false.pdf}
-  \caption{\em  CDFs of nominal (black), max-$z$ (red), max-$t$ (green) and  saturated-model (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on testing a false null hypothesis at step $k$.
+  \caption{  CDFs of nominal (black), max-$z$ (red), max-$t$ (green) and  saturated-model (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on testing a false null hypothesis at step $k$.
 The max-$t$ approaches are much more powerful than the saturated-model test.
@@
   \includegraphics[width=1\textwidth]{figs/simulation_snr_5_alpha_05_null_true.pdf}
-  \caption{\em  CDFs of nominal (black), max-$z$ (red), max-$t$ (green) and  saturated-model (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on testing a true null hypothesis at step $k$.
+  \caption{  CDFs of nominal (black), max-$z$ (red), max-$t$ (green) and  saturated-model (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on testing a true null hypothesis at step $k$.
 The nominal test is badly anti-conservative, while all of the other methods  show uniform $p$-values as desired.}
@@
 \end{figure}
-
- \RJTcomment{The lines in Fig 6 are super thin, thinner than Fig 5?}
 
@@
   \includegraphics[width=1\textwidth]{figs/simulation_snr_5_alpha_05_noise_var.pdf}
-  \caption{\em CDFs of nominal (black), max-$z$ (red), max-$t$ $\sigma$ (green) and  saturated-model (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on the event that the variable added at step $k$ is a noise variable in the full model.
+  \caption{ CDFs of nominal (black), max-$z$ (red), max-$t$ $\sigma$ (green) and  saturated-model (blue) $p$-values in the simulation of Section~\ref{sec:sparseReg}, conditional on the event that the variable added at step $k$ is a noise variable in the full model.
 Here, none of the methods produce uniform $p$-values.
@@
 Table~\ref{tab:stopping05} reports several model-selection metrics for each combination of test and stopping rule carried out at level $\alpha=0.05$, while Table~\ref{tab:stopping20} reports the same results for the more liberal $\alpha=0.2$.
-The metrics are $p_{\text{screen}}$, the probability of selecting a correct model; the (model-wise) FWER, both marginally and conditional on screening, the model-wise FDR and variable-wise $\text{FDR}^{\text{full}}$, and the average number of signal variables selected.
+The max-$t$ test, which we could use in place of the max-$z$ test if $\sigma^2$ were unknown, is included for comparison.
+The max-$z$-identify test was also included in the simulation, but not shown in Tables~\ref{tab:stopping05} and \ref{tab:stopping20} because its performance on all metrics was virtually identical to the max-$z$ test.
+
+The metrics are $\P(\hk \geq k_0)$, the probability of selecting a correct model; the (model-wise) FWER; the FWER conditional on $\hk \geq k_0$, labeled cFWER; the model-wise FDR and variable-wise $\text{FDR}^{\text{full}}$; and the average number of full-model signal variables selected, labeled $\E[S^{\text{full}}]$ (where $S^{\text{full}} = R-V^{\text{full}}$).
 Note that $\text{FDR}^{\text{full}}$ is not explicitly controlled by any of the methods other than knockoff+, but we might nevertheless hope to perform reasonably well.
 Also, note that the model-wise FWER and FDR and not defined for knockoffs, since it does not select variables sequentially, and never formally tests any model.
-
-% latex table generated in R 3.0.2 by xtable 1.7-1 package
-% Fri May  1 09:29:28 2015
-%\begin{table}[ht]
- % \centering
- % \begin{tabular}{llcccc}
-  %  \hline
-  %  Method & Stopping Rule & $p_{\text{screen}}$ & $\text{FWER}_{\text{mod}}$ an
- %   & $\text{FDR}_{\text{mod}}$ 
-  %  & $\text{FDR}_{\text{var}}$ \\ 
- %   \hline
-    %Selected & Basic & .290 & .000 & .002 & .039 \\ 
-   % Selected & Forward & .559 & .027 & .020 & .066 \\ 
-   % Selected & Strong & .041 & .014 & .008 & .042 \\ 
-  %  \hline
-   % Saturated & Basic & .000 & .000 & .000 & .028 \\ 
-  %  Saturated & Forward & .014 & .000 & .000 & .030 \\ 
-  %  Saturated & Strong & .000 & .000 & .000 & .032 \\ 
-  %  \hline
-  %  Knockoffs & & .000 & --- & --- & .231 \\ 
-  %  \hline
- % \end{tabular}
- % \caption{\WFcomment{Comment on results once they are correct. These numbers are wrong right now.}}
-%\end{table}
 
@@
 \input{tables/error_rates_05.tex}
-\caption[tab:stopping]{Model-selection performance of various stopping rules applied to simulated data with 7 strong signals,  as described at the beginning of Section~\ref{sec:sparseReg}.
-Results that are theoretically guaranteed to be less than $\alpha$ are in \guarantee{bold} type.}
+\caption[tab:stopping]{Model-selection performance of various stopping rules applied to simulated data with 7 strong signals,  at level $\alpha=0.05$.
+Values theoretically guaranteed to be less than $\alpha$ are in \guarantee{bold} type.
+The largest standard error in each of the six columns is 0.01, 0.01, 0.02, 0.004, 0.003, and 0.05.}
 \label{tab:stopping05}
@@
 \input{tables/error_rates_20.tex}
-\caption[tab:stopping]{Model-selection performance of various stopping rules applied to simulated data with 7 strong signals,  as described at the beginning of Section~\ref{sec:sparseReg}.
-Results that are theoretically guaranteed to be less than $\alpha$ are in \guarantee{bold} type.}
+\caption[tab:stopping]{Model-selection performance of various stopping rules applied to simulated data with 7 strong signals, at level $\alpha=0.2$.
+Values theoretically guaranteed to be less than $\alpha$ are in \guarantee{bold} type.
+The largest standard error in each of the six columns is 0.01, 0.01, 0.03, 0.004, 0.004, and 0.07.}
 \label{tab:stopping20}
@@
 
-We see that BasicStop and ForwardStop control model-wise FWER and FDR, respectively, using $p$-values from max-$t$ or max-$t$-identify, as predicted by the theory, and the nominal $p$-values do not lead to control of FWER or FDR, as expected due to their anti-conservatism.
+We see that BasicStop and ForwardStop control model-wise FWER and FDR, respectively, using $p$-values from max-$z$ and max-$t$, as predicted by the theory, and the nominal $p$-values do not lead to control of FWER or FDR, as expected due to their anti-conservatism.
 Although not guaranteed by the theory, the $p$-values from the saturated model do control model-wise FDR in this example.
-The max-$t$ rules show higher probabilities of selecting the true model, especially using ForwardStop.
-Lastly, the knockoff method does not yield variable-wise FDR control.
+The max-$z$ method is more powerful, showing higher probabilities of selecting a correct model, especially when using ForwardStop.
+The max-$t$ method is also powerful but slightly less so than the max-$z$, owing to the additional nuisance variable $\sigma^2$.
+
+The knockoff method does not yield variable-wise FDR control.
 While the more conservative knockoff+ method does achieve variable-wise FDR control, this comes at a cost of very low power.
@@
   \includegraphics[width=\textwidth]{figs/comparison.pdf}
-  \caption{\em Simulated example with $n= 100, p=20$ and one strong signal, computing the max-$t$ $p$-value at Step 3.
+  \caption{ Simulated example with $n= 100, p=20$ and one strong signal, computing the max-$t$ $p$-value at Step 3.
 Shown are bias, variance and MSE of the estimates from accept/reject and hit and run, using between 1000 and 8000 samples, averaged over 10 different starting seeds.

### 2015-12-06 code/python/error_rates.py

@@ -10,6 +10,7 @@ for results, outbase in zip(['../snr_5_alpha_%s.csv' % alpha for alpha in ['05',
 
     names = []
     results = []
+    stderrs = []
     guarantees = []
 
     name_map = dict(zip(['nominalT', 'maxT', 'maxT_identify', 'maxT_unknown', 'saturated'],
@@ -19,13 +20,21 @@ for results, outbase in zip(['../snr_5_alpha_%s.csv' % alpha for alpha in ['05',
     for test, rule in product(['nominalT', 'maxT', 'maxT_identify', 'maxT_unknown', 'saturated'],
                               ['simple', 'forward']):
         name = '_'.join([test, rule])
+        n_sim = len(getvar("screen"))
         screen = np.mean(getvar("screen"))
+        screen_se = np.sqrt(np.var(getvar("screen"))/n_sim)
         FDR_var = np.mean(getvar("FDP_var"))
+        FDR_var_se = np.sqrt(np.var(getvar("FDP_var"))/n_sim)
         FDR_model = np.mean(getvar("FDP_model"))
+        FDR_model_se = np.sqrt(np.var(getvar("FDP_model"))/n_sim)
         FDR_model_cond = np.mean(getvar("FDP_model") * getvar("screen")) / screen
+        FDR_model_cond_se = np.nan #np.sqrt(np.mean((getvar("FDP_model") * getvar("screen") - FDR_model_cond)^2) / screen / (screen*n_sim))
         FWER_model = np.mean(getvar("FWER_model"))
+        FWER_model_se = np.sqrt(np.var(getvar("FWER_model"))/n_sim)
         FWER_model_cond = FWER_model / screen
+        FWER_model_cond_se = np.sqrt(FWER_model_cond * (1 - FWER_model_cond) / (n_sim * screen))
         S_var = np.mean(getvar("S_var"))
+        S_var_se = np.sqrt(np.var(getvar("S_var"))/n_sim)
 
         if test in ['maxT', 'maxT_identify', 'maxT_unknown']:
             if rule == 'simple':
@@ -37,7 +46,8 @@ for results, outbase in zip(['../snr_5_alpha_%s.csv' % alpha for alpha in ['05',
         else:
             guarantees.append((False, False, False, False, False, False))
         results.append((screen, FWER_model, FWER_model_cond, FDR_model, FDR_var, S_var))
-        names.append('%s %s' % (name_map[test], rule_map[rule]))
[…]

### 2015-12-07 paper.tex
minor edits, deleted comments, added discussion of computation for next-entry

@@
 \newcommand{\Gv}{\;\;\big|\;\;}
-%\newcommand{\cP}{\mathcal{P}}
 \newcommand{\proj}{\cP}
@@
 Many model selection procedures produce a sequence of increasingly complex models, leaving the data analyst to choose among them. 
-%While methods like cross-validation aim at minimizing out-of-sample prediction error, such methods tend to select a large fraction of noise variables. 
 Given such a path, we consider the problem of choosing the simplest model in the path that is not falsified by the available data.
@@
 \centering
-\input{tables/diabetes_edit.tex}
+{\small \input{tables/diabetes_edit.tex}}
 \caption[tab:diab]{Forward stepwise regression for the diabetes data: naive $p$-values, $p$-values from the saturated model, and our max-$t$ $p$-values.
@@
 \]
-\RJTcomment{Why does $M_0(Y)$ depend on $Y$ but $M_\infty$ does not?
-I'm thinking of $M_0(Y)$ as the model with intercept, no?
-For simplicity I've removed the dependence of $M_0$ on $Y$, but if I'm wrong about this, sorry, just change it back.} \\
-\WFcomment{In many cases $M_0$ is deterministic but (for example) the way we define the ever-active path, we have $E_0(Y) = \bigcup_{\lambda\in\Lambda} \tE_\lambda(Y)$, which could be random if $\Lambda$ is a fixed grid.
-$M_\infty$ doesn't really need to be selected since we can just take it as a union over all possible models.}\\
 By {\em model}, we mean a family of candidate probability distributions for $Y$. For example in  linear regression, a model specifies  the set of
@@
 \[
-\L(Y \;\mid\; U_{k-1}, \; M_{k-1}=M)
+\L_F(Y \;\mid\; U_{k-1}, \; M_{k-1}=M)
 \]
 is the same for every $F\in M$.
-Thus, we can construct selective tests by comparing the realized value of any test statistic $T(Y)$ to its known conditional distribution under the null.
-What remains is to choose the test statistic and compute its conditional null distribution, which can be challenging.
+Thus, we can construct selective tests and $p$-values by comparing the realized value of any test statistic $T_k(Y)$ to its known conditional distribution under the null.
+It remains only to choose the test statistic and compute its conditional null distribution, which can be challenging.
 See \citet{fithian2014optimal} for a general treatment.
@@
 But, unfortunately, they also carry a number of drawbacks: they require us to assume $\sigma^2$ is known, can result in a large reduction in power, create dependence between $p$-values at different steps, and are difficult to generalize beyond the case of linear regression.
-%See \citet{gsell2013false} for an in-depth discussion of the relative conceptual advantages and disadvantages between the three approaches.
 
@@
 
-\RJTcomment{I've not changed anything with the notation below that conditions on $k_0=k$.
-I'll leave it up to you Will, but as we discussed, I think it would be most useful to either write $P(p_k \leq \alpha | M_{k-1})$ on the set $\{ F \in M_{k-1}\}$, or to plainly explain what you mean by conditioning on an event involving $k_0$.} \\
 In our problem setting, the sequence of selected models is random; thus, the truth or falsehood of each $H_k$ is not fixed, as assumed in the previous section.
@@
 \[
-\sF_k = \sF(M_{0:k}, U_k).
-\]
-By our assumption of minimal sufficiency, $\sF_i \sub \sF_k$ for $i\leq k$.
+\sF_k = \sF(M_{0:k}, U_k),
+\]
+where $\sF(Z)$ denotes the $\sigma$-algebra generated by random variable $Z$. By our assumption of minimal sufficiency, $\sF_i \sub \sF_k$ for $i\leq k$.
 
@@
     \includegraphics[width=\textwidth]{figs/bivariateSelVSat_powCurves_0.pdf}
-    %\caption{\WFcomment{Write caption here.}}
-    %\label{fig:bv_powCurves_0}
   \end{subfigure}
@@
     \includegraphics[width=\textwidth]{figs/bivariateSelVSat_powCurves_4.pdf}
-    %\caption{\WFcomment{Write caption here.}}
   \end{subfigure}
@@
 Because $U_{k-1}$ is sufficient for $Y$ under the null, this law is completely known.
-Thus, we have reduced computation of the $p$-value at step $k$ to a sampling problem: if we can sample values of $Y$ from its null distribution given $M_{k-1}(Y), U_{k-1}(Y)$, we can numerically approximate the $p$-value to whatever precision we desire.
+Thus, we have reduced computation of the $p$-value at step $k$ to a sampling problem: if we can sample values of $Y$ from its null distribution given $M_{k-1}(Y), U_{k-1}(Y)$, and compute the test statistic $T_k(Y)$ for each sample, we can numerically approximate the $p$-value to whatever precision we desire.
 
@@
 
+\subsection{Computing the Next-Entry $p$-Values}\label{sec:next-entry-comp}
+
+For the next entry test, the test statistic $\lambda_k(Y)$ is a function of the regularization path.
+A na\"{i}ve algorithm would simply recompute the entire path for each new sample $Y^*$ from the conditional null distribution in~\eqref{eq:genDist}.
+However, if we are a bit more clever, we can avoid ever recomputing the path.
+
+Let $A$ denote the event $\{E_{k-1}=E, U_{k-1}=u\}$, and let $Y^*$ represent a sample from $F$, independent of $Y$. On $A$, the next-entry $p$-value $p_k$ may be written as
+\begin{align}
+  p_k(Y) 
+  &= \P(\lambda_k(Y^*) > \lambda_k(Y) \mid E_{k-1}(Y^*)=E, M_{k-1}(Y^*)=u)\\
+  &= 1 - \P\left( \bigcup_{\lambda > \lambda_k(Y)} \{\hat\theta^\lambda(Y^*) = \hat\theta^{(E, \lambda)}(Y^*)\} \mid E_{k-1}(Y^*)=E, U_{k-1}(Y^*)=u\right) \\
+  &= 1 - \P\left( \bigcup_{\lambda > \lambda_k(Y)} \hat\theta^\lambda(Y^*) = \hat\theta^{\lambda}(Y) \mid E_{k-1}(Y^*)=E, U_{k-1}(Y^*)=u\right),
+\end{align}
+where the third equality is justified by the fact that $\hat\theta^{(E, \lambda)}$ depends only on $U_E$, and $\hat\theta^{(E,\lambda)}(Y)=\hat\theta^\lambda(Y)$ for $\lambda > \lambda_k(Y)$.
+As a result, we do {\em not} need to compute the whole path for each new $Y^*$; instead, we only need to compute the regularization path for $Y$, and then check for each $\lambda > \lambda_k(Y)$ whether $\hat\theta^\lambda(Y)$ is also optimal for $Y^*$.
+If $-\ell$ and $P_\lambda$ are both convex, then, we merely need to verify the Karush-Kuhn-Tucker (KKT) conditions.
+
+\paragraph{Remark} If $\lambda_k$ has a discrete distribution --- as it would, for example, if $\Lambda$ is a discrete grid --- then we can either accept that $p_k$ is conservative, or randomize $p_k$ so it is exactly uniform under the null.
+For a randomized $p_k$, we also need the probability $\lambda_k(Y^*)$ {\em equals} $\lambda_k(Y)$, which requires computing $\hat\theta^{(E, \lambda_k)}(Y)$ and checking whether $\hat\theta^{(E, )}$.
+
 \section{Simulation: Sparse Linear Regression}
@@
   \centering
-  % source code: ??? for simulation, sparseSim.R for plotting
+  % source code: compare_methods.py for simulation, sparseSim_figs.R for plotting
   \includegraphics[width=1\textwidth]{figs/simulation_snr_5_alpha_05_null_false.pdf}
@@
   \centering
-  % source code: ??? for simulation, sparseSim.R for plotting
+  % source code: compare_methods.py for simulation, sparseSim_figs.R for plotting
   \includegraphics[width=1\textwidth]{figs/simulation_snr_5_alpha_05_null_true.pdf}
@@
   \centering
-  % source code: ??? for simulation, sparseSim.R for plotting
+  % source code: compare_methods.py for simulation, sparseSim_figs.R for plotting
   \includegraphics[width=1\textwidth]{figs/simulation_snr_5_alpha_05_noise_var.pdf}
@@
 Hence we can obtain independent, sequential $p$-values for each successive split in the decision tree.
-\RJTcomment{Just for what it's worth, I always think of CART as performing best-first splits.
-This is how Jerry teaches it, at least.}
 
@@
 
-Open-source code is available at this article's github repository at \texttt{github.com/wfithian/sequential-selection}.
-R and Python code are available to implement proposals in this paper are available at \texttt{github.com/selective-inference}.
+Open-source code is available at this article's github repository at:\\ \texttt{github.com/wfithian/sequential-selection}.
+
+R and Python code are available to implement proposals in this paper are available at:\\ \texttt{github.com/selective-inference}.
 
@@
 
-\JTcomment{I don't really see the point of Figure 5. It is abundandtly clear
-that if feasible an IID sample is preferable over an MCMC sample. It is
-just as easy to make plots where accept/reject will take forever to get even one sample point.}
-
 In detail, we employ a hybrid computational strategy; we use accept-reject sampling, and do so as long as we obtain at least 300 (say) samples in the first set of $B$ realizations ($75,000$ in our code)  This is an effective approach until we have reached a point in the model path where all predictors have little or no signal.

### 2015-12-08 paper.tex
rob
modified section comparing null hypotheses, and 'related work'

@@
 Given such a sequence and the data used to produce them, we consider the problem of choosing the least complex model that is not falsified by the data.
-Extending the selected-model tests of \citet{fithian2014optimal}, we construct $p$-values for each step in the path, accounting for the fact that the model path is determined adaptively using the data.
+Extending the selected-model tests of \citet{fithian2014optimal}, we construct $p$-values for each step in the path which account for the adaptive selection of the model path using the data.
 In the case of linear regression, we propose two specific tests, the max-$t$ test for forward stepwise regression (generalizing a proposal of \citet{buja2014}), and the next-entry test for the lasso.
@@
 Note that in our formulation of the problem, the type I error \smash{$V=(\hk-k_0)_+$} is defined in a ``model-centric'' fashion: at step $k$ in linear regression, we are testing  whether a particular linear model $M(E_{k-1})$ adequately describes the data $Y$.
-Even if the next selected variable $X_{j_k}$ is a noise variable (i.e., truly inactive) in the full model, it is not necessarily a mistake to reject $M(E_{k-1})$ provided there are signal variables (i.e., truly active ones) that have not yet been included.
-
-In some circumstances, we might want to define a type I error at step $k$ differently, by choosing a different null hypothesis to test.
+Even if the next selected variable $X_{j_k}$ has a zero coefficient in the full model, it is not a mistake to reject $M(E_{k-1})$ provided there are some signal variables that have not yet been included.
+
+Depending on the scientific context, we might want to define a type I error at step $k$ differently, by choosing a different null hypothesis to test.
 Let $\mu = \E Y$ and let $\theta^E$ denote the {\em least-squares coefficients} of active set $E$ --- the coefficients of the best linear predictor for the design matrix $X_E$:
@@
 \[
-H_k^{\text{comp}}:\;\mu = X_{E_{k-1}} \theta^{E_{k-1}}.
+H_k:\;\mu = X_{E_{k-1}} \theta^{E_{k-1}}.
 \]
@@
 
-While the complete null is the strongest hypothesis of the three, the incremental null is neither weaker nor stronger than the full-model null. Defining 
+While the complete null is the strongest null hypothesis of the three, the incremental null is neither weaker nor stronger than the full-model null. Defining 
 \begin{align*}
@@
 
-The full-model null is the most different conceptually from the other two, taking a ``variable-centric'' instead of ``model-centric'' viewpoint, where it is of primary scientific interest to discover variables that have nonzero coefficients after controlling for all $p-1$ other variables that were measured.
-
-To elucidate this distinction consider two methods we could apply in a classical setting where we have some fixed {\em a priori} ordering for the variables, inducing the fixed sequence of nested linear models
-\[
-M(\emptyset) \sub M(\{1\}) \sub M(\{1,2\}) \sub \cdots \sub M(\{1,\ldots, p\}).
-\]
-The goal of course is to select a model along this sequence.
-In the first procedure, we construct a (type I) analysis-of-variance (ANOVA) table performing an $F$- or $t$-test of each model against the next.
-We then stop and select model $M_k=M(\{1,\ldots,k\})$ the first time the $F$-test $p$-value is larger than $\alpha$.
-This procedure would control the complete-null and incremental-null FWER, but {\em not} the full-model-null FWER.
-The methods we consider in this article are essentially generalizations of this procedure to the setting where the model sequence is selected adaptively, and where we may wish to control FDR instead of FWER.
-
-In the second procedure, we fit the full model $M(\{1,\ldots,p\})$ and compute a $t$-statistic for each variable with respect to that model, yielding $t$-statistics $t_1, t_2, \ldots, t_p$ (not ordered by size).
-For $k=1,\ldots,p$, we compare each $|t_j|$ to the level-$\alpha$ cutoff for a $t$-test with $n-p$ degrees of freedom, and stop selecting variables the first time $|t_{k+1}|$ is not larger than this cutoff.
-We might use this procedure if the variables had a  fixed ordering.
-This procedure would control the full-model-null FWER, but {\em not} the incremental-null FWER.
-Note that the {\em model} $M_k$ plays no role in this procedure: its fit to the data is never evaluated and it is not really selected in any meaningful sense.
-Even if the fit of $M_k$ is far worse than that of $M_{k+1}$, we may not be able to select $X_{k+1}$ if (say) it is highly collinear with $X_{k+100}$.
-Conversely, even if the fit of $M_k$ is no better than that of $M_{k-1}$, we may still select $X_k$ if it is a highly useful predictor when used in concert with $X_{k+100}$.
-
-More generally, methods that control full-model error rates are best viewed not as {\em model-selection} procedures --- since all inferences are made with respect to the full model, which is specified ahead of time --- but {\em variable-selection} procedures that test multiple hypotheses with respect to a single fixed model.
-Note that the truth or falsehood of these hypotheses can depend sensitively on the set of $p$ predictors: rejecting $H_j^{\text{full}}$ has no meaning without reference to the list of all variables that we controlled for, which can make a rejection difficult to interpret when $p$ is large.
+The full-model null is the most different conceptually from the other two, taking a ``variable-centric'' instead of ``model-centric'' viewpoint, with the focus on discovering variables that have nonzero coefficients after adjusting for all $p-1$ other variables under consideration.
+To elucidate this distinction, consider a bivariate regression example in which the two predictors $X_1$ and $X_2$ are both highly correlated with $Y$, but are also nearly collinear with each other, making it impossible to distinguish which variable has the ``true'' effect.
+Any procedure that controls $\text{FWER}^{\text{full}}$ could not select either variable, and would return the empty set of predictors.
+By contrast, most of the methods presented in this article would select the first variable to enter (rejecting the global null model), and then stop at the univariate model.
+
+Similarly, consider a genomics model with quantitative phenotype $Y$ and predictors $X_j$, representing minor allele counts for each of $p$ single-nucleotide polymorphisms (SNPs).
+If correlation between neighboring SNPs (neighboring $X_j$'s) is high, it may be very difficult to identify SNPs that are highly correlated with $Y$, adjusting for all other SNPs; however, we might nevertheless be glad to select a model with a single SNP from each important {\em gene}, even if we cannot guarantee it is truly the ``best'' SNP from that gene.
+
+As the above examples illustrate, methods that control full-model error rates are best viewed not as {\em model-selection} procedures --- since all inferences are made with respect to the full model --- but instead as {\em variable-selection} procedures that test multiple hypotheses with respect to a single model, which is specified ahead of time.
+The ``model'' returned by such procedures is not selected or validated in any meaningful sense.
+Indeed, in the bivariate example above, of the four models under consideration ($\emptyset$, $\{1\}$, $\{2\}$, and $\{1,2\}$), only the global null model is clearly inconsistent with the data; and yet, a full-model procedure is bound not to return any predictors.
+
+Because the truth or falsehood of full-model hypotheses can depend sensitively on the set of $p$ predictors, rejecting $H_j^{\text{full}}$ has no meaning without reference to the list of all variables that we controlled for.
+As a result, rejections may be difficult to interpret when $p$ is large.
 Thus, error rates like $\text{FDR}^{\text{full}}$ are best motivated when the full model has some special scientific status.
-For example, the scientist may believe due to theoretical considerations that the linear model in $X_1,\ldots,X_p$ is fairly credible, and that a nonzero coefficient $\beta_j$, controlling for all of the other variables, would constitute evidence for a causal effect of $X_j$ on the response.
-
-In this article we will concern ourselves primarily with testing the complete null, reflecting our stated aim of choosing the least complex model that is not rejected by the data.
+For example, the scientist may believe, due to theoretical considerations, that the linear model in $X_1,\ldots,X_p$ is fairly credible, and that a nonzero coefficient $\beta_j$, controlling for all of the other variables, would constitute evidence for a causal effect of $X_j$ on the response.
+
+In this article we will concern ourselves primarily with testing the complete null, reflecting our stated aim of {\em choosing the least complex model that is consistent with the data}.
 As we discuss further in Section~\ref{sec:selective-reg}, the saturated-model tests of \citet{taylor2014exact} are valid selective tests of $H_k^{\text{inc}}$.
@@
 The stability selection approach of \citet{meinshausen2010stability} uses an approach based on splitting the data many times and offers asymptotic control of the FDR, but no finite-sample guarantees are available.
+
+If the full model is sparse, and the predictors are not too highly correlated, it may be possible asymptotically to recover the support of the full-model coefficients with high probability --- the property of {\em sparsistency}.
+Much recent model-selection literature focuses on characterizing the regime in which sparsistency is possible; see, e.g.
+\citet{bickel2009simultaneous}, \citet{meinshausen2006high}, \citet{negahban2009unified}, \citet{van2009conditions}, \citet{wainwright2009sharp}, \citet{zhao2006model}.
+Under this regime, there is no need to distinguish between the ``model-centric'' and ``variable-centric'' viewpoints.
+However, the required conditions for sparsistency can be difficult to verify, and in many applied settings they fail to hold.
+By contrast, the methods presented here require no assumptions about sparsity or about the design matrix $X$, and offer finite-sample guarantees.
 
@@
 
-\section{Futher Examples and Extensions}\label{sec:further-examples}
+\section{Further Examples and Extensions}\label{sec:further-examples}
 
@@
 R and Python code are available to implement proposals in this paper are available at:\\ \texttt{github.com/selective-inference}.
+They will also be added to an upcoming version of the  R package  \texttt{selectiveInference} on CRAN.
 

### 2015-12-08 paper.tex
fixed some minor things. This is arxiv v1

@@
 \newcommand{\cJ}{\mathcal{J}}
+\newcommand{\cL}{\mathcal{L}}
 \newcommand{\leqAS}{\overset{\textrm{a.s.}}{\leq}}
@@
 \[
-\L_F(Y \;\mid\; U_{k-1}, \; M_{k-1}=M)
+\cL_F(Y \;\mid\; U_{k-1}, \; M_{k-1}=M)
 \]
@@
 \begin{equation*}
-\L\left(G_k \;\mid\; M_{k-1}, U_{k-1}\right)
+\cL\left(G_k \;\mid\; M_{k-1}, U_{k-1}\right)
 \end{equation*}
@@
 \begin{equation*}
-\L\left(\lambda_k \;\mid\; U_{k-1}, M_{k-1}\right).
+\cL\left(\lambda_k \;\mid\; U_{k-1}, M_{k-1}\right).
 \end{equation*}
@@
 
-We now review several proposals for ordered testing of fixed null hypotheses along with their error guarantees in the traditional setting, where the sequence of null hypotheses is fixed.
+We now review several proposals for ordered hypotheses testing along with their error guarantees in the traditional setting, where the sequence of null hypotheses is fixed.
 In the next section we will extend the analysis to random null hypotheses.
@@
 The problem is that $k_0$ is no longer a fixed index.
-Consequently, even though $\L(p_k \mid H_k \text{ true})=U[0,1]$ for each fixed $k$, the $p$-value \smash{$p_{k_0}$} corresponding to the {\em first} true null hypothesis is stochastically smaller than uniform.
-In this counterexample, $k_0=2 \iff |Y_3|>z_{\alpha/2}$, giving $\text{FWER}\eqAS 1$ on the event $k_0=2$ and leaving no room for error when $k_0\neq 2$.
+Consequently, even though $\cL(p_k \mid H_k \text{ true})=U[0,1]$ for each fixed $k$, the $p$-value \smash{$p_{k_0}$} corresponding to the {\em first} true null hypothesis is stochastically smaller than uniform.
+In this counterexample, $k_0=2 \iff |Y_3|>z_{\alpha/2}$, giving $\text{FWER}= 1$ conditional on the event $k_0=2$ and leaving no room for error when $k_0\neq 2$.
 
@@
 Similarly, the error control guarantees of \citet{gsell2013sequential} and \citet{li2015accumulation} do not directly apply to case where the null hypotheses are random.
-However, we recover these guarantees if we have {\em conditional} independence and uniformity of null $p$-values given $k_0$:
+However, we recover these guarantees if we have {\em conditional} independence and uniformity of null $p$-values given $k_0$: that is, if for all $k=0,\ldots, d-1$ and $\alpha_{k+1},\ldots,\alpha_{d} \in [0,1]$, we have
 \begin{equation}\label{eq:indepCond_random_k0}
   \P_F(p_{k+1} \leq \alpha_{k+1}, \ldots, p_d \leq \alpha_d
-  \mid p_1, \ldots, p_k) \eqAS \left(\prod_{i=k+1}^d \alpha_i\right) \quad \text{on } \{k_0(Y, F) = k\}.
+  \mid p_1, \ldots, p_k) \eqAS \left(\prod_{i=k+1}^d \alpha_i\right) \quad \text{on } \{k_0(Y, F) = k\},
 \end{equation}
@@
     For nested hypotheses, $k_0$ completely determines the truth or falsehood of $H_k(Y)$ for every $k$. If \smash{$\hk$} controls $\Err_g$ in the 
-fixed-hypothesis case, it must in particular control $\Err_g$ conditional on any fixed values of $p_{1:k_0}$. Thus, \eqref{eq:indepCond_random_k0} implies
+fixed-hypothesis case, it must in particular control $\Err_g$ conditional on any fixed values of $p_{1:k_0}$, since we could set 
+\[
+p_{1:k_0} \sim \prod_{k=1}^{k_0} \delta_{a_k}
+\]
+for any sequence $a_1,\ldots, a_{k_0}$, where $\delta_a$ is a point mass at $a \in [0,1]$.
+
+Thus, \eqref{eq:indepCond_random_k0} implies
   \[
@@
 \begin{equation}\label{eq:maxTnull}
-\L\left(T_k \; \mid \; E_{k-1}, \,X_{E_{k-1}}'Y, \,\|Y\|^2 \right).
+\cL\left(T_k \; \mid \; E_{k-1}, \,X_{E_{k-1}}'Y, \,\|Y\|^2 \right).
 \end{equation}
@@
 \end{enumerate}
-Informally, the pair $(E_{k-1}, U_{k-1})$ forms a ``wall of separation'' between $p_k$ and $p_{1:(k-1)}$, guaranteeing that $\L(p_k \mid p_{1:(k-1)}) = U[0,1]$ whenever $H_k$ is true.
+Informally, the pair $(E_{k-1}, U_{k-1})$ forms a ``wall of separation'' between $p_k$ and $p_{1:(k-1)}$, guaranteeing that $\cL(p_k \mid p_{1:(k-1)}) = U[0,1]$ whenever $H_k$ is true.
 
@@
 
-Furthermore, for $i<k$, $p_i$ is computed by comparing $T_i=|t_{j_i, E_{i-1}}|$, which is a function of $(E_i, U_i)$, to the reference null distribution $\L_{H_i}(T_i \mid E_{i-1}, U_{i-1})$, which is a function of $(E_{i-1}, U_{i-1})$.
+Furthermore, for $i<k$, $p_i$ is computed by comparing $T_i=|t_{j_i, E_{i-1}}|$, which is a function of $(E_i, U_i)$, to the reference null distribution $\cL_{H_i}(T_i \mid E_{i-1}, U_{i-1})$, which is a function of $(E_{i-1}, U_{i-1})$.
 Having verified that under $H_k$, $p_k$ is uniform and independent of $p_{1:{k-1}}$, we can apply this conclusion iteratively to see that all remaining $p$-values are also uniform and independent.
 
-By contrast, the saturated model $p$-values use a law different from~\eqref{eq:maxTnull} to compute $p$-values, one that uses
+By contrast, the saturated model $p$-values are computed using a reference distribution different from~\eqref{eq:maxTnull}, one that depends on
 information not contained in $(E_k,U_k)$.
@@
   \end{align*}
-  Second, the null distribution $\L\left(G_k \,\mid\, M_{k-1}, U_{k-1}\right)$ against which we compare $G_k$ is also $\sF_k$-measurable because it depends only on $(M_{k-1}, U_{k-1})$, both of which are $\sF_{k-1}$-measurable.
+  Second, the null distribution $\cL\left(G_k \,\mid\, M_{k-1}, U_{k-1}\right)$ against which we compare $G_k$ is also $\sF_k$-measurable because it depends only on $(M_{k-1}, U_{k-1})$, both of which are $\sF_{k-1}$-measurable.
 \end{proof}
@@
 As a result, \smash{${\hat\theta^{\lambda} =\hat\theta^{(E_k,\lambda)}}$} for such $\lambda$, so $\tE_{\lambda}$ depends only on \smash{$\ell_{E_k}^{Y\mid U}(Y;\, U_k)$}, and therefore $\lambda_k$ can be computed from $(E_k, U_k)$.
-Second, the null distribution $\L\left(\lambda_k \,\mid\, M_{k-1}, U_{k-1}\right)$ against which we compare $\lambda_k$ is also $\sF_k$-measurable because it depends only on $(M_{k-1}, U_{k-1})$, both of which are $\sF_{k-1}$-measurable.
+Second, the null distribution $\cL\left(\lambda_k \,\mid\, M_{k-1}, U_{k-1}\right)$ against which we compare $\lambda_k$ is also $\sF_k$-measurable because it depends only on $(M_{k-1}, U_{k-1})$, both of which are $\sF_{k-1}$-measurable.
 \end{proof}
@@
 Let $\eta_k\in \R^n$ denote the vector for which \smash{$\theta_{j_k}^{E_k} = \eta_k'\mu$}, and let \smash{$\proj_{\eta_k}^\perp$} denote the projection operator into the space orthogonal to $\eta_k$.
-The UMPU saturated-model test for $\eta_k'\mu$ is based on the distribution \smash{$\L\left( \eta_k'Y \mid E_{k-1}, \;j_k, \;\proj_{\eta_k}^\perp Y \right)$}, or equivalently
+The UMPU saturated-model test for $\eta_k'\mu$ is based on the distribution \smash{$\cL\left( \eta_k'Y \mid E_{k-1}, \;j_k, \;\proj_{\eta_k}^\perp Y \right)$}, or equivalently
 \begin{equation}\label{eq:satModel}
-\L\left( X_k'Y \mid E_{k-1}, \;j_k, \;\proj_{\eta_k}^\perp Y \right)
+\cL\left( X_k'Y \mid E_{k-1}, \;j_k, \;\proj_{\eta_k}^\perp Y \right)
 \end{equation}
@@
 \begin{equation}\label{eq:selModel_cond}
-\L\left(X_{j_k}'Y \mid E_{k-1}, \; j_k, \; X_{E_{k-1}}'Y\right).
+\cL\left(X_{j_k}'Y \mid E_{k-1}, \; j_k, \; X_{E_{k-1}}'Y\right).
 \end{equation}
@@
 \begin{equation}\label{eq:selModel_marg}
-\L\left(X_{j_k}'Y \mid E_{k-1}, \; X_{E_{k-1}}'Y\right).
+\cL\left(X_{j_k}'Y \mid E_{k-1}, \; X_{E_{k-1}}'Y\right).
 \end{equation}
@@
   \caption{Contrast between the saturated-model and selected-model
-    tests in Example~\ref{ex:bivariate}. The selected-model test is based on  $\L(Y_1 \gv j_1=1)$,  whereas the saturated-model test is based on $\L(Y_1  \gv Y_2, \; j_1=1)$. 
+    tests in Example~\ref{ex:bivariate}. The selected-model test is based on  $\cL(Y_1 \gv j_1=1)$,  whereas the saturated-model test is based on $\cL(Y_1  \gv Y_2, \; j_1=1)$. 
     When $Y=(2.9, 2.5)$, the selected- and saturated-model $p$-values are 0.007 and 0.3, respectively.}
@@
 \begin{equation}\label{eq:genDist}
-\L(T_k(Y) \,\mid\, M_{k-1}(Y), U_{k-1}(Y)),
+\cL(T_k(Y) \,\mid\, M_{k-1}(Y), U_{k-1}(Y)),
 \end{equation}
@@
 \begin{equation}\label{eq:condPerm}
-\L\left(W_k^* \mid E_{k-1}, U_{k-1}\right),
+\cL\left(W_k^* \mid E_{k-1}, U_{k-1}\right),
 \end{equation}
@@
 \[
-\L(|z_{E_{k-1}}^*| \,\mid\, E_{k-1}, Y_{E_{k-1}}).
+\cL(|z_{E_{k-1}}^*| \,\mid\, E_{k-1}, Y_{E_{k-1}}).
 \]
@@
 \begin{equation}\label{eq:truncNorm}
-\L(Y \mid X_E'Y, Y\in A),
+\cL(Y \mid X_E'Y, Y\in A),
 \end{equation}
@@
 \begin{equation}\label{eq:truncUnif}
-\L(Y \mid X_E'Y,\, \|Y\|^2,\, Y\in A),
+\cL(Y \mid X_E'Y,\, \|Y\|^2,\, Y\in A),
 \end{equation}
@@
 \paragraph{Accept/Reject Sampler} In either of these two cases, sampling would be very easy if $A$ were the entire sample space $\R^n$; the only reason we might encounter difficulty is that $A$ could be an irregular set that is difficult to condition on.
-However, we can always sample from $\L(Y \mid U)$ and then keep each new sample $Y$ if it lies in the selection event.
+However, we can always sample from $\cL(Y \mid U)$ and then keep each new sample $Y$ if it lies in the selection event.
 

### 2015-12-09 paper.tex
Fixed bib file

@@
 In the case of linear regression, we propose two specific tests, the max-$t$ test for forward stepwise regression (generalizing a proposal of \citet{buja2014}), and the next-entry test for the lasso.
-These tests improve on the power of the saturated-model test of \citet{taylor2014exact}, sometimes dramatically.
+These tests improve on the power of the saturated-model test of \citet{tibshirani2014exact}, sometimes dramatically.
 In addition, our framework extends beyond linear regression to a much more general class of parametric and nonparametric model selection problems.
@@
 To select a model, we can feed our single-step $p$-values as inputs into sequential stopping rules such as those proposed by \citet{gsell2013sequential} and \citet{li2015accumulation}, achieving control of the familywise error rate or false discovery rate (FDR) as desired.
-The FDR-controlling rules require the null $p$-values to be independent of each other and of the non-null $p$-values, a condition not satisfied by the saturated-model $p$-values of \citet{taylor2014exact}.
+The FDR-controlling rules require the null $p$-values to be independent of each other and of the non-null $p$-values, a condition not satisfied by the saturated-model $p$-values of \citet{tibshirani2014exact}.
 We derive intuitive and general sufficient conditions for independence, and show that our proposed constructions yield independent $p$-values.
@@
 Examples of path algorithms include forward stepwise linear regression, least angle regression (LAR) and the ever-active path in lasso ($\ell_1$-regularized) regression.
-\citet{taylor2014exact} study methods for generating exact $p$-values at each step of these path algorithms, and their methods provide a starting point for our proposals.
+\citet{tibshirani2014exact} study methods for generating exact $p$-values at each step of these path algorithms, and their methods provide a starting point for our proposals.
 Other related works include \citet{loftus2014significance}, who describe $p$-values for path algorithms that add groups of variables (instead of individual variables) at each step, and \citet{choi2014selecting}, who describe $p$-values for steps of principle components analysis (each step marking the estimation of a principle component direction).
@@
 Both sets of rules require that once we have reached the first correct model in our path, the $p$-values in subsequent steps are uniform and independent.
-While this is not true of the $p$-values constructed in \citet{taylor2014exact}, in this paper we develop a theoretical framework for constructing sequential $p$-values satisfying these properties, and give explicit constructions.
+While this is not true of the $p$-values constructed in \citet{tibshirani2014exact}, in this paper we develop a theoretical framework for constructing sequential $p$-values satisfying these properties, and give explicit constructions.
 
@@
 In this article we will concern ourselves primarily with testing the complete null, reflecting our stated aim of {\em choosing the least complex model that is consistent with the data}.
-As we discuss further in Section~\ref{sec:selective-reg}, the saturated-model tests of \citet{taylor2014exact} are valid selective tests of $H_k^{\text{inc}}$.
+As we discuss further in Section~\ref{sec:selective-reg}, the saturated-model tests of \citet{tibshirani2014exact} are valid selective tests of $H_k^{\text{inc}}$.
 The advantage of these tests is that they are highly computationally efficient (they do not require sampling).
@@
 As we have shown, the selective max-$t$ test is equivalent to the selective greedy likelihood test in the case of linear regression.
-In the next section, we will compare and contrast the max-$t$ and next-entry, and other {\em selected-model tests}, with the {\em saturated-model tests} proposed by \citet{taylor2014exact} and others.
+In the next section, we will compare and contrast the max-$t$ and next-entry, and other {\em selected-model tests}, with the {\em saturated-model tests} proposed by \citet{tibshirani2014exact} and others.
 
@@
 Many conditional selective tests for linear regression use the saturated-model framework.
-See for example \citet{taylor2013tests}, \citet{taylor2014exact}, \citet{lee2013exact}, and \citet{loftus2014significance}.
+See for example \citet{taylor2013tests}, \citet{tibshirani2014exact}, \citet{lee2013exact}, and \citet{loftus2014significance}.
 Define the saturated model as
@@
 At each step $s<k$, the selection of variable $j_s$ is a result of comparing its $t$-statistic with the $t$-statistics of the other $p-k$ inactive variables.
-Based on this logic, we might expect to accumulate at least $O(p)$ constraints at each step, resulting in $O(kp)$ constraints after $k$ steps \citep{taylor2014exact}.
+Based on this logic, we might expect to accumulate at least $O(p)$ constraints at each step, resulting in $O(kp)$ constraints after $k$ steps \citep{tibshirani2014exact}.
 

### 2016-03-23 paper.tex

@@
 \end{description}
+See~\citet{gsell2013false} for further discussion contrasting these distinct hypotheses.
 
@@
 Much recent model-selection literature focuses on characterizing the regime in which sparsistency is possible; see, e.g.
-\citet{bickel2009simultaneous}, \citet{meinshausen2006high}, \citet{negahban2009unified}, \citet{van2009conditions}, \citet{wainwright2009sharp}, \citet{zhao2006model}.
-Under this regime, there is no need to distinguish between the ``model-centric'' and ``variable-centric'' viewpoints.
-However, the required conditions for sparsistency can be difficult to verify, and in many applied settings they fail to hold.
+\citet{bickel2009simultaneous}, \citet{meinshausen2006high}, \citet{negahban2009unified}, \citet{van2009conditions}, \citet{wainwright2009sharp}, \citet{zhao2006model}, and \citet{kozbur2015testing}.
+Under this regime, we need not distinguish between the ``model-centric'' and ``variable-centric'' viewpoints.
+However, the conditions for sparsistency can be difficult to verify, and in many applied settings they fail to hold.
 By contrast, the methods presented here require no assumptions about sparsity or about the design matrix $X$, and offer finite-sample guarantees.
@@
 \]
+
 The forward stepwise path and the lasso path both select $j_1=1$ if and only if $|Y_1|>|Y_2|$.
 The selection event $A_1=\{|Y_1| > |Y_2|\}$ is shown in yellow in Figure~\ref{fig:bv_condSets}.
-On $A_1$, the first two models are
+If $j_1=1$ the first two models are
 \[
